### 2020-05-04 code/PS904-Project.Rmd
commit

---
title: The Pen Mightier Than the Sword? Retired Flag Officers and Generals and Opinion-Editorials
  Since 2000
author: "Pete Erickson"
date: "4/22/2020"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library("here")       # file paths
library("magrittr")   # pipe operators!
library("tidyverse")  # workhorse package
library("lubridate")  # lubridate helps with the package
library("plyr")

```

First, I import the data set and using the here () function to build the file path.  
I replace all NA/s with ) in the data set.

```{r summary}
# Read in the Data
dataset <- read_csv(here("data", "MilEditorials23Apr.csv"))
# Replace NA with 0 in the data
dataset[is.na(dataset)] <- 0

# Let's view the data so we know what kind of dataframes each of the variables are.  
dataset

# Note that I may want to convert AuthSe, AuthRank, and Service to some form of numeric.  
# But I will come back to that.

#See the variable names.  
summary(dataset)
names(dataset) 
```

```{r manipulation}

# Create a Variable for Publishing Year called PubYr
adjdataset <- dataset %>% 
  mutate(
  PubYr = year(PubDate)
) %>% print()

# Show all articles and authors who published articles on Social Policy, Political Parties, and Administrations/Political Figures
adjdataset %>% arrange(PubYr) %>% 
  filter(PolicySocial==1 | PartyGen==1 | PartyName ==1 |PolAdminName==1 )  %>% 
  select (AuthName, ArtTitle, Source, PubYr) %>% 
  view()

# Create a Variable Called case-count that counts the number of times by publishing year an article published 
# regarding social policy or general parties or parties by name or administrations occurs
adjdataset %>% group_by(PubYr) %>% 
  mutate(
    case_count=sum(PolicySocial ==1 |PartyGen==1 | PartyName ==1 | PolAdminName==1) 
  ) %>% 
  print

# Count the Number of Articles Published By Type
adjdataset %>% tally(PolicySecurity) 
adjdataset %>% tally(PersTroopSpt)
adjdataset %>% tally(GenAdvice)
adjdataset %>% tally(PolicySocial)
adjdataset %>% tally(PartyGen)
adjdataset %>% tally(PartyName)
adjdataset %>% tally(PolAdminName)

# need to change the type of class these vectors are - to binary

```

```{r simple graphs}

```

### 2020-05-12 code/PS904-Project.Rmd
commit

@@
 ---
-title: The Pen Mightier Than the Sword? Retired Flag Officers and Generals and Opinion-Editorials
-  Since 2000
+title: "The Pen as a Sword: Retired Military Officers and Opinion-Editorial Usage"
 author: "Pete Erickson"
-date: "4/22/2020"
+date: |
+  `r format(Sys.time(), '%B %d, %Y')`
+abstract: |
+  _This paper seeks to explain how retired military opinion-editorial usage has varied in the past 20 years.
+Using an original data set that captures more than 200 opinion-editorial pieces published by retired military officers in The Wall Street Journal, The Washington Post, The New York Times, and the USA Today since 2000, I argue that recent opinion-editorial usage is weakening the norm of retired military officers remaining apolitical in two distinct ways.
+First, op-eds published in recent years by retired military officers address topics that are outside the scope of direct security or strategic concerns.
+Second, there is an increase in recently published op-eds which directly attack Presidential administrations, specifically that of Donald Trump.
+These two trends are concerning from the perspective of civil-military relations and raise further normative questions about how the American public receives information from military elites._
+
+geometry: margin = 1.15in              # margin widths
+fontsize: 12pt                         # text size
+bibliography: researchpaper-ps904.bib # bibliography file
+biblio-style: authoryear               # citation style
+urlcolor: blue                         # links to external URLs
+citecolor: black                       # links to bibliography
+linkcolor: magenta                     # internal figures, footnotes, etc.
+indent: true                          # paragraph format
+                                       # true for indented paragraphs
+
+header-includes:
+    - \usepackage{setspace}
+    - \usepackage{multirow}
+    - \usepackage{indentfirst}
+
 output: 
-  pdf_document: default
-  html_document: default
+  bookdown::pdf_document2:
+    keep_tex: true
+    latex_engine: pdflatex
+    number_sections: false
+    toc: false
+    fig_caption: true
+    citation_package: biblatex
+    includes: 
+      in_header: 
+        - latex-preamble.tex
+  bookdown::htmldocument2: default
+  bookdown::word_document2:
+      toc: false
 ---
+
+\doublespacing
 
@@
 
-```{r}
+```{r, include = FALSE, echo = FALSE}
 library("here")       # file paths
-library("magrittr")   # pipe operators!
 library("tidyverse")  # workhorse package
-library("lubridate")  # lubridate helps with the package
-library("plyr")
+library("tidylog")
+library("kableExtra")
+library("knitr")
+```
+
+# Introduction
+
+In modern day America, the average citizen does not need to look or go far to hear or see an elite voice attempting to persuade them of the rightness or wrongness of certain courses of action.
+On the television and on the internet, "experts" in various forms take up a position on a certain issue and attempt to persuade, warn, and otherwise impact public opinion on a host of issues.
+To a considerable extent, this is a normal and everyday occurrence in a thriving democracy.
+
+This paper examines the way in which a narrow subset of elites - retired military officers - has attempted to influence public opinion in contemporary America.
+It is an interesting question worthy of examination for several reasons and from a variety of perspectives.
+First, even after leaving active duty service, military officers remain part of an institution - the military - that maintains considerable trust among the American public.
+A 2018 Pew Research Center poll determined that 80% of the American public expects the military to act in the best interests of the American public, compared to business leaders (45%), the media (40%), and elected officials (25%) [@johnson_trust_2018].
+Therefore, when leaders of the most trusted institution in the country speak, the public seems to trust them to a high degree.
+Second, it is an interesting question because as an institution, the military prides itself on being apolitical.
+In an era of increasing political polarization [@layman_party_2006; @hare_polarization_2014], examining the extent to which opinion-editorial usage by retired military officers has changed may reveal one way in which the military institution has altered its behavior in a polarized political environment.
+
+To be clear, I do not examine in this paper the efficacy of the opinion pieces authored by retired military officers.
+I do not address whether the actual op-ed pieces published by retired military officers actually sway public opinion.
+I am primarily concerned with what trends are evident in opinion-editorial publication from the perspective of the military as an institution: what variation is evident over the past 20 years, if any?
+Are there patterns that indicate deviation from historical norms and expectations of the military as an apolitical institution, and if so, what might this mean for the military?
+
+This paper proceeds in three parts.
+First, I review the basis upon which the norm of apolitical behavior of the military as an institution is based, as well as describe how a polarized environment may impact these norms.
+Here I review academic literature on civil-military relations and public opinion, as well as current military doctrine.
+All three are relevant strands that inform our research question.
+Second, I describe the analytical approach and the data used in this project.
+Assembling an original data set that includes more than 200 opinion pieces published by retired military officers in major American newspapers since 2000, I conduct a mixed-methods approach.
+Finally, I describe my results as well as their implications.
+Overall, the data reveals that the vast majority of opinion pieces penned by retired military officers are in line with traditional civil-military norms.
+However, I find two trends that may reveal future problems for civil-military relations: first, retired military officers are writing about topics that are not immediately related to national security, and second, retired military officers are criticizing if not attacking political administrations more often than they did in previous eras.
+
+# The Norm of Remaining Apolitical
+
+As a norm, the military acting as an objective, apolitical institution is deep and storied in the United States.
+In a thriving democracy such as the US, an apolitical military can best serve the Republic by defending it regardless of which political party of group holds power.
+Unlike autocratic societies, where military coups are unfortunate yet frequent in history, the United States had not had a civil-military relations crisis that has risen to the point that a coup has been a real concern.
+This is not to say that that civil-military crises are infrequent in the US - indeed, they have been rather common - but that these crises have typically been resolved.
+
+The political scientist Samuel Huntington, in his work, _The Soldier and the State: The Theory and Politics of Civil-Military Relations_, argued for civil-military relations to take on a certain character.
+Written in 1957, during the Cold War, Huntington asserted that two distinct forces would always be in tension with one another in a liberal democracy.
+The first was a "functional imperative," which Huntington said drove the requirement that a military be strong enough and capable of defeating the military threats it faced [@huntington_soldier_1957].
+The second force was a "societal imperative," which Huntington said generated requirements for the military in that the military could never lose so many aspects of it such that it no longer represented the liberal democracy that it defended [@huntington_soldier_1957].
+Huntington's point was that America needed a strong military that could win on the battlefield, but cautioned that such a military could not be so foreign in character so as to alienate the tenets of an American liberal society, and that these forces were at odds with each other.
+Huntington proceeded to argue that the key to resolving the tension between these two opposing forces was to adopt a scheme of "objective civilian control" in which political leaders established the ends and goals of policy, while military leaders developed and implemented plans to accomplish these goals [@huntington_soldier_1957].
+To Huntington, separate operating spheres between the political leaders of a democracy and the military as an institution ensured that the military remained "professional," and thus, untainted by political interests and concerns.
+In this way, Huntington posited that the American military would be up to the task of defeating the very real threats it would face, especially during the Cold War context in which Huntington wrote.
+
+While Huntington's theory has enjoyed tremendous staying power (Huntington is still studied and taught in many American military institutions), he is not without critics.
+It would be too burdensome to unpack all of those here, but for clarity, several quick counterarguments to Huntington are worth highlighting.
+Eliot Cohen, in his work, _Supreme Command,_ argues that the very concept of separate spheres between political and military leaders is flawed.
+Pointing to leaders such as Lincoln, Clemenceau, and Churchill, and the actions that each took to ensure that the military operated in accordance with each's wishes, Cohen argues that political leaders are ultimately responsible for the outcomes of war, and therefore, are not obligated to grant the military any sort of distinct operating space [@cohen_supreme_2003].
+Another scholar, Peter Feaver, argues in his work, _Armed Servants: Agency, Oversight, and Civil-Military Relations_ that a principal-agent approach most accurately conveys how civil-military relations work [@feaver_armed_2005].
+American elected civilian leaders, as agents, so argues Huntington, have the responsibility to select military leaders as agents, and to oversee ensuing problems of agency that arise, such as delegating, monitoring, and ensuring the compliance of the military as agents [@feaver_armed_2005].
+The point worth highlighting here is that Cohen and Feaver argue that civilian elected leaders have the right to encroach onto what Huntington would call for, a separate military sphere.
+Note, however, that none of these scholars argue for the opposite - that military leaders should in any way encroach onto or towards the political space.
+
+It is not only in the academic literature that the tradition and worth of an apolitical military have been cultivated.
+Indeed, the operational military - those who prepare for fighting and dying on the battlefield - have also been central in upholding the military's apolitical traditions.
+Retired Army General Douglas MacArthur, a hero of World War Two in the Pacific who is also well known for the controversy surrounding his conduct in the Korean War and his eventual dismissal by President Truman, told the cadets at West Point in May of 1962.
+
+\singlespacing
+
+> **Yours is the profession of arms, the will to win, the sure knowledge that in war there is no substitute for victory, that if you lose, the Nation will be destroyed, that the very obsession of your public service must be Duty, Honor, Country.
+Others will debate the controversial issues, national and international, which divide men’s minds.
+But serene, calm, aloof, you stand as the Nation’s war guardians, as its lifeguards from the raging tides of international conflict, as its gladiators in the arena of battle.
+For a century and a half you have defended, guarded and protected its hallowed traditions of liberty and freedom, of right and justice.
+Let civilian voices argue the merits or demerits of our processes of government.
+Whether our strength is being sapped by deficit financing indulged in too long, by federal paternalism grown too mighty, by power groups grown too arrogant, by politics grown too corrupt, by crime grown too rampant, by morals grown too low, by taxes grown too high, by extremists grown too violent; whether our personal liberties are as firm and complete as they should be.
+These great national problems are not for your professional participation or military solution.
+Your guidepost stands out like a tenfold beacon in the night: Duty, Honor, Country** [@macarthur_duty_1962].
+
+\doublespacing
+
+Current military doctrine and practices likewise reinforce the notion of the military remaining apolitical.
+Prior to the 2016 Presidential Election, Marine Corps General (Retired) Joseph Dunford, who then served as the Chairman of the Joint Chiefs of Staff, said that the basis upon which the military is able to maintain the public's trust is it's ability to remain "apolitical and neutral" [@dunford_upholding_2016].
+General Martin Dempsey, in an opinion piece published in the Washington Post on July 31, 2016, specifically criticized retired Army Major General Mike Flynn and Marine General John Allen for participating in the Republican and Democratic conventions, respectively, stating that "As generals, they have an obligation to uphold our apolitical traditions.
+They have just made the task of their successors – who continue to serve in uniform and are accountable for our security - more complicated.
+It was a mistake for them to participate as they did.
+It was a mistake for our presidential candidates to ask them to do so" [@dempsey_military_2016].
+
+It is important to keep in mind, however, that although the notion of an apolitical military is the theoretical goal of the institution, the practical record remains far from perfect.
+From MacArthur's clash with President Truman in the early 1950's, to the so-called "Revolt of the Generals" in 2005-2006, military leaders, even those who are retired, have ventured into political territory from time to time.
+What we have established, though, is clear: an apolitical military that performs its military mission well is what enables the American public to put trust and confidence in the military institution.
+Having examined the nature of the norm of the military as an apolitical institution, let us now briefly examine the topics of the media, public opinion, and polarization.
+Each of these topics intersects with the norm of the military as an apolitical institution in a unique way.
+
+# The Media, Public Opinion, and Polarization
+
+Scholars have long examined the role of elites in impacting public opinion.
+In his influential study, _The Nature and Origins of Mass Opinion_, John Zaller argues that formations in mass public opinion were the result of three interacting components: "variation in the information carried in elite discourse, individual differences in attention to this information, and individual differences in political values and other predispositions" [@zaller_nature_1992].
+Zaller believes that throughout modern day America, most citizens, as non-elites, must rely heavily on the messages that come from elites regarding various topics and issue areas [@zaller_nature_1992].
+Zaller further argues that the messages that elites send is necessarily "stereotyped" in some way: that is, the messages that elites send, in order to be understood by regular citizens who are part of the public, must be "short, simple, and highly thematic" [@zaller_nature_1992].
+In developing this argument, Zaller is not claiming that regular citizens are uninformed or careless people who are incapable of processing complex information.
+He is rather acknowledging that most citizens care most about the world that is not "beyond their personal experience," that is, the lives which they must interact with each and every day [@zaller_nature_1992].
+His point is that elites play a significant role in helping to form and sway public opinion.
+
+Michael Robinson further developed an argument in which he examines how elite cues are received and processed.
+Robinson argues that partisan groups, as a result of ongoing political polarization, are cognitively biased in distinct ways in how they receive and process information [@robinson_who_2019].
+As a result, Robinson found that partisans will express confidence in the military in vastly differing ways even after neutral, non-partisan information about various types of military failure is presented to these groups [@robinson_who_2019].
+If Republicans are less likely to express a reduction in confidence in the military as an institution than Democrats are, as Robinson alleges, then even accurate, unbiased information is potentially irrelevant, as these differing partisan groups are pre-programmed to selectively see a particular situation as they wish to see it [@robinson_who_2019].
+His main point serves to warn us regarding confidence in the military as an institution: there may be real reasons to lose confidence in the military as a result of a particular incident, but partisan filters will prevent a reduction of confidence from actually being expressed.
+
+Zaller's and Robinson's works are important considerations in synchronizing the relevant literature for this project.
+Zaller's framework makes it clear that it is a perfectly reasonable assumption that retired military officers who choose to express their opinions through the form of opinion-editorial pieces constitutes a form of elite discourse.
+Retired military officers have at least twenty years of experience, and in the case of multi-star generals or admirals, usually close to 35 or 40 years of service.
+Indeed, most retired senior generals and admirals have spent their professional careers solely in the military, and have extensive experience while serving in leadership positions of large organizations and, in many cases, during times of war.
+Retired military officers are thus elites in the field of national security, and according to Zaller, they rightfully have a potentially significant contribution to make with respect to impacting public opinion in these areas.
+And yet, we also know that the American public will not necessarily evaluate the opinions of these military experts objectively, but rather through pre-conceived and potentially partisan lenses [@robinson_who_2019].
+This consideration will be helpful when we consider our findings and potential implications.
+
+# Hypotheses
+
+To test whether there has been variation in opinion-editorial usage by retried military officers in the United States, I will examine opinion-editorial pieces published in the four leading newspapers of the United States since January 1, 2000: the _Wall Street Journal_ (WSJ), the _New York Times_ (NYT), the _Washington Post_ (WaPo), and the _USA Today_.
+This date range is important for several reasons.
+
+First, let us consider some of the significant ways in which the military as an institution has changed in the past twenty years.
+Most notably, the post-9/11 US military is a battle-hardened military.
+The wars in Afghanistan and Iraq continue to the present day, and today's recently retired military generals and admirals are, in many cases, veterans of these conflicts several times over.
+Moreover, both of these conflicts have thus far failed to reach conclusive ends.
+Heidi Urben found in a 2009 study that the proportion of military officers who felt that criticizing civilian administrations constituted proper conduct had increased since 1998, reasoning that the likely reason for the change was the frustration encountered while serving in Iraq and Afghanistan [@urben_party_2013].
+
+Moreover, the military since 2000 has undergone significant social changes.
+The repeal of the "Don't Ask, Don't Tell" Policy, which was instituted under President Clinton in 1992 and prevented homosexual members of the Armed Forces from revealing their sexual preference, was repealed in 2010.
+Then, in 2015, under the Obama Administration, the Department of Defense opened up all military occupational specialties to women, many of which, particularly those in the combat arms and special forces, had previously been closed off to  women.
+These changes in the military are just a subset of the profound change that America as a nation has experienced in the past 20 years.
+
+Likewise, the media has also undergone significant changes since 2000.
+Since the mid-2000s, social media has taken over as the primary means in which a majority of Americans receive their news [@martin_how_2018].
+Print newspapers still exist, even as many of the leading newspapers have taken up digital offerings, but overall circulation is still down compared to previous era even after factoring in the digital options available.
+Estimated total daily newspaper circulation in the US in 2000 was 56 million and 59 million for weekday and weekend editions, respectively, while in 2018, the the same figures were 29 million and 31 million.
+Despite these trends, however, the four newspapers that I examine in this project are the largest of all print newspapers in the United States.
+
+Moreover, while there a multitude of ways for military elites to express their opinions and to influence public opinion, such as appearing participating in an interview on television or on the radio, published opinion pieces, especially those that run in a major newspaper, possess certain qualities.
+First, such types of opinion pieces are deliberately authored for the purpose of informing and persuading the public on a salient or timely topic.
+Second, publishing opinion pieces, especially for major newspapers, is a highly competitive process.
+For the newspapers that I examine in this study, it is not uncommon to see opinion pieces published by heads of state, members of congress, governmental figures, and heads of industry and academia.
+So, even though overall newspaper consumption may in fact be down over the past twenty years, there is no evidence that opinion pieces have dropped in popularity.
+
+Attempting to reconcile the facts that the military prides itself on being a non-partisan and apolitical institution, that military officers with experience in Iraq and Afghanistan are more willing to criticize civilian leaders over the conduct of war, that newspaper circulations (even after accounting for digital offerings) are substantially down over the past 20 years, and that the US political environment is increasingly polarized along partisan lines, I adopt the following hypotheses:
+
+_$H_1$: The majority of op-eds published by retired military officers will be apolitical in tone._
+
+_$H_2$: There will be an increase in the number of op-eds that criticize civilian administration handling regarding the conduct of war in Iraq and Afghanistan._
+
+# The Data
+
+I first assemble the data by creating a data set that consists of opinion-editorial pieces published by retired military officers in the four major newspapers previously identified since January 1, 2000.
+I employ a search using the Factiva Database and condition my search on several filters:
+
+  * sources (WSJ, NYT, WaPo, and USA Today)
+  * date range (January 1, 2000 - April 16, 2020)
+  * article type (I select editorial, opinion, and commentary)
+  * perhaps the most difficult filter item, free text that identifies retired military officers as the   authors.
+I include terms such as "retired military officer," "retired from the Army, Navy, Marine Corps, Air Force," "former Chairman of the Joint Chiefs of Staff," "general," "admiral," and so forth.
+This is done in order to capture the biographic description that usually accompanies an editorial or opinion piece, such as "General David Petraeus is a retired Army general and former commander of all forces in Afghanistan."
+
+Factiva allows for the option of selecting print only or an "all-sources" option in the source filter.
+I select the "all-sources" option because I want to capture any online editorials or opinion pieces, in addition to those that are ran in print editions only.
+Though digital offerings have certainly expanded in the past twenty years, including digital offerings will enable a better comparison of the data over the past twenty years given that overall newspaper circulation is down over the same time period.
+Factiva also checks for duplicate articles that are published in different versions of the same newspaper, such as the New York Times and the International version of the New York Times.
+When the same articles appeared in two different versions of the same newspaper, I counted the article only one time.
+
+I further notice that when collecting and reviewing the data, some authors are identified as military officers in some opinion pieces, but not in others.
+This was especially true regarding Andrew Bacevich, a retired Army Colonel who then became a prominent historian and author.^[
+This happened with several other authors, such as Air Force General (Ret) Michael Hayden, who also served as the Director of the National Security Agency.
+Where I could identify these authors as military officers, I included the article and did a separate search under the author's last name to see if there were any missing articles that were not caught in my search.
+] I decide to conduct an additional search for Andrew Bacevich and include any other opinion pieces that were published in the sources and over the date range I am examining where he was not identified as a retired military officer.
+My basis for this decision is that there is a reasonable expectation that some readers would know his background as a retired military officer since it had been previously self-divulged in previous opinion publications.
+
+Ultimately, the data set ends up containing 216 opinion-editorial publications, penned by 129 different authors or author teams.^[There are several co-authored pieces.
+Where General (Ret.) David Petraeus wrote one publication and another one was written by General (Ret.) David Petraeus and Michael O'Hanlon, I counted these as two different authors for coding purposes.] Additionally, 10 pieces are written unanimously, with nods by the authors within the article to their being retired from the military in some capacity.
+
+# Methodological Approach and First Level-Analysis
+
+To test my hypotheses, I conduct a mixed-methods approach that attempts to look for evident trends and patterns among the data.
+I plan to conduct two steps of analysis.
+The first is to categorize each observation by subject, and the second is to categorize each observation in terms of its deviation from traditional civil-military norms of the military as remaining an apolitical institution.
+
+In the first level of analysis, after capturing key biographical and publication information, such as author name, article title, source, publication date, and branch of service and rank of the author, I code each observation into one of several binary categories that corresponds to the overall subject of the publication.
+The goal is to create a coding scheme by subject that is sufficiently broad but also specific enough that is helpful in differentiating the overall main point of each publication.
+The following coding scheme is therefore devised and implemented (all variables are binary, taking on the value of 1 if the publication corresponds to the general definition of the variable, and 0 otherwise):
+
+\singlespace
+
+  * **PolicySecurity**- if the piece advocates for or denounces a particular course of action directly related to security or operational or strategic decisions of a clear military nature.
+Examples: the wisdom of troop increases, the Surge in Iraq, etc.
+  * **PersTroopSupport**- if the piece generally showcases or highlights the work and sacrifice of members of the armed forces, or to defend the work of specific members of the Armed Forces.
+Examples: remembering our fallen during Memorial Day.
+  * **GenAdvice**- if the piece is generally written to offer some sort of strategic advice for the nation. Examples: why the US needs to strengthen NATO, or build ties with China, etc. 
+  * **PolicySocial**- if the piece advocates or denounces a particular social policy for adoption by the Armed Forces.
+Examples: "Don't Ask, Don't Tell" policy, Transgender members of the armed forces, etc.
+  * **Party**- if the piece explicitly denounces or advocates a position that is explicitly associated with the platform position of a major political party.
+Examples: why the Republicans/Democrats are wrong on this particular issue, etc.
+  * **PolAdminName**- if the piece is advocating for or denouncing a particular position held by a political administration, or if the piece is explicitly defending or attacking a political figure by name.
+Examples: endorsing a Presidential candidate for office, attacking a President for a decision that was made, etc.
+
+\doublespace
+Separating opinion-editorial publications into one of several distinct categories is an imperfect process, and for purposes of this research project, are completed by hand.
+Two points are worth mentioning.
+First, I attempt to keep in mind the fact that opinion pieces are intended to argue for something, and thus are, to an extent, designed to be provocative.
+With this in mind, I did not take phrases such as "President Bush or Obama should do x and y in Iraq and Afghanistan" as attacks on administrations, and therefore, did not did not code such articles under the variable **PolAdminName**.
+Every attempt was made to read each article slowly and as objectively as possible, and afterwards, to assign the publication to a category.
+There were instances in which I had to choose between placing a publication into one of two categories, say, between **PolicySecurity** and **GenAdvice**, but very few if any instances in which I had to choose between categories that differed greatly, such as **PolAdminName** and **PersTroopSupport**.
+Table 1 reveals the breakdown of the data by article type.
+
+```{r summary, include = FALSE}
+# Read in the Data
+dataset <- read_csv(here("data", "MilEditorials.csv"))
+dataset %>% print()
+# Replace NA with 0 in the data for variables PolicySecurity thru PolAdminName
+adjdataset <- dataset %>%
+  mutate_at(
+    .vars = vars(PolicySecurity:PolAdminName),
+    .funs = ~ case_when(is.na(.) ~0, TRUE ~ 1)
+) %>% mutate(
+    PubYr = lubridate::year(PubDate)
+) %>% view()
+
+summary(adjdataset)
+names(adjdataset)
+
+# count the number of observations by type of binary variable and store as new dataframe
+countbytype  <- adjdataset %>% count(PolicySecurity, PersTroopSpt, GenAdvice, PolicySocial, Party, PolAdminName)
 
@@
 
-First, I import the data set and using the here () function to build the file path.  
-I replace all NA/s with ) in the data set.
-
-```{r summary}
-# Read in the Data
-dataset <- read_csv(here("data", "MilEditorials23Apr.csv"))
-# Replace NA with 0 in the data
-dataset[is.na(dataset)] <- 0
-
-# Let's view the data so we know what kind of dataframes each of the variables are.  
-dataset
-
-# Note that I may want to convert AuthSe, AuthRank, and Service to some form of numeric.  
-# But I will come back to that.
-
-#See the variable names.  
-summary(dataset)
-names(dataset) 
+```{r table, fig.align="center", include=TRUE, echo=FALSE}
+countbytype %>%
+  kable(format="latex", caption="Publications by Article Type") %>%
+  kable_styling(latex_options = "striped") %>% 
+  row_spec(0, bold=T)
 ```
 
-```{r manipulation}
+Table 1 reveals that an overwhelming majority (186 of 216, or 86%) of the total number of opinion pieces written since 2000 in the four major newspapers examined can be classified according to one of the following three categories: **PolicySecurity**, **PersTroopSupport**, and **GenAdvice**.
+The remaining 30 publications (14% of the total) fall into one of the other three categories: **PolicySocial**, **Party**, and **PolAdminName**.
 
-# Create a Variable for Publishing Year called PubYr
-adjdataset <- dataset %>% 
-  mutate(
-  PubYr = year(PubDate)
-) %>% print()
+I consider these 186 observations as being in alignment with traditional civil-military norms on the basis that they address topics which American citizens can reasonably expect retired military officers to address.
+In other words, I consider retired military officers writing on topics related to security policy, support for members of the Armed Forces, and general advice of a strategic nature not to be problematic.
+In line with Zaller's framework of elite discourse helping to frame public opinion, these 186 observations demonstrate military elites writing on areas of policy in which they reasonably possess considerable expertise.
+The remaining 30 observations - those that have been classified as **PolicySocial**, **Party**, or **PolAdminName**, require further examination.
+There may be nothing inappropriate from the perspective of traditional civil-military norms about these remaining observations, but they have been flagged on the basis of the topics that they address.
+
+# Second-Level Analysis and Results
+
+```{r remaining data, include = FALSE}
 
 # Show all articles and authors who published articles on Social Policy, Political Parties, and Administrations/Political Figures
-adjdataset %>% arrange(PubYr) %>% 
-  filter(PolicySocial==1 | PartyGen==1 | PartyName ==1 |PolAdminName==1 )  %>% 
-  select (AuthName, ArtTitle, Source, PubYr) %>% 
+rem30_obs <-adjdataset %>% arrange(PubYr) %>% 
+  filter(PolicySocial==1 | Party==1 | PolAdminName==1 )  %>% 
+  filter(AuthName != "UNK") %>% filter(AuthName != "Murtha, John") %>% 
+  select (AuthName, ArtTitle, Source, PubYr, PolicySocial, Party, PolAdminName) %>% 
   view()
+```
 
+```{r table2, fig.align = "center", include=TRUE, echo=FALSE}
+rem30_obs %>%
+  kable(format="latex", caption="Remaining Observations") %>%
+  column_spec(1, width="10em") %>% 
+  column_spec(2, width="13em") %>% 
+  column_spec(3:4, width="2.3em") %>%
+  column_spec(5, width="4.9em") %>%
+  column_spec(6:7, width="2.4em") %>% 
+  kable_styling(latex_options="striped", font_size=9) %>% 
+  row_spec(0, bold=T)
+```
+
+```{r code, include=FALSE, echo=FALSE}
 # Create a Variable Called case-count that counts the number of times by publishing year an article published 
 # regarding social policy or general parties or parties by name or administrations occurs
-adjdataset %>% group_by(PubYr) %>% 
+rem30_obs %>% group_by(PubYr) %>% 
   mutate(
-    case_count=sum(PolicySocial ==1 |PartyGen==1 | PartyName ==1 | PolAdminName==1) 
-  ) %>% 
-  print
+    case_count=sum(PolicySocial ==1 |Party==1 | PolAdminName==1) 
+  ) %>% view()
+```
+Of the thirty remaining observations, two are removed upon immediate inspection.
+One was written by an unknown author and is entitled, "Fists Raised at West Point," and was published in the _Washington Post_ on May 13, 2016.
+It is removed because the article is actually a compendium of two separate letters sent to the editor by two different authors, one arguing for and the other against female West Point graduates who raised their fists upon graduating, which ignited some controversy about soon-to-be Army lieutenants purportedly making a symbol associated with the "Black Lives Matter" movement [@noauthor_fists_2016].
+Because the two separate letters were written anonymously, and were clearly not the deliberate work of a retired military officer who wish to be identified, I removed the observation.
+The second piece was written by former Democratic Congressman John Murtha, who was also a retired Marine Corps officer.
+As a politician, the American public can and should expect him to write in that capacity and thus, to make partisan arguments.
+Not wanting to skew the remaining data, I remove his publication for the remaining analysis.
+Thus, the second level of analysis proceeds on a total of 28 observations, which are listed and printed in chronological publishing order, in Table 2.
 
-# Count the Number of Articles Published By Type
-adjdataset %>% tally(PolicySecurity) 
-adjdataset %>% tally(PersTroopSpt)
-adjdataset %>% tally(GenAdvice)
-adjdataset %>% tally(PolicySocial)
-adjdataset %>% tally(PartyGen)
-adjdataset %>% tally(PartyName)
-adjdataset %>% tally(PolAdminName)
+A couple of interesting points are worth noting about the data in Table 2.
+First, one can quickly observe that there appear to be two distinct clusters of observations with respect to publishing year.
+The first occurs in the time frame between roughly 2004 - 2007, and the second between 2016 - present day.
+Of the 28 observations listed in Table 2, there are only five observations - one each year in 2003, 2010, 2013, 2014, and 2015, that were not published within these two time periods.
+The second interesting observation is that of the editorials published between 2016 - present day in Table 2, all fifteen were published in either the New York Times or the Washington Post, and none were published in the Wall Street Journal.
+The clusters can also be seen in time in Figure 1.
 
-# need to change the type of class these vectors are - to binary
-
+```{r graph, fig.height=4, fig.width=5, fig.align="center", include=TRUE, echo=FALSE}
+datatograph <- data.frame(
+  year=c("03", "04", "06", "07", "10", "13","14", "15", "16", "17", "18", "19", "20"), count=c(1, 3, 3, 2, 1, 1, 1, 1, 1, 4, 3, 6, 1)
+)
+ggplot(datatograph, aes(x =year, y =count)) + geom_bar(stat = "identity", width=.2) +
+  labs(
+    x = "Year",
+    y = "Count",
+    color = NULL,
+    title = "Figure 1: Op-Eds Written By Military Retirees",
+    subtitle = "On Topics About Social Policy, Parties, or Administrations"
+  )
 ```
 
-```{r simple graphs}
+To truly determine whether there have been any normative changes in retired military op-ed usage requires a more qualitative approach.
+To that end, I re-read each of the 28 observations listed in Table 2 and attempt to determine the "appropriateness" of the op-ed.
+I use the military's own espoused beliefs about the necessity of remaining an apolitical institution to help develop another simple coding mechanism.
+Recall that Retired Army General Douglas MacArthur cautioned the cadets at West Point in 1962 about involving themselves in matters that do not require a military solution (problems of "politics grown to corrupt," "morals grown too low," and "whether our personal liberties are as firm and complete as they should be") [@macarthur_duty_1962].
+And recall more recently, in 2016, that Retired Army General Martin Dempsey lamented the participation of retired generals in the political process by endorsing political candidates and appearing at the conventions of political parties.
+A section of Dempsey's opinion piece published in the Washington Post bears repeating:
 
+\singlespace
+
+> **The American people should not wonder where their military leaders draw the line between military advice and political preference.
+And our nation's soldiers, sailors, airmen and Marines should not wonder about the political leanings and motivations of their leaders.
+As generals, they have an obligation to uphold our apolitical traditions.
+They have just made the task of their successors – who continue to serve in uniform and are accountable for our security - more complicated** [@dempsey_military_2016].
+
+\doublespace
+
+MacArthur is exhorting future military leaders to stay in their lanes and focus on military problems as opposed to political ones, while Dempsey is explicitly calling for active and retired military not to engage in expressly political activities.
+Combined with the arguments made by Samuel Huntington, who called for military leaders to be aware of the opposing forces inherent in the functional and societal imperatives that impact the military, as well as the fact that no prominent scholar of civil-military relations calls for an increase in participation of the military into political waters, allows us to develop a reasonable and clear coding scheme for the second level of analysis, with emphasis on the two characteristics of topic (is the author writing about a reasonable military topic) and political leaning (is the author expressly attacking or supporting specific political figures, from which one can reasonable infer the officer's political bent)?
+
+Using these characteristics as guidelines, I therefore develop two new variables.  The first is **Result**, which can take on the following values:
+
+\singlespace
+
+  * **Permissible** -if the reader cannot clearly draw conclusions regarding the author's political stances, and if the author is writing on a topic that is directly related to national security 
+  * **Questionable** -if the reader can draw some general conclusion regarding the author's political stance concerning support for or denouncement of sitting or prospective political candidates, or if the the author writes on topics that are indirectly related to national security (Note: I am aware that one can try and include just about any topic under the umbrella of national security.
+I take a reasonable but clear approach to this, in line with the sentiment expressed by MacArthur and the spirit of Huntington)
+  * **Inappropriate** -if the reader can clearly draw conclusions or determine the military officer's political stance; or if the author attacks or supports a sitting or prospective political figure; or if the author addresses topics that are clearly outside of the realm of national security, thus commenting on topics that are outside the officer's area of expertise as a retired military officer
+
+\doublespace
+The second new variable I develop is **Justification** in which I briefly comment on why I assigned the value of **Result** for the observation.
+I simply comment on whether concerns regarding the topics addressed and/or political leanings are raised (I note N/A for those I deem permissible).
+The results of this second level of analysis are presented in Table 3.
+
+```{r second level analysis, include = FALSE}
+
+# Present Results of Result and Justification
+secondlayer_obs <-adjdataset %>% arrange(PubYr) %>% 
+  filter(PolicySocial==1 | Party==1 | PolAdminName==1 )  %>% 
+  filter(AuthName != "UNK") %>% filter(AuthName != "Murtha, John") %>% 
+  select (AuthName, ArtTitle, Source, PubYr, Result, Justification) %>% 
+  view()
 ```
+
+```{r table3, include=TRUE, echo=FALSE}
+secondlayer_obs %>%
+  kable(format="latex", caption="Appropriateness of Opinion Pieces Authored By Retired Military Officers in Major US Newspapers, 2000-2020") %>%
+  column_spec(1, width="10em") %>% 
+  column_spec(2, width="14em") %>% 
+  column_spec(3:4, width="2.4em") %>%
+  column_spec(5, width="6.5em") %>%
+  column_spec(6, width="8em") %>%
+  kable_styling(latex_options="striped", font_size=9) %>% 
+  row_spec(0, bold=T)
+```
+
+Table 3 reveals several interesting points regarding the clusters previously identified in Table 2.
+Eight of the nine opinion pieces published between 2003 and 2007 are take values of **Result** as either questionable or inappropriate.
+Further, note that the six coded as inappropriate take on values of **Justification** revealed a clear political leaning.
+Further analysis indicates:
+
+\singlespace
+* Three publications support either President George W. Bush ("War of Words" and "Right Leader, Right Time) or Senator John Kerry ("Medals of Honor") ahead of the 2004 Presidential Election   
+* Three publications either attack ("A Top-Down Review of the Pentagon", "An Army of One Less") or defend ("In Defense of Donald Rumsfeld") then-Secretary of Defense Donald Rumsfeld^[
+Note that the so-called "Revolt of the Generals" occurred in the middle of this time period, referring to the way in which several retired generals and admirals vocally spoke out against the manner in which Secretary Rumsfeld had dealt with the military, particularly in the lead up to the war in Iraq [@kagan_let_2006].
+]
+
+\doublespace
+Despite the fact that these opinion pieces violated traditional civil-military norms, there is a degree of homogeneity in that these inappropriate pieces all revealed clear political preferences, either by endorsing Presidential candidates, or by condemning figures on the basis of what they perceived to be poor leadership by a senior government official related to combat operations.
+However, the second cluster of publications, written between 2016-2020, reveal a vastly different picture.
+Four results are worth noting.
+
+First, this second cluster is larger than the previous one examined  (15 publications total versus nine published between 2003-2007).
+Second, there is a relative increase in the number of publications that take values of questionable under the variable **Result** (8 of 15, or 53%, compared to 2 of 9, or 22.2% in the first cluster), and a proportionate decrease in the number of inappropriate publications (5 of 15, or 33%, compared to six of nine, or 66%, in the first cluster).
+Third, there is a sharp increase in the proportion of publications in this cluster that take values of questionable or inappropriate in **Result** due to concerns about the _topics_ that are being addressed in the particular opinion pieces.
+Indeed, ten of the thirteen publications that take the value of questionable or inappropriate under **Result** correspond to values of **Justification** that indicate the publication addressed a topic that is either indirectly related to national security, or clearly outside the realm of national security.
+Fourth, three of the pieces coded inappropriate under **Result** take on values of **Justification** of "Topic, Political Leaning," which indicates these publications revealed _both_ a  clear political leaning _and_ addressed a topic that is not directly related to national security.
+The third and fourth results deserve to be unpacked in further detail.
+
+## Increasing Diversity of Topics Addressed By Military Officers
+
+Some examples will help show how some of the how the topics of editorial in this second cluster may deviate from national security issues (the title of the article in Table 3 may also help the reader).
+Army General (Retired) Stanley McChrystal's _Washington Post_ editorial, "Home Should Not Be a War Zone," comments on the tragic nature of mass shootings in America.
+He states in the article, "...as a combat veteran and proud American, I believe we need a national response to the gun violence that threatens so many of our communities" [@mcchrystal_home_2016].
+In an article published in the _Washington Post_ in 2018 entitled "Good Riddance," McChrystal shares how his beliefs in Robert E.
+Lee had evolved over the course of his professional career, and why he now believes it is time to discard Confederate statues and other symbols.
+Admiral (Retired) Mike Mullen and co-authors Lieutenant General (Retired) Mark Hertling and Admiral (Retired) Robert Natter wrote publications ("The Refugees We Need" and "Cutting Refugee Admissions Will Have Severe Consequences for the US Military") explaining how President Trump's immigration policies on refugees harm military readiness.
+
+While many Americans agree with the stances taken by these retired military officers, the reality is that these topics - domestic violence, the fate of Confederate symbols, and policy stances regarding refugees - are not _directly related_ to national security policy.
+This is especially true according to the words and ethos espoused by retired Generals MacArthur and Dempsey previously examined.
+No doubt that all of these problems, and more, are pressing concerns, and issues which America needs to solve.
+However, it is not clear that the expertise possessed by retired military officers directly translates into helping to solve these issues, or that retired military officers are the right elites to weigh in on how to tackle these concerns.
+By way of comparison, when compared to publications in the first cluster published between 2003-2007, only one took a value of **Justification** due to concerns about the topic being addressed ("A Multicultural Military").
+There is therefore clear evidence that in recent years, retired military officers have addressed topics in opinion-editorial pieces that are more diverse and further from the realm of traditional national security policy than earlier retirees.
+
+## Increasing Boldness in Criticizing Administrations for Non-Warfighting Reasons
+
+Next, let us briefly examine the publications that are part of cluster 2 which were deemed inappropriate or questionable due to concerns about revealing political leanings or preferences.
+Three of these articles, "Take My Security Clearance, too, Mr.
+President," "Our Republic is Under Attack from the President," and If Good Men Can't Speak the Truth, We Should Be Deeply Afraid," were all authored by Admiral (Retired) William McRaven, and as their titles suggest, they are fairly obvious expressions of disdain for President Trump and/or his administration.
+Three other articles, "Trump's Border Stunt is a Profound Betrayal of Our Military," "Retired Generals Warned Us About Rumsfeld.
+Now They're Warning Us About Trump," and "Banning Transgender Troops Only Hurts Us," are likewise fairly clear expressions of disdain either about President Trump himself or policies which his administration has adopted.
+It would be difficult to classify any of these particular opinion pieces as apolitical and thus, in alignment with historical civil-military norms.
+
+As noted previously, the publications from cluster one which displayed clear political leanings either expressed support one of the candidates in the 2004 Presidential Election, or took a stance regarding Secretary Rumsfeld's abilities regarding the Iraq War.^[
+The publication, "Medals of Honor" was authored by retired General and former Democratic Presidential Candidate Wesley Clark.
+The fact that this piece was authored in April 2004, more than two months after he had dropped out of the race for President, indicates that he was writing from the perspective of a retired military officer.
+] The publications from cluster two that adopt a clear political tone do not attack President Trump for how he asserts himself in military affairs nor do they (yet) support or attack his reelection opportunities.
+Instead, the publications from cluster two that display a clear political leaning either attack the President's policies or his personal character.
+Like those that display clear partisan leaning from cluster one, these pieces clearly violate civil-military norms, but they do so in a manner that is qualitatively distinct from those which were published between 2003-2007.
+
+Thus, revisiting the hypotheses of the project, the following conclusions can be reached.
+The first hypothesis is confirmed - that the majority of op-eds published by retired military officers are in fact apolitical in tone.
+In total, 195 of 216 (90.2%) pieces used in this study were found to be within historical civil-military norms.
+The second hypotheses, that there would be an increase in the number of op-eds criticizing political administrations regarding the conduct of war in Iraq and Afghanistan, is found to be untrue.
+In fact, the data reveal the opposite: there is a decrease in the number of opinion pieces which criticize political administrations over the conduct of American wars.
+Rather, the data revealed that there is a sharp increase in the number of op-eds which address topics that are not clearly related to national security as well as an increase in the number of op-ed pieces which attack a political administrations, namely that of the current President.
+Further, the data revealed that these attacks and criticisms from retired military officers stem not from his or his administration's waging of war, but rather out of concern for his character and/or policies.
+
+These trends are potentially problematic from the perspective of civil-military relations for several reasons.
+First, they reveal that in recent years, the perception of military expertise is expanding into areas that fall outside of the scope of what has been traditionally considered to be true military expertise.
+If retired military officers are writing opinion pieces in major newspapers on topics that do not directly address national security, or topics that are the reasonable jurisdiction of elites with different backgrounds, this could indicate that either these retired military officers, specific media outlets, or American citizens are tapping into the high levels of institutional confidence of the military as an institution in order to argue for or against certain issue areas.
+Each of these scenarios would be troubling as it would indicate a manipulation of institutional confidence for political or private gain.
+
+Strictly from the perspective of the retired military officer and the military as an apolitical institution operating in polarized political times, it will require more caution - not less - to maintain its stance as a nonpartisan.
+This is easier said than done, and the fact is that as leaders who have to exercise moral agency in their daily responsibilities, military officers may act or behave in ways that can readily be perceived through partisan lenses among a public that is polarized.
+However, retired military officers, though still members of the military as an institution, can likely exercise increased caution even more readily than the active duty officer corps.
+This is especially true when engaging with the media, and certainly considering the publication of opinion-editorial pieces, which are deliberate and intentional instruments designed to impact public opinion.
+
+# Counterarguments
+
+One normative counterargument to the findings laid out in this paper is that Huntingtonian logic, and specifically, the concept of "objective civilian control" and the differentiation of separate civilian and military spheres, is outdated and no longer relevant, especially in light of the inconclusive wars fought in Iraq and Afghanistan [@huntington_soldier_1957].
+Proponents of this school point to the fact Huntingtonian norms can actually foster poor strategic thinking and that, to a certain degree, the military as a profession is to some degree an inherent political entity in that it has to lobby on its own behalf [@brooks_paradoxes_2020].
+Proponents of this school raise many valid concerns, particularly concerning the efficacy with which senior military and civilian leaders are able to form a constructive dialogue that together, enables the nation to achieve its strategic goals.
+No doubt this is a concern that both civilian and military leaders should address, and perhaps that does require a merging, at least to some degree, of the "separateness" of the professional operating spheres that exist under Huntingtonian norms.
+However, in the effort to explore new paradigms in how to structure and manage effective civil-military relations, there should be no result that enables the military to be less subordinate to the overall control of civilian leadership than it currently is.
+Indeed, I agree and join in on those who have made calls for developing senior military leaders who are capable of entering into a more constructive dialogue with civilian leadership for the purposes of implementing better strategy [@rapp_civil-military_2015].
+However, it is unreasonable to think that improvements cannot be made without significantly altering the military's apolitical norms.
+
+Another counterargument to the findings in this paper involves the role of retired military officers vis-a-vis those who are actively serving.
+Shouldn't military retirees have more liberty and freedom to express themselves and their opinions than those who wear the uniform?
+In other words, proponents of this counterargument may hold to the belief that active military officers ought to remain apolitical, but that the rules should be different with respect to those who no longer wear the uniform.
+Ultimately, the answer to this question must be determined by the American public.
+Yet is is notable that when General (Retired) Dempsey lamented the involvement of Generals (Retired) Allen and Flynn in the 2016 Democratic and Republican conventions, he stated, "They have just made the task of their successors – who continue to serve in uniform and are accountable for our security - more complicated" [@dempsey_military_2016].
+What did he mean when he said this?
+I think at the heart of what he said is the recognition that in the military, perhaps more than in any other institution or profession, those who serve truly stand on the work accomplished by those who have passed before them.
+Out of today's lieutenant colonels and colonels will come the next generation's three and four star generals and admirals.
+From a sociological perspective, these future leaders very much watch, observe, and adopt the behaviors of those they look up to.
+Said differently, I think Dempsey had in mind the notion that for the military officer who retires, his or her subordinates still serving always look to him or her as the officer, for that is how the individual was known to them.
+In other words, Dempsey is arguing that for the sake of those still serving and who knew you and who must now advise senior civilian policymakers, much in the same way the retired officer used to do, the retired officers ought to remain loyal and thus, apolitical, to the institution.
+
+A final counterargument or objection to the results in this paper deal with the data used to advance the argument that the nature of opinion pieces authored by retired military officers has qualitatively changed over the past twenty years.
+An objection can be made that the data examined only four newspapers, which comprise a very small percentage of the overall market for media when one also considers television, blogging sites, and other news outlets.
+A related objection can be raised by pointing out that the _New York Times_ and the _Washington Post_ in particular lean to the left politically, whereas the _Wall Street Journal_ is more of a centrist or even slightly right-leaning source [@noauthor_allsides_nodate], and therefore, that these sources are more likely to present liberal material than conservative arguments.
+I acknowledge these counterarguments and argue that more work should be done to gather opinion pieces from more diverse sources to include in the data for further analysis.
+It is, after all, more likely that a network such as Fox or a publication such as the _Washington Examiner_  would potentially display editorials that advance conservative causes.
+While having the data from these and other sources would help researchers make broader comparisons, I can still stand confidently by the results of this paper in that the comparison over time was completed over the four same sources.
+Moreover, I argue that the fact that specific trends occurred over a time period where overall newspaper circulation dropped sharply further enables the results reached to stand up to scrutiny.
+
+# Conclusion
+
+This paper examined how retired military officers use opinion-editorial pieces to persuade the opinion of the American public of certain issues.
+Specifically, the goal was to determine if there had been noticeable trends in how retired military officers who penned opinion publications showed any deviation from the traditional norm of the military remaining an apolitical institution.
+While the data showed that the vast majority (90%) of opinion pieces authored by retired military officers and published in America's four leading newspapers aligned with traditional civil-military norms, two recent trends emerged from the data, both of which should be further explored.
+
+First, there is a widening of topics which military officers are addressing in their works.
+This could reveal that recent retirees themselves feel more qualified than those who retired twenty years ago to address topics that are either tangentially or not at all related to national security, or that the media is more willing to publish works by retired military on these subjects, or that the American public is willing to receive and listen to these officers argue for and against these topics, or some combination of these possibilities.
+Regardless, this trend indicates that traditional military expertise is perceived to have widened, which is possible considering the high degree of institutional confidence the military institution enjoys.
+
+Second, the data indicates an increased willingness by retired military officers to criticize political figures or administrations, and notably, for reasons other than the figure's or administration's performance related to war fighting.
+Rather, retired military officers are attacking the current President and his administration for policies and stances on issues such as immigration, refugees, and the general political environment that prevails in the country.
+This trend differs from that which existed between 2003-2007, when military retirees displayed clear political preferences ahead of the 2004 Presidential Election, or regarding the conduct of the war in Iraq by Secretary of Defense Donald Rumsfeld.
+
+These two trends indicate that retired military officers are violating traditional civil-military norms, as articulated by retired Generals MacArthur and Dempsey, in ways that differed from previous military retirees.
+Their continuing to do so, especially in a contentious domestic political environment, stands to weaken the military as an institution, or perhaps worse, to manipulate the institutional confidence the military currently enjoys among the American public, for private or political gain, even if that is not the true intention of these retired officers.
+
+\singlespace
+\pagebreak

### 2020-06-27 code/Submissionv1.Rmd
commit

--- 2020-05-12 code/PS904-Project.Rmd
@@
 ---
-title: "The Pen as a Sword: Retired Military Officers and Opinion-Editorial Usage"
-author: "Pete Erickson"
+title: "When the Pen Becomes a Sword: Retired Military Officers and Opinion-Editorial Usage in the US"
+author: "anonymous"
 date: |
@@
 abstract: |
-  _This paper seeks to explain how retired military opinion-editorial usage has varied in the past 20 years.
-Using an original data set that captures more than 200 opinion-editorial pieces published by retired military officers in The Wall Street Journal, The Washington Post, The New York Times, and the USA Today since 2000, I argue that recent opinion-editorial usage is weakening the norm of retired military officers remaining apolitical in two distinct ways.
-First, op-eds published in recent years by retired military officers address topics that are outside the scope of direct security or strategic concerns.
-Second, there is an increase in recently published op-eds which directly attack Presidential administrations, specifically that of Donald Trump.
-These two trends are concerning from the perspective of civil-military relations and raise further normative questions about how the American public receives information from military elites._
+  _This paper seeks to explain how retired military officers have attempted to shape US public opinion through the publication of op-ed pieces in major newspapers.
+I construct and analysze a data set that captures more than 300 opinion-editorial pieces published by retired military officers in The Wall Street Journal, The Washington Post, The New York Times, The Los Angeles Times, and the USA Today since 1975, and ultimately find two trends that stand out.
+The first is that retired military officers are increasingly writing about topics that are outside of traditional security related topics.
+The second trend is that retired military officers are increasingly criticizing Presidential administrations, especially that of President Donald Trump, for reasons other than disputes regarding international uses of force.
+Given the polarizing political climate in the United States and that a Presidential election will occur this fall, these two trends are concerning from the perspective of civil-military relations and suggest that the norm of a professional, apolitical military establishment is under significant strain._
 
@@
 
-In modern day America, the average citizen does not need to look or go far to hear or see an elite voice attempting to persuade them of the rightness or wrongness of certain courses of action.
-On the television and on the internet, "experts" in various forms take up a position on a certain issue and attempt to persuade, warn, and otherwise impact public opinion on a host of issues.
-To a considerable extent, this is a normal and everyday occurrence in a thriving democracy.
+On June 1 2020, the President of the United States, Donald Trump, after giving remarks at the Rose Garden of the White House in Washington D.C., walked to St.
+John's Church after security forces visibly displaced throngs of protestors in neighboring Lafayette Square.
+Flanked by members of his cabinet, to include the Secretary of Defense, and with the Chairman of the Joint Chiefs of Staff not far behind, the President held up a Bible at the church, stared into the cameras of the media, and after several minutes, returned to the White House.
 
-This paper examines the way in which a narrow subset of elites - retired military officers - has attempted to influence public opinion in contemporary America.
+In the aftermath of what the American mainstream media alleged had amounted to nothing more than a publicity stunt, several members of the President's cabinet members and advisors sought to distance themselves from the President's actions, and more broadly, the President's threat of invoking the Insurrection Act of 1807 to use active-duty forces to quell domestic rioting.
+The Secretary of Defense, Dr.
+Mark Esper, stated that he "did not support" use of the 1807 Insurrection Act to deploy active duty military to establish order, and General Mark Milley, the Chairman of the Joint Chiefs of Staff, later told a graduating class from the National Defense University that he "should not have been [joining the President on the walk to St.
+John's Church]," and that his presence "created a perception of the military involved in domestic politics" [@esper_secretary_2020; @milley_official_2020].
+
+Multiple retired American generals also took to the press to denounce the Trump administration's responce to ongoing protests and riots occurring the aftermath of George Floyd's death in Minneapolis.
+But for those paying close attention to these retired military elites, it is clear that not all who spoke out did so uniformly.
+Across a spectrum, there were some retired officers who acknowledged that what had occurred was, at least in a general sense, negative in terms of its impact on civil military relations, while there were also several other officers who went farther in their criticism, directly blasting the President or members of his cabinet for actions taken.
+
+This paper examines the way in which a narrow subset of elites - retired military officers - has attempted to influence public opinion in contemporary America through the use of opinion-editorial pieces.
 It is an interesting question worthy of examination for several reasons and from a variety of perspectives.
@@
 Second, it is an interesting question because as an institution, the military prides itself on being apolitical.
-In an era of increasing political polarization [@layman_party_2006; @hare_polarization_2014], examining the extent to which opinion-editorial usage by retired military officers has changed may reveal one way in which the military institution has altered its behavior in a polarized political environment.
-
-To be clear, I do not examine in this paper the efficacy of the opinion pieces authored by retired military officers.
-I do not address whether the actual op-ed pieces published by retired military officers actually sway public opinion.
-I am primarily concerned with what trends are evident in opinion-editorial publication from the perspective of the military as an institution: what variation is evident over the past 20 years, if any?
-Are there patterns that indicate deviation from historical norms and expectations of the military as an apolitical institution, and if so, what might this mean for the military?
+In an era of increasing political polarization [@layman_party_2006; @hare_polarization_2014], examining the extent to which opinion-editorial usage by retired military officers has changed may reveal one way in which the military institution has altered its behavior in a polarized political environment, and especially, in the months leading up to the Presidential Election in November of this year.
 
 This paper proceeds in three parts.
-First, I review the basis upon which the norm of apolitical behavior of the military as an institution is based, as well as describe how a polarized environment may impact these norms.
-Here I review academic literature on civil-military relations and public opinion, as well as current military doctrine.
-All three are relevant strands that inform our research question.
+First, I briefly review the basis upon which the norm of apolitical behavior of the military as an institution is based, as well as describe how a polarized environment may impact these norms.
+Here I review some of the academic literature on civil-military relations and public opinion, as well as what current military doctrine states on the topic.
 Second, I describe the analytical approach and the data used in this project.
-Assembling an original data set that includes more than 200 opinion pieces published by retired military officers in major American newspapers since 2000, I conduct a mixed-methods approach.
-Finally, I describe my results as well as their implications.
-Overall, the data reveals that the vast majority of opinion pieces penned by retired military officers are in line with traditional civil-military norms.
-However, I find two trends that may reveal future problems for civil-military relations: first, retired military officers are writing about topics that are not immediately related to national security, and second, retired military officers are criticizing if not attacking political administrations more often than they did in previous eras.
+Assembling an original data set that includes more than 300 opinion pieces published by retired military officers in major American newspapers since 1975, I conduct a mixed-methods approach which ultimately reveals two important findings: the first is that although the vast majority of opinion pieces published in major newspapers seem to reflect the military as a partisan institution, retired military officers are increasingly writing about topics that are not immediately related to national security, and second, retired military officers are criticizing if not altogether attacking political administrations more often than they did in previous eras.
+Third, I briefly explore the implications of these results on the military as an institution, and specifically, its ability to remain professional in a highly polarized political climate.
 
-# The Norm of Remaining Apolitical
+# The Military as an Apolitical Institution
 
@@
 In a thriving democracy such as the US, an apolitical military can best serve the Republic by defending it regardless of which political party of group holds power.
-Unlike autocratic societies, where military coups are unfortunate yet frequent in history, the United States had not had a civil-military relations crisis that has risen to the point that a coup has been a real concern.
-This is not to say that that civil-military crises are infrequent in the US - indeed, they have been rather common - but that these crises have typically been resolved.
-
-The political scientist Samuel Huntington, in his work, _The Soldier and the State: The Theory and Politics of Civil-Military Relations_, argued for civil-military relations to take on a certain character.
-Written in 1957, during the Cold War, Huntington asserted that two distinct forces would always be in tension with one another in a liberal democracy.
+One of the strongest origins of the norm of the military as an apolitical institution stems from the work of the political scientist, Samuel Huntington.
+Written in 1957 at the height of the Cold War, his book, _The Soldier and the State: The Theory and Politics of Civil-Military Relations_, asserted that two distinct forces would always be in tension with one another in a liberal democracy such as the United States.
 The first was a "functional imperative," which Huntington said drove the requirement that a military be strong enough and capable of defeating the military threats it faced [@huntington_soldier_1957].
@@
 Huntington's point was that America needed a strong military that could win on the battlefield, but cautioned that such a military could not be so foreign in character so as to alienate the tenets of an American liberal society, and that these forces were at odds with each other.
-Huntington proceeded to argue that the key to resolving the tension between these two opposing forces was to adopt a scheme of "objective civilian control" in which political leaders established the ends and goals of policy, while military leaders developed and implemented plans to accomplish these goals [@huntington_soldier_1957].
-To Huntington, separate operating spheres between the political leaders of a democracy and the military as an institution ensured that the military remained "professional," and thus, untainted by political interests and concerns.
-In this way, Huntington posited that the American military would be up to the task of defeating the very real threats it would face, especially during the Cold War context in which Huntington wrote.
+The key to resolving the tension generated by these two opposing forces was to adopt a scheme of "objective civilian control" in which political leaders established the ends and goals of policy, while military leaders developed and implemented the military plans to accomplish these goals [@huntington_soldier_1957].
+To Huntington, largely separate operating spheres between the political leaders of a democracy and the military as an institution ensured that the military remained "professional," and thus, untainted by political interests and concerns.
+In this way, Huntington posited that the American military would be up to the task of defeating the very real threats it would face while also remaining apolitical.
 
-While Huntington's theory has enjoyed tremendous staying power (Huntington is still studied and taught in many American military institutions), he is not without critics.
-It would be too burdensome to unpack all of those here, but for clarity, several quick counterarguments to Huntington are worth highlighting.
-Eliot Cohen, in his work, _Supreme Command,_ argues that the very concept of separate spheres between political and military leaders is flawed.
-Pointing to leaders such as Lincoln, Clemenceau, and Churchill, and the actions that each took to ensure that the military operated in accordance with each's wishes, Cohen argues that political leaders are ultimately responsible for the outcomes of war, and therefore, are not obligated to grant the military any sort of distinct operating space [@cohen_supreme_2003].
-Another scholar, Peter Feaver, argues in his work, _Armed Servants: Agency, Oversight, and Civil-Military Relations_ that a principal-agent approach most accurately conveys how civil-military relations work [@feaver_armed_2005].
-American elected civilian leaders, as agents, so argues Huntington, have the responsibility to select military leaders as agents, and to oversee ensuing problems of agency that arise, such as delegating, monitoring, and ensuring the compliance of the military as agents [@feaver_armed_2005].
-The point worth highlighting here is that Cohen and Feaver argue that civilian elected leaders have the right to encroach onto what Huntington would call for, a separate military sphere.
-Note, however, that none of these scholars argue for the opposite - that military leaders should in any way encroach onto or towards the political space.
+While Huntington's theory has enjoyed tremendous staying power (Huntington is still studied and taught in many American military institutions), he is not without critics [see, for example, @feaver_civil-military_1996; @cohen_supreme_2003; @feaver_armed_2005; and @brooks_paradoxes_2020].
+It would be too burdensome to unpack all of those here, but one of the earliest and richest critiques of Huntington, and perhaps one of the most relevant today, came from S.E.
+Finer, whose 1962 book, _The Man on Horseback: The Role of the Military in Politics_, argued that Huntington's concept of professionalism simply was not strong enough to keep the military out of politics [@finer_man_1962, 22].
+Finer's main concern was that of military professionals becoming so professional that they develop the idea that they somehow know better than their elected civilian leadership, particularly on thorny and controversial concerns.
+Indeed, Finer warned that "the moment the military draw this distinction between nation and the government in power, they begin to invent their own private notion of the national interest, and from this it is only a skip to the constrained substitution of this view for that of the civilian government…this purported care for the national interest as defined by the military is indeed one of their main reasons for intervening [@finer_man_1962].
+More important than viewing themselves as professionals, Finer argued that what the military needed in order to stay out of politics was a deep internalization of "the principle of civil supremacy" [@finer_man_1962]^[
+Importantly, neither Huntington's concept of professionalism nor Finer's critique of the concept are arguing for blind obedience by the military to civilian orders and directions, a point that is, in the author's opinion, often missed.
+On the contrary, both Huntington and Finer make room and acknowledge that there are instances when military disobedience is not only required, but justified.
+However, both agree that in general terms, these situation should be rare.]
 
 It is not only in the academic literature that the tradition and worth of an apolitical military have been cultivated.
-Indeed, the operational military - those who prepare for fighting and dying on the battlefield - have also been central in upholding the military's apolitical traditions.
+The operational military - those who prepare for fighting and dying on the battlefield - have also been central in upholding the military's non-partisan traditions.
 Retired Army General Douglas MacArthur, a hero of World War Two in the Pacific who is also well known for the controversy surrounding his conduct in the Korean War and his eventual dismissal by President Truman, told the cadets at West Point in May of 1962.
@@
 Prior to the 2016 Presidential Election, Marine Corps General (Retired) Joseph Dunford, who then served as the Chairman of the Joint Chiefs of Staff, said that the basis upon which the military is able to maintain the public's trust is it's ability to remain "apolitical and neutral" [@dunford_upholding_2016].
-General Martin Dempsey, in an opinion piece published in the Washington Post on July 31, 2016, specifically criticized retired Army Major General Mike Flynn and Marine General John Allen for participating in the Republican and Democratic conventions, respectively, stating that "As generals, they have an obligation to uphold our apolitical traditions.
+General Martin Dempsey, in an opinion piece published in the Washington Post on July 31, 2016, specifically criticized retired Army Major General Mike Flynn and Marine Corps General John Allen for participating in the Republican and Democratic conventions, respectively, stating that "As generals, they have an obligation to uphold our apolitical traditions.
 They have just made the task of their successors – who continue to serve in uniform and are accountable for our security - more complicated.
@@
 
-It is important to keep in mind, however, that although the notion of an apolitical military is the theoretical goal of the institution, the practical record remains far from perfect.
+Although the notion of an apolitical military is the theoretical goal of the institution, the practical record remains far from perfect.
 From MacArthur's clash with President Truman in the early 1950's, to the so-called "Revolt of the Generals" in 2005-2006, military leaders, even those who are retired, have ventured into political territory from time to time.
-What we have established, though, is clear: an apolitical military that performs its military mission well is what enables the American public to put trust and confidence in the military institution.
-Having examined the nature of the norm of the military as an apolitical institution, let us now briefly examine the topics of the media, public opinion, and polarization.
-Each of these topics intersects with the norm of the military as an apolitical institution in a unique way.
+What we have established, though, is clear: a non-partisan military that performs its military mission well has long been the goal of both civil-military relations theorists and what enables the American public to put trust and confidence in the military as an institution.
 
@@
 
-To test whether there has been variation in opinion-editorial usage by retried military officers in the United States, I will examine opinion-editorial pieces published in the four leading newspapers of the United States since January 1, 2000: the _Wall Street Journal_ (WSJ), the _New York Times_ (NYT), the _Washington Post_ (WaPo), and the _USA Today_.
-This date range is important for several reasons.
+To test whether there has been variation in opinion-editorial usage by retired military officers in the United States, I will examine opinion-editorial pieces published in the five leading newspapers of the United States since January 1, 1979: _The Wall Street Journal_ (WSJ), _The New York Times_ (NYT), _The Washington Post_ (WaPo), _The Los Angeles Times_, and _The USA Today_.
+This date range is important, interesting, and rich, as it captures a multitude of changes and debates that have involved the military in this time period, all of which are captured in the dataset constructed for this project.
 
-First, let us consider some of the significant ways in which the military as an institution has changed in the past twenty years.
-Most notably, the post-9/11 US military is a battle-hardened military.
-The wars in Afghanistan and Iraq continue to the present day, and today's recently retired military generals and admirals are, in many cases, veterans of these conflicts several times over.
-Moreover, both of these conflicts have thus far failed to reach conclusive ends.
-Heidi Urben found in a 2009 study that the proportion of military officers who felt that criticizing civilian administrations constituted proper conduct had increased since 1998, reasoning that the likely reason for the change was the frustration encountered while serving in Iraq and Afghanistan [@urben_party_2013].
+Consider first that in terms of external threats and wars, this period includes the final years of the Cold War against the Soviet Union, the first war in Iraq, operations in the Balkans, and the post-9/11 wars in Iraq and Afghanistan, and a host of smaller but important military efforts in places such as Somalia, Haiti, and Grenada.
+Each of these conflicts and wars have themselves varied significantly, including in intensity and and their conclusivity.
 
-Moreover, the military since 2000 has undergone significant social changes.
-The repeal of the "Don't Ask, Don't Tell" Policy, which was instituted under President Clinton in 1992 and prevented homosexual members of the Armed Forces from revealing their sexual preference, was repealed in 2010.
-Then, in 2015, under the Obama Administration, the Department of Defense opened up all military occupational specialties to women, many of which, particularly those in the combat arms and special forces, had previously been closed off to  women.
-These changes in the military are just a subset of the profound change that America as a nation has experienced in the past 20 years.
+Second, the military has undergone significant social and institutional changes in this time period.
+In the late 1970s, the military had ended the draft only somewhat recently (1973).
+The data includes discussion of institutional culture that rocked the Navy following the Tailhook Scandalin 1991, as well as debate over first instituting and ultimately repealing the "Don't Ask, Don't Tell" Policy.
+The data also includes discussions of changes occurring in the Obama Administration, such as the opening of military occupational specialties to women, and frustrations expressed by officers at much of President Trump's foreign policy.
 
-Likewise, the media has also undergone significant changes since 2000.
+Third, the media itself has undergone significant changes since the late 1970s.
 Since the mid-2000s, social media has taken over as the primary means in which a majority of Americans receive their news [@martin_how_2018].
-Print newspapers still exist, even as many of the leading newspapers have taken up digital offerings, but overall circulation is still down compared to previous era even after factoring in the digital options available.
-Estimated total daily newspaper circulation in the US in 2000 was 56 million and 59 million for weekday and weekend editions, respectively, while in 2018, the the same figures were 29 million and 31 million.
-Despite these trends, however, the four newspapers that I examine in this project are the largest of all print newspapers in the United States.
+Print newspapers still exist, but overall circulation is still down today compared to earlier years, even after factoring in expanse of digital options available today.
+As a means of rough comparison, estimated total daily newspaper circulation in the US in 2000 was 62 million and 54 million for weekday and weekend editions, respectively, while in 2018, the the same figures were 29 million and 31 million.
 
-Moreover, while there a multitude of ways for military elites to express their opinions and to influence public opinion, such as appearing participating in an interview on television or on the radio, published opinion pieces, especially those that run in a major newspaper, possess certain qualities.
-First, such types of opinion pieces are deliberately authored for the purpose of informing and persuading the public on a salient or timely topic.
+Given these significant changes involving the American military over the past forty years, and the substantial changes that have impacted the media and the manner in which Americans consume information, several specific observations are worth noting about opinion-editorial pieces.
+The first is that opinion pieces are deliberately authored for the purpose of informing and persuading the public on a salient or timely topic.
+They are different from interviews or commentary that may be asked because the author must take time to put forth his or her position in writing.
 Second, publishing opinion pieces, especially for major newspapers, is a highly competitive process.
-For the newspapers that I examine in this study, it is not uncommon to see opinion pieces published by heads of state, members of congress, governmental figures, and heads of industry and academia.
-So, even though overall newspaper consumption may in fact be down over the past twenty years, there is no evidence that opinion pieces have dropped in popularity.
+For the newspapers that I examine in this study, it is not uncommon to see opinion pieces published by figures that include heads of state, members of congress, governmental figures, and leaders of industry and academia.
 
-Attempting to reconcile the facts that the military prides itself on being a non-partisan and apolitical institution, that military officers with experience in Iraq and Afghanistan are more willing to criticize civilian leaders over the conduct of war, that newspaper circulations (even after accounting for digital offerings) are substantially down over the past 20 years, and that the US political environment is increasingly polarized along partisan lines, I adopt the following hypotheses:
+Attempting to reconcile the facts that the military prides itself on being a non-partisan and apolitical institution, that newspaper circulations (even after accounting for digital offerings) are substantially down over the past 40 years, and that the US political environment is increasingly polarized along partisan lines, I adopt the following hypotheses:
 
-_$H_1$: The majority of op-eds published by retired military officers will be apolitical in tone._
+_$H_1$: The majority of op-eds published by retired military officers will be apolitical and non-partisan in tone, but there will be an increase in the number of op-eds criticizing civilian administrations in the past several years, as polarization has dramatically worsened._
 
-_$H_2$: There will be an increase in the number of op-eds that criticize civilian administration handling regarding the conduct of war in Iraq and Afghanistan._
+_$H_2$: Because overall newspaper consumption has dropped in the past 40 years, there will be fewer opinion pieces over time authored by military officers._
 
@@
 
-I first assemble the data by creating a data set that consists of opinion-editorial pieces published by retired military officers in the four major newspapers previously identified since January 1, 2000.
-I employ a search using the Factiva Database and condition my search on several filters:
-
-  * sources (WSJ, NYT, WaPo, and USA Today)
-  * date range (January 1, 2000 - April 16, 2020)
-  * article type (I select editorial, opinion, and commentary)
-  * perhaps the most difficult filter item, free text that identifies retired military officers as the   authors.
-I include terms such as "retired military officer," "retired from the Army, Navy, Marine Corps, Air Force," "former Chairman of the Joint Chiefs of Staff," "general," "admiral," and so forth.
-This is done in order to capture the biographic description that usually accompanies an editorial or opinion piece, such as "General David Petraeus is a retired Army general and former commander of all forces in Afghanistan."
-
-Factiva allows for the option of selecting print only or an "all-sources" option in the source filter.
-I select the "all-sources" option because I want to capture any online editorials or opinion pieces, in addition to those that are ran in print editions only.
-Though digital offerings have certainly expanded in the past twenty years, including digital offerings will enable a better comparison of the data over the past twenty years given that overall newspaper circulation is down over the same time period.
-Factiva also checks for duplicate articles that are published in different versions of the same newspaper, such as the New York Times and the International version of the New York Times.
-When the same articles appeared in two different versions of the same newspaper, I counted the article only one time.
+To test my hyptheses, I first create a data set that includes all identifiable opinion-editorial pieces authored by retired military officers from January 1, 1979 through June 30, 2020.
+Using the Factiva and Nexus-Uni databases, I filter a search that utilizes appropriate parameters and uses specific keywords in the byline (author) field such as "retired military officer," "retired from the Army, Navy, Marine Corps, Air Force," "former Chairman of the Joint Chiefs of Staff," "general," "admiral," and so forth.
+I also include in my search both print and online offerings to ensure that I am capturing any opinion pieces published digitally.
 
@@
 # Read in the Data
-dataset <- read_csv(here("data", "MilEditorials.csv"))
+dataset <- read_csv(here("data", "MilEditorials24June.csv"))
 dataset %>% print()
@@
   mutate_at(
-    .vars = vars(PolicySecurity:PolAdminName),
+    .vars = vars(Warfighting_Ops:Admin_Fig_Name),
     .funs = ~ case_when(is.na(.) ~0, TRUE ~ 1)
@@
 # count the number of observations by type of binary variable and store as new dataframe
-countbytype  <- adjdataset %>% count(PolicySecurity, PersTroopSpt, GenAdvice, PolicySocial, Party, PolAdminName)
+countbytype  <- adjdataset %>% count(Warfighting_Ops, TroopSpt_Culture, Strat_Advice, MilPolicy_Social, Domestic_Policy, Admin_Fig_Name) 
+countbytype
 
@@
 
-Table 1 reveals that an overwhelming majority (186 of 216, or 86%) of the total number of opinion pieces written since 2000 in the four major newspapers examined can be classified according to one of the following three categories: **PolicySecurity**, **PersTroopSupport**, and **GenAdvice**.
-The remaining 30 publications (14% of the total) fall into one of the other three categories: **PolicySocial**, **Party**, and **PolAdminName**.
+Table 1 reveals that an overwhelming majority (289 of 334, or 86.5%) of the total number of opinion pieces written since 1979 in the major newspapers examined can be classified according to one of the following three categories: **Warfighting_Ops**, **TroopSpt_Culture**, and **Strat_Advice**.
+It is also entirely reasonable to conclude that these op-eds written are appropriate from the perspective of civil-military relations in that they address topics which American citizens can reasonably expect retired military officers to address and to bring their expertise to bear upon.
+In line with Zaller's framework of elite discourse helping to frame public opinion of substantive issue areas, the fact is that the overwhelming majority of opinion articles penned by retired military officers are offered on topics that military expertise can reasonably address based on the expertise they have acquired throughout their careers.
 
-I consider these 186 observations as being in alignment with traditional civil-military norms on the basis that they address topics which American citizens can reasonably expect retired military officers to address.
-In other words, I consider retired military officers writing on topics related to security policy, support for members of the Armed Forces, and general advice of a strategic nature not to be problematic.
-In line with Zaller's framework of elite discourse helping to frame public opinion, these 186 observations demonstrate military elites writing on areas of policy in which they reasonably possess considerable expertise.
+This is not to say that these pieces are entirely without controversy, or to suggest that opinion pieces authored by retired military ought to be free.
+A review of the data clearly shows, on multiple occasions, retired military officers passionately arguing for the adoption of certain courses of action through the American press.
+For example,....
+
+The remaining 45 opinion pieces (13.5% of the total) fall into one of the other three categories: **MilPolicy_Social**, **Domestic_Policy**, and **Admin_Fig_Name**. 
 The remaining 30 observations - those that have been classified as **PolicySocial**, **Party**, or **PolAdminName**, require further examination.

### 2020-07-01 code/Submissionv1.Rmd
made version changes

--- 2020-05-04 code/PS904-Project.Rmd
@@
 ---
-title: The Pen Mightier Than the Sword? Retired Flag Officers and Generals and Opinion-Editorials
-  Since 2000
-author: "Pete Erickson"
-date: "4/22/2020"
+title: "When the Pen Becomes a Sword: Retired US Military Officers and Op-Ed Use in the Contemporary Press"
+author: "Peter M. Erickson"
+date: |
+  `r format(Sys.time(), '%B %d, %Y')`
+abstract: |
+  _This paper seeks to explain how retired military officers have attempted to shape US public opinion through the publication of op-ed pieces in major newspapers.
+I construct and analyze a data set that captures more than 300 opinion-editorial pieces published by retired military officers in The Wall Street Journal, The Washington Post, The New York Times, The Los Angeles Times, and the USA Today since 1975, and ultimately find two trends that stand out.
+The first is that retired military officers are increasingly writing about topics that are outside of traditional security related topics.
+The second trend is that retired military officers are increasingly criticizing Presidential administrations, especially that of President Donald Trump, for reasons other than disputes regarding international uses of force.
+Given the polarizing political climate in the United States and that a Presidential election will occur this fall, these two trends are concerning from the perspective of civil-military relations and suggest that the norm of a professional, apolitical military establishment is under significant strain._
+
+geometry: margin = 1.15in              # margin widths
+fontsize: 12pt                         # text size
+bibliography: researchpaper-ps904.bib # bibliography file
+biblio-style: authoryear               # citation style
+urlcolor: blue                         # links to external URLs
+citecolor: black                       # links to bibliography
+linkcolor: magenta                     # internal figures, footnotes, etc.
+indent: true                          # paragraph format
+                                       # true for indented paragraphs
+
+header-includes:
+    - \usepackage{setspace}
+    - \usepackage{multirow}
+    - \usepackage{indentfirst}
+
 output: 
-  pdf_document: default
-  html_document: default
+  bookdown::pdf_document2:
+    keep_tex: true
+    latex_engine: pdflatex
+    number_sections: false
+    toc: false
+    fig_caption: true
+    citation_package: biblatex
+    includes: 
+      in_header: 
+        - latex-preamble.tex
+  bookdown::htmldocument2: default
+  bookdown::word_document2:
+      toc: false
 ---
+
+\doublespacing
 
@@
 
-```{r}
+```{r, include = FALSE, echo = FALSE}
 library("here")       # file paths
-library("magrittr")   # pipe operators!
 library("tidyverse")  # workhorse package
-library("lubridate")  # lubridate helps with the package
-library("plyr")
+library("tidylog")
+library("kableExtra")
+library("knitr")
+```
+
+# Introduction
+
+On June 1 2020, the President of the United States, Donald Trump, after giving remarks at the Rose Garden of the White House in Washington D.C., walked to St.
+John's Church after security forces visibly displaced throngs of protesters in neighboring Lafayette Square.
+Flanked by members of his cabinet, to include the Secretary of Defense, and with the Chairman of the Joint Chiefs of Staff not far behind, the President held up a Bible at the church, stared into the cameras of the media, and after several minutes, returned to the White House.
+
+In the days and weeks that followed, a full blown civil-military relations ensued [@feaver_military_2020].
+Several cabinet members and uniformed military leaders sought to distance themselves from the President's threat of invoking the Insurrection Act of 1807 to use active-duty forces to quell domestic riots that had initiated following the death of George Floyd in Minneapolis.
+The Secretary of Defense, Mark Esper, stated that he "did not support" use of the 1807 Insurrection Act to deploy active duty military to establish order, and General Mark Milley, the Chairman of the Joint Chiefs of Staff, told a graduating class from the National Defense University that he "should not have been [joining the President on the walk to St.
+John's Church]," and that his presence "created a perception of the military involved in domestic politics" [@esper_secretary_2020; @milley_official_2020].
+Multiple retired American generals also spoke out against the Trump administration's response  But for those paying close attention to what these retired military officers said, it is clear that there was some noticeable variation in their remarks.
+While all of them expressed deep concern over the way in which the President was threatening to use the military [@brooks_dismay_2020; @dempsey_former_2020 ], some attacked the President in much more forceful and personal terms [@goldberg_james_2020; @mullen_i_2020].
+
+This paper examines the way in which retired military officers have attempted to influence public opinion in contemporary America through the use of opinion-editorial pieces.
+It is an interesting question worthy of examination for at least three reasons.
+First, even after leaving active duty service, military officers remain part of the military institution - one that maintains considerable trust and credibility among the American public.
+A 2018 Pew Research Center poll determined that 80% of the American public expects the military to act in the best interests of the American public, compared to business leaders (45%), the media (40%), and elected officials (25%) [@johnson_trust_2018].
+The second reason this question is important is because as an institution, the military prides itself on being nonpartisan and apolitical, an ideal that is undoubtedly under significant stress in an era of increased political polarization [@layman_party_2006; @hare_polarization_2014].
+Third, this question is worth exploring because it may reveal the degree to which the first two factors - military credibility and political polarization - have influenced the concept and norm of professionalism among the military as an institution.
+All of these factors are worthy of exploration, especially ahead of what promises to be a very spirited presidential election this coming fall.
+
+This paper proceeds in three parts.
+First, I briefly review the basis upon which the norm of apolitical and nonpartisan behavior of the military as an institution is based.
+Here I briefly establish a basis for political polarization in contemporary America as well as the high degree of credibility the military as an institution presently enjoys.
+Second, I describe the analytical approach and the data used in this project.
+Assembling an original data set that includes more than 300 opinion pieces published by retired military officers in major American newspapers since the late 1970s, I conduct a mixed-methods approach which ultimately reveals that although the vast majority of opinion pieces published in major newspapers seem to uphold the ideal of the military as apolitical and non-partisan, this trend has been weakening in recent years.
+The data shows that retired military officers are increasingly writing about topics that are not immediately related to national security, as well as increasingly criticizing if not altogether attacking political administrations more often than they did in previous eras.
+Third, I briefly explore the implications of these results on the military as an institution, and specifically, its ability to remain professional in a highly polarized political climate.
+\singlespace
+# The Non-Partisan Military Institution, Credibility, and Polarization  
+\doublespace
+The ideal of the US military acting as an objective, apolitical and non-partisan institution is deep.
+One of the strongest origins of the norm stems from the work of the political scientist, Samuel Huntington.
+In his 1957 work, _The Soldier and the State: The Theory and Politics of Civil-Military Relations_, Huntington asserted that two distinct forces would always be in tension with one another in a liberal democracy such as the United States.
+The first was a "functional imperative," which Huntington said drove the requirement that a military be strong enough and capable of defeating the military threats it faced [@huntington_soldier_1957].
+The second force was a "societal imperative," which Huntington said generated requirements for the military in that the military could never lose so many aspects of it such that it no longer represented the liberal democracy that it defended [@huntington_soldier_1957].
+Huntington's point was that America needed a strong military that could win on the battlefield, but cautioned that such a military could not be so foreign in character so as to alienate the tenets of an American liberal society, and that these forces were at odds with each other.
+The key to resolving the tension generated by these two opposing forces was to adopt a scheme of "objective civilian control" in which political leaders established the ends and goals of policy, while military leaders developed and implemented the military plans to accomplish these goals [@huntington_soldier_1957].
+To Huntington, largely separate operating spheres between the political leaders of a democracy and the military as an institution would ensure that the military remained "professional," and thus, untainted by partisan interests and concerns.
+In this way, Huntington posited that the American military would best meet the task of defeating security threats while also remaining apolitical.
+
+While Huntington's theory has enjoyed tremendous staying power (Huntington is still studied and taught in many American military institutions), he is not without critics [see, for example, @feaver_civil-military_1996; @cohen_supreme_2003; @feaver_armed_2005; and @brooks_paradoxes_2020].
+It would be too burdensome to unpack all of those here, but one of the earliest and richest critiques of Huntington, and perhaps one of the most relevant given today's polarizing environment, came from S.E.
+Finer.
+Authored in 1962, Finer's book, _The Man on Horseback: The Role of the Military in Politics_, argued that Huntington's concept of professionalism simply was not strong enough to keep the military out of politics [@finer_man_1962].
+Finer's concern was that of military officers adopting an ethic of professionalism such that they would actually be more likely - not less likely - to become involved in political issues and concerns [@finer_man_1962, 22-23].
+Finer was especially concerned that the adoption of professionalism alone would lead to military professionals believing that they knew what was best for their country, over and above that of the specific, elected civilian leaders in office.
+Finer warned that "the moment the military draw this distinction between nation and the government in power, they begin to invent their own private notion of the national interest, and from this it is only a skip to the constrained substitution of this view for that of the civilian government…this purported care for the national interest as defined by the military is indeed one of their main reasons for intervening [@finer_man_1962].
+Finer's suggestion was that because of the perils of professionalism as espoused by Huntington, what was really needed to keep the military out of politics was a deep internalization of "the principle of civil supremacy" [@finer_man_1962]^[
+Importantly, neither Huntington's concept of professionalism nor Finer's critique of the concept are arguing for blind obedience by the military to civilian orders and directions, a point that is, in the author's opinion, often missed in debates about civil-military relations.
+On the contrary, both Huntington and Finer make room and acknowledge that there are instances when military disobedience is not only required, but justified.
+However, both agree that in general terms, these situations should be exceedingly rare.]
+
+It is not only in the academic literature that the tradition and worth of an apolitical military has been cultivated.
+The operational military has also been central in upholding the military's non-partisan traditions.
+Retired Army General Douglas MacArthur, a hero of World War Two in the Pacific who is also well known for the controversy surrounding his conduct in the Korean War and his eventual dismissal by President Truman, told the cadets at West Point in May of 1962.
+
+\singlespacing
+
+> **Yours is the profession of arms, the will to win, the sure knowledge that in war there is no substitute for victory, that if you lose, the Nation will be destroyed, that the very obsession of your public service must be Duty, Honor, Country.
+Others will debate the controversial issues, national and international, which divide men’s minds.
+But serene, calm, aloof, you stand as the Nation’s war guardians, as its lifeguards from the raging tides of international conflict, as its gladiators in the arena of battle.
+For a century and a half you have defended, guarded and protected its hallowed traditions of liberty and freedom, of right and justice.
+Let civilian voices argue the merits or demerits of our processes of government.
+Whether our strength is being sapped by deficit financing indulged in too long, by federal paternalism grown too mighty, by power groups grown too arrogant, by politics grown too corrupt, by crime grown too rampant, by morals grown too low, by taxes grown too high, by extremists grown too violent; whether our personal liberties are as firm and complete as they should be.
+These great national problems are not for your professional participation or military solution.
+Your guidepost stands out like a tenfold beacon in the night: Duty, Honor, Country** [@macarthur_duty_1962].
+
+\doublespacing
+
+Current military doctrine and practices likewise reinforce the notion of the military remaining apolitical.
+Prior to the 2016 Presidential Election, Marine Corps General Joseph Dunford, who then served as the Chairman of the Joint Chiefs of Staff, said that the basis upon which the military is able to maintain the public's trust is it's ability to remain "apolitical and neutral" [@dunford_upholding_2016].
+General Martin Dempsey, in an opinion piece published in the Washington Post on July 31, 2016, specifically criticized retired Army Major General Mike Flynn and Marine Corps General John Allen for participating in the Republican and Democratic conventions, respectively, stating that, "As generals, they have an obligation to uphold our apolitical traditions.
+They have just made the task of their successors – who continue to serve in uniform and are accountable for our security - more complicated.
+It was a mistake for them to participate as they did.
+It was a mistake for our presidential candidates to ask them to do so" [@dempsey_military_2016].
+
+Although the notion of an apolitical military is the theoretical goal of the institution, the practical record remains far from perfect.
+From MacArthur's clash with President Truman in the early 1950's, to the so-called "Revolt of the Generals" in 2005-2006, military leaders, even those who are retired, have ventured into political territory from time to time.
+What we have established, though, is clear: a non-partisan military that performs its military mission well has long been the goal of both civil-military relations theorists and military officers.
+More importantly, this is what fosters and ultimately earns the trust and confidence of the American people.
+
+## Political Polarization, Public Opinion, and Military Expertise
+
+Scholars of American Politics have studied political polarization in depth, and while they have offered differing explanations for the causes of polarization, ranging from race and religion [@abramowitz_great_2018], to debates about the size and role of government [@sides_two_2015], to fierce policy disagreements over policy [@mann_its_2012], virtually all agree that polarization, if not an entirely new phenomenon, has become more pronounced in recent decades [@hare_polarization_2014; @bonica_mapping_2014].
+Hare and Poole are among scholars who argue that modern American political polarization began in the 1970s as a result of redistribution along racial lines following key pieces of legislation in the 1960s [@hare_polarization_2014].
+They further argue that polarization disagreement has only intensified, something that a significant portion of modern Americans agree on [@schaeffer_far_nodate].
+
+Within this polarized political environment, what role does the military play - particularly retired military officers who possess considerable military expertise?
+In his influential study, _The Nature and Origins of Mass Opinion_, John Zaller argues that formations in mass public opinion are the result of three interacting components: "variation in the information carried in elite discourse, individual differences in attention to this information, and individual differences in political values and other predispositions" [@zaller_nature_1992].
+Zaller argued that throughout modern America, most citizens, as non-elites, must rely heavily on the messages that come from elites regarding various topics and issue areas [@zaller_nature_1992].
+Zaller further argues that the messages that elites send is necessarily "stereotyped" in some way: that is, the messages that elites send, in order to be understood by regular citizens who are part of the public, must be "short, simple, and highly thematic" [@zaller_nature_1992].
+In developing this argument, Zaller is not claiming that regular citizens are uninformed or careless people who are incapable of processing complex information.
+He is rather acknowledging that most citizens care most about the world that is not "beyond their personal experience" [@zaller_nature_1992].
+His point is that elites play a significant role in helping to form and sway public opinion.
+
+Michael Robinson further examined how elite cues are received and processed.
+Robinson argues that partisan groups, as a result of ongoing political polarization, are cognitively biased in distinct ways in how they receive and process information [@robinson_who_2019].
+As a result, Robinson found that partisans will express confidence in the military in vastly differing ways even after neutral, non-partisan information about various types of military failure is presented to these groups [@robinson_who_2019].
+If Republicans are less likely to express a reduction in confidence in the military as an institution than Democrats are, as Robinson alleges, then even accurate, unbiased information is potentially irrelevant, as these differing partisan groups are pre-programmed to selectively see a particular situation as they wish to see it [@robinson_who_2019].
+His main point serves to warn us regarding confidence in the military as an institution: there may be real reasons to lose confidence in the military as a result of a particular incident, but partisan filters will prevent a reduction of confidence from actually being expressed.
+
+Zaller's framework makes it clear that it is a perfectly reasonable assumption that retired military officers who choose to express their opinions through the form of opinion-editorial pieces constitutes a form of elite discourse.
+After all, retired military officers have at least twenty years of experience, and in the case of most multi-star generals or admirals, usually more than 30 and in some cases close to 40 years of military experience.
+Indeed, most retired senior generals and admirals have spent their professional careers solely in the military, and have extensive experience while serving in leadership positions of large organizations and, in many cases, during times of war.
+Retired military officers are thus elites in the field of national security, and in line with Zaller's framework, they possess the potential to significantly contribute to policy debates and thus, to influence public opinion.
+Yet we also know that the American public will not necessarily evaluate the opinions of these military experts objectively, but rather through pre-conceived and potentially partisan lenses [@robinson_who_2019].
+
+# Methodological Approach, Data, and Hypotheses
+
+Given these three parameters -- that retired military officers are part of an institution that highly values professionalism and non-partisanship, that political polarization in America has increased since at least the 1970s, and that retired military officers are elites who possess the potential to sway public opinion -- I now aim to see whether the discourse of retired military officers has changed over time, specifically since the 1970s, as polarization has worsened.
+To answer this question, I assemble and examine a data set that consists of opinion pieces authored by retired military officers in major US newspapers.
+My methodological approach involves two key steps.
+First, I determine if there have been shifts in the topics that retired military officers have been writing about.
+Second, I apply a normative lens to determine if there have been shifts in the tone and tenor of the pieces authored by these officers, particularly regarding the professional aspects of the military previously discussed.
+In simple terms, I want to answer the questions, what are these retired military officers writing about, and through the lens of professionalism, should they be saying what they are saying?
+
+Using the Factiva and Nexus-Uni databases, I conduct a search that focuses keywords in the byline (author field) such as "retired military officer," "retired from the Army, Navy, Marine Corps, Air Force," "former Chairman of the Joint Chiefs of Staff," "General," "Admiral," and so forth.
+For sources, I confine my search to the largest US newspapers that regularly run opinion-editorial pieces, which includes  _The Wall Street Journal_ (WSJ), _The New York Times_ (NYT), _The Washington Post_ (WaPo), _The Los Angeles Times_, and _The USA Today_.
+I begin the search for the data of January 1, 1979 - 30 June 2020^[I also open my search to include both print and online offerings to ensure that I am capturing any opinion pieces published digitally.
+Most of the newspapers initiated digital offerings in the late 1990s.].
+Using newspapers, and particularly these specific ones, provides some advantages as well as disadvantages.
+The primary advantage is that these sources have been around since the period I wish to examine, unlike many of the online only sites that have only been present in the digital age.
+This allows me to form a data set which covers more time.
+The primary disadvantage is that I am not capturing all of the opinion-pieces authored or given by retired military officers.
+For example, retired military officers are common on cable TV, often weighing in on topics, and I am not capturing those here.
+It is also true that over time,  social media has taken over as the primary means in which a majority of Americans receive their news [@martin_how_2018].
+Print newspapers still exist, but overall circulation is still down today compared to earlier years, even after factoring in the expanse of digital options available^[As a means of rough comparison, estimated total daily newspaper circulation in the US in 2000 was 62 million and 54 million for weekday and weekend editions, respectively, while in 2018, the the same figures were 29 million and 31 million [@barthel_leading_2019]].
+Nonetheless, I contend that the data I am able to assemble will still be informative.
+Publishing opinion-editorial pieces, especially in major, competitive newspapers, is a highly competitive process.
+For the newspapers that I examine in this study, it is not uncommon to see opinion pieces published by figures that include heads of state, members of congress, governmental figures, and leaders of industry and academia.
+Additionally, opinion pieces are deliberately authored for the purpose of informing and persuading the public on a salient or timely topic.
+They are different from interviews or commentary that may be asked by a questioner because the author must take time to put forth his or her position in writing.
+The data therefore allows the researcher to truly ascertain what the specific views and arguments of an author were on a given topic or policy.
+
+Of note, when assembling the data, some authors are explicitly identified in some of their publications as retired military officers, but not in others.^[This was true of several individuals, including Alexander Haig, a retired Army General who later served as Secretary of State under Ronald Reagan;  Robert McFarlane, a retired Marine Corps Lieutenant Colonel who later served as the National Security Adviser to President Reagan, Andrew Bacevich, a retired Army Colonel who then became a prominent historian and author; and Michael Hayden, a retired Air Force General who served as director of the National Security Agency and the Central Intelligence Agency].
+I conduct a separate search for all opinion pieces penned by these authors, regardless of whether they were identified as a retired military officer or not, concluding that it is best do so for coding purposes.
+Ultimately, the data set contains 334 opinion-editorial publications, penned by 178 different authors or author teams.^[There are several co-authored pieces.
+For example, where General (Retired) David Petraeus wrote one publication and another one was written by General (Retired) David Petraeus and Michael O'Hanlon, I counted these as two different authors for coding purposes.] 10 of these pieces are written unanimously, with nods by the authors to their previous service.
+
+From a macro perspective, consider for a moment all that the date range includes with respect to the US military.
+Indeed, numerous and substantive events, debates, and changes have occurred in this time period for retired military officers to weigh in on.
+In terms of external threats and wars, this period includes the final years of the Cold War against the Soviet Union, the first war in Iraq, operations in the Balkans, and the post-9/11 wars in Iraq and Afghanistan, as well as a host of smaller but important military operations in Grenada, Panama, Somalia, and Haiti.
+The military has also undergone significant social and institutional changes in the time period examined.
+In the late 1970s, the military had ended the draft only years before following the conclusion of the Vietnam War.
+The debates over "Don't Ask, Don't Tell" and policies regarding women in the combat arms are also included in the data, as are numerous discussions of military budgets, weapons systems, and alliances.
+Attempting to reconcile the fact that the military as an institution prides itself on being a non-partisan and apolitical institution, that newspaper circulations (even after accounting for digital offerings) are substantially down over the past 40 years, and that the US political environment is increasingly polarized along partisan lines, I adopt the following hypotheses:
+
+**$H_1$: The majority of op-eds published by retired military officers over the time period examine will be apolitical and non-partisan in tone.**
+
+**$H_2$: Because overall newspaper consumption has dropped in the past 40 years, there will be fewer opinion pieces over time authored by military officers.**
+
+**$H_3$: Given that polarization has become increasingly pronounced in the United States in recent years, there will be an increase in the number of op-eds which reveal political and partisan sentiments (indication of a drop in professionalism).**
+
+# First Layer: Topical Analysis
+
+In the first level of analysis, after capturing key biographical and publication information, such as author name, article title, source, publication date, and branch of service and rank of the author, I code each observation into one of several binary categories that corresponds to the overall subject matter of the publication.
+The goal is to create a coding scheme by subject that is sufficiently broad but also specific enough that is helpful in differentiating the overall main point of each publication.
+The following coding scheme is therefore devised and implemented (all variables are binary, taking on the value of 1 if the publication corresponds to the general definition of the variable, and 0 otherwise): \singlespace
+
+* **War_Ops**- if the piece advocates for or denounces a particular course of action directly related to operational or war fighting decisions of a clear military nature.
+Examples include: the wisdom of troop increases, the Surge in Iraq, whether to bomb Serbia, or even the role of the prison at Guantanamo.
+
+* **Trp_Clt**- if the piece generally showcases or highlights the work and sacrifice of members of the armed forces; to defend the work of specific members of the Armed Forces; or to call for policing up the institutional cultures of the services.
+Examples include: remembering our fallen soldiers on Memorial Day; improving the institutional culture of the Army or Navy; advocating on behalf of or remembering a specific flag officer.
+
+* **Str_Adv**- if the piece is generally written to offer some sort of strategic advice for the nation, to include the wisdom of pursuing particular alliances or agreements, or adopting a specific weapons platform or military budget.
+Examples include: why the US needs to strengthen NATO, build ties with China, increase defense spending, or not retire the A-10 combat plane.
+
+* **MilPSoc**- if the piece advocates for or denounces a particular social policy that _directly influences_ the armed forces.
+Examples include: debates over keeping or removing the "Don't Ask, Don't Tell" policy, opening up combat roles to women, the role of Transgender members of the armed forces.
+
+* **Dom_P**- if the piece advocates for or denounces a position on a policy that _does not directly_ influence the armed forces in its day to day activities.
+Examples include: support for the second amendment, immigration policy, and whether confederate statues should be taken down or left up.
+
+* **Fig_Name**- if the piece advocates for or denounces a particular position held by a political administration, or _if the main point of the piece_ is to explicitly defend or attack a political figure by name.
+Examples: endorsing a Presidential candidate for office, or for attacking a President or member of Congress for a decision that was made.
+
+\doublespace
+Separating opinion-editorial publications into one of several distinct categories is an imperfect process.
+To aid the coding process, I first identify the thesis statement of each opinion piece, which in almost every case, makes possible a reasonably clear classification of the publication.
+Two points are worth mentioning.
+First, I attempt to keep in mind the fact that opinion pieces are intended to argue for something, and thus are, to an extent, designed to be provocative.
+With this in mind, I did not automatically code pieces that generally advocated for or cautioned the President in the **Admin_Fig_Name** category.
+Second, there were some instances in which I had to choose between placing a publication into one of two categories, say, between **Warfighting_Ops** and **Strat_Advice**, but very few if any instances in which I had to choose between categories that differed greatly, such as **TroopSptCulture** and **Admin_Fig_Name**.
+Table 1 reveals the breakdown of the data by article type.
+
+```{r summary, include = FALSE}
+# Read in the Data
+dataset <- read_csv(here("data", "MilEditorials24June.csv"))
+dataset %>% print()
+# Replace NA with 0 in the data for variables War_Ops thru Fig_Nme
+adjdataset <- dataset %>%
+  mutate_at(
+    .vars = vars(War_Ops:Fig_Nme),
+    .funs = ~ case_when(is.na(.) ~0, TRUE ~ 1)
+) %>% mutate(
+    PubYr = lubridate::year(PubDate)
+) %>% view()
+
+summary(adjdataset)
+names(adjdataset)
+
+# count the number of observations by type of binary variable and store as new dataframe
+countbytype  <- adjdataset %>% count(War_Ops, Trp_Clt, Str_Adv, MilPSoc, Dom_P, Fig_Nme) 
+countbytype
+
+#count the number of times each source occurs in the data
+WSJcount <- adjdataset %>% count(Source=="WSJ")
+WSJcount
+
+NYTcount <- adjdataset %>% count(Source=="NYT")
+NYTcount
+
+WaPOcount <- adjdataset %>% count(Source=="WaPO")
+WaPOcount
+
+USATodaycount <- adjdataset %>% count(Source=="USA Today")
+USATodaycount
+
+LATimescount <- adjdataset %>% count(Source=="LA Times")
+LATimescount
 
@@
 
-First, I import the data set and using the here () function to build the file path.  
-I replace all NA/s with ) in the data set.
-
-```{r summary}
-# Read in the Data
-dataset <- read_csv(here("data", "MilEditorials23Apr.csv"))
-# Replace NA with 0 in the data
-dataset[is.na(dataset)] <- 0
-
-# Let's view the data so we know what kind of dataframes each of the variables are.  
-dataset
-
-# Note that I may want to convert AuthSe, AuthRank, and Service to some form of numeric.  
-# But I will come back to that.
-
-#See the variable names.  
-summary(dataset)
-names(dataset) 
+```{r table, fig.align="center", include=TRUE, echo=FALSE}
+countbytype %>%
+  kable(format="latex", caption="Publications by Subject") %>%
+  kable_styling(latex_options = "striped", font_size=10, full_width = T) %>% 
+  row_spec(0, bold=T)
 ```
 
-```{r manipulation}
+Table 1 reveals that an overwhelming majority (289 of 334, or 86.5%) of the total number of opinion pieces written since 1979 in the major newspapers examined can be classified according to one of the following three categories: **War_Ops**, **Trp_Clt**, and **Str_Adv**.
+It is reasonable to conclude that these op-eds written are appropriate from the perspective of civil-military relations in that they address topics which American citizens can reasonably expect retired military officers to bring their expertise to bear upon.
+The remaining 45 opinion pieces (13.5% of the total) fall into one of the other three categories: **MilPSoc**, **Dom_P**, and **Fig_Nme** and require further examination.
+They are not necessarily inappropriate, but they likely address topics that directly address or are closely related to issues that could be considered as partisan or at least beyond appropriate military expertise.
+This second layer of analysis will help expose whether there is any indication that professionalism as a norm among retired military officers appears to be weakening.
 
-# Create a Variable for Publishing Year called PubYr
-adjdataset <- dataset %>% 
-  mutate(
-  PubYr = year(PubDate)
-) %>% print()
+\singlespace
+# The Second Layer - Are Retired Military Officers Upholding Professionalism as a Norm?
+\doublespace
+Of the remaining observations, two are removed upon immediate inspection.
+One was written by an unknown author and is entitled, "Fists Raised at West Point," and was published in _The Washington Post_ on May 13, 2016.
+It is removed because the article is actually a compendium of two separate letters sent to the editor by two different authors, one arguing for and the other against female West Point graduates who raised their fists upon graduating, which ignited some controversy about soon-to-be Army lieutenants purportedly making a symbol associated with the "Black Lives Matter" movement [@noauthor_fists_2016].
+Because the two separate letters taken together do not take a firm position on the subject, but rather show a difference in opinion about a specific event happening in 2016, I remove the observation.
+The second piece I remove was written by former Democratic Congressman John Murtha, who was also a retired Marine Corps officer.
+As a politician, the American public can and should expect him to write in the capacity of a politician, and therefore, to make partisan arguments.
+Not wanting to skew the remaining data, I remove his publication for the remaining analysis.
+Thus, the second level of analysis proceeds on a total of 43 observations, which are listed and printed in chronological publishing order, in Table 2.
+
+\singlespace
+
+```{r remaining data, include = FALSE}
 
 # Show all articles and authors who published articles on Social Policy, Political Parties, and Administrations/Political Figures
-adjdataset %>% arrange(PubYr) %>% 
-  filter(PolicySocial==1 | PartyGen==1 | PartyName ==1 |PolAdminName==1 )  %>% 
-  select (AuthName, ArtTitle, Source, PubYr) %>% 
+rem_obs <-adjdataset %>% arrange(PubYr) %>% 
+  filter(MilPSoc==1 | Dom_P==1 | Fig_Nme==1 )  %>% 
+  filter(Author != "UNK") %>% filter(Author != "Murtha, John") %>% 
+  select (Author, Title, Source, PubYr, MilPSoc, Dom_P, Fig_Nme) %>% 
   view()
+```
 
+```{r table2, fig.align = "center", include=TRUE, echo=FALSE}
+  rem_obs %>%
+  kable(format="latex", caption="Remaining Observations", longtable=T) %>%
+  kable_styling(font_size=10, latex_options="striped", repeat_header_continued=T) %>%
+  column_spec(1, width="12em") %>% 
+  column_spec(2, width="16em") %>% 
+  column_spec(3:4, width="2.5em") %>% 
+  column_spec(5:7, width="3em") %>% 
+  row_spec(0, bold=T)
+```
+
+```{r code, include=FALSE, echo=FALSE}
 # Create a Variable Called case-count that counts the number of times by publishing year an article published 
 # regarding social policy or general parties or parties by name or administrations occurs
-adjdataset %>% group_by(PubYr) %>% 
+rem_obs %>% group_by(PubYr) %>% 
   mutate(
-    case_count=sum(PolicySocial ==1 |PartyGen==1 | PartyName ==1 | PolAdminName==1) 
-  ) %>% 
-  print
+    case_count=sum(MilPSoc ==1 | Dom_P==1 | Fig_Nme==1) 
+  ) %>% view()
+```
+\doublespace
 
-# Count the Number of Articles Published By Type
-adjdataset %>% tally(PolicySecurity) 
-adjdataset %>% tally(PersTroopSpt)
-adjdataset %>% tally(GenAdvice)
-adjdataset %>% tally(PolicySocial)
-adjdataset %>% tally(PartyGen)
-adjdataset %>% tally(PartyName)
-adjdataset %>% tally(PolAdminName)
+A couple of interesting observations are worth noting about the data in Table 2.
+First, one can quickly observe that there appear to be three distinct clusters of observations with respect to publishing year.
+The first occurs in the time frame between roughly 1998-1999, the second between 2004-2007, and the third between 2016-2019.
+The clusters can also be seen in time in Figure 1 on the following page.
 
-# need to change the type of class these vectors are - to binary
+In the first cluster, the publications refer to multiple disagreements certain retired military officers had about how President Clinton was planning or using military force.
+One observation, that written by retired Marine Corps General Carl Mundy, discussed implications of changing the policy of "Don't Ask, Don't Tell," which was in place at the time, while another observation, that written by retired Army General Barry McCaffrey, refers to domestic drug policy in the United States - a perfectly acceptable topic given that he, at the time, was serving as President Clinton's Director of the Office of National Drug Control Policy.
+The second cluster refers to pieces penned around the so-called "Revolt of the Generals," referring to sharp disagreements that entered the public arena between multiple US generals and admirals and President George W.
+Bush's Secretary of Defense, Donald Rumsfeld.
+Disagreement about Rumsfeld's leadership, particularly regarding the planning and execution of the early years of the war in Iraq, were central during this period of time.
+Retired Army General Tommy Franks penned two observations that essentially provided public support for the President at the time, George W.
+Bush, and Retired Army Major General Paul Eaton, who was a particularly vocal critic of Rumsfeld, authored two of the pieces published in this time period.
+Finally, the 2016-2019 cluster is indicative mainly of concerns over the candidate and later President Donald Trump.
+Retired Air Force General Michael Hayden, Navy Admirals Mike Mullen and William McRaven, and Army General Stanley McChrystal are among the retired officers penning pieces during this time period.
 
+```{r fig1, fig.height=5, fig.width=7, fig.align="center", include=TRUE, echo=FALSE}
+datatograph <- data.frame(
+  year=c("1994","1996","1998","1999","2000","2003", "2004", "2006", "2007", "2010", "2013","2014", "2016", "2017", "2018", "2019", "2020"), count=c(1, 1,2, 3, 1, 1, 3, 3, 3, 1, 1, 1, 6, 5, 4, 6, 1)
+)
+ggplot(datatograph, aes(x =year, y =count)) + geom_bar(stat = "identity", width=.2) +
+  labs(
+    x = "Year",
+    y = "Count",
+    color = NULL,
+    title = "Figure 1: # of Op-Eds Written By Military Retirees",
+    subtitle = "On Topics of Social Policy, Domestic Policy, or Individual Figures"
+  )
 ```
 
-```{r simple graphs}
+Evaluating a conclusive level of appropriateness of the opinion pieces listed in Table 2 cannot be definitively solved, precisely because it matters the perspective from which one views how military officers should interact with the American public, especially when attempting to shape public opinion.
+However, it is possible to determine whether there has been deviation from non-partisan and apolitical norms.
+Recall that Retired Army General Douglas MacArthur cautioned the cadets at West Point in 1962 about involving themselves in matters that do not require a military solution (in MacArthur's words, of "politics grown to corrupt," "morals grown too low," and "whether our personal liberties are as firm and complete as they should be") [@macarthur_duty_1962], and that more recently, in 2016, Retired Army General Martin Dempsey lamented the participation of retired generals in the political process by endorsing political candidates and appearing at the conventions of political parties:
 
+\singlespace
+
+> **The American people should not wonder where their military leaders draw the line between military advice and political preference.
+And our nation's soldiers, sailors, airmen and Marines should not wonder about the political leanings and motivations of their leaders.
+As generals, they have an obligation to uphold our apolitical traditions.
+They have just made the task of their successors – who continue to serve in uniform and are accountable for our security - more complicated** [@dempsey_military_2016].
+
+\doublespace
+
+Using these ideals as guidelines allows us to conduct a second level of analysis, with emphasis on the two characteristics of topic (is the author writing about a reasonable military topic) and political leaning (is the author expressly attacking or supporting specific political figures in a political way, from which one can reasonable infer the officer's political or partisan bent)?
+
+Using these characteristics as guidelines, I therefore develop two new variables. The first is **Result**, which can take on the following values:\singlespace
+* **Permissible** -if the reader cannot clearly draw conclusions regarding the author's political stances or partisan preferences; or if the topic addressed in the publication in fact does in fact fall underneath the area of clear military expertise.
+* **Questionable** -if the reader can draw some general, but not final, conclusions regarding the author's political stances or partisan preferences; or if the topic addressed in the publication is questionably an area of reasonable military expertise.
+* **Inappropriate** -if the reader can draw clear conclusions regarding the author's partisan preferences or political leanings; or if the author addresses a topic that is not an area considered to be directly relevant to military expertise, or a matter that the military should judge.
+
+\doublespace
+The second new variable I develop is **Justification** in which I briefly comment on why I assigned the value of **Result** for the observation.
+Here I simply comment on whether it is a concern of partisan preferences being revealed and/or topic addressed that are responsible for the value of **Result**.
+The results of this second level of analysis are presented in Table 3.
+Note that several publications are deemed to take on values of permissible under the variable **Result**.
+Importantly, three of these pieces occurred during the first previously identified cluster, 1998-1999.
+No pieces from the second cluster, 2004-2007 have been deemed permissible, whereas in the third cluster, 2016-2019, three take on the value of permissible.
+Thus, we now have a picture of how prevalent the publication is of opinion pieces authored by retired military officers in major newspapers which have failed to uphold the norm of non-partisanship and apolitical position.
+
+\singlespace
+
+```{r second level analysis, include = FALSE}
+
+# Present Results of Result and Justification
+secondlayer_obs <-adjdataset %>% arrange(PubYr) %>% 
+  filter(MilPSoc==1 | Dom_P==1 | Fig_Nme==1 )  %>% 
+  filter(Author != "UNK") %>% filter(Author != "Murtha, John") %>% 
+  select (Author, Title, Source, PubYr, Result, Justification) %>% 
+  view()
 ```
+
+```{r table3, include=TRUE, echo=FALSE}
+secondlayer_obs %>%
+  kable(format="latex", caption="Concern for Op-Ed Pieces Authored By Retired Military Officers in US Newspapers, 1979-2020", longtable=T) %>%
+  column_spec(1, width="9.5em") %>% 
+  column_spec(2, width="15em") %>% 
+  column_spec(3:4, width="2.5em") %>%
+  column_spec(5, width="5em") %>%
+  column_spec(6, width="6em") %>%
+  kable_styling(latex_options="striped", font_size=10) %>% 
+  row_spec(0, bold=T)
+```
+\doublespace
+
+Figure Two displays the number of opinion pieces authored by retired military officers which failed to uphold the strict standards of professionalism, as espoused by Retired Generals MacArthur and Dempsey, within each applicable year.
+The results reveal several conclusions.
+The first is that it clearly shows, in a broad sense, that from 1979 through the first half of 2020, opinion pieces penned at the hand of retired military officers which violate the concepts of professionalism in questionable or inappropriate ways are the exception, not the norm.
+Indeed, the entire count of questionable or inappropriate publications after conducting the second level of analysis drop to just 30 of 334 total pieces, or 8.9% of the entire sample.
+Thus, ample support exists to confirm the first hypothesis, that the vast majority of opinion pieces authored by retired military officers will be apolitical and non-partisan in nature.
+
+```{r fig2, fig.height=4, fig.width=5, fig.align="center", include=TRUE, echo=FALSE}
+datatograph <- data.frame(
+  year=c("1998","2000","2003", "2004", "2006", "2007", "2016", "2017", "2018", "2019", "2020"), count=c(2, 1, 1, 3, 3, 1, 5, 4, 4, 5, 1)
+)
+ggplot(datatograph, aes(x =year, y =count)) + geom_bar(stat = "identity", width=.2) +
+  labs(
+    x = "Year",
+    y = "Count",
+    color = NULL,
+    title = "Figure 2: # of Op-Eds Written By Military Retirees",
+    subtitle = "Failed to Uphold Professional Standards"
+  )
+```
+
+The second hypothesis, which inferred that because overall circulation among major US newspapers has been declining over the past several decades, the number of opinion pieces penned by retired military officers in these outlets would likewise decline, does not appear to be represented from the data.
+In fact, we find evidence for the opposite trend: retired military officers are penning opinion pieces at an increasing rate, even as overall circulation among the major newspapers are down in terms of overall circulation (even after accounting for digital offerings) over the same time period.
+
+The third hypothesis, that the number of published opinion pieces which have strained the concept of professionalism has increased, even as political polarization has worsened in the past several decades, is also confirmed, as evidenced by both by Table 3 and Figure 2.
+It is interesting to note that according to Figure 2, the first publishing year in which an opinion piece penned by a retired officer failed to uphold professional standards occurred in 1998, nearly twenty years into (or almost half) of the entire sample.
+Why is this the case?
+Retired military officers in the years preceding 1998 had plenty of reasons to go to the press and air disagreements with political leaders, or to address important topics that may have pushed the limits of professionalism as has been defined in this paper.
+In fact, many of these officers would have been veterans of the Vietnam and Korean Wars, and perhaps, like Maxwell Taylor, who is in our sample, veterans perhaps of World War Two.
+At this point, though it is only speculation, one can speculate that the factors of military credibility and political polarization, however imprecise and hazy at this point, may have played a role in keeping the majority of an earlier generation's retired officers out of the popular press.
+
+## Increasing Diversity of Topics Addressed By Military Officers
+
+It is also clear from the data presented in  Table 3 that several publications are deemed questionable or inappropriate from the perspective of professionalism in that they address topics that are potentially outside of the realm of proper military expertise to address.
+In his _Washington Post_ editorial, "Home Should Not Be a War Zone," Retired Army General Stanley McChrystal addresses the tragic nature of mass shootings in America.
+He states in the article, "...as a combat veteran and proud American, I believe we need a national response to the gun violence that threatens so many of our communities" [@mcchrystal_home_2016].
+In a later article published in the _Washington Post_ in 2018 entitled "Good Riddance," McChrystal shared how his beliefs in Robert E.
+Lee had evolved over the course of his professional career, and why he now believes it is now time to discard Confederate statues and other similar symbols.
+Admiral (Retired) Mike Mullen and co-authors Lieutenant General (Retired) Mark Hertling and Admiral (Retired) Robert Natter wrote publications ("The Refugees We Need" and "Cutting Refugee Admissions Will Have Severe Consequences for the US Military") explaining how President Trump's immigration policies on refugees harm military readiness.
+
+While many Americans agree with the stance these officers have taken on what are clearly important issues, one can simultaneously and not unreasonably ask whether it is appropriate for retired military officers to be considered appropriate elite voices on these specific issue areas.
+Again, the potential issue as it is related to the concept of military professionalism is not that these officers should not have an opinion on these topics, but whether they, as retired military officers, ought to be considered as appropriate mouthpieces by the American public on these issues.
+It is not unreasonable to insist that policy arenas involving appropriate responses to mass violence, the future of Confederate symbols, and refugee/immigration policy for the nation belong to non-military elites.
+
+```{r fig3, fig.height=4, fig.width=5, fig.align="center", include=TRUE, echo=FALSE}
+datatograph <- data.frame(
+  decade=c("1979-1989", "1989-1999", "1999-2009", "2009-2019" ), numcount=c(13, 51, 102, 161)
+)
+ggplot(datatograph, aes(x =decade, y =numcount)) + geom_bar(stat = "identity", width=.2) +
+  labs(
+    x = "Period",
+    y = "Published Opinion Pieces",
+    color = NULL,
+    title = "Figure 3: Total Number of Op-Eds Written By Military Retirees",
+    subtitle = "WSJ, NY Times, Washington Post, USA Today, and LA Times"
+  )
+```
+
+## Conclusion: Can the Military Remain Professional?
+
+The dynamics of the military institution engaging the American public in a polarized environment remains ripe for future scholarship.
+In the author's opinion, whether the American military can, and to what degree it will, be able to remain professional as an institution remains to be seen.
+As political polarization becomes increasingly salient, several dynamics are unfolding that are worth noting, as their full impact has yet to materialize.
+
+The first is that as political polarization worsens, the entire notion of loyalty is being reshaped, redefined, or at the very least, called into question.
+Consider that in June 2020, following the President's largely unpopular move of displacing protesters from Lafayette Square, a letter published by West Point alumni expressly called out prominent West Point alumni serving in the Trump administration.
+Though it did not mention individuals by name, the letter was intended and acknowledged by media as directed towards the Secretary of Defense, Mark Esper, and the Secretary of State, Mike Pompeo, who are both graduates of West Point's Class of 1986 [@gaouette_group_nodate].
+The charges of the letter were both damning and politically striking: "...We are concerned that fellow graduates serving in senior-level, public positions are failing to uphold their oath of office and their commitment to Duty, Honor, Country"[@line_letter_2020].
+The letter's clear implication is that if you are serving in the Trump administration, and if you graduated from the US Military Academy, you have forsaken your oath of office, and by extension, the US Constitution, as well as the values that West Point stands for.
+Increasing political polarization promises to make these charges of misplaced loyalty more common.
+Scholarly literature has looked at how the military faces pressure stemming from divided institutional loyalties [@betts_soldiers_1991], but polarization promises to add even more pressure and scrutiny on the loyalty of individual officers and the military as a whole.
+
+A second dynamic, that of maintaining credibility and trust with the American public, remains crucial for the military, but this dynamic will likely be tested in fierce ways should a polarized political climate endure.
+Even as high credibility with the American public can give the military support, it can also be abused, especially if military officers, to include retirees, use the high level of credibility they currently enjoy to take up and advocate for partisan positions, or worse, to jockey with civilian leaders for power [@brooks_shaping_2008-1].
+Unfortunately, as this paper illustrates, there is evidence that mild aspects of this are occurring, but the extent to which it is, and its resulting impacts, remains an area for future research.
+Credibility is not something that should be shunned.
+The military as an institution, to include military retirees, holds relevant and military expertise that should be afforded to the American public at large.
+And yet, just as retired officers were skilled in the management of violence when they wore the uniform every day, retirement in a polarized era means that they must so also be managers of American trust.
+The promising news is, as this paper makes clear, the great proportion of retirees who are speaking out in various settings know and internalize this.
+Nonetheless, this will become more difficult to do in a polarized environment.
+
+A third, and perhaps the most important dynamic of polarization as it concerns the military, is that the line between what is considered partisan or political and what is considered morally right and wrong is becoming seemingly blurred, if not altogether disappearing.
+Polarized sides tend to believe that each is acting "rightly" and that the other side is not just mistaken, but deeply wrong in a moral sense.
+
+These three factors are difficult in a purely domestic political environment, and one does not have to imagine how the potential snares of a polarized political environment and military credibility stewarded poorly could impede the military's ability to do it's main task, fighting and winning this nation's wars.
+Indeed, the impact of these characteristics on our nation's military effectiveness must also be considered in further research.
+Regardless, the importance of individual moral courage will continue to grow in value as an attribute of military leaders.
+Its importance as a trait will likely continue to grow in the years ahead.
+
+\singlespace
+\pagebreak

### 2020-07-05 code/Submissionv1.Rmd
commit changes

@@
 ---
-title: "When the Pen Becomes a Sword: Retired US Military Officers and Op-Ed Use in the Contemporary Press"
+title: "When the Pen Becomes a Sword: Retired US Military Officers and Op-Ed Use in Contemporary America"
 author: "Peter M. Erickson"
@@
 abstract: |
-  _This paper seeks to explain how retired military officers have attempted to shape US public opinion through the publication of op-ed pieces in major newspapers.
-I construct and analyze a data set that captures more than 300 opinion-editorial pieces published by retired military officers in The Wall Street Journal, The Washington Post, The New York Times, The Los Angeles Times, and the USA Today since 1975, and ultimately find two trends that stand out.
-The first is that retired military officers are increasingly writing about topics that are outside of traditional security related topics.
-The second trend is that retired military officers are increasingly criticizing Presidential administrations, especially that of President Donald Trump, for reasons other than disputes regarding international uses of force.
-Given the polarizing political climate in the United States and that a Presidential election will occur this fall, these two trends are concerning from the perspective of civil-military relations and suggest that the norm of a professional, apolitical military establishment is under significant strain._
+  _This paper explores variation in the professionalism of opinion pieces authored by retired US military officers and published in major US newspapers in the past 40 years.
+Assembling and analyzing more than 330 opinion-editorial pieces published by retired military officers in The Wall Street Journal, The Washington Post, The New York Times, The Los Angeles Times, and the USA Today since 1979 reveals that there has been an increasing trend in published opinion pieces which fail to uphold professional standards as defined by military norms.
+The paper further argues that this trend is primarily the result of two significant strains on the military's politically neutral culture: political polarization and military credibility.
+Finally, the paper develops a theory that shows how the forces of political polarization and military credibility interact, and argues that the norm of a professional, apolitical military will remain under further strain in the current US political environment._
 
@@
 
-On June 1 2020, the President of the United States, Donald Trump, after giving remarks at the Rose Garden of the White House in Washington D.C., walked to St.
+In late May and early June 2020, the American public was gripped by escalating racial tension and discussions about police brutality following the death of George Floyd, an unarmed black man, by police in Minneapolis.
+Protests, most of them peaceful, were accompanied in some cases by rioting and looting.
+On June 1, after giving remarks at the Rose Garden, President Donald Trump walked to St.
 John's Church after security forces visibly displaced throngs of protesters in neighboring Lafayette Square.
-Flanked by members of his cabinet, to include the Secretary of Defense, and with the Chairman of the Joint Chiefs of Staff not far behind, the President held up a Bible at the church, stared into the cameras of the media, and after several minutes, returned to the White House.
+Flanked by members of his cabinet, to include the Secretary of Defense, and with the Chairman of the Joint Chiefs of Staff not far behind, the President held up a Bible at the church, stared into the cameras of the media, and, several minutes later, returned to the White House.
 
-In the days and weeks that followed, a full blown civil-military relations ensued [@feaver_military_2020].
-Several cabinet members and uniformed military leaders sought to distance themselves from the President's threat of invoking the Insurrection Act of 1807 to use active-duty forces to quell domestic riots that had initiated following the death of George Floyd in Minneapolis.
-The Secretary of Defense, Mark Esper, stated that he "did not support" use of the 1807 Insurrection Act to deploy active duty military to establish order, and General Mark Milley, the Chairman of the Joint Chiefs of Staff, told a graduating class from the National Defense University that he "should not have been [joining the President on the walk to St.
-John's Church]," and that his presence "created a perception of the military involved in domestic politics" [@esper_secretary_2020; @milley_official_2020].
-Multiple retired American generals also spoke out against the Trump administration's response  But for those paying close attention to what these retired military officers said, it is clear that there was some noticeable variation in their remarks.
-While all of them expressed deep concern over the way in which the President was threatening to use the military [@brooks_dismay_2020; @dempsey_former_2020 ], some attacked the President in much more forceful and personal terms [@goldberg_james_2020; @mullen_i_2020].
+In the days and weeks that followed, a civil-military relations crisis ensued  [@feaver_military_2020].
+The Secretary of Defense, Mark Esper, publicly disagreed with the President's threat to invoke the Insurrection Act of 1807 to use active-duty forces to quell domestic riots [@esper_secretary_2020], and General Mark Milley, the Chairman of the Joint Chiefs of Staff, told senior military officers graduating from the National Defense University that he "should not have been there" (accompanying the President on the walk to St.
+John's Church), as his presence "created a perception of the military involved in domestic politics" [@milley_official_2020].
 
-This paper examines the way in which retired military officers have attempted to influence public opinion in contemporary America through the use of opinion-editorial pieces.
-It is an interesting question worthy of examination for at least three reasons.
-First, even after leaving active duty service, military officers remain part of the military institution - one that maintains considerable trust and credibility among the American public.
-A 2018 Pew Research Center poll determined that 80% of the American public expects the military to act in the best interests of the American public, compared to business leaders (45%), the media (40%), and elected officials (25%) [@johnson_trust_2018].
-The second reason this question is important is because as an institution, the military prides itself on being nonpartisan and apolitical, an ideal that is undoubtedly under significant stress in an era of increased political polarization [@layman_party_2006; @hare_polarization_2014].
-Third, this question is worth exploring because it may reveal the degree to which the first two factors - military credibility and political polarization - have influenced the concept and norm of professionalism among the military as an institution.
-All of these factors are worthy of exploration, especially ahead of what promises to be a very spirited presidential election this coming fall.
+Retired military officers were among the elites who weighed in to give their opinions not only on the Trump administration's handling of specific events near the White House, but also on the broader issues of race in America.
+But for those paying attention to what these retired officers said, it is also clear that there was substantial variation in the tone and tenor of each officer's remarks.
+While all of them expressed deep concern about the possibility of active-duty forces confronting peaceful protesters in clear violation of the First Amendment of the US Constitution, [@brooks_dismay_2020; @dempsey_former_2020], some officers went much further, leveling direct criticisms of the President's fitness to be President [@goldberg_james_2020], his administration's broader policies on race [@allen_moment_2020], and his general leadership abilities [@mullen_i_2020].
 
-This paper proceeds in three parts.
-First, I briefly review the basis upon which the norm of apolitical and nonpartisan behavior of the military as an institution is based.
-Here I briefly establish a basis for political polarization in contemporary America as well as the high degree of credibility the military as an institution presently enjoys.
-Second, I describe the analytical approach and the data used in this project.
-Assembling an original data set that includes more than 300 opinion pieces published by retired military officers in major American newspapers since the late 1970s, I conduct a mixed-methods approach which ultimately reveals that although the vast majority of opinion pieces published in major newspapers seem to uphold the ideal of the military as apolitical and non-partisan, this trend has been weakening in recent years.
-The data shows that retired military officers are increasingly writing about topics that are not immediately related to national security, as well as increasingly criticizing if not altogether attacking political administrations more often than they did in previous eras.
-Third, I briefly explore the implications of these results on the military as an institution, and specifically, its ability to remain professional in a highly polarized political climate.
+This paper examines the way in which retired military officers have attempted to influence public opinion in contemporary America through opinion-editorial publications.
+It is an interesting question worthy of examination because, as I argue, at least three important factors are interacting in unique ways to shape how and why retired military officers speak out on a variety of issues.
+These factors are professionalism, polarization, and credibility.
+In particular, polarization and credibility are creating tremendous strains on the military's professional nature, and in some cases, are causing words to become weaponized.
+
+This paper proceeds in four parts.
+First, I briefly review the concepts of military professionalism, polarization, and credibility in order to orient the reader to the theory that I build in part two.
+In the second section, I build a theory in which I argue that the contemporary forces of polarization and military credibility can strain the concept of professionalism such that words and actions become weaponized to the point of being antithetical to the very concept of professionalism itself.
+Third, I turn to the data, which consists of more than 330 opinion-editorial pieces published in major US newspapers by retired military officers since the late 1970s.
+While the data shows that the vast majority of these pieces are in fact in line with the notion of military professionalism, recent trends show that the retired military officers are weighing in on topics that fall outside of the realm of national security as well as expressing clear partisan preferences in their arguments.
+Fourth, I briefly explore the implications of these results on the military as an institution, and specifically, its ability to remain professional in a setting where polarization is high and the institution commands wide public support.
+
 \singlespace
-# The Non-Partisan Military Institution, Credibility, and Polarization  
+# Military Professionalism
 \doublespace
 The ideal of the US military acting as an objective, apolitical and non-partisan institution is deep.
-One of the strongest origins of the norm stems from the work of the political scientist, Samuel Huntington.
-In his 1957 work, _The Soldier and the State: The Theory and Politics of Civil-Military Relations_, Huntington asserted that two distinct forces would always be in tension with one another in a liberal democracy such as the United States.
-The first was a "functional imperative," which Huntington said drove the requirement that a military be strong enough and capable of defeating the military threats it faced [@huntington_soldier_1957].
-The second force was a "societal imperative," which Huntington said generated requirements for the military in that the military could never lose so many aspects of it such that it no longer represented the liberal democracy that it defended [@huntington_soldier_1957].
-Huntington's point was that America needed a strong military that could win on the battlefield, but cautioned that such a military could not be so foreign in character so as to alienate the tenets of an American liberal society, and that these forces were at odds with each other.
-The key to resolving the tension generated by these two opposing forces was to adopt a scheme of "objective civilian control" in which political leaders established the ends and goals of policy, while military leaders developed and implemented the military plans to accomplish these goals [@huntington_soldier_1957].
+One of the strongest origins of the norm stems from the work of the political scientist, Samuel Huntington, who, in his 1957 work, _The Soldier and the State: The Theory and Politics of Civil-Military Relations_, asserted that two distinct forces would always be in tension with one another in a liberal democracy such as the United States.
+The first was a "functional imperative," which Huntington said drove the requirement that a military be strong enough and capable of defeating the military threats it faced, and the second was a "societal imperative," which Huntington said generated requirements for the military in that the military could never lose so many aspects of it such that it no longer represented the liberal democracy that it defended [@huntington_soldier_1957, 1-3].
+Huntington's point was that America needed a strong military that could win on the battlefield, but cautioned that such a military could not be so foreign in character so as to alienate the tenets of an American liberal society.
+The key to resolving the tension generated by these two opposing forces was to adopt a scheme of "objective civilian control" in which political leaders established the ends and goals of policy, while military leaders developed and implemented the military plans to accomplish these goals [@huntington_soldier_1957, 80-85].
 To Huntington, largely separate operating spheres between the political leaders of a democracy and the military as an institution would ensure that the military remained "professional," and thus, untainted by partisan interests and concerns.
@@
 While Huntington's theory has enjoyed tremendous staying power (Huntington is still studied and taught in many American military institutions), he is not without critics [see, for example, @feaver_civil-military_1996; @cohen_supreme_2003; @feaver_armed_2005; and @brooks_paradoxes_2020].
-It would be too burdensome to unpack all of those here, but one of the earliest and richest critiques of Huntington, and perhaps one of the most relevant given today's polarizing environment, came from S.E.
+It would be too burdensome to unpack all of those here, but one of the earliest and richest critiques of Huntington, and perhaps one of the most relevant given the potential influences of polarization and credibility, came from S.E.
 Finer.
 Authored in 1962, Finer's book, _The Man on Horseback: The Role of the Military in Politics_, argued that Huntington's concept of professionalism simply was not strong enough to keep the military out of politics [@finer_man_1962].
-Finer's concern was that of military officers adopting an ethic of professionalism such that they would actually be more likely - not less likely - to become involved in political issues and concerns [@finer_man_1962, 22-23].
-Finer was especially concerned that the adoption of professionalism alone would lead to military professionals believing that they knew what was best for their country, over and above that of the specific, elected civilian leaders in office.
-Finer warned that "the moment the military draw this distinction between nation and the government in power, they begin to invent their own private notion of the national interest, and from this it is only a skip to the constrained substitution of this view for that of the civilian government…this purported care for the national interest as defined by the military is indeed one of their main reasons for intervening [@finer_man_1962].
+Finer's concern was that of military officers adopting an ethic of professionalism such that they would actually be more likely - not less likely - to become involved in political issues [@finer_man_1962, 22-23].
+Finer was especially concerned that the adoption of professionalism alone would lead to long-serving military professionals believing that they knew what was best for their country over and above that of temporary elected civilian leaders [@finer_man_1962, 22-23] Finer warned that "the moment the military draw this distinction between nation and the government in power, they begin to invent their own private notion of the national interest, and from this it is only a skip to the constrained substitution of this view for that of the civilian government…this purported care for the national interest as defined by the military is indeed one of their main reasons for intervening [@finer_man_1962].
 Finer's suggestion was that because of the perils of professionalism as espoused by Huntington, what was really needed to keep the military out of politics was a deep internalization of "the principle of civil supremacy" [@finer_man_1962]^[
-Importantly, neither Huntington's concept of professionalism nor Finer's critique of the concept are arguing for blind obedience by the military to civilian orders and directions, a point that is, in the author's opinion, often missed in debates about civil-military relations.
+Importantly, neither Huntington's concept of professionalism nor Finer's critique of the same are arguing for blind obedience by the military to civilian orders and directions, a point that is, in the author's opinion, often missed in debates about the perils of professionalism.
 On the contrary, both Huntington and Finer make room and acknowledge that there are instances when military disobedience is not only required, but justified.
 However, both agree that in general terms, these situations should be exceedingly rare.]
-
 It is not only in the academic literature that the tradition and worth of an apolitical military has been cultivated.
-The operational military has also been central in upholding the military's non-partisan traditions.
-Retired Army General Douglas MacArthur, a hero of World War Two in the Pacific who is also well known for the controversy surrounding his conduct in the Korean War and his eventual dismissal by President Truman, told the cadets at West Point in May of 1962.
+The operational military has also been central in upholding the military's non-partisan traditions, albeit with a far less than perfect record.
+Retired Army General Douglas MacArthur, a hero of World War Two in the Pacific and who later commanded the occupation of Japan, is well known for his disagreement with President Truman regarding American strategic goals during the Koran War.
+Yet his remarks to cadets at West Point as an 82 year old retired officer are also well known and applicable to the topic at hand:
 
@@
 
-Current military doctrine and practices likewise reinforce the notion of the military remaining apolitical.
+Current military doctrine and practices likewise reinforce the notion of the military institution remaining apolitical.
 Prior to the 2016 Presidential Election, Marine Corps General Joseph Dunford, who then served as the Chairman of the Joint Chiefs of Staff, said that the basis upon which the military is able to maintain the public's trust is it's ability to remain "apolitical and neutral" [@dunford_upholding_2016].
-General Martin Dempsey, in an opinion piece published in the Washington Post on July 31, 2016, specifically criticized retired Army Major General Mike Flynn and Marine Corps General John Allen for participating in the Republican and Democratic conventions, respectively, stating that, "As generals, they have an obligation to uphold our apolitical traditions.
+Retired General Martin Dempsey, in an opinion piece published in _The Washington Post_ on July 31, 2016, specifically criticized retired Army Major General Mike Flynn and Marine Corps General John Allen for participating in the Republican and Democratic conventions, respectively, stating that, "As generals, they have an obligation to uphold our apolitical traditions.
 They have just made the task of their successors – who continue to serve in uniform and are accountable for our security - more complicated.
@@
 
-## Political Polarization, Public Opinion, and Military Expertise
+## Military Credibility and Expertise
 
-Scholars of American Politics have studied political polarization in depth, and while they have offered differing explanations for the causes of polarization, ranging from race and religion [@abramowitz_great_2018], to debates about the size and role of government [@sides_two_2015], to fierce policy disagreements over policy [@mann_its_2012], virtually all agree that polarization, if not an entirely new phenomenon, has become more pronounced in recent decades [@hare_polarization_2014; @bonica_mapping_2014].
-Hare and Poole are among scholars who argue that modern American political polarization began in the 1970s as a result of redistribution along racial lines following key pieces of legislation in the 1960s [@hare_polarization_2014].
-They further argue that polarization disagreement has only intensified, something that a significant portion of modern Americans agree on [@schaeffer_far_nodate].
+Military credibility remains high among the US population, though this is not something that has come easy for the military.
+Consider that in the late 1970s, as a likely result of public opinion over the Vietnam War, 54% of Americans indicated that they trusted the military "a great deal or quite a lot," whereas in 2019, 73% of Americas expressed the same sentiments [@gallup_polls_confidence_2007].
+Additionally, military credibility is high relative to other institutions.
+According to a recent Pew Research Center poll, 80% of the American public expects the military to act in the best interests of the American public, compared to business leaders (45%), the media (40%), and elected officials (25%) [@johnson_trust_2018].
 
-Within this polarized political environment, what role does the military play - particularly retired military officers who possess considerable military expertise?
+When credibility between the American population and the military as in institution is high, it can be inferred that the public trusts the military - to accomplish its military mission most of all, but also to do more than that - to develop healthy institutional cultures, to vigorously address unethical conduct, and to exercise good leadership while taking care of American service members [@us_army_repairing_2014].
+Senior military leaders are aware of the importance of maintaining public trust, and are generally sensitive to issues or concerns which may lead to a loss in confidence of the military [@us_army_repairing_2014].
+
+Scholars have also shown that when elites possess a high degree of credibility, their ability to influence and shape public opinion is likewise high.
 In his influential study, _The Nature and Origins of Mass Opinion_, John Zaller argues that formations in mass public opinion are the result of three interacting components: "variation in the information carried in elite discourse, individual differences in attention to this information, and individual differences in political values and other predispositions" [@zaller_nature_1992].
-Zaller argued that throughout modern America, most citizens, as non-elites, must rely heavily on the messages that come from elites regarding various topics and issue areas [@zaller_nature_1992].
+Zaller argues that throughout modern America, most citizens, as non-elites, must rely heavily on the messages that come from elites regarding various topics and issue areas [@zaller_nature_1992].
 Zaller further argues that the messages that elites send is necessarily "stereotyped" in some way: that is, the messages that elites send, in order to be understood by regular citizens who are part of the public, must be "short, simple, and highly thematic" [@zaller_nature_1992].
-In developing this argument, Zaller is not claiming that regular citizens are uninformed or careless people who are incapable of processing complex information.
-He is rather acknowledging that most citizens care most about the world that is not "beyond their personal experience" [@zaller_nature_1992].
-His point is that elites play a significant role in helping to form and sway public opinion.
+In developing this argument, Zaller is not claiming that regular citizens are uninformed or careless people who are incapable of processing complex information, but rather acknowledging that most citizens care most about a world that is not "beyond their personal experience" [@zaller_nature_1992].
+Thus, his main point is simply that elites play a significant role in forming and swaying public opinion.
 
-Michael Robinson further examined how elite cues are received and processed.
-Robinson argues that partisan groups, as a result of ongoing political polarization, are cognitively biased in distinct ways in how they receive and process information [@robinson_who_2019].
+## Political Polarization
+
+Almost all accounts of polarization in the scholarly literature point to American society becoming increasingly polarized.
+While scholars have offered differing explanations for the causes of polarization, ranging from concerns over race and religion [@abramowitz_great_2018], to the size and role of government [@sides_two_2015], to fierce policy disagreements over policy issues [@mann_its_2012], virtually all scholars agree that polarization in America has become more pronounced in recent decades [@hare_polarization_2014; @bonica_mapping_2014], and by many measures, is worse than it has been during any period since the end of the American Civil War [@hare_polarization_2014].
+Scholars have also shown that whereas polarization in previous decades was pronounced over singular and specific issue areas, to include redistribution policies following key Civil Rights legislation passed in the 1960s [@hare_polarization_2014], recent polarization is marked by a phenomenon known as "conflict extension," [@layman_party_2002], referring to the notion that a far wider set of issues and policy preferences covering the full gamut of social welfare, racial, and cultural questions are taking on a polarizing character [@layman_party_2006].
+
+The 2016 Presidential Election, resulting in the Trump presidency, confirms the nature and extent of polarized politics in America.
+Political scientists, journalists, and reporters are still trying to piece together the specific factors that led to Donald Trump's surprise victory.
+Sides, Tesler, and Vavreck argue that Trump's election reveals just how split Americans are, particularly on core issues of identity and attitudes regarding race and religion [@sides_identity_2018].
+During the years of the Trump Presidency, moreover, the American people have indicated that polarization is an issue whose presence is being felt and worsening [@schaeffer_far_2020].
+Finally, given that Americans feel more strongly today about more issues and especially those issues from which individuals draw a sense of core identity, it is hardly surprising that individuals seem to apply partisan lenses, or biases, to what they hear.
+Michael Robinson showed as a result of political polarization, partisan groups are cognitively biased in how they receive and process information [@robinson_who_2019].
 As a result, Robinson found that partisans will express confidence in the military in vastly differing ways even after neutral, non-partisan information about various types of military failure is presented to these groups [@robinson_who_2019].
-If Republicans are less likely to express a reduction in confidence in the military as an institution than Democrats are, as Robinson alleges, then even accurate, unbiased information is potentially irrelevant, as these differing partisan groups are pre-programmed to selectively see a particular situation as they wish to see it [@robinson_who_2019].
-His main point serves to warn us regarding confidence in the military as an institution: there may be real reasons to lose confidence in the military as a result of a particular incident, but partisan filters will prevent a reduction of confidence from actually being expressed.
 
-Zaller's framework makes it clear that it is a perfectly reasonable assumption that retired military officers who choose to express their opinions through the form of opinion-editorial pieces constitutes a form of elite discourse.
-After all, retired military officers have at least twenty years of experience, and in the case of most multi-star generals or admirals, usually more than 30 and in some cases close to 40 years of military experience.
-Indeed, most retired senior generals and admirals have spent their professional careers solely in the military, and have extensive experience while serving in leadership positions of large organizations and, in many cases, during times of war.
-Retired military officers are thus elites in the field of national security, and in line with Zaller's framework, they possess the potential to significantly contribute to policy debates and thus, to influence public opinion.
-Yet we also know that the American public will not necessarily evaluate the opinions of these military experts objectively, but rather through pre-conceived and potentially partisan lenses [@robinson_who_2019].
+## When Words Become Weaponized - Theory
+
+The purpose of this section is to explain how the concepts of professionalism, credibility, and political polarization can interface to create a potentially dangerous atmosphere for both military professionals and the public.
+It must be stressed from the outset that none of these concepts in isolation impose a normative prohibition on the act of a military officer taking up the pen to write an opinion piece.
+Indeed, Huntington's concept of military professionalism recognizes the unique skills of military officers, and Zaller's framework of elite discourse tacitly affirms that retired military officers, especially generals and admirals, most of whom possess more than 30 and in some cases 40 years of professional experience, possess relevant expertise that the public would find helpful, relevant, and convincing as it pertains to a host of potential policy debates.
+Even the existence of political polarization itself is insufficient to suggest that retired military officers should never give their opinion on specific matters.
+
+One can reasonably infer that a professional opinion piece written by a military officer would then take on three very important qualities.
+First, stemming from the concept of military professionalism, the piece would address a relevant military topic.
+To be clear, multistar generals and admirals often are experienced in the realm of strategy, a word that is often misunderstood and misapplied in discussions of the topic [see for example @strachan_direction_2014].
+Without unpacking the concept of strategy further, the relevant point is that military elites, and especially those who achieved a high rank, have been privy to decisions and experiences occurring near the highest levels of government.
+Their body of experiences is not confined to mere tactics.
+They They are familiar with issues ranging from how wars should be best fought, to future security threats looming on the horizon, to fixing the institutional culture of a military service.
+Second, from a credibility and expertise perspective, a professional piece would be based on the officer's personal experiences and would be written from an individual's perspective so as not to infer that an argument is being made on behalf of the collective service or element within the military body.
+Third, an opinion made by a military officer would be free of any hint of partisan sentiment, whether advocacy or denouncement.
+If one of these three qualities - addressing a relevant military topic, from the perspective of an individual, and free of partisan expression - fails to be upheld, there is increasing likelihood that the officer's words can serve as a weapon, and, like other kinds of weapons, can do great harm.
+
+My theory holds that when military credibility is high, a demand signal is created by the American public such that future interactions between military elites and the military institution are encouraged and welcomed.
+There are numerous ways in which these interactions might satisfy the public's demand to observe or hear from the military.
+Institutionally, the military might respond to the public's demand signal for increased engagement by participating in public ceremonies, parades, and other showcases of military prowess (i.e.
+at sporting venues).
+Individually, military elites, especially retired military officers, would be more eager to dialogue with the American public when military credibility is high.
+Press and media outlets would be more eager to showcase elite military voices when military credibility is high because they represent an institution that commands a high degree of public respect.
+When military credibility is low, the opposite will occur.
+Since demand for interactions between the American public and the military is muted, the institution will not participate in a high number of public showcases of military prowess because there will not be as strong of a demand signal sent by the public.
+In short, military credibility and respect for military expertise impacts the likelihood of the military institution interacting with the American public.
+When respect for the military is high, the military is likely to speak out and to interact with the public, and when respect for the military is low, the military is less likely to do so.
+
+When political polarization is high, I contend that it ultimately places great strain on the concept and definition of loyalty because the distinction between what is considered morally right or wrong and partisan expression becomes blurred if not altogether removed.
+The result is a less unified military institution.
+I further contend that these mechanisms work as the logical and expected outworking of the characteristics of the polarized American political environment.
+While scholarly literature has looked at how the military faces pressure stemming from divided institutional loyalties [@betts_soldiers_1991], the dynamic of polarization promises to add even more pressure and scrutiny on the loyalty of individual officers and the military as a whole.
+
+I contend that the the line between what is considered by individuals to be morally right and wrong and sentiments of partisan expression becomes blurred in highly polarized environments.
+As discussed above, political polarization has caused Americans to move further apart ideologically on an increasing number of issues [@layman_party_2006], including on core attitudes regarding issues such as race and immigration that are central to one's identity [@sides_identity_2018].
+In this environment, individuals believe the stakes are raised to a very high level for a number pivotal concerns.
+Individuals in a polarized environment, I contend, also believe there are enormous consequences if the specific side with which they ally on any given issue loses to the other side because the other side is not just simply mistaken, but deeply and morally wrong.
+
+This dynamic impacts military officers in a unique way because military officers, unlike regular citizens, operate under strict codes of loyalty all of the time.
+The American military officer, for example, takes an oath to support and defend the Constitution of the United States.
+In a polarized environment, two different officers may adopt completely different stances on a contentious issue, each believing the other is morally wrong and operating outside of the framework of the constitution.
+Each officer, then, might accuse the other of failing to be loyal to the ideals of the military.
+As a result, when polarization is high, the likelihood of partisan expression is high because in doing so, members of the military institution may feel that doing so is the right and loyal thing to do.
+
+As a quick case in point, consider that in June 2020, following President Trump's largely unpopular move of displacing protesters from Lafayette Square, a letter published by West Point alumni expressly called out prominent West Point alumni serving in the Trump administration.
+Though it did not mention individuals by name, the letter was intended and acknowledged by media as directed towards the Secretary of Defense, Mark Esper, and the Secretary of State, Mike Pompeo, who are both graduates of West Point's Class of 1986 [@gaouette_group_nodate].
+The charges of the letter were both damning and politically striking: "...We are concerned that fellow graduates serving in senior-level, public positions are failing to uphold their oath of office and their commitment to Duty, Honor, Country"[@line_letter_2020].
+The letter's clear implication is that if you are serving in the Trump administration and if you graduated from the US Military Academy (i.e.
+Esper and Pompeo), you have forsaken your oath of office, the US Constitution, and the values that West Point stands for.
 
@@
 
-Given these three parameters -- that retired military officers are part of an institution that highly values professionalism and non-partisanship, that political polarization in America has increased since at least the 1970s, and that retired military officers are elites who possess the potential to sway public opinion -- I now aim to see whether the discourse of retired military officers has changed over time, specifically since the 1970s, as polarization has worsened.
+I now aim to see whether the discourse of retired military officers has changed over time, specifically since the 1970s, as polarization has worsened.
 To answer this question, I assemble and examine a data set that consists of opinion pieces authored by retired military officers in major US newspapers.
@@
 In the author's opinion, whether the American military can, and to what degree it will, be able to remain professional as an institution remains to be seen.
-As political polarization becomes increasingly salient, several dynamics are unfolding that are worth noting, as their full impact has yet to materialize.
+As political polarization becomes increasingly salient, several dynamics are unfolding, all of which deserve attention in future scholarship.
 
-The first is that as political polarization worsens, the entire notion of loyalty is being reshaped, redefined, or at the very least, called into question.
-Consider that in June 2020, following the President's largely unpopular move of displacing protesters from Lafayette Square, a letter published by West Point alumni expressly called out prominent West Point alumni serving in the Trump administration.
-Though it did not mention individuals by name, the letter was intended and acknowledged by media as directed towards the Secretary of Defense, Mark Esper, and the Secretary of State, Mike Pompeo, who are both graduates of West Point's Class of 1986 [@gaouette_group_nodate].
-The charges of the letter were both damning and politically striking: "...We are concerned that fellow graduates serving in senior-level, public positions are failing to uphold their oath of office and their commitment to Duty, Honor, Country"[@line_letter_2020].
-The letter's clear implication is that if you are serving in the Trump administration, and if you graduated from the US Military Academy, you have forsaken your oath of office, and by extension, the US Constitution, as well as the values that West Point stands for.
-Increasing political polarization promises to make these charges of misplaced loyalty more common.
-Scholarly literature has looked at how the military faces pressure stemming from divided institutional loyalties [@betts_soldiers_1991], but polarization promises to add even more pressure and scrutiny on the loyalty of individual officers and the military as a whole.
-
-A second dynamic, that of maintaining credibility and trust with the American public, remains crucial for the military, but this dynamic will likely be tested in fierce ways should a polarized political climate endure.
+The first dynamic, that of maintaining credibility and trust with the American public, remains crucial for the military, but this dynamic will likely be tested in fierce ways should a polarized political climate endure.
 Even as high credibility with the American public can give the military support, it can also be abused, especially if military officers, to include retirees, use the high level of credibility they currently enjoy to take up and advocate for partisan positions, or worse, to jockey with civilian leaders for power [@brooks_shaping_2008-1].
 Unfortunately, as this paper illustrates, there is evidence that mild aspects of this are occurring, but the extent to which it is, and its resulting impacts, remains an area for future research.
+
 Credibility is not something that should be shunned.
 The military as an institution, to include military retirees, holds relevant and military expertise that should be afforded to the American public at large.
-And yet, just as retired officers were skilled in the management of violence when they wore the uniform every day, retirement in a polarized era means that they must so also be managers of American trust.
+And yet, just as retired officers were skilled in the management of violence when they wore the uniform every day, retirement in a polarized era means that they must also become managers of American trust.
 The promising news is, as this paper makes clear, the great proportion of retirees who are speaking out in various settings know and internalize this.
@@
 
-A third, and perhaps the most important dynamic of polarization as it concerns the military, is that the line between what is considered partisan or political and what is considered morally right and wrong is becoming seemingly blurred, if not altogether disappearing.
-Polarized sides tend to believe that each is acting "rightly" and that the other side is not just mistaken, but deeply wrong in a moral sense.
-
-These three factors are difficult in a purely domestic political environment, and one does not have to imagine how the potential snares of a polarized political environment and military credibility stewarded poorly could impede the military's ability to do it's main task, fighting and winning this nation's wars.
-Indeed, the impact of these characteristics on our nation's military effectiveness must also be considered in further research.
-Regardless, the importance of individual moral courage will continue to grow in value as an attribute of military leaders.
-Its importance as a trait will likely continue to grow in the years ahead.
+The second, and perhaps more important dynamic, is understanding the impacts of the polarized environment on the military's ability to perform its primary military task, i.e.
+fight and win the nation's wars.
+Scholars, military officers, and practitioners must better understand what impacts, if any, are placed on military and civilian leaders as a result of political polarization on the development and implementation of war goals, aims, and strategies, and how military leaders and civilian policymakers can overcome these hurdles to generate effective policies [@rapp_civil-military_2015].
 

### 2020-07-09 code/Submissionv1.Rmd
commit changes

@@
 abstract: |
+
   _This paper explores variation in the professionalism of opinion pieces authored by retired US military officers and published in major US newspapers in the past 40 years.
@@
 Finally, the paper develops a theory that shows how the forces of political polarization and military credibility interact, and argues that the norm of a professional, apolitical military will remain under further strain in the current US political environment._
-
 geometry: margin = 1.15in              # margin widths
@@
 library("knitr")
+library("ggdag")
+library("dagitty")
 ```
@@
 
-Almost all accounts of polarization in the scholarly literature point to American society becoming increasingly polarized.
-While scholars have offered differing explanations for the causes of polarization, ranging from concerns over race and religion [@abramowitz_great_2018], to the size and role of government [@sides_two_2015], to fierce policy disagreements over policy issues [@mann_its_2012], virtually all scholars agree that polarization in America has become more pronounced in recent decades [@hare_polarization_2014; @bonica_mapping_2014], and by many measures, is worse than it has been during any period since the end of the American Civil War [@hare_polarization_2014].
+Although scholars have offered differing explanations for the causes of polarization, ranging from concerns over race and religion [@abramowitz_great_2018], to the size and role of government [@sides_two_2015], to fierce policy disagreements over policy issues [@mann_its_2012], virtually all scholars agree that polarization in America has become more pronounced in recent decades [@hare_polarization_2014; @bonica_mapping_2014], and by many measures, is worse than it has been during any period since the end of the American Civil War [@hare_polarization_2014].
 Scholars have also shown that whereas polarization in previous decades was pronounced over singular and specific issue areas, to include redistribution policies following key Civil Rights legislation passed in the 1960s [@hare_polarization_2014], recent polarization is marked by a phenomenon known as "conflict extension," [@layman_party_2002], referring to the notion that a far wider set of issues and policy preferences covering the full gamut of social welfare, racial, and cultural questions are taking on a polarizing character [@layman_party_2006].
@@
 If one of these three qualities - addressing a relevant military topic, from the perspective of an individual, and free of partisan expression - fails to be upheld, there is increasing likelihood that the officer's words can serve as a weapon, and, like other kinds of weapons, can do great harm.
+
+```{r, include=FALSE, echo=FALSE}
+problem_dag <- dagify(
+    Y ~ X1,
+    X1 ~ X2 + X3,
+    X2~X4,
+    X3~X4,
+    coords = tribble(
+      ~ name,      ~ x,    ~ y,
+      "Y", 1, .5,
+      "X1", .7, .5,
+      "X2", .3, 1,
+      "X3", .3,0,
+      "X4", 0, .5
+    ))
+
+problem_dag %>% 
+  ggplot(aes(x=x, y=y, xend=xend, yend=yend))+
+  geom_dag_edges()+
+  geom_dag_point()+
+  geom_dag_text_repel(aes(label=name), show.legend=FALSE)+
+  theme_dag()
+```
+
+```{r include=TRUE, echo=FALSE}
+
+problem_dag %>% tidy_dagitty() %>% 
+  dag_label(labels=c("Y"="Military\n Outcomes", 
+                     "X1"="Military\n Professionalism", 
+                     "X2"="Political\n Polarization", 
+                     "X3"="Institutional\n Credibility", 
+                     "X4"="Major Changes\n In American Society")) %>% 
+  ggplot(aes(x=x, y=y, xend=xend, yend=yend))+
+  geom_dag_edges()+
+  geom_dag_point()+
+  geom_dag_text()+ 
+  geom_dag_label_repel(aes(label=label, fill=label), col="white", show.legend=FALSE)+ theme_dag() + labs(title="Figure 1", subtitle="The Impact of Political Polarization and Institutional Credibility on Military Professionalism")
+
+```
 
@@
 Though it did not mention individuals by name, the letter was intended and acknowledged by media as directed towards the Secretary of Defense, Mark Esper, and the Secretary of State, Mike Pompeo, who are both graduates of West Point's Class of 1986 [@gaouette_group_nodate].
-The charges of the letter were both damning and politically striking: "...We are concerned that fellow graduates serving in senior-level, public positions are failing to uphold their oath of office and their commitment to Duty, Honor, Country"[@line_letter_2020].
+The charges of the letter were both damning and politically striking: "...We are concerned that fellow graduates serving in senior-level, public positions are failing to uphold their oath of office and their commitment to Duty, Honor, Country"[@noauthor_letter_2020].
 The letter's clear implication is that if you are serving in the Trump administration and if you graduated from the US Military Academy (i.e.
 Esper and Pompeo), you have forsaken your oath of office, the US Constitution, and the values that West Point stands for.
+
+Ultimately, four possible worlds exist when the impacts of military credibility and political polarization are considered in tandem (see Table 1, below).
+The first is...the second is...The third is....and the fourth is....
+
+\singlespacing
+\begin{table}[ht]
+\begin{footnotesize}
+\caption{Four Types of "Military Professionalism Worlds" Where Institutional Credibility and Political Polarization Vary} % title of Table
+\centering % used for centering table
+\begin{tabular}{|p{5cm}|p{5cm}|p{5cm}|p{5cm}|}
+    \hline
+    Influence of Variables & \textbf{Low Institutional Credibility} & \textbf{High Institutional Credibility}  \\ \hline
+    \textbf{Low Political Polarization} & Normal.
+Traditional Professionalism is Expected by the Military, Civilian Leaders, and the American Public & Possible Impacts.
+There is a demand for military voice, but little substantive matter that can cause crucial disagreement.
+\\ \hline
+    \textbf{High Political Polarization} & Possible Impacts.
+There are issues and concerns which could cause substantive disagreement between the military and civilian leaders, but military professionals are constrained from voicing their opinions and/or interfacing with the public due to low institutional credibility with the American public &  Highly Likely Impacts to Professionalism.
+Military professionals have a "platform with the public" due to high institutional credibility, and high polarization raises the likelihood of disagreement between the military and civilian leaders on a number of crucial topics.\\ \hline
+\end{tabular}
+\label{table:comparison} % is used to refer this table in the text
+\end{footnotesize}
+\end{table}
+
+\doublespacing
 
@@
 
-```{r fig1, fig.height=5, fig.width=7, fig.align="center", include=TRUE, echo=FALSE}
+```{r fig2, fig.height=5, fig.width=7, fig.align="center", include=TRUE, echo=FALSE}
 datatograph <- data.frame(
@@
     color = NULL,
-    title = "Figure 1: # of Op-Eds Written By Military Retirees",
+    title = "Figure 2: # of Op-Eds Written By Military Retirees",
     subtitle = "On Topics of Social Policy, Domestic Policy, or Individual Figures"
@@
 
-```{r fig2, fig.height=4, fig.width=5, fig.align="center", include=TRUE, echo=FALSE}
+```{r fig3, fig.height=4, fig.width=5, fig.align="center", include=TRUE, echo=FALSE}
 datatograph <- data.frame(
@@
     color = NULL,
-    title = "Figure 2: # of Op-Eds Written By Military Retirees",
+    title = "Figure 3: # of Op-Eds Written By Military Retirees",
     subtitle = "Failed to Uphold Professional Standards"
@@
 
-```{r fig3, fig.height=4, fig.width=5, fig.align="center", include=TRUE, echo=FALSE}
+```{r fig4, fig.height=4, fig.width=5, fig.align="center", include=TRUE, echo=FALSE}
 datatograph <- data.frame(
@@
     color = NULL,
-    title = "Figure 3: Total Number of Op-Eds Written By Military Retirees",
+    title = "Figure 4: Total Number of Op-Eds Written By Military Retirees",
     subtitle = "WSJ, NY Times, Washington Post, USA Today, and LA Times"

### 2020-07-09 code/Submissionv1.Rmd
commit changes

@@
 
+Figure 1, below, outlines the general theory and framework of the impact of political polarization and institutional credibility on military professionalism, and subsequently, on various military outcomes.
+The system of variables represented in Figure 1 constitute the "drive chain" that can lead to the words of retired military officers becoming militarized.
+
+When military credibility is high, a demand signal is created by the American public such that future interactions between the military institution and the American public are encouraged and welcomed.
+From an institutional perspective, the military might respond to the public's demand signal for increased military interaction by increasingly participating in public ceremonies, parades, and other showcases of military prowess (at sporting venues, for instance).
+From an individual perspective, military elites, especially retired military officers, would be psychologically primed to increasingly dialogue with the American public when military credibility is high.
+By extension, other outlets such as the Press would also be primed to showcase military elites when military credibility runs high, because these elites represent an institution that commands a high degree of public respect.
+
+When military credibility is low, I theorize the opposite will occur.
+Since demand for interactions between the American public and the military is muted when institutional credibility for the military is low, the institution will not engage with the public as frequently nor as intimately because the strong demand signal that existed when institutional credibility was high no longer exists.
+At the individual perspective, military elites will also not engage, nor will the Press ask these elites to, with the American public.
+In short, institutional credibility and respect for the military institution impacts the likelihood of the military institution and military elites interacting with the American public.
+Such interaction can take a variety of forms.
+When respect for the military is high, the military is likely to interact, speak to, and dialogue with the American public, and when institutional credibility is low, the military is less likely to do so.
+
 ```{r, include=FALSE, echo=FALSE}
@@
 ```
+The degree of political polarization evident also has the potential to impact the professionalism of the military as an institution.
+First, because of the phenomenon known as "conflict extension" [@layman_party_2002], which refers to the fact that the political parties have grown further apart from each other on an increasing number of issues, political polarization raises the number and scope of issues to disagree about.
+Research has consistently shown that the polarizing politics present in the United States today has caused voters who identify as liberals to be more liberal across multiple issue categories, and that the same holds for conservatives.
+The result is that in the past, the prospect of cooperation on issues that were not polarized remained high, whereas today, the prospect for cooperation in general is lower [@layman_activists_2010].
+As it concerns the military institution, "conflict extension" [@layman_party_2002] increases the likelihood that members of the military feel strongly about an increasing number of issues.
+In other words, I contend that today's polarized environment makes it more difficult for an institution that to be politically neutral when there are fewer issues to remain neutral over.
+At the individual level, the same is true: as the non-partisan space seems to shrink, individuals will be enticed to not remain neutral.
 
-My theory holds that when military credibility is high, a demand signal is created by the American public such that future interactions between military elites and the military institution are encouraged and welcomed.
-There are numerous ways in which these interactions might satisfy the public's demand to observe or hear from the military.
-Institutionally, the military might respond to the public's demand signal for increased engagement by participating in public ceremonies, parades, and other showcases of military prowess (i.e.
-at sporting venues).
-Individually, military elites, especially retired military officers, would be more eager to dialogue with the American public when military credibility is high.
-Press and media outlets would be more eager to showcase elite military voices when military credibility is high because they represent an institution that commands a high degree of public respect.
-When military credibility is low, the opposite will occur.
-Since demand for interactions between the American public and the military is muted, the institution will not participate in a high number of public showcases of military prowess because there will not be as strong of a demand signal sent by the public.
-In short, military credibility and respect for military expertise impacts the likelihood of the military institution interacting with the American public.
-When respect for the military is high, the military is likely to speak out and to interact with the public, and when respect for the military is low, the military is less likely to do so.
+The second impact of political polarization on military professionalism deals with the fact that the issues which are becoming polarized are not merely matters of simple policy preferences, but substantive issues which are directly related to notions of a person's core identity, such as race and immigration [@sides_identity_2018], religion [@white_values_2002], and the future of the country [@mann_its_2012].
+These issues do not lend themselves to compromise.
+They are, for many people, using an economics phrase, "indivisible goods." This also explains the notion of "negative partisanship," which refers to the very strong negative feelings that exist on each side of the political spectrum for those on the other [@abramowitz_great_2018, 5].
+Citizens feel and believe that the stakes are raised to an extremely high level for a number of pivotal issue areas, and that there are enormous consequences if the specific side with which they ally on any given issue loses to the other side.
+Polarization, then, raises the salience of right and wrong in a moral sense, which increases the odds of expressing partisan sentiment.
 
-When political polarization is high, I contend that it ultimately places great strain on the concept and definition of loyalty because the distinction between what is considered morally right or wrong and partisan expression becomes blurred if not altogether removed.
-The result is a less unified military institution.
-I further contend that these mechanisms work as the logical and expected outworking of the characteristics of the polarized American political environment.
-While scholarly literature has looked at how the military faces pressure stemming from divided institutional loyalties [@betts_soldiers_1991], the dynamic of polarization promises to add even more pressure and scrutiny on the loyalty of individual officers and the military as a whole.
+The third and most important way in which political polarization impacts military professionalism, I contend, extends from the first two reasons outlined above.
+First, because there are less issue areas in general to remain politically neutral about, and second, because issues are increasingly viewed in terms of moral right and wrong, political polarization impacts and distorts the notions of two vital character traits: loyalty and duty.
+As any soldier knows, few traits are more sacred in the military than loyalty and duty.
+In a very real sense, the military professional and the institution at large see loyalty and duty as incentives.
+By this, I simply mean that military professionals pride themselves on living up to mantras and phrases which extol the virtues of "doing the right thing," "leading with integrity," and, as the West Point Cadet Prayer so elegantly states, choosing the "harder right over the easier wrong."
 
-I contend that the the line between what is considered by individuals to be morally right and wrong and sentiments of partisan expression becomes blurred in highly polarized environments.
-As discussed above, political polarization has caused Americans to move further apart ideologically on an increasing number of issues [@layman_party_2006], including on core attitudes regarding issues such as race and immigration that are central to one's identity [@sides_identity_2018].
-In this environment, individuals believe the stakes are raised to a very high level for a number pivotal concerns.
-Individuals in a polarized environment, I contend, also believe there are enormous consequences if the specific side with which they ally on any given issue loses to the other side because the other side is not just simply mistaken, but deeply and morally wrong.
+Political polarization obfuscates these ideals, however, and places greater responsibility for interpreting the meaning and the associated action of these ideals on the individual, and less on the military institution.
+It is in this way, I contend, that political polarization raises the chance of the military professional acting in such a way that simultaneously expresses partisan sentiment, because in doing so, the military professional believes he or she is doing what is morally right.
 
-This dynamic impacts military officers in a unique way because military officers, unlike regular citizens, operate under strict codes of loyalty all of the time.
-The American military officer, for example, takes an oath to support and defend the Constitution of the United States.
-In a polarized environment, two different officers may adopt completely different stances on a contentious issue, each believing the other is morally wrong and operating outside of the framework of the constitution.
+This is why in a politically polarized environment, two different military officers may adopt completely different opinions of a contentious issue, each believing that the other's opinion is wrong.
+Should the officer act on that opinion, the other might believe he or she is operating outside of established military frameworks, the constitution, or any number of structures.
 Each officer, then, might accuse the other of failing to be loyal to the ideals of the military.
@@
 The letter's clear implication is that if you are serving in the Trump administration and if you graduated from the US Military Academy (i.e.
-Esper and Pompeo), you have forsaken your oath of office, the US Constitution, and the values that West Point stands for.
+Secretaries Esper and Pompeo), you have forsaken your oath of office, the US Constitution, and the values that West Point stands for.
 
 Ultimately, four possible worlds exist when the impacts of military credibility and political polarization are considered in tandem (see Table 1, below).
-The first is...the second is...The third is....and the fourth is....
+The first is when political polarization and military credibility are both low.
+When these conditions hold, there is no anticipated drop in military professionalism by the military institution nor its individual elites.
+With these conditions, the military does not have an out-sized platform to interact with the American public, nor are there outstanding incentives for the military to raise significant disagreement over any domestic political issues.
+
+The second possible set of conditions is when political polarization is low and institutional credibility is high.
+When these set of conditions exist, the military has a platform to interact with the public, as the public highly respects the military institution.
+The military enjoys this high level of trust and does not want to lose it.
+Further, because the degree of political polarization is low, there is little incentive for the military as an institution or military elites to use the trust they enjoy with the American public to adopt or denounce an issue associated with any partisan entity.
+The high institutional credibility increases the chance of the military interacting with the public in an unprofessional way, but the low political polarization removes the incentives of doing so.
+
+The third type occurs when political polarization is high, but institutional credibility of the military is low.
+Under these conditions, the military does not command the respect of the American public like it does when credibility of the institution is high, and thus, does not possess a considerable outlet to interact with the American public.
+However, the incentives of loyalty and duty still loom large in the minds of some military elites as a result of the high degree of political polarization, and thus, a dip in professionalism is still possible.
+
+The final set of conditions exists when political polarization and institutional credibility are high.
+It is in this world that I contend the American military finds itself in now in 2020.
+Under these conditions, the military faces dual dangers.
+High institutional credibility risks the military institution unilaterally acting or another body co-opting the military institution to speak out an a number of issue areas, and simultaneously, the highly polarized environment risks the military advocating for or denouncing partisan sentiments in the name of upholding the very traits it is most adhered to - loyalty and duty.
+Indeed, this set of conditions comprise a very dangerous world for the military institution and the military professional.
+I
 
@@
 The second, and perhaps more important dynamic, is understanding the impacts of the polarized environment on the military's ability to perform its primary military task, i.e.
-fight and win the nation's wars.
+fighting and winning the nation's wars.
 Scholars, military officers, and practitioners must better understand what impacts, if any, are placed on military and civilian leaders as a result of political polarization on the development and implementation of war goals, aims, and strategies, and how military leaders and civilian policymakers can overcome these hurdles to generate effective policies [@rapp_civil-military_2015].
+While scholarly literature has looked at how the military faces pressure stemming from divided institutional loyalties [@betts_soldiers_1991], none have explored the unique ways in which polarization as a domestic political force shapes individual and institutional behavior of the military.
 

### 2020-07-09 code/Submissionv1.Rmd
commit changes

@@
 
+\singlespacing
+\begin{table}[ht]
+\begin{footnotesize}
+\caption{Four Types of "Military Professionalism Worlds" Where Institutional Credibility and Political Polarization Vary} % title of Table
+\centering % used for centering table
+\begin{tabular}{|p{5cm}|p{5cm}|p{5cm}|p{5cm}|}
+    \hline
+    Influence of Variables & \textbf{Low Institutional Credibility} & \textbf{High Institutional Credibility}  \\ \hline
+    \textbf{Low Political Polarization} & Negligible anticipated impacts.
+Traditional Professionalism is Expected by the Military, Civilian Leaders, and the American Public & Possible Impacts.
+There is a demand for military voice, but little substantive matter stemming from US domestic politics that can cause crucial disagreement.
+Little incentive to deviate from traditional concepts of loyalty and duty.
+\\ \hline
+    \textbf{High Political Polarization} & Possible Impacts.
+Polarization raises the stakes of domestic politics, but military professionals are constrained from voicing their opinions and/or interfacing with the public due to low institutional credibility with the American public &  Highly Likely Impacts to Professionalism.
+Likely impacts.
+Military professionals have a "platform with the public" due to high institutional credibility, and high polarization raises the likelihood of disagreement between individual military elites and civilian leaders on a number of crucial topics.\\ \hline
+\end{tabular}
+\label{table:comparison} % is used to refer this table in the text
+\end{footnotesize}
+\end{table}
+\doublespacing
 The second possible set of conditions is when political polarization is low and institutional credibility is high.
@@
 Indeed, this set of conditions comprise a very dangerous world for the military institution and the military professional.
-I
-
-\singlespacing
-\begin{table}[ht]
-\begin{footnotesize}
-\caption{Four Types of "Military Professionalism Worlds" Where Institutional Credibility and Political Polarization Vary} % title of Table
-\centering % used for centering table
-\begin{tabular}{|p{5cm}|p{5cm}|p{5cm}|p{5cm}|}
-    \hline
-    Influence of Variables & \textbf{Low Institutional Credibility} & \textbf{High Institutional Credibility}  \\ \hline
-    \textbf{Low Political Polarization} & Normal.
-Traditional Professionalism is Expected by the Military, Civilian Leaders, and the American Public & Possible Impacts.
-There is a demand for military voice, but little substantive matter that can cause crucial disagreement.
-\\ \hline
-    \textbf{High Political Polarization} & Possible Impacts.
-There are issues and concerns which could cause substantive disagreement between the military and civilian leaders, but military professionals are constrained from voicing their opinions and/or interfacing with the public due to low institutional credibility with the American public &  Highly Likely Impacts to Professionalism.
-Military professionals have a "platform with the public" due to high institutional credibility, and high polarization raises the likelihood of disagreement between the military and civilian leaders on a number of crucial topics.\\ \hline
-\end{tabular}
-\label{table:comparison} % is used to refer this table in the text
-\end{footnotesize}
-\end{table}
-
-\doublespacing
 

### 2020-07-13 code/Submissionv1.Rmd
commit changes

@@
   _This paper explores variation in the professionalism of opinion pieces authored by retired US military officers and published in major US newspapers in the past 40 years.
-Assembling and analyzing more than 330 opinion-editorial pieces published by retired military officers in The Wall Street Journal, The Washington Post, The New York Times, The Los Angeles Times, and the USA Today since 1979 reveals that there has been an increasing trend in published opinion pieces which fail to uphold professional standards as defined by military norms.
-The paper further argues that this trend is primarily the result of two significant strains on the military's politically neutral culture: political polarization and military credibility.
-Finally, the paper develops a theory that shows how the forces of political polarization and military credibility interact, and argues that the norm of a professional, apolitical military will remain under further strain in the current US political environment._
+In doing so, this paper develops a theory that argues that two societal forces, political polarization and military credibility, interact in unique ways that ultimately strain the individual military officer's concepts of loyalty and duty in important ways.
+Assembling and analyzing more than 330 opinion-editorial pieces published by retired military officers in The Wall Street Journal, The Washington Post, The New York Times, The Los Angeles Times, and the USA Today since 1979, the paper argues there has been an increase in published opinion pieces which fail to uphold professional standards as defined by historical military norms.
+The paper concludes by arguing that today's political polarization reveals a profoundly changing society, and that the norm of a professional, apolitical military will remain under further strain in the current political environment._
 geometry: margin = 1.15in              # margin widths
@@
 
-In late May and early June 2020, the American public was gripped by escalating racial tension and discussions about police brutality following the death of George Floyd, an unarmed black man, by police in Minneapolis.
-Protests, most of them peaceful, were accompanied in some cases by rioting and looting.
+In late May and early June 2020, racial tension and concerns about police brutality gripped the American public following the death of George Floyd, an unarmed black man, by police in Minneapolis.
+In several cities and communities across the nation, mostly peaceful protests were followed by riots and looting.
 On June 1, after giving remarks at the Rose Garden, President Donald Trump walked to St.
@@
 
-Retired military officers were among the elites who weighed in to give their opinions not only on the Trump administration's handling of specific events near the White House, but also on the broader issues of race in America.
+Additionally, several retired military officers were among the elites who weighed in to give their opinions not only on the Trump administration's handling of specific events near the White House, but also on the broader issues of race in America.
 But for those paying attention to what these retired officers said, it is also clear that there was substantial variation in the tone and tenor of each officer's remarks.
-While all of them expressed deep concern about the possibility of active-duty forces confronting peaceful protesters in clear violation of the First Amendment of the US Constitution, [@brooks_dismay_2020; @dempsey_former_2020], some officers went much further, leveling direct criticisms of the President's fitness to be President [@goldberg_james_2020], his administration's broader policies on race [@allen_moment_2020], and his general leadership abilities [@mullen_i_2020].
+While all of them expressed deep concern about the prospect of active-duty forces confronting peaceful protesters, and thus, violating protections held by the First Amendment of the US Constitution [@brooks_dismay_2020; @dempsey_former_2020], some officers went much further in their criticism, leveling direct insults of the President's fitness to lead [@goldberg_james_2020], his administration's broader policies on race [@allen_moment_2020], and his general leadership abilities [@mullen_i_2020].
 
@@
 It is an interesting question worthy of examination because, as I argue, at least three important factors are interacting in unique ways to shape how and why retired military officers speak out on a variety of issues.
-These factors are professionalism, polarization, and credibility.
-In particular, polarization and credibility are creating tremendous strains on the military's professional nature, and in some cases, are causing words to become weaponized.
+These factors are professionalism, political polarization, and military credibility.
+I argue that the latter two factors, political polarization and institutional credibility, are interacting in unique ways that create tremendous strains on the military's professional nature, and that these factors will continue to do so.
 
 This paper proceeds in four parts.
-First, I briefly review the concepts of military professionalism, polarization, and credibility in order to orient the reader to the theory that I build in part two.
-In the second section, I build a theory in which I argue that the contemporary forces of polarization and military credibility can strain the concept of professionalism such that words and actions become weaponized to the point of being antithetical to the very concept of professionalism itself.
-Third, I turn to the data, which consists of more than 330 opinion-editorial pieces published in major US newspapers by retired military officers since the late 1970s.
-While the data shows that the vast majority of these pieces are in fact in line with the notion of military professionalism, recent trends show that the retired military officers are weighing in on topics that fall outside of the realm of national security as well as expressing clear partisan preferences in their arguments.
-Fourth, I briefly explore the implications of these results on the military as an institution, and specifically, its ability to remain professional in a setting where polarization is high and the institution commands wide public support.
+First, the paper briefly reviews the concept of military professionalism in order to orient the reader to the central role this concept has played in the ethos of the American military for the past several decades.
+Second, the paper builds a theory that demonstrates how the uniqueness of today's political polarization and institutional credibility are interacting to create an environment that threatens professional military behavior in ways that could be highly destructive.
+Third, the paper explores the data, which consists of more than 330 opinion-editorial pieces published in major US newspapers by retired military officers since the late 1970s.
+The analysis ultimately shows that although the vast majority of op-eds published by retired military officers maintain professional standards, there is clear evidence that retired military officers are using the pen to both argue about topics that fall outside of the realm of traditional national security topics, and to express clear partisan preferences when making their arguments.
+Fourth, the paper concludes by briefly exploring the implications of these results on the military as an institution, and specifically, on its ability to remain professional in a domestic setting marked by high political polarization in which the military commands wide public support.
 
 \singlespace
-# Military Professionalism
+# Military Professionalism as a Foundational Trait
 \doublespace
-The ideal of the US military acting as an objective, apolitical and non-partisan institution is deep.
+The ideal of the US military acting as an objective, apolitical and non-partisan institution is storied and long.
 One of the strongest origins of the norm stems from the work of the political scientist, Samuel Huntington, who, in his 1957 work, _The Soldier and the State: The Theory and Politics of Civil-Military Relations_, asserted that two distinct forces would always be in tension with one another in a liberal democracy such as the United States.
 The first was a "functional imperative," which Huntington said drove the requirement that a military be strong enough and capable of defeating the military threats it faced, and the second was a "societal imperative," which Huntington said generated requirements for the military in that the military could never lose so many aspects of it such that it no longer represented the liberal democracy that it defended [@huntington_soldier_1957, 1-3].
-Huntington's point was that America needed a strong military that could win on the battlefield, but cautioned that such a military could not be so foreign in character so as to alienate the tenets of an American liberal society.
-The key to resolving the tension generated by these two opposing forces was to adopt a scheme of "objective civilian control" in which political leaders established the ends and goals of policy, while military leaders developed and implemented the military plans to accomplish these goals [@huntington_soldier_1957, 80-85].
+Huntington's point was that America needed a strong military that could win on the battlefield, but cautioned that such a military could not be so foreign in character so as to alienate the character of American liberal society.
+Huntington argued that the key to resolving the tension generated by these two opposing forces was to adopt a scheme of "objective civilian control" in which political leaders established the ends and goals of policy, while military leaders developed and implemented the military plans to accomplish these goals [@huntington_soldier_1957, 80-85].
 To Huntington, largely separate operating spheres between the political leaders of a democracy and the military as an institution would ensure that the military remained "professional," and thus, untainted by partisan interests and concerns.
@@
 While Huntington's theory has enjoyed tremendous staying power (Huntington is still studied and taught in many American military institutions), he is not without critics [see, for example, @feaver_civil-military_1996; @cohen_supreme_2003; @feaver_armed_2005; and @brooks_paradoxes_2020].
-It would be too burdensome to unpack all of those here, but one of the earliest and richest critiques of Huntington, and perhaps one of the most relevant given the potential influences of polarization and credibility, came from S.E.
+It would be too burdensome to unpack all of those here, but one of the earliest and richest critiques of Huntington, and perhaps one of the most relevant given the potential influences of political polarization and institutional credibility, came from S.E.
 Finer.
 Authored in 1962, Finer's book, _The Man on Horseback: The Role of the Military in Politics_, argued that Huntington's concept of professionalism simply was not strong enough to keep the military out of politics [@finer_man_1962].
-Finer's concern was that of military officers adopting an ethic of professionalism such that they would actually be more likely - not less likely - to become involved in political issues [@finer_man_1962, 22-23].
-Finer was especially concerned that the adoption of professionalism alone would lead to long-serving military professionals believing that they knew what was best for their country over and above that of temporary elected civilian leaders [@finer_man_1962, 22-23] Finer warned that "the moment the military draw this distinction between nation and the government in power, they begin to invent their own private notion of the national interest, and from this it is only a skip to the constrained substitution of this view for that of the civilian government…this purported care for the national interest as defined by the military is indeed one of their main reasons for intervening [@finer_man_1962].
-Finer's suggestion was that because of the perils of professionalism as espoused by Huntington, what was really needed to keep the military out of politics was a deep internalization of "the principle of civil supremacy" [@finer_man_1962]^[
-Importantly, neither Huntington's concept of professionalism nor Finer's critique of the same are arguing for blind obedience by the military to civilian orders and directions, a point that is, in the author's opinion, often missed in debates about the perils of professionalism.
+Finer foresaw military officers adopting an ethic of professionalism such that they would actually be more likely, not less likely, to become involved in political issues [@finer_man_1962, 22-23].
+
+Finer was especially concerned that the adoption of professionalism would lead to long-serving military professionals believing that they knew what was best for their country over and above that of temporary elected civilian leaders [@finer_man_1962, 22-23].
+Finer warned that "the moment the military draw this distinction between nation and the government in power, they begin to invent their own private notion of the national interest, and from this it is only a skip to the constrained substitution of this view for that of the civilian government…this purported care for the national interest as defined by the military is indeed one of their main reasons for intervening" [@finer_man_1962].
+In Finer's view, what was really needed to keep the military out of politics was not only the concept of professionalism, but rather a deep internalization of "the principle of civil supremacy" [@finer_man_1962].^[
+Importantly, neither Huntington's concept of professionalism nor Finer's critique of the same are arguing for blind obedience by the military to all civilian orders and directions, a point that is, in the author's opinion, often missed in debates about military professionalism.
 On the contrary, both Huntington and Finer make room and acknowledge that there are instances when military disobedience is not only required, but justified.
 However, both agree that in general terms, these situations should be exceedingly rare.]
+
 It is not only in the academic literature that the tradition and worth of an apolitical military has been cultivated.
-The operational military has also been central in upholding the military's non-partisan traditions, albeit with a far less than perfect record.
-Retired Army General Douglas MacArthur, a hero of World War Two in the Pacific and who later commanded the occupation of Japan, is well known for his disagreement with President Truman regarding American strategic goals during the Koran War.
-Yet his remarks to cadets at West Point as an 82 year old retired officer are also well known and applicable to the topic at hand:
+The operational military has also been central in upholding the military's non-partisan traditions, albeit with a far less-than-perfect record.
+Retired Army General Douglas MacArthur, a hero of World War Two in the Pacific and later, Commander of the occupation of Japan, is well known for his disagreement with President Truman regarding American strategic goals during the Korean War.
+And yet, the remarks MacArthur made to cadets at West Point in 1962, as an 82 year old retired officer, speak to the primacy of professional military conduct:
 
@@
 
-Current military doctrine and practices likewise reinforce the notion of the military institution remaining apolitical.
-Prior to the 2016 Presidential Election, Marine Corps General Joseph Dunford, who then served as the Chairman of the Joint Chiefs of Staff, said that the basis upon which the military is able to maintain the public's trust is it's ability to remain "apolitical and neutral" [@dunford_upholding_2016].
-Retired General Martin Dempsey, in an opinion piece published in _The Washington Post_ on July 31, 2016, specifically criticized retired Army Major General Mike Flynn and Marine Corps General John Allen for participating in the Republican and Democratic conventions, respectively, stating that, "As generals, they have an obligation to uphold our apolitical traditions.
+Finally, current military doctrine and practices also reinforce the centrality of professional, apolitical, and non-partisan conduct.
+Prior to the 2016 Presidential Election, Marine Corps General Joseph Dunford, who at the time served as Chairman of the Joint Chiefs of Staff, said that the basis upon which the military is able to maintain the public's trust is it's ability to remain "apolitical and neutral" [@dunford_upholding_2016].
+At nearly the same time, his predecessor, Retired Army General Martin Dempsey, published an opinion-editorial piece in _The Washington Post_ which criticized retired Army Major General Mike Flynn and Marine Corps General John Allen for participating in the Republican and Democratic conventions.
+Dempsey wrote, "As generals, they have an obligation to uphold our apolitical traditions.
 They have just made the task of their successors – who continue to serve in uniform and are accountable for our security - more complicated.
@@
 
-Although the notion of an apolitical military is the theoretical goal of the institution, the practical record remains far from perfect.
-From MacArthur's clash with President Truman in the early 1950's, to the so-called "Revolt of the Generals" in 2005-2006, military leaders, even those who are retired, have ventured into political territory from time to time.
-What we have established, though, is clear: a non-partisan military that performs its military mission well has long been the goal of both civil-military relations theorists and military officers.
-More importantly, this is what fosters and ultimately earns the trust and confidence of the American people.
+Although a professional and non-partisan military is the theoretical goal of the institution, the practical record and results remain far from perfect.
+From MacArthur's clash with President Truman in the early 1950's, to the so-called "Revolt of the Generals" in 2005-2006, military leaders, even those who are retired, have ventured into political territory on more than one occasion.
+Nonetheless, academic theorists and military leaders alike have long recognized the importance of professional conduct rooted in non-partisan, apolitical conduct.
+At its core, such conduct ultimately earns the trust and confidence of the American people.
+
+It must be stressed that the concept of military professionalism does not impose a normative prohibition on the publication of an opinion piece, or the expression of other forms of public dialogue, by a military professional (though there are certainly practical differences between the manner these are employed by those who currently serve and those who are retired).
+Indeed, Huntington's concept of military professionalism recognizes the unique skills of military officers, some of whom, especially multistar generals and admirals, possess more than 30 and in some cases 40 years of professional experience.
+
+It is reasonable to conclude that a professional opinion piece written by a military officer would take on three very important qualities.
+First, the military professional would address a relevant military topic.
+To be clear, multistar generals and admirals often are experienced in the realm of strategy, a word that is often misunderstood and misapplied in discussions of the topic [see for example @strachan_direction_2014].
+Without unpacking the concept of strategy further, the relevant point is that military elites, and especially those who have achieved a high rank, have been privy to decisions and experiences occurring near the highest levels of government, and as such, their experiences are not confined to only tactical matters.
+They are familiar with issues ranging from how wars should be best fought, to future security threats looming on the horizon, to fixing the institutional culture of a military service.
+Second, from a credibility and expertise perspective, a professional piece would be based on the officer's personal experiences and would be written from an individual's perspective so as not to infer that an argument is being made on behalf of the collective service or element within the military body.
+Third, an opinion made by a military officer would be free of any hint of partisan sentiment, whether advocacy or denouncement.
+If one of these three qualities fails to be upheld, there is increasing likelihood that the officer's words can, either intentionally or unintentionally, serve as a weapon, and, like other kinds of weapons, cause great harm.
+To understand how, we will now turn our attention to understanding the impacts on military professionalism caused by two significant changes that have developed within American Society in the past 50 years.
+The first is increasing credibility of the military institution, and the second is the introduction of an unparalleled and divisive political polarization.
+The interaction of these forces is significantly straining the professional conduct of the military institution.
+
+# The Impact of Credibility and Polarization
 
@@
 
-Military credibility remains high among the US population, though this is not something that has come easy for the military.
+One of the two significant societal changes that has occurred in America during the past fifty years is the increase in credibility of the military institution.
 Consider that in the late 1970s, as a likely result of public opinion over the Vietnam War, 54% of Americans indicated that they trusted the military "a great deal or quite a lot," whereas in 2019, 73% of Americas expressed the same sentiments [@gallup_polls_confidence_2007].
-Additionally, military credibility is high relative to other institutions.
+Compared to other government institutions, the American public trusts the military to a remarkably high degree.
 According to a recent Pew Research Center poll, 80% of the American public expects the military to act in the best interests of the American public, compared to business leaders (45%), the media (40%), and elected officials (25%) [@johnson_trust_2018].
+It is accurate to say that America trusts its military more than any other entity.
 
-When credibility between the American population and the military as in institution is high, it can be inferred that the public trusts the military - to accomplish its military mission most of all, but also to do more than that - to develop healthy institutional cultures, to vigorously address unethical conduct, and to exercise good leadership while taking care of American service members [@us_army_repairing_2014].
-Senior military leaders are aware of the importance of maintaining public trust, and are generally sensitive to issues or concerns which may lead to a loss in confidence of the military [@us_army_repairing_2014].
+Military leaders are, rightfully so, cognizant of the benefits of holding the public's trust.
+In the past several years for example, there has been a concerted effort by the services to identify and punish senior leaders who engage in misconduct, to demonstrate to Congress that senior military leaders do not need assistance to handle sexual assault, and to otherwise prove that it can be trusted to accomplish its missions [@us_army_repairing_2014].
+Former Chief of Staff of the Army, Retired General Ray Odierno, told fellow Army generals that "of great importance within the Army Profession is the duty of each of us to be a steward of our profession and to maintain its essential and hard won trust with society [Ray Odierno in @us_army_repairing_2014].
 
@@
 
-## Political Polarization
+Therefore, as the public's trust in the military has grown in the years since Vietnam, so has the military's potential to impact public opinion.
+This is especially true of military elites, to include retired military officers who previously held substantial rank and positions of authority.
+An increase in credibility of the institution may not indicate that the military or particular military elites will actually engage in behavior to deliberately impact public opinion, but rather that the potential to do so has increased.
+The American people tend to listen to those whom they trust.
 
-Although scholars have offered differing explanations for the causes of polarization, ranging from concerns over race and religion [@abramowitz_great_2018], to the size and role of government [@sides_two_2015], to fierce policy disagreements over policy issues [@mann_its_2012], virtually all scholars agree that polarization in America has become more pronounced in recent decades [@hare_polarization_2014; @bonica_mapping_2014], and by many measures, is worse than it has been during any period since the end of the American Civil War [@hare_polarization_2014].
-Scholars have also shown that whereas polarization in previous decades was pronounced over singular and specific issue areas, to include redistribution policies following key Civil Rights legislation passed in the 1960s [@hare_polarization_2014], recent polarization is marked by a phenomenon known as "conflict extension," [@layman_party_2002], referring to the notion that a far wider set of issues and policy preferences covering the full gamut of social welfare, racial, and cultural questions are taking on a polarizing character [@layman_party_2006].
+##Political Polarization Today - Unique in America's History
 
-The 2016 Presidential Election, resulting in the Trump presidency, confirms the nature and extent of polarized politics in America.
-Political scientists, journalists, and reporters are still trying to piece together the specific factors that led to Donald Trump's surprise victory.
-Sides, Tesler, and Vavreck argue that Trump's election reveals just how split Americans are, particularly on core issues of identity and attitudes regarding race and religion [@sides_identity_2018].
-During the years of the Trump Presidency, moreover, the American people have indicated that polarization is an issue whose presence is being felt and worsening [@schaeffer_far_2020].
-Finally, given that Americans feel more strongly today about more issues and especially those issues from which individuals draw a sense of core identity, it is hardly surprising that individuals seem to apply partisan lenses, or biases, to what they hear.
-Michael Robinson showed as a result of political polarization, partisan groups are cognitively biased in how they receive and process information [@robinson_who_2019].
-As a result, Robinson found that partisans will express confidence in the military in vastly differing ways even after neutral, non-partisan information about various types of military failure is presented to these groups [@robinson_who_2019].
+The second and decisive societal change that has occurred in the past fifty years is the degree to which and character of the political polarization prevalent in the US political landscape today.
+Although scholars have offered differing explanations for the causes of this polarization, ranging from concerns over race and religion [@abramowitz_great_2018], to the size and role of government [@sides_two_2015], to fierce policy disagreements over policy issues [@mann_its_2012], virtually all scholars agree that polarization in America has become more pronounced in recent decades [@hare_polarization_2014; @bonica_mapping_2014], and that it is, by many measures, worse than it has been during any period since the end of the American Civil War [@hare_polarization_2014].
+The polarized political climate today extends to far more issues than previously polarized periods in American history [@layman_party_2002], is marked by stronger negative feelings towards those who hold opposing views [@abramowitz_great_2018], and, at root, involves fundamental issues of both personal and national identity that are not conducive to any reasonable form of compromise [@abramowitz_great_2018; @mohler_jr_gathering_2020].
+Through the lens of these characteristics, one can begin to adequately grasp the challenges of polarization to the professional character of the nation's military.
 
-## When Words Become Weaponized - Theory
+The first way in which today's highly polarizing political climate is different than that of previous periods in American history is its scope.
+Fewer and fewer issue areas and policy matters are outside of the "spell" of polarization, owing to a phenomenon known as "conflict extension" [@@layman_party_2002].
+In previous eras of polarization, a single issue area served as the "battleground" for partisan conflict.
+Today, though, this is no longer true.
+Research has consistently shown that the polarizing politics present in the United States today has caused voters who identify as liberals to be more liberal across multiple issue categories, and that the same holds for conservatives [@abramowitz_great_2018].
+This includes the full range of social welfare, cultural, and racial issues, all of which are important in the minds of the electorate and partisans [@layman_party_2006].
+As a result, the prospect for cooperation on issues that were not politically polarized remained high, whereas today, the prospect for cooperation in general is lower as a result of the shrinking, non-partisan space [@layman_activists_2010].
 
-The purpose of this section is to explain how the concepts of professionalism, credibility, and political polarization can interface to create a potentially dangerous atmosphere for both military professionals and the public.
-It must be stressed from the outset that none of these concepts in isolation impose a normative prohibition on the act of a military officer taking up the pen to write an opinion piece.
-Indeed, Huntington's concept of military professionalism recognizes the unique skills of military officers, and Zaller's framework of elite discourse tacitly affirms that retired military officers, especially generals and admirals, most of whom possess more than 30 and in some cases 40 years of professional experience, possess relevant expertise that the public would find helpful, relevant, and convincing as it pertains to a host of potential policy debates.
-Even the existence of political polarization itself is insufficient to suggest that retired military officers should never give their opinion on specific matters.
+The second important characteristic of today's polarizing climate involves the strong, negative feelings that each side espouses towards those who hold differing opinions on a variety of issues.
+Known as "the rise of negative partisanship" [@abramowitz_great_2018], these strongly negative feelings toward those who hold different political opinions is not a phenomenon that has occurred among political elites only; instead, this trend has been shown to have grown among voters since the late 1970s [@abramowitz_great_2018].
+Therefore, in addition to there being a diminished opportunity for compromise owing to the proliferation of issues that are polarized, the feelings produced by today's polarized climate are so strong that the other side believes that other is deeply and morally wrong.
 
-One can reasonably infer that a professional opinion piece written by a military officer would then take on three very important qualities.
-First, stemming from the concept of military professionalism, the piece would address a relevant military topic.
-To be clear, multistar generals and admirals often are experienced in the realm of strategy, a word that is often misunderstood and misapplied in discussions of the topic [see for example @strachan_direction_2014].
-Without unpacking the concept of strategy further, the relevant point is that military elites, and especially those who achieved a high rank, have been privy to decisions and experiences occurring near the highest levels of government.
-Their body of experiences is not confined to mere tactics.
-They They are familiar with issues ranging from how wars should be best fought, to future security threats looming on the horizon, to fixing the institutional culture of a military service.
-Second, from a credibility and expertise perspective, a professional piece would be based on the officer's personal experiences and would be written from an individual's perspective so as not to infer that an argument is being made on behalf of the collective service or element within the military body.
-Third, an opinion made by a military officer would be free of any hint of partisan sentiment, whether advocacy or denouncement.
-If one of these three qualities - addressing a relevant military topic, from the perspective of an individual, and free of partisan expression - fails to be upheld, there is increasing likelihood that the officer's words can serve as a weapon, and, like other kinds of weapons, can do great harm.
+The third characteristic of today's polarizing environment is that polarizing issues do not involve matters of simple policy difference, but rather deep and fundamental beliefs about individual and national identities and purposes [@abramowitz_great_2018; @sides_identity_2018; @mohler_jr_gathering_2020].
+The election of Donald Trump and his years in office as President have raised the salience on a host of issues, all of which are essential to various parts of the electorate, including attitudes about race, immigration, and religion  [@sides_identity_2018].
+Moreover, the American people are aware of the polarized environment, and in a recent survey, indicated that they are concerned the trend is not improving [@schaeffer_far_2020].
 
-Figure 1, below, outlines the general theory and framework of the impact of political polarization and institutional credibility on military professionalism, and subsequently, on various military outcomes.
-The system of variables represented in Figure 1 constitute the "drive chain" that can lead to the words of retired military officers becoming militarized.
+In summary, taken together, these three characteristics of contemporary US political polarization are sobering.
+As a large body of research indicates, it is right to understand the political climate which American finds itself in today not merely as another polarized period, but rather in a tumultuous era that may well challenge the very definition of what America is and what it stands for [@mann_its_2012; @sides_identity_2018; @abramowitz_great_2018; @mohler_jr_gathering_2020].
+How the military may be impacted by the forces of increasing institutional credibility and today's political polarization is the subject of the next section.
 
-When military credibility is high, a demand signal is created by the American public such that future interactions between the military institution and the American public are encouraged and welcomed.
+# When Words Become Weaponized - A Theory of Credibility and Polarization
+
+Figure 1, below, visually describes a general theory of how institutional credibility and political polarization interact to impact military professionalism.
+From left to right, two significant societal changes (node 'SC') have occurred since the late 1970s, both of which have been described, and as this theory contends, interact, to impact military professionalism.
+The first societal change, that of institutional credibility, is indicated by the node 'IC', and the second, that of political polarization, is indicated by the node 'PP'.
+Each of these, in turn, impact military professionalism, indicated by the node 'MP'.
+Last, though it is outside of the scope of this paper, I contend that changes in military professionalism as a result of the forces of increasing political polarization and institutional credibility, will impact various military outcomes (node 'MO').
+These military outcomes vary, and could include things such as military effectiveness, to include the military's ability to fight and win battles [@talmadge_dictators_2015], as well as strategy formulation, to include the military's ability to work with civilians to produce sound strategic advice [@rapp_civil-military_2015].
+Indeed, the full degree to which and in what ways polarization will impact the military are not yet fully known.
+
+The theory grounds the two independent variables discussed here, credibility and polarization, as changes that have occurred within American society in the past fifty years.
+Because these changes stem from changes that have their originate within broader American society, my theory contends that the military as an institution and the military professionals that comprise it are not impervious to their impacts.
+In other words, my theory assumes that maintaining professional conduct by the military institution requires a deliberate effort and is hard work, despite the fact that behaving in this manner is a long and storied norm.
+Professional conduct is not a guaranteed nor automatic outcome.
+
+When institutional credibility is high, a demand signal is created by the American public such that future interactions between the military institution and the American public are encouraged and welcomed.
 From an institutional perspective, the military might respond to the public's demand signal for increased military interaction by increasingly participating in public ceremonies, parades, and other showcases of military prowess (at sporting venues, for instance).
 From an individual perspective, military elites, especially retired military officers, would be psychologically primed to increasingly dialogue with the American public when military credibility is high.
-By extension, other outlets such as the Press would also be primed to showcase military elites when military credibility runs high, because these elites represent an institution that commands a high degree of public respect.
+By extension, other outlets such as the press would also be primed to showcase military elites when military credibility runs high, because these elites represent an institution that commands a high degree of public respect.
 
-When military credibility is low, I theorize the opposite will occur.
+When military credibility is low, the theory contends the opposite will occur.
 Since demand for interactions between the American public and the military is muted when institutional credibility for the military is low, the institution will not engage with the public as frequently nor as intimately because the strong demand signal that existed when institutional credibility was high no longer exists.
-At the individual perspective, military elites will also not engage, nor will the Press ask these elites to, with the American public.
-In short, institutional credibility and respect for the military institution impacts the likelihood of the military institution and military elites interacting with the American public.
+At the individual level, military elites will also not engage as often with the public, nor will the press ask these elites to.
+In sum, institutional credibility and respect for the military institution impacts the likelihood of the military institution and military elites interacting with the American public.
 Such interaction can take a variety of forms.
@@
 problem_dag <- dagify(
-    Y ~ X1,
-    X1 ~ X2 + X3,
-    X2~X4,
-    X3~X4,
+    MO ~ MP,
+    MP ~ PP + IC,
+    PP~SC,
+    IC~SC,
     coords = tribble(
       ~ name,      ~ x,    ~ y,
-      "Y", 1, .5,
-      "X1", .7, .5,
-      "X2", .3, 1,
-      "X3", .3,0,
-      "X4", 0, .5
+      "MO", 1, .5,
+      "MP", .7, .5,
+      "IC", .3, 1,
+      "PP", .3,0,
+      "SC", 0, .5
     ))
@@
 problem_dag %>% tidy_dagitty() %>% 
-  dag_label(labels=c("Y"="Military\n Outcomes", 
-                     "X1"="Military\n Professionalism", 
-                     "X2"="Political\n Polarization", 
-                     "X3"="Institutional\n Credibility", 
-                     "X4"="Major Changes\n In American Society")) %>% 
+  dag_label(labels=c("MO"="Military\n Outcomes", 
+                     "MP"="Military\n Professionalism", 
+                     "PP"="Political\n Polarization", 
+                     "IC"="Institutional\n Credibility", 
+                     "SC"="Major Changes\n In American Society")) %>% 
   ggplot(aes(x=x, y=y, xend=xend, yend=yend))+
@@
 ```
-The degree of political polarization evident also has the potential to impact the professionalism of the military as an institution.
-First, because of the phenomenon known as "conflict extension" [@layman_party_2002], which refers to the fact that the political parties have grown further apart from each other on an increasing number of issues, political polarization raises the number and scope of issues to disagree about.
-Research has consistently shown that the polarizing politics present in the United States today has caused voters who identify as liberals to be more liberal across multiple issue categories, and that the same holds for conservatives.
-The result is that in the past, the prospect of cooperation on issues that were not polarized remained high, whereas today, the prospect for cooperation in general is lower [@layman_activists_2010].
-As it concerns the military institution, "conflict extension" [@layman_party_2002] increases the likelihood that members of the military feel strongly about an increasing number of issues.
-In other words, I contend that today's polarized environment makes it more difficult for an institution that to be politically neutral when there are fewer issues to remain neutral over.
-At the individual level, the same is true: as the non-partisan space seems to shrink, individuals will be enticed to not remain neutral.
+The degree and character of political polarization in a society also has the potential to impact the professionalism of the military as an institution.
+To understand how, recall that the three salient characteristics of the political polarization prevalent in contemporary America include an increase in the number and scope of issues that have become polarized (the phenomenon referred to as "conflict extension" [@layman_party_2002]), an increase in the strength of negative feelings held by supporters of an issue toward the opposition (the rise of "negative partisanship" [@abramowitz_great_2018]), and an increase in the salience of core issues related to individual and national identity [@white_values_2002; @mann_its_2012; @sides_identity_2018].
+These three characteristics convey several implications for the military as an institution and its leaders.
 
-The second impact of political polarization on military professionalism deals with the fact that the issues which are becoming polarized are not merely matters of simple policy preferences, but substantive issues which are directly related to notions of a person's core identity, such as race and immigration [@sides_identity_2018], religion [@white_values_2002], and the future of the country [@mann_its_2012].
-These issues do not lend themselves to compromise.
-They are, for many people, using an economics phrase, "indivisible goods." This also explains the notion of "negative partisanship," which refers to the very strong negative feelings that exist on each side of the political spectrum for those on the other [@abramowitz_great_2018, 5].
-Citizens feel and believe that the stakes are raised to an extremely high level for a number of pivotal issue areas, and that there are enormous consequences if the specific side with which they ally on any given issue loses to the other side.
-Polarization, then, raises the salience of right and wrong in a moral sense, which increases the odds of expressing partisan sentiment.
+First, there is increasing difficulty for individuals, to include military professionals, to remain neutral as polarized politics subsumes more and more issues, many of which involve issues which individuals hold to dearly.
+In previous eras, when polarization was not as extreme as it is today, it was more possible to stay out of politics.
+Today, polarization seems to do the opposite: polarized politics seem to thrust issues and the unfolding political fight at the individual, even when the individual is not seeking confrontation.
+A second implication of polarization is that it raises the salience for individuals of right and wrong in a moral sense, which increases the odds of an individual expressing partisan sentiment.
+In other words, in saying or doing what one believes to be right, one is increasingly likely to also be taking a partisan position.
+The third and most important implication for the military professional as a result of contemporary political polarization is that it impacts and distorts the notions of two character traits that are held dearly by members of the military profession: loyalty and duty.
 
-The third and most important way in which political polarization impacts military professionalism, I contend, extends from the first two reasons outlined above.
-First, because there are less issue areas in general to remain politically neutral about, and second, because issues are increasingly viewed in terms of moral right and wrong, political polarization impacts and distorts the notions of two vital character traits: loyalty and duty.
 As any soldier knows, few traits are more sacred in the military than loyalty and duty.
-In a very real sense, the military professional and the institution at large see loyalty and duty as incentives.
-By this, I simply mean that military professionals pride themselves on living up to mantras and phrases which extol the virtues of "doing the right thing," "leading with integrity," and, as the West Point Cadet Prayer so elegantly states, choosing the "harder right over the easier wrong."
+These virtues are extolled in a host of codes and mantras that have played a central part of the American military since its inception.
+They are captured in phrases that military professionals say to themselves and to each other: "do the right thing," "lead with integrity," and, as the West Point Cadet Prayer so elegantly states, choose "the harder right over the easier wrong."
 
-Political polarization obfuscates these ideals, however, and places greater responsibility for interpreting the meaning and the associated action of these ideals on the individual, and less on the military institution.
-It is in this way, I contend, that political polarization raises the chance of the military professional acting in such a way that simultaneously expresses partisan sentiment, because in doing so, the military professional believes he or she is doing what is morally right.
-
-This is why in a politically polarized environment, two different military officers may adopt completely different opinions of a contentious issue, each believing that the other's opinion is wrong.
-Should the officer act on that opinion, the other might believe he or she is operating outside of established military frameworks, the constitution, or any number of structures.
+Political polarization, however, obfuscates these ideals, and places greater responsibility for interpreting the meaning and the associated action of these ideals on the individual and less on the corporate military institution.
+This stems from the fact that highly polarized political climate may involve specific issue areas that certain members of the military feel a stronger affinity towards than other members.
+Two different military officers may adopt completely different opinions of a contentious issue, each believing that the other's opinion is wrong.
+Should the officer act on that opinion, the other might believe he or she is operating outside of the proper military frameworks, or outside of what the right thing to do is according to the Constitution or other loyalty structures.
 Each officer, then, might accuse the other of failing to be loyal to the ideals of the military.
-As a result, when polarization is high, the likelihood of partisan expression is high because in doing so, members of the military institution may feel that doing so is the right and loyal thing to do.
 
@@
 Secretaries Esper and Pompeo), you have forsaken your oath of office, the US Constitution, and the values that West Point stands for.
+
+In summary, it can be said that high degrees of political polarization impact military professionalism in key ways.
+Polarization increases the chances of expressing partisan sentiment in the name of doing what is morally right.
+Polarization, when it is high, also increases the chances of speaking out on issue areas that are outside of the realm of traditional military expertise, especially as the degree of polarization and "conflict extension" [@layman_party_2002] spreads, which reduces the perceived non-partisan space available to military professionals.
+Finally, polarization muddies the concepts of loyalty and duty.
+Strong feelings about issue areas perceived as crucial and essential by some greatly increase the odds of charging others of failing to act loyally or to fulfill one's duty.
 
@@
 
-I now aim to see whether the discourse of retired military officers has changed over time, specifically since the 1970s, as polarization has worsened.
+I now aim to see whether the discourse of retired military officers has changed over time, specifically since the 1970s, as polarization has worsened and as the institutional credibility of the military has increased.
 To answer this question, I assemble and examine a data set that consists of opinion pieces authored by retired military officers in major US newspapers.
@@
 \singlespace
-# The Second Layer - Are Retired Military Officers Upholding Professionalism as a Norm?
+# The Second Layer - Failing to Uphold Professionalism?
 \doublespace
@@
 
-## Conclusion: Can the Military Remain Professional?
+# Conclusion: How Can the Military Remain Professional?
 
-The dynamics of the military institution engaging the American public in a polarized environment remains ripe for future scholarship.
-In the author's opinion, whether the American military can, and to what degree it will, be able to remain professional as an institution remains to be seen.
-As political polarization becomes increasingly salient, several dynamics are unfolding, all of which deserve attention in future scholarship.
+This paper set out to examine the ways in which political polarization and institutional credibility have impacted the professionalism of the military over the past fifty years.
+Examining opinion pieces authored by retired military officers in major US newspapers revealed that these forces, together, have generated changes in how the military interacts with and what it says to the American public.
+The data examined demonstrates that the vast majority of op-eds published by retired military officers in the past fifty years exhibited professional standards.
+The data also showed, however, that retired military officers are arguing about topics that fall outside of the traditional umbrella of national security, and that they are expressing clear partisan preferences at a greater rate than they previously did.
+Moreover, both of these trends have increased under the past four years of Donald Trump's Presidency.
 
-The first dynamic, that of maintaining credibility and trust with the American public, remains crucial for the military, but this dynamic will likely be tested in fierce ways should a polarized political climate endure.
-Even as high credibility with the American public can give the military support, it can also be abused, especially if military officers, to include retirees, use the high level of credibility they currently enjoy to take up and advocate for partisan positions, or worse, to jockey with civilian leaders for power [@brooks_shaping_2008-1].
-Unfortunately, as this paper illustrates, there is evidence that mild aspects of this are occurring, but the extent to which it is, and its resulting impacts, remains an area for future research.
+This paper also argued that it is the interaction of increasing institutional credibility and increasing political polarization that has the potential to harm the professional military ethic.
+Maintaining credibility and trust with the American public remains crucial for the military, but should a polarized political climate endure, which is suggested by the majority of polarization scholars, the military needs to also understand that credibility and trust can be abused if it is handled inappropriately.
+This is especially likely if military professionals, to include retirees, abuse the high level of credibility they currently enjoy to take up and advocate for partisan positions, or worse, to jockey with civilian leaders for power [@brooks_shaping_2008-1].
+Unfortunately, as this paper illustrates, there are hints of this occurring.
+The letter signed by concerned graduates of West Point, which was clearly aimed at Secretaries Pompeo and Esper, and discussed at greater length earlier in this paper, is clear evidence of this.
 
-Credibility is not something that should be shunned.
-The military as an institution, to include military retirees, holds relevant and military expertise that should be afforded to the American public at large.
-And yet, just as retired officers were skilled in the management of violence when they wore the uniform every day, retirement in a polarized era means that they must also become managers of American trust.
-The promising news is, as this paper makes clear, the great proportion of retirees who are speaking out in various settings know and internalize this.
-Nonetheless, this will become more difficult to do in a polarized environment.
+Gaining a better understanding of the impacts of political polarization on the military profession remains a critical area for future research.
+Other scholars such as Michael Robinson have explored the specific ways that polarization is impacting the interactions between the military and the public [@robinson_who_2019].
+Though this paper focused on the ways that polarization may impact professional military behavior, it recognized but stopped short of discussing polarization's subsequent impact on various "military outcomes," which may arguably include more important outcomes, such as achieving victory or defeat in battle, or the ability to formulate sound strategic advice.
+These outcomes and more, and the role of polarization in influencing them, deserve to be explored.
 
-The second, and perhaps more important dynamic, is understanding the impacts of the polarized environment on the military's ability to perform its primary military task, i.e.
-fighting and winning the nation's wars.
-Scholars, military officers, and practitioners must better understand what impacts, if any, are placed on military and civilian leaders as a result of political polarization on the development and implementation of war goals, aims, and strategies, and how military leaders and civilian policymakers can overcome these hurdles to generate effective policies [@rapp_civil-military_2015].
-While scholarly literature has looked at how the military faces pressure stemming from divided institutional loyalties [@betts_soldiers_1991], none have explored the unique ways in which polarization as a domestic political force shapes individual and institutional behavior of the military.
-
+Finally, if one believes that today's polarization is different in degree and kind than previous eras of polarization, which has been argued in this paper, one must also recognize that the military will continue to face challenges as a result of political polarization.
+The events following the death of George Floyd certainly have revealed this.
 \singlespace

### 2020-07-14 code/Submissionv1.Rmd
commit

@@
 
-Finally, if one believes that today's polarization is different in degree and kind than previous eras of polarization, which has been argued in this paper, one must also recognize that the military will continue to face challenges as a result of political polarization.
-The events following the death of George Floyd certainly have revealed this.
+Finally, if one believes that today's polarized climate in America is different in degree and kind than previous eras of polarization, which has been a supporting claim of this paper, one must recognize that the military will face continued institutional challenges stemming from polarization's underlying causes.
+The most portending aspect regarding today's polarizing climate is that fundamental values are at stake, values that speak directly to the heritage of the United States as one nation under God and to the military as the nation's guarantor of rights and freedoms.
+As a non-partisan institution, the actions of the military will undoubtedly remain under scrutiny and close watch by the American public.
+No matter what it does, some will likely accuse the military of acting outside the bounds of the soldierly virtues of loyalty and duty.
+It is during these tough moments that moral courage will remain perhaps the chief attribute of military leaders, now and in the years to come.
 \singlespace

### 2020-07-14 code/Submissionv1.Rmd
commit changes

--- 2020-07-13 code/Submissionv1.Rmd
@@
 
-## Military Credibility and Expertise
+## _Military Credibility and Expertise_
 
@@
 
-##Political Polarization Today - Unique in America's History
+## _Political Polarization Today - Unique in America's History_
 
@@
 
-```{r include=TRUE, echo=FALSE}
+```{r include=TRUE, echo=FALSE, fig.height=5, fig.width=7}
 
@@
 
-```{r fig2, fig.height=5, fig.width=7, fig.align="center", include=TRUE, echo=FALSE}
+```{r fig2, fig.height=4, fig.width=5, fig.align="center", include=TRUE, echo=FALSE}
 datatograph <- data.frame(
@@
 
-This paper set out to examine the ways in which political polarization and institutional credibility have impacted the professionalism of the military over the past fifty years.
-Examining opinion pieces authored by retired military officers in major US newspapers revealed that these forces, together, have generated changes in how the military interacts with and what it says to the American public.
-The data examined demonstrates that the vast majority of op-eds published by retired military officers in the past fifty years exhibited professional standards.
-The data also showed, however, that retired military officers are arguing about topics that fall outside of the traditional umbrella of national security, and that they are expressing clear partisan preferences at a greater rate than they previously did.
+The goal of this paper was to explore and examine the ways in which political polarization and institutional credibility have impacted the professionalism of the military over the past fifty years.
+Examining the published opinion pieces authored by retired military officers in major US newspapers revealed that together, institutional credibility and political polarization have generated changes in the professionalism of some retired military officers, and more broadly, in how they as elites are communicating with the American public.
+Though the data examined demonstrates that the vast majority of op-eds published by retired military officers in the past fifty years exhibited professional standards, the data also showed two worrisome trends: that retired military officers are arguing about topics that fall outside of the traditional umbrella of national security, and that they are expressing clear partisan preferences at a greater rate than they previously did.
 Moreover, both of these trends have increased under the past four years of Donald Trump's Presidency.
 
-This paper also argued that it is the interaction of increasing institutional credibility and increasing political polarization that has the potential to harm the professional military ethic.
-Maintaining credibility and trust with the American public remains crucial for the military, but should a polarized political climate endure, which is suggested by the majority of polarization scholars, the military needs to also understand that credibility and trust can be abused if it is handled inappropriately.
+This paper also developed a theory for how the interaction of increasing institutional credibility and increasing political polarization has potential to harm the professional military ethic.
+Maintaining credibility and trust with the American public remains crucial for the military, but should a polarized political climate endure, which is suggested by the majority of American politics scholars who study polarization, the military needs to also understand that credibility and trust can be abused if it is handled inappropriately - a legitimate risk in a polarized environment.
 This is especially likely if military professionals, to include retirees, abuse the high level of credibility they currently enjoy to take up and advocate for partisan positions, or worse, to jockey with civilian leaders for power [@brooks_shaping_2008-1].
@@
 Other scholars such as Michael Robinson have explored the specific ways that polarization is impacting the interactions between the military and the public [@robinson_who_2019].
-Though this paper focused on the ways that polarization may impact professional military behavior, it recognized but stopped short of discussing polarization's subsequent impact on various "military outcomes," which may arguably include more important outcomes, such as achieving victory or defeat in battle, or the ability to formulate sound strategic advice.
-These outcomes and more, and the role of polarization in influencing them, deserve to be explored.
+Though this paper focused on the ways that polarization may impact professional military behavior, it recognized but stopped short of discussing polarization's subsequent impact on various "military outcomes," which may arguably include more decisive events, such as achieving victory or defeat in battle, or the ability to formulate sound strategic advice.
+These outcomes and more, and the role of polarization in influencing them, deserve to be explored further.
 
-Finally, if one believes that today's polarization is different in degree and kind than previous eras of polarization, which has been argued in this paper, one must also recognize that the military will continue to face challenges as a result of political polarization.
-The events following the death of George Floyd certainly have revealed this.
+Finally, if one believes that today's polarized climate in America is different in degree and kind than previous eras of polarization, which has been a supporting claim of this paper, one must recognize that the military will face continued institutional challenges stemming from polarization's underlying causes, and so will individual military leaders.
+The most portending aspect regarding today's polarizing climate is that fundamental values are at stake, values that speak directly to the heritage of the United States as one nation under God and to the military as the nation's guarantor of individual rights and freedoms.
+As a non-partisan institution that should theoretically be free from partisan entanglement, the actions of the military will remain under close scrutiny by the American public.
+In this type of environment, in some sense, the military is doomed: regardless of how the institution or its leaders act or respond to pressing polarized debates, some will accuse it of failing to do its duty and to exercise genuine loyalty.
+It is during these tough moments that another soldierly virtue - moral courage - will perhaps stand out as the chief attribute of military leaders, now and in the years to come.
 \singlespace

### 2020-07-15 code/Submissionv1.Rmd
commit changes

@@
 library("dagitty")
+library("gridExtra")
 ```
+\newpage
 
@@
 ```
+
 The degree and character of political polarization in a society also has the potential to impact the professionalism of the military as an institution.
@@
 
-**$H_1$: The majority of op-eds published by retired military officers over the time period examine will be apolitical and non-partisan in tone.**
+_$H_1$: The majority of op-eds published by retired military officers over the time period examine will be apolitical and non-partisan in tone._
 
-**$H_2$: Because overall newspaper consumption has dropped in the past 40 years, there will be fewer opinion pieces over time authored by military officers.**
+_$H_2$: Because overall newspaper consumption has dropped in the past 40 years, there will be fewer opinion pieces over time authored by military officers._
 
-**$H_3$: Given that polarization has become increasingly pronounced in the United States in recent years, there will be an increase in the number of op-eds which reveal political and partisan sentiments (indication of a drop in professionalism).**
+_$H_3$: Given that polarization has become increasingly pronounced in the United States in recent years, there will be an increase in the number of op-eds which reveal political and partisan sentiments (indication of a drop in professionalism)._
 
@@
 # regarding social policy or general parties or parties by name or administrations occurs
-rem_obs %>% group_by(PubYr) %>% 
+adj2dataset <- adjdataset %>% group_by(PubYr) %>% 
   mutate(
-    case_count=sum(MilPSoc ==1 | Dom_P==1 | Fig_Nme==1) 
+    case_count=sum(MilPSoc ==1 | Dom_P==1 | Fig_Nme==1), 
+    yr_total_count=sum(War_Ops==1 | Trp_Clt==1| Str_Adv==1 | MilPSoc ==1 | Dom_P==1 | Fig_Nme==1),
+    prop_yr=(case_count/yr_total_count)
   ) %>% view()
@@
 The first occurs in the time frame between roughly 1998-1999, the second between 2004-2007, and the third between 2016-2019.
-The clusters can also be seen in time in Figure 1 on the following page.
+The clusters can also be seen in time in Figure 2 on the following page.
 
@@
 
-```{r fig2, fig.height=4, fig.width=5, fig.align="center", include=TRUE, echo=FALSE}
+```{r fig2, fig.height=4, fig.width=6, fig.align="center", include=TRUE, echo=FALSE}
 datatograph <- data.frame(
@@
     title = "Figure 2: # of Op-Eds Written By Military Retirees",
-    subtitle = "On Topics of Social Policy, Domestic Policy, or Individual Figures"
+    subtitle = "On Topics of Social Policy, Domestic Policy, or Political Administrations/Figures"
   )
+```
+
+```{r practice, include = TRUE, echo=FALSE}
+
+plot1<- ggplot(data=adj2dataset, mapping=aes(x=PubYr, y=case_count)) + geom_point() + geom_smooth() + labs(x="Year", y="Count", color=NULL, title="Figure 2: Yearly # of Op-Eds Published by Retired Military Officers", subtitle="On Topics of Social Policy, Domestic Policy, or Political Administrations/Figures")
+plot2 <- ggplot(data=adj2dataset, mapping=aes(x=PubYr, y=prop_yr)) + geom_point() + geom_smooth() + labs(x="Year", y="Count", color=NULL, title="Figure 2: Proportion of Total Op-Eds Published by Retired Military Officers ", subtitle="On Topics of Social Policy, Domestic Policy, or Political Administrations/Figures")
+grid.arrange(plot1, plot2, ncol=2)
 ```
@@
 
-Using these characteristics as guidelines, I therefore develop two new variables. The first is **Result**, which can take on the following values:\singlespace
+Using these characteristics as guidelines, I therefore develop two new variables. The first is **Result**, which can take on the following values:
+\singlespace
 * **Permissible** -if the reader cannot clearly draw conclusions regarding the author's political stances or partisan preferences; or if the topic addressed in the publication in fact does in fact fall underneath the area of clear military expertise.
@@
     color = NULL,
-    title = "Figure 4: Total Number of Op-Eds Written By Military Retirees",
-    subtitle = "WSJ, NY Times, Washington Post, USA Today, and LA Times"
+    title = "Figure 4: # of Op-Eds Written By Military Retirees",
+    subtitle = "Published in WSJ, NY Times, Wa Post, USA Today, and LA Times"
   )

### 2020-07-21 code/Submissionv1.Rmd
commit changes

@@
 
-# First Layer: Topical Analysis
+# First Layer Analysis: Topic Addressed
 
@@
 
-```{r table, fig.align="center", include=TRUE, echo=FALSE}
+```{r table2, fig.align="center", include=TRUE, echo=FALSE}
 countbytype %>%
@@
 
-Table 1 reveals that an overwhelming majority (289 of 334, or 86.5%) of the total number of opinion pieces written since 1979 in the major newspapers examined can be classified according to one of the following three categories: **War_Ops**, **Trp_Clt**, and **Str_Adv**.
+Table 2 reveals that an overwhelming majority (289 of 334, or 86.5%) of the total number of opinion pieces written since 1979 in the major newspapers examined can be classified according to one of the following three categories: **War_Ops**, **Trp_Clt**, and **Str_Adv**.
 It is reasonable to conclude that these op-eds written are appropriate from the perspective of civil-military relations in that they address topics which American citizens can reasonably expect retired military officers to bring their expertise to bear upon.
 The remaining 45 opinion pieces (13.5% of the total) fall into one of the other three categories: **MilPSoc**, **Dom_P**, and **Fig_Nme** and require further examination.
-They are not necessarily inappropriate, but they likely address topics that directly address or are closely related to issues that could be considered as partisan or at least beyond appropriate military expertise.
+They are not necessarily inappropriate, but they likely address topics that may be beyond reasonable military expertise or which reveal partisan preferences.
 This second layer of analysis will help expose whether there is any indication that professionalism as a norm among retired military officers appears to be weakening.
@@
 \doublespace
+
 Of the remaining observations, two are removed upon immediate inspection.
@@
 
-```{r table2, fig.align = "center", include=TRUE, echo=FALSE}
+```{r table3, fig.align = "center", include=TRUE, echo=FALSE}
   rem_obs %>%
@@
 # Create a Variable Called case-count that counts the number of times by publishing year an article published 
-# regarding social policy or general parties or parties by name or administrations occurs
+# regarding social policy, domestic policy, or political figures/administrations occurs
 adj2dataset <- adjdataset %>% group_by(PubYr) %>% 
@@
 
-A couple of interesting observations are worth noting about the data in Table 2.
+A couple of interesting observations are worth noting about the data in Table 3.
 First, one can quickly observe that there appear to be three distinct clusters of observations with respect to publishing year.
 The first occurs in the time frame between roughly 1998-1999, the second between 2004-2007, and the third between 2016-2019.
-The clusters can also be seen in time in Figure 2 on the following page.
 
@@
 Retired Air Force General Michael Hayden, Navy Admirals Mike Mullen and William McRaven, and Army General Stanley McChrystal are among the retired officers penning pieces during this time period.
-
-```{r fig2, fig.height=4, fig.width=6, fig.align="center", include=TRUE, echo=FALSE}
-datatograph <- data.frame(
-  year=c("1994","1996","1998","1999","2000","2003", "2004", "2006", "2007", "2010", "2013","2014", "2016", "2017", "2018", "2019", "2020"), count=c(1, 1,2, 3, 1, 1, 3, 3, 3, 1, 1, 1, 6, 5, 4, 6, 1)
-)
-ggplot(datatograph, aes(x =year, y =count)) + geom_bar(stat = "identity", width=.2) +
-  labs(
-    x = "Year",
-    y = "Count",
-    color = NULL,
-    title = "Figure 2: # of Op-Eds Written By Military Retirees",
-    subtitle = "On Topics of Social Policy, Domestic Policy, or Political Administrations/Figures"
-  )
-```
-
-```{r practice, include = TRUE, echo=FALSE}
-
-plot1<- ggplot(data=adj2dataset, mapping=aes(x=PubYr, y=case_count)) + geom_point() + geom_smooth() + labs(x="Year", y="Count", color=NULL, title="Figure 2: Yearly # of Op-Eds Published by Retired Military Officers", subtitle="On Topics of Social Policy, Domestic Policy, or Political Administrations/Figures")
-plot2 <- ggplot(data=adj2dataset, mapping=aes(x=PubYr, y=prop_yr)) + geom_point() + geom_smooth() + labs(x="Year", y="Count", color=NULL, title="Figure 2: Proportion of Total Op-Eds Published by Retired Military Officers ", subtitle="On Topics of Social Policy, Domestic Policy, or Political Administrations/Figures")
-grid.arrange(plot1, plot2, ncol=2)
-```
 
@@
 
-Using these characteristics as guidelines, I therefore develop two new variables. The first is **Result**, which can take on the following values:
-\singlespace
-* **Permissible** -if the reader cannot clearly draw conclusions regarding the author's political stances or partisan preferences; or if the topic addressed in the publication in fact does in fact fall underneath the area of clear military expertise.
-* **Questionable** -if the reader can draw some general, but not final, conclusions regarding the author's political stances or partisan preferences; or if the topic addressed in the publication is questionably an area of reasonable military expertise.
-* **Inappropriate** -if the reader can draw clear conclusions regarding the author's partisan preferences or political leanings; or if the author addresses a topic that is not an area considered to be directly relevant to military expertise, or a matter that the military should judge.
+Using these characteristics as guidelines, I therefore develop two new variables.
+The first is **Result**, which can take on one of three values: the following values: **Permissible,** if the reader cannot clearly draw conclusions regarding the author's political leaning or partisan preferences; or if the topic addressed in the publication, after a second reading, clearly falls within military expertise; **Questionable,** if the reader can draw some but not complete conclusions regarding the author's political leaning or partisan preferences, or if the topic addressed in the publication is questionably an area of reasonable military expertise; and **Inappropriate,** if the reader can draw clear conclusions regarding the author's partisan preferences or political leanings, or if the author addresses a topic that clearly falls outside of traditional military expertise.
+The second new variable I develop is **Justification.** In this variable, a brief explanation is placed which addresses why an observation is assigned a particular value of **Result.**
 
-\doublespace
-The second new variable I develop is **Justification** in which I briefly comment on why I assigned the value of **Result** for the observation.
-Here I simply comment on whether it is a concern of partisan preferences being revealed and/or topic addressed that are responsible for the value of **Result**.
 The results of this second level of analysis are presented in Table 3.
@@
 
-```{r table3, include=TRUE, echo=FALSE}
+```{r table4, include=TRUE, echo=FALSE}
 secondlayer_obs %>%
-  kable(format="latex", caption="Concern for Op-Ed Pieces Authored By Retired Military Officers in US Newspapers, 1979-2020", longtable=T) %>%
+  kable(format="latex", caption="Op-Ed Pieces By Retired Military Officers which failed to Uphold Professional Standards, 1979-2020", longtable=T) %>%
   column_spec(1, width="9.5em") %>% 
@@
 
-Figure Two displays the number of opinion pieces authored by retired military officers which failed to uphold the strict standards of professionalism, as espoused by Retired Generals MacArthur and Dempsey, within each applicable year.
+Figure 2 displays the number of opinion pieces authored by retired military officers which failed to uphold the strict standards of professionalism, as espoused by Retired Generals MacArthur and Dempsey, within each applicable year.
 The results reveal several conclusions.
@@
 
-```{r fig3, fig.height=4, fig.width=5, fig.align="center", include=TRUE, echo=FALSE}
+```{r fig2, fig.height=4, fig.width=5, fig.align="center", include=TRUE, echo=FALSE}
 datatograph <- data.frame(
@@
     color = NULL,
-    title = "Figure 3: # of Op-Eds Written By Military Retirees",
+    title = "Figure 2: # of Op-Eds Written By Military Retirees",
     subtitle = "Failed to Uphold Professional Standards"
@@
 
-The third hypothesis, that the number of published opinion pieces which have strained the concept of professionalism has increased, even as political polarization has worsened in the past several decades, is also confirmed, as evidenced by both by Table 3 and Figure 2.
-It is interesting to note that according to Figure 2, the first publishing year in which an opinion piece penned by a retired officer failed to uphold professional standards occurred in 1998, nearly twenty years into (or almost half) of the entire sample.
-Why is this the case?
+The third hypothesis, that the number of published opinion pieces which have strained the concept of professionalism has increased, even as political polarization has worsened in the past several decades, is also confirmed, as evidenced by both by Table 3 and Figure 3.
+It is interesting to note that according to Figure 3, the first publishing year in which an opinion piece penned by a retired officer failed to uphold professional standards occurred in 1998, nearly twenty years into (or almost half) of the entire sample.
+But why is this the case?
 Retired military officers in the years preceding 1998 had plenty of reasons to go to the press and air disagreements with political leaders, or to address important topics that may have pushed the limits of professionalism as has been defined in this paper.
@@
 
-```{r fig4, fig.height=4, fig.width=5, fig.align="center", include=TRUE, echo=FALSE}
+```{r fig3, fig.height=4, fig.width=5, fig.align="center", include=TRUE, echo=FALSE}
 datatograph <- data.frame(
@@
     color = NULL,
-    title = "Figure 4: # of Op-Eds Written By Military Retirees",
+    title = "Figure 3: # of Op-Eds Written By Military Retirees",
     subtitle = "Published in WSJ, NY Times, Wa Post, USA Today, and LA Times"
@@
 The goal of this paper was to explore and examine the ways in which political polarization and institutional credibility have impacted the professionalism of the military over the past fifty years.
-Examining the published opinion pieces authored by retired military officers in major US newspapers revealed that together, institutional credibility and political polarization have generated changes in the professionalism of some retired military officers, and more broadly, in how they as elites are communicating with the American public.
+Examining the published opinion pieces authored by retired military officers in major US newspapers revealed that together, institutional credibility and political polarization have generated changes in the professionalism of some retired military officers, and more broadly, in how military elites communicate with the American public.
 Though the data examined demonstrates that the vast majority of op-eds published by retired military officers in the past fifty years exhibited professional standards, the data also showed two worrisome trends: that retired military officers are arguing about topics that fall outside of the traditional umbrella of national security, and that they are expressing clear partisan preferences at a greater rate than they previously did.
 Moreover, both of these trends have increased under the past four years of Donald Trump's Presidency.
+These trends are captured in Figure 4, below, which shows the number (left) and proportion (right) of opinion pieces authored by retired military officers over the past 40 years that addressed issues of social policy, domestic policy, or specific political figures or administrations.
+The graphs tell an interesting story: the number and proportion were both basically 0 40 years ago, then increased slightly until about 2003, at which point, both decreased until roughly 2011.
+After that, both the number and proportion have risen sharply.
+Nearly 40% of all opinion pieces in 2019 published in major US newspapers by retired military officers were written about topics that fell outside of traditional national security topics.
+
+```{r build plots, include = FALSE, echo=FALSE}
+plot1<- ggplot(data=adj2dataset, mapping=aes(x=PubYr, y=case_count)) + geom_point() + geom_smooth() + labs(x="Year", y="Number", color=NULL)
+plot2 <- ggplot(data=adj2dataset, mapping=aes(x=PubYr, y=prop_yr)) + geom_point() + geom_smooth() + labs(x="Year", y="Proportion", color=NULL)
+```
+
+```{r figure 4, include=TRUE, warning= FALSE, message = FALSE, echo=FALSE}
+grid.arrange(plot1, plot2, ncol=2, top="Figure 4: #/Proportion of Op-Eds Published by Retired Military Officers, 1979-2020", bottom="On Topics of Social Policy, Domestic Policy, and Political Administrations/Figures")
+```
 

### 2020-08-12 code/Submissionv2.Rmd
commit

--- 2020-07-21 code/Submissionv1.Rmd
@@
 library("gridExtra")
+library("broom")
+library("dplyr")
 ```
@@
 # Read in the Data
-dataset <- read_csv(here("data", "MilEditorials24June.csv"))
+dataset <- read_csv(here("data", "MilEditorials10Aug.csv"))
 dataset %>% print()
@@
 LATimescount
+
+#adjust the data to account for polarization by creating a new dataset called polaradjdataset
+#create variables that capture ideological distance in each the House and Senate along
+#each of the two dimensions according to Poole and Rosenthal's data
+polaradjdataset <-adjdataset %>% mutate (
+  House_distance_dim1=abs(House_Dem_mean_nom1-House_Rep_mean_nom1),
+  Sen_distance_dim1=abs(Sen_Dem_mean_nom1-Sen_Rep_mean_nom1),
+  House_distance_dim2=abs(House_Dem_mean_nom2-House_Rep_mean_nom2),
+  Sen_distance_dim2=abs(Sen_Dem_mean_nom2-Sen_Rep_mean_nom2)
+) %>% view()
 
@@
 ```{r second level analysis, include = FALSE}
-
 # Present Results of Result and Justification
@@
   filter(Author != "UNK") %>% filter(Author != "Murtha, John") %>% 
-  select (Author, Title, Source, PubYr, Result, Justification) %>% 
+  select (Author, Title, Source, PubYr, Ques_Prof, Justification) %>% 
   view()
+
+#Now Prep the data that I need to analyze polarization levels.  I will try first by 
+# eliminating the NA values but I will do this by indexing the first 334 rows of the 
+# data, which corresponds to the number of observations that I have.  I do this by #creating a new data set called slimdata
+
+slimdata <- polaradjdataset[1:334,] %>% 
+  view()
+
+#next, conduct a logit regression where the DV is values of Ques_Prof along
+#several IVs
+
+logmodel <- glm(
+  Ques_Prof ~ Inst_Cred_Gallup + House_distance_dim1 + MilPSoc + Dom_P + Fig_Nme, 
+  family=binomial(link="logit"), 
+  data=slimdata
+  )
+
+tidy(logmodel)
+glance(logmodel)
+
+# now  make prediction data using the mean institutional credibility scores.  Vary #polarization and the three topics I am interested in.
+
+predictiondata <- tibble(House_polarization_level=seq(min(slimdata$House_distance_dim1), max(slimdata$House_distance_dim1), by=.01), Credibility = mean(slimdata$Inst_Cred_Gallup)) %>%
+crossing(Ques_Prof = c(0,1), MilPSoc = c(0,1), Dom_P = c(0,1), Fig_Nme = c(0,1))
+
+predictiondata
+
+# actual prediction data now with code.  Create variables that tie the model together. #Create Variables .prob, conf.low, conf.high, and pred_final_rec
+
+augment(logmodel, newdata = predictiondata)
+
+predictions <- augment(logmodel, newdata=predictiondata) %>% 
+  mutate(
+    .prob=plogis(.fitted),
+    conf.low = plogis(.fitted -(qnorm(.975) * .se.fit)),
+    conf.high = plogis(.fitted + (qnorm(.975) * .se.fit)),
+    pred_final_rec = case_when(
+      .prob >= .5 ~ "Likely Violation of Professionalism",
+      .prob < .5 ~ "Violation of Professionalism not likely"
+    )
+  ) 
+predictions %>% arrange(.prob, .fitted, Report, Night, desc(.fitted)) %>% 
+  view()
+
 ```
@@
 These trends are captured in Figure 4, below, which shows the number (left) and proportion (right) of opinion pieces authored by retired military officers over the past 40 years that addressed issues of social policy, domestic policy, or specific political figures or administrations.
-The graphs tell an interesting story: the number and proportion were both basically 0 40 years ago, then increased slightly until about 2003, at which point, both decreased until roughly 2011.
+The graphs tell an interesting story: the number and proportion were both basically zero 40 years ago, then increased slightly until about 2003, at which point, both decreased until roughly 2011.
 After that, both the number and proportion have risen sharply.

### 2020-08-13 code/Submissionv2.Rmd
commit

@@
   view()
+```
 
-#Now Prep the data that I need to analyze polarization levels.  I will try first by 
+```{r polarization logit analysis, include=FALSE}
+
+#Now I Prep the data that I need to analyze polarization levels.  I will try first by 
 # eliminating the NA values but I will do this by indexing the first 334 rows of the 
-# data, which corresponds to the number of observations that I have.  I do this by #creating a new data set called slimdata
+# data, which corresponds to the number of observations that I have.  I do this by #creating a new data set called slimdata.
+
+# Note: this is sloppy.  There has to be a better way to get rid of NA values.
 
@@
 
-predictiondata <- tibble(House_polarization_level=seq(min(slimdata$House_distance_dim1), max(slimdata$House_distance_dim1), by=.01), Credibility = mean(slimdata$Inst_Cred_Gallup)) %>%
+predictiondata <- tibble(House_distance_dim1=seq(min(slimdata$House_distance_dim1), max(slimdata$House_distance_dim1), by=.01), Inst_Cred_Gallup = mean(slimdata$Inst_Cred_Gallup)) %>%
 crossing(Ques_Prof = c(0,1), MilPSoc = c(0,1), Dom_P = c(0,1), Fig_Nme = c(0,1))
@@
   ) 
-predictions %>% arrange(.prob, .fitted, Report, Night, desc(.fitted)) %>% 
+predictions %>% arrange(.prob, .fitted, MilPSoc, Dom_P, Fig_Nme, desc(.fitted)) %>% 
   view()
 
+# stopping point here.  now let's try some graphs
+# notice that the predictions are not in (0, 1)
+ggplot(predictions) +
+  aes(x = .fitted) +
+  geom_histogram()
+
+# if we wanted predictions on the probability scale,
+#   we predict on the log scale,
+#   then apply inverse-logit (logistic CDF)
+# use plogis(): cumulative probability of the logistic distribution
+
+# demonstration: scatter log odds (x) and predicted probability (y)
+# with 95% CIs
+# (CIs don't look smooth in this example because .se.fit depends on X data)
+predictions %>%
+  ggplot() +
+  aes(x = .fitted, y = .prob) +
+  geom_pointrange(aes(ymin = conf.low, ymax = conf.high)) +
+  geom_smooth() 
+```
 ```

### 2020-08-25 code/Submissionv2.Rmd
commit

@@
 
-#next, conduct a logit regression where the DV is values of Ques_Prof along
-#several IVs
+#next, conduct logit regressions where the DV is values of Ques_Prof along
+#two IVs, political polarization and institutional credibility.
 
-logmodel <- glm(
-  Ques_Prof ~ Inst_Cred_Gallup + House_distance_dim1 + MilPSoc + Dom_P + Fig_Nme, 
+#model 1
+
+logmodel1 <- glm(
+  Ques_Prof ~ Inst_Cred_Gallup + House_distance_dim1, 
   family=binomial(link="logit"), 
@@
 
-tidy(logmodel)
-glance(logmodel)
+tidy(logmodel1)
+glance(logmodel1)
 
-# now  make prediction data using the mean institutional credibility scores.  Vary #polarization and the three topics I am interested in.
+# now  make prediction data using the mean institutional credibility scores.  Vary #polarization and hold institutional credibility constant.
 
 predictiondata <- tibble(House_distance_dim1=seq(min(slimdata$House_distance_dim1), max(slimdata$House_distance_dim1), by=.01), Inst_Cred_Gallup = mean(slimdata$Inst_Cred_Gallup)) %>%
-crossing(Ques_Prof = c(0,1), MilPSoc = c(0,1), Dom_P = c(0,1), Fig_Nme = c(0,1))
+crossing(Ques_Prof = c(0,1))
 
@@
 
-augment(logmodel, newdata = predictiondata)
+augment(logmodel1, newdata = predictiondata)
 
-predictions <- augment(logmodel, newdata=predictiondata) %>% 
+predictions <- augment(logmodel1, newdata=predictiondata) %>% 
   mutate(
@@
   ) 
-predictions %>% arrange(.prob, .fitted, MilPSoc, Dom_P, Fig_Nme, desc(.fitted)) %>% 
+predictions %>% arrange(.prob, .fitted, Ques_Prof, desc(.fitted)) %>% 
   view()
 
-# stopping point here.  now let's try some graphs
+# now let's try some graphs
 # notice that the predictions are not in (0, 1)
@@
   geom_smooth()
+
+# model2
+
+logmodel2 <- glm(
+  Ques_Prof ~ House_distance_dim1 + Inst_Cred_Gallup, 
+  family=binomial(link="logit"), 
+  data=slimdata
+  )
+
+tidy(logmodel2)
+glance(logmodel2)
+
+# now  make prediction data using the mean institutional credibility scores.  Vary #institutional credibilty and hold political polarization constant.
+
+predictiondata2 <- tibble(Inst_Cred_Gallup=seq(min(slimdata$Inst_Cred_Gallup), max(slimdata$Inst_Cred_Gallup), by=1), House_distance_dim1 = mean(slimdata$House_distance_dim1)) %>%
+crossing(Ques_Prof = c(0,1))
+
+predictiondata2
+
+# actual prediction data now with code.  Create variables that tie the model together. #Create Variables .prob, conf.low, conf.high, and pred_final_rec
+
+augment(logmodel2, newdata = predictiondata2)
+
+predictions2 <- augment(logmodel2, newdata=predictiondata2) %>% 
+  mutate(
+    .prob=plogis(.fitted),
+    conf.low = plogis(.fitted -(qnorm(.975) * .se.fit)),
+    conf.high = plogis(.fitted + (qnorm(.975) * .se.fit)),
+    pred_final_rec = case_when(
+      .prob >= .5 ~ "Likely Violation of Professionalism",
+      .prob < .5 ~ "Violation of Professionalism not likely"
+    )
+  ) 
+predictions2 %>% arrange(.prob, .fitted, Ques_Prof, desc(.fitted)) %>% 
+  view()
+
+# stopping point here.  now let's try some graphs
+# notice that the predictions are not in (0, 1)
+ggplot(predictions2) +
+  aes(x = .fitted) +
+  geom_histogram()
+
+# if we wanted predictions on the probability scale,
+#   we predict on the log scale,
+#   then apply inverse-logit (logistic CDF)
+# use plogis(): cumulative probability of the logistic distribution
+
+# demonstration: scatter log odds (x) and predicted probability (y)
+# with 95% CIs
+# (CIs don't look smooth in this example because .se.fit depends on X data)
+predictions2 %>%
+  ggplot() +
+  aes(x = .fitted, y = .prob) +
+  geom_pointrange(aes(ymin = conf.low, ymax = conf.high)) +
+  geom_smooth() 
 ```
+
 ```

### 2020-08-25 code/Submissionv3.Rmd
commit

--- 2020-08-25 code/Submissionv2.Rmd
@@
   _This paper explores variation in the professionalism of opinion pieces authored by retired US military officers and published in major US newspapers in the past 40 years.
-In doing so, this paper develops a theory that argues that two societal forces, political polarization and military credibility, interact in unique ways that ultimately strain the individual military officer's concepts of loyalty and duty in important ways.
-Assembling and analyzing more than 330 opinion-editorial pieces published by retired military officers in The Wall Street Journal, The Washington Post, The New York Times, The Los Angeles Times, and the USA Today since 1979, the paper argues there has been an increase in published opinion pieces which fail to uphold professional standards as defined by historical military norms.
-The paper concludes by arguing that today's political polarization reveals a profoundly changing society, and that the norm of a professional, apolitical military will remain under further strain in the current political environment._
+In doing so, this paper develops a theory which argues that increasing political polarization in the United States impacts the military as a profession by straining the individual military officer's concept of loyalty in important ways.
+Drawing on an original data set of more than 330 opinion-editorial pieces published by retired military officers in The Wall Street Journal, The Washington Post, The New York Times, The Los Angeles Times, and the USA Today since 1979, the paper argues there has been an increase in published opinion pieces which fail to uphold professional standards as defined by historical military norms and standards.
+The paper concludes by arguing that the concept of "politicization of the military" --- something that previous civil-military relations scholars have identified and described --- must be widened to account not only for efforts by politicians to use the military for political advantages, but also for the novel phenomenon of retired military officers using the concept of loyalty to achieve political ends.
+The author argues that such a development, if uncorrected, will cause significant damage to the military as a profession._
 geometry: margin = 1.15in              # margin widths
@@
 In late May and early June 2020, racial tension and concerns about police brutality gripped the American public following the death of George Floyd, an unarmed black man, by police in Minneapolis.
-In several cities and communities across the nation, mostly peaceful protests were followed by riots and looting.
+In many cities and communities across the nation, largely-peaceful protests were followed by periods of rioting and looting.
 On June 1, after giving remarks at the Rose Garden, President Donald Trump walked to St.
@@
 
-Additionally, several retired military officers were among the elites who weighed in to give their opinions not only on the Trump administration's handling of specific events near the White House, but also on the broader issues of race in America.
-But for those paying attention to what these retired officers said, it is also clear that there was substantial variation in the tone and tenor of each officer's remarks.
-While all of them expressed deep concern about the prospect of active-duty forces confronting peaceful protesters, and thus, violating protections held by the First Amendment of the US Constitution [@brooks_dismay_2020; @dempsey_former_2020], some officers went much further in their criticism, leveling direct insults of the President's fitness to lead [@goldberg_james_2020], his administration's broader policies on race [@allen_moment_2020], and his general leadership abilities [@mullen_i_2020].
+Furthermore, several retired military officers weighed in not only on the Trump administration's handling of specific events near the White House, but also on the broader issues of race in America.
+For those paying attention to what these retired officers said, it is clear that there was substantial variation in the tone and tenor of each of these officer's remarks.
+While all of them expressed deep concern about the prospect of active-duty forces confronting peaceful protesters, and thus, violating protections held by the First Amendment of the US Constitution [@brooks_dismay_2020; @dempsey_former_2020], some officers went much further in their criticisms, leveling direct insults of the President's fitness to lead [@goldberg_james_2020], his administration's broader policies on race in America [@allen_moment_2020], and his general leadership abilities [@mullen_i_2020].
+Moreover, in the weeks and months since George Floyd's death, retired and currently-serving military officers have further taken up the pen to argue for a variety of topical issues that are in one way or form related to President Donald Trump and his administration's response to a variety of current themes in domestic politics, including race in America [@askew_anti-racist_2020] and the US Postal Service and concerns about the upcoming Presidential Election [@nagl__2020].
 
-This paper examines the way in which retired military officers have attempted to influence public opinion in contemporary America through opinion-editorial publications.
-It is an interesting question worthy of examination because, as I argue, at least three important factors are interacting in unique ways to shape how and why retired military officers speak out on a variety of issues.
-These factors are professionalism, political polarization, and military credibility.
-I argue that the latter two factors, political polarization and institutional credibility, are interacting in unique ways that create tremendous strains on the military's professional nature, and that these factors will continue to do so.
+This paper examines the way in which retired military officers have attempted to influence public opinion in contemporary America by authoring opinion-editorial publications.
+It is a question worthy of examination for many reasons, including that at its root, at least three factors are involved, each of which may shape civil-military relations: military professionalism, political polarization, and institutional credibility of the military.
+The main argument of this paper is that a weakening of the concept of military professionalism can clearly be seen in the way retired military officers have made arguments in the public square over the past four decades, that the main variable that can explain this variation is increasing political polarization within American Society.
+Further, this paper argues that the military institution is particularly susceptible to increasing political polarization because of the way the concept of loyalty is used to make arguments that contain a clear partisan or political leaning.
+These developments reveal the extent to which increasing political polarization is synonymous with a fractured domestic public, and highlight the extraordinary challenge that the US military will llikely face to remain a  professional in an enduring polarized political climate.
 
 This paper proceeds in four parts.
-First, the paper briefly reviews the concept of military professionalism in order to orient the reader to the central role this concept has played in the ethos of the American military for the past several decades.
-Second, the paper builds a theory that demonstrates how the uniqueness of today's political polarization and institutional credibility are interacting to create an environment that threatens professional military behavior in ways that could be highly destructive.
+First, the paper briefly reviews the concept of military professionalism in order to orient the reader to the central role this concept has played in the ethos of the modern American military.
+Second, the paper builds a theory that demonstrates how certain features of today's polarized political landscape challenge the concept of loyalty for the individual military professional.
 Third, the paper explores the data, which consists of more than 330 opinion-editorial pieces published in major US newspapers by retired military officers since the late 1970s.
-The analysis ultimately shows that although the vast majority of op-eds published by retired military officers maintain professional standards, there is clear evidence that retired military officers are using the pen to both argue about topics that fall outside of the realm of traditional national security topics, and to express clear partisan preferences when making their arguments.
-Fourth, the paper concludes by briefly exploring the implications of these results on the military as an institution, and specifically, on its ability to remain professional in a domestic setting marked by high political polarization in which the military commands wide public support.
+This analysis ultimately shows that although the vast majority of op-eds published by retired military officers maintain professional standards, there is clear evidence that retired military officers are using the pen to both argue about topics that fall outside of the realm of traditional national security topics, and to express clear partisan preferences when making their arguments.
+The analysis also shows that this trend has worsened as the American political climate has become increasingly polarized.
+Fourth, the paper concludes by briefly exploring the implications of these results on the military as an institution, and specifically, on its ability to remain professional in a highly polarized setting.
 
@@
 \doublespace
-The ideal of the US military acting as an objective, apolitical and non-partisan institution is storied and long.
-One of the strongest origins of the norm stems from the work of the political scientist, Samuel Huntington, who, in his 1957 work, _The Soldier and the State: The Theory and Politics of Civil-Military Relations_, asserted that two distinct forces would always be in tension with one another in a liberal democracy such as the United States.
+The ideal of the US military acting as an objective, apolitical and non-partisan institution is long and storied.
+As a concept, the idea of a professional military stems from the post-WW2 era, when the US military was locked in the Cold War struggle agains the Soviet Union.
+Perhaps the strongest and certainly one of the original intellectual founders of a professional military is the political scientist, Samuel Huntington, who, in his 1957 work, _The Soldier and the State: The Theory and Politics of Civil-Military Relations_, asserted that two distinct forces would always be in tension with one another in a liberal democracy such as the United States.
 The first was a "functional imperative," which Huntington said drove the requirement that a military be strong enough and capable of defeating the military threats it faced, and the second was a "societal imperative," which Huntington said generated requirements for the military in that the military could never lose so many aspects of it such that it no longer represented the liberal democracy that it defended [@huntington_soldier_1957, 1-3].
 Huntington's point was that America needed a strong military that could win on the battlefield, but cautioned that such a military could not be so foreign in character so as to alienate the character of American liberal society.
-Huntington argued that the key to resolving the tension generated by these two opposing forces was to adopt a scheme of "objective civilian control" in which political leaders established the ends and goals of policy, while military leaders developed and implemented the military plans to accomplish these goals [@huntington_soldier_1957, 80-85].
-To Huntington, largely separate operating spheres between the political leaders of a democracy and the military as an institution would ensure that the military remained "professional," and thus, untainted by partisan interests and concerns.
+Huntington further argued that the key to resolving the tension generated by these two opposing forces was to adopt a scheme of "objective civilian control" in which political leaders established the ends and goals of policy, while military leaders developed and implemented the military plans to accomplish these goals [@huntington_soldier_1957, 80-85].
+To Huntington, separate operating spheres between the political leaders of a democracy and the military as an institution would ensure that the military remained "professional," and thus, untainted by partisan interests and concerns.
 In this way, Huntington posited that the American military would best meet the task of defeating security threats while also remaining apolitical.
@@
 While Huntington's theory has enjoyed tremendous staying power (Huntington is still studied and taught in many American military institutions), he is not without critics [see, for example, @feaver_civil-military_1996; @cohen_supreme_2003; @feaver_armed_2005; and @brooks_paradoxes_2020].
-It would be too burdensome to unpack all of those here, but one of the earliest and richest critiques of Huntington, and perhaps one of the most relevant given the potential influences of political polarization and institutional credibility, came from S.E.
+It would be too burdensome to unpack all of those here, but one of the earliest and richest critiques of Huntington came from S.E.
 Finer.
-Authored in 1962, Finer's book, _The Man on Horseback: The Role of the Military in Politics_, argued that Huntington's concept of professionalism simply was not strong enough to keep the military out of politics [@finer_man_1962].
-Finer foresaw military officers adopting an ethic of professionalism such that they would actually be more likely, not less likely, to become involved in political issues [@finer_man_1962, 22-23].
+Finer's 1962 book, _The Man on Horseback: The Role of the Military in Politics_, argued that Huntington's concept of professionalism simply was not strong enough by itself to keep the military out of politics [@finer_man_1962].
+In fact, Finer foresaw military officers adopting an ethic of professionalism such that they would actually be more likely, not less likely, to become involved in political issues [@finer_man_1962, 22-23].
 
 Finer was especially concerned that the adoption of professionalism would lead to long-serving military professionals believing that they knew what was best for their country over and above that of temporary elected civilian leaders [@finer_man_1962, 22-23].
-Finer warned that "the moment the military draw this distinction between nation and the government in power, they begin to invent their own private notion of the national interest, and from this it is only a skip to the constrained substitution of this view for that of the civilian government…this purported care for the national interest as defined by the military is indeed one of their main reasons for intervening" [@finer_man_1962].
-In Finer's view, what was really needed to keep the military out of politics was not only the concept of professionalism, but rather a deep internalization of "the principle of civil supremacy" [@finer_man_1962].^[
+In fact, Finer warned that "the moment the military draw this distinction between nation and the government in power, they begin to invent their own private notion of the national interest, and from this it is only a skip to the constrained substitution of this view for that of the civilian government…this purported care for the national interest as defined by the military is indeed one of their main reasons for intervening" [@finer_man_1962].
+Therefore, in Finer's view, what was really needed to keep the military out of politics was not merely adoption of an ethic that relied on professionalism, but rather a deep internalization of "the principle of civil supremacy" [@finer_man_1962].^[
 Importantly, neither Huntington's concept of professionalism nor Finer's critique of the same are arguing for blind obedience by the military to all civilian orders and directions, a point that is, in the author's opinion, often missed in debates about military professionalism.
@@
 The operational military has also been central in upholding the military's non-partisan traditions, albeit with a far less-than-perfect record.
-Retired Army General Douglas MacArthur, a hero of World War Two in the Pacific and later, Commander of the occupation of Japan, is well known for his disagreement with President Truman regarding American strategic goals during the Korean War.
-And yet, the remarks MacArthur made to cadets at West Point in 1962, as an 82 year old retired officer, speak to the primacy of professional military conduct:
+Retired Army General Douglas MacArthur, a hero of World War Two in the Pacific and later, Commander of the military occupation of Japan, is well known for his disagreement with President Truman regarding American strategic goals during the Korean War.
+And yet, MacArthur's address to the cadets at West Point in 1962, as an 82 year old retired officer, speaks to the importance that MacArthur felt of the military professional focusing on appropriate military conduct:
 
@@
 
-Finally, current military doctrine and practices also reinforce the centrality of professional, apolitical, and non-partisan conduct.
-Prior to the 2016 Presidential Election, Marine Corps General Joseph Dunford, who at the time served as Chairman of the Joint Chiefs of Staff, said that the basis upon which the military is able to maintain the public's trust is it's ability to remain "apolitical and neutral" [@dunford_upholding_2016].
+Finally, current military doctrine and norms further reinforce the centrality of professional, apolitical, and non-partisan conduct by the military.
+For example, prior to the 2016 Presidential Election, Marine Corps General Joseph Dunford, who at the time served as Chairman of the Joint Chiefs of Staff, said that the basis upon which the military is able to maintain the public's trust is it's ability to remain "apolitical and neutral" [@dunford_upholding_2016].
 At nearly the same time, his predecessor, Retired Army General Martin Dempsey, published an opinion-editorial piece in _The Washington Post_ which criticized retired Army Major General Mike Flynn and Marine Corps General John Allen for participating in the Republican and Democratic conventions.
-Dempsey wrote, "As generals, they have an obligation to uphold our apolitical traditions.
+Dempsey wrote, "As generals, they [Flynn and Allen] have an obligation to uphold our apolitical traditions.
 They have just made the task of their successors – who continue to serve in uniform and are accountable for our security - more complicated.
@@
 Although a professional and non-partisan military is the theoretical goal of the institution, the practical record and results remain far from perfect.
-From MacArthur's clash with President Truman in the early 1950's, to the so-called "Revolt of the Generals" in 2005-2006, military leaders, even those who are retired, have ventured into political territory on more than one occasion.
-Nonetheless, academic theorists and military leaders alike have long recognized the importance of professional conduct rooted in non-partisan, apolitical conduct.
-At its core, such conduct ultimately earns the trust and confidence of the American people.
+From MacArthur's clash with President Truman in the early 1950's, to the so-called "Revolt of the Generals" in 2005-2006, military leaders, even those who are retired, have ventured into political territory, in one form or another, on more than one occasion.
+Nonetheless, academic theorists, military tradition, and contemporary military leaders have long argued for the importance of the military remaining grounded in professional conduct that is both non-partisan and apolitical.
+At its core, it is this type of professional conduct which ultimately postures the military to earn the trust and confidence of the American people.
 
-It must be stressed that the concept of military professionalism does not impose a normative prohibition on the publication of an opinion piece, or the expression of other forms of public dialogue, by a military professional (though there are certainly practical differences between the manner these are employed by those who currently serve and those who are retired).
-Indeed, Huntington's concept of military professionalism recognizes the unique skills of military officers, some of whom, especially multistar generals and admirals, possess more than 30 and in some cases 40 years of professional experience.
+To be clear, the concept of military professionalism does not impose a normative prohibition on the publication of an opinion piece, or the expression of other forms of public dialogue, by a military professional.
+In fact, one of the foundations of Huntington's concept of military professionalism is the fact that military leaders possess unique and specialized skills that civilians simply do not have.
+Moreover, multistar generals and admirals in the contemporary military possess in many cases upwards of 30 and sometimes 40 years of service, and it is entirely reasonable to conclude that these officers in particular have been privy to decisions and experiences occurring at or near the highest levels of government.
+In other words, their experiences are not confined to only tactical matters.
+They are familiar with issues ranging from how wars should be best fought, to future security threats looming on the horizon, to fixing the institutional culture of a military service, and more.
 
-It is reasonable to conclude that a professional opinion piece written by a military officer would take on three very important qualities.
-First, the military professional would address a relevant military topic.
-To be clear, multistar generals and admirals often are experienced in the realm of strategy, a word that is often misunderstood and misapplied in discussions of the topic [see for example @strachan_direction_2014].
-Without unpacking the concept of strategy further, the relevant point is that military elites, and especially those who have achieved a high rank, have been privy to decisions and experiences occurring near the highest levels of government, and as such, their experiences are not confined to only tactical matters.
-They are familiar with issues ranging from how wars should be best fought, to future security threats looming on the horizon, to fixing the institutional culture of a military service.
-Second, from a credibility and expertise perspective, a professional piece would be based on the officer's personal experiences and would be written from an individual's perspective so as not to infer that an argument is being made on behalf of the collective service or element within the military body.
-Third, an opinion made by a military officer would be free of any hint of partisan sentiment, whether advocacy or denouncement.
-If one of these three qualities fails to be upheld, there is increasing likelihood that the officer's words can, either intentionally or unintentionally, serve as a weapon, and, like other kinds of weapons, cause great harm.
-To understand how, we will now turn our attention to understanding the impacts on military professionalism caused by two significant changes that have developed within American Society in the past 50 years.
-The first is increasing credibility of the military institution, and the second is the introduction of an unparalleled and divisive political polarization.
-The interaction of these forces is significantly straining the professional conduct of the military institution.
+It is reasonable, therefore, to conclude that a professional opinion piece written by a military officer would take on three qualities.
+First, a professionally-written opinion piece would address a relevant military topic consistent with the military expertise of the author.
+In the US context in particular, where the military conducts a variety of roles and performs a variety of missions, this would include a wide swath of topics, while at the same time, certain topics would be considered out of the realm of relevant military expertise, such as partisan domestic politics.
+Second, from the perspective of both expertise and institutional credibility, a professional opinion piece would be based on the officer's personal experiences and be written from an individual's perspective so as not to infer that an argument is being made on behalf of the military.
+Third, a professionally-written opinion piece would be free of of partisan sentiment or expression, whether advocacy or denouncement.
+If one of these three qualities -- addressing a relevant military topic, written from the perspective of an individual officer's expertise, free of partisan sentiment -- fails to be upheld, there is an increasing likelihood that the officer's words can be conveyed and/or received as unprofessional.
+
+If and when a military officer violates standards of military professionalism when conveying an opinion to the American public, his or her words can cause great harm, either to American society, the military as an instition, or a combination of both.
+To understand how and why this is the case, we will now turn our attention to understanding the impacts on military professionalism caused by two changes that have developed within American Society in the past 50 years.
+The first is a gradual increase in the American public's opinion of the military as an institution, and the second, more significant change is the steady increase in political polarization within American society.
 
@@
 
-One of the two significant societal changes that has occurred in America during the past fifty years is the increase in credibility of the military institution.
-Consider that in the late 1970s, as a likely result of public opinion over the Vietnam War, 54% of Americans indicated that they trusted the military "a great deal or quite a lot," whereas in 2019, 73% of Americas expressed the same sentiments [@gallup_polls_confidence_2007].
-Compared to other government institutions, the American public trusts the military to a remarkably high degree.
-According to a recent Pew Research Center poll, 80% of the American public expects the military to act in the best interests of the American public, compared to business leaders (45%), the media (40%), and elected officials (25%) [@johnson_trust_2018].
-It is accurate to say that America trusts its military more than any other entity.
+The credibility of the American military as an institution has increased during the past fifty years.
+In the late 1970s, likely as a result of public opinion over the Vietnam War, 54% of Americans indicated that they trusted the military "a great deal or quite a lot," whereas in 2019, 73% of Americas expressed the same sentiments [@gallup_polls_confidence_2007].
+Between these time periods, there have been some shorter and localized changes, but by and large, the increase that has occurred has been gradual and steady.
+Today in fact, relative to other government institutions, the American public trusts the military to a remarkably high degree.
+For example, according to a recent Pew Research Center poll, 80% of the American public expects the military to act in the best interests of the American public, compared to tho percentage who believe the same about business leaders (45%), the media (40%), and elected officials (25%) [@johnson_trust_2018].
+Military leaders are, rightfully so, cognizant of the benefits of holding the American public's trust.
+Former Chief of Staff of the Army, Retired General Ray Odierno, told fellow Army generals that "of great importance within the Army Profession is the duty of each of us to be a steward of our profession and to maintain its essential and hard won trust with society [Ray Odierno in @us_army_repairing_2014].
+For this reason, for example, there has been a concerted effort by the services to identify and punish senior leaders who engage in activities that may damage the "hard won" trust Odierno is referring to, activities such as sexual or financial misconduct [@us_army_repairing_2014].
 
-Military leaders are, rightfully so, cognizant of the benefits of holding the public's trust.
-In the past several years for example, there has been a concerted effort by the services to identify and punish senior leaders who engage in misconduct, to demonstrate to Congress that senior military leaders do not need assistance to handle sexual assault, and to otherwise prove that it can be trusted to accomplish its missions [@us_army_repairing_2014].
-Former Chief of Staff of the Army, Retired General Ray Odierno, told fellow Army generals that "of great importance within the Army Profession is the duty of each of us to be a steward of our profession and to maintain its essential and hard won trust with society [Ray Odierno in @us_army_repairing_2014].
-
-Scholars have also shown that when elites possess a high degree of credibility, their ability to influence and shape public opinion is likewise high.
+Scholars have shown that when elites possess a high degree of credibility, their ability to influence and shape public opinion is likewise high.
 In his influential study, _The Nature and Origins of Mass Opinion_, John Zaller argues that formations in mass public opinion are the result of three interacting components: "variation in the information carried in elite discourse, individual differences in attention to this information, and individual differences in political values and other predispositions" [@zaller_nature_1992].
@@
 
-The second and decisive societal change that has occurred in the past fifty years is the degree to which and character of the political polarization prevalent in the US political landscape today.
-Although scholars have offered differing explanations for the causes of this polarization, ranging from concerns over race and religion [@abramowitz_great_2018], to the size and role of government [@sides_two_2015], to fierce policy disagreements over policy issues [@mann_its_2012], virtually all scholars agree that polarization in America has become more pronounced in recent decades [@hare_polarization_2014; @bonica_mapping_2014], and that it is, by many measures, worse than it has been during any period since the end of the American Civil War [@hare_polarization_2014].
+The second and more influential societal change that has occurred over the past fifty years in America is the rise of a polarized political landscape.
+Though this is not the first time in America's history that political polarization has been a dominant feature of domestic politics, multiple scholars have argued that polarization has worsened in recent decades [@hare_polarization_2014; @bonica_mapping_2014]], and some have argued that polarization is at the worse level since the end of the American Civil War [@hare_polarization_2014].
+Although scholars have offered differing explanations for the cause of the political polarization evident today, ranging from concerns over race and religion [@abramowitz_great_2018], to the size and role of government [@sides_two_2015], to fierce policy disagreements over policy issues [@mann_its_2012], there is a fair amount of common ground among scholars regarding how polarization has impacted contemporary domestic politics.
+It is worth unpacking several features of America's polarized landscape, particularly for how these features might alter professional military behavior specifically and civil-military relations more broadly.
+
 The polarized political climate today extends to far more issues than previously polarized periods in American history [@layman_party_2002], is marked by stronger negative feelings towards those who hold opposing views [@abramowitz_great_2018], and, at root, involves fundamental issues of both personal and national identity that are not conducive to any reasonable form of compromise [@abramowitz_great_2018; @mohler_jr_gathering_2020].
@@
 
-The first way in which today's highly polarizing political climate is different than that of previous periods in American history is its scope.
-Fewer and fewer issue areas and policy matters are outside of the "spell" of polarization, owing to a phenomenon known as "conflict extension" [@@layman_party_2002].
-In previous eras of polarization, a single issue area served as the "battleground" for partisan conflict.
-Today, though, this is no longer true.
-Research has consistently shown that the polarizing politics present in the United States today has caused voters who identify as liberals to be more liberal across multiple issue categories, and that the same holds for conservatives [@abramowitz_great_2018].
-This includes the full range of social welfare, cultural, and racial issues, all of which are important in the minds of the electorate and partisans [@layman_party_2006].
-As a result, the prospect for cooperation on issues that were not politically polarized remained high, whereas today, the prospect for cooperation in general is lower as a result of the shrinking, non-partisan space [@layman_activists_2010].
+The first important characteristic of the polarized political climate is its all-encompassing scope.
+Fewer and fewer issue areas and policy matters are outside of the reaches of polarization, a phenomenon that American politics scholars have dubbed "conflict extension" [@layman_party_2002].
+These scholars note that as polarization has increased, more issue areas have become divisive in a partisan sense.
+The practical effect of "conflict extension" is that the perceived non-partisan or apolitical space has significantly shrunk.
+In fact, scholars have also shown that consistent with trends of increasing polarization, voters who identify as liberals have become more more liberal across more categories of issues, and that the same trend holds for conservatives [@abramowitz_great_2018].
+This includes a range of social welfare, cultural, and racial issues, all of which are important in the minds of the electorate and partisans [@layman_party_2006].
+Whereas in previous political eras, the prospect for cooperation on issues that were not politically polarized remained high, the prospect for cooperation across partisan sides is lower as a result of the shrinking, non-partisan space [@layman_activists_2010].
 
-The second important characteristic of today's polarizing climate involves the strong, negative feelings that each side espouses towards those who hold differing opinions on a variety of issues.
-Known as "the rise of negative partisanship" [@abramowitz_great_2018], these strongly negative feelings toward those who hold different political opinions is not a phenomenon that has occurred among political elites only; instead, this trend has been shown to have grown among voters since the late 1970s [@abramowitz_great_2018].
-Therefore, in addition to there being a diminished opportunity for compromise owing to the proliferation of issues that are polarized, the feelings produced by today's polarized climate are so strong that the other side believes that other is deeply and morally wrong.
+The second important characteristic of today's polarizing climate involves the rise of strong, negative feelings that each side espouses towards those who hold differing opinions on a variety of issues.
+Known as the rise of "negative partisanship" [@abramowitz_great_2018], citizens holding strongly negative feelings toward those who hold different political opinions is not a phenomenon that has occurred only among political elites; instead, this trend has been shown to have grown among all voters since at least the late 1970s [@abramowitz_great_2018].
+The effect of "conflict extension" and "negative partisanship" is a reduction in perceived or actual nonpartisan space as well as a genuine belief that opposing partisans are deeply wrong and flawed.
 
-The third characteristic of today's polarizing environment is that polarizing issues do not involve matters of simple policy difference, but rather deep and fundamental beliefs about individual and national identities and purposes [@abramowitz_great_2018; @sides_identity_2018; @mohler_jr_gathering_2020].
-The election of Donald Trump and his years in office as President have raised the salience on a host of issues, all of which are essential to various parts of the electorate, including attitudes about race, immigration, and religion  [@sides_identity_2018].
-Moreover, the American people are aware of the polarized environment, and in a recent survey, indicated that they are concerned the trend is not improving [@schaeffer_far_2020].
+The third characteristic of today's polarizing environment is that the issues which partisans disagree about do not involve matters of simple policy difference, but rather deep and fundamental beliefs about individual and national identities and purposes [@abramowitz_great_2018; @sides_identity_2018; @mohler_jr_gathering_2020].
+Indeed, the election of President Donald Trump and his years in office as President have raised the salience on a host of fundamental issues, many of which are dearly held by various parts of the electorate, including attitudes about race, immigration, and religion  [@sides_identity_2018].
+The impact of this particular feature of contemporary political polarization in America is that real moral issues are at stake for partisans on each respective side of the political divide.
 
-In summary, taken together, these three characteristics of contemporary US political polarization are sobering.
-As a large body of research indicates, it is right to understand the political climate which American finds itself in today not merely as another polarized period, but rather in a tumultuous era that may well challenge the very definition of what America is and what it stands for [@mann_its_2012; @sides_identity_2018; @abramowitz_great_2018; @mohler_jr_gathering_2020].
-How the military may be impacted by the forces of increasing institutional credibility and today's political polarization is the subject of the next section.
+Taken together, these three characteristics of contemporary US political polarization offer a sobering assessment of the contemporary political landscape.
+As a large body of research indicates, it is right to understand that America finds itself today not merely in yet another polarized period, but rather in a tumultuous era that may well challenge the very definition of what America is and what it stands for [@mann_its_2012; @sides_identity_2018; @abramowitz_great_2018; @mohler_jr_gathering_2020].
+For anyone - citizens, military officers, or elected officials - to assume that the military will automatically escape these dynamics of polarization simply because of the legacy of the military's professional ethos is to underestimate the strength and all-encompassing nature of the divisions in American society that political polarization underscores.
+Let us now turn our attention to understanding how the features of today's politically polarized climate may impact the professional nature of the military institution.
 
@@
 
-Figure 1, below, visually describes a general theory of how institutional credibility and political polarization interact to impact military professionalism.
-From left to right, two significant societal changes (node 'SC') have occurred since the late 1970s, both of which have been described, and as this theory contends, interact, to impact military professionalism.
-The first societal change, that of institutional credibility, is indicated by the node 'IC', and the second, that of political polarization, is indicated by the node 'PP'.
-Each of these, in turn, impact military professionalism, indicated by the node 'MP'.
-Last, though it is outside of the scope of this paper, I contend that changes in military professionalism as a result of the forces of increasing political polarization and institutional credibility, will impact various military outcomes (node 'MO').
-These military outcomes vary, and could include things such as military effectiveness, to include the military's ability to fight and win battles [@talmadge_dictators_2015], as well as strategy formulation, to include the military's ability to work with civilians to produce sound strategic advice [@rapp_civil-military_2015].
+Figure 1, below, shows describes a theory for how the variables of political polarization, institutional credibility, military professionalism, and other military outcomes may relate to each other.
+From left to right, the theory contends that increasing political polarization ultimately reduces military professionalism.
+At the same time, however, there is an intervening variable, institutional credibility, that may actually increase as a result of increasing political polarization.
+Finally, on the far right of the diagram, are changes in military outcomes.
+Though it is outside of the scope of this paper, I contend that eventually, increases in political polarization may at some point cause discernible changes in various military outcomes.
+These military outcomes may vary, but could include things such as military effectiveness, to include the military's ability to fight and win battles [@talmadge_dictators_2015] and to formulate sound strategic advice [@rapp_civil-military_2015].
 Indeed, the full degree to which and in what ways polarization will impact the military are not yet fully known.
@@
     MP ~ PP + IC,
-    PP~SC,
-    IC~SC,
+    IC~PP,
     coords = tribble(
@@
       "MP", .7, .5,
-      "IC", .3, 1,
-      "PP", .3,0,
-      "SC", 0, .5
+      "IC", .3, .2,
+      "PP", .1,.5,
     ))
@@
 problem_dag %>% tidy_dagitty() %>% 
-  dag_label(labels=c("MO"="Military\n Outcomes", 
-                     "MP"="Military\n Professionalism", 
-                     "PP"="Political\n Polarization", 
-                     "IC"="Institutional\n Credibility", 
-                     "SC"="Major Changes\n In American Society")) %>% 
+  dag_label(labels=c("MO"="Changes in\n Military\n Outcomes", 
+                     "MP"="Reduction in\n Military\n Professionalism", 
+                     "PP"="Increasing\n Political\n Polarization", 
+                     "IC"="Rising\n Institutional\n Credibility")) %>% 
   ggplot(aes(x=x, y=y, xend=xend, yend=yend))+
@@
   geom_dag_text()+ 
-  geom_dag_label_repel(aes(label=label, fill=label), col="white", show.legend=FALSE)+ theme_dag() + labs(title="Figure 1", subtitle="The Impact of Political Polarization and Institutional Credibility on Military Professionalism")
+  geom_dag_label_repel(aes(label=label, fill=label), col="white", show.legend=FALSE)+ theme_dag() + labs(title="Figure 1", subtitle="The Impact of Political Polarization on Military Professionalism")
 
@@
 
+predictiondatafut <- tibble(House_distance_dim1=seq(min(slimdata$House_distance_dim1), 1, by=.01), Inst_Cred_Gallup = mean(slimdata$Inst_Cred_Gallup)) %>%
+crossing(Ques_Prof = c(0,1))
+
+predictiondatafut
+
 # actual prediction data now with code.  Create variables that tie the model together. #Create Variables .prob, conf.low, conf.high, and pred_final_rec
@@
 augment(logmodel1, newdata = predictiondata)
+
+augment(logmodel1, newdata = predictiondatafut)
 
@@
 
+predictions2 <- augment(logmodel1, newdata=predictiondatafut) %>% 
+  mutate(
+    .prob=plogis(.fitted),
+    conf.low = plogis(.fitted -(qnorm(.975) * .se.fit)),
+    conf.high = plogis(.fitted + (qnorm(.975) * .se.fit)),
+    pred_final_rec = case_when(
+      .prob >= .5 ~ "Likely Violation of Professionalism",
+      .prob < .5 ~ "Violation of Professionalism not likely"
+    )
+  ) 
+predictions2 %>% arrange(.prob, .fitted, Ques_Prof, desc(.fitted)) %>% 
+  view()
+
 # now let's try some graphs
@@
 
+ggplot(predictions2) +
+  aes(x = .fitted) +
+  geom_histogram()
 # if we wanted predictions on the probability scale,
@@
 predictions %>%
+  ggplot() +
+  aes(x = .fitted, y = .prob) +
+  geom_pointrange(aes(ymin = conf.low, ymax = conf.high)) +
+  geom_smooth()
+
+predictions2 %>%
   ggplot() +

### 2020-08-28 code/Submissionv3.Rmd
commit

@@
 
-# When Words Become Weaponized - A Theory of Credibility and Polarization
+# When Words Become Weaponized - A Theory of Political Polarization and Military Professionalism
 
-Figure 1, below, shows describes a theory for how the variables of political polarization, institutional credibility, military professionalism, and other military outcomes may relate to each other.
-From left to right, the theory contends that increasing political polarization ultimately reduces military professionalism.
-At the same time, however, there is an intervening variable, institutional credibility, that may actually increase as a result of increasing political polarization.
-Finally, on the far right of the diagram, are changes in military outcomes.
-Though it is outside of the scope of this paper, I contend that eventually, increases in political polarization may at some point cause discernible changes in various military outcomes.
-These military outcomes may vary, but could include things such as military effectiveness, to include the military's ability to fight and win battles [@talmadge_dictators_2015] and to formulate sound strategic advice [@rapp_civil-military_2015].
+Figure One, below, shows describes a theory for how the variables of political polarization, institutional credibility, military professionalism, and other military outcomes may relate to each other.
+The theory contends that increasing political polarization ultimately reduces military professionalism, but that there may also be an intervening variable, rising institutional credibility, that also is impacted by political polarization and yet further impacts military professionalism.
+Finally, though it is outside the scope of this paper, I contend that an increase in political polarization may at some point cause discernible changes in various military outcomes.
+Such outcomes may include a multitude of topics, but would include items such as military effectiveness, to include the military's ability to fight and win battles [@talmadge_dictators_2015], and to formulate sound strategic advice [@rapp_civil-military_2015].
 Indeed, the full degree to which and in what ways polarization will impact the military are not yet fully known.
-
-The theory grounds the two independent variables discussed here, credibility and polarization, as changes that have occurred within American society in the past fifty years.
-Because these changes stem from changes that have their originate within broader American society, my theory contends that the military as an institution and the military professionals that comprise it are not impervious to their impacts.
-In other words, my theory assumes that maintaining professional conduct by the military institution requires a deliberate effort and is hard work, despite the fact that behaving in this manner is a long and storied norm.
-Professional conduct is not a guaranteed nor automatic outcome.
-
-When institutional credibility is high, a demand signal is created by the American public such that future interactions between the military institution and the American public are encouraged and welcomed.
-From an institutional perspective, the military might respond to the public's demand signal for increased military interaction by increasingly participating in public ceremonies, parades, and other showcases of military prowess (at sporting venues, for instance).
-From an individual perspective, military elites, especially retired military officers, would be psychologically primed to increasingly dialogue with the American public when military credibility is high.
-By extension, other outlets such as the press would also be primed to showcase military elites when military credibility runs high, because these elites represent an institution that commands a high degree of public respect.
-
-When military credibility is low, the theory contends the opposite will occur.
-Since demand for interactions between the American public and the military is muted when institutional credibility for the military is low, the institution will not engage with the public as frequently nor as intimately because the strong demand signal that existed when institutional credibility was high no longer exists.
-At the individual level, military elites will also not engage as often with the public, nor will the press ask these elites to.
-In sum, institutional credibility and respect for the military institution impacts the likelihood of the military institution and military elites interacting with the American public.
-Such interaction can take a variety of forms.
-When respect for the military is high, the military is likely to interact, speak to, and dialogue with the American public, and when institutional credibility is low, the military is less likely to do so.
 
@@
   geom_dag_text()+ 
-  geom_dag_label_repel(aes(label=label, fill=label), col="white", show.legend=FALSE)+ theme_dag() + labs(title="Figure 1", subtitle="The Impact of Political Polarization on Military Professionalism")
+  geom_dag_label_repel(aes(label=label, fill=label), col="white", show.legend=FALSE)+ theme_dag() + labs(title="Figure One", subtitle="The Impact of Political Polarization on Military Professionalism")
 
@@
 
-The degree and character of political polarization in a society also has the potential to impact the professionalism of the military as an institution.
-To understand how, recall that the three salient characteristics of the political polarization prevalent in contemporary America include an increase in the number and scope of issues that have become polarized (the phenomenon referred to as "conflict extension" [@layman_party_2002]), an increase in the strength of negative feelings held by supporters of an issue toward the opposition (the rise of "negative partisanship" [@abramowitz_great_2018]), and an increase in the salience of core issues related to individual and national identity [@white_values_2002; @mann_its_2012; @sides_identity_2018].
-These three characteristics convey several implications for the military as an institution and its leaders.
+What mechanisms might be at work through increasing political polarization that ultimately impact military professionalism?
+To answer this important question, I rely on the three major characteristics of the contemporary polarized political landscape previously identified by scholars of American Politics.
+These included, first, a sharp increase in the number and scope of issues that have been subsumed by partisan debate and disagreement (the phenomenon referred to as "conflict extension" [@layman_party_2002]); second, the increase in the strength of negative feelings held by one partisan group towards those of the other side (the rise of "negative partisanship" [@abramowitz_great_2018]); and third, an increase in the salience of debates over issues that individual partisans consider core to their individual identities [@white_values_2002; @mann_its_2012; @sides_identity_2018].
+Each of these three features of the politically polarized landscape strain the concept of military professionalism in important ways, and taken together, comprise a significant challenge to the individual military professional and thus, to the professional nature of the military as a whole.
+These challenges are summarized in Table One, below.
 
-First, there is increasing difficulty for individuals, to include military professionals, to remain neutral as polarized politics subsumes more and more issues, many of which involve issues which individuals hold to dearly.
-In previous eras, when polarization was not as extreme as it is today, it was more possible to stay out of politics.
-Today, polarization seems to do the opposite: polarized politics seem to thrust issues and the unfolding political fight at the individual, even when the individual is not seeking confrontation.
-A second implication of polarization is that it raises the salience for individuals of right and wrong in a moral sense, which increases the odds of an individual expressing partisan sentiment.
-In other words, in saying or doing what one believes to be right, one is increasingly likely to also be taking a partisan position.
-The third and most important implication for the military professional as a result of contemporary political polarization is that it impacts and distorts the notions of two character traits that are held dearly by members of the military profession: loyalty and duty.
+\singlespacing
+\begin{table}[ht]
+\begin{footnotesize}
+\caption{Mechanisms of Rising Political Polarization Which Impact Military Professionalism} % title of Table
+\centering % used for centering table
+\begin{tabular}{|p{5cm}|p{5cm}|p{5cm}|p{5cm}|}
+    \hline
+    \textbf{Characteristics of Increasing Political Polarization} & \textbf{Description} & \textbf{Resulting Strain on Military Professionalism}  \\ \hline
+    \textbf{Conflict Extension} & Less politically-neutral space exists as partisan disagreements subsume a greater number of issues & Military Professionals face an increasingly difficult prospect of remaining apolitical.
+\\ \hline
+    \textbf{Negative Partisanship} & Strong feelings of personal dislike and disdain for those who hold to opposing political views and who are on the "other side" & Greater willingness to confront or take issue with those with opposing views\\ \hline
+    \textbf{Core and dearly-held issues are increasingly at stake} & The salience of one's identity and belonging to a particular group is raised; Belief that moral right and wrong are at stake & Changes in one's concept of loyalty \\ \hline
+\end{tabular}
+\label{table:comparison} % is used to refer this table in the text
+\end{footnotesize}
+\end{table}
+\doublespacing
 
-As any soldier knows, few traits are more sacred in the military than loyalty and duty.
-These virtues are extolled in a host of codes and mantras that have played a central part of the American military since its inception.
-They are captured in phrases that military professionals say to themselves and to each other: "do the right thing," "lead with integrity," and, as the West Point Cadet Prayer so elegantly states, choose "the harder right over the easier wrong."
+The first way in which the concept of military professionalism is challenged as a result of increasing political polarization is that there is increasing difficulty for individuals to remain politically neutral, non-partisan, or otherwise apolitical - a hallmark of military professionalism.
+This occurs because as "conflict extension" [@layman_party_2002] occurs and an increasing number of issues are swallowed up by partisan battlegrounds, there is less politically neutral space for someone to stand on.
+When polarization levels in America were not as high as they are today, it was more possible for individuals to either stay out of politics, avoid partisan squabbles, and otherwise to stay politically neutral.
+Today's polarizing environment seems to have the opposite impact: polarized politics seem to thrust issues and the unfolding political fight at the individual citizen, even when the individual is not deliberately seeking confrontation.
+There is no theoretical reason to suspect that this mechanism would not impact military leaders.
+Military leaders are citizens of the Republic, and though the institutional ethos of the military profession is to remain apolitical, such an ideal is difficult to uphold if less and less neutral political space exists.
 
-Political polarization, however, obfuscates these ideals, and places greater responsibility for interpreting the meaning and the associated action of these ideals on the individual and less on the corporate military institution.
-This stems from the fact that highly polarized political climate may involve specific issue areas that certain members of the military feel a stronger affinity towards than other members.
-Two different military officers may adopt completely different opinions of a contentious issue, each believing that the other's opinion is wrong.
-Should the officer act on that opinion, the other might believe he or she is operating outside of the proper military frameworks, or outside of what the right thing to do is according to the Constitution or other loyalty structures.
-Each officer, then, might accuse the other of failing to be loyal to the ideals of the military.
+A second challenge to military professionalism stems from the "negative partisanship"[@abramowitz_great_2018] that dominates the polarized American political landscape.
+Strong feelings towards those on a host of other issues are becoming increasingly hostile.
+These feelings are growing increasingly strong likely because of the other prominent features of political polarization: a decrease in the availability of neutral political space, and the importance of issues that are at stake in many of today's political debates.
+
+A third, and I contend a more insidious, mechanism through which increasing political polarization works to impact military professionalism is by altering one's sense of loyalty and identity.
+This occurs because the issues that are at stake as political polarization increases are in many instances involving issues that individuals associate with core concepts of one's identity, and thus, one's sense of moral right and wrong.
+Because individuals care about their identities and what they believe to be right and wrong, they are willing to defend these values, even at a cost.
+The rise of "identity politics" in America as political polarization has worsened illustrates this point.
+
+However, it is also precisely on this point that the military institution is perhaps most vulnerable to increasing political polarization.
+I contend that the rise of identity politics can and will be harmful if a military professional's identity and the loyalty associated with that identity directly conflicts with his or her identity and loyalty as a professional soldier.
+As any soldier knows, and as civilians can reasonably understand,few traits are more sacred in the military than loyalty.
+Indeed, the virtue of loyalty is extolled in a host of codes and mantras that have played a central part of the American military since its inception, and loyalty is captured, either in word or in spirit, in a host of phrases and professional codes.
+In the middle of the the West Point Cadet Prayer is the eloquently written plea, "make us to choose the harder right instead of the easier wrong" [@wheat_west_1918].
+
+High levels of political polarization obfuscate the ideal of loyalty, however.
+As core issues that involve individual perceptions of moral right and wrong are at stake, individuals, to include military officers, are less likely to not become involved, attached, or otherwise invested in these issues.
+This is especially true when the language and appeal of concepts of loyalty are employed to justify one is or is not taking action over an issue.
+In a highly polarized political climate, there is increasing likelihood that two different people military officers may adopt completely different opinions regarding a contentious issue, each believing that the other is deeply and perhaps even morally wrong.
 
@@
 
-In summary, it can be said that high degrees of political polarization impact military professionalism in key ways.
-Polarization increases the chances of expressing partisan sentiment in the name of doing what is morally right.
-Polarization, when it is high, also increases the chances of speaking out on issue areas that are outside of the realm of traditional military expertise, especially as the degree of polarization and "conflict extension" [@layman_party_2002] spreads, which reduces the perceived non-partisan space available to military professionals.
-Finally, polarization muddies the concepts of loyalty and duty.
-Strong feelings about issue areas perceived as crucial and essential by some greatly increase the odds of charging others of failing to act loyally or to fulfill one's duty.
+# Accounting for Institutional Credibility
 
-Ultimately, four possible worlds exist when the impacts of military credibility and political polarization are considered in tandem (see Table 1, below).
+The previous section described how increasing political polarization presents significant challenges to the concept of military professionalism.
+But when will these challenges become public knowledge, and perhaps more relevant, how does increasing political polarization impact the credibility of the military institution?
+
+One can imagine several possible scenarios.
+It is reasonable to consider that as political polarization increases, respect and trust in seemingly partisan institutions would decrease, while trust in nonpartisan and apolitical institutions would rise.
+But why?
+Would such a trend reflect a genuine increase in trust in the nonpartisan institution, or would it reflect the simple consequence that as one institution becomes less respected, others might simply gain as a result?
+Certainly, more work needs to be done in this area.
+
+Nonetheless, I contend that when institutional credibility is high, a demand signal is created by the American public such that future interactions between the military institution and the American public are encouraged and welcomed.
+From an institutional perspective, the military might respond to the public's demand signal for increased military interaction by increasingly participating in public ceremonies, parades, and other showcases of military prowess (at sporting venues, for instance).
+From an individual perspective, military elites, especially retired military officers, would be psychologically primed to increasingly dialogue with the American public when military credibility is high.
+By extension, other outlets such as the press would also be primed to showcase military elites when military credibility runs high, because these elites represent an institution that commands a high degree of public respect.
+
+When military credibility is low, the theory contends the opposite will occur.
+Since demand for interactions between the American public and the military is muted when institutional credibility for the military is low, the institution will not engage with the public as frequently nor as intimately because the strong demand signal that existed when institutional credibility was high no longer exists.
+At the individual level, military elites will also not engage as often with the public, nor will the press ask these elites to.
+In sum, institutional credibility and respect for the military institution impacts the likelihood of the military institution and military elites interacting with the American public.
+Such interaction can take a variety of forms.
+When respect for the military is high, the military is likely to interact, speak to, and dialogue with the American public, and when institutional credibility is low, the military is less likely to do so.
+
+Ultimately, four possible worlds exist when the impacts of military credibility and political polarization are considered in tandem (see Table Two, below).
 The first is when political polarization and military credibility are both low.
@@
 \begin{footnotesize}
-\caption{Four Types of "Military Professionalism Worlds" Where Institutional Credibility and Political Polarization Vary} % title of Table
+\caption{Table Two: Four Types of "Military Professionalism Worlds" Where Institutional Credibility and Political Polarization Vary} % title of Table
 \centering % used for centering table
@@
 Indeed, this set of conditions comprise a very dangerous world for the military institution and the military professional.
-
 # Methodological Approach, Data, and Hypotheses
 
-I now aim to see whether the discourse of retired military officers has changed over time, specifically since the 1970s, as polarization has worsened and as the institutional credibility of the military has increased.
-To answer this question, I assemble and examine a data set that consists of opinion pieces authored by retired military officers in major US newspapers.
-My methodological approach involves two key steps.
-First, I determine if there have been shifts in the topics that retired military officers have been writing about.
-Second, I apply a normative lens to determine if there have been shifts in the tone and tenor of the pieces authored by these officers, particularly regarding the professional aspects of the military previously discussed.
-In simple terms, I want to answer the questions, what are these retired military officers writing about, and through the lens of professionalism, should they be saying what they are saying?
+I now aim to see whether the discourse of retired military officers has changed over time, specifically since the 1970s, as polarization has worsened dramatically.
+To answer this question, I assemble and examine a data set that consists of opinion-editorial publications authored by retired military officers in major US newspapers.
+My methodological approach involves two fundamental steps.
+First, I determine what topics retired military officers have been writing about over time.
+Second, I apply a normative lens to determine if there have been shifts in the tone and tenor of the pieces authored by these officers, particularly regarding the professional aspects of the military previously discussed (apolitical in nature, non-partisan, relying on relevant military expertise, etc.).
+In essense, I want to answer the questions, what are these retired military officers writing about, and through the lens of professionalism, should they be saying what they are saying?
+Is there any indication that increasing political polarization is altering what and how these retired officers are communicating with the American public?
 
-Using the Factiva and Nexus-Uni databases, I conduct a search that focuses keywords in the byline (author field) such as "retired military officer," "retired from the Army, Navy, Marine Corps, Air Force," "former Chairman of the Joint Chiefs of Staff," "General," "Admiral," and so forth.
-For sources, I confine my search to the largest US newspapers that regularly run opinion-editorial pieces, which includes  _The Wall Street Journal_ (WSJ), _The New York Times_ (NYT), _The Washington Post_ (WaPo), _The Los Angeles Times_, and _The USA Today_.
-I begin the search for the data of January 1, 1979 - 30 June 2020^[I also open my search to include both print and online offerings to ensure that I am capturing any opinion pieces published digitally.
-Most of the newspapers initiated digital offerings in the late 1990s.].
-Using newspapers, and particularly these specific ones, provides some advantages as well as disadvantages.
-The primary advantage is that these sources have been around since the period I wish to examine, unlike many of the online only sites that have only been present in the digital age.
-This allows me to form a data set which covers more time.
-The primary disadvantage is that I am not capturing all of the opinion-pieces authored or given by retired military officers.
-For example, retired military officers are common on cable TV, often weighing in on topics, and I am not capturing those here.
+To assemble the data, I use the Factiva and Nexus-Uni databases, conducting a search that focuses keywords in the byline (author field) such as "retired military officer," "retired from the Army, Navy, Marine Corps, Air Force," "former Chairman of the Joint Chiefs of Staff," "General," "Admiral," and so forth.
+I confine my search to the largest US newspapers that regularly run opinion-editorial pieces, which includes  _The Wall Street Journal_ (WSJ), _The New York Times_ (NYT), _The Washington Post_ (WaPo), _The Los Angeles Times_, and _The USA Today_.
+The time period for the data includes the period January 1, 1979 - 30 June 2020, the main limitation being that not all of the Newspaper sources are available for previous dates.
+I also open my search to include both print and online offerings to ensure that I am capturing any opinion pieces published digitally.
+
+The sources that I rely upon provide some relative advantages, but they are not without limitations.
+The primary advantage is that these sources have some longevity; they are not modern blogging sites or online-only offerings.
+As a result, I can truly examine a larger period of time and thus, examine whether how political polarization has impacted the way in which military officers engage with the American Public.
+Simultaneously, though, this presents a disadvantage.
+Retired military officers engage with the American public through a host of mediums, to include social media, blogging sites and cable television, neither of which I am capturing in my data set.
 It is also true that over time, social media has taken over as the primary means in which a majority of Americans receive their news [@martin_how_2018].
@@
 I conduct a separate search for all opinion pieces penned by these authors, regardless of whether they were identified as a retired military officer or not, concluding that it is best do so for coding purposes.
+
 Ultimately, the data set contains 334 opinion-editorial publications, penned by 178 different authors or author teams.^[There are several co-authored pieces.
 For example, where General (Retired) David Petraeus wrote one publication and another one was written by General (Retired) David Petraeus and Michael O'Hanlon, I counted these as two different authors for coding purposes.] 10 of these pieces are written unanimously, with nods by the authors to their previous service.
-
 From a macro perspective, consider for a moment all that the date range includes with respect to the US military.
@@
 The debates over "Don't Ask, Don't Tell" and policies regarding women in the combat arms are also included in the data, as are numerous discussions of military budgets, weapons systems, and alliances.
-Attempting to reconcile the fact that the military as an institution prides itself on being a non-partisan and apolitical institution, that newspaper circulations (even after accounting for digital offerings) are substantially down over the past 40 years, and that the US political environment is increasingly polarized along partisan lines, I adopt the following hypotheses:
+
+Attempting to reconcile the fact that the military as an institution prides itself on being a non-partisan and apolitical institution, that newspaper circulations (even after accounting for digital offerings) are substantially down over the past 40 years, and that the US political environment is increasingly polarized,  I adopt the following hypotheses:
 
@@
 
-_$H_2$: Because overall newspaper consumption has dropped in the past 40 years, there will be fewer opinion pieces over time authored by military officers._
+_$H_2$: Because overall newspaper consumption has dropped in the past 40 years, there will be fewer opinion pieces over time authored by retired military officers._
 
-_$H_3$: Given that polarization has become increasingly pronounced in the United States in recent years, there will be an increase in the number of op-eds which reveal political and partisan sentiments (indication of a drop in professionalism)._
+_$H_3$: Given that polarization has become increasingly pronounced in the United States over the time period examined, there will be an increase in recent years in the number of op-eds which reveal political and partisan sentiments, indicating a drop in military professionalism._
 
-# First Layer Analysis: Topic Addressed
+# First Layer Analysis: Topics Addressed by Retired Military Officers
 
-In the first level of analysis, after capturing key biographical and publication information, such as author name, article title, source, publication date, and branch of service and rank of the author, I code each observation into one of several binary categories that corresponds to the overall subject matter of the publication.
-The goal is to create a coding scheme by subject that is sufficiently broad but also specific enough that is helpful in differentiating the overall main point of each publication.
-The following coding scheme is therefore devised and implemented (all variables are binary, taking on the value of 1 if the publication corresponds to the general definition of the variable, and 0 otherwise): \singlespace
+In the first level of analysis, I code each observation into one of several binary categories that corresponds to the overall subject of the publication.
+The goal is to create a coding scheme that captures a concise number of topical descriptions that are still sufficiently broad enough to reveal variation in the topic addressed by retired military authors.
+I develop a total of six topical variables, and each observation takes on the value of 1 if the publication corresponds to the general definition of the variable, and 0 otherwise.
 
-* **War_Ops**- if the piece advocates for or denounces a particular course of action directly related to operational or war fighting decisions of a clear military nature.
-Examples include: the wisdom of troop increases, the Surge in Iraq, whether to bomb Serbia, or even the role of the prison at Guantanamo.
+These topical variables are: warfighting and operations, if the piece advocates for or denounces a particular course of action directly related to operational or war fighting decisions of a clear military nature; troop support and institutional culture, if the publication generally highlights the work and sacrifice of members of the armed forces, defends the work of specific members of the Armed Forces, or makes a call for action to improve a service's institutional culture; strategic advice, if the piece is generally written to offer some sort of strategic advice for the nation, to include the wisdom of pursuing particular alliances or agreements, or adopting a specific weapons platform or military budget; social policy directly related to the military, if the piece advocates for or denounces a particular social policy that _directly influences_ the armed forces (Examples include: debates over keeping or removing the "Don't Ask, Don't Tell" policy, opening up combat roles to women, the role of Transgender members of the armed forces); domestic policy, if the piece advocates for or denounces a position on a policy that _does not directly_ influence the armed forces in its day to day activities (Examples include: support for the second amendment, immigration policy, and whether confederate statues should be taken down or left up); figures by name, if the piece advocates for or denounces a particular position held by a political administration, or _if the main point of the piece_ is to explicitly defend or attack a political figure by name (Examples: endorsing a Presidential candidate for office, or for attacking a President or member of Congress for a decision that was made).
 
-* **Trp_Clt**- if the piece generally showcases or highlights the work and sacrifice of members of the armed forces; to defend the work of specific members of the Armed Forces; or to call for policing up the institutional cultures of the services.
-Examples include: remembering our fallen soldiers on Memorial Day; improving the institutional culture of the Army or Navy; advocating on behalf of or remembering a specific flag officer.
+Separating opinion-editorial publications into one of several distinct categories is an imperfect process, and requires as objective a reading of the publication as possible.
+To aid the coding process, I identify the thesis statement of each opinion piece, which in most cases is identifiable in either the lead or concluding paragraph of the opinion piece.
+In most cases, identifying and re-reading the thesis of each op-ed makes it possible to reasonably classify the publication by topic.
+Two additional points, however, are worth emphasizing here.
 
-* **Str_Adv**- if the piece is generally written to offer some sort of strategic advice for the nation, to include the wisdom of pursuing particular alliances or agreements, or adopting a specific weapons platform or military budget.
-Examples include: why the US needs to strengthen NATO, build ties with China, increase defense spending, or not retire the A-10 combat plane.
-
-* **MilPSoc**- if the piece advocates for or denounces a particular social policy that _directly influences_ the armed forces.
-Examples include: debates over keeping or removing the "Don't Ask, Don't Tell" policy, opening up combat roles to women, the role of Transgender members of the armed forces.
-
-* **Dom_P**- if the piece advocates for or denounces a position on a policy that _does not directly_ influence the armed forces in its day to day activities.
-Examples include: support for the second amendment, immigration policy, and whether confederate statues should be taken down or left up.
-
-* **Fig_Name**- if the piece advocates for or denounces a particular position held by a political administration, or _if the main point of the piece_ is to explicitly defend or attack a political figure by name.
-Examples: endorsing a Presidential candidate for office, or for attacking a President or member of Congress for a decision that was made.
-
-\doublespace
-Separating opinion-editorial publications into one of several distinct categories is an imperfect process.
-To aid the coding process, I first identify the thesis statement of each opinion piece, which in almost every case, makes possible a reasonably clear classification of the publication.
-Two points are worth mentioning.
-First, I attempt to keep in mind the fact that opinion pieces are intended to argue for something, and thus are, to an extent, designed to be provocative.
-With this in mind, I did not automatically code pieces that generally advocated for or cautioned the President in the **Admin_Fig_Name** category.
-Second, there were some instances in which I had to choose between placing a publication into one of two categories, say, between **Warfighting_Ops** and **Strat_Advice**, but very few if any instances in which I had to choose between categories that differed greatly, such as **TroopSptCulture** and **Admin_Fig_Name**.
-Table 1 reveals the breakdown of the data by article type.
+First, opinion pieces are intended to argue for something, and thus are, to an extent, designed to be at least mildy provocative.
+With this in mind, I did not necessarily code pieces that called for a political figure or instituton to take a specific action into the category of _figures by name_.
+Again, the identification of the publication's thesis statement aided significantly in identifying the main thrust of the argument, and thus, in classifying the argument by topic.
+Second, there were some instances in which I had to choose between placing a publication into one of two categories.
+However, in instances of choosing between two categories, the choice to classify between categories was usually between categories that were reasonably close.
+For example, there were a few instances of choosing to classify a publication between the warighting and operations category and the strategic advice category, but no instances of choosing between the warfighting and operations category and the domestic policy category.
+Table 3 reveals the breakdown of all examined publications by article type.
 
@@
 
-```{r table2, fig.align="center", include=TRUE, echo=FALSE}
+```{r table3, fig.align="center", include=TRUE, echo=FALSE}
 countbytype %>%
@@
 
-Table 2 reveals that an overwhelming majority (289 of 334, or 86.5%) of the total number of opinion pieces written since 1979 in the major newspapers examined can be classified according to one of the following three categories: **War_Ops**, **Trp_Clt**, and **Str_Adv**.
+Table 3 reveals that an overwhelming majority (289 of 334, or 86.5%) of the total number of opinion pieces written since 1979 in the major newspapers examined can be classified according to one of the following three categories: **War_Ops**, **Trp_Clt**, and **Str_Adv**.
 It is reasonable to conclude that these op-eds written are appropriate from the perspective of civil-military relations in that they address topics which American citizens can reasonably expect retired military officers to bring their expertise to bear upon.
@@
 
-```{r table3, fig.align = "center", include=TRUE, echo=FALSE}
+```{r table4, fig.align = "center", include=TRUE, echo=FALSE}
   rem_obs %>%
@@
 
-```{r table4, include=TRUE, echo=FALSE}
+```{r table5, include=TRUE, echo=FALSE}
 secondlayer_obs %>%

### 2020-08-29 code/Submissionv3.Rmd
commit

@@
 library("dplyr")
+library("reshape2")
 ```
@@
 They are familiar with issues ranging from how wars should be best fought, to future security threats looming on the horizon, to fixing the institutional culture of a military service, and more.
-
 It is reasonable, therefore, to conclude that a professional opinion piece written by a military officer would take on three qualities.
@@
 The credibility of the American military as an institution has increased during the past fifty years.
-In the late 1970s, likely as a result of public opinion over the Vietnam War, 54% of Americans indicated that they trusted the military "a great deal or quite a lot," whereas in 2019, 73% of Americas expressed the same sentiments [@gallup_polls_confidence_2007].
-Between these time periods, there have been some shorter and localized changes, but by and large, the increase that has occurred has been gradual and steady.
 Today in fact, relative to other government institutions, the American public trusts the military to a remarkably high degree.
 For example, according to a recent Pew Research Center poll, 80% of the American public expects the military to act in the best interests of the American public, compared to tho percentage who believe the same about business leaders (45%), the media (40%), and elected officials (25%) [@johnson_trust_2018].
+
 Military leaders are, rightfully so, cognizant of the benefits of holding the American public's trust.
@@
 The second and more influential societal change that has occurred over the past fifty years in America is the rise of a polarized political landscape.
-Though this is not the first time in America's history that political polarization has been a dominant feature of domestic politics, multiple scholars have argued that polarization has worsened in recent decades [@hare_polarization_2014; @bonica_mapping_2014]], and some have argued that polarization is at the worse level since the end of the American Civil War [@hare_polarization_2014].
+Though this is not the first time in America's history that political polarization has been a dominant feature of domestic politics, multiple scholars have argued that polarization has worsened in recent decades [@hare_polarization_2014; @bonica_mapping_2014], and some have argued that polarization is at the worse level since the end of the American Civil War [@hare_polarization_2014].
 Although scholars have offered differing explanations for the cause of the political polarization evident today, ranging from concerns over race and religion [@abramowitz_great_2018], to the size and role of government [@sides_two_2015], to fierce policy disagreements over policy issues [@mann_its_2012], there is a fair amount of common ground among scholars regarding how polarization has impacted contemporary domestic politics.
@@
 
-```{r, include=FALSE, echo=FALSE}
+```{r figure one, include=FALSE, echo=FALSE}
 problem_dag <- dagify(
@@
     \hline
-    \textbf{Characteristics of Increasing Political Polarization} & \textbf{Description} & \textbf{Resulting Strain on Military Professionalism}  \\ \hline
-    \textbf{Conflict Extension} & Less politically-neutral space exists as partisan disagreements subsume a greater number of issues & Military Professionals face an increasingly difficult prospect of remaining apolitical.
+    \textbf{Features of Increasing Political Polarization} & \textbf{Associated Mechanism} & \textbf{Potential Impact on Military Professionalism}  \\ \hline
+    \textbf{Conflict Extension} & Less politically-neutral space exists as partisan disagreements subsume a greater number of issues.
+It becomes increasingly difficult to remain "out of politics." & Military Professionals face an increasingly difficult prospect of remaining apolitical.
 \\ \hline
-    \textbf{Negative Partisanship} & Strong feelings of personal dislike and disdain for those who hold to opposing political views and who are on the "other side" & Greater willingness to confront or take issue with those with opposing views\\ \hline
-    \textbf{Core and dearly-held issues are increasingly at stake} & The salience of one's identity and belonging to a particular group is raised; Belief that moral right and wrong are at stake & Changes in one's concept of loyalty \\ \hline
+    \textbf{Negative Partisanship} & Strong feelings of personal dislike and disdain arise against those who hold to opposing political views.
+There is a belief that those on the "other side" are greatly wrong and flawed.
+& There is a greater willingness to confront or take issue with those who hold opposing views.\\ \hline
+    \textbf{Core and dearly-held issues are increasingly at stake} & As the issues to which one holds most dear are at stake, the salience of one's identity and sense of belonging to a particular group is raised.
+Associated with this phenomenon is a belief that not just preference, but moral right and wrong, are at stake.
+& Military professionals face potential changes in one's concept and notion of loyalty.
+There is a sense that failing to take action over a certain issue is equivalent to failing to uphold a sense of loyalty to an important group, idea, or concept.\\ \hline
 \end{tabular}
@@
 Indeed, this set of conditions comprise a very dangerous world for the military institution and the military professional.
+
 # Methodological Approach, Data, and Hypotheses
 
-I now aim to see whether the discourse of retired military officers has changed over time, specifically since the 1970s, as polarization has worsened dramatically.
-To answer this question, I assemble and examine a data set that consists of opinion-editorial publications authored by retired military officers in major US newspapers.
-My methodological approach involves two fundamental steps.
-First, I determine what topics retired military officers have been writing about over time.
-Second, I apply a normative lens to determine if there have been shifts in the tone and tenor of the pieces authored by these officers, particularly regarding the professional aspects of the military previously discussed (apolitical in nature, non-partisan, relying on relevant military expertise, etc.).
-In essense, I want to answer the questions, what are these retired military officers writing about, and through the lens of professionalism, should they be saying what they are saying?
-Is there any indication that increasing political polarization is altering what and how these retired officers are communicating with the American public?
+```{r introduction, include = FALSE}
 
-To assemble the data, I use the Factiva and Nexus-Uni databases, conducting a search that focuses keywords in the byline (author field) such as "retired military officer," "retired from the Army, Navy, Marine Corps, Air Force," "former Chairman of the Joint Chiefs of Staff," "General," "Admiral," and so forth.
-I confine my search to the largest US newspapers that regularly run opinion-editorial pieces, which includes  _The Wall Street Journal_ (WSJ), _The New York Times_ (NYT), _The Washington Post_ (WaPo), _The Los Angeles Times_, and _The USA Today_.
-The time period for the data includes the period January 1, 1979 - 30 June 2020, the main limitation being that not all of the Newspaper sources are available for previous dates.
-I also open my search to include both print and online offerings to ensure that I am capturing any opinion pieces published digitally.
+# Read in the Data
 
-The sources that I rely upon provide some relative advantages, but they are not without limitations.
-The primary advantage is that these sources have some longevity; they are not modern blogging sites or online-only offerings.
-As a result, I can truly examine a larger period of time and thus, examine whether how political polarization has impacted the way in which military officers engage with the American Public.
-Simultaneously, though, this presents a disadvantage.
-Retired military officers engage with the American public through a host of mediums, to include social media, blogging sites and cable television, neither of which I am capturing in my data set.
-It is also true that over time, social media has taken over as the primary means in which a majority of Americans receive their news [@martin_how_2018].
-Print newspapers still exist, but overall circulation is still down today compared to earlier years, even after factoring in the expanse of digital options available^[As a means of rough comparison, estimated total daily newspaper circulation in the US in 2000 was 62 million and 54 million for weekday and weekend editions, respectively, while in 2018, the the same figures were 29 million and 31 million [@barthel_leading_2019]].
-Nonetheless, I contend that the data I am able to assemble will still be informative.
-Publishing opinion-editorial pieces, especially in major, competitive newspapers, is a highly competitive process.
-For the newspapers that I examine in this study, it is not uncommon to see opinion pieces published by figures that include heads of state, members of congress, governmental figures, and leaders of industry and academia.
-Additionally, opinion pieces are deliberately authored for the purpose of informing and persuading the public on a salient or timely topic.
-They are different from interviews or commentary that may be asked by a questioner because the author must take time to put forth his or her position in writing.
-The data therefore allows the researcher to truly ascertain what the specific views and arguments of an author were on a given topic or policy.
-
-Of note, when assembling the data, some authors are explicitly identified in some of their publications as retired military officers, but not in others.^[This was true of several individuals, including Alexander Haig, a retired Army General who later served as Secretary of State under Ronald Reagan;  Robert McFarlane, a retired Marine Corps Lieutenant Colonel who later served as the National Security Adviser to President Reagan, Andrew Bacevich, a retired Army Colonel who then became a prominent historian and author; and Michael Hayden, a retired Air Force General who served as director of the National Security Agency and the Central Intelligence Agency].
-I conduct a separate search for all opinion pieces penned by these authors, regardless of whether they were identified as a retired military officer or not, concluding that it is best do so for coding purposes.
-
-Ultimately, the data set contains 334 opinion-editorial publications, penned by 178 different authors or author teams.^[There are several co-authored pieces.
-For example, where General (Retired) David Petraeus wrote one publication and another one was written by General (Retired) David Petraeus and Michael O'Hanlon, I counted these as two different authors for coding purposes.] 10 of these pieces are written unanimously, with nods by the authors to their previous service.
-From a macro perspective, consider for a moment all that the date range includes with respect to the US military.
-Indeed, numerous and substantive events, debates, and changes have occurred in this time period for retired military officers to weigh in on.
-In terms of external threats and wars, this period includes the final years of the Cold War against the Soviet Union, the first war in Iraq, operations in the Balkans, and the post-9/11 wars in Iraq and Afghanistan, as well as a host of smaller but important military operations in Grenada, Panama, Somalia, and Haiti.
-The military has also undergone significant social and institutional changes in the time period examined.
-In the late 1970s, the military had ended the draft only years before following the conclusion of the Vietnam War.
-The debates over "Don't Ask, Don't Tell" and policies regarding women in the combat arms are also included in the data, as are numerous discussions of military budgets, weapons systems, and alliances.
-
-Attempting to reconcile the fact that the military as an institution prides itself on being a non-partisan and apolitical institution, that newspaper circulations (even after accounting for digital offerings) are substantially down over the past 40 years, and that the US political environment is increasingly polarized,  I adopt the following hypotheses:
-
-_$H_1$: The majority of op-eds published by retired military officers over the time period examine will be apolitical and non-partisan in tone._
-
-_$H_2$: Because overall newspaper consumption has dropped in the past 40 years, there will be fewer opinion pieces over time authored by retired military officers._
-
-_$H_3$: Given that polarization has become increasingly pronounced in the United States over the time period examined, there will be an increase in recent years in the number of op-eds which reveal political and partisan sentiments, indicating a drop in military professionalism._
-
-# First Layer Analysis: Topics Addressed by Retired Military Officers
-
-In the first level of analysis, I code each observation into one of several binary categories that corresponds to the overall subject of the publication.
-The goal is to create a coding scheme that captures a concise number of topical descriptions that are still sufficiently broad enough to reveal variation in the topic addressed by retired military authors.
-I develop a total of six topical variables, and each observation takes on the value of 1 if the publication corresponds to the general definition of the variable, and 0 otherwise.
-
-These topical variables are: warfighting and operations, if the piece advocates for or denounces a particular course of action directly related to operational or war fighting decisions of a clear military nature; troop support and institutional culture, if the publication generally highlights the work and sacrifice of members of the armed forces, defends the work of specific members of the Armed Forces, or makes a call for action to improve a service's institutional culture; strategic advice, if the piece is generally written to offer some sort of strategic advice for the nation, to include the wisdom of pursuing particular alliances or agreements, or adopting a specific weapons platform or military budget; social policy directly related to the military, if the piece advocates for or denounces a particular social policy that _directly influences_ the armed forces (Examples include: debates over keeping or removing the "Don't Ask, Don't Tell" policy, opening up combat roles to women, the role of Transgender members of the armed forces); domestic policy, if the piece advocates for or denounces a position on a policy that _does not directly_ influence the armed forces in its day to day activities (Examples include: support for the second amendment, immigration policy, and whether confederate statues should be taken down or left up); figures by name, if the piece advocates for or denounces a particular position held by a political administration, or _if the main point of the piece_ is to explicitly defend or attack a political figure by name (Examples: endorsing a Presidential candidate for office, or for attacking a President or member of Congress for a decision that was made).
-
-Separating opinion-editorial publications into one of several distinct categories is an imperfect process, and requires as objective a reading of the publication as possible.
-To aid the coding process, I identify the thesis statement of each opinion piece, which in most cases is identifiable in either the lead or concluding paragraph of the opinion piece.
-In most cases, identifying and re-reading the thesis of each op-ed makes it possible to reasonably classify the publication by topic.
-Two additional points, however, are worth emphasizing here.
-
-First, opinion pieces are intended to argue for something, and thus are, to an extent, designed to be at least mildy provocative.
-With this in mind, I did not necessarily code pieces that called for a political figure or instituton to take a specific action into the category of _figures by name_.
-Again, the identification of the publication's thesis statement aided significantly in identifying the main thrust of the argument, and thus, in classifying the argument by topic.
-Second, there were some instances in which I had to choose between placing a publication into one of two categories.
-However, in instances of choosing between two categories, the choice to classify between categories was usually between categories that were reasonably close.
-For example, there were a few instances of choosing to classify a publication between the warighting and operations category and the strategic advice category, but no instances of choosing between the warfighting and operations category and the domestic policy category.
-Table 3 reveals the breakdown of all examined publications by article type.
-
-```{r summary, include = FALSE}
-# Read in the Data
 dataset <- read_csv(here("data", "MilEditorials10Aug.csv"))
 dataset %>% print()
+# Only read in the first 334 rows - is there a better way to get rid of N/A data?
 # Replace NA with 0 in the data for variables War_Ops thru Fig_Nme
-adjdataset <- dataset %>%
+adjdataset <- dataset[1:334,] %>%
   mutate_at(
@@
 
-#adjust the data to account for polarization by creating a new dataset called polaradjdataset
-#create variables that capture ideological distance in each the House and Senate along
-#each of the two dimensions according to Poole and Rosenthal's data
-polaradjdataset <-adjdataset %>% mutate (
-  House_distance_dim1=abs(House_Dem_mean_nom1-House_Rep_mean_nom1),
-  Sen_distance_dim1=abs(Sen_Dem_mean_nom1-Sen_Rep_mean_nom1),
-  House_distance_dim2=abs(House_Dem_mean_nom2-House_Rep_mean_nom2),
-  Sen_distance_dim2=abs(Sen_Dem_mean_nom2-Sen_Rep_mean_nom2)
-) %>% view()
+```
 
+Ultimately, I aim to see whether the discourse of retired military officers has changed over time, specifically since the 1970s, as polarization has worsened dramatically.
+To answer this question, I assemble and examine a data set that consists of opinion-editorial publications authored by retired military officers in major US newspapers.
+My methodological approach involves two fundamental steps.
+First, I determine what topics retired military officers have been writing about over time.
+Second, I apply a normative lens to determine if there have been shifts in the tone and tenor of the pieces authored by these officers, particularly regarding the professional aspects of the military previously discussed (apolitical in nature, non-partisan, relying on relevant military expertise, etc.).
+In essense, I want to answer the questions, what are these retired military officers writing about, and through the lens of professionalism, should they be saying what they are saying?
+Is there any indication that increasing political polarization is altering what and how these retired officers are communicating with the American public?
+
+To assemble the data, I use the Factiva and Nexus-Uni databases, conducting a search that focuses keywords in the byline (author field) such as "retired military officer," "retired from the Army, Navy, Marine Corps, Air Force," "former Chairman of the Joint Chiefs of Staff," "General," "Admiral," and so forth.
+I confine my search to the largest US newspapers that regularly run opinion-editorial pieces, which includes  _The Wall Street Journal_ (WSJ), _The New York Times_ (NYT), _The Washington Post_ (WaPo), _The Los Angeles Times_, and _The USA Today_.
+The time period for the data includes the period January 1, 1979 - 30 June 2020, the main limitation being that not all of the Newspaper sources are available for previous dates.
+I also open my search to include both print and online offerings in order to ensure that I am capturing any opinion pieces published digitally.
+
+The data set contains 334 opinion-editorial publications, penned by 178 different authors or author teams.^[There are several co-authored pieces.
+For example, where General (Retired) David Petraeus wrote one publication and another one was written by General (Retired) David Petraeus and Michael O'Hanlon, I counted these as two different authors for coding purposes.] Ten publications were written anonymously, with at best general recognition by the authors to their previous military service and/or rank.
+From a macro perspective, consider for a moment all that the date range includes with respect to the US military.
+Indeed, numerous and substantive events, debates, and changes have occurred in this time period for retired military officers to weigh in on.
+In terms of external threats and wars, this period includes the final years of the Cold War against the Soviet Union, the first war in Iraq, operations in the Balkans, and the post-9/11 wars in Iraq and Afghanistan, as well as a host of smaller but important military operations in Grenada, Panama, Somalia, and Haiti.
+The military has also undergone significant social and institutional changes in the time period examined.
+In the late 1970s, the military had ended the draft only years before following the conclusion of the Vietnam War.
+The debates over "Don't Ask, Don't Tell" and policies regarding women in the combat arms are also included in the data, as are numerous discussions of military budgets, weapons systems, and alliances.
+It is also interesting to note, as shown in Figure Two, below, that retired military officers have increasingly written and successfully published opinion-editorial pieces in the sources examined over the past forty years.
+
+```{r figure two, fig.height=4, fig.width=5, fig.align="center", include=TRUE, echo=FALSE}
+
+#Need to adjust this section depending on the number of observations I have
+# Also need to figure out how to code this properly.  There has to be a better way. 
+# Adjust using the geom_text functions to label each bar properly.
+
+datatograph <- data.frame(
+  decade=c("1979-1988", "1989-1998", "1999-2008", "2009-2018", "2019-Present" ), numcount=c(13, 37, 109, 152, 23)
+)
+ggplot(datatograph, aes(x =decade, y =numcount)) + geom_bar(stat = "identity", width=.2) + geom_text(aes(label=numcount), position=position_stack(vjust=1.08))+
+  labs(
+    x = "Time Interval",
+    y = "Number of Publications",
+    color = NULL,
+    title = "Figure Two: Op-Eds Written By Retired Military Officers",
+    subtitle = "Published in WSJ, NY Times, Wa Post, USA Today, and LA Times"
+  )
 ```
+
+There are both advantages and disadvantages to confining my data to select US newspapers, as I have.
+The primary advantage is that these sources offer longevity across a time period of forty-one years; in doing so, they provide the researcher with an opportunity to detect patterns across several decades.
+The primary disadvantage is that by limiting myself to only the five newspapers I examine, I am not including a large number of modern social media and blogging sites, contemporary webpages, and other digital offerings that presumably showcase the opinions of retired military officers.
+In fact, it is likely that more opinion pieces are available through these other types of media outlets.
+Is this a serious limitation?
+
+After all, it is true that over time, social media has taken over as the primary means in which a majority of Americans receive their news [@martin_how_2018].
+Print newspapers still exist, but overall circulation is still down today compared to earlier years, even after factoring in the expanse of digital options available today^[As a means of rough comparison, estimated total daily newspaper circulation in the US in 2000 was 62 million and 54 million for weekday and weekend editions, respectively, while in 2018, the the same figures were 29 million and 31 million [@barthel_leading_2019]].
+But these trends do not seem to impact the story that Figure Two reveals so clearly, that retired military officers are publishing their opinions in at least the newspapers examined for this paper at an increasingly greater frequency than in previous decades.
+
+I contend that the data I am able to assemble will still be very informative for several reasons.
+First, publishing opinion-editorial pieces, especially in a major, competitive newspaper such as the ones from which I gather the data, is a highly competitive process.
+In the newspapers that I examine in this study, for instance, it is not uncommon to see opinion pieces regularly published by heads of state, members of congress, governmental figures, and leaders of industry and academia.
+Additionally, opinion pieces are deliberately authored for the purpose of informing and persuading the public on a salient or timely topic.
+They are different from interviews or commentary that may be asked by a questioner because the author must take time to put forth his or her position in writing.
+Limiting the data to publications in the five major US newspapers allows the researcher to truly examine the views and arguments an author intended to convey to a significant swath of the American Public, as opposed to a niche group or smaller following, on a given topic or policy.
+
+Of note, when assembling the data, some authors are explicitly identified in some of their publications as retired military officers, but not in others.^[This was true of several individuals, including Alexander Haig, a retired Army General who later served as Secretary of State under Ronald Reagan;  Robert McFarlane, a retired Marine Corps Lieutenant Colonel who later served as the National Security Adviser to President Reagan, Andrew Bacevich, a retired Army Colonel who then became a prominent historian and author; and Michael Hayden, a retired Air Force General who served as director of the National Security Agency and the Central Intelligence Agency].
+I conduct a separate search for all opinion pieces penned by these authors, regardless of whether they were identified as a retired military officer or not, concluding that it is best do so for coding purposes.
+
+Attempting to reconcile the fact that the military as an institution prides itself on being a non-partisan and apolitical institution, that newspaper circulations (even after accounting for digital offerings) are substantially down over the past 40 years, and that the US political environment is increasingly polarized,  I adopt the two following hypotheses:
+
+_$H_1$: The majority of op-eds published by retired military officers over the time period examine will be apolitical and non-partisan in tone._
+
+_$H_3$: Given that polarization has become increasingly pronounced in the United States over the time period examined, there will be an increase in recent years in the number of op-eds which reveal political and partisan sentiments, indicating a potential drop in military professionalism across the ranks of retired military officers across the past several decades._
+
+# First Layer Analysis: Topics Addressed by Retired Military Officers
+
+In the first level of analysis, I code each observation into one of several binary categories that corresponds to the overall topic or subject of the publication.
+The goal is to create a coding scheme that captures a concise number of topical descriptions that are still sufficiently broad enough to reveal variation in the topic addressed by retired military authors.
+I develop a total of six topical variables, and each observation takes on the value of 1 if the publication corresponds to the general definition of the variable, and 0 otherwise.
+
+These topical variables are: warfighting and operations, if the piece advocates for or denounces a particular course of action directly related to operational or war fighting decisions of a clear military nature; troop support and institutional culture, if the publication generally highlights the work and sacrifice of members of the armed forces, defends the work of specific members of the Armed Forces, or makes a call for action to improve a service's institutional culture; strategic advice, if the piece is generally written to offer some sort of strategic advice for the nation, to include the wisdom of pursuing particular alliances or agreements, or adopting a specific weapons platform or military budget; social policy directly related to the military, if the piece advocates for or denounces a particular social policy that _directly influences_ the armed forces (Examples include: debates over keeping or removing the "Don't Ask, Don't Tell" policy, opening up combat roles to women, the role of Transgender members of the armed forces); domestic policy, if the piece advocates for or denounces a position on a policy that _does not directly_ influence the armed forces in its day to day activities (Examples include: support for the second amendment, immigration policy, and whether confederate statues should be taken down or left up); figures by name, if the piece advocates for or denounces a particular position held by a political administration, or _if the main point of the piece_ is to explicitly defend or attack a political figure by name (Examples: endorsing a Presidential candidate for office, or for attacking a President or member of Congress for a decision that was made).
+
+Separating opinion-editorial publications into one of several distinct categories is an imperfect process, and requires as objective a reading of the publication as possible.
+To aid the coding process, I identify the thesis statement of each opinion piece, which in most cases is identifiable in either the lead or concluding paragraph of the opinion piece.
+In most cases, identifying and re-reading the thesis of each op-ed makes it possible to reasonably classify the publication by topic.
+Two additional points, however, are worth emphasizing here.
+
+First, opinion pieces are intended to argue for something, and thus are, to an extent, designed to be at least mildy provocative.
+With this in mind, I did not necessarily code pieces that called for a political figure or instituton to take a specific action into the category of _figures by name_.
+Again, the identification of the publication's thesis statement aided significantly in identifying the main thrust of the argument, and thus, in classifying the argument by topic.
+Second, there were some instances in which I had to choose between placing a publication into one of two categories.
+However, in instances of choosing between two categories, the choice to classify between categories was usually between categories that were reasonably close.
+For example, there were a few instances of choosing to classify a publication between the warighting and operations category and the strategic advice category, but no instances of choosing between the warfighting and operations category and the domestic policy category.
+Table 3 reveals the breakdown of all examined publications by article type.
 
@@
 ```
-
-Table 3 reveals that an overwhelming majority (289 of 334, or 86.5%) of the total number of opinion pieces written since 1979 in the major newspapers examined can be classified according to one of the following three categories: **War_Ops**, **Trp_Clt**, and **Str_Adv**.
-It is reasonable to conclude that these op-eds written are appropriate from the perspective of civil-military relations in that they address topics which American citizens can reasonably expect retired military officers to bring their expertise to bear upon.
-The remaining 45 opinion pieces (13.5% of the total) fall into one of the other three categories: **MilPSoc**, **Dom_P**, and **Fig_Nme** and require further examination.
-They are not necessarily inappropriate, but they likely address topics that may be beyond reasonable military expertise or which reveal partisan preferences.
-This second layer of analysis will help expose whether there is any indication that professionalism as a norm among retired military officers appears to be weakening.
+An overwhelming majority (289 of 334, or 86.5%) of the total number of opinion pieces in the dataset fall into the categories of warfighting and operations, troop support and institutional culture, and strategic advice.
+Moreover, I conclude that these observations pose no problem for norms of military professionalism, as they are written on topics that I contend are clearly related to relevant military experience.
+The remaining 45 publications (13.5% of the observations) fall under the other three categories of social policy directly related to the military, domestic policy, and figure by name.
+These obsercations require a second and more-thorough examination to determine whether any of these observations indicate a drop in the norm of military professionalism.
+This second layer of analysis will allow us to examine these instances in greater detail.
 
@@
 
-Of the remaining observations, two are removed upon immediate inspection.
-One was written by an unknown author and is entitled, "Fists Raised at West Point," and was published in _The Washington Post_ on May 13, 2016.
-It is removed because the article is actually a compendium of two separate letters sent to the editor by two different authors, one arguing for and the other against female West Point graduates who raised their fists upon graduating, which ignited some controversy about soon-to-be Army lieutenants purportedly making a symbol associated with the "Black Lives Matter" movement [@noauthor_fists_2016].
-Because the two separate letters taken together do not take a firm position on the subject, but rather show a difference in opinion about a specific event happening in 2016, I remove the observation.
-The second piece I remove was written by former Democratic Congressman John Murtha, who was also a retired Marine Corps officer.
-As a politician, the American public can and should expect him to write in the capacity of a politician, and therefore, to make partisan arguments.
-Not wanting to skew the remaining data, I remove his publication for the remaining analysis.
-Thus, the second level of analysis proceeds on a total of 43 observations, which are listed and printed in chronological publishing order, in Table 2.
-
-\singlespace
-
-```{r remaining data, include = FALSE}
-
-# Show all articles and authors who published articles on Social Policy, Political Parties, and Administrations/Political Figures
-rem_obs <-adjdataset %>% arrange(PubYr) %>% 
-  filter(MilPSoc==1 | Dom_P==1 | Fig_Nme==1 )  %>% 
-  filter(Author != "UNK") %>% filter(Author != "Murtha, John") %>% 
-  select (Author, Title, Source, PubYr, MilPSoc, Dom_P, Fig_Nme) %>% 
-  view()
-```
-
-```{r table4, fig.align = "center", include=TRUE, echo=FALSE}
-  rem_obs %>%
-  kable(format="latex", caption="Remaining Observations", longtable=T) %>%
-  kable_styling(font_size=10, latex_options="striped", repeat_header_continued=T) %>%
-  column_spec(1, width="12em") %>% 
-  column_spec(2, width="16em") %>% 
-  column_spec(3:4, width="2.5em") %>% 
-  column_spec(5:7, width="3em") %>% 
-  row_spec(0, bold=T)
-```
-
-```{r code, include=FALSE, echo=FALSE}
-# Create a Variable Called case-count that counts the number of times by publishing year an article published 
-# regarding social policy, domestic policy, or political figures/administrations occurs
-adj2dataset <- adjdataset %>% group_by(PubYr) %>% 
-  mutate(
-    case_count=sum(MilPSoc ==1 | Dom_P==1 | Fig_Nme==1), 
-    yr_total_count=sum(War_Ops==1 | Trp_Clt==1| Str_Adv==1 | MilPSoc ==1 | Dom_P==1 | Fig_Nme==1),
-    prop_yr=(case_count/yr_total_count)
-  ) %>% view()
-```
-\doublespace
-
-A couple of interesting observations are worth noting about the data in Table 3.
-First, one can quickly observe that there appear to be three distinct clusters of observations with respect to publishing year.
-The first occurs in the time frame between roughly 1998-1999, the second between 2004-2007, and the third between 2016-2019.
-
-In the first cluster, the publications refer to multiple disagreements certain retired military officers had about how President Clinton was planning or using military force.
-One observation, that written by retired Marine Corps General Carl Mundy, discussed implications of changing the policy of "Don't Ask, Don't Tell," which was in place at the time, while another observation, that written by retired Army General Barry McCaffrey, refers to domestic drug policy in the United States - a perfectly acceptable topic given that he, at the time, was serving as President Clinton's Director of the Office of National Drug Control Policy.
-The second cluster refers to pieces penned around the so-called "Revolt of the Generals," referring to sharp disagreements that entered the public arena between multiple US generals and admirals and President George W.
-Bush's Secretary of Defense, Donald Rumsfeld.
-Disagreement about Rumsfeld's leadership, particularly regarding the planning and execution of the early years of the war in Iraq, were central during this period of time.
-Retired Army General Tommy Franks penned two observations that essentially provided public support for the President at the time, George W.
-Bush, and Retired Army Major General Paul Eaton, who was a particularly vocal critic of Rumsfeld, authored two of the pieces published in this time period.
-Finally, the 2016-2019 cluster is indicative mainly of concerns over the candidate and later President Donald Trump.
-Retired Air Force General Michael Hayden, Navy Admirals Mike Mullen and William McRaven, and Army General Stanley McChrystal are among the retired officers penning pieces during this time period.
-
-Evaluating a conclusive level of appropriateness of the opinion pieces listed in Table 2 cannot be definitively solved, precisely because it matters the perspective from which one views how military officers should interact with the American public, especially when attempting to shape public opinion.
-However, it is possible to determine whether there has been deviation from non-partisan and apolitical norms.
+Conclusively determining whether an observation failed to uphold military professionalism is once again an imperfect and somewhat subjective evaluation.
+However, the work of academic theorists and decades of military norms make clear that at a minimum, military professionalism emphasizes behavior by military officers that is relevant to one's military expertise, as well as absent of partisan expression or preferences.
 Recall that Retired Army General Douglas MacArthur cautioned the cadets at West Point in 1962 about involving themselves in matters that do not require a military solution (in MacArthur's words, of "politics grown to corrupt," "morals grown too low," and "whether our personal liberties are as firm and complete as they should be") [@macarthur_duty_1962], and that more recently, in 2016, Retired Army General Martin Dempsey lamented the participation of retired generals in the political process by endorsing political candidates and appearing at the conventions of political parties:
@@
 
-Using these ideals as guidelines allows us to conduct a second level of analysis, with emphasis on the two characteristics of topic (is the author writing about a reasonable military topic) and political leaning (is the author expressly attacking or supporting specific political figures in a political way, from which one can reasonable infer the officer's political or partisan bent)?
-
-Using these characteristics as guidelines, I therefore develop two new variables.
-The first is **Result**, which can take on one of three values: the following values: **Permissible,** if the reader cannot clearly draw conclusions regarding the author's political leaning or partisan preferences; or if the topic addressed in the publication, after a second reading, clearly falls within military expertise; **Questionable,** if the reader can draw some but not complete conclusions regarding the author's political leaning or partisan preferences, or if the topic addressed in the publication is questionably an area of reasonable military expertise; and **Inappropriate,** if the reader can draw clear conclusions regarding the author's partisan preferences or political leanings, or if the author addresses a topic that clearly falls outside of traditional military expertise.
-The second new variable I develop is **Justification.** In this variable, a brief explanation is placed which addresses why an observation is assigned a particular value of **Result.**
-
-The results of this second level of analysis are presented in Table 3.
-Note that several publications are deemed to take on values of permissible under the variable **Result**.
-Importantly, three of these pieces occurred during the first previously identified cluster, 1998-1999.
-No pieces from the second cluster, 2004-2007 have been deemed permissible, whereas in the third cluster, 2016-2019, three take on the value of permissible.
-Thus, we now have a picture of how prevalent the publication is of opinion pieces authored by retired military officers in major newspapers which have failed to uphold the norm of non-partisanship and apolitical position.
-
-\singlespace
+Relying on these two central ideals of military professionalism, relevant military expertise and non-partisan/apolitical expression, allows the researcher to ask two important questions: first, is the author addressing a topic related and applicable to the officer's military expertise, and second, is the author making an argument without clearly conveying or expressing political or partisan preferences?
+If the answer to either of these two questions is "no," then the publication might violate the concept of military professionalism.
+I therefore develop two new variables, _questionable professionalism_ and _justification_.
+The first variable is binary and codes as '1' the observations which may violate the concept of military professionalism, and the second variable briefly captures the basis for a possible violation, i.e.
+whether it is a violation of topic or of expression of political sentiment.
+The results of this analysis on the remaining 45 observations are displayed in Table 4, below.
+Several observations are worth noting.
 
 ```{r second level analysis, include = FALSE}
+
+# Show all articles and authors who published articles on Social Policy, Political Parties, and Administrations/Political Figures
+rem_obs <-adjdataset %>% arrange(PubYr) %>% 
+  filter(MilPSoc==1 | Dom_P==1 | Fig_Nme==1 )  %>% 
+  select (Author, Title, Source, PubYr, MilPSoc, Dom_P, Fig_Nme) %>% 
+  view()
+
+# Create a Variable Called case-count that counts the number of times by publishing year an article published 
+# regarding social policy, domestic policy, or political figures/administrations occurs
+
+adj2dataset <- adjdataset %>% group_by(PubYr) %>% 
+  mutate(
+    case_count=sum(MilPSoc ==1 | Dom_P==1 | Fig_Nme==1), 
+    yr_total_count=sum(War_Ops==1 | Trp_Clt==1| Str_Adv==1 | MilPSoc ==1 | Dom_P==1 | Fig_Nme==1),
+    prop_yr=(case_count/yr_total_count)
+  ) %>% view()
+
 # Present Results of Result and Justification
@@
   filter(MilPSoc==1 | Dom_P==1 | Fig_Nme==1 )  %>% 
-  filter(Author != "UNK") %>% filter(Author != "Murtha, John") %>% 
   select (Author, Title, Source, PubYr, Ques_Prof, Justification) %>% 
@@
 
-```{r polarization logit analysis, include=FALSE}
+\singlespace
+```{r table 4, fig.align = "center", include=TRUE, echo=FALSE}
+#Need to adjust the width and spacing of this table
+secondlayer_obs %>%
+  kable(format="latex", caption="Potentially Unprofessional Op-Ed Pieces By Retired Military Officers, 1979-2020", longtable=T) %>%
+  kable_styling(font_size=10, latex_options="striped", repeat_header_continued=T) %>% 
+  column_spec(1, width="9.5em") %>% 
+  column_spec(2, width="15em") %>% 
+  column_spec(3:4, width="2.5em") %>%
+  column_spec(5, width="5em") %>%
+  column_spec(6, width="6em") %>%
+  row_spec(0, bold=T)
+```
+\doublespace
+
+First, 15 of the 45 publications are, upon a closer reading, determined to not reflect possible violations of military professionalism.
+This is worth showcasing in order to draw attention to the fact that there are instances of retired military officers authoring opinion pieces which address the subjects of social policy related to the military, domestic policy, and figures by name while simultaneously upholding standards of military professionalism.
+However, the data reveals that in practice, such instances are relatively uncommon.
+The practical implication is simple and straightforward: if military officers decide to write about topics other than warfighting and operations, troop support and institutional culture, or strategic advice, they need to carefully craft their argument so as to not violate norms of military professionalism.
+
+Second, and resulting from the first, is that across the entire data set, instances of potential violations of military professionalism in the op-eds published by retired military officers in major US newspapers is a somewhat rare occurrence.
+Indeed, the entire count of potentially questionable publications with respect to military professionalism across the data set is just 30 of 334 total pieces, or 8.9% of the entire sample.
+Thus, ample support exists to confirm the first hypothesis, that the vast majority of opinion pieces authored by retired military officers uphold standards of military professionalism.
+
+A third and interesting finding, however, is that the first publishing year in which an opinion piece penned by a retired officer potentially failed to uphold professional standards is 1998, nearly halfway across the sample in terms of publishing year.
+Why is this the case?
+Certainly, retired military officers before 1998 had plenty of reasons to go to the press and air disagreements with political leaders, or to address important topics that may have pushed the limits of military professionalism.
+In fact, many of these officers would have been veterans of the Vietnam and Korean Wars, and perhaps, like Maxwell Taylor, an author in the sample, veterans of World War Two.
+
+Fourth, Table 4 reveals that several retired military officers wrote publications not from the perspective of a retired military officer, but from the perspective of other professional and political roles that these officers served in after retiring from the military.
+These include op-eds by: Retired Air Force Lieutenant General Claudius Watts, who wrote an op-ed regarding the advantages of all-male education while serving as the president of the Citadel; Retired Army General Barry McCaffrey, who penned an opinion piece while serving in the administration of Bill Clinton as the Director of the Office of National Drug Control Policy; and Retired Marine Corps Colonel John Murtha, who also served in the US House of Representatives.
+Each of these authors addressed topics that one should reasonably expect them to on the basis of the roles each was serving in.
+This would include, in the case of John Murtha, making obviously partisan agreements and thus, revealing political preferences.
+
+Table 3 reveals, fifth, that several publications are deemed as possibly having violated military professionalism in that they address topics that may be considered outside of the realm of proper military expertise.
+It is worth briefly commenting on several of these examples.
+In his _Washington Post_ editorial, "Home Should Not Be a War Zone," Retired Army General Stanley McChrystal addresses the tragic nature of mass shootings in America.
+He states in the article, "...as a combat veteran and proud American, I believe we need a national response to the gun violence that threatens so many of our communities" [@mcchrystal_home_2016].
+In a later article published in the _Washington Post_ in 2018 entitled "Good Riddance," McChrystal shared how his beliefs in Robert E.
+Lee had evolved over the course of his professional career, and why he now believes it is now time to discard Confederate statues and other similar symbols in America.
+Other examples include publications by Admiral (Retired) Mike Mullen and co-authors Lieutenant General (Retired) Mark Hertling and Admiral (Retired) Robert Natter, who respectively wrote the publications, "The Refugees We Need" and "Cutting Refugee Admissions Will Have Severe Consequences for the US Military."
+
+These articles are not unreasonable in that they, in the opinion of the author of this article, express reasonable thoughts and ideas to many of the nation's pressing problems, to include gun violence, the legacy of the Confederacy and our national history, and immigration policy.
+Moreover, many Americans agree with the stance these officers have taken in the articles they have published.
+One can even make a strong, perhaps even a compelling argument, why these officers might be uniquely qualified to address some of these topics, in that they possess expertise commanding large and powerful organizations charged to carry out violence (on the issue of gun control), have served in the nation's military (and know what it is like to, for example, lead servicemen and women who serve on bases named after Confederate generals), and have considerable experience serving alongside soldiers of multiple ethnicities as well as considerable experience in Iraq and Afghanistan (and thus, can speak about how immigration policy might impact the military).
+
+At the same time, however, one is also not wrong to ask whether these and other contentious and large issues are best considered to be political issues in that they are best solved by politicians, as opposed to retired military officers.
+The important point as is related to the concept of military professionalism is not that these officers should not have an opinion on these and other important topics, but whether they, as retired military officers, ought to be considered as appropriate mouthpieces on these issues by the American public.
+Interestingly, in July of 2020, General Mark Milley, the Chairman of the Joint Chiefs of Staff, was asked by US Representative Anthony Brown, D-MD, during a hearing held by the House Armed Services Committee, to "comment on the naming of Army installations after Confederate soldiers," and whether the continued use of the names of these bases "reflects the values" that the military is trying to instill in the nation's armed forces [@milley_mark_general_2020]?
+Milley's response was instructive.
+He offered the Congressmen the idea that even as the Department of Defense continues to take a hard look at the issue, the decision to remove those names should reflect the decision to adopt those names in the first place in that such a decision was "a political decision" [@milley_mark_general_2020].
+It is not unreasonable to insist that policy arenas involving appropriate responses to mass violence, the future of Confederate symbols, and refugee/immigration policy for the nation belong to non-military elites.
+
+A sixth finding is that there are roughly three clusters of observations in Table Four with respect to the year of publication.
+In the first cluster (1998-1999), the publications reveal multiple grievances between certain retired military officers and the manner in which President Clinton was planning or using military force abroad.
+Only two of the five observations in this first cluster are coded as possibly having violated standards of military professionalism.
+
+The second cluster refers to publications penned between 2004-2006.
+Most of these comprise part of or are related to the so-called "Revolt of the Generals," which occurred near the 2004 US midterm elections and referred to sharp disagreements that entered the public arena between multiple US generals and admirals and President George W.
+Bush's Secretary of Defense, Donald Rumsfeld.
+These disagreements were directly related to these military leaders' disagreement about Rumsfeld's leadership regarding the planning and execution of the early years of the war in Iraq.
+Retired Army General Tommy Franks penned two observations that essentially provided public support for the President, George W.
+Bush, while Retired Army Major General Paul Eaton, a critic of Rumsfeld, authored two of the pieces published in this cluster.
+Of the seven observations published in Table 4 between 2004-2006, all but one contained possible violations of the concept of military professionalism.
+
+The third and largest cluster occurs between 2016-2019.
+The 23 observations in this cluster reveals concern over the candidate and later President Donald Trump.
+The titles alone of many of these opinion pieces reveal, at a minimum, clear partisan preferences by many of the authors.
+Retired Air Force General Michael Hayden, Navy Admirals Mike Mullen and William McRaven, and Army General Stanley McChrystal are among the retired officers who authored several publications during this time period.
+All but four of the observations in this cluster are coded as having potentially violated standards of military professionalism.
+Of note, one observation in this cluster entitled, "Fists Raised at West Point" and published in the Washington Post is actually a compendium of two anonymous letters, each one taking an opposite stance on the appropriateness of an incident in which a photograph at West Point's 2016 graduation ceremony captured a graduating female cadet raising a clenched a fist, an action which some saw as an innapproprite political gesture related to the growing Black Lives Matter movement [@unknown_fists_2016].
+
+## Variation in Military Professionalism Over Time
+
+```{r build plots, include = FALSE, echo=FALSE}
+plot1<- ggplot(data=adj2dataset, mapping=aes(x=PubYr, y=case_count)) + geom_point() + geom_smooth() + labs(x="Year", y="Number", color=NULL)
+plot2 <- ggplot(data=adj2dataset, mapping=aes(x=PubYr, y=prop_yr)) + geom_point() + geom_smooth() + labs(x="Year", y="Proportion", color=NULL)
+```
+
+Figure 3, below, plots both the total number and proportion of yearly opinion pieces authored by retired military officers that address topics of social policy affecting the military, domestic policy, and political administrations or figures.
+The points are the actual terms (number and proportion, respectively), while the blue line represents a smoothed line of best fit for the data.
+Both plots reveal that beginning with the first year of the data, 1979, there has a moderate increase until approximately 2005, followed by a decline until approximately 2011, followed by a sharp increase that has persisted until the present day.
+What explains these dynamics - both the increase that occurred between 1979 - 2005 as well as the sharper increase between 2012 - today - and the decline that occurred between roughly 2005 - 2011?
+According to the theory developed at the beginning of this paper, a significant reason stems from changes in political polarization and institutional credibility that have occurred over the past four decades.
+
+```{r figure 3, include=TRUE, warning= FALSE, message = FALSE, echo=FALSE}
+grid.arrange(plot1, plot2, ncol=2, top="Figure 3: #/Proportion of Op-Eds Published\n by Retired Military Officers, 1979-2020", bottom="On Topics of Social Policy which Impacts the Military,\n Domestic Policy, and Political Administrations/Figures")
+```
+
+To understand how measurements of political polarization have changed over time, I rely on the DW-NOMINATE (dynamic-weight, nominal three step estimation) scores that were originally developed by the political scientists Keith Poole and Howard Rosenthal in the early 1980s [@jeffrey_b_lewis_voteview_2020].
+The DW-NOMINATE data are one measure that uses roll-call voting data for each member of congress to assign a score to the member of congress on a liberal-conservative dimension, ranging from -1 (extreme liberal) to 1 (extreme conservative) [@jeffrey_b_lewis_voteview_2020].
+The data is frequently relied upon by academics, journalists, and the media to understand how a variety of political bodies, to include Congress and political parties, have changed over time.
+Over the time period examined in this paper, polarization in each chamber of Congress, the House of Representatives and the Senate, has increased [@jeffrey_b_lewis_voteview_2020].
+The distance in party mean DW-NOMINATE scores in the late 1970s was just less than .6, while the distance today for both chambers is above .8 [@jeffrey_b_lewis_voteview_2020].
+
+Changes in the institutional credibility of the American military institution can be seen by examining public opinion data.
+Pollsters survey the American public annually to gauge their trust in a variety of instutions.
+Using Gallup opinion polling, one sees that since the late 1970s, the credibility of the American military as an institution has gradually increased [@gallup_polls_confidence_2020].
+In 1979, 54% of those surveyed expressed either a great deal or quite a lot of confidence in the military, while in 2020, 72% expressed the same [@gallup_polls_confidence_2020].
+These trends in political polarization and confidence in the military are displayed in Figure 4, below.
+The top and middle graph show changes in means between Republicans and Democrats in the House of Representatives (top graph) and Senate (middle graph) since 1979, while the bottom graph shows changes in the percentage of surveyed Americans who reported either a great deal of or a lot of confidence in the military institution over the same time period.
+
+```{r figure 4, include=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
+
+#adjust the data to account for polarization by creating a new dataset called polaradjdataset.  
+#In this dataset, create variables that capture ideological distance in each the House and Senate along
+#each of the two dimensions according to Poole and Rosenthal's DW-NOMINATE data
+
+polaradjdataset <-adjdataset %>% mutate (
+  House_distance_dim1=abs(House_Dem_mean_nom1-House_Rep_mean_nom1),
+  Sen_distance_dim1=abs(Sen_Dem_mean_nom1-Sen_Rep_mean_nom1),
+  House_distance_dim2=abs(House_Dem_mean_nom2-House_Rep_mean_nom2),
+  Sen_distance_dim2=abs(Sen_Dem_mean_nom2-Sen_Rep_mean_nom2)
+) %>% view()
+
+# Create a variable called polar_and_cred that uses the melt function (from the library reshape2")
+#note that the id.var will serve as the common axis for these 
+polar_and_cred <- melt(subset(polaradjdataset, select=c(PubYr, House_distance_dim1, Sen_distance_dim1, Inst_Cred_Gallup)), id.var="PubYr")
+
+# 
+ggplot(polar_and_cred, aes(x=PubYr, y=value)) + geom_line(aes(color=variable)) + facet_grid(variable~., scales="free_y") + theme(legend.position = "none") +
+  labs(
+    title = "Figure 4: Distance Between Party Means in the House,\n the US Senate,\n and American Confidence in the Military, 1979-2020"
+  )
+```
+
+While it is possible that the trend in increasing instititional credibility of the US military over the past four decades may be at least partially responsible for the increasing trend in the number of opinion pieces published by retired military officers, a trend that the data supported earlier as shown in Figure 2, rising institutional credibility alone does not adequately explain the trends shown in Figure 3: that there has been a sharp increase over the past four decades in retired military officers penning pieces that address non-traditional military topics that lay outside the bounds of relevant military expertise, or pieces that express partisan sentiment.
+The theory laid out earlier in this paper predicted that as political polarization increases, three mechanisms are at work which have strong impacts on the military professional: it becomes increasingly difficult to remain politically neutral as partisan conflicts subsume a greater number of issues; as this occurs, strong feelings of dislike or disdain surface against those who hold on opposing view; and alterations in one's sense of loyalty arise, particularly as issues that individuals hold to most dearly raise the salience of one's identity.
+
+To test if changes in political polarization adequately explain the changes in the qualitative nature (increasing trends of potential violations of military professionalism) of opinion pieces authored by retired military officers, I conduct a logit regression of the variable _questionable_professionalism_ while holding institutinal trust in the military constant at its mean (71 percent) and varying the value of the mean ideological distance between both major political parties in the House between 1979-2020.
+Conducting such a regression allows one to examine, given the data assembled, the predictive likelihood of publishing an opinion piece that violates military professionalism given varying levels of political polarization and constant institutional credibility in the military.
+These results are displayed in Figure 5, below.
+
+```{r logit analysis, include=FALSE, echo=FALSE}
 
@@
 
-#next, conduct logit regressions where the DV is values of Ques_Prof along
-#two IVs, political polarization and institutional credibility.
+#next, conduct logit regressions where the DV is values of Ques_Prof
+# include two IVs: house_distance_dim1 and Inst_Cred_Gallup
 
@@
 
-# now  make prediction data using the mean institutional credibility scores.  Vary #polarization and hold institutional credibility constant.
+# now  make predictions.
 
+# the first prediction data keeps credibility constant and varies house_distance_dim1 from it's min to its max
 predictiondata <- tibble(House_distance_dim1=seq(min(slimdata$House_distance_dim1), max(slimdata$House_distance_dim1), by=.01), Inst_Cred_Gallup = mean(slimdata$Inst_Cred_Gallup)) %>%
@@
 
-predictiondatafut <- tibble(House_distance_dim1=seq(min(slimdata$House_distance_dim1), 1, by=.01), Inst_Cred_Gallup = mean(slimdata$Inst_Cred_Gallup)) %>%
+# the second prediction keeps credibility constant but varies house_distance_dim1 from its min to .99, which is a 
+# perfectly reasonable expectation of the next 10-20 years of increasing political polarization in Congress. 
+predictiondatafut <- tibble(House_distance_dim1=seq(min(slimdata$House_distance_dim1), .99, by=.01), Inst_Cred_Gallup = mean(slimdata$Inst_Cred_Gallup)) %>%
 crossing(Ques_Prof = c(0,1))
@@
 
+#augment first prediction
 augment(logmodel1, newdata = predictiondata)
 
+#augment second prediction
 augment(logmodel1, newdata = predictiondatafut)
 
+# first prediction: house_distance_dim1 from min to max
 predictions <- augment(logmodel1, newdata=predictiondata) %>% 
@@
   )
+
 predictions %>% arrange(.prob, .fitted, Ques_Prof, desc(.fitted)) %>% 
@@
 
+# second prediction: house_distance_dim1 from min to a predicted value of .99
 predictions2 <- augment(logmodel1, newdata=predictiondatafut) %>% 
@@
   view()
+```
 
-# stopping point here.  now let's try some graphs
-# notice that the predictions are not in (0, 1)
-ggplot(predictions2) +
-  aes(x = .fitted) +
-  geom_histogram()
-
-# if we wanted predictions on the probability scale,
-#   we predict on the log scale,
-#   then apply inverse-logit (logistic CDF)
-# use plogis(): cumulative probability of the logistic distribution
-
-# demonstration: scatter log odds (x) and predicted probability (y)
-# with 95% CIs
-# (CIs don't look smooth in this example because .se.fit depends on X data)
-predictions2 %>%
+```{r fig5, include=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
+predictions%>%
   ggplot() +
@@
   geom_pointrange(aes(ymin = conf.low, ymax = conf.high)) +
-  geom_smooth() 
-```
-
-```
-
-```{r table5, include=TRUE, echo=FALSE}
-secondlayer_obs %>%
-  kable(format="latex", caption="Op-Ed Pieces By Retired Military Officers which failed to Uphold Professional Standards, 1979-2020", longtable=T) %>%
-  column_spec(1, width="9.5em") %>% 
-  column_spec(2, width="15em") %>% 
-  column_spec(3:4, width="2.5em") %>%
-  column_spec(5, width="5em") %>%
-  column_spec(6, width="6em") %>%
-  kable_styling(latex_options="striped", font_size=10) %>% 
-  row_spec(0, bold=T)
-```
-\doublespace
-
-Figure 2 displays the number of opinion pieces authored by retired military officers which failed to uphold the strict standards of professionalism, as espoused by Retired Generals MacArthur and Dempsey, within each applicable year.
-The results reveal several conclusions.
-The first is that it clearly shows, in a broad sense, that from 1979 through the first half of 2020, opinion pieces penned at the hand of retired military officers which violate the concepts of professionalism in questionable or inappropriate ways are the exception, not the norm.
-Indeed, the entire count of questionable or inappropriate publications after conducting the second level of analysis drop to just 30 of 334 total pieces, or 8.9% of the entire sample.
-Thus, ample support exists to confirm the first hypothesis, that the vast majority of opinion pieces authored by retired military officers will be apolitical and non-partisan in nature.
-
-```{r fig2, fig.height=4, fig.width=5, fig.align="center", include=TRUE, echo=FALSE}
-datatograph <- data.frame(
-  year=c("1998","2000","2003", "2004", "2006", "2007", "2016", "2017", "2018", "2019", "2020"), count=c(2, 1, 1, 3, 3, 1, 5, 4, 4, 5, 1)
-)
-ggplot(datatograph, aes(x =year, y =count)) + geom_bar(stat = "identity", width=.2) +
-  labs(
-    x = "Year",
-    y = "Count",
-    color = NULL,
-    title = "Figure 2: # of Op-Eds Written By Military Retirees",
-    subtitle = "Failed to Uphold Professional Standards"
+  geom_smooth() + labs(
+    x="log Mean Distance Between Republicans and Democrats\n in the US House of Representatives, 1979-2020",
+    y="Predicted Probability of A Retired Military Officer\n Publishing an Op-Ed that Violates Military Professionalism",
+    title="Figure 5: The Predicted Impact of Political Polarization\n on Military Professionalism"
   )
@@
 
-The second hypothesis, which inferred that because overall circulation among major US newspapers has been declining over the past several decades, the number of opinion pieces penned by retired military officers in these outlets would likewise decline, does not appear to be represented from the data.
-In fact, we find evidence for the opposite trend: retired military officers are penning opinion pieces at an increasing rate, even as overall circulation among the major newspapers are down in terms of overall circulation (even after accounting for digital offerings) over the same time period.
+The graph, which is modeled off of predicted data using a logit regression, clearly shows a non-linear relationship between increasing political polarization and the likelihood that an opinion piece authored by a retired military officer will potentially violate aspects of military professionalism.
+While the confidence intervals at each point estimate are relatively large, the graph clearly shows that the predicted probability of violating military professionalism is different between extremes along the x axis.
+In other words, the political polarization that exists today, as measured by the ideological distance between the political parties in the house, is predicted to have resoundingly different impacts on military professionalism than the polarization levels that existed several decades ago.
+Whereas the polarization levels that existed several decades ago may have had negligible, if any impact, on the likelihood of a retired military officer violating military professionalism in the content of a published opinion piece, the polarization levels that exist today have, at least insofar as the predictive power of the model is concerned, changed the landscape: today's polarization levels reveal a roughly 15% chance of violating military professionalism, with confidence intervals that range from roughly 9% to nearly 23%.
 
-The third hypothesis, that the number of published opinion pieces which have strained the concept of professionalism has increased, even as political polarization has worsened in the past several decades, is also confirmed, as evidenced by both by Table 3 and Figure 3.
-It is interesting to note that according to Figure 3, the first publishing year in which an opinion piece penned by a retired officer failed to uphold professional standards occurred in 1998, nearly twenty years into (or almost half) of the entire sample.
-But why is this the case?
-Retired military officers in the years preceding 1998 had plenty of reasons to go to the press and air disagreements with political leaders, or to address important topics that may have pushed the limits of professionalism as has been defined in this paper.
-In fact, many of these officers would have been veterans of the Vietnam and Korean Wars, and perhaps, like Maxwell Taylor, who is in our sample, veterans perhaps of World War Two.
-At this point, though it is only speculation, one can speculate that the factors of military credibility and political polarization, however imprecise and hazy at this point, may have played a role in keeping the majority of an earlier generation's retired officers out of the popular press.
+## Does Combat Attenuate Polarization's Impact on Military Professionalism?
 
-## Increasing Diversity of Topics Addressed By Military Officers
+Before concluding, it is worth commenting about why Figure 3 revealed a sharp drop in the number and proportion of observations that were written about topics such as social policy which impacts the military, domestic policy, and political administrations/figures between roughly 2005 - 2011.
+One compelling possibility is that between these years, the military was heavily engaged in combat operations overseas to a point that the institution was taxed tremendously: repeated deployments, surges of forces to both Iraq (2007-2008) and Afghanistan (2010-2011), and sustained US casualties from these conflicts.
+This is not to suggest that the military was not engaged in conflicts around the globe in other periods of time, but to instead suggest that even retired military officers will be impacted by the institutional requirements of the active military.
+When such requirements include repeated deployments and an operational tempo that spans multiple years, perhaps retired military officers acknowledge that their voice is best served advancing these interests, as opposed to weighing in on other topics.
 
-It is also clear from the data presented in Table 3 that several publications are deemed questionable or inappropriate from the perspective of professionalism in that they address topics that are potentially outside of the realm of proper military expertise to address.
-In his _Washington Post_ editorial, "Home Should Not Be a War Zone," Retired Army General Stanley McChrystal addresses the tragic nature of mass shootings in America.
-He states in the article, "...as a combat veteran and proud American, I believe we need a national response to the gun violence that threatens so many of our communities" [@mcchrystal_home_2016].
-In a later article published in the _Washington Post_ in 2018 entitled "Good Riddance," McChrystal shared how his beliefs in Robert E.
-Lee had evolved over the course of his professional career, and why he now believes it is now time to discard Confederate statues and other similar symbols.
-Admiral (Retired) Mike Mullen and co-authors Lieutenant General (Retired) Mark Hertling and Admiral (Retired) Robert Natter wrote publications ("The Refugees We Need" and "Cutting Refugee Admissions Will Have Severe Consequences for the US Military") explaining how President Trump's immigration policies on refugees harm military readiness.
-
-While many Americans agree with the stance these officers have taken on what are clearly important issues, one can simultaneously and not unreasonably ask whether it is appropriate for retired military officers to be considered appropriate elite voices on these specific issue areas.
-Again, the potential issue as it is related to the concept of military professionalism is not that these officers should not have an opinion on these topics, but whether they, as retired military officers, ought to be considered as appropriate mouthpieces by the American public on these issues.
-It is not unreasonable to insist that policy arenas involving appropriate responses to mass violence, the future of Confederate symbols, and refugee/immigration policy for the nation belong to non-military elites.
-
-```{r fig3, fig.height=4, fig.width=5, fig.align="center", include=TRUE, echo=FALSE}
-datatograph <- data.frame(
-  decade=c("1979-1989", "1989-1999", "1999-2009", "2009-2019" ), numcount=c(13, 51, 102, 161)
-)
-ggplot(datatograph, aes(x =decade, y =numcount)) + geom_bar(stat = "identity", width=.2) +
-  labs(
-    x = "Period",
-    y = "Published Opinion Pieces",
-    color = NULL,
-    title = "Figure 3: # of Op-Eds Written By Military Retirees",
-    subtitle = "Published in WSJ, NY Times, Wa Post, USA Today, and LA Times"
-  )
-```
-
-# Conclusion: How Can the Military Remain Professional?
+# Conclusion: Will the Military Remain Professional in a Polarized Setting?
 
@@
 Nearly 40% of all opinion pieces in 2019 published in major US newspapers by retired military officers were written about topics that fell outside of traditional national security topics.
-
-```{r build plots, include = FALSE, echo=FALSE}
-plot1<- ggplot(data=adj2dataset, mapping=aes(x=PubYr, y=case_count)) + geom_point() + geom_smooth() + labs(x="Year", y="Number", color=NULL)
-plot2 <- ggplot(data=adj2dataset, mapping=aes(x=PubYr, y=prop_yr)) + geom_point() + geom_smooth() + labs(x="Year", y="Proportion", color=NULL)
-```
-
-```{r figure 4, include=TRUE, warning= FALSE, message = FALSE, echo=FALSE}
-grid.arrange(plot1, plot2, ncol=2, top="Figure 4: #/Proportion of Op-Eds Published by Retired Military Officers, 1979-2020", bottom="On Topics of Social Policy, Domestic Policy, and Political Administrations/Figures")
-```
 

### 2020-09-21 code/Submissionv3.Rmd
commit

@@
 
-  _This paper explores variation in the professionalism of opinion pieces authored by retired US military officers and published in major US newspapers in the past 40 years.
+  _This paper explores variation in the professionalism of opinion pieces authored by retired US military officers and published in major US newspapers since 1979.
 In doing so, this paper develops a theory which argues that increasing political polarization in the United States impacts the military as a profession by straining the individual military officer's concept of loyalty in important ways.
-Drawing on an original data set of more than 330 opinion-editorial pieces published by retired military officers in The Wall Street Journal, The Washington Post, The New York Times, The Los Angeles Times, and the USA Today since 1979, the paper argues there has been an increase in published opinion pieces which fail to uphold professional standards as defined by historical military norms and standards.
-The paper concludes by arguing that the concept of "politicization of the military" --- something that previous civil-military relations scholars have identified and described --- must be widened to account not only for efforts by politicians to use the military for political advantages, but also for the novel phenomenon of retired military officers using the concept of loyalty to achieve political ends.
+Drawing on an original data set of more than 340 opinion-editorial pieces published by retired military officers in The Wall Street Journal, The Washington Post, The New York Times, The Los Angeles Times, and the USA Today since 1979, the paper argues there has been an increase in published opinion pieces which fail to uphold the concept of military professionalism as defined by historical military norms and standards.
+The paper concludes by arguing that the concept of "politicization of the military" --- something that previous civil-military relations scholars have identified and described --- must be widened to account not only for efforts by politicians to use the military for political advantages, but also for the novel phenomenon of retired military officers appealing to one's concept of loyalty to achieve political ends.
 The author argues that such a development, if uncorrected, will cause significant damage to the military as a profession._
@@
 In late May and early June 2020, racial tension and concerns about police brutality gripped the American public following the death of George Floyd, an unarmed black man, by police in Minneapolis.
-In many cities and communities across the nation, largely-peaceful protests were followed by periods of rioting and looting.
+In many cities and communities across the nation, peaceful protests were followed in some cases by periods of rioting and looting.
 On June 1, after giving remarks at the Rose Garden, President Donald Trump walked to St.
@@
 Further, this paper argues that the military institution is particularly susceptible to increasing political polarization because of the way the concept of loyalty is used to make arguments that contain a clear partisan or political leaning.
-These developments reveal the extent to which increasing political polarization is synonymous with a fractured domestic public, and highlight the extraordinary challenge that the US military will llikely face to remain a  professional in an enduring polarized political climate.
+These developments reveal the extent to which increasing political polarization is synonymous with a fractured domestic public, and highlight the extraordinary challenge that the US military will likely face to remain a  professional in an enduring polarized political climate.
 
@@
 Second, the paper builds a theory that demonstrates how certain features of today's polarized political landscape challenge the concept of loyalty for the individual military professional.
-Third, the paper explores the data, which consists of more than 330 opinion-editorial pieces published in major US newspapers by retired military officers since the late 1970s.
-This analysis ultimately shows that although the vast majority of op-eds published by retired military officers maintain professional standards, there is clear evidence that retired military officers are using the pen to both argue about topics that fall outside of the realm of traditional national security topics, and to express clear partisan preferences when making their arguments.
-The analysis also shows that this trend has worsened as the American political climate has become increasingly polarized.
-Fourth, the paper concludes by briefly exploring the implications of these results on the military as an institution, and specifically, on its ability to remain professional in a highly polarized setting.
+Third, the paper explores and analyzes a data set which consists of more than 340 opinion-editorial pieces published in major US newspapers by retired military officers since the late 1970s.
+This analysis ultimately shows that although the vast majority of op-eds published by retired military officers maintain professional standards, there is clear evidence that retired military officers are using the pen both to argue about topics that fall outside of the realm of traditional national security topics, and to express clear partisan preferences when making their arguments.
+The analysis also shows that this trend has worsened as the American political climate has become increasingly polarized over the past several decades.
+Finally, the paper concludes by briefly exploring the implications of these results on the military as an institution, and specifically, on its ability to remain professional in a highly polarized setting that will likely endure.
 
@@
 The ideal of the US military acting as an objective, apolitical and non-partisan institution is long and storied.
-As a concept, the idea of a professional military stems from the post-WW2 era, when the US military was locked in the Cold War struggle agains the Soviet Union.
+As a concept, the idea of a professional military stems from the post-WW2 era, when the US military was locked in the Cold War struggle against the Soviet Union.
 Perhaps the strongest and certainly one of the original intellectual founders of a professional military is the political scientist, Samuel Huntington, who, in his 1957 work, _The Soldier and the State: The Theory and Politics of Civil-Military Relations_, asserted that two distinct forces would always be in tension with one another in a liberal democracy such as the United States.
@@
 Nonetheless, academic theorists, military tradition, and contemporary military leaders have long argued for the importance of the military remaining grounded in professional conduct that is both non-partisan and apolitical.
-At its core, it is this type of professional conduct which ultimately postures the military to earn the trust and confidence of the American people.
+At its core, it is this type of professional conduct which ultimately postures the military to earn the trust and confidence of the American people^[There are certainly other aspects that impact the military's ability to earn the public's trust, such as performing its military missions well, winning in combat, etc.].
 
@@
 In fact, one of the foundations of Huntington's concept of military professionalism is the fact that military leaders possess unique and specialized skills that civilians simply do not have.
-Moreover, multistar generals and admirals in the contemporary military possess in many cases upwards of 30 and sometimes 40 years of service, and it is entirely reasonable to conclude that these officers in particular have been privy to decisions and experiences occurring at or near the highest levels of government.
+Moreover, multistar generals and admirals in the contemporary military possess in many cases upwards of 30 and sometimes 40 years of service, and it is entirely reasonable to conclude that these high-ranking officers in particular have been privy to decisions and experiences occurring at or near the highest levels of government.
 In other words, their experiences are not confined to only tactical matters.
-They are familiar with issues ranging from how wars should be best fought, to future security threats looming on the horizon, to fixing the institutional culture of a military service, and more.
+They are familiar with issues ranging from how wars should be best fought, to future security threats looming on the horizon, to fixing the institutional culture of a military service, and much more.
+
 It is reasonable, therefore, to conclude that a professional opinion piece written by a military officer would take on three qualities.
@@
 
-If and when a military officer violates standards of military professionalism when conveying an opinion to the American public, his or her words can cause great harm, either to American society, the military as an instition, or a combination of both.
+If and when a military officer violates standards of military professionalism when conveying an opinion to the American public, his or her words can cause great harm, either to American society, the military as an institution, or a combination of both.
 To understand how and why this is the case, we will now turn our attention to understanding the impacts on military professionalism caused by two changes that have developed within American Society in the past 50 years.
@@
 
+\singlespace
 # When Words Become Weaponized - A Theory of Political Polarization and Military Professionalism
+\doublespace
 
@@
 Each of these three features of the politically polarized landscape strain the concept of military professionalism in important ways, and taken together, comprise a significant challenge to the individual military professional and thus, to the professional nature of the military as a whole.
-These challenges are summarized in Table One, below.
+These challenges are summarized in Table 1, below.
 
@@
 The first way in which the concept of military professionalism is challenged as a result of increasing political polarization is that there is increasing difficulty for individuals to remain politically neutral, non-partisan, or otherwise apolitical - a hallmark of military professionalism.
-This occurs because as "conflict extension" [@layman_party_2002] occurs and an increasing number of issues are swallowed up by partisan battlegrounds, there is less politically neutral space for someone to stand on.
+This happens because as "conflict extension" [@layman_party_2002] occurs, an increasing number of issues are brought into partisan battlegrounds, and thus, there is less politically-neutral territory.
 When polarization levels in America were not as high as they are today, it was more possible for individuals to either stay out of politics, avoid partisan squabbles, and otherwise to stay politically neutral.
@@
 However, it is also precisely on this point that the military institution is perhaps most vulnerable to increasing political polarization.
-I contend that the rise of identity politics can and will be harmful if a military professional's identity and the loyalty associated with that identity directly conflicts with his or her identity and loyalty as a professional soldier.
+I contend that the rise of identity politics can and will be harmful if a military professional's identity and the loyalties associated with that identity directly conflict with his or her identity and loyalty as a professional soldier.
 As any soldier knows, and as civilians can reasonably understand,few traits are more sacred in the military than loyalty.
 Indeed, the virtue of loyalty is extolled in a host of codes and mantras that have played a central part of the American military since its inception, and loyalty is captured, either in word or in spirit, in a host of phrases and professional codes.
-In the middle of the the West Point Cadet Prayer is the eloquently written plea, "make us to choose the harder right instead of the easier wrong" [@wheat_west_1918].
+In the middle of the West Point Cadet Prayer is the eloquently written plea, "make us to choose the harder right instead of the easier wrong" [@wheat_west_1918].
+Even Huntington spent a great deal talking about the importance of loyalty as a trait to the very essence of the military profession.
+He recognized that in order for civilian control to be effective, the military profession had to organize itself as a "hierarchy of obedience," which required "loyalty and obedience" to stand out as "the highest military virtues" [@huntington_soldier_1957, 73].
 
-High levels of political polarization obfuscate the ideal of loyalty, however.
-As core issues that involve individual perceptions of moral right and wrong are at stake, individuals, to include military officers, are less likely to not become involved, attached, or otherwise invested in these issues.
-This is especially true when the language and appeal of concepts of loyalty are employed to justify one is or is not taking action over an issue.
-In a highly polarized political climate, there is increasing likelihood that two different people military officers may adopt completely different opinions regarding a contentious issue, each believing that the other is deeply and perhaps even morally wrong.
+High levels of political polarization obfuscates the ideal of loyalty, however, and thus pose a significant risk to the professional nature of the military.
+As core issues that involve individual perceptions of moral right and wrong are at stake in an increasing number of partisan battles, individuals, to include military officers, are more likely to become attached to or otherwise invested in these issues.
+This is especially true when the justification for taking or not taking action invokes the concept of loyalty.
+As a result, highly polarized political climates increase the likelihood that two different military officers may adopt completely different opinions or stances on an issue, each believing that the other is not only morally wrong, but failing to act out of loyalty in accordance with a specific oath or moral code.
+A couple of recent cases illustrate how military officers, active and retired, are invoking the concept of loyalty to make an appeal to action that is, if not outright hostile to, certainly not helpful for, the ideal of military professionalism.
 
-As a quick case in point, consider that in June 2020, following President Trump's largely unpopular move of displacing protesters from Lafayette Square, a letter published by West Point alumni expressly called out prominent West Point alumni serving in the Trump administration.
+First, in June 2020, following President Trump's largely unpopular move of displacing protesters from Lafayette Square, a letter published by West Point alumni expressly called out prominent West Point alumni serving in the Trump administration.
 Though it did not mention individuals by name, the letter was intended and acknowledged by media as directed towards the Secretary of Defense, Mark Esper, and the Secretary of State, Mike Pompeo, who are both graduates of West Point's Class of 1986 [@gaouette_group_nodate].
-The charges of the letter were both damning and politically striking: "...We are concerned that fellow graduates serving in senior-level, public positions are failing to uphold their oath of office and their commitment to Duty, Honor, Country"[@noauthor_letter_2020].
-The letter's clear implication is that if you are serving in the Trump administration and if you graduated from the US Military Academy (i.e.
-Secretaries Esper and Pompeo), you have forsaken your oath of office, the US Constitution, and the values that West Point stands for.
+The charges of the letter were both damning and striking: "...We are concerned that fellow graduates serving in senior-level, public positions are failing to uphold their oath of office and their commitment to Duty, Honor, Country"[@anonymous_authors_letter_2020].
+The letter's clear implication is that Secretaries Esper and Pompeo have forsaken their oaths of office, the US Constitution, and the values that West Point stands for (the values of duty, honor, and country).
+
+Then, in July 2020, several recent West Point graduates, who are now officers in the US Army, released an open letter addressed to the Superintendent, Commandant, Academic Dean, and Athletic Director of West Point.
+The letter's authors expressed concern that West Point had not done all it could do to be a "anti-racist" institution, and made a list of requests of the Academy's leadership, including "releasing a public statement acknowledging the existence of systemic racism at West Point, demonstrating commitment to anti-racism..., outlining anti-racist education..., and detailing plans to eradicate Confederate memorialization" [@askew_anti-racist_2020].
+The authors further stated that, "we hope Academy leadership will agree that, by writing this proposal, we are espousing the same values that West Point seeks to instill in its graduates: leadership, duty, respect, selfless service, honor, integrity, and personal courage" [@askew_anti-racist_2020].
+
+And in August, 2020, yet another open letter, this time written by a pair of retired Army Lieutenant Colonels, John Nagl and Paul Yingling, addressed the Chairman of the Joint Chiefs of Staff, General Mark Milley, invoked the concept of loyalty to frame a very disconcerting prospect: active duty military forces removing President Trump from office if, in January, he fails to accept the results of the election [@nagl__2020].
+Nagl and Yingling assert that Milley has a choice to make.
+As the letter states to Milley, "the choice between these two options [_removing the President from office, or nothing_] lies with you...your duty is to give unambiguous orders directing U.S.
+military forces to support the Constitutional transfer of power.
+Should you remain silent, you will be complicit in a coup d’état [@nagl__2020,_emphasis mine].
+
+Each of these three instances could deserve a much longer exposition and analysis, but the important thing to notice here as it relates to political polarization is how in each case, the concept of loyalty is used to justify taking action, but even more than that, to use the concept of loyalty to bring disdain and perhaps even shame on someone for failing to act in accordance with the letter's demands.
+This is remarkable in that in effect, political polarization has served to draw up battle lines for America's warriors, its venerated and historically apolitical institution.
 
@@
 
-Ultimately, four possible worlds exist when the impacts of military credibility and political polarization are considered in tandem (see Table Two, below).
+Ultimately, four possible worlds exist when the impacts of military credibility and political polarization are considered in tandem (see Table 2, below).
 The first is when political polarization and military credibility are both low.
@@
 With these conditions, the military does not have an out-sized platform to interact with the American public, nor are there outstanding incentives for the military to raise significant disagreement over any domestic political issues.
-
 \singlespacing
@@
 \begin{footnotesize}
-\caption{Table Two: Four Types of "Military Professionalism Worlds" Where Institutional Credibility and Political Polarization Vary} % title of Table
+\caption{Four Types of "Military Professionalism Worlds" Where Institutional Credibility and Political Polarization Vary} % title of Table
 \centering % used for centering table
@@
 dataset %>% print()
-# Only read in the first 334 rows - is there a better way to get rid of N/A data?
+# Only read in the first 344 rows - is there a better way to get rid of N/A data?
 # Replace NA with 0 in the data for variables War_Ops thru Fig_Nme
-adjdataset <- dataset[1:334,] %>%
+adjdataset <- dataset[1:344,] %>%
   mutate_at(
@@
 
+# Create a new variable called Yrcount, which is a count of the yearly number of published op-eds
+
+adjdataset <- adjdataset %>% 
+  group_by(PubYr) %>% 
+  mutate(Yrcount=n()) 
 ```
@@
 Second, I apply a normative lens to determine if there have been shifts in the tone and tenor of the pieces authored by these officers, particularly regarding the professional aspects of the military previously discussed (apolitical in nature, non-partisan, relying on relevant military expertise, etc.).
-In essense, I want to answer the questions, what are these retired military officers writing about, and through the lens of professionalism, should they be saying what they are saying?
-Is there any indication that increasing political polarization is altering what and how these retired officers are communicating with the American public?
+In essence, I want to answer the questions, what are these retired military officers writing about, and then, using the normative approach to military professionalism just discussed, are these op-ed's deviating from the concept of military professionalism, and how?
 
@@
 I confine my search to the largest US newspapers that regularly run opinion-editorial pieces, which includes  _The Wall Street Journal_ (WSJ), _The New York Times_ (NYT), _The Washington Post_ (WaPo), _The Los Angeles Times_, and _The USA Today_.
-The time period for the data includes the period January 1, 1979 - 30 June 2020, the main limitation being that not all of the Newspaper sources are available for previous dates.
+The time period for the data includes the period January 1, 1979 - 30 August, 2020, the main limitation being that not all of the Newspaper sources are available for previous date periods.
 I also open my search to include both print and online offerings in order to ensure that I am capturing any opinion pieces published digitally.
+The data only include publications classified as either editorials or commentaries/opinions by the databases.
+Thus, other types of pieces that may still be authored by retired military officers, such as letters to the editor and/or published interviews, are not included.
 
-The data set contains 334 opinion-editorial publications, penned by 178 different authors or author teams.^[There are several co-authored pieces.
-For example, where General (Retired) David Petraeus wrote one publication and another one was written by General (Retired) David Petraeus and Michael O'Hanlon, I counted these as two different authors for coding purposes.] Ten publications were written anonymously, with at best general recognition by the authors to their previous military service and/or rank.
+The data set contains 344 observations, penned by 185 different authors or author teams.^[There are several co-authored pieces.
+For example, where General (Retired) David Petraeus wrote one publication and another one was written by General (Retired) David Petraeus and Michael O'Hanlon, I counted these as two different authors for coding purposes.] Several authors penned multiple pieces, including Retired Army Lieutenant General (23 publications) and Retired Air Force General Michael Hayden (20 publications).
+Ten publications were written anonymously, with at best general recognition by the authors to their previous military service and/or rank.
 From a macro perspective, consider for a moment all that the date range includes with respect to the US military.
@@
 
-```{r figure two, fig.height=4, fig.width=5, fig.align="center", include=TRUE, echo=FALSE}
+```{r figure two, fig.height=4, fig.width=5, fig.align="center", warning=FALSE, include=TRUE, echo=FALSE}
 
-#Need to adjust this section depending on the number of observations I have
-# Also need to figure out how to code this properly.  There has to be a better way. 
-# Adjust using the geom_text functions to label each bar properly.
+# construct a scatter plot with line and smoothing
 
-datatograph <- data.frame(
-  decade=c("1979-1988", "1989-1998", "1999-2008", "2009-2018", "2019-Present" ), numcount=c(13, 37, 109, 152, 23)
-)
-ggplot(datatograph, aes(x =decade, y =numcount)) + geom_bar(stat = "identity", width=.2) + geom_text(aes(label=numcount), position=position_stack(vjust=1.08))+
-  labs(
-    x = "Time Interval",
-    y = "Number of Publications",
-    color = NULL,
-    title = "Figure Two: Op-Eds Written By Retired Military Officers",
-    subtitle = "Published in WSJ, NY Times, Wa Post, USA Today, and LA Times"
-  )
+ggplot(data=adjdataset, aes(x=PubYr, y=Yrcount)) + geom_point() + geom_line() + geom_smooth(method="loess") + labs(
+  x="Year", 
+  y="count", 
+  title="Figure Two: Number of Op-Eds Published by Retired\n Military Officers (WSJ, NYT, WaPo, USA Today, and LA Times",
+  subtitle= "1979-2020",
+  caption="Note: As of 31 August 2020"
+  ) + scale_x_continuous(limits=c(1979, 2020)) + scale_y_continuous(limits=c(-7, 28)) + 
+geom_vline(xintercept = 1981, size=.2, colour="tomato") + 
+  geom_vline(xintercept=1989, size=.2, colour="tomato") + 
+  geom_vline(xintercept=1993, size=.2, colour="dodgerblue") + geom_vline(xintercept=2001, size=.2, colour="tomato") + 
+  geom_vline(xintercept=2009, size=.2, colour="dodgerblue") +
+  geom_vline(xintercept=2017, size=.2, colour="tomato") +
+  geom_text(aes(x=1981, label="Reagan",y=-5), colour="tomato", angle=90, vjust=1, text=element_text(size=9)) + 
+  geom_text(aes(x=1989, label="Bush 41",y=-5), colour="tomato", angle=90, vjust=1, text=element_text(size=9)) + 
+  geom_text(aes(x=1993, label="Clinton",y=-5), colour="dodgerblue", angle=90, vjust=1, text=element_text(size=9)) + 
+  geom_text(aes(x=2001, label="Bush 43",y=-5), colour="tomato", angle=90, vjust=1, text=element_text(size=9)) + 
+  geom_text(aes(x=2009, label="Obama",y=-5), colour="dodgerblue", angle=90, vjust=1, text=element_text(size=9)) + 
+  geom_text(aes(x=2017, label="Trump",y=-5), colour="tomato", angle=90, vjust=1, text=element_text(size=9))
+
 ```
@@
 
-_$H_3$: Given that polarization has become increasingly pronounced in the United States over the time period examined, there will be an increase in recent years in the number of op-eds which reveal political and partisan sentiments, indicating a potential drop in military professionalism across the ranks of retired military officers across the past several decades._
+_$H_2$: Given that polarization has become increasingly pronounced in the United States over the time period examined, there will be an increase in recent years in the number of op-eds written about topics that are not relevant to military expertise, or that reveal inappropriate partisan sentiments.
+Such a trend may indicate that increasing political polarization is impacting military professionalism among a small but important part of the military institution, retired military officers._
 
-# First Layer Analysis: Topics Addressed by Retired Military Officers
+# First Layer Analysis: Topics Addressed
 
@@
 
-These topical variables are: warfighting and operations, if the piece advocates for or denounces a particular course of action directly related to operational or war fighting decisions of a clear military nature; troop support and institutional culture, if the publication generally highlights the work and sacrifice of members of the armed forces, defends the work of specific members of the Armed Forces, or makes a call for action to improve a service's institutional culture; strategic advice, if the piece is generally written to offer some sort of strategic advice for the nation, to include the wisdom of pursuing particular alliances or agreements, or adopting a specific weapons platform or military budget; social policy directly related to the military, if the piece advocates for or denounces a particular social policy that _directly influences_ the armed forces (Examples include: debates over keeping or removing the "Don't Ask, Don't Tell" policy, opening up combat roles to women, the role of Transgender members of the armed forces); domestic policy, if the piece advocates for or denounces a position on a policy that _does not directly_ influence the armed forces in its day to day activities (Examples include: support for the second amendment, immigration policy, and whether confederate statues should be taken down or left up); figures by name, if the piece advocates for or denounces a particular position held by a political administration, or _if the main point of the piece_ is to explicitly defend or attack a political figure by name (Examples: endorsing a Presidential candidate for office, or for attacking a President or member of Congress for a decision that was made).
+These topical variables are: war fighting and operations, if the piece advocates for or denounces a particular course of action directly related to operational or war fighting decisions of a clear military nature; troop support and institutional culture, if the publication generally highlights the work and sacrifice of members of the armed forces, defends the work of specific members of the Armed Forces, or makes a call for action to improve a service's institutional culture; strategic advice, if the piece is generally written to offer some sort of strategic advice for the nation, to include the wisdom of pursuing particular alliances or agreements, or adopting a specific weapons platform or military budget; military social policy, if the piece advocates for or denounces a particular social policy that _directly influences_ the armed forces (Examples include: debates over keeping or removing the "Don't Ask, Don't Tell" policy, opening up combat roles to women, the role of Transgender members of the armed forces); domestic policy, if the piece advocates for or denounces a position on a policy that _does not directly_ influence the armed forces in its day to day activities (Examples include: support for the second amendment, national immigration policy, and whether confederate statues should be taken down or left up); and political figures, if the piece advocates for or denounces a particular position held by a political administration, or _if the main point of the piece_ is to explicitly defend or attack a political figure by name (Examples: endorsing a Presidential candidate for office, or for attacking a President or member of Congress for a decision that was made).
 
-Separating opinion-editorial publications into one of several distinct categories is an imperfect process, and requires as objective a reading of the publication as possible.
+Up front, separating opinion-editorial publications into one of several distinct categories is an imperfect process.
+It requires a thoroughly objective reading of the publication as possible.
 To aid the coding process, I identify the thesis statement of each opinion piece, which in most cases is identifiable in either the lead or concluding paragraph of the opinion piece.
@@
 
-First, opinion pieces are intended to argue for something, and thus are, to an extent, designed to be at least mildy provocative.
-With this in mind, I did not necessarily code pieces that called for a political figure or instituton to take a specific action into the category of _figures by name_.
+First, opinion pieces are intended to argue for something, and thus are, to an extent, designed to be at least mildly provocative.
+With this in mind, I did not necessarily code pieces that called for a political figure or institution to take a specific action into the category of _figures by name_.
 Again, the identification of the publication's thesis statement aided significantly in identifying the main thrust of the argument, and thus, in classifying the argument by topic.
@@
 However, in instances of choosing between two categories, the choice to classify between categories was usually between categories that were reasonably close.
-For example, there were a few instances of choosing to classify a publication between the warighting and operations category and the strategic advice category, but no instances of choosing between the warfighting and operations category and the domestic policy category.
+For example, there were a few instances of choosing to classify a publication between the war fighting and operations category and the strategic advice category, but no instances of choosing between the war fighting and operations category and the domestic policy category.
 Table 3 reveals the breakdown of all examined publications by article type.
@@
 ```
-An overwhelming majority (289 of 334, or 86.5%) of the total number of opinion pieces in the dataset fall into the categories of warfighting and operations, troop support and institutional culture, and strategic advice.
-Moreover, I conclude that these observations pose no problem for norms of military professionalism, as they are written on topics that I contend are clearly related to relevant military experience.
-The remaining 45 publications (13.5% of the observations) fall under the other three categories of social policy directly related to the military, domestic policy, and figure by name.
-These obsercations require a second and more-thorough examination to determine whether any of these observations indicate a drop in the norm of military professionalism.
-This second layer of analysis will allow us to examine these instances in greater detail.
+An overwhelming majority (293 of 344, or 85.2%) of the total number of opinion pieces in the data set fall into the categories of war fighting and operations, troop support and institutional culture, and strategic advice.
+Moreover, I conclude that these observations pose no problem for norms of military professionalism, as they are written on topics that I contend are clearly related to one's relevant military experience.
+The remaining 51 publications (14.8% of the observations) are coded as falling under the other three categories of topics: social policy directly related to the military, domestic policy, and political figures by name.
+These observations require a second and more-thorough examination to determine whether the content of these publications indicate a drop in the norm of military professionalism.
 
@@
 whether it is a violation of topic or of expression of political sentiment.
-The results of this analysis on the remaining 45 observations are displayed in Table 4, below.
+The results of this analysis on the remaining 51 observations are displayed in Table 4, below.
 Several observations are worth noting.
@@
 \singlespace
-```{r table 4, fig.align = "center", include=TRUE, echo=FALSE}
+```{r table4, fig.align = "center", include=TRUE, echo=FALSE}
 #Need to adjust the width and spacing of this table
@@
 
-First, 15 of the 45 publications are, upon a closer reading, determined to not reflect possible violations of military professionalism.
+First, 17 of the 49 publications are, upon a closer reading, determined to not reflect possible violations of military professionalism.
 This is worth showcasing in order to draw attention to the fact that there are instances of retired military officers authoring opinion pieces which address the subjects of social policy related to the military, domestic policy, and figures by name while simultaneously upholding standards of military professionalism.
-However, the data reveals that in practice, such instances are relatively uncommon.
-The practical implication is simple and straightforward: if military officers decide to write about topics other than warfighting and operations, troop support and institutional culture, or strategic advice, they need to carefully craft their argument so as to not violate norms of military professionalism.
+However, the data reveals that in practice, such instances are questionable.
+The practical implication is simple and straightforward: if military officers decide to write about topics other than war fighting and operations, troop support and institutional culture, or strategic advice, they need to carefully craft their argument so as to not violate norms of military professionalism.
 
 Second, and resulting from the first, is that across the entire data set, instances of potential violations of military professionalism in the op-eds published by retired military officers in major US newspapers is a somewhat rare occurrence.
-Indeed, the entire count of potentially questionable publications with respect to military professionalism across the data set is just 30 of 334 total pieces, or 8.9% of the entire sample.
+Indeed, the entire count of potentially questionable publications with respect to military professionalism across the data set is just 32 of 344 total observations, or 9.3% of the entire sample.
 Thus, ample support exists to confirm the first hypothesis, that the vast majority of opinion pieces authored by retired military officers uphold standards of military professionalism.
@@
 Why is this the case?
-Certainly, retired military officers before 1998 had plenty of reasons to go to the press and air disagreements with political leaders, or to address important topics that may have pushed the limits of military professionalism.
-In fact, many of these officers would have been veterans of the Vietnam and Korean Wars, and perhaps, like Maxwell Taylor, an author in the sample, veterans of World War Two.
+We will deal more with the suspected culprits of this trend, rising political polarization and rising institutional credibility of the military, later in this paper.
+The point to make is that without question, retired military officers before 1998 had plenty of opportunities either to air partisan preferences and/or to address topics that may have similarly pushed the limits of military professionalism.
+In fact, many of these retired officers would have been veterans of the Vietnam and Korean Wars, and perhaps, like Maxwell Taylor, an author who is in the sample data, veterans of World War Two.
+Moreover, the data shows that these retirees wrote opinion pieces ( a total of 46 observations were written before the first observation in Table 4 raises a question about military professionalism), but that in doing so, they stayed clearly within the bounds of military professionalism.
 
@@
 
-Table 3 reveals, fifth, that several publications are deemed as possibly having violated military professionalism in that they address topics that may be considered outside of the realm of proper military expertise.
+Table 4 reveals, fifth, that several publications are deemed as possibly having violated military professionalism in that they address topics that may be considered outside of the realm of proper military expertise.
 It is worth briefly commenting on several of these examples.
@@
 He states in the article, "...as a combat veteran and proud American, I believe we need a national response to the gun violence that threatens so many of our communities" [@mcchrystal_home_2016].
-In a later article published in the _Washington Post_ in 2018 entitled "Good Riddance," McChrystal shared how his beliefs in Robert E.
-Lee had evolved over the course of his professional career, and why he now believes it is now time to discard Confederate statues and other similar symbols in America.
-Other examples include publications by Admiral (Retired) Mike Mullen and co-authors Lieutenant General (Retired) Mark Hertling and Admiral (Retired) Robert Natter, who respectively wrote the publications, "The Refugees We Need" and "Cutting Refugee Admissions Will Have Severe Consequences for the US Military."
+He later, in 2018, pens an op-ed entitled "Good Riddance" in which he discusses the evolution of his views on Robert E.
+Lee and other Confederate soldiers and symbols [@mcchyrstal_stanley_good_2018].
+Retired Army Brigadier General Ty Seidule similarly, in 2020, advocates for the renaming of several Army installations [@seidule_ty_what_2020].
+Other examples of observations coded as potentially violating military professionalism based on the topic of the publication include articles penned by Retired Navy Admiral Mike Mullen (entitled "The Refugees We Need") and co-authors Retired Army Lieutenant General Mark Hertling and Retired Naval Admiral Robert Natter (entitled "Cutting Refugee Admissions Will Have Severe Consequences for the US Military").
 
-These articles are not unreasonable in that they, in the opinion of the author of this article, express reasonable thoughts and ideas to many of the nation's pressing problems, to include gun violence, the legacy of the Confederacy and our national history, and immigration policy.
-Moreover, many Americans agree with the stance these officers have taken in the articles they have published.
-One can even make a strong, perhaps even a compelling argument, why these officers might be uniquely qualified to address some of these topics, in that they possess expertise commanding large and powerful organizations charged to carry out violence (on the issue of gun control), have served in the nation's military (and know what it is like to, for example, lead servicemen and women who serve on bases named after Confederate generals), and have considerable experience serving alongside soldiers of multiple ethnicities as well as considerable experience in Iraq and Afghanistan (and thus, can speak about how immigration policy might impact the military).
+These articles are certainly not unreasonable in that they, in the opinion of the author of this article, express reasonable thoughts and ideas to many of the nation's pressing problems, including gun violence, the appropriate legacy of Confederate symbols, and national immigration policy.
+To be clear, many Americans agree with the stance these officers take in the articles they have published.
+One can even make a strong, perhaps even a compelling argument, why these officers might be uniquely qualified to address aspects of some of these topics in that they possess expertise commanding large and powerful organizations charged to carry out violence on behalf of the country (related to the issue of gun control), that they may have experience serving and leading troops on bases named after Confederate generals (related to the issue of Confederate symbols and bases named after Confederate generals), and that they have have considerable experience serving alongside soldiers of multiple ethnicities as well as considerable experience in Iraq and Afghanistan (related to the issue of improving American immigration policy).
 
-At the same time, however, one is also not wrong to ask whether these and other contentious and large issues are best considered to be political issues in that they are best solved by politicians, as opposed to retired military officers.
-The important point as is related to the concept of military professionalism is not that these officers should not have an opinion on these and other important topics, but whether they, as retired military officers, ought to be considered as appropriate mouthpieces on these issues by the American public.
+But I contend that it is also not wrong to insist that these and other contentious national issues are most appropriately solved by elected civilian officials and not by retired military officers.
 Interestingly, in July of 2020, General Mark Milley, the Chairman of the Joint Chiefs of Staff, was asked by US Representative Anthony Brown, D-MD, during a hearing held by the House Armed Services Committee, to "comment on the naming of Army installations after Confederate soldiers," and whether the continued use of the names of these bases "reflects the values" that the military is trying to instill in the nation's armed forces [@milley_mark_general_2020]?
@@
 He offered the Congressmen the idea that even as the Department of Defense continues to take a hard look at the issue, the decision to remove those names should reflect the decision to adopt those names in the first place in that such a decision was "a political decision" [@milley_mark_general_2020].
-It is not unreasonable to insist that policy arenas involving appropriate responses to mass violence, the future of Confederate symbols, and refugee/immigration policy for the nation belong to non-military elites.
+Milley's answer was the appropriate one, I contend, because it recognized that regardless of how he feels about this issue, the issue is not for him (or the military) to solve.
+Many other topics addressed by publications listed in Table 4, to include gun violence and immigration, fall into the same realm.
 
@@
 All but four of the observations in this cluster are coded as having potentially violated standards of military professionalism.
-Of note, one observation in this cluster entitled, "Fists Raised at West Point" and published in the Washington Post is actually a compendium of two anonymous letters, each one taking an opposite stance on the appropriateness of an incident in which a photograph at West Point's 2016 graduation ceremony captured a graduating female cadet raising a clenched a fist, an action which some saw as an innapproprite political gesture related to the growing Black Lives Matter movement [@unknown_fists_2016].
+Of note, one observation in this cluster entitled, "Fists Raised at West Point" and published in the Washington Post is actually a compendium of two anonymous letters, each one taking an opposite stance on the appropriateness of an incident in which a photograph at West Point's 2016 graduation ceremony captured a graduating female cadet raising a clenched a fist, an action which some saw as an inappropriate political gesture related to the growing Black Lives Matter movement [@unknown_fists_2016].
 
@@
 ```{r build plots, include = FALSE, echo=FALSE}
-plot1<- ggplot(data=adj2dataset, mapping=aes(x=PubYr, y=case_count)) + geom_point() + geom_smooth() + labs(x="Year", y="Number", color=NULL)
-plot2 <- ggplot(data=adj2dataset, mapping=aes(x=PubYr, y=prop_yr)) + geom_point() + geom_smooth() + labs(x="Year", y="Proportion", color=NULL)
+plot1<- ggplot(data=adj2dataset, mapping=aes(x=PubYr, y=case_count)) + geom_point() + geom_smooth() + labs(x="Year", y="Number", color=NULL) + geom_vline(xintercept = 1981, size=.2, colour="tomato") +
+  geom_vline(xintercept=1989, size=.2, colour="tomato") + 
+  geom_vline(xintercept=1993, size=.2, colour="dodgerblue") + geom_vline(xintercept=2001, size=.2, colour="tomato") + 
+  geom_vline(xintercept=2009, size=.2, colour="dodgerblue") +
+  geom_vline(xintercept=2017, size=.2, colour="tomato") +
+  geom_text(aes(x=1981, label="Reagan",y=3), colour="tomato", angle=90, vjust=1, text=element_text(size=9)) + 
+  geom_text(aes(x=1989, label="Bush 41",y=3), colour="tomato", angle=90, vjust=1, text=element_text(size=9)) + 
+  geom_text(aes(x=1993, label="Clinton",y=3), colour="dodgerblue", angle=90, vjust=1, text=element_text(size=9)) + 
+  geom_text(aes(x=2001, label="Bush 43",y=5), colour="tomato", angle=90, vjust=1, text=element_text(size=9)) + 
+  geom_text(aes(x=2009, label="Obama",y=5), colour="dodgerblue", angle=90, vjust=1, text=element_text(size=9)) + 
+  geom_text(aes(x=2017, label="Trump",y=-1), colour="tomato", angle=90, vjust=1, text=element_text(size=9))
+
+plot2 <- ggplot(data=adj2dataset, mapping=aes(x=PubYr, y=prop_yr)) + geom_point() + geom_smooth() + labs(x="Year", y="Proportion", color=NULL)+ geom_vline(xintercept = 1981, size=.2, colour="tomato") +
+  geom_vline(xintercept=1989, size=.2, colour="tomato") + 
+  geom_vline(xintercept=1993, size=.2, colour="dodgerblue") + geom_vline(xintercept=2001, size=.2, colour="tomato") + 
+  geom_vline(xintercept=2009, size=.2, colour="dodgerblue") +
+  geom_vline(xintercept=2017, size=.2, colour="tomato") +
+  geom_text(aes(x=1981, label="Reagan",y=.25), colour="tomato", angle=90, vjust=1, text=element_text(size=9)) + 
+  geom_text(aes(x=1989, label="Bush 41",y=.25), colour="tomato", angle=90, vjust=1, text=element_text(size=9)) + 
+  geom_text(aes(x=1993, label="Clinton",y=.25), colour="dodgerblue", angle=90, vjust=1, text=element_text(size=9)) + 
+  geom_text(aes(x=2001, label="Bush 43",y=.5), colour="tomato", angle=90, vjust=1, text=element_text(size=9)) + 
+  geom_text(aes(x=2009, label="Obama",y=.5), colour="dodgerblue", angle=90, vjust=1, text=element_text(size=9)) + 
+  geom_text(aes(x=2017, label="Trump",y=.55), colour="tomato", angle=90, vjust=1, text=element_text(size=9))
+
 ```
@@
 ```{r figure 3, include=TRUE, warning= FALSE, message = FALSE, echo=FALSE}
-grid.arrange(plot1, plot2, ncol=2, top="Figure 3: #/Proportion of Op-Eds Published\n by Retired Military Officers, 1979-2020", bottom="On Topics of Social Policy which Impacts the Military,\n Domestic Policy, and Political Administrations/Figures")
+grid.arrange(plot1, plot2, ncol=2, top="Figure 3: Number(L) and Proportion (R) of Op-Eds Published\n by Retired Military Officers on Non-Traditional Military Topics, 1979-2020***", bottom="*** Topics include: Social Policy which Impacts the Military,\n Domestic Policy, and Political Administrations/Figures")
 ```
@@
 Changes in the institutional credibility of the American military institution can be seen by examining public opinion data.
-Pollsters survey the American public annually to gauge their trust in a variety of instutions.
+Pollsters survey the American public annually to gauge their trust in a variety of institutions.
 Using Gallup opinion polling, one sees that since the late 1970s, the credibility of the American military as an institution has gradually increased [@gallup_polls_confidence_2020].
 In 1979, 54% of those surveyed expressed either a great deal or quite a lot of confidence in the military, while in 2020, 72% expressed the same [@gallup_polls_confidence_2020].
-These trends in political polarization and confidence in the military are displayed in Figure 4, below.
+These trends in political polarization and confidence in the military are displayed in Figure Four, below.
 The top and middle graph show changes in means between Republicans and Democrats in the House of Representatives (top graph) and Senate (middle graph) since 1979, while the bottom graph shows changes in the percentage of surveyed Americans who reported either a great deal of or a lot of confidence in the military institution over the same time period.
@@
   labs(
-    title = "Figure 4: Distance Between Party Means in the House,\n the US Senate,\n and American Confidence in the Military, 1979-2020"
+    title = "Figure Four: Distance Between Party Means in the House,\n the US Senate,\n and American Confidence in the Military, 1979-2020"
   )
@@
 
-While it is possible that the trend in increasing instititional credibility of the US military over the past four decades may be at least partially responsible for the increasing trend in the number of opinion pieces published by retired military officers, a trend that the data supported earlier as shown in Figure 2, rising institutional credibility alone does not adequately explain the trends shown in Figure 3: that there has been a sharp increase over the past four decades in retired military officers penning pieces that address non-traditional military topics that lay outside the bounds of relevant military expertise, or pieces that express partisan sentiment.
+While it is possible that the trend in increasing institutional credibility of the US military over the past four decades may be at least partially responsible for the increasing trend in the number of opinion pieces published by retired military officers, a trend that the data supported earlier as shown in Figure 2, rising institutional credibility alone does not adequately explain the trends shown in Figure 3: that there has been a sharp increase over the past four decades in retired military officers penning pieces that address non-traditional military topics that lay outside the bounds of relevant military expertise, or pieces that express partisan sentiment.
 The theory laid out earlier in this paper predicted that as political polarization increases, three mechanisms are at work which have strong impacts on the military professional: it becomes increasingly difficult to remain politically neutral as partisan conflicts subsume a greater number of issues; as this occurs, strong feelings of dislike or disdain surface against those who hold on opposing view; and alterations in one's sense of loyalty arise, particularly as issues that individuals hold to most dearly raise the salience of one's identity.
 
-To test if changes in political polarization adequately explain the changes in the qualitative nature (increasing trends of potential violations of military professionalism) of opinion pieces authored by retired military officers, I conduct a logit regression of the variable _questionable_professionalism_ while holding institutinal trust in the military constant at its mean (71 percent) and varying the value of the mean ideological distance between both major political parties in the House between 1979-2020.
+To test if changes in political polarization adequately explain the changes in the qualitative nature (increasing trends of potential violations of military professionalism) of opinion pieces authored by retired military officers, I conduct a logit regression of the variable _questionable_professionalism_ while holding institutional trust in the military constant at its mean (71 percent) and varying the value of the mean ideological distance between both major political parties in the House between 1979-2020.
 Conducting such a regression allows one to examine, given the data assembled, the predictive likelihood of publishing an opinion piece that violates military professionalism given varying levels of political polarization and constant institutional credibility in the military.
@@
 
-slimdata <- polaradjdataset[1:334,] %>% 
+slimdata <- polaradjdataset[1:344,] %>% 
   view()
@@
 
-The graph, which is modeled off of predicted data using a logit regression, clearly shows a non-linear relationship between increasing political polarization and the likelihood that an opinion piece authored by a retired military officer will potentially violate aspects of military professionalism.
-While the confidence intervals at each point estimate are relatively large, the graph clearly shows that the predicted probability of violating military professionalism is different between extremes along the x axis.
-In other words, the political polarization that exists today, as measured by the ideological distance between the political parties in the house, is predicted to have resoundingly different impacts on military professionalism than the polarization levels that existed several decades ago.
-Whereas the polarization levels that existed several decades ago may have had negligible, if any impact, on the likelihood of a retired military officer violating military professionalism in the content of a published opinion piece, the polarization levels that exist today have, at least insofar as the predictive power of the model is concerned, changed the landscape: today's polarization levels reveal a roughly 15% chance of violating military professionalism, with confidence intervals that range from roughly 9% to nearly 23%.
+The graph clearly shows a non-linear relationship between increasing political polarization and the likelihood that an opinion piece authored by a retired military officer will potentially violate aspects of military professionalism.
+While the confidence intervals at each point estimate are relatively large, the graph demonstrates that the predicted probability of violating military professionalism is different between relative extremes along the x axis.
+In other words, the political polarization levels that exist today, as measured by the ideological distance between members of the two prominent political parties in the house, is predicted to have resoundingly different impacts on military professionalism than the polarization levels that existed several decades ago.
+The polarization levels that existed several decades ago, in the late 1970s and 1980s, may have had negligible, if any impact, on the likelihood of a retired military officer violating military professionalism in the content of a published opinion piece, but the political landscape is simply different today.
+Contemporary polarization levels, insofar as the predictive power of the model is concerned, changed the environment: at contemporary polarization levels (the far right of the x axis), there is roughly a 17% chance of a retired military officer's published opinion piece violating the standards and norms of military professionalism, with confidence intervals that range from roughly 11% to nearly 25%.
+
+Given these dynamics, as well as the results of both Figures 3 and 5, ample evidence exists to confirm the second hypothesis, which stated that as a result of increasing political polarization, there will be an increase in recent years in the number of op-eds written about topics that are not relevant to military expertise, or that reveal inappropriate partisan sentiments, among authors which are a vital component of the institutional military, retired military officers.
 
@@
 
-Before concluding, it is worth commenting about why Figure 3 revealed a sharp drop in the number and proportion of observations that were written about topics such as social policy which impacts the military, domestic policy, and political administrations/figures between roughly 2005 - 2011.
-One compelling possibility is that between these years, the military was heavily engaged in combat operations overseas to a point that the institution was taxed tremendously: repeated deployments, surges of forces to both Iraq (2007-2008) and Afghanistan (2010-2011), and sustained US casualties from these conflicts.
-This is not to suggest that the military was not engaged in conflicts around the globe in other periods of time, but to instead suggest that even retired military officers will be impacted by the institutional requirements of the active military.
-When such requirements include repeated deployments and an operational tempo that spans multiple years, perhaps retired military officers acknowledge that their voice is best served advancing these interests, as opposed to weighing in on other topics.
+Why did Figure 3 reveal a sharp drop between approximately 2005-2011 in both the real number and proportion of annual opinion pieces authored by retired military officers that address non-traditional military topics, such as social policy that impacts the military, domestic policy, and political administrations/figures?
+What was it about this time period that made it unlike any other span of years across the past four decades?
+
+One compelling answer is that between these years, roughly 2005-2011, the military was heavily involved in combat operations to a point such that the institution was taxed tremendously, thereby necessitating that it remain focused on its wartime mission.
+Repeated deployments during the Global War on Terrorism, which included substantial surges of forces to both Iraq (2007-2008) and Afghanistan (2010-2011), as well as sustained US casualties resulting from these conflicts, possibly shaped the military institution so strongly such that even retired military officers acknowledged that the military's institutional energy required a laser-sharp focus on combat requirements.
+This is not to suggest the military was not heavily engaged in conflicts nor active around the glove at other times throughout the past four decades (which we know is not the case), but to instead suggest that even retired military officers are perhaps more sensitive to combat requirements and missions undertaken by the military as higher levels of continued operational strain and duress are endured.
+Though more data is required to fully test this argument, other scholars have argued similarly in that periods of transition, especially away from major combat operations and into eras of downsizing and relative peace, often seem to test the very essence of military professions in significant ways [@snider_once_2012].
 
@@
 The goal of this paper was to explore and examine the ways in which political polarization and institutional credibility have impacted the professionalism of the military over the past fifty years.
-Examining the published opinion pieces authored by retired military officers in major US newspapers revealed that together, institutional credibility and political polarization have generated changes in the professionalism of some retired military officers, and more broadly, in how military elites communicate with the American public.
-Though the data examined demonstrates that the vast majority of op-eds published by retired military officers in the past fifty years exhibited professional standards, the data also showed two worrisome trends: that retired military officers are arguing about topics that fall outside of the traditional umbrella of national security, and that they are expressing clear partisan preferences at a greater rate than they previously did.
-Moreover, both of these trends have increased under the past four years of Donald Trump's Presidency.
-These trends are captured in Figure 4, below, which shows the number (left) and proportion (right) of opinion pieces authored by retired military officers over the past 40 years that addressed issues of social policy, domestic policy, or specific political figures or administrations.
-The graphs tell an interesting story: the number and proportion were both basically zero 40 years ago, then increased slightly until about 2003, at which point, both decreased until roughly 2011.
-After that, both the number and proportion have risen sharply.
-Nearly 40% of all opinion pieces in 2019 published in major US newspapers by retired military officers were written about topics that fell outside of traditional national security topics.
+Examining the published opinion pieces authored by retired military officers in major US newspapers revealed that although the vast number of total opinion publications uphold the military's norms and traditions concerning military professionalism, a couple of worrisome trends are evident in the data: first, retired military officers are increasingly writing about topics outside of the umbrella of traditional national security topics, and second, retired officers are expressing partisan preferences at a greater rate than they previously did.
+What do these trends mean, and should they cause concern?
+If so, what should be done about it?
 
-This paper also developed a theory for how the interaction of increasing institutional credibility and increasing political polarization has potential to harm the professional military ethic.
-Maintaining credibility and trust with the American public remains crucial for the military, but should a polarized political climate endure, which is suggested by the majority of American politics scholars who study polarization, the military needs to also understand that credibility and trust can be abused if it is handled inappropriately - a legitimate risk in a polarized environment.
-This is especially likely if military professionals, to include retirees, abuse the high level of credibility they currently enjoy to take up and advocate for partisan positions, or worse, to jockey with civilian leaders for power [@brooks_shaping_2008-1].
-Unfortunately, as this paper illustrates, there are hints of this occurring.
-The letter signed by concerned graduates of West Point, which was clearly aimed at Secretaries Pompeo and Esper, and discussed at greater length earlier in this paper, is clear evidence of this.
+First, these findings illuminate the power and pervasiveness of increasing political polarization, and in particular, its impact not only on political institutions, but also on professions, which the military prides itself on being.
+Thus, if the military, which has for decades adopted and promulgated norms of apolitical and nonpartisan expression, has been impacted by rising political polarization, it stands to reason that other political institutions and professions also have been impacted.
+While the specific manifestations of political polarization's increase and spread in non-military institutions and professions might be different, the fact that the military is experiencing a noticeable impact as a result of increasing political polarization speaks to its significant power and influence.
 
+Indeed, evident in the content of opinion pieces authored by retired military officers over the past four decades are the three features of contemporary American political polarization developed in the theory proffered earlier in this paper: the loss of politically neutral space as more issues enter the partisan battleground, the creation of strong feelings of dislike for those who hold to opposing views on these issues, and perhaps worst for the military, alterations in an individual's concept of loyalty, which arises as the salience of one's identity is raised as the issues that an individual holds most dearly to are increasingly at stake.
+Moreover, the data supported what most scholars of American politics have argued for: a realization that polarizing politics across America's political landscape will likely not disappear anytime soon, regardless of which party holds the nation's highest office.
+
+Second, these findings underscore the challenge that the military as a profession has in "reigning in" its retired officers and compelling them to stay out of partisan battles, and out of politics more broadly.
+Among the features and traits of a profession is that it is self-policing; however, the data indicates that current efforts by active military leaders to police and clean up the behavior of its retired elites is not working.
+Indeed, the letters and pleas of previous military leaders to retired military officers, including those written by now Retired Army General Martin Dempsey and Retired Marine Corps General Joseph Dunford, both of which were referenced in this paper, are simply not working.
+This suggests that stronger and more creative efforts are required.
+
+Third, these findings demonstrate that the phrase "politicization of the military," which has long been a proper concern of academics, journalists, the military, and citizens alike, should not be intellectually confined to notions of how politicians or other political leaders political leaders attempt to harness or exploit the military for political purposes or advantages.
+Though such efforts do occur and are a genuine cause for concern that continue to deserve both scholarly and practical attention, the evidence presented in this paper shows that "politicization of the military" is also occurring at the hands and through the words of particular retired military officers.
 Gaining a better understanding of the impacts of political polarization on the military profession remains a critical area for future research.
-Other scholars such as Michael Robinson have explored the specific ways that polarization is impacting the interactions between the military and the public [@robinson_who_2019].
-Though this paper focused on the ways that polarization may impact professional military behavior, it recognized but stopped short of discussing polarization's subsequent impact on various "military outcomes," which may arguably include more decisive events, such as achieving victory or defeat in battle, or the ability to formulate sound strategic advice.
+Other scholars such as Michael Robinson have explored the specific ways that polarization is impacting the interactions between the military and the public [@robinson_michael_danger_2018].
+Though this paper focused on the ways that polarization may impact professional military behavior, it recognized but stopped short of discussing polarization's subsequent impact on various "military outcomes," which may arguably include even more decisive events, such as achieving victory or defeat in battle, or the ability to formulate sound strategic advice.
 These outcomes and more, and the role of polarization in influencing them, deserve to be explored further.
 
-Finally, if one believes that today's polarized climate in America is different in degree and kind than previous eras of polarization, which has been a supporting claim of this paper, one must recognize that the military will face continued institutional challenges stemming from polarization's underlying causes, and so will individual military leaders.
-The most portending aspect regarding today's polarizing climate is that fundamental values are at stake, values that speak directly to the heritage of the United States as one nation under God and to the military as the nation's guarantor of individual rights and freedoms.
-As a non-partisan institution that should theoretically be free from partisan entanglement, the actions of the military will remain under close scrutiny by the American public.
-In this type of environment, in some sense, the military is doomed: regardless of how the institution or its leaders act or respond to pressing polarized debates, some will accuse it of failing to do its duty and to exercise genuine loyalty.
-It is during these tough moments that another soldierly virtue - moral courage - will perhaps stand out as the chief attribute of military leaders, now and in the years to come.
+Maintaining credibility and trust with the American public is crucial for the military, but should a polarized political climate endure, the military needs to also understand that credibility and trust can be abused to the point that they also can constitute significant risks.
+One such risk is that as polarization worsens, military leaders, to include military retirees, abuse the high level of credibility they currently enjoy with the American public to advocate for obviously partisan positions.
+Such actions, though, in a polarized setting, might actually result in an increase in the credibility of the military institution, at least among citizens who share the same partisan sentiments advocated for by military leaders, which might, somewhat perversely, entice more of the same behavior.
+A second risk, stemming from the first, is that a polarized setting sets the stage for the military institution to eventually openly jockey with civilian leaders, either for power or for control, a concern that generations of scholars have warned us about [@finer_man_1962; @brooks_shaping_2008-1].
+Were either of these risks to materialize, the character and nature of the American military as an institution, and likely, American democracy, would be irrevocably changed.
+
+There are no simple solutions for the military with respect to the risks posed by rising political polarization.
+Two possible ideas, however, are apropos to this discussion.
+One involves identity, and the other, credibility.
+First, our nation's military is composed of soldiers who come from every walk of life throughout America, and accordingly, in the years to come, many future soldiers and leaders will bear the features of a polarized society.
+Thus, the military's challenge, and one which it must embrace itself for immediately, is integrating the multiple identities each soldier carries with them as people in such a way that each can still proudly wear the uniform and bear the identity of an American service member.
+Though this challenge is not new for the US military, increasing political polarization has significantly raised the level of difficulty the military faces in tackling such a vital challenge.
+Scholars who have studied multi-ethnic militaries across the world have argued that it is military leaders, not civilians, who ultimately devise effective solutions to the problems posed by soldiers who carry multiple loyalties, and interestingly, only as a result of combat requirements [@peled_question_1998].
+Perhaps the American challenge is even greater in that at least for the near term future, there is no major war that will necessitate the coming together of disparate loyalties and identities among future soldiers.
+
+The second idea, and one that may sound like anathema to modern military leaders, is that perhaps the level of trust and credibility between the American public and the military institution should not be the ultimate indicator of the military's performance as a profession.
+In suggesting this, I am not suggesting that credibility and trust are not important or that the concept should be disregarded entirely.
+The American public is, after all, the client of the American military and ultimately, it is their opinion that counts.I am suggesting, however, that for the reasons laid out in this paper, credibility and trust are influenced by the mechanisms through which polarization works, to include the loss of politically-neutral space, increased negative feelings toward those who hold differing opinions, and most sinister of all, alterations in one's concept of loyalty.
+For these reasons, I am simply offering the notion that _credibility can be deceiving_ in polarized times.
+Now, more than ever before, military leaders will need to rely on individual moral courage over and above any other trait in order to perform their duties as military professionals.
 \singlespace

### 2020-09-21 code/prospectusv1.Rmd
commit

---
title: "The Limits of Loyalty: Postmodernism, Polarization, and Military Professionalism  (Prospectus)"
author: "Peter M. Erickson"
date: |
  `r format(Sys.time(), '%B %d, %Y')`
abstract: |

  _This dissertation seeks to explore the relationship of postmodernism and political polarization and their impact on the US military profession. The fundamental question that I seek to answer is, "how is rising political polarization shaping the military as a profession in the United States, and what are the implications of these changes, both in terms of national security and the future of the military profession?" In answering this question, I develop a theory that, at its core, argues that the nature of rising political polarization fundamentally challenges the conception of loyalty among members of the military profession. I further argue that the strain of loyalty is especially problematic for military, as it is a profession that by its nature holds mutiple loyalties simultaneously: to the Constitution, to the President as Commander-in-Chief, to Congress, and to the American People. Balancing these competing loyalties becomes more difficult, I argue, as political polarization worsens, because of the fundamental values - not mere policy differences - that are at the root of many polarized political battles today. In the dissertation, I seek to employ a mixed-methods approach that incorporates several case studies, as well as quantitative analysis, to support my theory. This dissertation will contribute to the fields of American politics, international security, and civil-military relations by further illuminating the mechanisms through which political polarization is impacting the military institution. Specifically, it will provide a contrarian view to recent civil-military relations scholars who have argued that the traditional concept of military professionalism should be significantly redesigned, if not abandoned entirely. This dissertation will argue the opposite: that what is needed most in politically polarizing times, such as those the US is in right now, is a return to and embrace of traditional conceptions of military professionalism - an effort that will require deliberate action by not just military leaders, but elected civilian leaders as well._   
geometry: margin = 1.15in              # margin widths
fontsize: 12pt                         # text size
bibliography: researchpaper-ps904.bib # bibliography file
biblio-style: authoryear               # citation style
urlcolor: blue                         # links to external URLs
citecolor: black                       # links to bibliography
linkcolor: magenta                     # internal figures, footnotes, etc.
indent: true                          # paragraph format
                                       # true for indented paragraphs

header-includes:
    - \usepackage{setspace}
    - \usepackage{multirow}
    - \usepackage{indentfirst}

output: 
  bookdown::pdf_document2:
    keep_tex: true
    latex_engine: pdflatex
    number_sections: false
    toc: false
    fig_caption: true
    citation_package: biblatex
    includes: 
      in_header: 
        - latex-preamble.tex
  bookdown::htmldocument2: default
  bookdown::word_document2:
      toc: false
---

\doublespacing

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include = FALSE, echo = FALSE}
library("here")       # file paths
library("tidyverse")  # workhorse package
library("tidylog")
library("kableExtra")
library("knitr")
library("ggdag")
library("dagitty")
library("gridExtra")
library("broom")
library("dplyr")
library("reshape2")
```
\newpage

## Introduction and Research Question

In late May and early June 2020, racial tension and concerns about police brutality gripped the American public following the death of George Floyd, an unarmed black man, by police in Minneapolis. In many cities and communities across the nation, peaceful protests were followed in some cases by periods of rioting and looting. On June 1, after giving remarks at the Rose Garden, President Donald Trump walked to St. John's Church after security forces visibly displaced throngs of protesters in neighboring Lafayette Square. Flanked by members of his cabinet, to include the Secretary of Defense, and with the Chairman of the Joint Chiefs of Staff not far behind, the President held up a Bible at the church, stared into the cameras of the media, and, several minutes later, returned to the White House.

In the days and weeks that followed, a full-blown civil-military relations crisis ensued  [@feaver_military_2020]. The Secretary of Defense, Mark Esper, publicly disagreed with a threat made by President Trump to invoke the Insurrection Act of 1807 to use active-duty forces to quell domestic riots [@esper_secretary_2020], and General Mark Milley, the Chairman of the Joint Chiefs of Staff, told senior military officers graduating from the National Defense University that he "should not have been there (accompanying the President on the walk to St. John's Church)", as his presence "created a perception of the military involved in domestic politics" [@milley_official_2020].

Several retired military officers also weighed in, not only on the Trump administration's handling of specific events near the White House, but on a host of broader topics that dominated the popular dialogue in America after the death of George Floyd, to include the state of American racial relations. Upon paying close attention to what these retired military officers said, though, one can notice that there was in fact fairly substantial variation in both the substance and tone of their remarks. While all of them expressed concerns about the President potentially ordering active-duty military forces to confront what many Americans believed were peaceful protesters, and thus, violating Constitutional protections afforded by First Amendment [@brooks_dismay_2020; @dempsey_former_2020], some officers went much further in their criticisms. Indeed, some of these retired military officers accused the President of failing to possess the overall fitness to lead [@goldberg_james_2020], promote harmonious racial policies  [@allen_moment_2020], and demonstrate adequate leadership abilities [@mullen_i_2020]. In the months following George Floyd's death, moreover, several retired and currently-serving military officers have further made public arguments on a variety of topics that are directly related to President Donald Trump, including race in America [@askew_anti-racist_2020] and voter fraud as it relates to the upcoming Presidential Election [@nagl__2020].

While America has witnessed numerous incidents in its history in which tension between civilian and military elites has been present, the events of the Summer of 2020 between the Trump administration and some - though certainly not all - retired  military officers raises important questions, particularly I argue, because the context did not involve combat overseas.  These questions include: why did several retired military elites speak out so strongly against President Trump? Why did some of these elites confine their criticism to concerns over the Constitution, while others sounded more like representatives of the Democratic Party? Are these events that occurred during the Summer of 2020, and the military response generated in response to these events, an isolated chapter in American history, or is there something more disturbing underneath the surface regarding contemporary American civil-military relations?

This dissertation seeks to explore the impacts of rising domestic political polarization on the US military as an institution and profession. The primary argument espoused in this dissertation is that certain features of the polarized environment impact how military officers view and exercise the concept of loyalty, a claim that I argue holds significant ramifications for US national security. Using a mixed-methods approach, I develop a theory and then test it using a mixture of quantitative analysis and several qualitative case studies.  I conclude with discussing the implications for such an argument.

# The Scholarly Literature and Why This Question Matters

The question of how political polarization affects changes in the concept of military professionalism spans several related literatures, most notable of which are American Politics and International Relations. Broadly speaking, this project is focused on how one aspect of domestic politics, political polarization, impacts a specific but important aspect of international relations, civil-military relations, which itself spans a wide and diverse literature. Because I intend to focus my work on the military institution, my dissertation will also rely on literature from the fields of sociology, mainly regarding the development of organizations and professions, and ethics, to include military ethics.

Within the American Politics literature, there is broad recognition that political polarization levels in contemporary America are higher now than at any point since the end of the American Civil War [@hare_polarization_2014], but there are multiple explanations for what has caused this trend, ranging from a growing divide over matters pertaining to race and religion [@abramowitz_great_2018], the size and role of government [@sides_two_2015], and disagreements over specific policy issues [@mann_its_2012]. Despite the herculean task of disentangling the roots of the causes of polarization, virtually all scholars agree that political polarization is a genuine problem and should spark concern.

One of these major concerns is the impact of polarized domestic politics on American foreign policy in a broad sense, and specifically, on America's ability to lead and impact the world. Kenneth Schultz, for instance, underscored four challenges that have emerged as a result of America's polarized domestic politics, including the ability to obtain bipartisan support for "risky undertakings," the ability to learn the "lessons of foreign policy failure," the inability to "make long term commitments to allies and adversaries" as a result of "dramatic policy swings from one administration to another," and the "vulnerability" of America's electoral system to foreign third party adversaries [@schultz_perils_2017]. Because the military is a central actor in shaping America's foreign policy, obtaining a better understanding of how political polarization shapes various military institutions specifically will yield findings that may bear on helping policymakers and academics help solve each of the problems Schultz identifies.

Thankfully, a small but growing number of civil-military relations scholars have looked at aspects of this question in recent years. But the focus of these works has been, I believe, not on the impact of polarization on the military itself but on other important but peripheral components of the military's relationship with American society. One recent dissertation completed by Michael Robinson at Stanford University in 2018, for instance, found that partisan affiliation significantly affects public credibility and trust in the military perhaps resulting in or as a consequence of media "echo chambers" which may incentivize military leaders to offer opinions partisan opinions without fear of sanction [@robinson_danger_2018]. Robinson's work builds off of previous work by Heidi Urben in which she found that active military officers believe less strongly that retired military officers should remain apolitical or non-partisan in retirement [@urben_party_2013], a norm that has existed for decades or more. Together, Robinson and Urben join other scholars in providing a wealth of information regarding public opinion and the military, and in particular, how aspects of political polarization shape partisan differences regarding topics such as the use of force [@golby_jim_elite_2017] and military elite endorsement of political candidates [@golby_military_2012].

In their own right, these scholarly developments are interesting and important, but another debate in the civil-military relations field has recently emerged, or to be more correct, reawakened. At the center of this debate is the degree to which the concept of military professionalism should be altered if not discarded altogether [@brooks_paradoxes_2020]. The debate is not new because in actuality, it began as soon as Samuel Huntington published his seminal work in 1957, _The Soldier and the State_. While I do not lay out the full synopsis of the book nor the the various critiques here explicitly, the central question in this debate is whether the concept of military professionalism, which Huntington espoused as essential to achieving military security during the Cold War and as the US military has largely inculcated among its officer corps for the last couple of generations, separates the military too much from civilians, leading to gaps in the development and execution of strategic outcomes [@rapp_civil-military_2015; @dubik_just_2016; @brooks_paradoxes_2020]. These criticisms largely point to the American failures in Iraq and Afghanistan to produce and implement strategy and achieve a better and lasting peace - criticisms which this author, for what it is worth, shares. But I contend that something is missing in this debate, namely, a thorough exploration and theoretical outworking of the impacts of political polarization on the military institution.

Why does this matter? I contend that the impacts of contemporary political polarization on civil-military relations and on the military institution specifically matters and should be studied more thoroughly because, as I intend to show, portions of the military institution, though not all, have started to argue and disagree with civilian leaders in ways that are substantively different than previous civil-military feuds. To be sure, military and civilian relationships are in tension by their very nature in that the leaders of a democracy allow a most undemocratic institution, the military, to protect the polity [@feaver_civil-military_1996]. And as American history proves, there have been countless instances in which civilian and military leaders disagreed. Many of these rise to the level of common knowledge: Lincoln fired McClellan because he lost confidence in him as a wartime Commander; Truman sacked MacArthur because the latter was insubordinate regarding the future of war in Asia; and Obama and McChrystal clashed about troop levels in Afghanistan.

Beyond these episodes in which President and wartime commander locked horns, consider several other instances of "civilians" and "military" clashing. One episode is certainly the way in which the Vietnam war escalated in ways that, at least according to H.R. McMaster, military generals at the time knew were destined to fail [@mcmaster_dereliction_1998]. A more recent example, and one much studied and commented upon by civil-military relations scholars, is the so-called "Revolt of the Generals" that occurred following the early years of the invasion of Iraq, which involved several recently-retired generals and admirals sharply criticizing Secretary of Defense Rumsfeld's leadership dynamics [@cook_revolt_2008]. The point here is that civil-military conflict is anything but novel, but that when it occurs, it is usually, for understandable reasons, involving a clear national security issue, such as a war, combat operations, or items of a related sort, such as military budgets, weapons programs, etc.

While these types of disagreements are important and will certainly feature in my intended dissertation (I hope!), my motivation in choosing my intended topic is because we are now seeing portions of the military institution (particular retired military officers) openly disagree with civilian leaders on topics that I believe are, even by the widest of academic definitions, way outside of appropriate military expertise. In other words, there is a world of difference between retired generals criticizing a President for conduct of a war, but entirely another when generals criticize (or applaud) the President for his domestic policies that have little if anyting to do with national security. I contend that the latter is taking place, and that the reasons for such developments, as well as their implications, deserve to be further explored. I further believe that while such developments are catching the attention of civil-military relations scholars, most of the work done on this topic is impressionistic [@golby_jim_keep_2020] as opposed to systematic. In other words, civil-military relations scholars have identified that polarized partisan politics serve as a risk for the military, but few have explored in a systematic way what these risks are, how serious they are, the mechanisms through which they work, and what the implications are for the future of the military. I hope to fill this gap. In doing so, I hope to also contribute to the reawakened debate regarding military professionalism and its role vis-a-vis the American military institution.

## Polarization and Military Professionalism

how increasing domestic political polarization impacts conceptions of loyalty in military professions, and to examine the implications of these changes in loyalty on broader security outcomes. For at least the past 75 years[need footnote here], the US military has prided itself on being not merely an organization charged with the defense of the nation, but a profession. That is, on being an institution marked by several unique traits, including: providing a unique service that requires special skills (the application and management of violence); establishing and maintaining standards of behavior and performance, as well as policing itself when members of that profession fail to uphold these standards; and building and maintaining trust between itself, the profession, with the client whom it serves (the American people) [insert source here]. The US military, further, is both by statute and by normative standards an apolitical and non-partisan institution with multiple, simultaneous loyalties: to the Constitution of the United States; to the President of the United States as the Commander-in-Chief; and to Congress. But what happens if and when the objects of the military's multiple loyalties not merely disagree, but become severely divided ideologically? Can such an outcome exist, and if it can, is there empirical proof that such a trend is occurring? If so, what are the normative implications, both for the military as a profession, and for the nation's security, and are there actions that can be taken to halt such a trend?

## The Concept and Structure of Loyalty

Figure One, below, shows one possible way to conceptualize the system of problems generated by increasing political polarization and its impact on the military profession, as well as the roles of other structural variables, including the institutional credibility of the military profession, and other various military outcomes. From left to right, this conceptual model suggests that aspects of political polarization are simulataneously driving changes in the institutional credibility of the military as well as changes in the professional character/nature of the military. This is made more complicated, as the model suggests, that the intervening variable of institutional credibility is also driving some changes in the professional character of the military.

In my dissertation, I situate changing conceptions of loyalty as a major subcomponent of the arrow running from political polarization to changes in military professionalism. At the same time, I do think that political polarization is impacting the credibility of the military institution, which in turn may likewise be driving changes in military professionalism. The specific mechanisms through which this is occurring, if at all, however, need to be developed, proposed, and tested.

```{r figure one, include=FALSE, echo=FALSE}
problem_dag <- dagify(
    MO ~ MP,
    MP ~ PP + IC,
    IC~PP,
    coords = tribble(
      ~ name,      ~ x,    ~ y,
      "MO", 1, .5,
      "MP", .7, .5,
      "IC", .3, .2,
      "PP", .1,.5,
    ))

problem_dag %>% 
  ggplot(aes(x=x, y=y, xend=xend, yend=yend))+
  geom_dag_edges()+
  geom_dag_point()+
  geom_dag_text_repel(aes(label=name), show.legend=FALSE)+
  theme_dag()
```

```{r include=TRUE, echo=FALSE, fig.height=5, fig.width=7}

problem_dag %>% tidy_dagitty() %>% 
  dag_label(labels=c("MO"="Changes in\n Military\n Outcomes", 
                     "MP"="Military\n Professionalism", 
                     "PP"="Increasing\n Political\n Polarization", 
                     "IC"="Institutional\n Credibility")) %>% 
  ggplot(aes(x=x, y=y, xend=xend, yend=yend))+
  geom_dag_edges()+
  geom_dag_point()+
  geom_dag_text()+ 
  geom_dag_label_repel(aes(label=label, fill=label), col="white", show.legend=FALSE)+ theme_dag() + labs(title="Figure One", subtitle="The Impact of Political Polarization on Military Professionalism")

```

At this stage in the dissertation phase, I do not know how polarization is impacting conceptions of loyalty, or even if it is, for sure. I strongly suspect that it is, however, for a couple of reasons. First, the military is, I believe, susceptible to charges of failing to uphold the ideal of loyalty based on the nature of the organization in that it ultimately is charged to enact violence on behalf of the state. Harold Lasswell said of military officers that their skill is the "management of violence" [need quote from Huntington]. As the bonds of loyalty forged on the battlefield are strong and different in ways than loyalty forged in other environments, I suspect that the military and its leaders will take very seriously any accusations that the corporate entity or individual officers are failing to be loyal to a number of ideals. In the Army, for example, the first "Army Value," as it is called by the militar service,  which is summarized as the acronym, "LDRSHIP" (short for "leadership"), is the word "loyalty." [need cite for the Army values here].

A second reason that I suspect polarization is impacting the conception of loyalty among the military professions stems from some of the mechanisms that we know regarding how political polarization operates. Politics of American scholars, though not united in agreement regarding the causes of polarization, have agreed that polarization is working through at least three mechanisms, all of which, I believe, impact the military in its normative role qua non-partisan and apolitical American institution. First, there is a sharp increase in the number and scope of issues that have been subsumed by partisan debate and disagreement (a phenomenon referred to among scholars of American Politics as "conflict extension" [@layman_party_2002]). A second mechanism exists through the increase in the strength of negative feelings held by one partisan group towards those of the other side, a phenomenon referred to as "negative partisanship" [@abramowitz_great_2018]. Third, there is an increase in the salience of debates over issues that individual partisans consider core to their individual identities [@white_values_2002; @mann_its_2012; @sides_identity_2018]. Each of these three features of the politically polarized landscape strain the concept of military professionalism in important ways, and taken together, comprise a significant challenge to the individual military professional and thus, to the professional nature of the military as a whole. These challenges are theorized in Table 1, below.

\singlespacing
\begin{table}[ht]
\begin{footnotesize}
\caption{Mechanisms of Rising Political Polarization Which Impact Military Professionalism} % title of Table
\centering % used for centering table
\begin{tabular}{|p{5cm}|p{5cm}|p{5cm}|p{5cm}|}
    \hline
    \textbf{Features of Increasing Political Polarization} & \textbf{Associated Mechanism} & \textbf{Potential Impact on Military Professionalism}  \\ \hline
    \textbf{Conflict Extension} & Less politically-neutral space exists as partisan disagreements subsume a greater number of issues. It becomes increasingly difficult to remain "out of politics." & Military Professionals face an increasingly difficult prospect of remaining apolitical. \\ \hline
    \textbf{Negative Partisanship} & Strong feelings of personal dislike and disdain arise against those who hold to opposing political views. There is a belief that those on the "other side" are greatly wrong and flawed. & There is a greater willingness to confront or take issue with those who hold opposing views.\\ \hline
    \textbf{Core and dearly-held issues are increasingly at stake} & As the issues to which one holds most dear are at stake, the salience of one's identity and sense of belonging to a particular group is raised. Associated with this phenomenon is a belief that not just preference, but moral right and wrong, are at stake. & Military professionals face potential changes in one's concept and notion of loyalty. There is a sense that failing to take action over a certain issue is equivalent to failing to uphold a sense of loyalty to an important group, idea, or concept.\\ \hline
\end{tabular}
\label{table:comparison} % is used to refer this table in the text
\end{footnotesize}
\end{table}
\doublespacing

The first way in which the concept of military professionalism is challenged as a result of increasing political polarization is that there is increasing difficulty for individuals to remain politically neutral, non-partisan, or otherwise apolitical - a hallmark of military professionalism. This happens because as "conflict extension" [@layman_party_2002] occurs, an increasing number of issues are brought into partisan battlegrounds, and thus, there is less politically-neutral territory. When polarization levels in America were not as high as they are today, it was more possible for individuals to either stay out of politics, avoid partisan squabbles, and otherwise to stay politically neutral. Today's polarizing environment seems to have the opposite impact: polarized politics seem to thrust issues and the unfolding political fight at the individual citizen, even when the individual is not deliberately seeking confrontation. There is no theoretical reason to suspect that this mechanism would not impact military leaders. Military leaders are citizens of the Republic, and though the institutional ethos of the military profession is to remain apolitical, such an ideal is difficult to uphold if less and less neutral political space exists.

A second challenge to military professionalism stems from the "negative partisanship"[@abramowitz_great_2018] that dominates the polarized American political landscape. Strong feelings towards those on a host of other issues are becoming increasingly hostile. These feelings are growing increasingly strong likely because of the other prominent features of political polarization: a decrease in the availability of neutral political space, and the importance of issues that are at stake in many of today's political debates.

A third, and I contend a more insidious, mechanism through which increasing political polarization works to impact military professionalism is by altering one's sense of loyalty and identity. This occurs because the issues that are at stake as political polarization increases are in many instances involving issues that individuals associate with core concepts of one's identity, and thus, one's sense of moral right and wrong. Because individuals care about their identities and what they believe to be right and wrong, they are willing to defend these values, even at a cost. The rise of "identity politics" in America as political polarization has worsened illustrates this point.

However, it is also precisely on this point that the military institution is perhaps most vulnerable to increasing political polarization. I contend that the rise of identity politics can and will be harmful if a military professional's identity and the loyalties associated with that identity directly conflict with his or her identity and loyalty as a professional soldier. As any soldier knows, and as civilians can reasonably understand,few traits are more sacred in the military than loyalty. Indeed, the virtue of loyalty is extolled in a host of codes and mantras that have played a central part of the American military since its inception, and loyalty is captured, either in word or in spirit, in a host of phrases and professional codes. In the middle of the West Point Cadet Prayer is the eloquently written plea, "make us to choose the harder right instead of the easier wrong" [@wheat_west_1918]. Even Huntington spent a great deal talking about the importance of loyalty as a trait to the very essence of the military profession. He recognized that in order for civilian control to be effective, the military profession had to organize itself as a "hierarchy of obedience," which required "loyalty and obedience" to stand out as "the highest military virtues" [@huntington_soldier_1957, 73].

High levels of political polarization obfuscates the ideal of loyalty, however, and thus pose a significant risk to the professional nature of the military. As core issues that involve individual perceptions of moral right and wrong are at stake in an increasing number of partisan battles, individuals, to include military officers, are more likely to become attached to or otherwise invested in these issues. This is especially true when the justification for taking or not taking action invokes the concept of loyalty. As a result, highly polarized political climates increase the likelihood that two different military officers may adopt completely different opinions or stances on an issue, each believing that the other is not only morally wrong, but failing to act out of loyalty in accordance with a specific oath or moral code.

# Accounting for Institutional Credibility

The previous section described how increasing political polarization presents significant challenges to the concept of military professionalism. But when will these challenges become public knowledge, and perhaps more relevant, how does increasing political polarization impact the credibility of the military institution?

One can imagine several possible scenarios. It is reasonable to consider that as political polarization increases, respect and trust in seemingly partisan institutions would decrease, while trust in nonpartisan and apolitical institutions would rise. But why? Would such a trend reflect a genuine increase in trust in the nonpartisan institution, or would it reflect the simple consequence that as one institution becomes less respected, others might simply gain as a result? Certainly, more work needs to be done in this area.

Nonetheless, I contend that when institutional credibility is high, a demand signal is created by the American public such that future interactions between the military institution and the American public are encouraged and welcomed. From an institutional perspective, the military might respond to the public's demand signal for increased military interaction by increasingly participating in public ceremonies, parades, and other showcases of military prowess (at sporting venues, for instance). From an individual perspective, military elites, especially retired military officers, would be psychologically primed to increasingly dialogue with the American public when military credibility is high. By extension, other outlets such as the press would also be primed to showcase military elites when military credibility runs high, because these elites represent an institution that commands a high degree of public respect.

When military credibility is low, the theory contends the opposite will occur. Since demand for interactions between the American public and the military is muted when institutional credibility for the military is low, the institution will not engage with the public as frequently nor as intimately because the strong demand signal that existed when institutional credibility was high no longer exists. At the individual level, military elites will also not engage as often with the public, nor will the press ask these elites to. In sum, institutional credibility and respect for the military institution impacts the likelihood of the military institution and military elites interacting with the American public. Such interaction can take a variety of forms. When respect for the military is high, the military is likely to interact, speak to, and dialogue with the American public, and when institutional credibility is low, the military is less likely to do so.

Ultimately, four possible worlds exist when the impacts of military credibility and political polarization are considered in tandem (see Table 2, below). The first is when political polarization and military credibility are both low. When these conditions hold, there is no anticipated drop in military professionalism by the military institution nor its individual elites. With these conditions, the military does not have an out-sized platform to interact with the American public, nor are there outstanding incentives for the military to raise significant disagreement over any domestic political issues. 
\singlespacing
\begin{table}[ht]
\begin{footnotesize}
\caption{Four Types of "Military Professionalism Worlds" Where Institutional Credibility and Political Polarization Vary} % title of Table
\centering % used for centering table
\begin{tabular}{|p{5cm}|p{5cm}|p{5cm}|p{5cm}|}
    \hline
    Influence of Variables & \textbf{Low Institutional Credibility} & \textbf{High Institutional Credibility}  \\ \hline
    \textbf{Low Political Polarization} & Negligible anticipated impacts. Traditional Professionalism is Expected by the Military, Civilian Leaders, and the American Public & Possible Impacts. There is a demand for military voice, but little substantive matter stemming from US domestic politics that can cause crucial disagreement. Little incentive to deviate from traditional concepts of loyalty and duty. \\ \hline
    \textbf{High Political Polarization} & Possible Impacts. Polarization raises the stakes of domestic politics, but military professionals are constrained from voicing their opinions and/or interfacing with the public due to low institutional credibility with the American public &  Highly Likely Impacts to Professionalism. Likely impacts. Military professionals have a "platform with the public" due to high institutional credibility, and high polarization raises the likelihood of disagreement between individual military elites and civilian leaders on a number of crucial topics.\\ \hline
\end{tabular}
\label{table:comparison} % is used to refer this table in the text
\end{footnotesize}
\end{table}
\doublespacing
The second possible set of conditions is when political polarization is low and institutional credibility is high. When these set of conditions exist, the military has a platform to interact with the public, as the public highly respects the military institution. The military enjoys this high level of trust and does not want to lose it. Further, because the degree of political polarization is low, there is little incentive for the military as an institution or military elites to use the trust they enjoy with the American public to adopt or denounce an issue associated with any partisan entity. The high institutional credibility increases the chance of the military interacting with the public in an unprofessional way, but the low political polarization removes the incentives of doing so.

The third type occurs when political polarization is high, but institutional credibility of the military is low. Under these conditions, the military does not command the respect of the American public like it does when credibility of the institution is high, and thus, does not possess a considerable outlet to interact with the American public. However, the incentives of loyalty and duty still loom large in the minds of some military elites as a result of the high degree of political polarization, and thus, a dip in professionalism is still possible.

The final set of conditions exists when political polarization and institutional credibility are high. It is in this world that I contend the American military finds itself in now in 2020. Under these conditions, the military faces dual dangers. High institutional credibility risks the military institution unilaterally acting or another body co-opting the military institution to speak out an a number of issue areas, and simultaneously, the highly polarized environment risks the military advocating for or denouncing partisan sentiments in the name of upholding the very traits it is most adhered to - loyalty and duty. Indeed, this set of conditions comprise a very dangerous world for the military institution and the military professional.

# Plan of the Dissertation and the Data Generating Process

To complete my dissertation, I plan on incorporating a mixed-methods approach that relies on first, proposing a theory for how political polarization impacts and affects the overarching structure of loyalty that military officers - and the military as an institution - face. I then seek to test that theory using both empirical and quantitative data, as well as a series of qualitative case studies. I plan to conclude by discussing the implications of my findings. My dissertatino will be a book style projecte composed of at least seven chapters.

In Chapter 1, I plan to frame my research question, lay out my research agenda, and review the relevant literature for my research project. In this introductory chapter, I will motivate the project by pointing out recent cases in which the American military institution has been at the center of partisan political struggles. My point in making this claim is not to argue that this place of political contention is new; indeed, it is not. There have been many times and instances in which the military has been at odds with other American political institutions and actors. My motivation, rather, is to argue that the behavior of the military institution itself - to include active and retired military officers who serve as elites - are demonstrating behavioral patterns that are at odds with traditional norms of military professionalism. This chapter will include an extensive review of the concept of military professionalism, a concept first coined by Samuel Huntington, and an extensive review of the concept of military professionalism, first coined by Samuel Huntington, and a series of critiques that Huntington's concept has endured over the years [@janowitz_professional_1960; @finer_man_1962; @feaver_civil-military_1996; @brooks_paradoxes_2020].

In Chapter 2, I plan to empirically show the deviation in military professionalism that Chapter 1 describes. Using quantitative analysis of an original data set that I collected this past Summer of more than 340 opinion-editorial pieces written by retired military officers since 1979 and published in top American newspapers, in print and online, I will demonstrate that the topics upon which retired military officers are writing, as well as the tone and tenor of their remarks, have changed in a statistically meaningful way.

In Chapter 3, I want to develop a theory for why these changes have occurred. My theory will argue that increases in the levels of political polarization and institutional credibility/prestige have driven these changes. At the heart of my theory is the idea that certain features of political polarization have so altered the political preferences of some military elites such that significant changes have occurred to fundamental concepts and structures of loyalty that exist within and outside of the military as an institution. My theory up until this point does not explain, however, why these divergences in political preferences have risen to the level of public knowledge, which I attempt to show in Chapter 2. Therefore, the other key feature that I need to incorporate into my theory is that the public credibility or prestige of these military actors has also risen, enabling them to make arguments to the public without fear of backlash. This portion of the theory is an argument about the prestige of military elites as well as of the military institution, and how that prestige translates into the ability to message to the public.

### 2020-09-23 code/prospectusv1.Rmd
commit

@@
 
-# The Scholarly Literature and Why This Question Matters
+# Framing Within the Scholarly Literature
 
-The question of how political polarization affects changes in the concept of military professionalism spans several related literatures, most notable of which are American Politics and International Relations.
+## Political Polarization
+
+The question of if and how political polarization drives changes in the professional identity of the military spans several related literatures, most notable of which are American Politics and International Relations.
 Broadly speaking, this project is focused on how one aspect of domestic politics, political polarization, impacts a specific but important aspect of international relations, civil-military relations, which itself spans a wide and diverse literature.
-Because I intend to focus my work on the military institution, my dissertation will also rely on literature from the fields of sociology, mainly regarding the development of organizations and professions, and ethics, to include military ethics.
+Because I intend to focus my work on the military as an institution, my dissertation will also rely on literature from the fields of sociology, organizational psychology, and ethics.
+I will frame my research question within each of these fields briefly before addressing why this question matters and should be studied in depth.
 
 Within the American Politics literature, there is broad recognition that political polarization levels in contemporary America are higher now than at any point since the end of the American Civil War [@hare_polarization_2014], but there are multiple explanations for what has caused this trend, ranging from a growing divide over matters pertaining to race and religion [@abramowitz_great_2018], the size and role of government [@sides_two_2015], and disagreements over specific policy issues [@mann_its_2012].
-Despite the herculean task of disentangling the roots of the causes of polarization, virtually all scholars agree that political polarization is a genuine problem and should spark concern.
+Despite the herculean task of disentangling the roots of the causes of polarization, scholars agree that political polarization is a problem of sorts.
 
-One of these major concerns is the impact of polarized domestic politics on American foreign policy in a broad sense, and specifically, on America's ability to lead and impact the world.
+One of the major concerns regarding political polarization is its on impact on American foreign policy in a broad sense, and specifically, on America's ability to lead and impact the world.
 Kenneth Schultz, for instance, underscored four challenges that have emerged as a result of America's polarized domestic politics, including the ability to obtain bipartisan support for "risky undertakings," the ability to learn the "lessons of foreign policy failure," the inability to "make long term commitments to allies and adversaries" as a result of "dramatic policy swings from one administration to another," and the "vulnerability" of America's electoral system to foreign third party adversaries [@schultz_perils_2017].
@@
 
+## Civil Military Relations
+
 Thankfully, a small but growing number of civil-military relations scholars have looked at aspects of this question in recent years.
 But the focus of these works has been, I believe, not on the impact of polarization on the military itself but on other important but peripheral components of the military's relationship with American society.
+These scholars have made important findings that speak to the state of current American civil-military relations.
 One recent dissertation completed by Michael Robinson at Stanford University in 2018, for instance, found that partisan affiliation significantly affects public credibility and trust in the military perhaps resulting in or as a consequence of media "echo chambers" which may incentivize military leaders to offer opinions partisan opinions without fear of sanction [@robinson_danger_2018].
@@
 In their own right, these scholarly developments are interesting and important, but another debate in the civil-military relations field has recently emerged, or to be more correct, reawakened.
-At the center of this debate is the degree to which the concept of military professionalism should be altered if not discarded altogether [@brooks_paradoxes_2020].
-The debate is not new because in actuality, it began as soon as Samuel Huntington published his seminal work in 1957, _The Soldier and the State_.
-While I do not lay out the full synopsis of the book nor the the various critiques here explicitly, the central question in this debate is whether the concept of military professionalism, which Huntington espoused as essential to achieving military security during the Cold War and as the US military has largely inculcated among its officer corps for the last couple of generations, separates the military too much from civilians, leading to gaps in the development and execution of strategic outcomes [@rapp_civil-military_2015; @dubik_just_2016; @brooks_paradoxes_2020].
-These criticisms largely point to the American failures in Iraq and Afghanistan to produce and implement strategy and achieve a better and lasting peace - criticisms which this author, for what it is worth, shares.
-But I contend that something is missing in this debate, namely, a thorough exploration and theoretical outworking of the impacts of political polarization on the military institution.
+At the center of this debate is the degree to which the concept of _military professionalism_ should be altered, if not discarded altogether [@brooks_paradoxes_2020], by the collective civilian-military leadership, in part due to concerns about the tense domestic political environment that exists in contemporary America.
+The central question in this debate is whether the concept of _military professionalism_ actually accomplishes its intended goals: to simultaneously achieve military security from external threats, and to maintain civilian control of the military -- an argument developed in Samuel Huntington's 1957 work, _The Soldier and the State: The Theory and Politics of Civil-Military Relations_.
 
-Why does this matter?
-I contend that the impacts of contemporary political polarization on civil-military relations and on the military institution specifically matters and should be studied more thoroughly because, as I intend to show, portions of the military institution, though not all, have started to argue and disagree with civilian leaders in ways that are substantively different than previous civil-military feuds.
+In this seminal work, Huntington asserts that two distinct forces would always be in tension with one another in a liberal democracy such as the United States.
+The first was a "functional imperative," which Huntington said drove the requirement that a military be strong enough and capable of defeating the military threats it faced, and the second was a "societal imperative," which Huntington claims constrains the military in that the military can never lose certain aspects such that the military no longer represents the liberal democracy it defends [@huntington_soldier_1957, 1-3].
+Writing in the context of the seemingly long and dangerous Cold War, Huntington's point is that America needs a strong military that can win on the battlefield, but cautions that such a military cannot fundamentally alienate the character of American liberal society.
+
+Huntington asserts that the key to resolving the tension generated by these two opposing forces - to win on the battlefield and to maintain itself as a liberal democracy - is to adopt a scheme of "objective civilian control" whereby political leaders maximize the professional autonomy of the military by establishing the ends and goals of policy, but allowing military leaders to develop and implement the military plans to accomplish these goals [@huntington_soldier_1957, 80-85].
+The creation and maintenance of largely separate operating spheres between the political leaders of a democracy and the military as an institution would ensure that the military would win on the battlefield, but also stay out of politics and thus, untainted by partisan interests and concerns.
+
+Numerous critiques surrounding the Huntingtonian ideal of _military professionalism_ have been levied over the years.
+Most of these critiques, in one way or another, charge that a strict theoretical and practical separation between military and political operating spheres is fundamentally flawed.
+Many, if not all, of these critiques cite Clausewitz's famous dictum that war is a "mere continuation of policy by other means" [@clausewitz_war_1976, 69, 605], charging that war by its very nature requires the highest integration - not separation - of military and political activities.
+In essence, these critiques argue that the strict separation of the responsibilities between the military and civilians leads to poor strategic outcomes [@rapp_civil-military_2015; @brooks_paradoxes_2020].
+Moreover, these criticisms can and do point largely to American failures in Iraq and Afghanistan to secure, at least as of yet, a more-lasting peace - failures that are ultimately, according to one scholar and former general, moral in nature, and shared between both civilians and the military [@dubik_just_2016].
+
+## The Concept and Norm of Organizational Loyalty
+
+An important value among militaries, especially professional militaries, is loyalty.
+Indeed, the US Army's first "Army Value" is "loyalty," followed by six other traits that in totality, spell out the acronym, l-d-r-s-h-i-p, or "leadership": "integrity," "duty," "respect," selfless service," "honor," integrity," and "personal courage" [need cite here].
+The US Army states in its doctrinal manual on Army Leadership and the Profession, ADP 6-22, that the Army Values "embody the practical application of the Army Ethic.
+They encompass the enduring moral principles, beliefs, and laws that guide Army professionals in accomplishing the mission as well as their conduct in all aspects of life" [need cite here for ADP 6-22, p.
+1-12].
+The manual defines loyalty as, "loyalty: bear true faith and allegiance to the Constitution of the United States, the Army, your unit and other Soldiers" [need cite here for ADP 6-22.]
+
+# Where My Puzzle Fits and Why It Matters
+
+One area of work that remains under-developed, both theoretically and empirically, is an exploration of the relationship between polarizing domestic politics and the professional identity of the military, both as an American institution, but also as a profession.
+Exploring this relationship will, I believe, help scholars in the field of civil-military relations and international relations more broadly better understand if and how the military, as a unique body charged with the defense of the nation, changes in fundamental ways as a result of a polarized political climate.
+On one hand, one would expect that it would, in the sense that the military, at least in America, is comprised of people from all walks of life.
+On the other hand, one might expect that the unique purpose and "professional" nature of the military might insulate it from some of the impacts of polarized politics.
+I am especially interested, to be more specific, in if and how political polarization impacts the concept of organizational loyalty in the military, particularly because one of the hallmarks of the military's professional nature is that it ought to be apolitical and non-partisan.
+Yet in an environment where the ability to be apolitical and non-partisan is becoming more and more difficult, due to rising political polarization, it is important that scholars and policymakers understand the mechanisms through which the military is being shaped, if there any, resulting in changes to its professional character.
+
+To be clear, scholars have long studied how domestic politics impact the military.
 To be sure, military and civilian relationships are in tension by their very nature in that the leaders of a democracy allow a most undemocratic institution, the military, to protect the polity [@feaver_civil-military_1996].
@@
 Many of these rise to the level of common knowledge: Lincoln fired McClellan because he lost confidence in him as a wartime Commander; Truman sacked MacArthur because the latter was insubordinate regarding the future of war in Asia; and Obama and McChrystal clashed about troop levels in Afghanistan.
-
 Beyond these episodes in which President and wartime commander locked horns, consider several other instances of "civilians" and "military" clashing.
@@
 A more recent example, and one much studied and commented upon by civil-military relations scholars, is the so-called "Revolt of the Generals" that occurred following the early years of the invasion of Iraq, which involved several recently-retired generals and admirals sharply criticizing Secretary of Defense Rumsfeld's leadership dynamics [@cook_revolt_2008].
-The point here is that civil-military conflict is anything but novel, but that when it occurs, it is usually, for understandable reasons, involving a clear national security issue, such as a war, combat operations, or items of a related sort, such as military budgets, weapons programs, etc.
 
-While these types of disagreements are important and will certainly feature in my intended dissertation (I hope!), my motivation in choosing my intended topic is because we are now seeing portions of the military institution (particular retired military officers) openly disagree with civilian leaders on topics that I believe are, even by the widest of academic definitions, way outside of appropriate military expertise.
-In other words, there is a world of difference between retired generals criticizing a President for conduct of a war, but entirely another when generals criticize (or applaud) the President for his domestic policies that have little if anyting to do with national security.
-I contend that the latter is taking place, and that the reasons for such developments, as well as their implications, deserve to be further explored.
+My personal motivation in choosing the intended topic, however, stems from what I believe are changes in the type of civil-military disagreement that have surfaced in recent years.
+For example, I believe we are seeing portions of the military institution (particular retired military officers) openly disagree with civilian leaders on topics that I believe are, even by the widest of academic definitions, way outside of appropriate military expertise.^[I hope to prove this, empirically, in or as part of a chapter in my dissertation] More interestingly, these conflicts are occurring not in the context of major war, but the opposite: a period in which the levels of US Troops deployed in active combat zones is less than it has been in recent years.
 I further believe that while such developments are catching the attention of civil-military relations scholars, most of the work done on this topic is impressionistic [@golby_jim_keep_2020] as opposed to systematic.
-In other words, civil-military relations scholars have identified that polarized partisan politics serve as a risk for the military, but few have explored in a systematic way what these risks are, how serious they are, the mechanisms through which they work, and what the implications are for the future of the military.
-I hope to fill this gap.
-In doing so, I hope to also contribute to the reawakened debate regarding military professionalism and its role vis-a-vis the American military institution.
+In other words, civil-military relations scholars have identified that polarized partisan politics serve as a risk for the military, but none to my knowledge have have specifically set out to systematically explore if and how polarized partisan politics are impacting the military's professional identity and character.
 
+My intended dissertation will help answer several related questions, including: if personal and partisan loyalties are changing as a result of political polarization [need cite], is this having an impact within the military, i.e.
+is the concept of organizational loyalty changing as a result of polarization?
+Are these changes in organizational loyalty having an impact on the military's professional identity and character?
+Through what mechanisms do these changes operate?
+What sorts of changes are they driving, if any, in the military institution?
+in the military changing as a result of polarized domestic politics?
+If so, through what mechanisms, and how much?
+Is this problematic for the military profession, and why?
+The answers to these questions, I believe, will help inform the broader field of international relations on the broader and much more holistic issue of how changes in political polarization impact the military's ability to defend the nation and achieve the goals of American foreign policy.
+
+# The Analytical Framework
 ## Polarization and Military Professionalism

### 2020-09-26 code/prospectusv1.Rmd
commit

@@
 The fundamental question that I seek to answer is, "how is rising political polarization shaping the military as a profession in the United States, and what are the implications of these changes, both in terms of national security and the future of the military profession?" In answering this question, I develop a theory that, at its core, argues that the nature of rising political polarization fundamentally challenges the conception of loyalty among members of the military profession.
-I further argue that the strain of loyalty is especially problematic for military, as it is a profession that by its nature holds mutiple loyalties simultaneously: to the Constitution, to the President as Commander-in-Chief, to Congress, and to the American People.
+I further argue that the strain of loyalty is especially problematic for military, as it is a profession that by its nature holds multiple loyalties simultaneously: to the Constitution, to the President as Commander-in-Chief, to Congress, and to the American People.
 Balancing these competing loyalties becomes more difficult, I argue, as political polarization worsens, because of the fundamental values - not mere policy differences - that are at the root of many polarized political battles today.
@@
 These scholars have made important findings that speak to the state of current American civil-military relations.
-One recent dissertation completed by Michael Robinson at Stanford University in 2018, for instance, found that partisan affiliation significantly affects public credibility and trust in the military perhaps resulting in or as a consequence of media "echo chambers" which may incentivize military leaders to offer opinions partisan opinions without fear of sanction [@robinson_danger_2018].
+One recent dissertation completed by Michael Robinson at Stanford University in 2018, for instance, found that partisan affiliation significantly affects public credibility and trust in the military, perhaps resulting in or as a consequence of media "echo chambers" which may incentivize military leaders to offer opinions partisan opinions without fear of sanction [@robinson_danger_2018].
 Robinson's work builds off of previous work by Heidi Urben in which she found that active military officers believe less strongly that retired military officers should remain apolitical or non-partisan in retirement [@urben_party_2013], a norm that has existed for decades or more.
@@
 In their own right, these scholarly developments are interesting and important, but another debate in the civil-military relations field has recently emerged, or to be more correct, reawakened.
-At the center of this debate is the degree to which the concept of _military professionalism_ should be altered, if not discarded altogether [@brooks_paradoxes_2020], by the collective civilian-military leadership, in part due to concerns about the tense domestic political environment that exists in contemporary America.
+At the center of this debate is the degree to which the concept of _military professionalism_ should be altered, if not discarded altogether [@brooks_paradoxes_2020].
 The central question in this debate is whether the concept of _military professionalism_ actually accomplishes its intended goals: to simultaneously achieve military security from external threats, and to maintain civilian control of the military -- an argument developed in Samuel Huntington's 1957 work, _The Soldier and the State: The Theory and Politics of Civil-Military Relations_.
 
-In this seminal work, Huntington asserts that two distinct forces would always be in tension with one another in a liberal democracy such as the United States.
-The first was a "functional imperative," which Huntington said drove the requirement that a military be strong enough and capable of defeating the military threats it faced, and the second was a "societal imperative," which Huntington claims constrains the military in that the military can never lose certain aspects such that the military no longer represents the liberal democracy it defends [@huntington_soldier_1957, 1-3].
+In this seminal work, which often serves as a point of departure for many discussions and debates within the field of civil-military relations, Huntington asserts that two distinct forces would always be in tension with one another in a liberal democracy such as the United States.
+The first was a "functional imperative," which Huntington said drove the requirement that a military be strong enough and capable of defeating the military threats it faced, and the second was a "societal imperative," which Huntington claims actually constrains the military in that the military can never become so alienated from society such that the military no longer represents the liberal democracy it defends [@huntington_soldier_1957, 1-3].
+The societal imperative, Huntington claimed, was a further function of two distinct features of the United States.
+The first was America's liberal ideology [need to confirm this name of ideology], which Huntington saw as fundamentally suspicious of standing military forces, dating back to the Revolutionary War period [need cite].
+Moreover, American liberal ideology for Huntington also values the individual extensively.
+The second feature of the societal imperative, claimed Huntington, was the US constitution, which clearly states that the military will always be controlled by civilians.
 Writing in the context of the seemingly long and dangerous Cold War, Huntington's point is that America needs a strong military that can win on the battlefield, but cautions that such a military cannot fundamentally alienate the character of American liberal society.
 
-Huntington asserts that the key to resolving the tension generated by these two opposing forces - to win on the battlefield and to maintain itself as a liberal democracy - is to adopt a scheme of "objective civilian control" whereby political leaders maximize the professional autonomy of the military by establishing the ends and goals of policy, but allowing military leaders to develop and implement the military plans to accomplish these goals [@huntington_soldier_1957, 80-85].
-The creation and maintenance of largely separate operating spheres between the political leaders of a democracy and the military as an institution would ensure that the military would win on the battlefield, but also stay out of politics and thus, untainted by partisan interests and concerns.
+Figure one, below, shows a visual diagram of Huntington's logic in a political science framework (though as @feaver_civil-military_1996 points out, Huntington did not use such language when articulating his concepts).
+According to Huntington, the ultimate goal, or dependent variable, of civil-military relations in the American context is achieving both military security and maintaining civilian supremacy over the military.
+Huntington viewed the key independent variable that impacts and enables these goals while also resolving the tension generated by the opposing functional and societal imperatives, as described above, as adoption of a model of "objective civilian control" of civil-military relations, whereby political leaders maximize the professional autonomy of the military by establishing the ends and goals of policy, but allow military leaders to develop and implement the military plans to accomplish these goals [@huntington_soldier_1957, 80-85].
+
+In making his argument, Huntington went to great lengths to describe an alternate world in which the pattern of civilian control, the key independent variable impacting military security and civilian supremacy of the military, was not "objective control," but "subjective control" [need cite for Huntington].
+In this conception, Huntington saw civilian leaders making the military a political object through a variety of means, including the selection of military leaders sympathetic to partisan political causes [need cite], to the point that there was no longer any distinction between the military and civilian leadership [need cite].
+To Huntington, adopting this form of control, and thus erasing what he saw as a necessary distinction between both the military and civilian leadership and society, would endanger American security.
+By virtue of its distinct mission and functional goals, Huntington asserted that the military needed to operate apart from partisan politics - something it could not do if a pattern of subjective control were to be adopted and implemented - in order to maintain military security.
+
+```{r figure one, include=FALSE, echo=FALSE}
+huntington_dag <- dagify(
+    MSCS ~ CC,
+    CC ~ FI + SI,
+    SI~I + S,
+    coords = tribble(
+      ~ name,      ~ x,    ~ y,
+      "MSCS", 1, .5,
+      "CC", .7, .5,
+      "FI", .5, .6,
+      "SI", .5, .4,
+      "I", .3, .3,
+      "S",.7, .3,
+    ))
+
+huntington_dag %>% 
+  ggplot(aes(x=x, y=y, xend=xend, yend=yend))+
+  geom_dag_edges()+
+  geom_dag_point()+
+  geom_dag_text_repel(aes(label=name), show.legend=FALSE)+
+  theme_dag()
+```
+
+```{r include=TRUE, echo=FALSE, fig.height=5, fig.width=7}
+
+huntington_dag %>% tidy_dagitty() %>% 
+  dag_label(labels=c("MSCS"="Military Security and\n Civilian Supremacy", 
+                     "CC"="Patterns of\n Civilian Control\n 'Objective Control'", 
+                     "FI"="Functional\n Imperative", 
+                     "SI"="Societal\n Imperative",
+                     "I"="Liberal Ideology",
+                     "S"="Constitutional Structure")) %>% 
+  ggplot(aes(x=x, y=y, xend=xend, yend=yend))+
+  geom_dag_edges()+
+  geom_dag_point()+
+  geom_dag_text()+ 
+  geom_dag_label_repel(aes(label=label, fill=label), col="white", show.legend=FALSE)+ theme_dag() + labs(title="Figure One: Samuel Huntington's Theory\n Of Civilan Control of the Military", subtitle="As espoused in The Soldier and The State, 1957")
+
+```
 
@@
 
-## The Concept and Norm of Organizational Loyalty
+As an extension of the critiques of Huntington, there has also been renewed focus and attention among civil-military relations scholars on numerous professional normative standards of the military.
+One of the most important yet difficult debates is an attempt by scholars to define where appropriate political behavior ends and partisan activity begins [@owens_military_2015; @golby_beyond_2015].
+In the Huntingtonian framework, this is essentially the opposite side of the debate over the same coin in attempting to define for civilians what constitutes meddling in military affairs.
+One of the reasons this debate has become even more difficult, I believe, is because as political polarization increases, the lines between partisanship and politics, as well as politics and everything not politics, have blurred, as so much of the literature on American politics and polarization affirms.
 
-An important value among militaries, especially professional militaries, is loyalty.
-Indeed, the US Army's first "Army Value" is "loyalty," followed by six other traits that in totality, spell out the acronym, l-d-r-s-h-i-p, or "leadership": "integrity," "duty," "respect," selfless service," "honor," integrity," and "personal courage" [need cite here].
+## Military Professions and Organizational Loyalty
+
+Loyalty features prominently in the international relations and civil-military relations literatures, often in a comparative sense and in the context regime exclusion and a host of related phenomena, including military coups [need cite], civil wars [@mclauchlin_loyalty_2018], and the difficulties that multi-ethnic states face in maintaining a professional military [@peled_question_1998].
+This literature deals with a host of important issues that stem from problems associated with ethnic identification and loyalty, either to a regime, cause, or leader.
+Though the US context is very different from the contexts usually explored in this literature, its insights may help me identify related dynamics and/or mechanisms that perhaps are unfolding in the US: how are changes in partisan or ideological identification (instead of ethnic) as a result of political polarization impacting conceptions of loyalty within the military.
+Because appropriate to explore partisan or ideological (instead of ethnic) identification stemming from However, this literature often assumes a definition of loyalty and rarely develops how the reader should think of loyalty in the first place.
+For this fundamental starting point of defining loyalty, the fields of ethics, management, and sociology are helpful.
+
+Loyalty has been conceptualized as both a virtue [@coleman_problems_2009] and as a moral emotion [@connor_military_2019].
+Though the differences between the two approaches are important, both assert that loyalty is something felt by an individual and subsequently reflected in action, such as, in the military context, "the desire to protect each other from harm and to defend each other from attack" [@coleman_problems_2009, p.
+109].
+Oldenquist, a philosopher, describes loyalty as a devotion or attachment that an individual has to someone or something specific such that one refers to the object of loyalty as "mine" and is inclined to feel "pride when it [the object of one's loyalty] prospers, shame when it declines, and anger or indignation when it is harmed" [@oldenquist_loyalties_1982, p.
+175, brackets mine].
+
+Loyalty further generates a willingness in the individual to assume "obligations" [@oldenquist_loyalties_1982, p.
+175] to the object of one's loyalty, and these obligations can take multiple forms, such as a willingness of the individual to endure hardships and make sacrifices [@fielder_organizational_1992].
+Scholars agree that individuals can hold multiple loyalties simultaneously, and recognize that this often can and does cause problems [@coleman_problems_2009].
+For organizations, and by extension, professions, the concept of loyalty is also very important.
+Conner et al.
+state, for example, that "having loyal members is vital to perpetuating the group" [@connor_military_2019].
+
+We also know that the military services in the United States place a premium on loyalty as a trait and concept.
+Indeed, of the different military services, the US Army in particular has extensively developed, in a normative sense, a framework entitled "The Army Ethic," in which loyalty constitutes a significant part of the Army profession [@us_department_of_the_army_adp_2019].
+This ethic, which will be described in detail in later parts of this dissertation, is formally laid out and described in the Army's doctrinal publication entitled, _Army Leadership and the Profession, ADP 6-22_, and holds that as "stewards of the Army Profession" and as "committed professionals," Army leaders "embrace and uphold the Army Values and standards of the profession" [@us_department_of_the_army_adp_2019].
+There are a total of seven "Army Values," which together spell the acronym, l-d-r-s-h-i-p, or "leadership," and the first of these values, loyalty, is followed by six others: duty, respect, selfless service, honor, integrity, and personal courage [@us_department_of_the_army_adp_2019, p.
+1-9.].
 The US Army states in its doctrinal manual on Army Leadership and the Profession, ADP 6-22, that the Army Values "embody the practical application of the Army Ethic.
@@
 
+Loyalty is further important in a military setting in a broader and strategic sense, especially when one considers that the US military, both by statute and by normative standards, serves as an apolitical and non-partisan institutions with multiple, simultaneous loyalties.
+The Constitution of the United States, to which military officers and soldiers alike ascribe allegiance when they take their respective oaths of commissioning and enlistment, deliberately contrives the military as an institution possessing multiple loyalties.
+One of the most basic forms of these multiple loyalties is captured in the fact that the Constitution empowers the President as "the Commander in Chief of the Army and the Navy of the United States" [@noauthor_us_1787], but also gives Congress an array of tremendous powers and responsibility pertaining to the Armed Forces @[noauthor_us_1787-1].
+Indeed, the tensions that inherently exist in American's constitutional framework are explored in a robust literature that examines the multiple ways in which the Congress, the President, and the military interact to implement, but also to influence, American foreign policy [for instance, see @howell_while_2007].
+
+In the context of the US military as an institution, loyalty has rarely been used as a conceptual construct in the political science literature, likely and thankfully because the US military has not yet considered, or been involved, in a coup.
+Nonetheless, I contend that there is a rich opportunity to study strategic loyalty as a concept, especially in a contemporary context in the United States, where political polarization has increased extensively.
+
 # Where My Puzzle Fits and Why It Matters
 
-One area of work that remains under-developed, both theoretically and empirically, is an exploration of the relationship between polarizing domestic politics and the professional identity of the military, both as an American institution, but also as a profession.
+One area of work that remains under-developed, both theoretically and empirically, is an exploration of the relationship between polarizing domestic politics and the professional identity of the military, both as an American institution and as a profession.
 Exploring this relationship will, I believe, help scholars in the field of civil-military relations and international relations more broadly better understand if and how the military, as a unique body charged with the defense of the nation, changes in fundamental ways as a result of a polarized political climate.
-On one hand, one would expect that it would, in the sense that the military, at least in America, is comprised of people from all walks of life.
-On the other hand, one might expect that the unique purpose and "professional" nature of the military might insulate it from some of the impacts of polarized politics.
-I am especially interested, to be more specific, in if and how political polarization impacts the concept of organizational loyalty in the military, particularly because one of the hallmarks of the military's professional nature is that it ought to be apolitical and non-partisan.
-Yet in an environment where the ability to be apolitical and non-partisan is becoming more and more difficult, due to rising political polarization, it is important that scholars and policymakers understand the mechanisms through which the military is being shaped, if there any, resulting in changes to its professional character.
+On one hand, one would expect that it might, in the sense that the American military is, after all, a microcosm of American society.
+Therefore, one could expect that if society truly is becoming polarized, then at least to some degree, so is the military, or at least those who volunteer for service and at the point in time that they join.
+On the other hand, one might reasonably expect that the unique purpose of the military, combined with its long-standing professional ethic, might insulate the military institution from some of the potentially negative impacts of polarized politics.
+These differing expectations need to be more clearly articulated and explored.
+
+I am specifically interested in the question of if and how political polarization impacts the concept of organizational loyalty in the military.
+This is an interesting research question because, as noted above, one of the most fundamental normative hallmarks of the military's professional nature is that it ought to be apolitical and non-partisan, and one of the features of militaries in general is their tendency to foster a concept of loyalty among those who serve.
+Yet a puzzle emerges: within an institution that is normatively apolitical and non-partisan, can it remain so in an environment marked by increasingly polarized domestic politics, where partisan loyalties are a salient feature of the political landscape [need cite here for partisan loyalties].
+
+My intended dissertation will help answer several related questions, including: if personal and partisan loyalties are changing as a result of political polarization [need cite], is this also happening within the military institution, and why?
+Are partisan/ideological loyalties displacing other loyalty relationships within the military?
+If so, are these changes in organizational loyalty having an impact on the military's professional identity and character?
+Through what mechanisms do these changes operate?
+What sorts of changes are they driving, if any, in the military institution?
+Is this problematic for the military profession, and why?
+
+The answers to these questions, which I intend to include as part of my dissertation, will help inform scholars and policymakers alike on number of important issues.
+First, it will help illuminate in a broad sense an important aspect of how changes in political polarization impact American institutions, and in particular, the military.
+Second, work on this topic will attempt to explore how a fundamental concept and value within the military - loyalty - is both impacted as a result of political polarization but also how loyalty changes may be driving changes in the military's professional character, identity, and nature.
+In doing so, it will help enrich a scholarly understanding of what loyalty is to begin with, and how it features prominently in the context of the American military.
 
 To be clear, scholars have long studied how domestic politics impact the military.
-To be sure, military and civilian relationships are in tension by their very nature in that the leaders of a democracy allow a most undemocratic institution, the military, to protect the polity [@feaver_civil-military_1996].
+And as scholars have pointed out, military and civilian relationships are in tension by their very nature in that the leaders of a democracy allow a most undemocratic institution, the military, to protect the polity [@feaver_civil-military_1996].
 And as American history proves, there have been countless instances in which civilian and military leaders disagreed.
@@
 For example, I believe we are seeing portions of the military institution (particular retired military officers) openly disagree with civilian leaders on topics that I believe are, even by the widest of academic definitions, way outside of appropriate military expertise.^[I hope to prove this, empirically, in or as part of a chapter in my dissertation] More interestingly, these conflicts are occurring not in the context of major war, but the opposite: a period in which the levels of US Troops deployed in active combat zones is less than it has been in recent years.
-I further believe that while such developments are catching the attention of civil-military relations scholars, most of the work done on this topic is impressionistic [@golby_jim_keep_2020] as opposed to systematic.
-In other words, civil-military relations scholars have identified that polarized partisan politics serve as a risk for the military, but none to my knowledge have have specifically set out to systematically explore if and how polarized partisan politics are impacting the military's professional identity and character.
+While I believe that such developments are catching the attention of civil-military relations scholars, most of the work done on this topic is, in my opinion, impressionistic [@golby_jim_keep_2020] as opposed to systematic.
+In other words, civil-military relations scholars have identified that polarized partisan politics serve as a risk for the military, but none to my knowledge have specifically set out to systematically explore if and how polarized partisan politics are impacting the military's professional identity and character, and how.
 
-My intended dissertation will help answer several related questions, including: if personal and partisan loyalties are changing as a result of political polarization [need cite], is this having an impact within the military, i.e.
-is the concept of organizational loyalty changing as a result of polarization?
-Are these changes in organizational loyalty having an impact on the military's professional identity and character?
-Through what mechanisms do these changes operate?
-What sorts of changes are they driving, if any, in the military institution?
-in the military changing as a result of polarized domestic politics?
-If so, through what mechanisms, and how much?
-Is this problematic for the military profession, and why?
-The answers to these questions, I believe, will help inform the broader field of international relations on the broader and much more holistic issue of how changes in political polarization impact the military's ability to defend the nation and achieve the goals of American foreign policy.
+# The Analytical Framework - Initial Arguments and Hypotheses
 
-# The Analytical Framework
-## Polarization and Military Professionalism
+Given all that the existing scholarship says about the topics of political polarization, military professionalism, and concepts of loyalty and identity, the goal in my dissertation is to examine how these concepts are inter-related, and specifically, if and how polarization drives changes in the way that these concepts are inter-related.
+Further, I want to determine if these changes impact national security, and if so, in what ways?
 
-how increasing domestic political polarization impacts conceptions of loyalty in military professions, and to examine the implications of these changes in loyalty on broader security outcomes.
-For at least the past 75 years[need footnote here], the US military has prided itself on being not merely an organization charged with the defense of the nation, but a profession.
-That is, on being an institution marked by several unique traits, including: providing a unique service that requires special skills (the application and management of violence); establishing and maintaining standards of behavior and performance, as well as policing itself when members of that profession fail to uphold these standards; and building and maintaining trust between itself, the profession, with the client whom it serves (the American people) [insert source here].
-The US military, further, is both by statute and by normative standards an apolitical and non-partisan institution with multiple, simultaneous loyalties: to the Constitution of the United States; to the President of the United States as the Commander-in-Chief; and to Congress.
-But what happens if and when the objects of the military's multiple loyalties not merely disagree, but become severely divided ideologically?
-Can such an outcome exist, and if it can, is there empirical proof that such a trend is occurring?
-If so, what are the normative implications, both for the military as a profession, and for the nation's security, and are there actions that can be taken to halt such a trend?
+Figure two below, shows an initial way to link some of the variables that I am interested in exploring throughout my dissertation.
+From left to right, this conceptual model suggests that aspects of political polarization are simultaneously driving changes in both the institutional credibility of the military (how society views the military), as well as changes in the conceptions of loyalty and identity (within the military).
+These two variables are, in turn, driving changes in the concept of military professionalism, my main dependent variable, within the military, and I suspect that these changes may, if not yet, potentially impact various military outcomes.
+By various military outcomes, I mean the very same military outcomes that other scholars do and have already explored, such as effectiveness on the battlefield, i.e winning wars and battles [@reiter_democracies_2002; @biddle_democracy_2004] and achieving strategic effectivness [@brooks_creating_2007; @brooks_shaping_2008].
+While I intend to examine this system of variables, my focus will be on the link between increasing political polarization, the concepts of loyalty and identity, and military professionalism, both in order to focus my research but also because there has been more work done recently examining other variables, to include the impact of increasing political polarization on institutional credibility [@robinson_michael_danger_2018].
 
-## The Concept and Structure of Loyalty
-
-Figure One, below, shows one possible way to conceptualize the system of problems generated by increasing political polarization and its impact on the military profession, as well as the roles of other structural variables, including the institutional credibility of the military profession, and other various military outcomes.
-From left to right, this conceptual model suggests that aspects of political polarization are simulataneously driving changes in the institutional credibility of the military as well as changes in the professional character/nature of the military.
-This is made more complicated, as the model suggests, that the intervening variable of institutional credibility is also driving some changes in the professional character of the military.
-
-In my dissertation, I situate changing conceptions of loyalty as a major subcomponent of the arrow running from political polarization to changes in military professionalism.
-At the same time, I do think that political polarization is impacting the credibility of the military institution, which in turn may likewise be driving changes in military professionalism.
-The specific mechanisms through which this is occurring, if at all, however, need to be developed, proposed, and tested.
-
-```{r figure one, include=FALSE, echo=FALSE}
+```{r figure two, include=FALSE, echo=FALSE}
 problem_dag <- dagify(
     MO ~ MP,
-    MP ~ PP + IC,
+    MP ~ LI + IC,
     IC~PP,
+    LI~PP,
     coords = tribble(
@@
       "MP", .7, .5,
-      "IC", .3, .2,
+      "IC", .3, .3,
+      "LI", .3, .7,
       "PP", .1,.5,
@@
                      "PP"="Increasing\n Political\n Polarization", 
-                     "IC"="Institutional\n Credibility")) %>% 
+                     "IC"="Institutional\n Credibility",
+                     "LI"="Concepts of\n Loyalty and Identity")) %>% 
   ggplot(aes(x=x, y=y, xend=xend, yend=yend))+
@@
 
+A primary argument is...
+
+how increasing domestic political polarization impacts conceptions of loyalty in military professions, and to examine the implications of these changes in loyalty on broader security outcomes.
+For at least the past 75 years[need footnote here], the US military has prided itself on being not merely an organization charged with the defense of the nation, but a profession.
+That is, on being an institution marked by several unique traits, including: providing a unique service that requires special skills (the application and management of violence); establishing and maintaining standards of behavior and performance, as well as policing itself when members of that profession fail to uphold these standards; and building and maintaining trust between itself, the profession, with the client whom it serves (the American people) [insert source here].
+But what happens if and when the objects of the military's multiple loyalties not merely disagree, but become severely divided ideologically?
+Can such an outcome exist, and if it can, is there empirical proof that such a trend is occurring?
+If so, what are the normative implications, both for the military as a profession, and for the nation's security, and are there actions that can be taken to halt such a trend?
+
+## The Concept and Structure of Loyalty
+
+In my dissertation, I situate changing conceptions of loyalty as a major sub component of the arrow running from political polarization to changes in military professionalism.
+At the same time, I do think that political polarization is impacting the credibility of the military institution, which in turn may likewise be driving changes in military professionalism.
+The specific mechanisms through which this is occurring, if at all, however, need to be developed, proposed, and tested.
+
 At this stage in the dissertation phase, I do not know how polarization is impacting conceptions of loyalty, or even if it is, for sure.
@@
 As the bonds of loyalty forged on the battlefield are strong and different in ways than loyalty forged in other environments, I suspect that the military and its leaders will take very seriously any accusations that the corporate entity or individual officers are failing to be loyal to a number of ideals.
-In the Army, for example, the first "Army Value," as it is called by the militar service,  which is summarized as the acronym, "LDRSHIP" (short for "leadership"), is the word "loyalty." [need cite for the Army values here].
+In the Army, for example, the first "Army Value," as it is called by the military service,  which is summarized as the acronym, "LDRSHIP" (short for "leadership"), is the word "loyalty." [need cite for the Army values here].
 
@@
 I plan to conclude by discussing the implications of my findings.
-My dissertatino will be a book style projecte composed of at least seven chapters.
+My dissertation will be a book style project composed of at least seven chapters.
 

### 2020-10-01 code/prospectusv1.Rmd
commit

@@
 ---
-title: "The Limits of Loyalty: Postmodernism, Polarization, and Military Professionalism  (Prospectus)"
+title: "Loyalty's Limits: The Impact of Postmodernism and Political Polarization on Military Professionalism  (Prospectus)"
 author: "Peter M. Erickson"
@@
 
-  _This dissertation seeks to explore the relationship of postmodernism and political polarization and their impact on the US military profession.
-The fundamental question that I seek to answer is, "how is rising political polarization shaping the military as a profession in the United States, and what are the implications of these changes, both in terms of national security and the future of the military profession?" In answering this question, I develop a theory that, at its core, argues that the nature of rising political polarization fundamentally challenges the conception of loyalty among members of the military profession.
-I further argue that the strain of loyalty is especially problematic for military, as it is a profession that by its nature holds multiple loyalties simultaneously: to the Constitution, to the President as Commander-in-Chief, to Congress, and to the American People.
-Balancing these competing loyalties becomes more difficult, I argue, as political polarization worsens, because of the fundamental values - not mere policy differences - that are at the root of many polarized political battles today.
-In the dissertation, I seek to employ a mixed-methods approach that incorporates several case studies, as well as quantitative analysis, to support my theory.
-This dissertation will contribute to the fields of American politics, international security, and civil-military relations by further illuminating the mechanisms through which political polarization is impacting the military institution.
-Specifically, it will provide a contrarian view to recent civil-military relations scholars who have argued that the traditional concept of military professionalism should be significantly redesigned, if not abandoned entirely.
-This dissertation will argue the opposite: that what is needed most in politically polarizing times, such as those the US is in right now, is a return to and embrace of traditional conceptions of military professionalism - an effort that will require deliberate action by not just military leaders, but elected civilian leaders as well._
+  _This dissertation seeks to explore the impact of increasing political polarization on the professional identity of military officers and the military institution.
+I develop a theory that argues that the nature of rising political polarization fundamentally challenges the conception of loyalty among members of the military profession, and that such a strain is problematic for an institution such as the military that holds the virtue of loyalty in high regard, and by statutes possesses multiple, simultaneous loyalties: to the Constitution of the United States, to the President qua Commander-in-Chief, to Congress, to the American People, to one's fellow service members, and to one's self.
+Balancing these competing loyalties becomes more difficult as political polarization worsens, I argue, because of the fundamental and core values - not mere policy differences - that are at the root of many contemporary polarized political battles.
+Employing a mixed-methods approach that incorporates quantitative analysis and several historical case studies, this dissertation argues that the challenges to conceptions of loyalty among members of the military institution brought about by political polarization have the potential to impact normative concepts of military professionalism, which could be problematic not only for the character and identity of the military as an institution, but for American national security at large._
 geometry: margin = 1.15in              # margin widths
@@
 Because I intend to focus my work on the military as an institution, my dissertation will also rely on literature from the fields of sociology, organizational psychology, and ethics.
-I will frame my research question within each of these fields briefly before addressing why this question matters and should be studied in depth.
+I will frame my research question within each of these fields briefly, and to define important concepts as required, before addressing why this question matters and should be studied in depth.
 
 Within the American Politics literature, there is broad recognition that political polarization levels in contemporary America are higher now than at any point since the end of the American Civil War [@hare_polarization_2014], but there are multiple explanations for what has caused this trend, ranging from a growing divide over matters pertaining to race and religion [@abramowitz_great_2018], the size and role of government [@sides_two_2015], and disagreements over specific policy issues [@mann_its_2012].
-Despite the herculean task of disentangling the roots of the causes of polarization, scholars agree that political polarization is a problem of sorts.
+Despite the wide-ranging scholarship exploring the many facets of political polarization, virtually all scholars agree that contemporary levels of political polarization in the United States are historically abnormal and pose myriad problems.
 
@@
 Kenneth Schultz, for instance, underscored four challenges that have emerged as a result of America's polarized domestic politics, including the ability to obtain bipartisan support for "risky undertakings," the ability to learn the "lessons of foreign policy failure," the inability to "make long term commitments to allies and adversaries" as a result of "dramatic policy swings from one administration to another," and the "vulnerability" of America's electoral system to foreign third party adversaries [@schultz_perils_2017].
-Because the military is a central actor in shaping America's foreign policy, obtaining a better understanding of how political polarization shapes various military institutions specifically will yield findings that may bear on helping policymakers and academics help solve each of the problems Schultz identifies.
+Because the military is a central actor in shaping America's foreign policy, obtaining a better understanding of how political polarization shapes various military institutions will yield findings that may help policymakers and academics help solve the problems Schultz identifies.
 
@@
 
-Thankfully, a small but growing number of civil-military relations scholars have looked at aspects of this question in recent years.
-But the focus of these works has been, I believe, not on the impact of polarization on the military itself but on other important but peripheral components of the military's relationship with American society.
-These scholars have made important findings that speak to the state of current American civil-military relations.
-One recent dissertation completed by Michael Robinson at Stanford University in 2018, for instance, found that partisan affiliation significantly affects public credibility and trust in the military, perhaps resulting in or as a consequence of media "echo chambers" which may incentivize military leaders to offer opinions partisan opinions without fear of sanction [@robinson_danger_2018].
-Robinson's work builds off of previous work by Heidi Urben in which she found that active military officers believe less strongly that retired military officers should remain apolitical or non-partisan in retirement [@urben_party_2013], a norm that has existed for decades or more.
-Together, Robinson and Urben join other scholars in providing a wealth of information regarding public opinion and the military, and in particular, how aspects of political polarization shape partisan differences regarding topics such as the use of force [@golby_jim_elite_2017] and military elite endorsement of political candidates [@golby_military_2012].
+Civil-military relations comprises an immense area of scholarship and interest, and deals with a correspondingly large number of questions and debates that have occupied the minds of political scientists, theorists, historians, ethicists, and sociologists for hundreds if not thousands of years.
+At its root, the study of civil-military relations examines what Peter Feaver calls the "problematique," or how to ensure that the military is strong enough to defeat threats to the polity without threatening the polity itself [@feaver_civil-military_1996; @feaver_civil-military_1999].
+Feaver describes his "problematique" as "vexing" and a great "paradox" in that the military must be "strong enough to prevail in war," yet submissive to authority such that it cannot and will not "destroy or prey on the society it is intended to protect" [@feaver_civil-military_1999, p.
+214].
 
-In their own right, these scholarly developments are interesting and important, but another debate in the civil-military relations field has recently emerged, or to be more correct, reawakened.
-At the center of this debate is the degree to which the concept of _military professionalism_ should be altered, if not discarded altogether [@brooks_paradoxes_2020].
-The central question in this debate is whether the concept of _military professionalism_ actually accomplishes its intended goals: to simultaneously achieve military security from external threats, and to maintain civilian control of the military -- an argument developed in Samuel Huntington's 1957 work, _The Soldier and the State: The Theory and Politics of Civil-Military Relations_.
+Civil-military relations involves a wide angle of significant and important theoretical questions that involve the pull and push of power between institutions, the delineation of authority, and a host of other massive and consequential issues impacting how militaries perform in war, their relationship to the societies they serve, and other important questions.
+In the context of the US, which has never experienced or come close to a coup (thankfully), the study of civil-military relations often focuses on the relationship between military and government elites, or what Risa Brooks calls "leadership at the state's apex" [@brooks_risa_integrating_2019, p.
+380].
+Indeed, it is at this level that vital decisions are made, including to go to war, how to prosecute war, how to achieve and implement effective strategy, how to prepare for the next conflict, and how large the military and its budget should be.
+Each of these questions, and many, many more, involve questions regarding the proper role of the military and its leaders relative to both society and, perhaps more important, the leaders of government.
+A number of scholars have examined this question on normative grounds, but none has been more lasting and consequential than the late Samuel Huntington, whose 1957 seminal work, _The Soldier and the State: The Theory and Politics of Civil-Military Relations_, often serves as the point of departure for many scholarly works on civil-military relations.
+Although there have been many critiques of Huntington over the years [for example, see @janowitz_professional_1960; @finer_man_1962; @cohen_supreme_2003; @feaver_armed_2003; @brooks_paradoxes_2020], it is absolutely essential to understand Huntington's concept of normative civil-military relations, and what lies at the heart of it - a theory of civilian control of the military that maximizes the concept of what Huntington called "military professionalism" [@huntington_soldier_1957].
 
-In this seminal work, which often serves as a point of departure for many discussions and debates within the field of civil-military relations, Huntington asserts that two distinct forces would always be in tension with one another in a liberal democracy such as the United States.
+In _The Soldier and the State_, Huntington asserts that two distinct forces would always be in tension with one another in a liberal democracy such as the United States.
 The first was a "functional imperative," which Huntington said drove the requirement that a military be strong enough and capable of defeating the military threats it faced, and the second was a "societal imperative," which Huntington claims actually constrains the military in that the military can never become so alienated from society such that the military no longer represents the liberal democracy it defends [@huntington_soldier_1957, 1-3].
 The societal imperative, Huntington claimed, was a further function of two distinct features of the United States.
-The first was America's liberal ideology [need to confirm this name of ideology], which Huntington saw as fundamentally suspicious of standing military forces, dating back to the Revolutionary War period [need cite].
-Moreover, American liberal ideology for Huntington also values the individual extensively.
-The second feature of the societal imperative, claimed Huntington, was the US constitution, which clearly states that the military will always be controlled by civilians.
-Writing in the context of the seemingly long and dangerous Cold War, Huntington's point is that America needs a strong military that can win on the battlefield, but cautions that such a military cannot fundamentally alienate the character of American liberal society.
+The first was an "ideological constant," which referred to America's political ideology of liberalism [@huntington_soldier_1957, 143-162].
+Huntington argued that liberalism as a political ideology stressed "individualism" [@huntington_soldier_1957, 90], and that American society was, since its founding, suspicious of standing military organizations and professional military forces.
+The second feature of the societal imperative, however, was the US Constitution, which Huntington claimed to be "conservative," and written by "men who feared concentrated political power and who provided for the widespread dispersion of that power among numerous governmental units" [@huntington_soldier_1957, 143].
+The tension raised by these two imperatives - one functional stemming from external threat, the other societal stemming from the character of American political ideology and structure - needed to be solved, claimed Huntington.
+Writing in the context of the seemingly long and dangerous Cold War, Huntington's point is that America needs strong military that can win on the battlefield, but cautions that such a military cannot fundamentally alienate the character of American liberal society.
+Both goals - military security and civilian control - had to be obtained.
 
-Figure one, below, shows a visual diagram of Huntington's logic in a political science framework (though as @feaver_civil-military_1996 points out, Huntington did not use such language when articulating his concepts).
+Figure 1, below, shows a visual diagram of Huntington's logic in a political science framework [@feaver_civil-military_1996].
+As Peter Feaver points out in a 1996 article, Huntington did not use such language when developing and explaining his concept.
 According to Huntington, the ultimate goal, or dependent variable, of civil-military relations in the American context is achieving both military security and maintaining civilian supremacy over the military.
-Huntington viewed the key independent variable that impacts and enables these goals while also resolving the tension generated by the opposing functional and societal imperatives, as described above, as adoption of a model of "objective civilian control" of civil-military relations, whereby political leaders maximize the professional autonomy of the military by establishing the ends and goals of policy, but allow military leaders to develop and implement the military plans to accomplish these goals [@huntington_soldier_1957, 80-85].
-
-In making his argument, Huntington went to great lengths to describe an alternate world in which the pattern of civilian control, the key independent variable impacting military security and civilian supremacy of the military, was not "objective control," but "subjective control" [need cite for Huntington].
-In this conception, Huntington saw civilian leaders making the military a political object through a variety of means, including the selection of military leaders sympathetic to partisan political causes [need cite], to the point that there was no longer any distinction between the military and civilian leadership [need cite].
-To Huntington, adopting this form of control, and thus erasing what he saw as a necessary distinction between both the military and civilian leadership and society, would endanger American security.
-By virtue of its distinct mission and functional goals, Huntington asserted that the military needed to operate apart from partisan politics - something it could not do if a pattern of subjective control were to be adopted and implemented - in order to maintain military security.
+Huntington viewed the key independent variable that impacts and enables these goals while also resolving the tension generated by the opposing functional and societal imperatives, as described above, the adoption of a model of "objective civilian control" of civil-military relations, whereby political and military leaders "maximize military professionalism," which includes granting the military a high degree of professional autonomy [@huntington_soldier_1957, 80-85].
 
@@
   geom_dag_text()+ 
-  geom_dag_label_repel(aes(label=label, fill=label), col="white", show.legend=FALSE)+ theme_dag() + labs(title="Figure One: Samuel Huntington's Theory\n Of Civilan Control of the Military", subtitle="As espoused in The Soldier and The State, 1957")
+  geom_dag_label_repel(aes(label=label, fill=label), col="white", show.legend=FALSE)+ theme_dag() + labs(title="Figure 1: Samuel Huntington's Theory\n Of Civilan Control of the Military", subtitle="As espoused in The Soldier and The State, 1957")
 
@@
 
-Numerous critiques surrounding the Huntingtonian ideal of _military professionalism_ have been levied over the years.
+Huntington's argument in support of "objective control" of the military stands in contra-distinction, as he explains, to an alternate model of "subjective control" in which civilian leaders limit the power of the military by treating it as any other civilian interest group [@huntington_soldier_1957, 80-83].
+In fact, Huntington admits that the US Constitution actually envisioned this "subjective" scheme of civilian control of the military [@huntington_soldier_1957, 163-168].
+What the framers did not see, nor could they, argues Huntington, was the rise of the military as a profession in response to ever-increasing and dangerous external threats.
+It was because of these pernicious threats, argues Huntington, which the US military had witnessed, confronted, and battled during the 19th and 20th centuries during World War One, World War Two, and now the Cold War, that the military profession had risen as a separate and distinct body that could focus on the demands of military service.
+Such a separation and distinctness, claims Huntington, would need to be embraced by civilians and the militarym going forward into the future.
+In so doing, claims Huntington, is the only way that both military security and civilian control of the military - two often-competing imperatives - will be safeguarded.
+
+Numerous critiques surrounding the Huntingtonian ideal of _military professionalism_ have been levied over the years, as Huntington did not attempt to precisely define it.
 Most of these critiques, in one way or another, charge that a strict theoretical and practical separation between military and political operating spheres is fundamentally flawed.
@@
 As an extension of the critiques of Huntington, there has also been renewed focus and attention among civil-military relations scholars on numerous professional normative standards of the military.
-One of the most important yet difficult debates is an attempt by scholars to define where appropriate political behavior ends and partisan activity begins [@owens_military_2015; @golby_beyond_2015].
+One of the most important -  yet difficult - debates is an attempt by scholars to define where appropriate political behavior ends and partisan activity begins [@owens_military_2015; @golby_beyond_2015].
 In the Huntingtonian framework, this is essentially the opposite side of the debate over the same coin in attempting to define for civilians what constitutes meddling in military affairs.
@@
 The US Army states in its doctrinal manual on Army Leadership and the Profession, ADP 6-22, that the Army Values "embody the practical application of the Army Ethic.
-They encompass the enduring moral principles, beliefs, and laws that guide Army professionals in accomplishing the mission as well as their conduct in all aspects of life" [need cite here for ADP 6-22, p.
+They encompass the enduring moral principles, beliefs, and laws that guide Army professionals in accomplishing the mission as well as their conduct in all aspects of life" [@us_department_of_the_army_adp_2019, p.
 1-12].
-The manual defines loyalty as, "loyalty: bear true faith and allegiance to the Constitution of the United States, the Army, your unit and other Soldiers" [need cite here for ADP 6-22.]
+The manual defines loyalty as, "loyalty: bear true faith and allegiance to the Constitution of the United States, the Army, your unit and other Soldiers" [@us_department_of_the_army_adp_2019].
+The virtue of loyalty is likewise extolled in a host of codes and mantras that have played a central part of the American military since its inception.
+In the middle of the West Point Cadet Prayer is the eloquently written plea, "make us to choose the harder right instead of the easier wrong" [@wheat_west_1918].
+Even Huntington spent a great deal talking about the importance of loyalty as a trait to the very essence of the military profession.
+He recognized that in order for civilian control to be effective, the military profession had to organize itself as a "hierarchy of obedience," which required "loyalty and obedience" to stand out as "the highest military virtues" [@huntington_soldier_1957, 73].
 
@@
 
-# The Analytical Framework - Initial Arguments and Hypotheses
+Thankfully, a growing number of civil-military relations scholars have looked at various aspects of this question - the impact of the political polarization on civil-military relations - in recent years.
+But the focus of these works has been, I believe, not on the impact of polarization on the military itself but on other important but peripheral components of the military's relationship with American society.
+One recent dissertation completed by Michael Robinson at Stanford University in 2018, for instance, found that partisan affiliation significantly affects public credibility and trust in the military, perhaps resulting in or as a consequence of media "echo chambers" which may incentivize military leaders to offer opinions partisan opinions without fear of sanction [@robinson_danger_2018].
+Robinson's work builds off of previous work by Heidi Urben in which she found that active military officers believe less strongly that retired military officers should remain apolitical or non-partisan in retirement [@urben_party_2013], a norm that has existed for decades or more.
+Together, Robinson and Urben join other scholars in providing a wealth of information regarding public opinion and the military, and in particular, how aspects of political polarization shape partisan differences regarding topics such as the use of force [@golby_jim_elite_2017] and military elite endorsement of political candidates [@golby_military_2012].
 
-Given all that the existing scholarship says about the topics of political polarization, military professionalism, and concepts of loyalty and identity, the goal in my dissertation is to examine how these concepts are inter-related, and specifically, if and how polarization drives changes in the way that these concepts are inter-related.
-Further, I want to determine if these changes impact national security, and if so, in what ways?
+# The Analytical Framework - Initial Argument
 
-Figure two below, shows an initial way to link some of the variables that I am interested in exploring throughout my dissertation.
-From left to right, this conceptual model suggests that aspects of political polarization are simultaneously driving changes in both the institutional credibility of the military (how society views the military), as well as changes in the conceptions of loyalty and identity (within the military).
-These two variables are, in turn, driving changes in the concept of military professionalism, my main dependent variable, within the military, and I suspect that these changes may, if not yet, potentially impact various military outcomes.
-By various military outcomes, I mean the very same military outcomes that other scholars do and have already explored, such as effectiveness on the battlefield, i.e winning wars and battles [@reiter_democracies_2002; @biddle_democracy_2004] and achieving strategic effectivness [@brooks_creating_2007; @brooks_shaping_2008].
-While I intend to examine this system of variables, my focus will be on the link between increasing political polarization, the concepts of loyalty and identity, and military professionalism, both in order to focus my research but also because there has been more work done recently examining other variables, to include the impact of increasing political polarization on institutional credibility [@robinson_michael_danger_2018].
+## How Polarization Impacts Loyalty as a Concept
+
+In this section of the prospectus, I will lay out an initial argument, as well as accompanying hypotheses, which I intend to explore and test in my dissertation.
+Figure 2, below, shows an initial conception of the linkages between several of the variables discussed in the previous section.
+From left to right, this conceptual model suggests that worsening political polarization is driving changes in both the institutional credibility of the military (how society views the military) and changes in the conceptions of loyalty and identity (within the military).
+These two independent variables are, I suggest, driving subsequently driving changes in the concept of military professionalism, the main dependent variable of analysis in my dissertation.
+Following Huntington's model, I also suggest that changes in military professionalism can impact a variety of outcomes, by which I mean the very same outcomes that other scholars do and have already explored, such as effectiveness on the battlefield [@reiter_democracies_2002; @biddle_democracy_2004] and achieving strategic effectivness [@brooks_creating_2007; @brooks_shaping_2008].
+While I intend to examine the broad system of variables shown in Figure 2, the focus of my dissertation, as well as the least-explored linkage in this model, is the relationship between political polarization and changes in the concept of loyalty and identity, and how such changes manifest themselves in the variation of military professionalism.
+For emphasis and clarity, I do not think that the other link - that between political polarization and institutional credibility - is unimportant.
+Rather, this link seems to be the focus of more-recent scholarship [@robinson_michael_danger_2018].
 
@@
   geom_dag_text()+ 
-  geom_dag_label_repel(aes(label=label, fill=label), col="white", show.legend=FALSE)+ theme_dag() + labs(title="Figure One", subtitle="The Impact of Political Polarization on Military Professionalism")
+  geom_dag_label_repel(aes(label=label, fill=label), col="white", show.legend=FALSE)+ theme_dag() + labs(title="Figure 2", subtitle="The Impact of Political Polarization on Military Professionalism")
 
@@
 
-A primary argument is...
-
-how increasing domestic political polarization impacts conceptions of loyalty in military professions, and to examine the implications of these changes in loyalty on broader security outcomes.
-For at least the past 75 years[need footnote here], the US military has prided itself on being not merely an organization charged with the defense of the nation, but a profession.
-That is, on being an institution marked by several unique traits, including: providing a unique service that requires special skills (the application and management of violence); establishing and maintaining standards of behavior and performance, as well as policing itself when members of that profession fail to uphold these standards; and building and maintaining trust between itself, the profession, with the client whom it serves (the American people) [insert source here].
-But what happens if and when the objects of the military's multiple loyalties not merely disagree, but become severely divided ideologically?
-Can such an outcome exist, and if it can, is there empirical proof that such a trend is occurring?
-If so, what are the normative implications, both for the military as a profession, and for the nation's security, and are there actions that can be taken to halt such a trend?
-
-## The Concept and Structure of Loyalty
-
-In my dissertation, I situate changing conceptions of loyalty as a major sub component of the arrow running from political polarization to changes in military professionalism.
-At the same time, I do think that political polarization is impacting the credibility of the military institution, which in turn may likewise be driving changes in military professionalism.
-The specific mechanisms through which this is occurring, if at all, however, need to be developed, proposed, and tested.
-
-At this stage in the dissertation phase, I do not know how polarization is impacting conceptions of loyalty, or even if it is, for sure.
-I strongly suspect that it is, however, for a couple of reasons.
-First, the military is, I believe, susceptible to charges of failing to uphold the ideal of loyalty based on the nature of the organization in that it ultimately is charged to enact violence on behalf of the state.
-Harold Lasswell said of military officers that their skill is the "management of violence" [need quote from Huntington].
-As the bonds of loyalty forged on the battlefield are strong and different in ways than loyalty forged in other environments, I suspect that the military and its leaders will take very seriously any accusations that the corporate entity or individual officers are failing to be loyal to a number of ideals.
-In the Army, for example, the first "Army Value," as it is called by the military service,  which is summarized as the acronym, "LDRSHIP" (short for "leadership"), is the word "loyalty." [need cite for the Army values here].
-
-A second reason that I suspect polarization is impacting the conception of loyalty among the military professions stems from some of the mechanisms that we know regarding how political polarization operates.
-Politics of American scholars, though not united in agreement regarding the causes of polarization, have agreed that polarization is working through at least three mechanisms, all of which, I believe, impact the military in its normative role qua non-partisan and apolitical American institution.
-First, there is a sharp increase in the number and scope of issues that have been subsumed by partisan debate and disagreement (a phenomenon referred to among scholars of American Politics as "conflict extension" [@layman_party_2002]).
-A second mechanism exists through the increase in the strength of negative feelings held by one partisan group towards those of the other side, a phenomenon referred to as "negative partisanship" [@abramowitz_great_2018].
-Third, there is an increase in the salience of debates over issues that individual partisans consider core to their individual identities [@white_values_2002; @mann_its_2012; @sides_identity_2018].
-Each of these three features of the politically polarized landscape strain the concept of military professionalism in important ways, and taken together, comprise a significant challenge to the individual military professional and thus, to the professional nature of the military as a whole.
-These challenges are theorized in Table 1, below.
+A primary contention from the system visualized above is that as political polarization in the domestic sphere increases, the military faces increased challenges in how both individual military leaders as well as the institution conceptualizes, adopts, and exercises loyalty, and by extension, professional identity.
+This expectation stems from several features that scholars of American politics have raised regarding the current politically polarized climate within the United States.
+First, scholars have noted that political polarization raises the number and scope of issues that are subsumed by partisan debate and disagreement (a phenomenon referred to among scholars of American Politics as "conflict extension" [@layman_party_2002]).
+A second feature of contemporary political polarization is the prominence of  negative feelings held by one partisan group towards another, a phenomenon referred to as "negative partisanship" [@abramowitz_great_2018].
+A third feature is an increase in the salience of debates over issues that individual partisans consider core to their individual identities [@white_values_2002; @mann_its_2012; @sides_identity_2018].
+I contend that these features impact individuals serving in the military, including military leaders, as they do members of the domestic public writ large, and that these impacts are problematic for members of an institution that normatively exists as an apolitical and non-partisan actor.
+Table 1, below, lays out how these three prominent features of domestic political polarilzation may impact the non-partisan and apolitical nature of the normative military professional and institution.
 
@@
 \begin{footnotesize}
-\caption{Mechanisms of Rising Political Polarization Which Impact Military Professionalism} % title of Table
+\caption{How Features of Contemporary Political Polarization Impact Normative Military Professionalism} % title of Table
 \centering % used for centering table
@@
     \textbf{Features of Increasing Political Polarization} & \textbf{Associated Mechanism} & \textbf{Potential Impact on Military Professionalism}  \\ \hline
-    \textbf{Conflict Extension} & Less politically-neutral space exists as partisan disagreements subsume a greater number of issues.
-It becomes increasingly difficult to remain "out of politics." & Military Professionals face an increasingly difficult prospect of remaining apolitical.
+    \textbf{Conflict Extension} & Less politically-neutral space exists as partisan disagreements subsume a greater number/scope of issues.
+& Military Professionals and the Institution find it increasingly difficult to "stay out" of partisan politics as the territory of what is considered partisan politics grows.
 \\ \hline
     \textbf{Negative Partisanship} & Strong feelings of personal dislike and disdain arise against those who hold to opposing political views.
-There is a belief that those on the "other side" are greatly wrong and flawed.
-& There is a greater willingness to confront or take issue with those who hold opposing views.\\ \hline
-    \textbf{Core and dearly-held issues are increasingly at stake} & As the issues to which one holds most dear are at stake, the salience of one's identity and sense of belonging to a particular group is raised.
-Associated with this phenomenon is a belief that not just preference, but moral right and wrong, are at stake.
-& Military professionals face potential changes in one's concept and notion of loyalty.
-There is a sense that failing to take action over a certain issue is equivalent to failing to uphold a sense of loyalty to an important group, idea, or concept.\\ \hline
+There is a belief that those on the other side of the political aisle are not just mistaken over policy, but greatly wrong and flawed as individuals.
+& Even among military leaders, there is a greater willingness to confront or take issue with those who hold opposing views.\\ \hline
+    \textbf{Individuals believe that "core" issues are increasingly at stake} & The salience of one's identity and core values is raised as partisans believe essential issues are at stake in the political arena.
+Associated with this is a belief that not just policy preference, but moral right and wrong, are at stake.
+& Military professionals, who hold multiple institutional loyalties, face an increasingly difficult task of balancing these loyalties, especially as the objects of these loyalties are increasingly associated with or attached to notions of moral right and wrong.
+\\ \hline
 \end{tabular}
@@
 
-The first way in which the concept of military professionalism is challenged as a result of increasing political polarization is that there is increasing difficulty for individuals to remain politically neutral, non-partisan, or otherwise apolitical - a hallmark of military professionalism.
-This happens because as "conflict extension" [@layman_party_2002] occurs, an increasing number of issues are brought into partisan battlegrounds, and thus, there is less politically-neutral territory.
-When polarization levels in America were not as high as they are today, it was more possible for individuals to either stay out of politics, avoid partisan squabbles, and otherwise to stay politically neutral.
-Today's polarizing environment seems to have the opposite impact: polarized politics seem to thrust issues and the unfolding political fight at the individual citizen, even when the individual is not deliberately seeking confrontation.
-There is no theoretical reason to suspect that this mechanism would not impact military leaders.
-Military leaders are citizens of the Republic, and though the institutional ethos of the military profession is to remain apolitical, such an ideal is difficult to uphold if less and less neutral political space exists.
+The first way in which the apolitical and non-partisan ethos of the military is challenged as a result of increasing political polarization is that there is increasing difficulty for individuals to remain politically neutral, non-partisan, or otherwise apolitical - a hallmark of military professionalism.
+This happens because as "conflict extension" [@layman_party_2002] occurs, an increasing number of issues are brought into partisan battlegrounds, and thus, there is simply less politically-neutral territory available.
+Thus, an actor who wishes to remain non-partisan, such as the military, may be increasingly less able to do so, as a result of this feature.
 
-A second challenge to military professionalism stems from the "negative partisanship"[@abramowitz_great_2018] that dominates the polarized American political landscape.
-Strong feelings towards those on a host of other issues are becoming increasingly hostile.
-These feelings are growing increasingly strong likely because of the other prominent features of political polarization: a decrease in the availability of neutral political space, and the importance of issues that are at stake in many of today's political debates.
+A second challenge military leaders and the institution faces in a polarized climate is from the "negative partisanship"[@abramowitz_great_2018] that dominates the polarized American political landscape.
+As citizens hold increasingly strong and negative feelings towards those who side differently on a host of issues, there may be a tendency for members of the military institution to harbor similar feelings.
+Even if impact of this feature on the military and its leaders, due to training and professional development that emphasizes the importance of upholding apolitical and non-partisan behavioral norms, is somewhat muted, the point is that members of the military institution are likely experiencing the same phenomenon - even if it is to a lesser degree - as the country's citizens.
 
-A third, and I contend a more insidious, mechanism through which increasing political polarization works to impact military professionalism is by altering one's sense of loyalty and identity.
-This occurs because the issues that are at stake as political polarization increases are in many instances involving issues that individuals associate with core concepts of one's identity, and thus, one's sense of moral right and wrong.
-Because individuals care about their identities and what they believe to be right and wrong, they are willing to defend these values, even at a cost.
-The rise of "identity politics" in America as political polarization has worsened illustrates this point.
+A third, and I contend the most impactful, challenge that normative military professionalism faces stems from the belief by individuals that "core" issues are at stake in the current polarized political climate.
+"Core" issues are difficult to define precisely [need cite], but the main point is that these issues often involve perceptions of moral right and wrong, and not just simple policy preferences.
+Moreover, some of the most polarizing issues are often associated with issues related to an individual's identity, such as race, gender, immigration [need cite].
+The rise of "identity politics" in America as political polarization has worsened illustrates this point [need cite].
 
-However, it is also precisely on this point that the military institution is perhaps most vulnerable to increasing political polarization.
-I contend that the rise of identity politics can and will be harmful if a military professional's identity and the loyalties associated with that identity directly conflict with his or her identity and loyalty as a professional soldier.
-As any soldier knows, and as civilians can reasonably understand,few traits are more sacred in the military than loyalty.
-Indeed, the virtue of loyalty is extolled in a host of codes and mantras that have played a central part of the American military since its inception, and loyalty is captured, either in word or in spirit, in a host of phrases and professional codes.
-In the middle of the West Point Cadet Prayer is the eloquently written plea, "make us to choose the harder right instead of the easier wrong" [@wheat_west_1918].
-Even Huntington spent a great deal talking about the importance of loyalty as a trait to the very essence of the military profession.
-He recognized that in order for civilian control to be effective, the military profession had to organize itself as a "hierarchy of obedience," which required "loyalty and obedience" to stand out as "the highest military virtues" [@huntington_soldier_1957, 73].
+I argue that this third challenge is especially relevant for the concept of loyalty.
+As an increasing number of what individuals believe to be "core" issues are raised and debated in the partisan arena, individuals face decisions of how to prioritize the persons or objects of their loyalty and further, what actions to take on the basis of loyalty to these persons or objects.
+I further theorize (though this has to be tested, of course) that while all individuals will likely face some version of this challenge of loyalty in a polarizing age, the challenges will be more acute for military leaders.
+Not only is the de-facto setting in which military leaders operate marked by multiple and often conflicting loyalties due to the structure of and statutes set forth in the US Constitution, but also, the potential consequences are higher for military leaders than average citizens should they fail to discharge the duties that correspond to their roles as professional officers.
+These consequences include failure to achieve tangible national security goals, such as loss on the battlefield, or less tangible but still important outcomes such as maintaining trust with the American public.
 
-High levels of political polarization obfuscates the ideal of loyalty, however, and thus pose a significant risk to the professional nature of the military.
-As core issues that involve individual perceptions of moral right and wrong are at stake in an increasing number of partisan battles, individuals, to include military officers, are more likely to become attached to or otherwise invested in these issues.
-This is especially true when the justification for taking or not taking action invokes the concept of loyalty.
-As a result, highly polarized political climates increase the likelihood that two different military officers may adopt completely different opinions or stances on an issue, each believing that the other is not only morally wrong, but failing to act out of loyalty in accordance with a specific oath or moral code.
+Moreover, as important as these consequences are, they do not speak to the potentially high personal and individualistic consequences military leaders may face if confronted with a situation in which they have to execute orders which are technically legal, but morally borderline at best.
+Military leaders are still humans who have to live with themselves at the end of the day.
+Therefore, I argue that if polarization continues to erode boundaries of what is considered moral by individuals to begin with, there is an increased likelihood of military leaders having to choose between conflicting loyalties.
+What happens if and when the objects of a military leader's multiple loyalties not merely disagree, but become stubbornly divided ideologically?
+Can such an outcome exist, and if it can, is there empirical proof that such a trend is occurring?
+If so, what are the normative implications, both for the military as a profession, and for the nation's security, and are there actions that can be taken to halt such a trend?
 
-# Accounting for Institutional Credibility
+## How Polarization Interacts with Institutional Credibility
 
-The previous section described how increasing political polarization presents significant challenges to the concept of military professionalism.
-But when will these challenges become public knowledge, and perhaps more relevant, how does increasing political polarization impact the credibility of the military institution?
+The previous section described how increasing political polarization presents significant challenges to aspects of the normative concept of military professionalism.
+But whether these challenges enter the public domain, or remain private challenges that specific individual actors contend with on a personal level only, matters a great deal because it "changes the character...from a matter of principle to a political matter" [@dubik_principled_2014] In fact, in much of the field of civil-military relations, this dynamic of private disagreement or discontent potentially entering the public realm is seen in the scholarly literature on senior officer resignation.
+This debate examines various aspects of resignation, including whether there can ever be conditions under which resignation is justified or appropriate, with scholars taking various sides on the issue, both for [@dubik_principled_2014; @snider_strategic_2014] and against [@feaver_should_2014; @golby_beyond_2015].
+I contend that polarization and credibility of the military interact, however, in at least two important ways.
+First, as loyalties are challenged as a result of increasing political polarization, individual military leaders may not necessarily remain quiet despite the existence of a strong professional norm of doing so because the objects of loyalty may demand that their adherents publicly speak out.
+This of course, depends specifically on what the object of loyalty is and what it demands from those who hold closely to it.
+The second way in which polarization and credibility may interact arises more directly from what can be termed a credibility effect, and has more to do with the place of prominence (or lack thereof) that the military or individual leaders hold in the perception of American society.
+[need to blend this in with recent scholarship and where gaps are - bring in Urben, Golby, Dropp and Feaver, etc., and Robinson.
+Also need to figure out where to put my hypotheses]
 
-One can imagine several possible scenarios.
-It is reasonable to consider that as political polarization increases, respect and trust in seemingly partisan institutions would decrease, while trust in nonpartisan and apolitical institutions would rise.
-But why?
-Would such a trend reflect a genuine increase in trust in the nonpartisan institution, or would it reflect the simple consequence that as one institution becomes less respected, others might simply gain as a result?
-Certainly, more work needs to be done in this area.
-
-Nonetheless, I contend that when institutional credibility is high, a demand signal is created by the American public such that future interactions between the military institution and the American public are encouraged and welcomed.
+I contend that when institutional credibility is high, a demand signal is created by the American public such that future interactions between the military institution and the American public are encouraged and welcomed.
 From an institutional perspective, the military might respond to the public's demand signal for increased military interaction by increasingly participating in public ceremonies, parades, and other showcases of military prowess (at sporting venues, for instance).
@@
 \doublespacing
+
 The second possible set of conditions is when political polarization is low and institutional credibility is high.
@@
 
+# The Analytical Framework - Hypotheses
+
+In order to test my argument, I will have to examine and test several hypotheses, which are inter-related and linked. They include the following:
+
+_H1: In highly polarizing times, there will be an increase in the frequency of military officers breaking normative civil-military relationship behaviors, such as criticizing the President, members of Congress, or those who hold to different political views on particularly-polarizing issues outside of the conduct of war._
+
+_H2: These breaks in normative civil-military behaviors will occur because political polarized periods feature events or situations which test or challenge the military's ability to juggle and balance multiple simultaneous loyalties: to the President as Commander in Chief, to Congress in its oversight role of the military, to the underlying ideologies or ideas inherent in issues that prevail during polarized times, to the normative professional ideals of the military, and to the individual's bonds of loyalty to his or fellow soldiers and the mission._
+
+_H3: Because of the normative strength and prominence of place that the ideal and concept of loyalty has within the profession of arms, military leaders are more sensitive than regular civilians to appeals of loyalty._
+
+_H4: Following from the above hypothesis, polarized periods will feature events in which governmental actors appeal to the military's sense of loyalty in order to persuade them to adopt a stance on a particular issue area._
+
+_H5: In highly polarized time periods, it is the senior military officers as the "stewards of the profession" [@us_department_of_the_army_adp_2019, p.1-9] whose actions will most influence to what degree and how the military institution changes.
+If changes occur to the military institution, they are bound to occur within two primary spheres: the military's professional identity, and the military's professional coherence._
+
 # Plan of the Dissertation and the Data Generating Process
 
-To complete my dissertation, I plan on incorporating a mixed-methods approach that relies on first, proposing a theory for how political polarization impacts and affects the overarching structure of loyalty that military officers - and the military as an institution - face.
-I then seek to test that theory using both empirical and quantitative data, as well as a series of qualitative case studies.
-I plan to conclude by discussing the implications of my findings.
-My dissertation will be a book style project composed of at least seven chapters.
+To complete my dissertation, I plan on incorporating a mixed-methods approach.
+In broad conceptual terms, I plan to propose a theory and then test the theory using a combination of quantitative analyses and qualitative case studies.
+The one sentence thesis of my dissertation is that political polarization significantly impacts how the military thinks about and exercises loyalty, a chief military trait, and that these impacts can and often do result in changes to the military's professional identity and character.
+I plan to execute a book style project comprised of at least seven chapters.
 
-In Chapter 1, I plan to frame my research question, lay out my research agenda, and review the relevant literature for my research project.
-In this introductory chapter, I will motivate the project by pointing out recent cases in which the American military institution has been at the center of partisan political struggles.
-My point in making this claim is not to argue that this place of political contention is new; indeed, it is not.
-There have been many times and instances in which the military has been at odds with other American political institutions and actors.
-My motivation, rather, is to argue that the behavior of the military institution itself - to include active and retired military officers who serve as elites - are demonstrating behavioral patterns that are at odds with traditional norms of military professionalism.
-This chapter will include an extensive review of the concept of military professionalism, a concept first coined by Samuel Huntington, and an extensive review of the concept of military professionalism, first coined by Samuel Huntington, and a series of critiques that Huntington's concept has endured over the years [@janowitz_professional_1960; @finer_man_1962; @feaver_civil-military_1996; @brooks_paradoxes_2020].
+In the first chapter, I will frame my research question, lay out my research agenda, and review the relevant literature for my project.
+This will involve synthesizing several of the key aspects and debates within the field of civil-military relations, much in the same way I have done in the first part of this dissertation prospectus.
+I will focus my attention on introducing and summarizing the concepts of military professionalism, organizational loyalty, and political polarization, and spend time orienting the reader to the debate within scholarly literature regarding a Huntingtonian conception of military professionalism.
+I will emphasize the long-running nature of this debate going all the way back to when Huntington first published his book, but point to some of the prominent contemporary critiques regarding his theory [@feaver_civil-military_1996-1; @rapp_civil-military_2015; @golby_why_2017; @karlin_case_2020; @brooks_paradoxes_2020].
+The main point of chapter 1 is to set the stage for my dissertation and signal to the reader that modern critiques of Huntington may be overlooking a key part of his concept of military professionalism: it may insulate the military from the most pernicious effects of polarizing politics.
 
-In Chapter 2, I plan to empirically show the deviation in military professionalism that Chapter 1 describes.
-Using quantitative analysis of an original data set that I collected this past Summer of more than 340 opinion-editorial pieces written by retired military officers since 1979 and published in top American newspapers, in print and online, I will demonstrate that the topics upon which retired military officers are writing, as well as the tone and tenor of their remarks, have changed in a statistically meaningful way.
+In the second chapter, I will introduce, develop, and explain my theory that links increasing political polarization to conceptions of loyalty and military professionalism (see Figure 2).
+I will argue that specific features of the polarized environment strain and challenge both individual and institutional conceptions of loyalty, and by extension, normative aspects of military professionalism (Table 1).
+I will further argue that rising political polarization can interact with changes in the military's institutional credibility to increase the likelihood that the public becomes aware of changes in the military's professional conduct.
+Using Table 2 as a framework, I will place epochs of historical American civil-military relations into one of four distinct quadrants, and signal that these quadrants will comprise the qualitative case studies that I later examine in my dissertation.
 
-In Chapter 3, I want to develop a theory for why these changes have occurred.
-My theory will argue that increases in the levels of political polarization and institutional credibility/prestige have driven these changes.
-At the heart of my theory is the idea that certain features of political polarization have so altered the political preferences of some military elites such that significant changes have occurred to fundamental concepts and structures of loyalty that exist within and outside of the military as an institution.
-My theory up until this point does not explain, however, why these divergences in political preferences have risen to the level of public knowledge, which I attempt to show in Chapter 2.
-Therefore, the other key feature that I need to incorporate into my theory is that the public credibility or prestige of these military actors has also risen, enabling them to make arguments to the public without fear of backlash.
-This portion of the theory is an argument about the prestige of military elites as well as of the military institution, and how that prestige translates into the ability to message to the public.
+In the third chapter, I will argue that today, in 2020, the American public is witnessing a decline in professional military conduct among retired military officers.
+I plan to demonstrate this empirically by drawing upon an original dataset consisting of more than 340 op-ed pieces authored by retired military officers since the late 1970s and published in the top five major print newspapers in the United States, including the Wall Street Journal, the New York Times, the Washington Post, the USA Today, and the Los Angeles Times.
+I believe that I can show several things from this analysis, including that there has been a sharp uptick in the number of op-eds published by retired military officers during this time period, and more importantly, that there has been a decline in the normative professional ideals espoused and argued for in these op-eds.
+Critical to this chapter is acknowledgement that coding for what is professional versus what is not is subjective to a large extent, and thus, the need for me to truly develop clear working definitions of the concept of military professionalism in Chapter 1.
+However, I will show that whether one adopts a strict Huntingtonian version of military professionalism, or a less strict version that allows for more encroachment by the military institution into the political sphere, by both standards the retired military officer corps is not living up to the ideals of normative military professionalism.
+The question, then, becomes why aren't they?
+I will argue that while there are certainly other confounding variables, to include changes in the conservative-liberal tilt of newspapers over time, the chief explanation for the drop in professionalism stems from the combined effects of increasing political polarization and rising institutional credibility of the military since the late 1970s, and the challenges that these changes introduce to a military leader's conception of loyalty.
+Moreover, my ultimate point is to point out a blatant inconsistency in the way that the military currently thinks about the political behavior by those within the ranks in that the institution has rules that governs the conduct of those on active duty [@department_of_defense_political_2008], but those same rules, and the spirit behind those rules, are disregarded by the institution's vangaurds, its retired multistar officers and "stewards of the profession" [@us_department_of_the_army_adp_2019, p.
+1-9] of arms.
+
+In the fourth chapter, I want to show that the institutional challenges to loyalty have been present as political polarization has worsened.
+To do this, I plan to examine, via textual analysis, the records of Congressional Committee hearings in which military officers, active and retired, have played a part, and detect if and how patterns of questioning and providing answers have changed over time.
+I am especially interested in the way that individuals have invoked language associated with the concept of loyalty in asking and answering questions.
+An example of this might be analyzing the ways in which terms such as "moral obligation" or similar language is used in these hearings.
+I hope to detect patterns that will help support my overall argument.
+I have yet to flesh this particular chapter out some more.
+
+In Chapter 5, I plan to analyze the results of an original survey experiment in which I survey a nationally representative sample of civilians and compare the results to a sample of either a) veterans, or b) active duty military officers.
+In this survey experiment, I plan to test whether members of the military (which I proxy by veterans) genuinely are more sensitive to appeals to loyalty, as I have hypothesized in _H3._ The survey, after controlling for gender, partisan identification, and age, will present the respondent with 6-8 hypothetical situations and a corresponding argument made by a military elite.
+Members of the treatment group will receive an appeal to loyalty as part of the elite's argument as to why a certain course of action should be followed regarding the hypothetical situation presented.
+
+I plan to structure the survey such that the scenario and corresponding course of action contain references to issues that are highly correlated, in some cases, to partisan / ideological affiliation.
+An example of a question might be as follows:
+\singlespace
+* Situation: The President of the United States, a [Republican, Democrat] recently commented that he is seriously contemplating ordering active duty military Troops to assist local police and National Guard units in several US cities in maintaining law and order as a result of initially-peaceful protests which turned violent in some, but not all, cases.
+These protests have been going on for more than two weeks and are largely, but not exclusively, affiliated with [the Black Lives Matter Organization, a pro-2nd Amendment organization].
+
+* Elite Voice: A retired US Army 4 star general, who has extensive combat experience in Panama, Iraq, and Afghanistan, recently penned the following op-ed piece in a prominent newspaper:
+
+* “The military must not be ordered to put down riots and protests, even if a small number of them have, unfortunately, turned violent.
+Even if the President is legally able to issue such an order, the military’s fundamental purpose is to defend the nation from external threats, and not to endanger the Constitutional rights of everyday Americans.
+[Treatment group, add ]: Should the President order the active military to deploy to cities where these protests have turned violent, military officers should resist such an order and if necessary, resign their commissions.
+They owe a higher loyalty to the Constitution, and the President’s order would clearly violate the oath of office these officers have sworn to uphold.”
+
+* Do you find this argument compelling (7 point scale of (1) not at all - (7) absolutely, very compelling).
+
+\doublespace
+Such a survey will enable me to measure two things.
+First, it will allow me to measure whether appeals to loyalty are truly more effective than standard appeals to action.
+This result can be measured by comparing the answers between the treatment and control groups.
+Second, it will allow me to measure whether members of the military (or veterans) react differently to these appeals to loyalty than regular civilians.
+As I have hypothesized in _H3_, I suspect that they do.
+Of note, I am planning to also pursue a survey among actively-serving military officers, but the lead time for obtaining approval to do this can be very long, perhaps 9-12 months or more.
+If I am able to obtain approval for this, I will then have three different populations to measure: civilians, veterans, and actively-serving military officers.
+Of note, if I discover throughout the course of my survey that loyalty does not in fact have the strong explanatory power that I think it does, or if veterans and/or active military officers do not respond to appeals to loyalty any differently than civilians, I think such a finding is also interesting.
+
+In the sixth chapter, I will turn to a series of qualitative case studies to explore and test my theory.
+Key in this chapter is finding and examining historical periods in American civil-military relations in which there existed varying levels of political polarization and institutional credibility, and then, to walk the reader through how challenges of loyalty during these periods manifested themselves in different ways among specific members of the military and the institional military at large, in line with the theory that I propose in Chapter 2.
+My initial projection of what these case studies will be and how they will align with varying levels of political polarization and instititional credibility are as follows:
+
+* High Polarization, High Institutional Credibility: civil-military relations in the Andrew Johnson administration after the American Civil War.
+I will explore how General Ulysses Grant confronted challenges of loyalty to multiple institutions and objects as he dealt with a President who was by many accounts sympathetic to southern interests as the US Army played a prominent role in reconstruction in the south [@hyman_santon_1960].
+
+* High Polarization, Low institutional credibility: civil-military relations in the post-Vietnam period, in the late 1970s.
+
+* Low Polarization, High Institutional Credibility: civil-military relations after World War Two, in the late 1940s and early 1950s.
+
+* Low Polarization, Low Institutional Credibility: TBD.
+
+In the seventh and final chapter, I will conclude by exploring the implications of my argument and by
+
+\singlespace
+\pagebreak

### 2020-10-02 code/prospectusv1.Rmd
no message

@@
 According to Huntington, the ultimate goal, or dependent variable, of civil-military relations in the American context is achieving both military security and maintaining civilian supremacy over the military.
-Huntington viewed the key independent variable that impacts and enables these goals while also resolving the tension generated by the opposing functional and societal imperatives, as described above, the adoption of a model of "objective civilian control" of civil-military relations, whereby political and military leaders "maximize military professionalism," which includes granting the military a high degree of professional autonomy [@huntington_soldier_1957, 80-85].
+Huntington viewed the key independent variable that enables the achievement of these goals while also resolving the tension generated by the opposing functional and societal imperatives as the adoption of a type of civil-military relations framework which he calls, "objective civilian control" of the military [@huntington_soldier_1957, 83].
+In such a model of "objective civilian control," Huntington envisioned civilians granting the military a high degree of professional autonomy, which he believed would simultaneously be "most conducive to the emergence of professional attitudes and bahvior among members of the officer corps" as well as the best way to minimize the military's political power by rendering the military "politically sterile and neutral" [@huntington_soldier_1957, 83-84], and thus preserving civilian supremacy.
 
@@
   geom_dag_text()+ 
-  geom_dag_label_repel(aes(label=label, fill=label), col="white", show.legend=FALSE)+ theme_dag() + labs(title="Figure 1: Samuel Huntington's Theory\n Of Civilan Control of the Military", subtitle="As espoused in The Soldier and The State, 1957")
+  geom_dag_label_repel(aes(label=label, fill=label), col="white", show.legend=FALSE)+ theme_dag() + labs(title="Figure 1: Samuel Huntington's Theory Of Civilan Control of the Military", subtitle="As Explained in The Soldier and The State, 1957")
 
@@
 Huntington's argument in support of "objective control" of the military stands in contra-distinction, as he explains, to an alternate model of "subjective control" in which civilian leaders limit the power of the military by treating it as any other civilian interest group [@huntington_soldier_1957, 80-83].
-In fact, Huntington admits that the US Constitution actually envisioned this "subjective" scheme of civilian control of the military [@huntington_soldier_1957, 163-168].
+In fact, Huntington admits that the US Constitution actually envisioned and established a form of subjective vice objective scheme of civilian control of the military [@huntington_soldier_1957, 163-168].
 What the framers did not see, nor could they, argues Huntington, was the rise of the military as a profession in response to ever-increasing and dangerous external threats.
-It was because of these pernicious threats, argues Huntington, which the US military had witnessed, confronted, and battled during the 19th and 20th centuries during World War One, World War Two, and now the Cold War, that the military profession had risen as a separate and distinct body that could focus on the demands of military service.
-Such a separation and distinctness, claims Huntington, would need to be embraced by civilians and the militarym going forward into the future.
-In so doing, claims Huntington, is the only way that both military security and civilian control of the military - two often-competing imperatives - will be safeguarded.
+It was because of these pernicious threats, argues Huntington, which the US military had witnessed, confronted, and overcome during the 19th and 20th centuries in World War One, World War Two, and as Huntington writes, the Cold War, as well as the destructive capability of modern weapons and warfare in general that the military profession had emerged [@huntington_soldier_1957].
+Huntington claims that unlike during the nation's founding, military officers constitute a profession, and thus, are unique and distinct from other occupations in several important ways: professions practice and develop a special body of knowledge and expertise; professions exercise a  "social responsibility" by providing a vital service to society; and professions have a degree of "corporateness" in that they develop and formalize standards by which members of the profession are certified to practice [@huntington_soldier_1957, 11-18].
 
-Numerous critiques surrounding the Huntingtonian ideal of _military professionalism_ have been levied over the years, as Huntington did not attempt to precisely define it.
-Most of these critiques, in one way or another, charge that a strict theoretical and practical separation between military and political operating spheres is fundamentally flawed.
-Many, if not all, of these critiques cite Clausewitz's famous dictum that war is a "mere continuation of policy by other means" [@clausewitz_war_1976, 69, 605], charging that war by its very nature requires the highest integration - not separation - of military and political activities.
-In essence, these critiques argue that the strict separation of the responsibilities between the military and civilians leads to poor strategic outcomes [@rapp_civil-military_2015; @brooks_paradoxes_2020].
-Moreover, these criticisms can and do point largely to American failures in Iraq and Afghanistan to secure, at least as of yet, a more-lasting peace - failures that are ultimately, according to one scholar and former general, moral in nature, and shared between both civilians and the military [@dubik_just_2016].
+Numerous critiques surrounding the Huntingtonian conception of _military professionalism_ have been levied over the years.
+Most of these critiques, in one way or another, charge that a strict theoretical and practical separation between military and political operating spheres, which Huntington to a certain degree certainly argued for, is fundamentally flawed.
+Many, if not all, of these critiques cite Clausewitz's famous dictum that war is a "mere continuation of policy by other means" [@clausewitz_war_1976, 69, 605], charging that by its very nature, war requires the highest integration, as opposed to segregation, of political and military activities.
+Moreover, these critiques often point out, correctly in this author's opinion, that the conduct of war is more complicated than civilians simply telling the military the goals it wishes to achieve, and then allowing the military to determine how best to accomplish these goals.
+Rather, the development and implementation of strategy, especially during a time of war, requires an "iterative" dialogue between military leaders and civilians to ensure that effective policies are adopted [@rapp_civil-military_2015, 20].
 
-As an extension of the critiques of Huntington, there has also been renewed focus and attention among civil-military relations scholars on numerous professional normative standards of the military.
-One of the most important -  yet difficult - debates is an attempt by scholars to define where appropriate political behavior ends and partisan activity begins [@owens_military_2015; @golby_beyond_2015].
-In the Huntingtonian framework, this is essentially the opposite side of the debate over the same coin in attempting to define for civilians what constitutes meddling in military affairs.
-One of the reasons this debate has become even more difficult, I believe, is because as political polarization increases, the lines between partisanship and politics, as well as politics and everything not politics, have blurred, as so much of the literature on American politics and polarization affirms.
+Others argue that Huntington's conception of military professionalism is too narrow to be of practical use in the contemporary period in that such a model actually enables, rather than constrains, inappropriate political activity by military officers [@brooks_paradoxes_2020].
+Risa Brooks charges that military officers who strongly identify with an ethic that stresses remaining apolitical at all costs, which she alleges follows from a Huntingtonian conception of military professionalism, may develop "blind spots such that they fail to recognize the political content or impact of their actions" [@brooks_paradoxes_2020, 17].
+Whether military leaders recognize it or not, Brooks argues, military leaders comprise a profession that is inherently political, and such a reality ought to be acknowledged and embraced [@brooks_paradoxes_2020].
+
+Regardless of whether and the degree to which one agrees or disagrees with Huntington's concept of military professionalism, a significant number of scholars note that a significant issue involves attempting to define where the line between appropriate political expression and inappropriate partistan activity should be drawn [@owens_military_2015; @golby_beyond_2015; @brooks_paradoxes_2020].
+It is a difficult task for many reasons because, as Mackubin Thomas Owens points out, civil-military relations in the United States can be viewed as a "bargain" [@owens_us_2011] between the American people, the military, and the government that has to account for constantly-changing "political, social, technological, or geopolitical changes" [@owens_is_2016].
+In other words, the ground is always moving for the actors involved in civil-military relations.
+
+One of these recent changes, I argue, is increasing political polarization that has engulfed not only government elites, but the nation at large.
+Brooks (2020) points out three challenges posed by the contemporary period to the military's apolitical ethic, all of which are likely related to the growing problem of political polarization: rising acceptance among military officers of behaviors that contradict an apolitical ethos; increasing "pressures and opporunities" for military officers to express partisan sentiment; and increased politicization of the military by government elites [@brooks_paradoxes_2020, 21-23].
+Indeed, as political polarization increases, the lines between partisanship and politics, as well as politics and everything not politics, have blurred, as so much of the literature on American politics and polarization affirms, promising that this challenge with respect to the military institution and its members will likely persist for some time.
 
@@
 
-Loyalty features prominently in the international relations and civil-military relations literatures, often in a comparative sense and in the context regime exclusion and a host of related phenomena, including military coups [need cite], civil wars [@mclauchlin_loyalty_2018], and the difficulties that multi-ethnic states face in maintaining a professional military [@peled_question_1998].
+As a subject or concept, loyalty features prominently in the international relations and civil-military relations literatures, often in a comparative setting and in the context of coups [@marcum_overthrowing_2016; @johnson_squeaky_2018], civil wars [@mclauchlin_loyalty_2018], and the difficulties that multi-ethnic states face in maintaining a professional military [@peled_question_1998].
 This literature deals with a host of important issues that stem from problems associated with ethnic identification and loyalty, either to a regime, cause, or leader.
-Though the US context is very different from the contexts usually explored in this literature, its insights may help me identify related dynamics and/or mechanisms that perhaps are unfolding in the US: how are changes in partisan or ideological identification (instead of ethnic) as a result of political polarization impacting conceptions of loyalty within the military.
-Because appropriate to explore partisan or ideological (instead of ethnic) identification stemming from However, this literature often assumes a definition of loyalty and rarely develops how the reader should think of loyalty in the first place.
-For this fundamental starting point of defining loyalty, the fields of ethics, management, and sociology are helpful.
+Though the US context is very different from the contexts usually explored in this literature, its insights may help me identify related dynamics and/or mechanisms that are unfolding in the US: how are changes in partisan or ideological identification (instead of ethnic) as a result of political polarization impacting conceptions of loyalty within the US military, if at all?
+To even begin answering this question, one has to first define what loyalty is and how it works.
+For this fundamental task, the fields of psychology, ethics, management, and sociology are helpful.
 
-Loyalty has been conceptualized as both a virtue [@coleman_problems_2009] and as a moral emotion [@connor_military_2019].
-Though the differences between the two approaches are important, both assert that loyalty is something felt by an individual and subsequently reflected in action, such as, in the military context, "the desire to protect each other from harm and to defend each other from attack" [@coleman_problems_2009, p.
-109].
-Oldenquist, a philosopher, describes loyalty as a devotion or attachment that an individual has to someone or something specific such that one refers to the object of loyalty as "mine" and is inclined to feel "pride when it [the object of one's loyalty] prospers, shame when it declines, and anger or indignation when it is harmed" [@oldenquist_loyalties_1982, p.
-175, brackets mine].
-
-Loyalty further generates a willingness in the individual to assume "obligations" [@oldenquist_loyalties_1982, p.
-175] to the object of one's loyalty, and these obligations can take multiple forms, such as a willingness of the individual to endure hardships and make sacrifices [@fielder_organizational_1992].
+Loyalty has been conceptualized as both a virtue [@coleman_problems_2009] and as a moral emotion [@connor_military_2019], the main difference between the two being that the former is a normative aspirational trait, while the latter is a primed behavioral response.
+For the purposes of my dissertation, these differences are not consequential, because both assert that loyalty is something that is ultimately felt by an individual and subsequently reflected in action, such as, in the military context, "the desire to protect each other from harm and to defend each other from attack" [@coleman_problems_2009, 109].
+Oldenquist, a philosopher, further describes loyalty as a devotion or attachment that an individual has to someone or something specific such that one refers to the object of loyalty as "mine" and is inclined to feel "pride when it prospers, shame when it declines, and anger or indignation when it is harmed" [@oldenquist_loyalties_1982, 175].
+Loyalty also generates a willingness in the individual to assume "obligations" [@oldenquist_loyalties_1982, p.
+175] on behalf of the object of one's loyalty, and these obligations can take multiple forms, such as a willingness of the individual to endure hardships and make sacrifices [@fielder_organizational_1992].
 Scholars agree that individuals can hold multiple loyalties simultaneously, and recognize that this often can and does cause problems [@coleman_problems_2009].
@@
 We also know that the military services in the United States place a premium on loyalty as a trait and concept.
-Indeed, of the different military services, the US Army in particular has extensively developed, in a normative sense, a framework entitled "The Army Ethic," in which loyalty constitutes a significant part of the Army profession [@us_department_of_the_army_adp_2019].
-This ethic, which will be described in detail in later parts of this dissertation, is formally laid out and described in the Army's doctrinal publication entitled, _Army Leadership and the Profession, ADP 6-22_, and holds that as "stewards of the Army Profession" and as "committed professionals," Army leaders "embrace and uphold the Army Values and standards of the profession" [@us_department_of_the_army_adp_2019].
-There are a total of seven "Army Values," which together spell the acronym, l-d-r-s-h-i-p, or "leadership," and the first of these values, loyalty, is followed by six others: duty, respect, selfless service, honor, integrity, and personal courage [@us_department_of_the_army_adp_2019, p.
+Indeed, of the different military services, the US Army in particular has extensively developed, in a normative sense, a framework entitled "The Army Ethic," in which loyalty plays a significant role within the Army profession [@us_department_of_the_army_adp_2019].
+This ethic, which will be described in detail in later parts of this dissertation, is formally laid out and described in the Army's doctrinal publication entitled, _Army Leadership and the Profession, ADP 6-22_, and holds that as "stewards of the Army profession" and as "committed professionals," Army leaders "embrace and uphold the Army Values and standards of the profession" [@us_department_of_the_army_adp_2019].
+There are a total of seven "Army Values," which together spell the acronym, l-d-r-s-h-i-p, which is short for the word "leadership," and the first of these values, loyalty, is followed by six others: duty, respect, selfless service, honor, integrity, and personal courage [@us_department_of_the_army_adp_2019, p.
 1-9.].
-The US Army states in its doctrinal manual on Army Leadership and the Profession, ADP 6-22, that the Army Values "embody the practical application of the Army Ethic.
+
+ADP 6-22 further states that the Army Values "embody the practical application of the Army Ethic.
 They encompass the enduring moral principles, beliefs, and laws that guide Army professionals in accomplishing the mission as well as their conduct in all aspects of life" [@us_department_of_the_army_adp_2019, p.
@@
 The virtue of loyalty is likewise extolled in a host of codes and mantras that have played a central part of the American military since its inception.
-In the middle of the West Point Cadet Prayer is the eloquently written plea, "make us to choose the harder right instead of the easier wrong" [@wheat_west_1918].
-Even Huntington spent a great deal talking about the importance of loyalty as a trait to the very essence of the military profession.
-He recognized that in order for civilian control to be effective, the military profession had to organize itself as a "hierarchy of obedience," which required "loyalty and obedience" to stand out as "the highest military virtues" [@huntington_soldier_1957, 73].
+In the middle of the West Point Cadet Prayer is the eloquently written plea, "make us to choose the harder right instead of the easier wrong" [@wheat_west_1918], indicating that to be loyal to what is right is not only preferable but often a difficult endeavor.
+Even Samuel Huntington recognized that in order for his model of civilian control to be effective, the military profession had to organize itself as a "hierarchy of obedience," which required "loyalty and obedience" to stand out as "the highest military virtues" [@huntington_soldier_1957, 73].
 
-Loyalty is further important in a military setting in a broader and strategic sense, especially when one considers that the US military, both by statute and by normative standards, serves as an apolitical and non-partisan institutions with multiple, simultaneous loyalties.
-The Constitution of the United States, to which military officers and soldiers alike ascribe allegiance when they take their respective oaths of commissioning and enlistment, deliberately contrives the military as an institution possessing multiple loyalties.
-One of the most basic forms of these multiple loyalties is captured in the fact that the Constitution empowers the President as "the Commander in Chief of the Army and the Navy of the United States" [@noauthor_us_1787], but also gives Congress an array of tremendous powers and responsibility pertaining to the Armed Forces @[noauthor_us_1787-1].
-Indeed, the tensions that inherently exist in American's constitutional framework are explored in a robust literature that examines the multiple ways in which the Congress, the President, and the military interact to implement, but also to influence, American foreign policy [for instance, see @howell_while_2007].
-
-In the context of the US military as an institution, loyalty has rarely been used as a conceptual construct in the political science literature, likely and thankfully because the US military has not yet considered, or been involved, in a coup.
-Nonetheless, I contend that there is a rich opportunity to study strategic loyalty as a concept, especially in a contemporary context in the United States, where political polarization has increased extensively.
+Loyalty is further important in a military setting in a broad and strategic sense, especially when one considers that the US military, both by statute and by normative standards, serves as institution with multiple, simultaneous loyalties.
+The Constitution of the United States, to which military officers and soldiers alike ascribe allegiance when they take their respective oaths of commissioning and enlistment, contrives the military as an institution possessing multiple loyalties: to the President as "the Commander in Chief of the Army and the Navy of the United States" [@noauthor_us_1787], but also to Congress in that they likewise hold unique powers and responsibilities pertaining to the Armed Forces [@noauthor_us_1787-1].
+The tensions that inherently exist in American's constitutional framework with respect to national security are explored in a robust scholarly literature [for instance, see @howell_while_2007; @milner_sailing_2015].
 
@@
 
-One area of work that remains under-developed, both theoretically and empirically, is an exploration of the relationship between polarizing domestic politics and the professional identity of the military, both as an American institution and as a profession.
-Exploring this relationship will, I believe, help scholars in the field of civil-military relations and international relations more broadly better understand if and how the military, as a unique body charged with the defense of the nation, changes in fundamental ways as a result of a polarized political climate.
-On one hand, one would expect that it might, in the sense that the American military is, after all, a microcosm of American society.
-Therefore, one could expect that if society truly is becoming polarized, then at least to some degree, so is the military, or at least those who volunteer for service and at the point in time that they join.
-On the other hand, one might reasonably expect that the unique purpose of the military, combined with its long-standing professional ethic, might insulate the military institution from some of the potentially negative impacts of polarized politics.
-These differing expectations need to be more clearly articulated and explored.
+In the context of the US military, to my knowledge, loyalty has rarely been used as a conceptual construct in the political science literature, likely and thankfully because the US military has not yet considered, or been involved, in a coup.
+Nonetheless, I contend that there is a rich opportunity to study strategic loyalty as a concept, especially in the contemporary context of the United States' domestic political climate, as political polarization has increased extensively.
 
-I am specifically interested in the question of if and how political polarization impacts the concept of organizational loyalty in the military.
-This is an interesting research question because, as noted above, one of the most fundamental normative hallmarks of the military's professional nature is that it ought to be apolitical and non-partisan, and one of the features of militaries in general is their tendency to foster a concept of loyalty among those who serve.
-Yet a puzzle emerges: within an institution that is normatively apolitical and non-partisan, can it remain so in an environment marked by increasingly polarized domestic politics, where partisan loyalties are a salient feature of the political landscape [need cite here for partisan loyalties].
+One potential area of work that remains under-developed, both theoretically and empirically, is an exploration of the relationship between polarizing domestic politics and the professional identity and character of the military.
+I am specifically interested in the question of if and how political polarization impacts the concept of loyalty in the military.
+This is an interesting research question because it brings together at least three relevant strands of literatures which, when combined, generates an interesting set of puzzles and questions.
+First, one of the most fundamental normative hallmarks of the military's professional nature, at least by Huntingtonian standards, is that it ought to be an apolitical and non-partisan institution.
+Second, one of the features of the US military is that it prizes and elevates the trait of loyalty among those who serve and, by statute as well as normative tradition, exercises simulataneous dual loyalties to its civilian heads, the President as the Commander in Chief, and Congress.
+And third, we know from the American politics literature that ideological polarization and partisan loyalty are connected in that those citizens who hold to stronger ideological convictions tend to be the most loyal partisans in terms of voting and other political behaviors [@abramowitz_great_2018, 112-116].
 
-My intended dissertation will help answer several related questions, including: if personal and partisan loyalties are changing as a result of political polarization [need cite], is this also happening within the military institution, and why?
+If these things are true, then a question is whether rising political polarization will generate instances in which military officers must choose between which objects of loyalty to ultimately elevate and side with in the event that there is conflict between and among these objects.
+Moreover, at the level of the military institution, a similar problem may also emerge: will the military institution have to choose between competing objects of loyalty, such as the President as the Commander in Chief, the Congress, and even its own professional ethic?
+^ [I believe we have already seen such instances of exactly this situation playing out over the past several years as political polarization has worsened.As examples, consider the case of now retired Army Lieutenant Colonel Alex Vindman, who testified before Congress against the US President, Donald Trump, regarding what be believed to be behavior that may have been illegal by the President regarding aid to Ukraine, or the case of Navy Captain Brett Crozier, who leaked a letter to the press about how his pleas for help in battling the coranavirus aboard the ship he commanded, the USS Theodore Roosevelt.
+Certainly these instances involved individuals making choices about to whom or to what they would be most loyal to - the President, the Constitution, the ideals and obligations of the military, themselves, and perhaps more.]
+
+Other questions I wish to explore include whether there patterns in the way that loyalty obligations held by the military are in fact challenged by governmental actors during periods of increased political polarization, and if so, what are they, and how is the military reacting to these patterns?
+Are there impacts on the military's professional identity as a result of political polarization, and if so, from who or what?
 Are partisan/ideological loyalties displacing other loyalty relationships within the military?
-If so, are these changes in organizational loyalty having an impact on the military's professional identity and character?
 Through what mechanisms do these changes operate?
@@
 Is this problematic for the military profession, and why?
+These and other questions will shed light on how rising political polarization is impacting the professional identity of the military.
 
-The answers to these questions, which I intend to include as part of my dissertation, will help inform scholars and policymakers alike on number of important issues.
-First, it will help illuminate in a broad sense an important aspect of how changes in political polarization impact American institutions, and in particular, the military.
-Second, work on this topic will attempt to explore how a fundamental concept and value within the military - loyalty - is both impacted as a result of political polarization but also how loyalty changes may be driving changes in the military's professional character, identity, and nature.
-In doing so, it will help enrich a scholarly understanding of what loyalty is to begin with, and how it features prominently in the context of the American military.
+The gaps in the current literature, which I hope to advance and fill, involve the impact of political polarization directly on the military institution itself, and why the military experiences such impacts.
+A number of excellent and recent scholarly works have looked at aspects of this question, but from an angle that examines not the military directly, but rather society's perception of the military or various aspects of the military as a result of polarized politics.
+A number of these works have directly examined the impact of partisan affiliation on the credibility of military actors, resulting in some very interesting findings.
+Michael Robinson, for instance, found that members of the public did not sanction military elites for making public messages that accorded with a member's partisan views, and that the public only sanctioned those elites whose messages were in line with the opposite political party [@robinson_danger_2018].
+Previously, Heidi Urben found that among active military officers, there was weakening support for the norm that retired military officers should remain apolitical or non-partisan in retirement [@urben_party_2013].
+These scholars' findings suggest, as others have pointed out [@brooks_paradoxes_2020], that political polarization at least coincides with, if not causes, a weakening in support among both the public and among service members for the military's normative apolitical ethos.
+Other scholars have also unpacked a trove of information regarding public opinion and the military, including how partisan affiliation shapes differences in support for the use of military force [@golby_jim_elite_2017] and military elite endorsement of political candidates [@golby_military_2012].
+
+Focusing attention on the military institution will help scholars in the field of civil-military relations and international relations more broadly better understand if and how the military, as a unique body charged with the defense of the nation, changes in fundamental ways as a result of a polarized political climate.
+On one hand, one would expect that it might, in the sense that the American military is, after all, a microcosm of American society.
+Therefore, one could expect that if society truly is becoming polarized, then at least to some degree, so is the military, or at least those who volunteer for service and at the point in time that they join.
+On the other hand, one might reasonably expect that given the unique purpose of the military, combined with its long-standing professional ethic, might insulate the military institution from some of the potentially negative impacts of polarized politics.
+These differing expectations need to be more clearly articulated and explored.
 
@@
 And as American history proves, there have been countless instances in which civilian and military leaders disagreed.
-Many of these rise to the level of common knowledge: Lincoln fired McClellan because he lost confidence in him as a wartime Commander; Truman sacked MacArthur because the latter was insubordinate regarding the future of war in Asia; and Obama and McChrystal clashed about troop levels in Afghanistan.
+Many of these rise to the level of common knowledge: Lincoln fired McClellan because he lost confidence in him as a wartime Commander; Truman sacked MacArthur because the latter was insubordinate regarding the future of war in Asia; and Obama and McChrystal clashed about troop levels in Afghanistan, ultimately resulting in the former firing the latter.
 Beyond these episodes in which President and wartime commander locked horns, consider several other instances of "civilians" and "military" clashing.
@@
 
-My personal motivation in choosing the intended topic, however, stems from what I believe are changes in the type of civil-military disagreement that have surfaced in recent years.
-For example, I believe we are seeing portions of the military institution (particular retired military officers) openly disagree with civilian leaders on topics that I believe are, even by the widest of academic definitions, way outside of appropriate military expertise.^[I hope to prove this, empirically, in or as part of a chapter in my dissertation] More interestingly, these conflicts are occurring not in the context of major war, but the opposite: a period in which the levels of US Troops deployed in active combat zones is less than it has been in recent years.
-While I believe that such developments are catching the attention of civil-military relations scholars, most of the work done on this topic is, in my opinion, impressionistic [@golby_jim_keep_2020] as opposed to systematic.
-In other words, civil-military relations scholars have identified that polarized partisan politics serve as a risk for the military, but none to my knowledge have specifically set out to systematically explore if and how polarized partisan politics are impacting the military's professional identity and character, and how.
-
-Thankfully, a growing number of civil-military relations scholars have looked at various aspects of this question - the impact of the political polarization on civil-military relations - in recent years.
-But the focus of these works has been, I believe, not on the impact of polarization on the military itself but on other important but peripheral components of the military's relationship with American society.
-One recent dissertation completed by Michael Robinson at Stanford University in 2018, for instance, found that partisan affiliation significantly affects public credibility and trust in the military, perhaps resulting in or as a consequence of media "echo chambers" which may incentivize military leaders to offer opinions partisan opinions without fear of sanction [@robinson_danger_2018].
-Robinson's work builds off of previous work by Heidi Urben in which she found that active military officers believe less strongly that retired military officers should remain apolitical or non-partisan in retirement [@urben_party_2013], a norm that has existed for decades or more.
-Together, Robinson and Urben join other scholars in providing a wealth of information regarding public opinion and the military, and in particular, how aspects of political polarization shape partisan differences regarding topics such as the use of force [@golby_jim_elite_2017] and military elite endorsement of political candidates [@golby_military_2012].
+My intended topic is different than these other, still very important topics in which disagreements between the executive and the military occur during wartime.
+It is my assessment that polarization is causing conflict, or certainly has the potential to do so, between the military and civilian leaders, but in most instances, the subject is about topics other than the use of military force, which is somewhat puzzling, given the military's reason d'etre.
+For example, I believe we are seeing portions of the military institution (particular retired military officers) openly disagree with civilian leaders on topics that I believe are, even by the widest of academic definitions, outside the bounds of what constitutes appropriate military expertise.^[I hope to prove this, empirically, in or as part of a chapter in my dissertation] While I believe that such developments are catching the attention of civil-military relations scholars, none to my knowledge have specifically set out to systematically explore if and how polarized partisan politics are impacting the military's professional identity and character, and how.
 

### 2020-10-22 code/prospectusv1.Rmd
commit

@@
   _This dissertation seeks to explore the impact of increasing political polarization on the professional identity of military officers and the military institution.
-I develop a theory that argues that the nature of rising political polarization fundamentally challenges the conception of loyalty among members of the military profession, and that such a strain is problematic for an institution such as the military that holds the virtue of loyalty in high regard, and by statutes possesses multiple, simultaneous loyalties: to the Constitution of the United States, to the President qua Commander-in-Chief, to Congress, to the American People, to one's fellow service members, and to one's self.
+I develop a theory that argues that the nature of rising political polarization fundamentally challenges the conception of loyalty among members of the military profession, and that such a strain is problematic for an institution such as the military that holds the virtue of loyalty in high regard, and by statute possesses multiple, simultaneous loyalties: to the Constitution of the United States, to the President qua Commander-in-Chief, to Congress, to the American People, to one's fellow service members, and to one's self.
 Balancing these competing loyalties becomes more difficult as political polarization worsens, I argue, because of the fundamental and core values - not mere policy differences - that are at the root of many contemporary polarized political battles.
@@
 If these things are true, then a question is whether rising political polarization will generate instances in which military officers must choose between which objects of loyalty to ultimately elevate and side with in the event that there is conflict between and among these objects.
-Moreover, at the level of the military institution, a similar problem may also emerge: will the military institution have to choose between competing objects of loyalty, such as the President as the Commander in Chief, the Congress, and even its own professional ethic?
-^ [I believe we have already seen such instances of exactly this situation playing out over the past several years as political polarization has worsened.As examples, consider the case of now retired Army Lieutenant Colonel Alex Vindman, who testified before Congress against the US President, Donald Trump, regarding what be believed to be behavior that may have been illegal by the President regarding aid to Ukraine, or the case of Navy Captain Brett Crozier, who leaked a letter to the press about how his pleas for help in battling the coranavirus aboard the ship he commanded, the USS Theodore Roosevelt.
-Certainly these instances involved individuals making choices about to whom or to what they would be most loyal to - the President, the Constitution, the ideals and obligations of the military, themselves, and perhaps more.]
+Moreover, at the level of the military institution, a similar problem may also emerge: will the military institution have to choose between competing objects of loyalty, such as the President as the Commander in Chief, the Congress, and even its own professional ethic?^[Loyalty had to be a prime consideration in several prominent, recent cases involving the military and the government.
+Consider the case of now retired Army Lieutenant Colonel Alex Vindman, who testified before Congress against the US President, Donald Trump, regarding what Vindman believed to be behavior by the President that may have been illegal, or the case of Navy Captain Brett Crozier, who leaked a letter to the press after his pleas for help in battling the coranavirus aboard the ship he commanded, the USS Theodore Roosevelt, went unheeded for what he felt was too long.
+Certainly these instances involved individuals making choices about to whom or to what they would be most loyal to - the President, the Constitution, the ideals and obligations of the military, the operational mission, themselves as individual moral agents, and perhaps even more objects of loyalty.]
 
@@
 
-My intended topic is different than these other, still very important topics in which disagreements between the executive and the military occur during wartime.
+My intended research topic is different than these other, still very important topics in which disagreements between the executive and the military occur during wartime.
 It is my assessment that polarization is causing conflict, or certainly has the potential to do so, between the military and civilian leaders, but in most instances, the subject is about topics other than the use of military force, which is somewhat puzzling, given the military's reason d'etre.
-For example, I believe we are seeing portions of the military institution (particular retired military officers) openly disagree with civilian leaders on topics that I believe are, even by the widest of academic definitions, outside the bounds of what constitutes appropriate military expertise.^[I hope to prove this, empirically, in or as part of a chapter in my dissertation] While I believe that such developments are catching the attention of civil-military relations scholars, none to my knowledge have specifically set out to systematically explore if and how polarized partisan politics are impacting the military's professional identity and character, and how.
+For example, I believe we are seeing portions of the military institution (particular retired military officers) openly disagree with civilian leaders on topics that I believe are, even by the widest of academic definitions, outside the bounds of what constitutes appropriate military expertise.^[I hope to prove this, empirically, in or as part of a chapter in my dissertation]
+
+This raises, then, an important question: does the military's essential purpose risk evolving or changing significantly during polarized times?
+While I believe that these potential developments are catching the attention of civil-military relations scholars, none to my knowledge have specifically set out to systematically explore if and how polarized partisan politics are impacting the military's professional identity and character, and in particular, the role of the military institution itself in facilitating these developments.
+Most scholars acknowledge the potentially damaging impacts of the politicization of the military [@golby_listening_2013, @brooks_paradoxes_2020] at the hands of civilian leaders, and I share these concerns.
+Fewer scholars, if any, are exploring the potential role of the military in politicizing itself, and the mechanisms through which these developments occur.
 
@@
 
-## How Polarization Impacts Loyalty as a Concept
+## Political Polarization Impacts Conceptions of Loyalty
 
 In this section of the prospectus, I will lay out an initial argument, as well as accompanying hypotheses, which I intend to explore and test in my dissertation.
-Figure 2, below, shows an initial conception of the linkages between several of the variables discussed in the previous section.
-From left to right, this conceptual model suggests that worsening political polarization is driving changes in both the institutional credibility of the military (how society views the military) and changes in the conceptions of loyalty and identity (within the military).
-These two independent variables are, I suggest, driving subsequently driving changes in the concept of military professionalism, the main dependent variable of analysis in my dissertation.
-Following Huntington's model, I also suggest that changes in military professionalism can impact a variety of outcomes, by which I mean the very same outcomes that other scholars do and have already explored, such as effectiveness on the battlefield [@reiter_democracies_2002; @biddle_democracy_2004] and achieving strategic effectivness [@brooks_creating_2007; @brooks_shaping_2008].
-While I intend to examine the broad system of variables shown in Figure 2, the focus of my dissertation, as well as the least-explored linkage in this model, is the relationship between political polarization and changes in the concept of loyalty and identity, and how such changes manifest themselves in the variation of military professionalism.
+Figure 2, below, shows an initial conception of the linkages between several of the variables discussed in the previous section of the prospectus.
+From left to right, this conceptual model suggests that worsening political polarization is driving changes in both the institutional credibility of the military (how society views the military) and changes in the way members of the military institution think about and exercise loyalty, and by extension, how they think of their professional identity.
+These two independent variables are, I suggest, driving subsequent changes in the concept of military professionalism, the main dependent variable in my analysis.
+
+In line with Huntington's model as well as a host of other scholarship, I suggest that changes in military professionalism can impact a variety of military outcomes, varying from effectiveness on the battlefield [@reiter_democracies_2002; @biddle_democracy_2004] to achieving strategic effectivness [@brooks_creating_2007; @brooks_shaping_2008].
+While I intend to examine the broad system of variables shown in Figure 2, the focus of my dissertation is the relationship between political polarization and changes in the concept of loyalty and identity, and how such changes manifest themselves in the variation of military professionalism.
 For emphasis and clarity, I do not think that the other link - that between political polarization and institutional credibility - is unimportant.
-Rather, this link seems to be the focus of more-recent scholarship [@robinson_michael_danger_2018].
+Rather, this link seems to have been the focus of several recent scholarship efforts [@robinson_michael_danger_2018; ].
 
@@
 
-A primary contention from the system visualized above is that as political polarization in the domestic sphere increases, the military faces increased challenges in how both individual military leaders as well as the institution conceptualizes, adopts, and exercises loyalty, and by extension, professional identity.
+A primary contention from the system visualized above is that as political polarization in the domestic sphere increases, the military faces increased challenges in how both individual military leaders as well as the institution conceptualizes, adopts, and exercises loyalty, and by extension, how it sees its own professional identity.
 This expectation stems from several features that scholars of American politics have raised regarding the current politically polarized climate within the United States.
@@
 This happens because as "conflict extension" [@layman_party_2002] occurs, an increasing number of issues are brought into partisan battlegrounds, and thus, there is simply less politically-neutral territory available.
-Thus, an actor who wishes to remain non-partisan, such as the military, may be increasingly less able to do so, as a result of this feature.
+Thus, an actor such as the military, who may wish to remain non-partisan and out of politics to the extent that it can, may simply be less able to do so.
 
-A second challenge military leaders and the institution faces in a polarized climate is from the "negative partisanship"[@abramowitz_great_2018] that dominates the polarized American political landscape.
-As citizens hold increasingly strong and negative feelings towards those who side differently on a host of issues, there may be a tendency for members of the military institution to harbor similar feelings.
-Even if impact of this feature on the military and its leaders, due to training and professional development that emphasizes the importance of upholding apolitical and non-partisan behavioral norms, is somewhat muted, the point is that members of the military institution are likely experiencing the same phenomenon - even if it is to a lesser degree - as the country's citizens.
+A second challenge military leaders and the institution faces in a polarized climate is from the rise of "negative partisanship"[@abramowitz_great_2018] that dominates the polarized American political landscape.
+As citizens hold increasingly strong and negative feelings towards those who side differently on a host of issues, there may be a tendency for members of the military institution to harbor similar feelings and attitudes.
+Even if the impact of this feature on the military and its leaders, due to training and professional development that emphasizes the importance of upholding apolitical and non-partisan behavioral norms is somewhat muted, the point is that members of the military institution are likely experiencing the same phenomenon - even if it is to a lesser degree - as the country's citizens.
+One important question about "negative partisanship" in a military setting, though, is whether service members hold strong negative feelings toward only civilians who are on the opposite side of the ideological spectrum, or whether such feelings also permeate within the ranks.
 
-A third, and I contend the most impactful, challenge that normative military professionalism faces stems from the belief by individuals that "core" issues are at stake in the current polarized political climate.
-"Core" issues are difficult to define precisely [need cite], but the main point is that these issues often involve perceptions of moral right and wrong, and not just simple policy preferences.
-Moreover, some of the most polarizing issues are often associated with issues related to an individual's identity, such as race, gender, immigration [need cite].
-The rise of "identity politics" in America as political polarization has worsened illustrates this point [need cite].
+A third, and I contend the most impactful, challenge that normative military professionalism faces stems from the belief by individuals that core issues - such as those that involve attitudes about race, religion, and multiple layers of identity - are at stake in the current polarized political climate.
+"Core" issues are difficult to define precisely, but the main point is that these issues often involve perceptions of identity [] and moral right and wrong, and not just simple policy preferences.
+Moreover, some of the most polarizing issues are often associated with issues related to an individual's identity, such as race, gender, immigration, as the the rise of "identity politics" in America illustrates this point [@sides_identity_2018].
 
-I argue that this third challenge is especially relevant for the concept of loyalty.
-As an increasing number of what individuals believe to be "core" issues are raised and debated in the partisan arena, individuals face decisions of how to prioritize the persons or objects of their loyalty and further, what actions to take on the basis of loyalty to these persons or objects.
+I argue that this third challenge is especially relevant for the contemporary military, and specifically, with respect to how members of the military exercise the concept of loyalty.
+As an increasing number of what individuals believe to be core issues are raised and debated in the partisan arena, individuals face decisions of how to prioritize the persons or objects of their loyalty and further, what actions to take on the basis of loyalty to these persons or objects.
 I further theorize (though this has to be tested, of course) that while all individuals will likely face some version of this challenge of loyalty in a polarizing age, the challenges will be more acute for military leaders.
@@
 
-Moreover, as important as these consequences are, they do not speak to the potentially high personal and individualistic consequences military leaders may face if confronted with a situation in which they have to execute orders which are technically legal, but morally borderline at best.
-Military leaders are still humans who have to live with themselves at the end of the day.
-Therefore, I argue that if polarization continues to erode boundaries of what is considered moral by individuals to begin with, there is an increased likelihood of military leaders having to choose between conflicting loyalties.
+As important as these consequences are, they do not speak to another set of consequences individual military leaders also may face in a polarizing age.
+Indeed, in addition to consequences to the country that may occur should the military fail to discharge its duties to safeguard America, military leaders also potential consequences if confronted with a situation in which they have to execute orders which are technically legal, but morally borderline at best.
+Military leaders are still "moral agents" who "remain responsible for the decisions and actions they take" [@dubik_just_2016, 160].
+But the considerations of what constitutes moral right and wrong to begin with may be changing as polarization worsens [@graham_liberals_2009].
+
+Therefore, I argue that if polarization continues to heighten the identities of groups that people associate with, as well as come closer to the boundaries of how individuals think about moral right and wrong, there is an increased likelihood of military leaders having to choose between conflicting loyalties.
 What happens if and when the objects of a military leader's multiple loyalties not merely disagree, but become stubbornly divided ideologically?
-Can such an outcome exist, and if it can, is there empirical proof that such a trend is occurring?
+Can such an outcome exist?
+Is there empirical proof that such a trend is happening?
 If so, what are the normative implications, both for the military as a profession, and for the nation's security, and are there actions that can be taken to halt such a trend?
@@
 The second way in which polarization and credibility may interact arises more directly from what can be termed a credibility effect, and has more to do with the place of prominence (or lack thereof) that the military or individual leaders hold in the perception of American society.
-[need to blend this in with recent scholarship and where gaps are - bring in Urben, Golby, Dropp and Feaver, etc., and Robinson.
-Also need to figure out where to put my hypotheses]
+As recent scholarship has shown, the credibility of the military institution in a variety of contexts depends heavily on partisan affiliation [@urben_party_2013; @golby_listening_2013; @golby_jim_elite_2017; @robinson_danger_2018].
 
@@
 Under these conditions, the military faces dual dangers.
-High institutional credibility risks the military institution unilaterally acting or another body co-opting the military institution to speak out an a number of issue areas, and simultaneously, the highly polarized environment risks the military advocating for or denouncing partisan sentiments in the name of upholding the very traits it is most adhered to - loyalty and duty.
-Indeed, this set of conditions comprise a very dangerous world for the military institution and the military professional.
+High institutional credibility risks the military institution unilaterally acting or another body co-opting the military institution to speak out an a number of issue areas, and simultaneously, the highly polarized environment risks the military advocating for or denouncing partisan sentiments in the name of upholding the very virtue it is most normatively adhered to - loyalty.
 
@@
 In order to test my argument, I will have to examine and test several hypotheses, which are inter-related and linked. They include the following:
+
+\singlespace
 
@@
 
-_H4: Following from the above hypothesis, polarized periods will feature events in which governmental actors appeal to the military's sense of loyalty in order to persuade them to adopt a stance on a particular issue area._
+_H4: Following from the above hypothesis, polarized periods will feature events in which governmental actors appeal to the military's sense of loyalty in order to persuade them to adopt a stance on a particular issue area.
+These can be viewed as "exploits of loyalty."_
 
-_H5: In highly polarized time periods, it is the senior military officers as the "stewards of the profession" [@us_department_of_the_army_adp_2019, p.1-9] whose actions will most influence to what degree and how the military institution changes.
-If changes occur to the military institution, they are bound to occur within two primary spheres: the military's professional identity, and the military's professional coherence._
+_H5: In highly polarized time periods, it is the senior military officers as the "stewards of the profession" [@us_department_of_the_army_adp_2019, p.1-9] whose actions will most influence to what degree and how the military institution reacts to and changes as a result of "exploits of loyalty." If changes occur to the military institution, they are bound to occur to the professional identity of the military, which includes the degree to which it functions as a profession (vice bureaucracy [@snider_once_2012]) and the degree to which it is obedient to its civilian leaders (vice acting as an activist on behalf of a particular cause)._
 
+\doublespace
 # Plan of the Dissertation and the Data Generating Process
@@
 The one sentence thesis of my dissertation is that political polarization significantly impacts how the military thinks about and exercises loyalty, a chief military trait, and that these impacts can and often do result in changes to the military's professional identity and character.
-I plan to execute a book style project comprised of at least seven chapters.
+I plan to execute a book style project comprised of seven chapters.
 
@@
 I will emphasize the long-running nature of this debate going all the way back to when Huntington first published his book, but point to some of the prominent contemporary critiques regarding his theory [@feaver_civil-military_1996-1; @rapp_civil-military_2015; @golby_why_2017; @karlin_case_2020; @brooks_paradoxes_2020].
-The main point of chapter 1 is to set the stage for my dissertation and signal to the reader that modern critiques of Huntington may be overlooking a key part of his concept of military professionalism: it may insulate the military from the most pernicious effects of polarizing politics.
+The main point of chapter 1 is to set the stage for my dissertation and signal to the reader that modern critiques of Huntington may be overlooking a key part of his concept of military professionalism: imperfect as it may be, it may also insulate the military from the most pernicious effects of polarizing politics.
 
-In the second chapter, I will introduce, develop, and explain my theory that links increasing political polarization to conceptions of loyalty and military professionalism (see Figure 2).
+In the second chapter, I will introduce, develop, and explain my theory that links increasing political polarization to conceptions of loyalty and military professionalism (see Figure 2, above).
 I will argue that specific features of the polarized environment strain and challenge both individual and institutional conceptions of loyalty, and by extension, normative aspects of military professionalism (Table 1).
-I will further argue that rising political polarization can interact with changes in the military's institutional credibility to increase the likelihood that the public becomes aware of changes in the military's professional conduct.
+I will further argue that rising political polarization can interact with changes in the military's institutional credibility to increase the likelihood that the public becomes aware of how the military chooses to respond to the dynamics brought on as the military faces challenges to how it conceptualizes and practices loyalty.
 Using Table 2 as a framework, I will place epochs of historical American civil-military relations into one of four distinct quadrants, and signal that these quadrants will comprise the qualitative case studies that I later examine in my dissertation.
@@
 I plan to demonstrate this empirically by drawing upon an original dataset consisting of more than 340 op-ed pieces authored by retired military officers since the late 1970s and published in the top five major print newspapers in the United States, including the Wall Street Journal, the New York Times, the Washington Post, the USA Today, and the Los Angeles Times.
-I believe that I can show several things from this analysis, including that there has been a sharp uptick in the number of op-eds published by retired military officers during this time period, and more importantly, that there has been a decline in the normative professional ideals espoused and argued for in these op-eds.
-Critical to this chapter is acknowledgement that coding for what is professional versus what is not is subjective to a large extent, and thus, the need for me to truly develop clear working definitions of the concept of military professionalism in Chapter 1.
-However, I will show that whether one adopts a strict Huntingtonian version of military professionalism, or a less strict version that allows for more encroachment by the military institution into the political sphere, by both standards the retired military officer corps is not living up to the ideals of normative military professionalism.
-The question, then, becomes why aren't they?
-I will argue that while there are certainly other confounding variables, to include changes in the conservative-liberal tilt of newspapers over time, the chief explanation for the drop in professionalism stems from the combined effects of increasing political polarization and rising institutional credibility of the military since the late 1970s, and the challenges that these changes introduce to a military leader's conception of loyalty.
+I believe that I can show several things from this analysis, including that there has been a sharp uptick in the number of op-eds published by retired military officers during this time period, and more relevant to the theory I propose in the dissertation, that these published op-eds reveal a decline in the normative professional ideals of the military institution.
+Unlike previous recent works which focus on how polarization impacts society's view of the military, this chapter introduces an empirical record of actors within the military institution (retired senior officers) also changing as a result of increasing political polarization.
+
+Critical to this chapter is acknowledgement that coding for what is professional versus what is not is a somewhat subjective process.
+Thus, I will need to develop and implement a clear working definition of the concept of military professionalism in Chapter 1.
+However, I believe that I can show in Chapter 3 that whether one adopts a strictly Huntingtonian version of military professionalism (strict separation of political and military spheres) or a less strict version that allows for an increasing shared space, the retired military officer corps is not living up to the non-partisan ethic that the military espouses.
+
+I will argue that while there are certainly other confounding variables, to include changes in the conservative-liberal tilt of newspapers over time, the chief explanation for the variation (decline) in professionalism stems from the combined effects of increasing political polarization and rising institutional credibility of the military since the late 1970s, and the challenges that these changes introduce to a military leader's conception of loyalty.
 Moreover, my ultimate point is to point out a blatant inconsistency in the way that the military currently thinks about the political behavior by those within the ranks in that the institution has rules that governs the conduct of those on active duty [@department_of_defense_political_2008], but those same rules, and the spirit behind those rules, are disregarded by the institution's vangaurds, its retired multistar officers and "stewards of the profession" [@us_department_of_the_army_adp_2019, p.
 1-9] of arms.
+I believe that this chapter will result in finding support for _H1._
 
-In the fourth chapter, I want to show that the institutional challenges to loyalty have been present as political polarization has worsened.
-To do this, I plan to examine, via textual analysis, the records of Congressional Committee hearings in which military officers, active and retired, have played a part, and detect if and how patterns of questioning and providing answers have changed over time.
-I am especially interested in the way that individuals have invoked language associated with the concept of loyalty in asking and answering questions.
-An example of this might be analyzing the ways in which terms such as "moral obligation" or similar language is used in these hearings.
-I hope to detect patterns that will help support my overall argument.
-I have yet to flesh this particular chapter out some more.
-
-In Chapter 5, I plan to analyze the results of an original survey experiment in which I survey a nationally representative sample of civilians and compare the results to a sample of either a) veterans, or b) active duty military officers.
+In Chapter four, I plan to analyze the results of an original survey experiment in which I survey a nationally representative sample of civilians and compare the results to a sample of either a) veterans, or b) active duty military officers.
 In this survey experiment, I plan to test whether members of the military (which I proxy by veterans) genuinely are more sensitive to appeals to loyalty, as I have hypothesized in _H3._ The survey, after controlling for gender, partisan identification, and age, will present the respondent with 6-8 hypothetical situations and a corresponding argument made by a military elite.
@@
 \doublespace
+
 Such a survey will enable me to measure two things.
@@
 
+In the fifth chapter, I want to show that the military has also faced institutional challenges to the concept of loyalty as political polarization has worsened.
+To do this, I plan to examine, via textual analysis, the records of Congressional Committee hearings in which military officers, active and retired, have participated.
+My goal is to detect if and how patterns of questioning and providing answers have changed, particularly in how language related to the concept of loyalty has been used.
+I still have some work to do in fleshing this out some more, but my initial idea is to examine how the military is both asked and answers questions by members of Congress, and how these questions and answers invoke language that speaks to various loyalty obligations.
+An example of this might be analyzing the ways in which terms such as "moral obligation" and "sacred duty" have been used over time.
+I might be able to associate causes or subjects that are associated with this type of language, and correlate these causes and subjects to the level of polarization present in the country, as well as to the partisan identification of the specific member of Congress asking the question.
+My goal is to detect patterns that will help support my overall argument, and specifically, _H2_.
+
 In the sixth chapter, I will turn to a series of qualitative case studies to explore and test my theory.
-Key in this chapter is finding and examining historical periods in American civil-military relations in which there existed varying levels of political polarization and institutional credibility, and then, to walk the reader through how challenges of loyalty during these periods manifested themselves in different ways among specific members of the military and the institional military at large, in line with the theory that I propose in Chapter 2.
+Key in this chapter is finding and examining historical periods in American civil-military relations in which there existed varying levels of political polarization and institutional credibility, and then, to walk the reader through how challenges of loyalty during these periods manifested themselves in different ways among specific members of the military and the institional military at large, in line with the theory that I propose in Chapter 2 and specifically, _H2_.I then hope to find evidence for how the actions of the military and its leaders during these periods impacted the military's professional identity and character, in line with _H4_ and _H5_.
+
+I have not yet settled on what the specific case studies be, particularly because I am wresting with whether I need to confine my examination to the time period in which the US military has generally been considered a profession (depending on who you ask, this is generally around World War One, or perhaps a little bit earlier, at the tail end of the 19th Century).
+My initial reaction is that I do not need to confine myself to this period.
 My initial projection of what these case studies will be and how they will align with varying levels of political polarization and instititional credibility are as follows:
@@
 
-* High Polarization, Low institutional credibility: civil-military relations in the post-Vietnam period, in the late 1970s.
+* High Polarization, Low institutional credibility: civil-military relations after World War One, during the 1920s.
 
@@
 
-* Low Polarization, Low Institutional Credibility: TBD.
+* Low Polarization, Low Institutional Credibility: civil-military relations in the post-Vietnam years, roughly 1972-1980.
 
-In the seventh and final chapter, I will conclude by exploring the implications of my argument and by
+In the seventh and final chapter, I will conclude by exploring the implications of my argument.
+Note: one idea I am considering is doing an extra chapter or at least an extended portion of Chapter 6 on how the Black Lives Matter movement is impacting the military profession.
+I think there could be some interesting work here.
+
+# Initital Concerns and Timeline
+
+My main concern at this point is that I harness the relevant outside literature, from the fields of psychology, ethics, sociology, and management, and still focus my puzzle on an empirical, political science question.
+To that end, I am concerned that I have a topic that is not sufficiently narrow, and thus, I may need to hone in on a specific subset of questions that I have presented here.
+
+A second concern is that I do my best to make this dissertation relevant and interesting for international relations, and not only for the military itself (though that is certainly a focus of my intended work).
+To that end, I may need to adopt a more comparative approach that examines if and how other militaries have been impacted by political polarization in their countries, and thus attempt some sort of cross-national study, should one of my intended chapters be deemed insufficient.
+I note this especially because the concept of loyalty, as mentioned in this prospectus, features so prominentaly in international relations in the context of coups, ethnic exclusion, etc., and I want to make sure I am not missing an opportunity to bring these matters to bear on my research project.
+
+In terms of timeline, and this is still very fluid, I intend to reach the following milestones:
+
+* Between now and 15 December 2021: Continue to work on my prospectus.
+Work with Jonathan Renshon and Jon Pevehouse to develop and field my intended survey experiment (I have already obtained IRB approval through UW-Madison, but I need to apply for funding (through the Army), finalize the experiment, and prepare for data collection.) Plan to take my methods (second field) comprehensive exam.
+
+* No later than 15 January 2021: defend my dissertation prospectus.
+
+* No later than 15 May 2021: fully write Chapters 1-3 of my dissertation and submit for review to the dissertation committee.
+Prepare to field the original survey experiment for Chapter 4.
+Begin data collection for Chapter 5 of the dissertation, and settle on which case studies to write for Chapter 6 of the dissertation.
+
+* Summer 2021: continue working on Chapters 4 and 5 of the dissertation.
+
+* Fall 2021: submit drafts of Chapters 4 and 5 for review to the committee, and begin writing Chapters 6 and 7.
+
+* January 2022: submit initial drafts of Chapters 6 and 7 to the committee for review.
+
+*February - May 2022: improve all chapters of the dissertation.
+
+* No later than May 2022: defend dissertation.
+
+# Conclusion
+
+The purpose of my dissertation is to explore how polarizing domestic politics affects normative concepts of military professionalism, particularly in the context of the United States and among the leaders of the US military.
+Certain features of increasing political polarization - in particular, the way in which individuals associate their partisan beliefs with salient features of identity [@sides_identity_2018] and right and wrong [@graham_liberals_2009], I argue, introduce immense challenges to the way military leaders conceive of and exercise loyalty, which has long been a foundational trait of military professions [@us_department_of_the_army_adp_2019].
+In particular, military officers in polarized time periods face challenges of conflicting loyalties [@coleman_problems_2009] that are more acute than during non-polarized periods, resulting in situations where military officers must prioritize the multiple objects of their loyalty (to the President as Commander in Chief, to Congress, to the operational mission, to subordinates and superiors along the chain-of-command, and to one's self).
+
+These decisions, though made by individual military leaders, often result in changes to the military as a profession and as an institution, with ramifications that bear on civil-military relations, the military's ability to defend the nation, and other important areas.
+By focusing my dissertation on this topic, I will contribute to scholarship by examining the specific ways in which polarization impacts the military institution, thus helping scholars of civil-military relations and international relations better understand the mechanisms through which polarization changes American institutions, particularly among the one institution charged with maintaining national security.
 

### 2020-10-23 code/prospectusv2.Rmd
commit

--- 2020-10-22 code/prospectusv1.Rmd
@@
 ---
-title: "Loyalty's Limits: The Impact of Postmodernism and Political Polarization on Military Professionalism  (Prospectus)"
+title: "Saviors or Syndicates?: The Rise of Domestic Political Interventionism by Retired Military Leaders (Prospectus)"
 author: "Peter M. Erickson"
@@
 
-  _This dissertation seeks to explore the impact of increasing political polarization on the professional identity of military officers and the military institution.
-I develop a theory that argues that the nature of rising political polarization fundamentally challenges the conception of loyalty among members of the military profession, and that such a strain is problematic for an institution such as the military that holds the virtue of loyalty in high regard, and by statute possesses multiple, simultaneous loyalties: to the Constitution of the United States, to the President qua Commander-in-Chief, to Congress, to the American People, to one's fellow service members, and to one's self.
-Balancing these competing loyalties becomes more difficult as political polarization worsens, I argue, because of the fundamental and core values - not mere policy differences - that are at the root of many contemporary polarized political battles.
-Employing a mixed-methods approach that incorporates quantitative analysis and several historical case studies, this dissertation argues that the challenges to conceptions of loyalty among members of the military institution brought about by political polarization have the potential to impact normative concepts of military professionalism, which could be problematic not only for the character and identity of the military as an institution, but for American national security at large._
+  _This dissertation explores the changes in how a powerful block of military professionals - retired senior military leaders - intervene in domestic contemporary politics.
+I advance a theory which argues that changes in political activism by retired senior leaders is a result of two interacting factors: the level of domestic political polarization within a society, and the degree of prestige and credibility that military leaders enjoy within the society.
+To support my argument, I trace changes in levels and types of distinct forms of political activism, including authoring opinion-editorials in major newspapers, and participation in presidential endorsements.
+I further advance the theory by conducting a survey experiment that tests for differences between civilian and military responses after exposure to situations that evoke one's partisan or ideological identity.
+I ultimately argue that the increasing political activism of retired military leaders is not merely a response to the hyper-partisanship present in the current political atmosphere, but rather the logical outworking of trends that have been active for some time in which both the ideological differences between Americans and the prestige of the military have increased.
+I conclude by describing the effect of the implications of growing politicla activism by senior military leaders on the core beliefs of the military regarding its very own essence and role in society._
+
 geometry: margin = 1.15in              # margin widths
@@
 
-## Introduction and Research Question
+# Introduction and Research Question
 
-In late May and early June 2020, racial tension and concerns about police brutality gripped the American public following the death of George Floyd, an unarmed black man, by police in Minneapolis.
-In many cities and communities across the nation, peaceful protests were followed in some cases by periods of rioting and looting.
-On June 1, after giving remarks at the Rose Garden, President Donald Trump walked to St.
-John's Church after security forces visibly displaced throngs of protesters in neighboring Lafayette Square.
-Flanked by members of his cabinet, to include the Secretary of Defense, and with the Chairman of the Joint Chiefs of Staff not far behind, the President held up a Bible at the church, stared into the cameras of the media, and, several minutes later, returned to the White House.
+Across the world and in almost nearly every country, discussions regarding the participation of the military in politics are nothing new.
+Indeed, few topics evoke such strong opinions and emotions than whether and to what degree the military, both as a corporate body and as individual military officers, should be political vice remain separate from certain aspects of and activities associated with politics.
+These normative questions rightly form a significant portion of the study of civil-military relations, and intellectuals, practitioners, and pundits have frequently weighed in on these topics over the years.
 
-In the days and weeks that followed, a full-blown civil-military relations crisis ensued  [@feaver_military_2020].
-The Secretary of Defense, Mark Esper, publicly disagreed with a threat made by President Trump to invoke the Insurrection Act of 1807 to use active-duty forces to quell domestic riots [@esper_secretary_2020], and General Mark Milley, the Chairman of the Joint Chiefs of Staff, told senior military officers graduating from the National Defense University that he "should not have been there (accompanying the President on the walk to St.
-John's Church)", as his presence "created a perception of the military involved in domestic politics" [@milley_official_2020].
+Thankfully, within the United States, unlike many other countries across the world and throughout history, there has never been a coup, nor a serious attempt made by the military to dislodge political control from their duly-constituted elected leaders and to hold it themselves.
+However, as almost every civil-military relations scholar points out, the absence of coups does not mean that the relationship between civilian and military leaders is healthy, nor absent of conflict altogether.
+Indeed, a cursory examination of the relationships between Lincoln and MacClellan, Truman and MacArthur, and Obama and McChrystal reveals that disagreement between those who serve at the highest levels of civilian government and military organizations is neither rare nor understudied.
 
-Several retired military officers also weighed in, not only on the Trump administration's handling of specific events near the White House, but on a host of broader topics that dominated the popular dialogue in America after the death of George Floyd, to include the state of American racial relations.
-Upon paying close attention to what these retired military officers said, though, one can notice that there was in fact fairly substantial variation in both the substance and tone of their remarks.
-While all of them expressed concerns about the President potentially ordering active-duty military forces to confront what many Americans believed were peaceful protesters, and thus, violating Constitutional protections afforded by First Amendment [@brooks_dismay_2020; @dempsey_former_2020], some officers went much further in their criticisms.
-Indeed, some of these retired military officers accused the President of failing to possess the overall fitness to lead [@goldberg_james_2020], promote harmonious racial policies  [@allen_moment_2020], and demonstrate adequate leadership abilities [@mullen_i_2020].
-In the months following George Floyd's death, moreover, several retired and currently-serving military officers have further made public arguments on a variety of topics that are directly related to President Donald Trump, including race in America [@askew_anti-racist_2020] and voter fraud as it relates to the upcoming Presidential Election [@nagl__2020].
+What has changed, however, is the degree to which a very specific subset of the American military institution - retired senior military officials - have participated in the domestic political process post retirement.
+Indeed, whereas General Omar Bradley, one of the military's five star generals and a hero of World War Two, famously remarked in 1959, "I am convinced that the best service a retired general can perform is to turn in his tongue along with his suit and mothball his opinions," [@times_bradley_1959], retired generals and admirals today appear frequently in the media, endorse Presidential candidates at higher frequency [@griffiths_not_2019], and frequently serve in government after their retirement from military service [@brooks_dont_2016].
+To provide one such example, following protests across the United States in the summer of 2020 in which President Donald Trump nearly ordered the use of active duty military forces in American cities, former Chairman of the Joint Chiefs of Staff Admiral (Ret) Michael Mullen was one of at least ten four star retired officers who publicly aired concern about the President's potential use of force.
+^[Other officers included General (Ret) James Mattis, General (Ret) Mark Kelly, Admiral (Ret) William McRaven, General (Ret) Martin Dempsey, General (Ret) Colin Powell, General (Ret) Michael Hayden, General (Ret) David Petraeus, General (Ret) John Allen, and General (Ret) Vincent Brooks.].
+During an interview in the days which followed, when asked by NPR reporter why he spoke out, Mullen stated, "we were getting to a point where the country was in jeopardy, which is why I spoke out" [@isaacson_ret_2020; for further analysis, see also @brooks_let_2020].
 
-While America has witnessed numerous incidents in its history in which tension between civilian and military elites has been present, the events of the Summer of 2020 between the Trump administration and some - though certainly not all - retired  military officers raises important questions, particularly I argue, because the context did not involve combat overseas.
-These questions include: why did several retired military elites speak out so strongly against President Trump?
-Why did some of these elites confine their criticism to concerns over the Constitution, while others sounded more like representatives of the Democratic Party?
-Are these events that occurred during the Summer of 2020, and the military response generated in response to these events, an isolated chapter in American history, or is there something more disturbing underneath the surface regarding contemporary American civil-military relations?
+Mullen's words are incredible from a number of important perspectives, and raise important normative questions regarding the military's role in a democratic society [@brooks_let_2020] and the conditions under which the military can and should express dissent to its civilian bosses [@cohn_what_2019].
+Few, if any, scholars have attempted to systematically examine and explain the rise of political activism by retired military officers over time.
+Such an effort would better illumine the understanding of scholars, as well as the American military and society, of why a powerful group of what were once non-political actors have, at least in some cases, suddenly become one.
+More important, a systematic study of this trend would contribute to a much-needed discussion regarding the role of the military institution and its relationship to the government and American society in a contentious political age.
 
-This dissertation seeks to explore the impacts of rising domestic political polarization on the US military as an institution and profession.
-The primary argument espoused in this dissertation is that certain features of the polarized environment impact how military officers view and exercise the concept of loyalty, a claim that I argue holds significant ramifications for US national security.
-Using a mixed-methods approach, I develop a theory and then test it using a mixture of quantitative analysis and several qualitative case studies.
-I conclude with discussing the implications for such an argument.
+The central argument made in this dissertation is that increasing political activism at the hands of retired military officers is the result of two longstanding and interconnected trends: rising political polarization within the domestic political sphere, and rising prestige and credibility of military institutions afforded by American society.
+The interaction of these trends explains the remarkable increase in political activity by retired senior military leaders, and portends a shift in the perceived role that these senior military leaders fill vis-a-vis civilian authority - a shift away from professional and obedient servants and towards "guardians of democracy" who pick and choose when and how to intervene in American politics.
+Such a shift builds on previous work regarding normative "role beliefs" [@fitch_armed_1998] of the military institution, and has significant ramifications on the organizational culture of the military, its relationship with government and society, and American national security.
 
-# Framing Within the Scholarly Literature
+# The Field of Civil-Military Relations
 
-## Political Polarization
-
-The question of if and how political polarization drives changes in the professional identity of the military spans several related literatures, most notable of which are American Politics and International Relations.
-Broadly speaking, this project is focused on how one aspect of domestic politics, political polarization, impacts a specific but important aspect of international relations, civil-military relations, which itself spans a wide and diverse literature.
-Because I intend to focus my work on the military as an institution, my dissertation will also rely on literature from the fields of sociology, organizational psychology, and ethics.
-I will frame my research question within each of these fields briefly, and to define important concepts as required, before addressing why this question matters and should be studied in depth.
-
-Within the American Politics literature, there is broad recognition that political polarization levels in contemporary America are higher now than at any point since the end of the American Civil War [@hare_polarization_2014], but there are multiple explanations for what has caused this trend, ranging from a growing divide over matters pertaining to race and religion [@abramowitz_great_2018], the size and role of government [@sides_two_2015], and disagreements over specific policy issues [@mann_its_2012].
-Despite the wide-ranging scholarship exploring the many facets of political polarization, virtually all scholars agree that contemporary levels of political polarization in the United States are historically abnormal and pose myriad problems.
-
-One of the major concerns regarding political polarization is its on impact on American foreign policy in a broad sense, and specifically, on America's ability to lead and impact the world.
-Kenneth Schultz, for instance, underscored four challenges that have emerged as a result of America's polarized domestic politics, including the ability to obtain bipartisan support for "risky undertakings," the ability to learn the "lessons of foreign policy failure," the inability to "make long term commitments to allies and adversaries" as a result of "dramatic policy swings from one administration to another," and the "vulnerability" of America's electoral system to foreign third party adversaries [@schultz_perils_2017].
-Because the military is a central actor in shaping America's foreign policy, obtaining a better understanding of how political polarization shapes various military institutions will yield findings that may help policymakers and academics help solve the problems Schultz identifies.
-
-## Civil Military Relations
-
-Civil-military relations comprises an immense area of scholarship and interest, and deals with a correspondingly large number of questions and debates that have occupied the minds of political scientists, theorists, historians, ethicists, and sociologists for hundreds if not thousands of years.
+Civil-military relations involve a wide host of significant and important theoretical questions that involve the pull and push of power between institutions, the delineation of authority between military and civilian leaders, and a number of other consequential issues.
 At its root, the study of civil-military relations examines what Peter Feaver calls the "problematique," or how to ensure that the military is strong enough to defeat threats to the polity without threatening the polity itself [@feaver_civil-military_1996; @feaver_civil-military_1999].
-Feaver describes his "problematique" as "vexing" and a great "paradox" in that the military must be "strong enough to prevail in war," yet submissive to authority such that it cannot and will not "destroy or prey on the society it is intended to protect" [@feaver_civil-military_1999, p.
+Feaver describes this "problematique" as "vexing" and a great "paradox" in that the military must be "strong enough to prevail in war," yet submissive to authority such that it cannot and will not "destroy or prey on the society it is intended to protect" [@feaver_civil-military_1999, p.
 214].
-
-Civil-military relations involves a wide angle of significant and important theoretical questions that involve the pull and push of power between institutions, the delineation of authority, and a host of other massive and consequential issues impacting how militaries perform in war, their relationship to the societies they serve, and other important questions.
-In the context of the US, which has never experienced or come close to a coup (thankfully), the study of civil-military relations often focuses on the relationship between military and government elites, or what Risa Brooks calls "leadership at the state's apex" [@brooks_risa_integrating_2019, p.
+Within the context of the US, the study of civil-military relations often focuses on the relationship between military and government elites, or what Risa Brooks calls "leadership at the state's apex" [@brooks_risa_integrating_2019, p.
 380].
 Indeed, it is at this level that vital decisions are made, including to go to war, how to prosecute war, how to achieve and implement effective strategy, how to prepare for the next conflict, and how large the military and its budget should be.
-Each of these questions, and many, many more, involve questions regarding the proper role of the military and its leaders relative to both society and, perhaps more important, the leaders of government.
-A number of scholars have examined this question on normative grounds, but none has been more lasting and consequential than the late Samuel Huntington, whose 1957 seminal work, _The Soldier and the State: The Theory and Politics of Civil-Military Relations_, often serves as the point of departure for many scholarly works on civil-military relations.
+A number of scholars have offered solutions to Feaver's "problematique," but none has been more lasting and consequential than the late Samuel Huntington, whose 1957 seminal work, _The Soldier and the State: The Theory and Politics of Civil-Military Relations_, often serves as the point of departure for many scholarly works on civil-military relations.
 Although there have been many critiques of Huntington over the years [for example, see @janowitz_professional_1960; @finer_man_1962; @cohen_supreme_2003; @feaver_armed_2003; @brooks_paradoxes_2020], it is absolutely essential to understand Huntington's concept of normative civil-military relations, and what lies at the heart of it - a theory of civilian control of the military that maximizes the concept of what Huntington called "military professionalism" [@huntington_soldier_1957].
@@
 The second feature of the societal imperative, however, was the US Constitution, which Huntington claimed to be "conservative," and written by "men who feared concentrated political power and who provided for the widespread dispersion of that power among numerous governmental units" [@huntington_soldier_1957, 143].
-The tension raised by these two imperatives - one functional stemming from external threat, the other societal stemming from the character of American political ideology and structure - needed to be solved, claimed Huntington.
+Huntington claimed that the tension generated by these two imperatives - one functional stemming from external threats, the other societal stemming from the character of American political ideology and structure - needed to be solved.
 Writing in the context of the seemingly long and dangerous Cold War, Huntington's point is that America needs strong military that can win on the battlefield, but cautions that such a military cannot fundamentally alienate the character of American liberal society.
@@
 Huntington viewed the key independent variable that enables the achievement of these goals while also resolving the tension generated by the opposing functional and societal imperatives as the adoption of a type of civil-military relations framework which he calls, "objective civilian control" of the military [@huntington_soldier_1957, 83].
-In such a model of "objective civilian control," Huntington envisioned civilians granting the military a high degree of professional autonomy, which he believed would simultaneously be "most conducive to the emergence of professional attitudes and bahvior among members of the officer corps" as well as the best way to minimize the military's political power by rendering the military "politically sterile and neutral" [@huntington_soldier_1957, 83-84], and thus preserving civilian supremacy.
+In such a model of "objective civilian control," Huntington envisioned civilians granting the military a high degree of professional autonomy, which he believed would simultaneously be "most conducive to the emergence of professional attitudes and behavior among members of the officer corps" as well as the best way to minimize the military's political power by rendering the military "politically sterile and neutral" [@huntington_soldier_1957, 83-84], and thus preserving civilian supremacy.
 
@@
   geom_dag_text()+ 
-  geom_dag_label_repel(aes(label=label, fill=label), col="white", show.legend=FALSE)+ theme_dag() + labs(title="Figure 1: Samuel Huntington's Theory Of Civilan Control of the Military", subtitle="As Explained in The Soldier and The State, 1957")
+  geom_dag_label_repel(aes(label=label, fill=label), col="white", show.legend=FALSE)+ theme_dag() + labs(title="Figure 1: Samuel Huntington's Theory Of Civilian Control of the Military", subtitle="As Explained in The Soldier and The State, 1957")
 
@@
 What the framers did not see, nor could they, argues Huntington, was the rise of the military as a profession in response to ever-increasing and dangerous external threats.
-It was because of these pernicious threats, argues Huntington, which the US military had witnessed, confronted, and overcome during the 19th and 20th centuries in World War One, World War Two, and as Huntington writes, the Cold War, as well as the destructive capability of modern weapons and warfare in general that the military profession had emerged [@huntington_soldier_1957].
-Huntington claims that unlike during the nation's founding, military officers constitute a profession, and thus, are unique and distinct from other occupations in several important ways: professions practice and develop a special body of knowledge and expertise; professions exercise a  "social responsibility" by providing a vital service to society; and professions have a degree of "corporateness" in that they develop and formalize standards by which members of the profession are certified to practice [@huntington_soldier_1957, 11-18].
+It was because of these pernicious threats, argues Huntington, which the US military had witnessed, confronted, and overcome during the 19th and 20th centuries in World War One, World War Two, and as Huntington writes, the Cold War, as well as the destructive capability of modern weapons and warfare in general, that the military profession had emerged [@huntington_soldier_1957].
+Huntington claims that Cold-War military officers, unlike the corporate body of officers that existed at the nation's founding, constitute a profession and thus, are part of a unique and distinct organization that differs from other occupations: professions practice and develop a special body of knowledge and expertise; professions exercise a  "social responsibility" by providing a vital service to society; and professions have a degree of "corporateness" in that they develop and formalize standards by which members of the profession are certified to practice [@huntington_soldier_1957, 11-18].
 
@@
 Most of these critiques, in one way or another, charge that a strict theoretical and practical separation between military and political operating spheres, which Huntington to a certain degree certainly argued for, is fundamentally flawed.
-Many, if not all, of these critiques cite Clausewitz's famous dictum that war is a "mere continuation of policy by other means" [@clausewitz_war_1976, 69, 605], charging that by its very nature, war requires the highest integration, as opposed to segregation, of political and military activities.
+Many, if not all, of these critiques cite Clausewitz's famous dictum that war is a "mere continuation of policy by other means" [@clausewitz_war_1976, 69, 605], and thus point out that by its very nature, war requires the highest integration - not segregation - of political and military activities.
 Moreover, these critiques often point out, correctly in this author's opinion, that the conduct of war is more complicated than civilians simply telling the military the goals it wishes to achieve, and then allowing the military to determine how best to accomplish these goals.
@@
 Whether military leaders recognize it or not, Brooks argues, military leaders comprise a profession that is inherently political, and such a reality ought to be acknowledged and embraced [@brooks_paradoxes_2020].
-
-Regardless of whether and the degree to which one agrees or disagrees with Huntington's concept of military professionalism, a significant number of scholars note that a significant issue involves attempting to define where the line between appropriate political expression and inappropriate partistan activity should be drawn [@owens_military_2015; @golby_beyond_2015; @brooks_paradoxes_2020].
+Regardless of whether and the degree to which one agrees or disagrees with Huntington's concept of military professionalism, a significant number of scholars note that a significant issue involves attempting to define where the line between appropriate political expression and inappropriate partisan activity should be drawn [@owens_military_2015; @golby_beyond_2015; @brooks_paradoxes_2020].
 It is a difficult task for many reasons because, as Mackubin Thomas Owens points out, civil-military relations in the United States can be viewed as a "bargain" [@owens_us_2011] between the American people, the military, and the government that has to account for constantly-changing "political, social, technological, or geopolitical changes" [@owens_is_2016].
@@
 
-One of these recent changes, I argue, is increasing political polarization that has engulfed not only government elites, but the nation at large.
-Brooks (2020) points out three challenges posed by the contemporary period to the military's apolitical ethic, all of which are likely related to the growing problem of political polarization: rising acceptance among military officers of behaviors that contradict an apolitical ethos; increasing "pressures and opporunities" for military officers to express partisan sentiment; and increased politicization of the military by government elites [@brooks_paradoxes_2020, 21-23].
-Indeed, as political polarization increases, the lines between partisanship and politics, as well as politics and everything not politics, have blurred, as so much of the literature on American politics and polarization affirms, promising that this challenge with respect to the military institution and its members will likely persist for some time.
+## The Dependent Variable: Military Intervention in Politics and Military "Role Beliefs"
+
+Given the normative model espoused by Huntington and the critiques his model has been subjected to, it is important to understand what frameworks exist for understanding why the military would intervene in politics.
+Numerous scholars, mainly writing from a comparative perspective (although not exclusively), have enriched our understanding of what possibly motivates the military to interject itself into the political sphere.
+To be clear, these scholars are focusing mainly on active-duty military organizations, and not on the specific block of retired military officers, as I plan to examine, but the motivations and mechanisms for intervention that they put forward enrich our understanding of the phenomenon of political intervention by the military.
+
+The British political scientist, S.E.
+Finer, argues that the military's intervention in politics is a function of the opportunity and desire to do so [@finer_man_1962].
+Finer identifies several factors which should normatively inhibit the military from intervening in politics.
+These factors are military professionalism, the principle of civilian supremacy, and fear that the military would fail to remain a cohesive fighting force should it fracture internally upon intervening in political matters [@finer_man_1962].
+Finer critiques the concept of Huntingtonian military professionalism somewhat harshly, arguing that although military professionalism should act to focus the military on itself and thus not on political activities, those who claim to be professional might and have used the concept of professionalism as an excuse to not obey civilian authorities [@finer_man_1962].
+Finer is clear that the principle of civilian control - not military professionalism - is the true and only bulwark that can keep the military out of politics.
+
+Describing the various motives that militaries might have for intervening in politics, Finer goes on to describe at least three.
+He first lists the motive of the "manifest destiny of the soldiers," in which the military acts as "saviors" of the country [@finer_man_1962, 28-29].
+This motive, Finer claims, arises because of the confluence of the military's unique mission, sense of power, and claim of being "outside party politics" [@finer_man_1962, 29].
+A second motive that militaries might possess which disposes them to intervene in politics is what Finer calls the "unique identification with the national interest" [@finer_man_1962, 30].
+Acting from such a motive, Finer conceptualizes the military acting as "custodians" of the "national interest" [@finer_man_1962, 30].
+Finer acknowledges that "custodianship" can take on different forms in different contexts, but points out that some militaries might view "custodianship" as a "duty to arbitrate or veto...to exercise it when some convulsion or decision of the civil autorities seems to them [the military] to threaten what they think are the permanent interests of the nation" [@finer_man_1962, 31].
+The final motive Finer spends the most time on, calling it "the sectional interest," which refers to a more narrow subset of interests than the "national interest," and includes a wide ranging set of interests based on social class, regional affiliation, personal career motivations, or organizational interests [@finer_man_1962, 34-53].
+
+Ultimately, Finer develops a typology of four levels, or degress of, military intervention.
+The first he calls "influence," and argues that at this level of intervention, the military makes appeals "to reason or emotion" in an attempt to persuade the civilian government [@finer_man_1962, 77].
+Finer adds that intervening at this level is, in his mind at least, not problematic and entirely "consistent with the supremacy of the civil power" [@finer_man_1962, 77].
+The next level of intervention is "pressures or blackmail," describing efforts by the military to "convince the civil power by threat of some sanction" [@finer_man_1962, 77].
+Finer adds that military intervention via this level is "on the margin," referring to his belief that while some cases may be unproblematic, many other forms, such as the "threat to stage a coup," would be normatively wrong [@finer_man_1962, 77].
+Finer points out a third level of intervention, that of "displacement," referring to the "removal" of government officials by the military as a result of "violence or the threat of violence" [@finer_man_1962, 78].
+The fourth level, which is basically displacement on a wider scale, Finer calls "supplantment," which "sweeps away the civilian regime and establishes the military in its place" [@finer_man_1962, 78].
+Finer goes on to explain the level of intervention as a function of the level of "political culture" that is prevalent in a state and society  [do I need more here?
+See Finer, pp.
+79 and following]
+
+Another helpful conceptualization of military intervention in politics is the idea of "role beliefs," a concept articulated by the political scientist, J.
+Samuel Fitch, in his book, _The Armed Forces and Democracy in Latin America._
+
+A third, and equally rich, study of the degrees of military intervention in politics is available from Brian Taylor's book, _Politics and the Russian Army: Civil-Military Relations, 1689-2000.
+
+## Political Polarization
+
+The question of if and how political polarization drives changes in the professional identity of the military spans several related literatures, most notable of which are American Politics and International Relations.
+Broadly speaking, this project is focused on how one aspect of domestic politics, political polarization, impacts a specific but important aspect of international relations, civil-military relations, which itself spans a wide and diverse literature.
+Because I intend to focus my work on the military as an institution, my dissertation will also rely on literature from the fields of sociology, organizational psychology, and ethics.
+I will frame my research question within each of these fields briefly, and to define important concepts as required, before addressing why this question matters and should be studied in depth.
+
+Within the American Politics literature, there is broad recognition that political polarization levels in contemporary America are higher now than at any point since the end of the American Civil War [@hare_polarization_2014], but there are multiple explanations for what has caused this trend, ranging from a growing divide over matters pertaining to race and religion [@abramowitz_great_2018], the size and role of government [@sides_two_2015], and disagreements over specific policy issues [@mann_its_2012].
+Despite the wide-ranging scholarship exploring the many facets of political polarization, virtually all scholars agree that contemporary levels of political polarization in the United States are historically abnormal and pose myriad problems.
+
+One of the major concerns regarding political polarization is its on impact on American foreign policy in a broad sense, and specifically, on America's ability to lead and impact the world.
+Kenneth Schultz, for instance, underscored four challenges that have emerged as a result of America's polarized domestic politics, including the ability to obtain bipartisan support for "risky undertakings," the ability to learn the "lessons of foreign policy failure," the inability to "make long term commitments to allies and adversaries" as a result of "dramatic policy swings from one administration to another," and the "vulnerability" of America's electoral system to foreign third party adversaries [@schultz_perils_2017].
+Because the military is a central actor in shaping America's foreign policy, obtaining a better understanding of how political polarization shapes various military institutions will yield findings that may help policymakers and academics help solve the problems Schultz identifies.
 
@@
 First, one of the most fundamental normative hallmarks of the military's professional nature, at least by Huntingtonian standards, is that it ought to be an apolitical and non-partisan institution.
-Second, one of the features of the US military is that it prizes and elevates the trait of loyalty among those who serve and, by statute as well as normative tradition, exercises simulataneous dual loyalties to its civilian heads, the President as the Commander in Chief, and Congress.
+Second, one of the features of the US military is that it prizes and elevates the trait of loyalty among those who serve and, by statute as well as normative tradition, exercises simultaneous dual loyalties to its civilian heads, the President as the Commander in Chief, and Congress.
 And third, we know from the American politics literature that ideological polarization and partisan loyalty are connected in that those citizens who hold to stronger ideological convictions tend to be the most loyal partisans in terms of voting and other political behaviors [@abramowitz_great_2018, 112-116].
@@
 Moreover, at the level of the military institution, a similar problem may also emerge: will the military institution have to choose between competing objects of loyalty, such as the President as the Commander in Chief, the Congress, and even its own professional ethic?^[Loyalty had to be a prime consideration in several prominent, recent cases involving the military and the government.
-Consider the case of now retired Army Lieutenant Colonel Alex Vindman, who testified before Congress against the US President, Donald Trump, regarding what Vindman believed to be behavior by the President that may have been illegal, or the case of Navy Captain Brett Crozier, who leaked a letter to the press after his pleas for help in battling the coranavirus aboard the ship he commanded, the USS Theodore Roosevelt, went unheeded for what he felt was too long.
+Consider the case of now retired Army Lieutenant Colonel Alex Vindman, who testified before Congress against the US President, Donald Trump, regarding what Vindman believed to be behavior by the President that may have been illegal, or the case of Navy Captain Brett Crozier, who leaked a letter to the press after his pleas for help in battling the coronavirus aboard the ship he commanded, the USS Theodore Roosevelt, went unheeded for what he felt was too long.
 Certainly these instances involved individuals making choices about to whom or to what they would be most loyal to - the President, the Constitution, the ideals and obligations of the military, the operational mission, themselves as individual moral agents, and perhaps even more objects of loyalty.]
@@
 While I believe that these potential developments are catching the attention of civil-military relations scholars, none to my knowledge have specifically set out to systematically explore if and how polarized partisan politics are impacting the military's professional identity and character, and in particular, the role of the military institution itself in facilitating these developments.
-Most scholars acknowledge the potentially damaging impacts of the politicization of the military [@golby_listening_2013, @brooks_paradoxes_2020] at the hands of civilian leaders, and I share these concerns.
+Most scholars acknowledge the potentially damaging impacts of the politicization of the military [@golby_listening_2013; @brooks_paradoxes_2020] at the hands of civilian leaders, and I share these concerns.
 Fewer scholars, if any, are exploring the potential role of the military in politicizing itself, and the mechanisms through which these developments occur.
@@
 
-In line with Huntington's model as well as a host of other scholarship, I suggest that changes in military professionalism can impact a variety of military outcomes, varying from effectiveness on the battlefield [@reiter_democracies_2002; @biddle_democracy_2004] to achieving strategic effectivness [@brooks_creating_2007; @brooks_shaping_2008].
+In line with Huntington's model as well as a host of other scholarship, I suggest that changes in military professionalism can impact a variety of military outcomes, varying from effectiveness on the battlefield [@reiter_democracies_2002; @biddle_democracy_2004] to achieving strategic effectiveness [@brooks_creating_2007; @brooks_shaping_2008].
 While I intend to examine the broad system of variables shown in Figure 2, the focus of my dissertation is the relationship between political polarization and changes in the concept of loyalty and identity, and how such changes manifest themselves in the variation of military professionalism.
@@
 I contend that these features impact individuals serving in the military, including military leaders, as they do members of the domestic public writ large, and that these impacts are problematic for members of an institution that normatively exists as an apolitical and non-partisan actor.
-Table 1, below, lays out how these three prominent features of domestic political polarilzation may impact the non-partisan and apolitical nature of the normative military professional and institution.
+Table 1, below, lays out how these three prominent features of domestic political polarization may impact the non-partisan and apolitical nature of the normative military professional and institution.
 
@@
 With these conditions, the military does not have an out-sized platform to interact with the American public, nor are there outstanding incentives for the military to raise significant disagreement over any domestic political issues.
+
 \singlespacing
@@
 \end{tabular}
-\label{table:comparison} % is used to refer this table in the text
+\label{table:comparison1} % is used to refer this table in the text
 \end{footnotesize}
@@
 _H1: In highly polarizing times, there will be an increase in the frequency of military officers breaking normative civil-military relationship behaviors, such as criticizing the President, members of Congress, or those who hold to different political views on particularly-polarizing issues outside of the conduct of war._
+\doublespace
+\singlespace
 
 _H2: These breaks in normative civil-military behaviors will occur because political polarized periods feature events or situations which test or challenge the military's ability to juggle and balance multiple simultaneous loyalties: to the President as Commander in Chief, to Congress in its oversight role of the military, to the underlying ideologies or ideas inherent in issues that prevail during polarized times, to the normative professional ideals of the military, and to the individual's bonds of loyalty to his or fellow soldiers and the mission._
+\doublespace
+\singlespace
 
 _H3: Because of the normative strength and prominence of place that the ideal and concept of loyalty has within the profession of arms, military leaders are more sensitive than regular civilians to appeals of loyalty._
+\doublespace
+\singlespace
 
@@
 These can be viewed as "exploits of loyalty."_
+\doublespace
+\singlespace
 
 _H5: In highly polarized time periods, it is the senior military officers as the "stewards of the profession" [@us_department_of_the_army_adp_2019, p.1-9] whose actions will most influence to what degree and how the military institution reacts to and changes as a result of "exploits of loyalty." If changes occur to the military institution, they are bound to occur to the professional identity of the military, which includes the degree to which it functions as a profession (vice bureaucracy [@snider_once_2012]) and the degree to which it is obedient to its civilian leaders (vice acting as an activist on behalf of a particular cause)._
+\doublespace
 
-\doublespace
 # Plan of the Dissertation and the Data Generating Process
@@
 I will emphasize the long-running nature of this debate going all the way back to when Huntington first published his book, but point to some of the prominent contemporary critiques regarding his theory [@feaver_civil-military_1996-1; @rapp_civil-military_2015; @golby_why_2017; @karlin_case_2020; @brooks_paradoxes_2020].
-The main point of chapter 1 is to set the stage for my dissertation and signal to the reader that modern critiques of Huntington may be overlooking a key part of his concept of military professionalism: imperfect as it may be, it may also insulate the military from the most pernicious effects of polarizing politics.
+The main point of chapter one is to set the stage for my dissertation and signal to the reader that modern critiques of Huntington may be overlooking a key part of his concept of military professionalism: imperfect as it may be, it may also insulate the military from the most pernicious effects of polarizing politics.
 
@@
 In the third chapter, I will argue that today, in 2020, the American public is witnessing a decline in professional military conduct among retired military officers.
-I plan to demonstrate this empirically by drawing upon an original dataset consisting of more than 340 op-ed pieces authored by retired military officers since the late 1970s and published in the top five major print newspapers in the United States, including the Wall Street Journal, the New York Times, the Washington Post, the USA Today, and the Los Angeles Times.
+I plan to demonstrate this empirically by drawing upon an original data set consisting of more than 340 op-ed pieces authored by retired military officers since the late 1970s and published in the top five major print newspapers in the United States, including the Wall Street Journal, the New York Times, the Washington Post, the USA Today, and the Los Angeles Times.
 I believe that I can show several things from this analysis, including that there has been a sharp uptick in the number of op-eds published by retired military officers during this time period, and more relevant to the theory I propose in the dissertation, that these published op-eds reveal a decline in the normative professional ideals of the military institution.
@@
 I will argue that while there are certainly other confounding variables, to include changes in the conservative-liberal tilt of newspapers over time, the chief explanation for the variation (decline) in professionalism stems from the combined effects of increasing political polarization and rising institutional credibility of the military since the late 1970s, and the challenges that these changes introduce to a military leader's conception of loyalty.
-Moreover, my ultimate point is to point out a blatant inconsistency in the way that the military currently thinks about the political behavior by those within the ranks in that the institution has rules that governs the conduct of those on active duty [@department_of_defense_political_2008], but those same rules, and the spirit behind those rules, are disregarded by the institution's vangaurds, its retired multistar officers and "stewards of the profession" [@us_department_of_the_army_adp_2019, p.
+Moreover, my ultimate point is to point out a blatant inconsistency in the way that the military currently thinks about the political behavior by those within the ranks in that the institution has rules that governs the conduct of those on active duty [@department_of_defense_political_2008], but those same rules, and the spirit behind those rules, are disregarded by the institution's vanguards, its retired multistar officers and "stewards of the profession" [@us_department_of_the_army_adp_2019, p.
 1-9] of arms.
@@
 
-In Chapter four, I plan to analyze the results of an original survey experiment in which I survey a nationally representative sample of civilians and compare the results to a sample of either a) veterans, or b) active duty military officers.
-In this survey experiment, I plan to test whether members of the military (which I proxy by veterans) genuinely are more sensitive to appeals to loyalty, as I have hypothesized in _H3._ The survey, after controlling for gender, partisan identification, and age, will present the respondent with 6-8 hypothetical situations and a corresponding argument made by a military elite.
-Members of the treatment group will receive an appeal to loyalty as part of the elite's argument as to why a certain course of action should be followed regarding the hypothetical situation presented.
+In Chapter four, I plan to analyze the results of an original survey experiment in which I survey a nationally representative sample of civilians and compare the results to a sample of either a) veterans, or b) active duty military officers (pending a lengthy approval process through the Army).
+I intend to design a survey experiment that measures whether arguments that appeal to one's individual or collective sense of loyalty are more effective than those messages that do not incorporate such language, and more directly, whether such arguments are acutely effective among military audiences.
+Such an experiment will build off of recent surveys that have examined the impacts of "moral reframing" [@feinberg_moral_2019] and "moral messaging" [@luttrell_challenging_2019], which describe efforts senders of messages employ to persuade their audiences by appealing to the audience's sense of morality.
 
-I plan to structure the survey such that the scenario and corresponding course of action contain references to issues that are highly correlated, in some cases, to partisan / ideological affiliation.
-An example of a question might be as follows:
+Moral reframing is an extension of the broader _Moral Foundations Theory_, a paradigm developed by psychologists several years ago [@graham_liberals_2009] that seeks to explain the moral domain in terms of five distinct moral foundations: "harm/care, fairness/reciprocity, ingroup/loyalty, authority/respect, and purity/sanctity" [@graham_liberals_2009].
+Several follow on studies have confirmed that political liberals and conservatives rely on these five moral foundations in profoundly different ways, particularly when making "moral judgements" [@graham_liberals_2009]: liberals tend to rely on a morality based primarily on the foundations of harm/care and fairness/reciprocity, while conservatives tend to rely on a "more even distribution" of the five values, to include the "binding" foundations of ingroup/loyalty, authority/respect, and purity/sanctity [@graham_liberals_2009].
+My intended survey experiment will advance the field because it will explore the degree to which one particular moral foundation - ingroup/loyalty - resonates within an organization that emphasizes and values this particular moral foundation, as I have laid out in the preceeding pages.
+To my knowledge, while there have been many studies that examine the extent to which partisans rely on differing moral foundations [need cite], there have not been studies that examine the degree to which one particular moral foundation binds members of an organization.
+The military serves as a potentially very interesting case study, particularly concerning the moral foundation of ingroup/loyalty.
+
+Therefore, I will test whether an appeal to loyalty works stronger among members of the military than among members of the civilian population, as I have hypothesized in _H3._ The survey, after controlling for gender, partisan identification, and age (experience for military), will present the respondent with 6-8 hypothetical situations and a corresponding argument made by military voice (the biographical particulars of which I will also vary).
+Members of the treatment group will receive an appeal to loyalty as part of the military voice's argument as to why a certain course of action should be followed regarding the hypothetical situation presented.
+
+The survey will be structured such that the scenario and corresponding course of action contain references to issues that are, at least in some of the cases, highly correlated with partisan / ideological affiliation.
+I will also vary the treatment such that while each treatment will be an appeal to loyalty, the specific object of loyalty within the treatment condition will be varied.
+For example, some questions will appeal to the military member's sense of loyalty to the Constitution, while another may appeal to the military member's sense of loyalty to the mission and/or the member's fellow soldiers.
+An example of one of these questions is as follows:
+
 \singlespace
@@
 
-Such a survey will enable me to measure two things.
-First, it will allow me to measure whether appeals to loyalty are truly more effective than standard appeals to action.
-This result can be measured by comparing the answers between the treatment and control groups.
-Second, it will allow me to measure whether members of the military (or veterans) react differently to these appeals to loyalty than regular civilians.
-As I have hypothesized in _H3_, I suspect that they do.
-Of note, I am planning to also pursue a survey among actively-serving military officers, but the lead time for obtaining approval to do this can be very long, perhaps 9-12 months or more.
-If I am able to obtain approval for this, I will then have three different populations to measure: civilians, veterans, and actively-serving military officers.
-Of note, if I discover throughout the course of my survey that loyalty does not in fact have the strong explanatory power that I think it does, or if veterans and/or active military officers do not respond to appeals to loyalty any differently than civilians, I think such a finding is also interesting.
+Such a survey will enable me to measure at least three things.
+First, it will allow me to measure whether appeals to loyalty are truly more effective than standard arguments that do not invoke obligations of loyalty.
+This result can be measured by comparing the answers between the treatment and control groups for both populations.
+The second object this survey will enable me to measure is whether members of the military (or veterans) react differently to appeals based in the trait of loyalty than regular civilians, as I have hypothesized in _H3_.
+The third thing this survey will allow me measure is the relative strength, if any, of appeals made to differing objects of loyalty (to the Constitution, to the Commander in Chief, to the sense of mission, to one's fellow soldiers, etc.) in relation to a particular course of action.
+The results of this chapter will inform a micro-foundational understanding of the strength of loyalty as it pertains to the decision-making of individuals.^[Of particular note, I am planning to also pursue a survey among actively-serving military officers, but the lead time for obtaining approval to do this can be very long, perhaps 9-12 months or more.
+If I am able to obtain approval for this, I will then have three different populations to measure: civilians, veterans, and actively-serving military officers.]
 
@@
 In the sixth chapter, I will turn to a series of qualitative case studies to explore and test my theory.
-Key in this chapter is finding and examining historical periods in American civil-military relations in which there existed varying levels of political polarization and institutional credibility, and then, to walk the reader through how challenges of loyalty during these periods manifested themselves in different ways among specific members of the military and the institional military at large, in line with the theory that I propose in Chapter 2 and specifically, _H2_.I then hope to find evidence for how the actions of the military and its leaders during these periods impacted the military's professional identity and character, in line with _H4_ and _H5_.
+Key in this chapter is finding and examining historical periods in American civil-military relations in which there existed varying levels of political polarization and institutional credibility, and then, to walk the reader through how challenges of loyalty during these periods manifested themselves in different ways among specific members of the military and the institutional military at large, in line with the theory that I propose in Chapter 2 and specifically, _H2_.I then hope to find evidence for how the actions of the military and its leaders during these periods impacted the military's professional identity and character, in line with _H4_ and _H5_.
 
@@
 My initial reaction is that I do not need to confine myself to this period.
-My initial projection of what these case studies will be and how they will align with varying levels of political polarization and instititional credibility are as follows:
+My initial projection of what these case studies will be and how they will align with varying levels of political polarization and institutional credibility are as follows:
 
@@
 
-# Initital Concerns and Timeline
+# Initial Concerns and Timeline
 
@@
 To that end, I may need to adopt a more comparative approach that examines if and how other militaries have been impacted by political polarization in their countries, and thus attempt some sort of cross-national study, should one of my intended chapters be deemed insufficient.
-I note this especially because the concept of loyalty, as mentioned in this prospectus, features so prominentaly in international relations in the context of coups, ethnic exclusion, etc., and I want to make sure I am not missing an opportunity to bring these matters to bear on my research project.
+I note this especially because the concept of loyalty, as mentioned in this prospectus, features so prominently in international relations in the context of coups, ethnic exclusion, etc., and I want to make sure I am not missing an opportunity to bring these matters to bear on my research project.
 

### 2020-11-18 code/Submissionv4.Rmd
commit

--- 2020-09-21 code/Submissionv3.Rmd
@@
     - \usepackage{indentfirst}
+    - \usepackage{caption}
 
@@
 
+\captionsetup[table]{labelformat=empty}
+
 ```{r setup, include=FALSE}
@@
 library("reshape2")
+library(stargazer)
 ```
@@
 
-dataset <- read_csv(here("data", "MilEditorials10Aug.csv"))
+dataset <- read_csv(here("data", "MilEditorials.csv"))
 dataset %>% print()
-# Only read in the first 344 rows - is there a better way to get rid of N/A data?
-# Replace NA with 0 in the data for variables War_Ops thru Fig_Nme
-adjdataset <- dataset[1:344,] %>%
+# Only read in the first 354 rows - is there a better way to get rid of N/A data?
+
+# Replace NA with 0 in the data for variables letter_ed to online_only, and leg_act #to role
+adjdataset <- dataset[1:353,] %>%
   mutate_at(
-    .vars = vars(War_Ops:Fig_Nme),
+    .vars = vars(letter_ed:online_only, Leg_Act:Role),
     .funs = ~ case_when(is.na(.) ~0, TRUE ~ 1)
@@
 # count the number of observations by type of binary variable and store as new dataframe
-countbytype  <- adjdataset %>% count(War_Ops, Trp_Clt, Str_Adv, MilPSoc, Dom_P, Fig_Nme) 
-countbytype
+countbytype  <- adjdataset %>% 
+  count(Topic) %>% print()
+
+# count the number of observations by source
+countbysource <- adjdataset %>% 
+  count(Source) %>% print()
 
@@
   group_by(PubYr) %>% 
-  mutate(Yrcount=n()) 
+  mutate(Yrcount=n()) %>% 
+  view()
 ```
@@
   x="Year", 
-  y="count", 
-  title="Figure Two: Number of Op-Eds Published by Retired\n Military Officers (WSJ, NYT, WaPo, USA Today, and LA Times",
+  y="yearly count", 
+  title="Figure Two: Number of Op-Eds Published by Retired Military\n Officers (WSJ, NYT, WaPo, USA Today, LA Times)",
   subtitle= "1979-2020",
-  caption="Note: As of 31 August 2020"
+  caption="Note: As of 29 October 2020"
   ) + scale_x_continuous(limits=c(1979, 2020)) + scale_y_continuous(limits=c(-7, 28)) + 
@@
   geom_vline(xintercept=2017, size=.2, colour="tomato") +
-  geom_text(aes(x=1981, label="Reagan",y=-5), colour="tomato", angle=90, vjust=1, text=element_text(size=9)) + 
-  geom_text(aes(x=1989, label="Bush 41",y=-5), colour="tomato", angle=90, vjust=1, text=element_text(size=9)) + 
-  geom_text(aes(x=1993, label="Clinton",y=-5), colour="dodgerblue", angle=90, vjust=1, text=element_text(size=9)) + 
-  geom_text(aes(x=2001, label="Bush 43",y=-5), colour="tomato", angle=90, vjust=1, text=element_text(size=9)) + 
-  geom_text(aes(x=2009, label="Obama",y=-5), colour="dodgerblue", angle=90, vjust=1, text=element_text(size=9)) + 
-  geom_text(aes(x=2017, label="Trump",y=-5), colour="tomato", angle=90, vjust=1, text=element_text(size=9))
+  geom_text(aes(x=1981, label="Reagan",y=10), colour="tomato", angle=90, vjust=1, text=element_text(size=9)) + 
+  geom_text(aes(x=1989, label="Bush 41",y=15), colour="tomato", angle=90, vjust=1, text=element_text(size=9)) + 
+  geom_text(aes(x=1993, label="Clinton",y=20), colour="dodgerblue", angle=90, vjust=1, text=element_text(size=9)) + 
+  geom_text(aes(x=2001, label="Bush 43",y=-2), colour="tomato", angle=90, vjust=1, text=element_text(size=9)) + 
+  geom_text(aes(x=2009, label="Obama",y=-2), colour="dodgerblue", angle=90, vjust=1, text=element_text(size=9)) + 
+  geom_text(aes(x=2017, label="Trump",y=-2), colour="tomato", angle=90, vjust=1, text=element_text(size=9))
 
@@
 
-# Show all articles and authors who published articles on Social Policy, Political Parties, and Administrations/Political Figures
-rem_obs <-adjdataset %>% arrange(PubYr) %>% 
-  filter(MilPSoc==1 | Dom_P==1 | Fig_Nme==1 )  %>% 
-  select (Author, Title, Source, PubYr, MilPSoc, Dom_P, Fig_Nme) %>% 
+# Show all articles and authors who published articles on Domestic Policy and Social Policy 
+adjdataset %>% arrange(PubYr) %>% 
+  filter(Topic=="DomesticPolicy" | Topic=="SocialPolicy")  %>% 
+  select (Author, Title, Source, PubYr, Topic) %>% 
+  print()
+
+# Show all articles where Leg_Act, Crit_Insub, or Endorse_Partisan was equal to 1
+## (in other words, articles in which the author called for legislative activity, criticized, or made
+## an explicitly partisan statement)
+
+adjdataset %>% arrange(PubYr) %>% 
+  filter(Leg_Act== 1 | Crit_Insub== 1 | Endorse_Partisan ==1)  %>% 
+  select (Author, Title, Source, PubYr, Topic, Leg_Act, Crit_Insub, Endorse_Partisan) %>% 
+  print()
+
+# create a df called quest_prof that contains articles that either criticized or were #expressly partisan, as well as whether someone was in a role at the time that may have # made such behavior ok
+
+quest_prof <-adjdataset %>% arrange(PubYr, Topic) %>% 
+  filter(Crit_Insub== 1 | Endorse_Partisan ==1)  %>% 
+  select (Author, Title, Source, PubYr, Topic, Crit_Insub, Endorse_Partisan, Role) %>% 
   view()
 
-# Create a Variable Called case-count that counts the number of times by publishing year an article published 
-# regarding social policy, domestic policy, or political figures/administrations occurs
+## Do a couple of things here.  Create several new variables.  
+# The first is a variable called case-count that counts the number of times by publishing #year an article published that calls for legislative action, criticizes or is #insubordinate, or is blatantly partisan.
+The second variable, prop_yr, is simply a
+# proportion of annual publications where these conditions hold true.
+# The third variable is called dom_pol_count which counts the number of publications
+# by publishing year that are written on topics of domestic policy, and the fourth 
+# variable is a proportion of this type of article by year.
 
-adj2dataset <- adjdataset %>% group_by(PubYr) %>% 
+adjdataset <- adjdataset %>% group_by(PubYr) %>% 
   mutate(
-    case_count=sum(MilPSoc ==1 | Dom_P==1 | Fig_Nme==1), 
-    yr_total_count=sum(War_Ops==1 | Trp_Clt==1| Str_Adv==1 | MilPSoc ==1 | Dom_P==1 | Fig_Nme==1),
-    prop_yr=(case_count/yr_total_count)
+    case_count=sum(Leg_Act ==1 | Crit_Insub==1 | Endorse_Partisan==1), 
+    prop_yr=(case_count/Yrcount),
+    dom_pol_count=sum(Topic=="DomesticPolicy"),
+    dom_pol_prop=(dom_pol_count/Yrcount)
   ) %>% view()
 
-# Present Results of Result and Justification
-secondlayer_obs <-adjdataset %>% arrange(PubYr) %>% 
-  filter(MilPSoc==1 | Dom_P==1 | Fig_Nme==1 )  %>% 
-  select (Author, Title, Source, PubYr, Ques_Prof, Justification) %>% 
+# Now create two variables, leg_crit_part and crit_part that take a 1 if certain 
+# conditions are met.  This will come in handy for the glm regressions that
+# will be conducted later.
+
+adjdataset <- adjdataset %>% 
+  mutate(
+    leg_crit_part=case_when(
+      (Leg_Act==1 | Crit_Insub ==1 | Endorse_Partisan==1 & Role ==0) ~ 1,
+      (Leg_Act != 1 & Crit_Insub != 1 & Endorse_Partisan !=1 | Role==1) ~ 0
+    ),
+    crit_part=case_when(
+      (Crit_Insub ==1 | Endorse_Partisan==1 & Role ==0) ~ 1,
+      (Crit_Insub != 1 & Endorse_Partisan != 1 | Role ==1) ~ 0
+    )
+  ) %>% 
   view()
+
 ```
@@
 #Need to adjust the width and spacing of this table
-secondlayer_obs %>%
-  kable(format="latex", caption="Potentially Unprofessional Op-Ed Pieces By Retired Military Officers, 1979-2020", longtable=T) %>%
+quest_prof %>%
+  kable(format="latex", caption="Op-Eds authored by Retired Military Officers that Strongly Criticized Civilian Leaders, or were Expressly Partisan, 1979-2020", longtable=T) %>%
   kable_styling(font_size=10, latex_options="striped", repeat_header_continued=T) %>% 
@@
 ```{r build plots, include = FALSE, echo=FALSE}
-plot1<- ggplot(data=adj2dataset, mapping=aes(x=PubYr, y=case_count)) + geom_point() + geom_smooth() + labs(x="Year", y="Number", color=NULL) + geom_vline(xintercept = 1981, size=.2, colour="tomato") +
+plot1<- ggplot(data=adjdataset, mapping=aes(x=PubYr, y=prop_yr)) + geom_point() + geom_smooth() + labs(x="Year", y="Annual Proportion", color=NULL) + geom_vline(xintercept = 1981, size=.2, colour="tomato") +
   geom_vline(xintercept=1989, size=.2, colour="tomato") + 
@@
   geom_vline(xintercept=2017, size=.2, colour="tomato") +
-  geom_text(aes(x=1981, label="Reagan",y=3), colour="tomato", angle=90, vjust=1, text=element_text(size=9)) + 
-  geom_text(aes(x=1989, label="Bush 41",y=3), colour="tomato", angle=90, vjust=1, text=element_text(size=9)) + 
-  geom_text(aes(x=1993, label="Clinton",y=3), colour="dodgerblue", angle=90, vjust=1, text=element_text(size=9)) + 
-  geom_text(aes(x=2001, label="Bush 43",y=5), colour="tomato", angle=90, vjust=1, text=element_text(size=9)) + 
-  geom_text(aes(x=2009, label="Obama",y=5), colour="dodgerblue", angle=90, vjust=1, text=element_text(size=9)) + 
-  geom_text(aes(x=2017, label="Trump",y=-1), colour="tomato", angle=90, vjust=1, text=element_text(size=9))
+  geom_text(aes(x=1981, label="Reagan",y=.2), colour="tomato", angle=90, vjust=1, text=element_text(size=9)) + 
+  geom_text(aes(x=1989, label="Bush 41",y=.2), colour="tomato", angle=90, vjust=1, text=element_text(size=9)) + 
+  geom_text(aes(x=1993, label="Clinton",y=.4), colour="dodgerblue", angle=90, vjust=1, text=element_text(size=9)) + 
+  geom_text(aes(x=2001, label="Bush 43",y=.3), colour="tomato", angle=90, vjust=1, text=element_text(size=9)) + 
+  geom_text(aes(x=2009, label="Obama",y=.4), colour="dodgerblue", angle=90, vjust=1, text=element_text(size=9)) + 
+  geom_text(aes(x=2017, label="Trump",y=0), colour="tomato", angle=90, vjust=1, text=element_text(size=9))
 
-plot2 <- ggplot(data=adj2dataset, mapping=aes(x=PubYr, y=prop_yr)) + geom_point() + geom_smooth() + labs(x="Year", y="Proportion", color=NULL)+ geom_vline(xintercept = 1981, size=.2, colour="tomato") +
+plot2 <- ggplot(data=adjdataset, mapping=aes(x=PubYr, y=dom_pol_prop)) + geom_point() + geom_smooth() + labs(x="Year", y="Proportion", color=NULL)+ geom_vline(xintercept = 1981, size=.2, colour="tomato") +
   geom_vline(xintercept=1989, size=.2, colour="tomato") + 
@@
   geom_text(aes(x=1993, label="Clinton",y=.25), colour="dodgerblue", angle=90, vjust=1, text=element_text(size=9)) + 
-  geom_text(aes(x=2001, label="Bush 43",y=.5), colour="tomato", angle=90, vjust=1, text=element_text(size=9)) + 
-  geom_text(aes(x=2009, label="Obama",y=.5), colour="dodgerblue", angle=90, vjust=1, text=element_text(size=9)) + 
-  geom_text(aes(x=2017, label="Trump",y=.55), colour="tomato", angle=90, vjust=1, text=element_text(size=9))
+  geom_text(aes(x=2001, label="Bush 43",y=.3), colour="tomato", angle=90, vjust=1, text=element_text(size=9)) + 
+  geom_text(aes(x=2009, label="Obama",y=.3), colour="dodgerblue", angle=90, vjust=1, text=element_text(size=9)) + 
+  geom_text(aes(x=2017, label="Trump",y=-.05), colour="tomato", angle=90, vjust=1, text=element_text(size=9))
 
@@
 ```{r figure 3, include=TRUE, warning= FALSE, message = FALSE, echo=FALSE}
-grid.arrange(plot1, plot2, ncol=2, top="Figure 3: Number(L) and Proportion (R) of Op-Eds Published\n by Retired Military Officers on Non-Traditional Military Topics, 1979-2020***", bottom="*** Topics include: Social Policy which Impacts the Military,\n Domestic Policy, and Political Administrations/Figures")
+grid.arrange(plot1, plot2, ncol=2, top="Figure 3: Proportion of Op-Eds Published Which\n (L) Call for Legislative Action, Criticize Civilian Leaders, or are Partisan\n in nature; or (R) Address Domestic Political Topics")
 ```
@@
 ```{r figure 4, include=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
+#adjust the data to account for polarization by creating a new dataset called polaradjdataset.  
+#In this dataset, create variables that capture ideological distance in each the House and Senate along each of the two dimensions according to Poole and Rosenthal's DW-NOMINATE data
 
-#adjust the data to account for polarization by creating a new dataset called polaradjdataset.  
-#In this dataset, create variables that capture ideological distance in each the House and Senate along
-#each of the two dimensions according to Poole and Rosenthal's DW-NOMINATE data
+# scale the data by multiplying by 100
 
 polaradjdataset <-adjdataset %>% mutate (
-  House_distance_dim1=abs(House_Dem_mean_nom1-House_Rep_mean_nom1),
-  Sen_distance_dim1=abs(Sen_Dem_mean_nom1-Sen_Rep_mean_nom1),
-  House_distance_dim2=abs(House_Dem_mean_nom2-House_Rep_mean_nom2),
-  Sen_distance_dim2=abs(Sen_Dem_mean_nom2-Sen_Rep_mean_nom2)
-) %>% view()
+  House_distance_dim1=abs(House_Dem_mean_nom1-House_Rep_mean_nom1)*100,
+  Sen_distance_dim1=abs(Sen_Dem_mean_nom1-Sen_Rep_mean_nom1)*100,
+  House_distance_dim2=abs(House_Dem_mean_nom2-House_Rep_mean_nom2)*100,
+  Sen_distance_dim2=abs(Sen_Dem_mean_nom2-Sen_Rep_mean_nom2)*100
+) %>% view ()
 
@@
 
-#Now I Prep the data that I need to analyze polarization levels.  I will try first by 
-# eliminating the NA values but I will do this by indexing the first 334 rows of the 
-# data, which corresponds to the number of observations that I have.  I do this by #creating a new data set called slimdata.
-
-# Note: this is sloppy.  There has to be a better way to get rid of NA values.
-
-slimdata <- polaradjdataset[1:344,] %>% 
-  view()
-
-#next, conduct logit regressions where the DV is values of Ques_Prof
-# include two IVs: house_distance_dim1 and Inst_Cred_Gallup
+#next, conduct logit regressions where the DV is first, values of leg_crit_part and
+# then, where the DV is crit_part
 
@@
 
-logmodel1 <- glm(
-  Ques_Prof ~ Inst_Cred_Gallup + House_distance_dim1, 
+mod1 <- glm(leg_crit_part ~ Inst_Cred_Gallup + House_distance_dim1 + AuthSex + AuthRank + Topic, 
   family=binomial(link="logit"), 
-  data=slimdata
+  data=polaradjdataset
   )
 
-tidy(logmodel1)
-glance(logmodel1)
+summary(mod1)
+tidy(mod1)
+glance(mod1)
 
-# now  make predictions.
-
-# the first prediction data keeps credibility constant and varies house_distance_dim1 from it's min to its max
-predictiondata <- tibble(House_distance_dim1=seq(min(slimdata$House_distance_dim1), max(slimdata$House_distance_dim1), by=.01), Inst_Cred_Gallup = mean(slimdata$Inst_Cred_Gallup)) %>%
-crossing(Ques_Prof = c(0,1))
-
-predictiondata
-
-# the second prediction keeps credibility constant but varies house_distance_dim1 from its min to .99, which is a 
-# perfectly reasonable expectation of the next 10-20 years of increasing political polarization in Congress. 
-predictiondatafut <- tibble(House_distance_dim1=seq(min(slimdata$House_distance_dim1), .99, by=.01), Inst_Cred_Gallup = mean(slimdata$Inst_Cred_Gallup)) %>%
-crossing(Ques_Prof = c(0,1))
-
-predictiondatafut
-
-# actual prediction data now with code.  Create variables that tie the model together. #Create Variables .prob, conf.low, conf.high, and pred_final_rec
-
-#augment first prediction
-augment(logmodel1, newdata = predictiondata)
-
-#augment second prediction
-augment(logmodel1, newdata = predictiondatafut)
-
-# first prediction: house_distance_dim1 from min to max
-predictions <- augment(logmodel1, newdata=predictiondata) %>% 
-  mutate(
-    .prob=plogis(.fitted),
-    conf.low = plogis(.fitted -(qnorm(.975) * .se.fit)),
-    conf.high = plogis(.fitted + (qnorm(.975) * .se.fit)),
-    pred_final_rec = case_when(
-      .prob >= .5 ~ "Likely Violation of Professionalism",
-      .prob < .5 ~ "Violation of Professionalism not likely"
-    )
+# model 2
+mod2 <- glm(crit_part ~ Inst_Cred_Gallup + House_distance_dim1 + AuthSex + AuthRank + Topic, 
+    family=binomial(link="logit"),
+    data=polaradjdataset
     )
 
-predictions %>% arrange(.prob, .fitted, Ques_Prof, desc(.fitted)) %>% 
-  view()
+summary(mod2)
+tidy(mod2)
+glance(mod2)
 
-# second prediction: house_distance_dim1 from min to a predicted value of .99
-predictions2 <- augment(logmodel1, newdata=predictiondatafut) %>% 
-  mutate(
-    .prob=plogis(.fitted),
-    conf.low = plogis(.fitted -(qnorm(.975) * .se.fit)),
-    conf.high = plogis(.fitted + (qnorm(.975) * .se.fit)),
-    pred_final_rec = case_when(
-      .prob >= .5 ~ "Likely Violation of Professionalism",
-      .prob < .5 ~ "Violation of Professionalism not likely"
-    )
-  ) 
-predictions2 %>% arrange(.prob, .fitted, Ques_Prof, desc(.fitted)) %>% 
-  view()
+# model 3 - remove topic as an IV because there is significant correlation between
+# one type of topic (domestic policy) and the DV
 
-# now let's try some graphs
-# notice that the predictions are not in (0, 1)
-ggplot(predictions) +
-  aes(x = .fitted) +
-  geom_histogram()
-
-ggplot(predictions2) +
-  aes(x = .fitted) +
-  geom_histogram()
-# if we wanted predictions on the probability scale,
-#   we predict on the log scale,
-#   then apply inverse-logit (logistic CDF)
-# use plogis(): cumulative probability of the logistic distribution
-
-# demonstration: scatter log odds (x) and predicted probability (y)
-# with 95% CIs
-# (CIs don't look smooth in this example because .se.fit depends on X data)
-predictions %>%
-  ggplot() +
-  aes(x = .fitted, y = .prob) +
-  geom_pointrange(aes(ymin = conf.low, ymax = conf.high)) +
-  geom_smooth()
-
-predictions2 %>%
-  ggplot() +
-  aes(x = .fitted, y = .prob) +
-  geom_pointrange(aes(ymin = conf.low, ymax = conf.high)) +
-  geom_smooth()
-
-# model2
-
-logmodel2 <- glm(
-  Ques_Prof ~ House_distance_dim1 + Inst_Cred_Gallup, 
+mod3 <- glm(leg_crit_part ~ Inst_Cred_Gallup + House_distance_dim1 + AuthSex + AuthRank,
     family=binomial(link="logit"),
-  data=slimdata
+    data=polaradjdataset
     )
 
-tidy(logmodel2)
-glance(logmodel2)
+summary(mod3)
+tidy(mod3)
+glance(mod3)
 
-# now  make prediction data using the mean institutional credibility scores.  Vary #institutional credibilty and hold political polarization constant.
+# model 4 - same as mod3, but change the DV to crit_part
 
-predictiondata2 <- tibble(Inst_Cred_Gallup=seq(min(slimdata$Inst_Cred_Gallup), max(slimdata$Inst_Cred_Gallup), by=1), House_distance_dim1 = mean(slimdata$House_distance_dim1)) %>%
-crossing(Ques_Prof = c(0,1))
+mod4 <- glm(crit_part ~ Inst_Cred_Gallup + House_distance_dim1 + AuthSex + AuthRank,
+    family=binomial(link="logit"),
+    data=polaradjdataset
+    )
 
-predictiondata2
-
-# actual prediction data now with code.  Create variables that tie the model together. #Create Variables .prob, conf.low, conf.high, and pred_final_rec
-
-augment(logmodel2, newdata = predictiondata2)
-
-predictions2 <- augment(logmodel2, newdata=predictiondata2) %>% 
-  mutate(
-    .prob=plogis(.fitted),
-    conf.low = plogis(.fitted -(qnorm(.975) * .se.fit)),
-    conf.high = plogis(.fitted + (qnorm(.975) * .se.fit)),
-    pred_final_rec = case_when(
-      .prob >= .5 ~ "Likely Violation of Professionalism",
-      .prob < .5 ~ "Violation of Professionalism not likely"
-    )
-  ) 
-predictions2 %>% arrange(.prob, .fitted, Ques_Prof, desc(.fitted)) %>% 
-  view()
+summary(mod4)
+tidy(mod4)
+glance(mod4)
 ```
 
-```{r fig5, include=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
-predictions%>%
-  ggplot() +
-  aes(x = .fitted, y = .prob) +
-  geom_pointrange(aes(ymin = conf.low, ymax = conf.high)) +
-  geom_smooth() + labs(
-    x="log Mean Distance Between Republicans and Democrats\n in the US House of Representatives, 1979-2020",
-    y="Predicted Probability of A Retired Military Officer\n Publishing an Op-Ed that Violates Military Professionalism",
-    title="Figure 5: The Predicted Impact of Political Polarization\n on Military Professionalism"
-  )
+```{r regression table, include=TRUE, echo=FALSE, results="asis"}
+stargazer(mod3, mod4, 
+          title="Table 3: Logistic Regression Results", 
+          column.labels = c('Legislative Activity, Criticize, or Partisan','Criticize or Partisan'),
+          covariate.labels = c('Institutional Credibility', 'Congressional Polarization', 'Male', 'Mixed Authorship', 'Unknown Authorship', 'Rank'),
+          align=TRUE)
+
 ```

### 2020-11-18 code/prospectusv2.Rmd
commit

@@
 ---
-title: "Saviors or Syndicates?: The Rise of Domestic Political Interventionism by Retired Military Leaders (Prospectus)"
+title: "Servants or Syndicates?: The Rise of Domestic Political Interventionism by Senior Military Leaders (Prospectus)"
 author: "Peter M. Erickson"
@@
 
-  _This dissertation explores the changes in how a powerful block of military professionals - retired senior military leaders - intervene in domestic contemporary politics.
-I advance a theory which argues that changes in political activism by retired senior leaders is a result of two interacting factors: the level of domestic political polarization within a society, and the degree of prestige and credibility that military leaders enjoy within the society.
-To support my argument, I trace changes in levels and types of distinct forms of political activism, including authoring opinion-editorials in major newspapers, and participation in presidential endorsements.
-I further advance the theory by conducting a survey experiment that tests for differences between civilian and military responses after exposure to situations that evoke one's partisan or ideological identity.
-I ultimately argue that the increasing political activism of retired military leaders is not merely a response to the hyper-partisanship present in the current political atmosphere, but rather the logical outworking of trends that have been active for some time in which both the ideological differences between Americans and the prestige of the military have increased.
-I conclude by describing the effect of the implications of growing politicla activism by senior military leaders on the core beliefs of the military regarding its very own essence and role in society._
+  _This dissertation explores changes in how the contemporary military institution intervenes in domestic politics.
+I advance a theory which argues that normative beliefs of military officers regarding political activism is a result of two interacting factors: the level of domestic political polarization within a society, and the degree of prestige and credibility the military holds.
+To support my argument, I trace changes in levels and types of distinct forms of political activism by military officers, including authoring opinion-editorials in major newspapers, as domestic polarization and military institutional credibility rises.
+I ultimately argue that the increasing political activism of retired military leaders is not merely a response to the hyper-partisanship present in the current political atmosphere, but rather the logical outworking of historical trends that have been building for decades.
+I conclude by describing the implications of changing beliefs among military leaders regarding the role the military should have in domestic politics.
+This dissertation contributes to the fields of international relations and civil-military relations by advanving an empirical and historical analysis of a growing phenomenon: military leaders intervening in domestic American politics.
+_
 
@@
 
-The central argument made in this dissertation is that increasing political activism at the hands of retired military officers is the result of two longstanding and interconnected trends: rising political polarization within the domestic political sphere, and rising prestige and credibility of military institutions afforded by American society.
+The central argument made in this dissertation is that increasing political activism at the hands of military officers is the result of two longstanding and interconnected trends: rising political polarization within the domestic political sphere, and rising prestige and credibility of military institutions afforded by American society.
 The interaction of these trends explains the remarkable increase in political activity by retired senior military leaders, and portends a shift in the perceived role that these senior military leaders fill vis-a-vis civilian authority - a shift away from professional and obedient servants and towards "guardians of democracy" who pick and choose when and how to intervene in American politics.
@@
 
-Civil-military relations involve a wide host of significant and important theoretical questions that involve the pull and push of power between institutions, the delineation of authority between military and civilian leaders, and a number of other consequential issues.
-At its root, the study of civil-military relations examines what Peter Feaver calls the "problematique," or how to ensure that the military is strong enough to defeat threats to the polity without threatening the polity itself [@feaver_civil-military_1996; @feaver_civil-military_1999].
+My proposed dissertation topic features within the subfield of civil-military relations, a broad academic subfield that is generally placed within the International Relations and Comparative Politics subfield (though within the American context, the subfield of American Politics also certainly applies).
+Civil-military relations involve a wide host of significant and important theoretical questions that involve the pull and push of power between institutions, the delineation of authority between military and civilian leaders, and the relationship between a state's military and society.
+
+At its root, the study of civil-military relations examines what Peter Feaver calls the "problematique," or the question of how to ensure that a military is strong enough to defeat threats to the polity without threatening the polity itself [@feaver_civil-military_1996; @feaver_civil-military_1999].
 Feaver describes this "problematique" as "vexing" and a great "paradox" in that the military must be "strong enough to prevail in war," yet submissive to authority such that it cannot and will not "destroy or prey on the society it is intended to protect" [@feaver_civil-military_1999, p.
 214].
-Within the context of the US, the study of civil-military relations often focuses on the relationship between military and government elites, or what Risa Brooks calls "leadership at the state's apex" [@brooks_risa_integrating_2019, p.
+Moreover, the study of civil-military relations often focuses on the relationship between military and government elites, or at what Risa Brooks calls "leadership at the state's apex" [@brooks_risa_integrating_2019, p.
 380].
-Indeed, it is at this level that vital decisions are made, including to go to war, how to prosecute war, how to achieve and implement effective strategy, how to prepare for the next conflict, and how large the military and its budget should be.
-A number of scholars have offered solutions to Feaver's "problematique," but none has been more lasting and consequential than the late Samuel Huntington, whose 1957 seminal work, _The Soldier and the State: The Theory and Politics of Civil-Military Relations_, often serves as the point of departure for many scholarly works on civil-military relations.
-Although there have been many critiques of Huntington over the years [for example, see @janowitz_professional_1960; @finer_man_1962; @cohen_supreme_2003; @feaver_armed_2003; @brooks_paradoxes_2020], it is absolutely essential to understand Huntington's concept of normative civil-military relations, and what lies at the heart of it - a theory of civilian control of the military that maximizes the concept of what Huntington called "military professionalism" [@huntington_soldier_1957].
+Indeed, it is at this level of interaction between government and military leaders that vital decisions are made, including decisions whether to go to war, how to achieve and implement effective strategy, how to prepare for the next conflict, and how large the military and its budget should be.
+
+A number of scholars have offered solutions to the"problematique" raised by Feaver, but none has been more lasting and consequential than the late Samuel Huntington, whose 1957 seminal work, _The Soldier and the State: The Theory and Politics of Civil-Military Relations_, addresses this fundamental problem.
+Indeed, even more than 60 years after he wrote his book, Huntington's work often serves as the point of departure for many scholarly works on topics concerning civil-military relations.
+Although there have been many critiques of Huntington over the years [for example, see @janowitz_professional_1960; @finer_man_1962; @cohen_supreme_2003; @feaver_armed_2003; @brooks_paradoxes_2020], it is absolutely essential to understand Huntington's normative conception of effective civil-military relations, and what lies at the heart of it - a theory of civilian control of the military that maximizes a concept that Huntington called "military professionalism" [@huntington_soldier_1957].
 
@@
 Rather, the development and implementation of strategy, especially during a time of war, requires an "iterative" dialogue between military leaders and civilians to ensure that effective policies are adopted [@rapp_civil-military_2015, 20].
-
 Others argue that Huntington's conception of military professionalism is too narrow to be of practical use in the contemporary period in that such a model actually enables, rather than constrains, inappropriate political activity by military officers [@brooks_paradoxes_2020].
@@
 Whether military leaders recognize it or not, Brooks argues, military leaders comprise a profession that is inherently political, and such a reality ought to be acknowledged and embraced [@brooks_paradoxes_2020].
-Regardless of whether and the degree to which one agrees or disagrees with Huntington's concept of military professionalism, a significant number of scholars note that a significant issue involves attempting to define where the line between appropriate political expression and inappropriate partisan activity should be drawn [@owens_military_2015; @golby_beyond_2015; @brooks_paradoxes_2020].
+
+Regardless of whether and the degree to which one agrees or disagrees with Huntington's concept of military professionalism, a significant number of scholars note that an enduring issue involves attempting to define where the line between appropriate political expression and inappropriate partisan activity should be drawn [@owens_military_2015; @golby_beyond_2015; @brooks_paradoxes_2020].
 It is a difficult task for many reasons because, as Mackubin Thomas Owens points out, civil-military relations in the United States can be viewed as a "bargain" [@owens_us_2011] between the American people, the military, and the government that has to account for constantly-changing "political, social, technological, or geopolitical changes" [@owens_is_2016].
@@
 
-## The Dependent Variable: Military Intervention in Politics and Military "Role Beliefs"
+# Framing the DV: Military Intervention in Politics and Military "Role Beliefs"
 
-Given the normative model espoused by Huntington and the critiques his model has been subjected to, it is important to understand what frameworks exist for understanding why the military would intervene in politics.
-Numerous scholars, mainly writing from a comparative perspective (although not exclusively), have enriched our understanding of what possibly motivates the military to interject itself into the political sphere.
-To be clear, these scholars are focusing mainly on active-duty military organizations, and not on the specific block of retired military officers, as I plan to examine, but the motivations and mechanisms for intervention that they put forward enrich our understanding of the phenomenon of political intervention by the military.
+Given Huntington's normative model and the critiques his model has been subjected to, it is important to understand what frameworks exist for understanding the debates surrounding why the military would or should intervene in politics, and to what degree.
+Here again, the literature is vast.
+Scholars from the field of comparative politics have generally tackled the question of why militaries intervene in political matters, likely as a result of the variation that is available in country-specific or cross-national studies.
+It is vital to briefly understand a couple of these important frameworks.
 
 The British political scientist, S.E.
-Finer, argues that the military's intervention in politics is a function of the opportunity and desire to do so [@finer_man_1962].
-Finer identifies several factors which should normatively inhibit the military from intervening in politics.
-These factors are military professionalism, the principle of civilian supremacy, and fear that the military would fail to remain a cohesive fighting force should it fracture internally upon intervening in political matters [@finer_man_1962].
-Finer critiques the concept of Huntingtonian military professionalism somewhat harshly, arguing that although military professionalism should act to focus the military on itself and thus not on political activities, those who claim to be professional might and have used the concept of professionalism as an excuse to not obey civilian authorities [@finer_man_1962].
-Finer is clear that the principle of civilian control - not military professionalism - is the true and only bulwark that can keep the military out of politics.
+Finer, argues that military intervention in politics is ultimately a function of both the opportunity and desire to do so [@finer_man_1962].
+Finer first identifies several factors that should inhibit the military from intervening in politics, including the concept of military professionalism (as espoused by Huntington), strict adherence to the principle of civilian supremacy, and the military's fear that intervention in politics would lead to it fracturing internally, and thus failing to remain a cohesive fighting force [@finer_man_1962].
+Finer delivers a rather stinging critique of Huntingtonian military professionalism, and argues that although professionalism should work to focus the military on itself and thus, away from politics, the danger of the concept of military professionalism is that it can also be abused, particularly by military leaders who adopt the view that they serve the larger concept of "the state" rather than a particular civilian government [@finer_man_1962, 22-23].
+Finer argues that military leaders throughout history, including von Seeckt and Gröner under the German Weimar Republic, and MacArthur under Truman in the United States, inappropriately intervened in politics precisely because their conceptions of "professionalism" demanded that they did so [@finer_man_1962], a critique of military professionalism that Risa Brooks shares [@brooks_paradoxes_2020].
+Finer's whole point is that the principle of civilian control - not military professionalism - is and ought to serve as the only bulwark to keep the military out of politics.
 
-Describing the various motives that militaries might have for intervening in politics, Finer goes on to describe at least three.
-He first lists the motive of the "manifest destiny of the soldiers," in which the military acts as "saviors" of the country [@finer_man_1962, 28-29].
+Finer further describes at least three motives that militaries might have for intervening in politics.
+He argues that soldiers might adopt a specific form of "manifest destiny," or a worldview in which the military acts as "saviors" of the country [@finer_man_1962, 28-29].
 This motive, Finer claims, arises because of the confluence of the military's unique mission, sense of power, and claim of being "outside party politics" [@finer_man_1962, 29].
 A second motive that militaries might possess which disposes them to intervene in politics is what Finer calls the "unique identification with the national interest" [@finer_man_1962, 30].
-Acting from such a motive, Finer conceptualizes the military acting as "custodians" of the "national interest" [@finer_man_1962, 30].
-Finer acknowledges that "custodianship" can take on different forms in different contexts, but points out that some militaries might view "custodianship" as a "duty to arbitrate or veto...to exercise it when some convulsion or decision of the civil autorities seems to them [the military] to threaten what they think are the permanent interests of the nation" [@finer_man_1962, 31].
-The final motive Finer spends the most time on, calling it "the sectional interest," which refers to a more narrow subset of interests than the "national interest," and includes a wide ranging set of interests based on social class, regional affiliation, personal career motivations, or organizational interests [@finer_man_1962, 34-53].
+Finer argues that militaries which act from such a motive might view themselves as the "custodians" of the "national interest" [@finer_man_1962, 30], a role that can take different forms, including a "duty to arbitrate or veto...to exercise it when some convulsion or decision of the civil authorities seems to them [the military] to threaten what they think are the permanent interests of the nation" [@finer_man_1962, 31].
+Finally, Finer posits that militaries intervene in politics because of "the sectional interest," a wide-ranging set of reasons based on social class, regional affiliation, personal career motivations, or organizational interests [@finer_man_1962, 34-53].
 
-Ultimately, Finer develops a typology of four levels, or degress of, military intervention.
-The first he calls "influence," and argues that at this level of intervention, the military makes appeals "to reason or emotion" in an attempt to persuade the civilian government [@finer_man_1962, 77].
-Finer adds that intervening at this level is, in his mind at least, not problematic and entirely "consistent with the supremacy of the civil power" [@finer_man_1962, 77].
-The next level of intervention is "pressures or blackmail," describing efforts by the military to "convince the civil power by threat of some sanction" [@finer_man_1962, 77].
-Finer adds that military intervention via this level is "on the margin," referring to his belief that while some cases may be unproblematic, many other forms, such as the "threat to stage a coup," would be normatively wrong [@finer_man_1962, 77].
-Finer points out a third level of intervention, that of "displacement," referring to the "removal" of government officials by the military as a result of "violence or the threat of violence" [@finer_man_1962, 78].
-The fourth level, which is basically displacement on a wider scale, Finer calls "supplantment," which "sweeps away the civilian regime and establishes the military in its place" [@finer_man_1962, 78].
-Finer goes on to explain the level of intervention as a function of the level of "political culture" that is prevalent in a state and society  [do I need more here?
-See Finer, pp.
-79 and following]
+Finer further classifies military intervention as one of four ascending levels.
+At the bottom or first level, which he calls "influence," Finer views the military as making appeals "to reason or emotion" in an attempt to persuade the civilian government to adopt a certain course of action [@finer_man_1962, 77].
+Finer adds that intervention by the military at this level is not problematic and entirely "consistent with the supremacy of the civil power" [@finer_man_1962, 77].
+The next level of intervention Finer labels "pressures or blackmail," describing efforts by the military to "convince the civil power by threat of some sanction" [@finer_man_1962, 77].
+Finer adds that military intervention via this level is "on the margin," referring to his belief that while some cases may be unproblematic, many other forms, such as the "threat to stage a coup," are normatively wrong [@finer_man_1962, 77].
+The third level of intervention, "displacement," refers to the "removal" of government officials by the military as a result of "violence or the threat of violence," and the fourth level, "supplantment," occurs when the military "sweeps away the civilian regime and establishes the military in its place" [@finer_man_1962, 78].
 
-Another helpful conceptualization of military intervention in politics is the idea of "role beliefs," a concept articulated by the political scientist, J.
-Samuel Fitch, in his book, _The Armed Forces and Democracy in Latin America._
+Another helpful conceptualization of military intervention in politics is the concept of military "role beliefs," introduced by J.
+Samuel Fitch in his book, _The Armed Forces and Democracy in Latin America._ Through rich interview data conducted in the 1990s, Fitch traces how leaders within the militaries of Argentina and Ecuador adopted one of five distinct "role beliefs," which describe the perceptions held by these officers of their normative role in politics [@fitch_armed_1998, 61].
+The first type of role belief is "democratic professionalism," and involves both "a subjective commitment to democracy as the preferred political system," and a "belief in a professional ethic that differentiates and separates military and political affairs" [@fitch_armed_1998, 66].
+A second "role belief" is "classical professionalism," and differs from the first in that while it "argues for the subordination of the military, in both political and policy terms, to the constitutionally designated civilian authorities", it does not espouse a "normative commitment to a democratic regime" [@fitch_armed_1998, 66-67].
+A third type of "role belief" that military offices can adopt is that of a "power factor," or a belief that within the regime, "the military should have some degree of influence in deference to its vital societal role" [@fitch_armed_1998, 67].
+The fourth type of "role belief" militaries can adopt is one of "conditional separation" to civilian leaders, and refers to the notion that while the military should stay out of politics "in normal circumstances", the advent of "crisis situations" can and does warrant the military to "safeguard the highest interests of the nation," and, as Fitch adds, "even if that entails overthrowing the incumbent government and assuming power temporarily" [@fitch_armed_1998, 68].
+A final possible role belief militaries can adopt is a "tutelary" model in which the civilian government grants the military both a "high degree of autonomy" and the ability to become a regular and "institutional voice in policy matters" [@fitch_armed_1998, 68-69].
+According to this "role belief," the military believes itself to be the final "guardian of national security," and thus that it has "the right" to participate in a wide variety of policy debates that impact the state [@fitch_armed_1998, 69].
+Fitch argues that the particular "role beliefs" militaries adopt are a function of three primary factors: the lessons militaries draw from previous domestic intervention in politics, a wide-ranging assessment of current domestic indicators (the economy, the effectiveness of institutions, etc.), and the prevailing  international context [@fitch_armed_1998].
 
-A third, and equally rich, study of the degrees of military intervention in politics is available from Brian Taylor's book, _Politics and the Russian Army: Civil-Military Relations, 1689-2000.
+Within the US context, scholars have long been interested in the specific political activities that the military engages in.
+These studies have tended to examine specific political behaviors such as voting [@dempsey_our_2010], social media use [@urben_like_2017], political endorsements [@griffiths_not_2019], and attitudes towards political engagement [@urben_party_2013; @urben_wearing_2014], but they have not attempted to systematically explore these changes over time.
+Additionally, while Risa Brooks has created a helpful typology of political activities the US military engages in, ranging from "public appeals" that engage the public, to "shoulder tapping" focused exclusively on Congress [@brooks_militaries_2009], her typology also does not explain variation in these activities over time.
+While the domestic environment in Latin America is vastly different than that in the United States, Fitch's work underscores and points to the importance of civilian and military leaders committing themselves to the supremacy of both democratic institutions and democratic rule [@fitch_armed_1998].
+And while there has arguably been more variation in Latin American militaries intervening in politics, I believe that there is sufficient evidence that indicates a subtle but important shift in the US military's beliefs regarding its central role and purpose has occurred, and that a systematic examination of this shift is warranted.
+
+# Framing the IVs: Political Polarization and Military Prestige
 
@@
 
-The question of if and how political polarization drives changes in the professional identity of the military spans several related literatures, most notable of which are American Politics and International Relations.
-Broadly speaking, this project is focused on how one aspect of domestic politics, political polarization, impacts a specific but important aspect of international relations, civil-military relations, which itself spans a wide and diverse literature.
-Because I intend to focus my work on the military as an institution, my dissertation will also rely on literature from the fields of sociology, organizational psychology, and ethics.
-I will frame my research question within each of these fields briefly, and to define important concepts as required, before addressing why this question matters and should be studied in depth.
+Within the American Politics literature, there is broad recognition that political polarization levels in contemporary America are higher now than at any point since the end of the American Civil War [@hare_polarization_2014], and there are multiple explanations for what has caused this trend, ranging from a growing divide over matters pertaining to race and religion [@abramowitz_great_2018], the size and role of government [@sides_two_2015], and disagreements over specific policy issues [@mann_its_2012].
+Despite the wide-ranging scholarship within the field of American Politics that explores the facets of political polarization, virtually all scholars agree that contemporary levels of political polarization pose myriad problems.
 
-Within the American Politics literature, there is broad recognition that political polarization levels in contemporary America are higher now than at any point since the end of the American Civil War [@hare_polarization_2014], but there are multiple explanations for what has caused this trend, ranging from a growing divide over matters pertaining to race and religion [@abramowitz_great_2018], the size and role of government [@sides_two_2015], and disagreements over specific policy issues [@mann_its_2012].
-Despite the wide-ranging scholarship exploring the many facets of political polarization, virtually all scholars agree that contemporary levels of political polarization in the United States are historically abnormal and pose myriad problems.
-
-One of the major concerns regarding political polarization is its on impact on American foreign policy in a broad sense, and specifically, on America's ability to lead and impact the world.
+Scholars of international relations have, more recently, examined the pernicious impacts of polarization on foreign policy.
 Kenneth Schultz, for instance, underscored four challenges that have emerged as a result of America's polarized domestic politics, including the ability to obtain bipartisan support for "risky undertakings," the ability to learn the "lessons of foreign policy failure," the inability to "make long term commitments to allies and adversaries" as a result of "dramatic policy swings from one administration to another," and the "vulnerability" of America's electoral system to foreign third party adversaries [@schultz_perils_2017].
@@
 
-## Military Professions and Organizational Loyalty
+But a more central concern of mine is understanding how and why political polarization might impact how the military views its central role in American politics.
+I argue that there are at least three ways through which increasing polarization can shape the apolitical and non-partisan ethos of the military, and thus, can alter how the military views its role in politics.
+The first is that there is increasing difficulty for military leaders to remain politically neutral, non-partisan, or otherwise apolitical - a hallmark of Huntingtonian military professionalism.
+Fewer and fewer issue areas and policy matters are outside of the "spell" of polarization, owing to a phenomenon schoalrs of American politis refer to as "conflict extension" [@layman_party_2002].
+In previous eras of polarization, whereas a single issue may have served as the "battleground" for partisan conflict, this is no longer true today, as research has consistently shown that voters who identify as liberals are more liberal across multiple issue categories and that the same trend holds for conservatives [@abramowitz_great_2018].
+As a result, whereas the prospect for cooperation on issues that were not politically polarized remained high in the past, today the prospect for cooperation in general is far lower [@layman_activists_2010].
 
-As a subject or concept, loyalty features prominently in the international relations and civil-military relations literatures, often in a comparative setting and in the context of coups [@marcum_overthrowing_2016; @johnson_squeaky_2018], civil wars [@mclauchlin_loyalty_2018], and the difficulties that multi-ethnic states face in maintaining a professional military [@peled_question_1998].
-This literature deals with a host of important issues that stem from problems associated with ethnic identification and loyalty, either to a regime, cause, or leader.
-Though the US context is very different from the contexts usually explored in this literature, its insights may help me identify related dynamics and/or mechanisms that are unfolding in the US: how are changes in partisan or ideological identification (instead of ethnic) as a result of political polarization impacting conceptions of loyalty within the US military, if at all?
-To even begin answering this question, one has to first define what loyalty is and how it works.
-For this fundamental task, the fields of psychology, ethics, management, and sociology are helpful.
+A second challenge posed by rising polarization stems from the drastic increase of negative feelings toward those on the other side of the political spectrum, a phenomenon known as affective polarization or "negative partisanship"[@abramowitz_great_2018].
+Just as contemporary polarization is marked by citizens who hold increasingly strong and negative feelings towards those of a different political stripe, there is likely a tendency for members of the military institution to harbor similar feelings and attitudes.
+One might object, pointing to the training and professional development military leaders receive that emphasizes the importance of upholding an apolitical and non-partisan ethic, but even a cursory review of the evidence (members of the military increasingly making presidential endorsements [@griffiths_not_2019] and posting political messages on social media [@urben_like_2017] ) suggests this is not the case.
 
-Loyalty has been conceptualized as both a virtue [@coleman_problems_2009] and as a moral emotion [@connor_military_2019], the main difference between the two being that the former is a normative aspirational trait, while the latter is a primed behavioral response.
-For the purposes of my dissertation, these differences are not consequential, because both assert that loyalty is something that is ultimately felt by an individual and subsequently reflected in action, such as, in the military context, "the desire to protect each other from harm and to defend each other from attack" [@coleman_problems_2009, 109].
-Oldenquist, a philosopher, further describes loyalty as a devotion or attachment that an individual has to someone or something specific such that one refers to the object of loyalty as "mine" and is inclined to feel "pride when it prospers, shame when it declines, and anger or indignation when it is harmed" [@oldenquist_loyalties_1982, 175].
-Loyalty also generates a willingness in the individual to assume "obligations" [@oldenquist_loyalties_1982, p.
-175] on behalf of the object of one's loyalty, and these obligations can take multiple forms, such as a willingness of the individual to endure hardships and make sacrifices [@fielder_organizational_1992].
-Scholars agree that individuals can hold multiple loyalties simultaneously, and recognize that this often can and does cause problems [@coleman_problems_2009].
-For organizations, and by extension, professions, the concept of loyalty is also very important.
-Conner et al.
-state, for example, that "having loyal members is vital to perpetuating the group" [@connor_military_2019].
+A third, and I contend the most impactful, challenge posed by polarization to the normative conceptions of the military's role in politics stems from the fact that our polarized climate is marked by the belief among individuals that core issues - such as those that involve attitudes about race, religion, and multiple layers of not just individual, but also national identity - are at stake in the current polarized political climate.
+As the rise of "identity politics" illustrates, some of the most polarizing issues today are those that are related to an individual's identity, such as race, gender, and immigration [@sides_identity_2018].
+And at the national level, the same phenomenon holds: as Vice President Joe Biden and President Donald Trump wind down the final days of their presidential campaigns, both claim that "the soul of the nation" is at stake [@dias_biden_2020].
 
-We also know that the military services in the United States place a premium on loyalty as a trait and concept.
-Indeed, of the different military services, the US Army in particular has extensively developed, in a normative sense, a framework entitled "The Army Ethic," in which loyalty plays a significant role within the Army profession [@us_department_of_the_army_adp_2019].
-This ethic, which will be described in detail in later parts of this dissertation, is formally laid out and described in the Army's doctrinal publication entitled, _Army Leadership and the Profession, ADP 6-22_, and holds that as "stewards of the Army profession" and as "committed professionals," Army leaders "embrace and uphold the Army Values and standards of the profession" [@us_department_of_the_army_adp_2019].
-There are a total of seven "Army Values," which together spell the acronym, l-d-r-s-h-i-p, which is short for the word "leadership," and the first of these values, loyalty, is followed by six others: duty, respect, selfless service, honor, integrity, and personal courage [@us_department_of_the_army_adp_2019, p.
-1-9.].
+## Military Prestige
 
-ADP 6-22 further states that the Army Values "embody the practical application of the Army Ethic.
-They encompass the enduring moral principles, beliefs, and laws that guide Army professionals in accomplishing the mission as well as their conduct in all aspects of life" [@us_department_of_the_army_adp_2019, p.
-1-12].
-The manual defines loyalty as, "loyalty: bear true faith and allegiance to the Constitution of the United States, the Army, your unit and other Soldiers" [@us_department_of_the_army_adp_2019].
-The virtue of loyalty is likewise extolled in a host of codes and mantras that have played a central part of the American military since its inception.
-In the middle of the West Point Cadet Prayer is the eloquently written plea, "make us to choose the harder right instead of the easier wrong" [@wheat_west_1918], indicating that to be loyal to what is right is not only preferable but often a difficult endeavor.
-Even Samuel Huntington recognized that in order for his model of civilian control to be effective, the military profession had to organize itself as a "hierarchy of obedience," which required "loyalty and obedience" to stand out as "the highest military virtues" [@huntington_soldier_1957, 73].
+With respect to the military institution itself, one of the most massive and influential changes that has occurred over the past roughly seventy years is the increase in overall prestige, credibility, and influence of the military institution.
+While key legislative acts passed since World War Two, to include the National Security Act of 1947 and the Goldwater-Nichols Act of 1946, may have fixed important gaps in America's ability to meet its national security requirements [@inboden_national_2017], they also cemented the military as one of if not the dominant institution and actor within American politics.
+Over the same time period, while there have been spikes as a result of exogenous events such as wars and scandals that have involved the military, public opinion towards the military has risen sharpy.
+Consider that in the late 1970s, as a likely result of public opinion over the Vietnam War, 54% of Americans indicated that they trusted the military "a great deal or quite a lot," whereas in 2019, 73% of Americas expressed the same sentiments [@gallup_polls_confidence_2007].
+Unsurprisingly given these developments, compared to other government institutions, today the American public trusts the military to a remarkably high degree.
+According to a recent Pew Research Center poll, 80% of the American public expects the military to act in the best interests of the American public, compared to business leaders (45%), the media (40%), and elected officials (25%) [@johnson_trust_2018].
 
-Loyalty is further important in a military setting in a broad and strategic sense, especially when one considers that the US military, both by statute and by normative standards, serves as institution with multiple, simultaneous loyalties.
-The Constitution of the United States, to which military officers and soldiers alike ascribe allegiance when they take their respective oaths of commissioning and enlistment, contrives the military as an institution possessing multiple loyalties: to the President as "the Commander in Chief of the Army and the Navy of the United States" [@noauthor_us_1787], but also to Congress in that they likewise hold unique powers and responsibilities pertaining to the Armed Forces [@noauthor_us_1787-1].
-The tensions that inherently exist in American's constitutional framework with respect to national security are explored in a robust scholarly literature [for instance, see @howell_while_2007; @milner_sailing_2015].
+While military leaders are rightfully cognizant of the benefits of holding the public's trust, a number of scholars and practitioners have noted that the trend of "military worship" is not only unhealthy [@fallows_tragedy_2014], but likely harmful for democracy [@friend_military_2018].
+Esteeming the military more than it ought can perpetuate, according to Mara Karlin and Alice Friend, a number of harmful myths, including the idea that military service is somehow superior than other, non-military forms of service [@friend_military_2018].
+Over the long haul, a military that is perceived to be more prestigious than its normative civilian superiors can, according to numerous scholars, be problematic for civilian control of the miltiary [@friend_military_2018].
 
-# Where My Puzzle Fits and Why It Matters
+Scholars have also shown that when elites possess a high degree of credibility, their ability to influence and shape public opinion is likewise high.
+In his influential study, _The Nature and Origins of Mass Opinion_, John Zaller argues that formations in mass public opinion are the result of three interacting components: "variation in the information carried in elite discourse, individual differences in attention to this information, and individual differences in political values and other predispositions" [@zaller_nature_1992].
+Zaller argues that throughout modern America, most citizens, as non-elites, must rely heavily on the messages that come from elites regarding various topics and issue areas [@zaller_nature_1992].
+Zaller further argues that the messages that elites send is necessarily "stereotyped" in some way: that is, the messages that elites send, in order to be understood by regular citizens who are part of the public, must be "short, simple, and highly thematic" [@zaller_nature_1992].
+In developing this argument, Zaller is not claiming that regular citizens are uninformed or careless people who are incapable of processing complex information, but rather acknowledging that most citizens care most about a world that is not "beyond their personal experience" [@zaller_nature_1992].
+Thus, his main point is simply that elites play a significant role in forming and swaying public opinion.
+
+Therefore, as the public's trust in the military has grown in the years since Vietnam, so has the military's potential to impact public opinion.
+This is especially true of military elites, who hold substantial positions of authority.
+An increase in credibility of the institution may not indicate that the military or particular military elites will actually engage in behavior to deliberately impact public opinion, but rather that the potential to do so has increased, as the American people tend to listen to those whom they trust and revere.
+
+# My Puzzle, Where it Fits, and Why It Matters
+
+The central question I wish to explore in my dissertation is whether and why the US military, or portions of it, have become increasingly politically active, and whether such a shift portends a shift in the military's own "role beliefs" concerning how it relates to American government and society.
+As such, my dissertation will center on fundamental and longstanding debates about the concepts of authority, loyalty, and the proper role of the military with respect to government.
+What is different from previous works in the field is that I wish to explore changes in these concepts, particularly within the US military, over time and in historical perspective.
+
+While there have been several vital studies, most of them incorporating original survey data, exploring the beliefs that military members have regarding forms of political behavior [@dempsey_our_2010; @urben_party_2013; @urben_wearing_2014], these studies have not sought to explain these changes.
+Similarly, there have also been a number of recent and important studies regarding the impacts of political polarization on the credibility society places in the military as an instition [@robinson_michael_danger_2018; @robinson_who_2019], as well as a host of impressionistic pieces noting the potentially detrimental impacts of polarization and politicization on the military institution [@feaver_we_2016; @barno_how_2016; @golby_jim_keep_2020; @reid_retired_2020 ].
+However, these studies also have not empirically explored the impact of increasing polarization and institutioncal prestige on the military as an institution over time.
+I wish to close that gap, as I think doing so will help not only scholars, but also practitioners, to include members of the military, understand the answers to some very important questions.
+These questions include:
+
+* Among whom within the military institution is a shift in normative "role beliefs" occurring, and why?
+* When did such a shift start, and why? 
+* If there is a shift in the military's normative "role beliefs," how will that impact the military's organizational culture?
+* Are some of the most polarized domestic issues compatible with historically normative conceptions of military professionalism, such as the principles of civilian supremacy and a non-partisan ethos?
+why or why not?
+* What are the key challenges or obstacles to the military regaining and sustaining an effective non-partisan ethos?
+
+Specifically, I desire to explore the relationship between polarizing domestic politics and institutional credibility on the professional identity and character of the military.
+In the post World War Two era in particular, the military's institutional power, social prestige, and level of credibility has generally grown, making the military a recognized and respected actor within American politics.
+And, especially since the 1970s, political polarization has also increased dramatically, to the point that America is now as polarized as it was at the end of the American Civil-War [@hare_polarization_2014].
+I believe that these two trends, occurring at least in a broad sense simultaneously, have significantly impacted the military's beliefs about its proper role in politics, and that a systematic investigation of this relationship should be undertaken.
+
+---
+### Figure this out
 
@@
 
-One potential area of work that remains under-developed, both theoretically and empirically, is an exploration of the relationship between polarizing domestic politics and the professional identity and character of the military.
+I argue that this third challenge is especially relevant for the contemporary military, and specifically, with respect to how members of the military exercise the concept of loyalty.
+As an increasing number of what individuals believe to be core issues are raised and debated in the partisan arena, individuals face decisions of how to prioritize the persons or objects of their loyalty and further, what actions to take on the basis of loyalty to these persons or objects.
+I further theorize (though this has to be tested, of course) that while all individuals will likely face some version of this challenge of loyalty in a polarizing age, the challenges will be more acute for military leaders.
+Not only is the de-facto setting in which military leaders operate marked by multiple and often conflicting loyalties due to the structure of and statutes set forth in the US Constitution, but also, the potential consequences are higher for military leaders than average citizens should they fail to discharge the duties that correspond to their roles as professional officers.
+These consequences include failure to achieve tangible national security goals, such as loss on the battlefield, or less tangible but still important outcomes such as maintaining trust with the American public.
+
+As important as these consequences are, they do not speak to another set of consequences individual military leaders also may face in a polarizing age.
+Indeed, in addition to consequences to the country that may occur should the military fail to discharge its duties to safeguard America, military leaders also potential consequences if confronted with a situation in which they have to execute orders which are technically legal, but morally borderline at best.
+Military leaders are still "moral agents" who "remain responsible for the decisions and actions they take" [@dubik_just_2016, 160].
+But the considerations of what constitutes moral right and wrong to begin with may be changing as polarization worsens [@graham_liberals_2009].
+
+Therefore, I argue that if polarization continues to heighten the identities of groups that people associate with, as well as come closer to the boundaries of how individuals think about moral right and wrong, there is an increased likelihood of military leaders having to choose between conflicting loyalties.
+What happens if and when the objects of a military leader's multiple loyalties not merely disagree, but become stubbornly divided ideologically?
+Can such an outcome exist?
+Is there empirical proof that such a trend is happening?
+If so, what are the normative implications, both for the military as a profession, and for the nation's security, and are there actions that can be taken to halt such a trend?
+
+ In the past several years for example, there has been a concerted effort by the services to identify and punish senior leaders who engage in misconduct, to demonstrate to Congress that senior military leaders do not need assistance to handle sexual assault, and to otherwise prove that it can be trusted to accomplish its missions [@us_army_repairing_2014].
+Former Chief of Staff of the Army, Retired General Ray Odierno, told fellow Army generals that "of great importance within the Army Profession is the duty of each of us to be a steward of our profession and to maintain its essential and hard won trust with society [Ray Odierno in @us_army_repairing_2014].
+
+---
+
  I am specifically interested in the question of if and how political polarization impacts the concept of loyalty in the military.
@@
 Fewer scholars, if any, are exploring the potential role of the military in politicizing itself, and the mechanisms through which these developments occur.
+
+## Military Professions and Organizational Loyalty
+
+As a subject or concept, loyalty features prominently in the international relations and civil-military relations literatures, often in a comparative setting and in the context of coups [@marcum_overthrowing_2016; @johnson_squeaky_2018], civil wars [@mclauchlin_loyalty_2018], and the difficulties that multi-ethnic states face in maintaining a professional military [@peled_question_1998].
+This literature deals with a host of important issues that stem from problems associated with ethnic identification and loyalty, either to a regime, cause, or leader.
+Though the US context is very different from the contexts usually explored in this literature, its insights may help me identify related dynamics and/or mechanisms that are unfolding in the US: how are changes in partisan or ideological identification (instead of ethnic) as a result of political polarization impacting conceptions of loyalty within the US military, if at all?
+To even begin answering this question, one has to first define what loyalty is and how it works.
+For this fundamental task, the fields of psychology, ethics, management, and sociology are helpful.
+
+Loyalty has been conceptualized as both a virtue [@coleman_problems_2009] and as a moral emotion [@connor_military_2019], the main difference between the two being that the former is a normative aspirational trait, while the latter is a primed behavioral response.
+For the purposes of my dissertation, these differences are not consequential, because both assert that loyalty is something that is ultimately felt by an individual and subsequently reflected in action, such as, in the military context, "the desire to protect each other from harm and to defend each other from attack" [@coleman_problems_2009, 109].
+Oldenquist, a philosopher, further describes loyalty as a devotion or attachment that an individual has to someone or something specific such that one refers to the object of loyalty as "mine" and is inclined to feel "pride when it prospers, shame when it declines, and anger or indignation when it is harmed" [@oldenquist_loyalties_1982, 175].
+Loyalty also generates a willingness in the individual to assume "obligations" [@oldenquist_loyalties_1982, p.
+175] on behalf of the object of one's loyalty, and these obligations can take multiple forms, such as a willingness of the individual to endure hardships and make sacrifices [@fielder_organizational_1992].
+Scholars agree that individuals can hold multiple loyalties simultaneously, and recognize that this often can and does cause problems [@coleman_problems_2009].
+For organizations, and by extension, professions, the concept of loyalty is also very important.
+Conner et al.
+state, for example, that "having loyal members is vital to perpetuating the group" [@connor_military_2019].
+
+We also know that the military services in the United States place a premium on loyalty as a trait and concept.
+Indeed, of the different military services, the US Army in particular has extensively developed, in a normative sense, a framework entitled "The Army Ethic," in which loyalty plays a significant role within the Army profession [@us_department_of_the_army_adp_2019].
+This ethic, which will be described in detail in later parts of this dissertation, is formally laid out and described in the Army's doctrinal publication entitled, _Army Leadership and the Profession, ADP 6-22_, and holds that as "stewards of the Army profession" and as "committed professionals," Army leaders "embrace and uphold the Army Values and standards of the profession" [@us_department_of_the_army_adp_2019].
+There are a total of seven "Army Values," which together spell the acronym, l-d-r-s-h-i-p, which is short for the word "leadership," and the first of these values, loyalty, is followed by six others: duty, respect, selfless service, honor, integrity, and personal courage [@us_department_of_the_army_adp_2019, p.
+1-9.].
+
+ADP 6-22 further states that the Army Values "embody the practical application of the Army Ethic.
+They encompass the enduring moral principles, beliefs, and laws that guide Army professionals in accomplishing the mission as well as their conduct in all aspects of life" [@us_department_of_the_army_adp_2019, p.
+1-12].
+The manual defines loyalty as, "loyalty: bear true faith and allegiance to the Constitution of the United States, the Army, your unit and other Soldiers" [@us_department_of_the_army_adp_2019].
+The virtue of loyalty is likewise extolled in a host of codes and mantras that have played a central part of the American military since its inception.
+In the middle of the West Point Cadet Prayer is the eloquently written plea, "make us to choose the harder right instead of the easier wrong" [@wheat_west_1918], indicating that to be loyal to what is right is not only preferable but often a difficult endeavor.
+Even Samuel Huntington recognized that in order for his model of civilian control to be effective, the military profession had to organize itself as a "hierarchy of obedience," which required "loyalty and obedience" to stand out as "the highest military virtues" [@huntington_soldier_1957, 73].
+
+Loyalty is further important in a military setting in a broad and strategic sense, especially when one considers that the US military, both by statute and by normative standards, serves as institution with multiple, simultaneous loyalties.
+The Constitution of the United States, to which military officers and soldiers alike ascribe allegiance when they take their respective oaths of commissioning and enlistment, contrives the military as an institution possessing multiple loyalties: to the President as "the Commander in Chief of the Army and the Navy of the United States" [@noauthor_us_1787], but also to Congress in that they likewise hold unique powers and responsibilities pertaining to the Armed Forces [@noauthor_us_1787-1].
+The tensions that inherently exist in American's constitutional framework with respect to national security are explored in a robust scholarly literature [for instance, see @howell_while_2007; @milner_sailing_2015].
 
@@
 \doublespacing
-
-The first way in which the apolitical and non-partisan ethos of the military is challenged as a result of increasing political polarization is that there is increasing difficulty for individuals to remain politically neutral, non-partisan, or otherwise apolitical - a hallmark of military professionalism.
-This happens because as "conflict extension" [@layman_party_2002] occurs, an increasing number of issues are brought into partisan battlegrounds, and thus, there is simply less politically-neutral territory available.
-Thus, an actor such as the military, who may wish to remain non-partisan and out of politics to the extent that it can, may simply be less able to do so.
-
-A second challenge military leaders and the institution faces in a polarized climate is from the rise of "negative partisanship"[@abramowitz_great_2018] that dominates the polarized American political landscape.
-As citizens hold increasingly strong and negative feelings towards those who side differently on a host of issues, there may be a tendency for members of the military institution to harbor similar feelings and attitudes.
-Even if the impact of this feature on the military and its leaders, due to training and professional development that emphasizes the importance of upholding apolitical and non-partisan behavioral norms is somewhat muted, the point is that members of the military institution are likely experiencing the same phenomenon - even if it is to a lesser degree - as the country's citizens.
-One important question about "negative partisanship" in a military setting, though, is whether service members hold strong negative feelings toward only civilians who are on the opposite side of the ideological spectrum, or whether such feelings also permeate within the ranks.
-
-A third, and I contend the most impactful, challenge that normative military professionalism faces stems from the belief by individuals that core issues - such as those that involve attitudes about race, religion, and multiple layers of identity - are at stake in the current polarized political climate.
-"Core" issues are difficult to define precisely, but the main point is that these issues often involve perceptions of identity [] and moral right and wrong, and not just simple policy preferences.
-Moreover, some of the most polarizing issues are often associated with issues related to an individual's identity, such as race, gender, immigration, as the the rise of "identity politics" in America illustrates this point [@sides_identity_2018].
-
-I argue that this third challenge is especially relevant for the contemporary military, and specifically, with respect to how members of the military exercise the concept of loyalty.
-As an increasing number of what individuals believe to be core issues are raised and debated in the partisan arena, individuals face decisions of how to prioritize the persons or objects of their loyalty and further, what actions to take on the basis of loyalty to these persons or objects.
-I further theorize (though this has to be tested, of course) that while all individuals will likely face some version of this challenge of loyalty in a polarizing age, the challenges will be more acute for military leaders.
-Not only is the de-facto setting in which military leaders operate marked by multiple and often conflicting loyalties due to the structure of and statutes set forth in the US Constitution, but also, the potential consequences are higher for military leaders than average citizens should they fail to discharge the duties that correspond to their roles as professional officers.
-These consequences include failure to achieve tangible national security goals, such as loss on the battlefield, or less tangible but still important outcomes such as maintaining trust with the American public.
-
-As important as these consequences are, they do not speak to another set of consequences individual military leaders also may face in a polarizing age.
-Indeed, in addition to consequences to the country that may occur should the military fail to discharge its duties to safeguard America, military leaders also potential consequences if confronted with a situation in which they have to execute orders which are technically legal, but morally borderline at best.
-Military leaders are still "moral agents" who "remain responsible for the decisions and actions they take" [@dubik_just_2016, 160].
-But the considerations of what constitutes moral right and wrong to begin with may be changing as polarization worsens [@graham_liberals_2009].
-
-Therefore, I argue that if polarization continues to heighten the identities of groups that people associate with, as well as come closer to the boundaries of how individuals think about moral right and wrong, there is an increased likelihood of military leaders having to choose between conflicting loyalties.
-What happens if and when the objects of a military leader's multiple loyalties not merely disagree, but become stubbornly divided ideologically?
-Can such an outcome exist?
-Is there empirical proof that such a trend is happening?
-If so, what are the normative implications, both for the military as a profession, and for the nation's security, and are there actions that can be taken to halt such a trend?
 

### 2020-12-04 code/Submissionv4.Rmd
commit

@@
 glance(mod4)
+
+#model 5 - now check for robustness using aff_polar as the polarization variable
+mod5 <- glm(leg_crit_part ~ Inst_Cred_Gallup + aff_polar + AuthSex + AuthRank,
+    family=binomial(link="logit"),
+    data=polaradjdataset
+    )
+
+summary(mod5)
+
+mod6 <- glm(crit_part ~ Inst_Cred_Gallup + aff_polar + AuthSex + AuthRank,
+    family=binomial(link="logit"),
+    data=polaradjdataset
+    )
+
+summary(mod6)
 ```
@@
 ```{r regression table, include=TRUE, echo=FALSE, results="asis"}
-stargazer(mod3, mod4, 
+stargazer(mod3, mod4, mod5, mod6,
           title="Table 3: Logistic Regression Results", 
-          column.labels = c('Legislative Activity, Criticize, or Partisan','Criticize or Partisan'),
-          covariate.labels = c('Institutional Credibility', 'Congressional Polarization', 'Male', 'Mixed Authorship', 'Unknown Authorship', 'Rank'),
+          column.labels = c('Model 3','Model 4','Model 5', 'Model 6'),
+          covariate.labels = c('Institutional Credibility', 'Congressional Polarization', 'Affective Polarization', 'Male', 'Mixed Authorship', 'Unknown Authorship', 'Rank'),
           align=TRUE)

### 2020-12-04 code/Tahkv1.tex
commit

\PassOptionsToPackage{unicode=true}{hyperref} % options for packages loaded elsewhere
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames*,x11names*}{xcolor}
%
\documentclass[12pt,]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provides euro and other symbols
\else % if luatex or xelatex
  \usepackage{unicode-math}
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage[]{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage{xcolor}
\usepackage{hyperref}
\hypersetup{
            pdftitle={When the Pen Becomes a Sword: Retired US Military Officers and Op-Ed Use in Contemporary America},
            pdfauthor={Peter M. Erickson},
            colorlinks=true,
            linkcolor=magenta,
            filecolor=Maroon,
            citecolor=black,
            urlcolor=blue,
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage[margin = 1.15in]{geometry}
\usepackage{longtable,booktabs}
% Fix footnotes in tables (requires footnote package)
\IfFileExists{footnote.sty}{\usepackage{footnote}\makesavenoteenv{longtable}}{}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

% set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

% there are a few ways to include latex options.
% 1: as a preamble file (like this one)
% 2: directly into the header-includes YAML variable (global)
% 3: output templates. For more info on templates,
%    see <https://bookdown.org/yihui/rmarkdown/rticles-templates.html>
%    and <https://bookdown.org/yihui/rmarkdown/rticles-usage.html>

% These packages control the fonts used in PDF output.
% The standard LaTeX engine (PDFLaTeX) has a limited selection of fonts.
% See <https://www.tug.org/FontCatalogue/>

\usepackage{libertine}                 % serif font
\usepackage{libertinust1math}          % matching math font
\usepackage[scaled = 0.95, varqu]{zi4} % teletype (monospace/code) font

% The XeLaTeX engine gives you more control to specify fonts
%   at the expense of having poorer typesetting principles
%   (but xelatex typesetting is getting better)

% this package enables decimal-aligned table columns
\usepackage{dcolumn}

% the default PDF template for R Markdown loads a lot of convenient packages
% so you don't actually have to mess with them all that much.
\usepackage{setspace}
\usepackage{multirow}
\usepackage{indentfirst}
\usepackage{caption}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\usepackage[style=authoryear,]{biblatex}
\addbibresource{researchpaper-ps904.bib}

\title{When the Pen Becomes a Sword: Retired US Military Officers and Op-Ed Use in Contemporary America}
\author{Peter M. Erickson}
\date{December 03, 2020}

\begin{document}
\maketitle
\begin{abstract}
\emph{This chapter explores variation in the professionalism of opinion pieces authored by retired US military officers and published in major US newspapers since 1979. Drawing on an original data set almost 400 opinion-editorial pieces published by retired military officers in The Wall Street Journal, The Washington Post, The New York Times, The Los Angeles Times, and the USA Today since 1979, the paper argues there has been an increase in published opinion pieces which fail to uphold the concept of military professionalism as defined by historical military norms and standards. What explains this variation? The author argues that increases in political polarization and the prestige of the military account for the increase in the number of op-eds which challenge traditional conceptions and norms of military professionalism. The paper concludes by arguing that the concept of ``politicization of the military'' --- something that previous civil-military relations scholars have identified and described --- must be widened to account not only for efforts by politicians to use the military for political advantages, but also for the phenomenon of retired military officers engaging the public to achieve political ends. The author argues that such a development, if uncorrected, will cause significant damage to the military as a profession.}\\
\end{abstract}

\doublespacing

\captionsetup[table]{labelformat=empty}

\newpage

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

In late May and early June 2020, racial tension and concerns about police brutality gripped the American public following the death of George Floyd, an unarmed black man, by police in Minneapolis. In some cities and communities across the nation, peaceful protests were followed by periods of rioting and looting. On June 1, after giving remarks at the Rose Garden, President Donald Trump walked to St.~John's Church after security forces visibly displaced throngs of protesters in neighboring Lafayette Square. Flanked by members of his cabinet, to include the Secretary of Defense, and with the Chairman of the Joint Chiefs of Staff not far behind, the President held up a Bible at the church, stared into the cameras of the media, and, several minutes later, returned to the White House. In the days and weeks that followed, a civil-military relations crisis ensued \autocite{feaver_military_2020}. The Secretary of Defense, Mark Esper, publicly disagreed with the President's threat to invoke the Insurrection Act of 1807 to use active-duty forces to quell domestic riots \autocite{esper_secretary_2020}, and General Mark Milley, the Chairman of the Joint Chiefs of Staff, later told senior military officers graduating from the National Defense University that he ``should not have been there'' (accompanying the President on the walk to St.~John's Church), as his presence ``created a perception of the military involved in domestic politics'' \autocite{milley_official_2020}.

Furthermore, several retired military officers weighed in not only on the Trump administration's handling of specific events near the White House, but also on the broader issues of race in America.\footnote{These officers included several retired four star generals and admirals}. However, there was marked variation in the tone and tenor of the remarks made by these officers: although each expressed deep concern about the prospect of active-duty forces confronting peaceful protesters, and thus, violating protections held by the First Amendment of the US Constitution \autocites{brooks_dismay_2020}{dempsey_former_2020}, some officers went much further in their criticisms, leveling direct insults of the President, including his fitness to lead \autocite{goldberg_james_2020}, his administration's broader policies on race in America \autocite{allen_moment_2020}, and his general leadership abilities \autocite{mullen_i_2020}. Moreover, the comments made by these retired senior leaders renewed interest, at least for a while, in the topic of retired military officer speaking out on current events - a topic that civil-military relations scholars have debated \autocites{kagan_let_2006}{brooks_let_2020}, but one that has yet, in the author's opinion, to be studied empirically and systematically. In a domestic political age in which polarization is high, the topic of retired military officer speech is important for a variety of reasons. Not only does the speech of retired military officers carry the potential to impact public opinion, but the very act of retired senior leaders speaking out may also compel those who are still in the military to follow suit. More importantly, the topics that military leaders choose to address in the public square, and the tone of their remarks, is reflective of the conceptualizations these officers have regarding their role in society and ultimately, their purpose and professional identity as military officers. Indeed, in 2020, retired as well as active duty military officers grabbed the pen and argued in support of a variety of topics that clearly pushed the traditional boundaries of professional conduct by military officers by openly arguing for a variety of topics related to President Donald Trump, including including race in America \autocite{askew_anti-racist_2020} and even removing the President from the White House \autocite{nagl__2020}.

This chapter examines variation in the publication of op-ed pieces by retired military officers since 1979. There are two main arguments of this chapter. The first argument is that the variation in the number, topic addressed, and tone of opinion-editorial publications published by retired military officers indicates a shift over the past 40 years in the way a number of retired military officers think about their normative roles as members of the military institution and how they interact with the public in America. The second argument is that the driver of the variation in the number, topic, and tone of op-eds authored by retired military officers is mainly the increase in political polarization that has occurred in America since the 1970s. While there are other factors that influence also responsible, such as the increase in institutional prestige of the military over the same time period, the empirical story reveals that political polarization holds more explanatory power than other factors in explaining the variation of retired military officer speech in America.

This paper proceeds in three parts. First, the paper connects the concept of military ``role beliefs'' to retired military officer speech. Second, the paper explores and analyzes an original data set of nearly 390 opinion-editorial pieces published in major US newspapers by retired military officers since the late 1970s. This analysis ultimately demonstrates that although the vast majority of op-eds published by retired military officers maintain professional standards, there is clear evidence that retired military officers are using the pen both to argue about topics that fall outside of the realm of traditional national security topics, and to violate the principle of civilian control of the military when doing so. The analysis also shows that these trends have worsened as the American political climate has become increasingly polarized over the past several decades. Third, thechapter concludes by exploring the implications of these results on the military as an institution, and specifically, on its ability to remain professional in a highly polarized setting that will likely endure.

\singlespace

\hypertarget{how-the-military-views-itself}{%
\section{How the Military Views Itself}\label{how-the-military-views-itself}}

\doublespace

\hypertarget{military-role-beliefs}{%
\subsection{Military Role Beliefs}\label{military-role-beliefs}}

The concept of military ``role beliefs'' was introduced by J. Samuel Fitch in his book, \emph{The Armed Forces and Democracy in Latin America.} In the book, Fitch describes military ``role beliefs'' as the perceptions held by military officers regarding their normative role in politics \autocite[61]{fitch_armed_1998}. Though Fitch's focus is on militaries in Latin America, it is possible to use the concept of ``role beliefs'' to develop a normative starting point applicable to military officers in the United States.

Two foundations for developing a starting point for military ``role beliefs'' in the United States are the tandem norms of civilian control of the military, and the idea of the military acting as a non-partisan institution. Though such norms have been episodically challenged throughout American history in a variety of ways and during specific instances, these norms have nonetheless been both foundational and aspirational, particularly in the post World War Two period \autocite{weigley_american_2001}. The norm of civilian control of the military stems directly from the US Constitution, which expressly grants control of the military (as well as specific responsibilities associated with maintaining control) to the President, Congress, and the individual states \autocite{noauthor_us_1787-2}. Unlike the norm of civilian control, the norm of the military functioning as a non-partisan institution is not expressly spelled out in the Constitution, but rather the product of the American military's collective experiences and traditions in fighting and prevailing in the nation's wars and conflicts.

Further, a number of scholars have underscored the importance of these two norms, and the relationship between these norms and the military's ability to prevail during times of war. In his influential book, \emph{The Soldier and the State: The Theory and Politics of Civil-Military Relations}, Samuel Huntington argues in favor of maintaining separate operating spheres between the political leaders of a democracy and the military as an institution in order to ensure civilian control of the military by rendering the military ``politically sterile and neutral,'' or non-partisan \autocite[84]{huntington_soldier_1957}. Huntington goes on to argue that only by such an arrangement, which he called ``objective control of the military,'' will the military be capable of defeating significant security threats while also remaining apolitical \autocite[83-84]{huntington_soldier_1957}.

Having established the foundational norms of civilian control of the military and the military operating as an apolitical institution, the next step is to extrapolate how these norms might impact the public speech of military officers. This requires balancing these norms with the fact that military officers are nonetheless servants of the Republic who have accumulated knowledge and experiences over the course of sometimes lengthy careers (in the case of retired military officers, at least 20 and in some cases, upwards of 40 or even more years of experience), and thus, in many cases are qualified to weigh in on circumstances and situations that impact the nation. For instance, in the aftermath of the ``Revolt of the Generals'' in 2006, in which several retired military officers publicly criticized and called for the removal of Defense Secretary Donald Rumsfeld, Frederick Kagan argued in support of retired military officers having the freedom to weigh in policy debates, particularly those involving an active armed conflict, stating, ``there is no danger to the republic in a handful of retired generals speaking their minds. There is great danger in making vital decisions about an on-going armed struggle without hearing the views of all available experts'' \autocite{kagan_let_2006}. Kagan's remarks lend support to the notion that in democracies in particular, transparent and flowing information is a precious resource that helps leaders make decisions and, more importantly, the public to gain awareness of the state's actions {[}need cite{]}.

Nonetheless, there have been both formal and informal boundaries placed on various forms of political activities by military officers. Formally speaking, Department of Defense (DoD) Directive 1344.10 contains a ``standing list of political behavior do's and don'ts for members of the active duty military'' \autocite[1]{urben_wearing_2014}. This list includes authorizing members of the military to engage in activities such as voting and joining a partisan or non-partisan club and to attend this club's meetings when not in uniform, and prohibits members of the military from activities such as participating in partisan fundraising, or advocating for or against a partisan cause or figure on radio or television \autocite{department_of_defense_political_2008}. Informally, a number of recent Chairmen of the Joint Chiefs of Staff have likewise urged members of the military to refrain from engaging in overly political activities, particularly in election years, when such activity is more likely to negatively reflect on the military. For example, prior to the 2016 Presidential Election, then-Chairman of the Joint Chiefs of Staff General Joseph Dunford reminded the military that the basis upon which it is able to maintain the public's trust is it's ability to remain ``apolitical and neutral'' \autocite{dunford_upholding_2016}.

Based on these formal and informal prohibitions on the political activities conducted by military officers, it is possible to conclude that visible behaviors that are overtly partisan are to be avoided at all costs, and are inappropriate for military officers to engage in. It is also possible to extrapolate that the norm of civilian control of the military results in a corresponding prohibition on members of the military criticizing political leaders to the point that civilian control of the military is jeopardized or fails to be upheld. Where this line should be drawn in practice is not clear, however, because as Kagan has pointed out, criticisms can and often do point out blind spots in a plan or otherwise reveal shortcomings that may not otherwise have been identified. However, there is a point at which a criticism levied at a political leader by a military officer goes beyond criticism. Finally, it is possible to conclude that activities which build on the appropriate expertise and knowledge of the military officer are appropriate, provided that these activities do not violate the stipulations previously mentioned, i.e.~a non-partisan military, and civilian control of the military.

\hypertarget{opinion-editorial-publication-as-a-political-activity}{%
\subsection{Opinion-Editorial Publication as a Political Activity?}\label{opinion-editorial-publication-as-a-political-activity}}

We can now focus on the actual behavior of the military authoring op-ed pieces. Should these considered a form of political activity? I argue that it depends on what message the op-ed actually sends or is intended to send. An op-ed that strictly points out the pitfalls of the nation's combat strategy written by an experienced military officer may not be inappropriate, whereas an op-ed written by a retired military officer that argues in favor of a specific Presidential candidate is, at least by the standards put forth in this paper, entirely inappropriate. Therefore, at least two questions should be asked when examining an opinion piece authored by a retired military officer: first, does the piece violate either the norm of civilian control of the military or the military as an apolitical institution, and second, does the op-ed address an appropriate topic that a military officer is qualified to address?

Beyond these two questions, however, opinion-editorials are important because they serve as an informative barometer of how the military institution sees itself, and particularly, its role with respect to the American public. Members of the active military do not frequently write op-eds, but there have been notable exceptions over the years, perhaps none so notable as several written by then-Chairman of the Joint Chiefs of Staff General Colin Powell in the early 1990s. For instance, Powell presented the public with his view regarding the potential use of American military force in several developing ``hot spots'' across the world, including the Balkans, Somalia, and other places, during an October 8, 1992 piece entitled, ``Why Generals Get Nervous.'' Published in the \emph{New York Times}, Powell wrote in part, ``decisive means and results are always to be preferred, even if they are not always possible. So you bet I get nervous when so-called experts suggest that all we need is a little surgical bombing or a limited attack. When the desired result isn't obtained, a new set of experts then comes forward with talk of a little esclalation. History has not been kind to this approach'' \autocite{powell_opinion_1992}.

Powell clearly felt compelled to inform the public of how he, as the military's top officer, viewed the potential use of American military force abroad during the prevailing international context at the time. What is less known, however, is that Powell was also writing against a domestic backdrop in which there was an election coming up in less than a month that pitted the incumbent, George W. Bush, against Democratic challenger Bill Clinton. The two candidates differed with respect to foreign policy on a number of issues, and Bill Clinton was not highly respected by the military establishment, a fact that would impact the military's relationship with Clinton during his early years in office {[}need cite{]}. Whether it was proper for Powell to write such an op-ed truly depends: one could argue that the nation's top general was simply informing the public based on his years of individual experience serving in the military, in which case, from a democratic transparency standpoint, the op-ed well served the Republic; alternately, one could also argue that by writing during an election year, Powell's remarks brazenly staked the claims of the military ahead of a potential Presidential transition (which turned out to be the case), and in the process, limited the President's options for potentially using force abroad during his early days in office. The first explanation argues that from the standpoint of civil-military relations, there was no problem with Powell's opinion-editorial, while the second argues that the op-ed was nothing but problematic.

\hypertarget{retired-military-officers}{%
\subsection{Retired Military Officers}\label{retired-military-officers}}

One might argue that the logic employed thus far is seemingly plausible for members of the active military, but that the cautionary tone and reservations regarding certain types of political activity should not apply for members of the retired military community. And indeed, there are obvious distinctions between those who are on active duty and those who have retired from the service. These distinctions are made clear in fact in portions of DoD 1344.10 \autocite{department_of_defense_political_2008}. But the fact of the matter is that, especially for retired military officers who achieved general or flag officer rank, these retired officers sometimes carry outsized influence and impact when they engage in activities that approach on the political. During the Presidential Campaign of 2016, Retired Army General Martin Dempsey published an opinion-editorial piece in \emph{The Washington Post} which criticized retired Army Major General Mike Flynn and Marine Corps General John Allen for participating in the respective Republican and Democratic conventions. Dempsey wrote, ``As generals, they have an obligation to uphold our apolitical traditions. They have just made the task of their successors -- who continue to serve in uniform and are accountable for our security - more complicated. It was a mistake for them to participate as they did. It was a mistake for our presidential candidates to ask them to do so'' \autocite{dempsey_military_2016}.

Of course, speaking at a political convention on behalf of a Presidential candidate is, if not formally, at least informally the textbook definition of a political activity. Not every opinion-editorial piece supports a political candidate's agenda, or disparages the political views of another. Thus, it is important to remember that although an op-ed can be written in such a way that it is a politically inappropriate form of political behavior, it need not be. It is precisely on this point that the norms of civilian control of the military, the military as an apolitical institution, and retired members of the military expressing views that are in accord with the knowledge and skills gained during their time in uniform are appropriate.

Moreover, because multistar generals and admirals in the contemporary military possess in many cases upwards of 30 and sometimes 40 years of service, it is entirely reasonable to conclude that these high-ranking officers in particular have been privy to decisions and experiences occurring at or near the highest levels of government. Having interacted with Presidents, members of the cabinet, members of Congress, and foreign dignitaries, the experiences of retired general and flag officers are not limited to only tactical matters. Though they possess an abundance of expertise in many tactical matters that pertain to warfighting, they are also familiar with issues ranging from how wars should be fought, to future security threats looming on the horizon, to fixing the institutional culture of large organizations such as the military services, and much more.

Nonetheless, they are not experts in everything, nor should they try to be. Even General of the Army Douglas MacArthur, one of several men who ever wore the five star rank, and a hero of World War Two who later clashed with President Truman regarding American strategic goals in Asia, recognized that the military had focus areas, even if he did not personally exhibit these limitations while on active duty. Addressing the Corps of Cadets at West Point in 1962, the then-82 year old retired General told the Corps:

\singlespacing

\begin{quote}
\textbf{Yours is the profession of arms, the will to win, the sure knowledge that in war there is no substitute for victory, that if you lose, the Nation will be destroyed, that the very obsession of your public service must be Duty, Honor, Country. Others will debate the controversial issues, national and international, which divide men's minds. But serene, calm, aloof, you stand as the Nation's war guardians, as its lifeguards from the raging tides of international conflict, as its gladiators in the arena of battle. For a century and a half you have defended, guarded and protected its hallowed traditions of liberty and freedom, of right and justice. Let civilian voices argue the merits or demerits of our processes of government. Whether our strength is being sapped by deficit financing indulged in too long, by federal paternalism grown too mighty, by power groups grown too arrogant, by politics grown too corrupt, by crime grown too rampant, by morals grown too low, by taxes grown too high, by extremists grown too violent; whether our personal liberties are as firm and complete as they should be. These great national problems are not for your professional participation or military solution. Your guidepost stands out like a tenfold beacon in the night: Duty, Honor, Country} \autocite{macarthur_duty_1962}.
\end{quote}

\doublespacing

By systematically examining the op-eds written by retired military officers in major US publications over time, it is possible, I contend, to draw inferences about how the concept of a ``role belief'' - conceptions of one's normative role in politics - has changed among retired US military officers (if at all) \autocite{fitch_armed_1998}. Op-eds are an interesting and appropriate medium to examine if differentiation in ``role beliefs'' is an object of interest. The first reason why this is the case is because as previously stated, the act of crafting and publishing an op-ed does not necessarily constitute a political activity, though depending on content, they can be. A second and releated reason is because unlike other forms of engaging with the press, such as giving interviews on television, crafting an op-ed requires the author to deliberately prepare and submit a viewpoint for publication. In major newspapers, this is often a competitive process. It is not uncommon to see opinion pieces in prominent newspapers published by heads of state, members of congress, governmental figures, and leaders of industry and academia. A deliberate and carefully-crafted message further underscores the degree to which the author believes the message he or she is sending. It is difficult to ``misspeak'' during or to be ``misquoted'' from an opinion-editorial. The author generally agrees with the message he or she is sending.

When examining op-eds authored by retired military officers over time to detect potential shifts in normative role beliefs, the key area of focus must be the degree to which retired military officers have upheld vice strayed from the norms of civilian control and non-partisanship of the military, as well as the degree to which retired officers have addressed topics that are appropriately rooted in relevant military expertise. If there are patterns of violating these norms that can be detected, it may indicate that the role beliefs of certain members of the retired military community have likewise changed. On the other hand, if there has been little deviation over time from these foundational norms, it is possible to conclude that military retirees view their normative roles in politics (role beliefs) in similar ways as previous generations of retired military officers.

\hypertarget{hypotheses}{%
\subsection{Hypotheses}\label{hypotheses}}

The hypotheses I wish to examine, then, are as follows:

\emph{H1: When political polarization is high, the military (or certain military figures) will increasingly criticize civilian leadership (indicating a drop in the principle of civilian control).}

\emph{H2: When political polarization is low, the military (or certain military figures) will publicly weigh in on traditional national security debates (and not on domestic political concerns or issues outside of relevant military expertise).}

\emph{H3: During periods of high prestige and credibility, the frequency of military involvement in quasi-political activities (such as writing op-eds) will increase relative to periods of low prestige and credibility.}

\emph{H4: When both political polarization and high prestige are present, the military will be increasingly vocal and at higher frequency about certain domestic ideological or partisan causes than when these factors are not present.}

\emph{H5: During periods of active ground combat, military elites will likely comment on the war, but with less frequency on other domestic issues.}

Some caution must be taken up front so as to not overpromise and under-deliver concerning what the data is actually capable of presenting. The first caution to consider is that opinion-editorial publications by nature reflect the demand of the press to publish the viewpoints of retired military officers. Moreover, because op-eds are generally provocative and at least to some degree designed to advocate for or against a certain course of action, it is possible that op-eds may generally present somewhat contentious material. Additionally, this means that under periods of harmonious civil-military relations, there is either less of a demand from the press to publish the viewpoints of retired military officers, or less of a supply side effect for retired military officers to write about topics in the first place. Certain exogenous events, such as wars, will impact whether retired military officers speak out. A second cautionary note is that over the course of several decades, underlying societal changes - such as political polarization - have occurred that likely impact either or both the willingess of the press to publish the viewpoints of military officers, or the desire of military officers to write about certain topics in the first place. It is important that these societal-level effects be identified and, if possible, controlled for.

\hypertarget{methodological-approach-and-the-data}{%
\section{Methodological Approach and the Data}\label{methodological-approach-and-the-data}}

The methodological approach used in this analysis consists of two steps. The first is to analyze the topics which retired military officers have been writing about, and the second is to determine, regardless of the topic addressed in a publication, whether the op-ed has violated norms of non-partisanship or civilian control of the military. The data consists of opinion-editorial and letters to the editor publications authored by retired military officers and published in major US newspapers from 1979-2020. These newspapers consist of \emph{The Wall Street Journal} (WSJ), \emph{The New York Times} (NYT), \emph{The Washington Post} (WaPo), \emph{The Los Angeles Times}, and \emph{The USA Today}.

I include both opinion-editorials and letters to the editor because both types of publications display a viewpoint that required a deliberate effort by the officer to craft and submit to the particular newspaper. I do not include published interviews with retired military officers, nor reprinted editorials that were first published before 1979, such as those reprinted for the benefit of readers sometime after 1979. To assemble the data, I use a series of key-word searches, such as ``retired military officer,'' ``retired from the Army, Navy, Marine Corps, Air Force,'' ``former Chairman of the Joint Chiefs of Staff,'' ``General,'' ``Admiral,'' and so forth. I conduct these key word searches in several different databases, including Factiva, Lexus-Uni, and ProQuest Historical Newspapers. \footnote{For exact replication of the database search parameters, please see the Appendix.} I include in my search publications that were made available both in print and online format, though I did not double-count: if the same op-ed was published both in print and online and published by the same author, I count the article only one time.

There are both advantages and disadvantages to confining my data to select US newspapers, as I have. The primary advantage is that these sources offer longevity across a time period of nearly forty-two years; in doing so, they provide the researcher with an opportunity to detect patterns across several decades. The primary disadvantage is that by limiting myself to only the five newspapers I examine, I am not including a large number of modern social media and blogging sites, contemporary webpages, and other digital offerings that showcase the opinions of retired military officers.

Although social media has taken over as the primary means in which a majority of Americans receive their news today \autocite{martin_how_2018}, the methodological approach adopted in this analysis allows for an examination of the same sources over time. More importantly, even after accounting for the advent of online editions of newspapers in the late 1990s and early 2000s, the overall circulation of newspapers (print and online editions combined) is still down today compared to earlier years. As a means of rough comparison, estimated total daily newspaper circulation in the US in 2000 was 62 million for weekday and 54 million for weekend editions, respectively, while in 2018, the same figures were 29 million and 31 million \autocite{barthel_leading_2019}.

In some instances, especially where the same retired military officers authored multiple publications, the officer did not always refer to him or herself as a retired military officer. This occurred in several instances in which the officer was, ostensibly, writing from the perspective of someone other than a retired military officer, such as that of a political appointee to a high level governmental position (after retirement from the military, Army General Barry McCaffrey, for instance, served as President Clinton's Director of the Office of National Drug Control Policy). Nonethless, because these officers had previously identified themselves as retired military officers in earlier publications, and thus, because members of the public might reasonably recognize the author as a prominent retired military figure, I include the subsequent publications in the data, and simply control for whether the author identified him or herself as a retired military officer in a specific publication.

\hypertarget{summary-statistics}{%
\subsection{Summary Statistics}\label{summary-statistics}}

The data set includes a total of 390 observations, penned by 216 distinct different authors or author teams. Eight of the observations are written by anonymous authors, but it is unknown whether the same unknown author penned several of these eight observations. Included in the data are cases in which at least one retired military officer co-authored a piece with one or more civilian authors. \footnote{There are several co-authored pieces. For example, where General (Retired) David Petraeus wrote one publication and another one was written by General (Retired) David Petraeus and Michael O'Hanlon, I counted these as two different authors for coding purposes.} There are also several retired military officers who penned several observations. The most prolific of these officers includes Retired Army Lieutenant General William Odom (22 publications in total), Retired Air Force General Michael Hayden (20 publications in total), and Retired Marine Lieutenant Colonel and Former National Security Advisor Robert McFarlane (18 publications).

Figure One, below, plots the yearly count of publications authored by retired military officers across all of the major newspapers. The smoothed trendline indicates that the average yearly count of publications rose steadily from 1979 until the early 2000s, at which time, a decline occurred until approximately 2008. From 2008 until the present day, the average yearly count of publications authored by retired military officers in the major US newspapers examined in this study rose heavily during the Obama administration years before briefly falling, but then resuming the high count during the Trump administration.

It is noteworthy, from a macro perspective, to consider all of the impacts on the US military over the past four decades and more. Numerous substantive events, debates, and changes have occurred in this time period for retired military officers to weigh in on. In terms of external threats and wars, this period includes the final years of the Cold War against the Soviet Union (through 1991), the Gulf War in Iraq (1990-1991), operations in the Balkans (primarily 1993-1995), and the post-9/11 wars in Iraq (2003-present day) and Afghanistan (2001-present day), as well as a host of smaller but important military operations in Grenada (1983), Panama (1989), Somalia (1992-1993), and Haiti (1994-1995). The military has also undergone significant social and institutional changes in the time period examined. In the late 1970s, the military had ended the draft only years before (1973) following the conclusion of the Vietnam War. The debates over ``Don't Ask, Don't Tell'' (instituted in 1994 and repealed in 2011), as well as policies regarding women in the combat arms (2015) are also included in the data, as are numerous annual discussions regarding the size of military budgets.

\begin{verbatim}
## `geom_smooth()` using formula 'y ~ x'
\end{verbatim}

\begin{center}\includegraphics{Tahkv1_files/figure-latex/figure one-1} \end{center}

In addition to collecting the yearly count of publications, biographical information of the author is collected, including the service in which the author served, the rank achieved by the author during his or her time in service, and the sex of the author. Summary statistics for these variables are included in Table 1, below.

Insert table of yearly count, rank, and sex, as well as a panel facet graph of Figure 1, above, by source.

Describe the other variables and coding scheme.

A bar graph by decade across all sources collectively of \# of publication by topic count. Or do this later.

Of note, when assembling the data, some authors are explicitly identified in some of their publications as retired military officers, but not in others.\footnote{This was true of several individuals, including Alexander Haig, a retired Army General who later served as Secretary of State under Ronald Reagan; Robert McFarlane, a retired Marine Corps Lieutenant Colonel who later served as the National Security Adviser to President Reagan, Andrew Bacevich, a retired Army Colonel who then became a prominent historian and author; and Michael Hayden, a retired Air Force General who served as director of the National Security Agency and the Central Intelligence Agency}. I conduct a separate search for all opinion pieces penned by these authors, regardless of whether they were identified as a retired military officer or not, concluding that it is best do so for coding purposes.

Attempting to reconcile the fact that the military as an institution prides itself on being a non-partisan and apolitical institution, that newspaper circulations (even after accounting for digital offerings) are substantially down over the past 40 years, and that the US political environment is increasingly polarized, I adopt the two following hypotheses:

\emph{\(H_1\): The majority of op-eds published by retired military officers over the time period examine will be apolitical and non-partisan in tone.}

\emph{\(H_2\): Given that polarization has become increasingly pronounced in the United States over the time period examined, there will be an increase in recent years in the number of op-eds written about topics that are not relevant to military expertise, or that reveal inappropriate partisan sentiments. Such a trend may indicate that increasing political polarization is impacting military professionalism among a small but important part of the military institution, retired military officers.}

\hypertarget{first-layer-analysis-topics-addressed}{%
\section{First Layer Analysis: Topics Addressed}\label{first-layer-analysis-topics-addressed}}

In the first level of analysis, I code each observation into one of several binary categories that corresponds to the overall topic or subject of the publication. The goal is to create a coding scheme that captures a concise number of topical descriptions that are still sufficiently broad enough to reveal variation in the topic addressed by retired military authors. I develop a total of six topical variables, and each observation takes on the value of 1 if the publication corresponds to the general definition of the variable, and 0 otherwise.

These topical variables are: war fighting and operations, if the piece advocates for or denounces a particular course of action directly related to operational or war fighting decisions of a clear military nature; troop support and institutional culture, if the publication generally highlights the work and sacrifice of members of the armed forces, defends the work of specific members of the Armed Forces, or makes a call for action to improve a service's institutional culture; strategic advice, if the piece is generally written to offer some sort of strategic advice for the nation, to include the wisdom of pursuing particular alliances or agreements, or adopting a specific weapons platform or military budget; military social policy, if the piece advocates for or denounces a particular social policy that \emph{directly influences} the armed forces (Examples include: debates over keeping or removing the ``Don't Ask, Don't Tell'' policy, opening up combat roles to women, the role of Transgender members of the armed forces); domestic policy, if the piece advocates for or denounces a position on a policy that \emph{does not directly} influence the armed forces in its day to day activities (Examples include: support for the second amendment, national immigration policy, and whether confederate statues should be taken down or left up); and political figures, if the piece advocates for or denounces a particular position held by a political administration, or \emph{if the main point of the piece} is to explicitly defend or attack a political figure by name (Examples: endorsing a Presidential candidate for office, or for attacking a President or member of Congress for a decision that was made).

Up front, separating opinion-editorial publications into one of several distinct categories is an imperfect process. It requires a thoroughly objective reading of the publication as possible. To aid the coding process, I identify the thesis statement of each opinion piece, which in most cases is identifiable in either the lead or concluding paragraph of the opinion piece. In most cases, identifying and re-reading the thesis of each op-ed makes it possible to reasonably classify the publication by topic. Two additional points, however, are worth emphasizing here.

First, opinion pieces are intended to argue for something, and thus are, to an extent, designed to be at least mildly provocative. With this in mind, I did not necessarily code pieces that called for a political figure or institution to take a specific action into the category of \emph{figures by name}. Again, the identification of the publication's thesis statement aided significantly in identifying the main thrust of the argument, and thus, in classifying the argument by topic. Second, there were some instances in which I had to choose between placing a publication into one of two categories. However, in instances of choosing between two categories, the choice to classify between categories was usually between categories that were reasonably close. For example, there were a few instances of choosing to classify a publication between the war fighting and operations category and the strategic advice category, but no instances of choosing between the war fighting and operations category and the domestic policy category. Table 3 reveals the breakdown of all examined publications by article type.

\begin{table}

\caption{\label{tab:table3}Publications by Subject}
\centering
\fontsize{10}{12}\selectfont
\begin{tabu} to \linewidth {>{\raggedright}X>{\raggedleft}X}
\hline
\textbf{Topic} & \textbf{n}\\
\hline
\cellcolor{gray!6}{BdgtWpnsTrps} & \cellcolor{gray!6}{37}\\
\hline
CivMilBalance & 19\\
\hline
\cellcolor{gray!6}{DomesticPolicy} & \cellcolor{gray!6}{38}\\
\hline
ForeignPolicy & 112\\
\hline
\cellcolor{gray!6}{ServiceCulture} & \cellcolor{gray!6}{13}\\
\hline
SocialPolicy & 9\\
\hline
\cellcolor{gray!6}{Support} & \cellcolor{gray!6}{44}\\
\hline
Warfighting & 118\\
\hline
\end{tabu}
\end{table}

An overwhelming majority (293 of 344, or 85.2\%) of the total number of opinion pieces in the data set fall into the categories of war fighting and operations, troop support and institutional culture, and strategic advice. Moreover, I conclude that these observations pose no problem for norms of military professionalism, as they are written on topics that I contend are clearly related to one's relevant military experience. The remaining 51 publications (14.8\% of the observations) are coded as falling under the other three categories of topics: social policy directly related to the military, domestic policy, and political figures by name. These observations require a second and more-thorough examination to determine whether the content of these publications indicate a drop in the norm of military professionalism.

\singlespace

\hypertarget{the-second-layer---failing-to-uphold-professionalism}{%
\section{The Second Layer - Failing to Uphold Professionalism?}\label{the-second-layer---failing-to-uphold-professionalism}}

\doublespace

Conclusively determining whether an observation failed to uphold military professionalism is once again an imperfect and somewhat subjective evaluation. However, the work of academic theorists and decades of military norms make clear that at a minimum, military professionalism emphasizes behavior by military officers that is relevant to one's military expertise, as well as absent of partisan expression or preferences. Recall that Retired Army General Douglas MacArthur cautioned the cadets at West Point in 1962 about involving themselves in matters that do not require a military solution (in MacArthur's words, of ``politics grown to corrupt,'' ``morals grown too low,'' and ``whether our personal liberties are as firm and complete as they should be'') \autocite{macarthur_duty_1962}, and that more recently, in 2016, Retired Army General Martin Dempsey lamented the participation of retired generals in the political process by endorsing political candidates and appearing at the conventions of political parties:

\singlespace

\begin{quote}
\textbf{The American people should not wonder where their military leaders draw the line between military advice and political preference. And our nation's soldiers, sailors, airmen and Marines should not wonder about the political leanings and motivations of their leaders. As generals, they have an obligation to uphold our apolitical traditions. They have just made the task of their successors -- who continue to serve in uniform and are accountable for our security - more complicated} \autocite{dempsey_military_2016}.
\end{quote}

\doublespace

Relying on these two central ideals of military professionalism, relevant military expertise and non-partisan/apolitical expression, allows the researcher to ask two important questions: first, is the author addressing a topic related and applicable to the officer's military expertise, and second, is the author making an argument without clearly conveying or expressing political or partisan preferences? If the answer to either of these two questions is ``no,'' then the publication might violate the concept of military professionalism. I therefore develop two new variables, \emph{questionable professionalism} and \emph{justification}. The first variable is binary and codes as `1' the observations which may violate the concept of military professionalism, and the second variable briefly captures the basis for a possible violation, i.e.~whether it is a violation of topic or of expression of political sentiment. The results of this analysis on the remaining 51 observations are displayed in Table 4, below. Several observations are worth noting.

\singlespace
\begingroup\fontsize{10}{12}\selectfont

\begin{longtable}[t]{>{\raggedright\arraybackslash}p{9.5em}|>{\raggedright\arraybackslash}p{15em}|>{\raggedright\arraybackslash}p{2.5em}|>{\raggedleft\arraybackslash}p{2.5em}|>{\raggedright\arraybackslash}p{5em}|>{\raggedleft\arraybackslash}p{6em}|r|r}
\caption{\label{tab:table4}Op-Eds authored by Retired Military Officers that Strongly Criticized Civilian Leaders, or were Expressly Partisan, 1979-2020}\\
\hline
\textbf{Author} & \textbf{Title} & \textbf{Source} & \textbf{PubYr} & \textbf{Topic} & \textbf{Crit\_Insub} & \textbf{Endorse\_Partisan} & \textbf{Role}\\
\hline
\cellcolor{gray!6}{Yancey, William} & \cellcolor{gray!6}{General Bashing, Then and Now} & \cellcolor{gray!6}{WSJ} & \cellcolor{gray!6}{1989} & \cellcolor{gray!6}{CivMilBalance} & \cellcolor{gray!6}{1} & \cellcolor{gray!6}{0} & \cellcolor{gray!6}{0}\\
\hline
Beckwith, Charlie & Somalia's Needless Deaths & WSJ & 1993 & Warfighting & 1 & 0 & 0\\
\hline
\cellcolor{gray!6}{Odom, William} & \cellcolor{gray!6}{Invade, Don't Bomb} & \cellcolor{gray!6}{WSJ} & \cellcolor{gray!6}{1994} & \cellcolor{gray!6}{Warfighting} & \cellcolor{gray!6}{1} & \cellcolor{gray!6}{0} & \cellcolor{gray!6}{0}\\
\hline
McCain, John & In Bosnia, Another Mistake & WSJ & 1994 & Warfighting & 1 & 0 & 1\\
\hline
\cellcolor{gray!6}{Odom, William} & \cellcolor{gray!6}{One Year? In Bosnia?} & \cellcolor{gray!6}{NYT} & \cellcolor{gray!6}{1995} & \cellcolor{gray!6}{Warfighting} & \cellcolor{gray!6}{1} & \cellcolor{gray!6}{0} & \cellcolor{gray!6}{0}\\
\hline
Odom, William & Chechnya, Freedom, and the Voice of Yeltsin Past & WaPO & 1996 & ForeignPolicy & 1 & 0 & 0\\
\hline
\cellcolor{gray!6}{Mcdonough, James} & \cellcolor{gray!6}{Clinton's Contempt for U.S. Soldiers} & \cellcolor{gray!6}{WSJ} & \cellcolor{gray!6}{1998} & \cellcolor{gray!6}{Warfighting} & \cellcolor{gray!6}{1} & \cellcolor{gray!6}{0} & \cellcolor{gray!6}{0}\\
\hline
Peters, Ralph & It's Wonks vs. Warlords, and Guess Who Wins Again? & WSJ & 1998 & Warfighting & 1 & 0 & 0\\
\hline
\cellcolor{gray!6}{Peters, Ralph} & \cellcolor{gray!6}{A Question of Leadership} & \cellcolor{gray!6}{WSJ} & \cellcolor{gray!6}{1998} & \cellcolor{gray!6}{Warfighting} & \cellcolor{gray!6}{1} & \cellcolor{gray!6}{0} & \cellcolor{gray!6}{0}\\
\hline
Krulak, Charles & Don't Politicize the Joint Chiefs & WSJ & 2000 & CivMilBalance & 0 & 1 & 0\\
\hline
\cellcolor{gray!6}{Odom, William} & \cellcolor{gray!6}{Buchanan Has it Backwards on Globalization} & \cellcolor{gray!6}{WSJ} & \cellcolor{gray!6}{2000} & \cellcolor{gray!6}{ForeignPolicy} & \cellcolor{gray!6}{0} & \cellcolor{gray!6}{1} & \cellcolor{gray!6}{0}\\
\hline
Shalikashvili, John & Kerry Proceeds With Caution & USA Today & 2004 & BdgtWpnsTrps & 0 & 1 & 0\\
\hline
\cellcolor{gray!6}{Clark, Wesley} & \cellcolor{gray!6}{Medals of Honor} & \cellcolor{gray!6}{NYT} & \cellcolor{gray!6}{2004} & \cellcolor{gray!6}{DomesticPolicy} & \cellcolor{gray!6}{0} & \cellcolor{gray!6}{1} & \cellcolor{gray!6}{0}\\
\hline
Shalikashvili, John & Old Soldiers Don't Have to Fade Away & WSJ & 2004 & DomesticPolicy & 0 & 1 & 0\\
\hline
\cellcolor{gray!6}{Franks, Tommy} & \cellcolor{gray!6}{Right Leader, Right Time} & \cellcolor{gray!6}{WSJ} & \cellcolor{gray!6}{2004} & \cellcolor{gray!6}{DomesticPolicy} & \cellcolor{gray!6}{0} & \cellcolor{gray!6}{1} & \cellcolor{gray!6}{0}\\
\hline
Whitlow, William & The Price of Giving Bad Advice & WaPO & 2004 & Warfighting & 1 & 0 & 0\\
\hline
\cellcolor{gray!6}{Franks, Tommy} & \cellcolor{gray!6}{War of Words} & \cellcolor{gray!6}{NYT} & \cellcolor{gray!6}{2004} & \cellcolor{gray!6}{Warfighting} & \cellcolor{gray!6}{0} & \cellcolor{gray!6}{1} & \cellcolor{gray!6}{0}\\
\hline
Clark, Wesley & Before It's Too Late in Iraq & WaPO & 2005 & Warfighting & 1 & 0 & 0\\
\hline
\cellcolor{gray!6}{Clark, Wesley} & \cellcolor{gray!6}{The Next Iraq Offensive} & \cellcolor{gray!6}{NYT} & \cellcolor{gray!6}{2005} & \cellcolor{gray!6}{Warfighting} & \cellcolor{gray!6}{1} & \cellcolor{gray!6}{0} & \cellcolor{gray!6}{0}\\
\hline
Eaton, Paul & A Top-Down Review for the Pentagon & NYT & 2006 & CivMilBalance & 1 & 0 & 0\\
\hline
\cellcolor{gray!6}{Crosby, John and McInerney, Thomas and Moore, Burton and Vallely, Paul} & \cellcolor{gray!6}{In Defense of Donald Rumsfeld} & \cellcolor{gray!6}{WSJ} & \cellcolor{gray!6}{2006} & \cellcolor{gray!6}{Support} & \cellcolor{gray!6}{1} & \cellcolor{gray!6}{0} & \cellcolor{gray!6}{0}\\
\hline
Murtha, John & Confessions of a 'Defeatocrat' & WaPO & 2006 & Warfighting & 1 & 0 & 1\\
\hline
\cellcolor{gray!6}{Eaton, Paul} & \cellcolor{gray!6}{An Army of One Less} & \cellcolor{gray!6}{NYT} & \cellcolor{gray!6}{2006} & \cellcolor{gray!6}{Warfighting} & \cellcolor{gray!6}{1} & \cellcolor{gray!6}{0} & \cellcolor{gray!6}{0}\\
\hline
Scales, Robert & A War the Pentagon Doesn't Want & WaPO & 2013 & Warfighting & 1 & 0 & 0\\
\hline
\cellcolor{gray!6}{Hayden, Michael and Mukasey, Michael} & \cellcolor{gray!6}{NSA Reform that Only ISIS Could Love} & \cellcolor{gray!6}{WSJ} & \cellcolor{gray!6}{2014} & \cellcolor{gray!6}{Warfighting} & \cellcolor{gray!6}{1} & \cellcolor{gray!6}{0} & \cellcolor{gray!6}{0}\\
\hline
McSally, Martha & Saving a Plane that Saves Lives & NYT & 2015 & BdgtWpnsTrps & 1 & 0 & 1\\
\hline
\cellcolor{gray!6}{Scales, Robert} & \cellcolor{gray!6}{Our Army is Breaking} & \cellcolor{gray!6}{WaPO} & \cellcolor{gray!6}{2015} & \cellcolor{gray!6}{BdgtWpnsTrps} & \cellcolor{gray!6}{1} & \cellcolor{gray!6}{0} & \cellcolor{gray!6}{0}\\
\hline
Petraeus, David and O'Hanlon, Michael & Afghanistan After Obama & WaPO & 2015 & Warfighting & 1 & 0 & 0\\
\hline
\cellcolor{gray!6}{Hayden, Michael} & \cellcolor{gray!6}{Classified Briefings and Candidates} & \cellcolor{gray!6}{NYT} & \cellcolor{gray!6}{2016} & \cellcolor{gray!6}{CivMilBalance} & \cellcolor{gray!6}{1} & \cellcolor{gray!6}{0} & \cellcolor{gray!6}{0}\\
\hline
Davis, Morris & Stop Meddling in the Bergdahl Case & NYT & 2016 & CivMilBalance & 1 & 0 & 0\\
\hline
\cellcolor{gray!6}{Hayden, Michael} & \cellcolor{gray!6}{Russia's Useful Fool} & \cellcolor{gray!6}{WaPO} & \cellcolor{gray!6}{2016} & \cellcolor{gray!6}{DomesticPolicy} & \cellcolor{gray!6}{1} & \cellcolor{gray!6}{0} & \cellcolor{gray!6}{0}\\
\hline
McCain, John & We Have a Stake in Syria, Yet We Have Done Nothing & WaPO & 2016 & ForeignPolicy & 1 & 0 & 0\\
\hline
\cellcolor{gray!6}{Hayden, Michael} & \cellcolor{gray!6}{Trump's Most Important New Partner} & \cellcolor{gray!6}{WaPO} & \cellcolor{gray!6}{2016} & \cellcolor{gray!6}{Warfighting} & \cellcolor{gray!6}{1} & \cellcolor{gray!6}{0} & \cellcolor{gray!6}{0}\\
\hline
Hayden, Michael & A Damaging Disregard for Intel & WaPO & 2016 & Warfighting & 1 & 0 & 0\\
\hline
\cellcolor{gray!6}{Hayden, Michael} & \cellcolor{gray!6}{The Travel Ban Hurts American Spies - and America} & \cellcolor{gray!6}{WaPO} & \cellcolor{gray!6}{2017} & \cellcolor{gray!6}{DomesticPolicy} & \cellcolor{gray!6}{1} & \cellcolor{gray!6}{0} & \cellcolor{gray!6}{0}\\
\hline
Hayden, Michael & Trump Proves He's Russia's Useful Fool & WaPO & 2017 & DomesticPolicy & 1 & 0 & 0\\
\hline
\cellcolor{gray!6}{Mullen, Mike} & \cellcolor{gray!6}{The Refugees We Need} & \cellcolor{gray!6}{NYT} & \cellcolor{gray!6}{2017} & \cellcolor{gray!6}{DomesticPolicy} & \cellcolor{gray!6}{1} & \cellcolor{gray!6}{0} & \cellcolor{gray!6}{0}\\
\hline
Mullen, Mike & Bannon Has No Place on the NSC & NYT & 2017 & ForeignPolicy & 1 & 0 & 0\\
\hline
\cellcolor{gray!6}{Hayden, Michael} & \cellcolor{gray!6}{Donald Trump is Undermining Intelligence Gathering} & \cellcolor{gray!6}{NYT} & \cellcolor{gray!6}{2017} & \cellcolor{gray!6}{Warfighting} & \cellcolor{gray!6}{1} & \cellcolor{gray!6}{0} & \cellcolor{gray!6}{0}\\
\hline
Hayden, Michael & The End of Intelligence & NYT & 2018 & DomesticPolicy & 1 & 0 & 0\\
\hline
\cellcolor{gray!6}{McRaven, William} & \cellcolor{gray!6}{Take My Security Clearance, too, Mr. President} & \cellcolor{gray!6}{WaPO} & \cellcolor{gray!6}{2018} & \cellcolor{gray!6}{DomesticPolicy} & \cellcolor{gray!6}{1} & \cellcolor{gray!6}{0} & \cellcolor{gray!6}{0}\\
\hline
Wilkerson, Lawrence and Wilson III, Isaiah and Adams, Gordon & Trump's Border Stunt is a Profound Betrayal of Our Military & NYT & 2018 & Warfighting & 1 & 0 & 0\\
\hline
\cellcolor{gray!6}{Nagl, John} & \cellcolor{gray!6}{Retired Generals Warned Us About Rumsfeld. Now They're Warning Us About Trump} & \cellcolor{gray!6}{WaPO} & \cellcolor{gray!6}{2019} & \cellcolor{gray!6}{DomesticPolicy} & \cellcolor{gray!6}{1} & \cellcolor{gray!6}{0} & \cellcolor{gray!6}{0}\\
\hline
Allen, John and Victor, David & Despite What Trump Says, Climate Change Threatens Our National Security & NYT & 2019 & DomesticPolicy & 1 & 0 & 0\\
\hline
\cellcolor{gray!6}{McRaven, William} & \cellcolor{gray!6}{Our Republic is Under Attack From the President} & \cellcolor{gray!6}{NYT} & \cellcolor{gray!6}{2019} & \cellcolor{gray!6}{DomesticPolicy} & \cellcolor{gray!6}{1} & \cellcolor{gray!6}{1} & \cellcolor{gray!6}{0}\\
\hline
McRaven, William & If Good Men Can’t Speak the Truth, We Should Be Deeply Afraid & WaPO & 2020 & DomesticPolicy & 1 & 0 & 0\\
\hline
\cellcolor{gray!6}{VanLandingham, Rachel and Corn, Geoffrey} & \cellcolor{gray!6}{Military Brass Finally Speaks Up on Trump} & \cellcolor{gray!6}{USA Today} & \cellcolor{gray!6}{2020} & \cellcolor{gray!6}{DomesticPolicy} & \cellcolor{gray!6}{1} & \cellcolor{gray!6}{0} & \cellcolor{gray!6}{0}\\
\hline
Seidule, Ty & West Point and its Cadets are not Campaign Props & WaPO & 2020 & DomesticPolicy & 1 & 0 & 0\\
\hline
\cellcolor{gray!6}{Vindman, Alexander} & \cellcolor{gray!6}{Coming Forward Ended My Career. I Still Believe Doing What's Right Matters} & \cellcolor{gray!6}{WaPO} & \cellcolor{gray!6}{2020} & \cellcolor{gray!6}{DomesticPolicy} & \cellcolor{gray!6}{1} & \cellcolor{gray!6}{0} & \cellcolor{gray!6}{0}\\
\hline
McRaven, William & Trump is Working to Actively Undermine the Postal Service - and Every Major U.S. Institution & WaPO & 2020 & DomesticPolicy & 1 & 1 & 0\\
\hline
\cellcolor{gray!6}{Vindman, Alexander and Gans, John} & \cellcolor{gray!6}{Trump Has Sold off America's Credibility for His Personal Gain} & \cellcolor{gray!6}{NYT} & \cellcolor{gray!6}{2020} & \cellcolor{gray!6}{DomesticPolicy} & \cellcolor{gray!6}{1} & \cellcolor{gray!6}{0} & \cellcolor{gray!6}{0}\\
\hline
McRaven, William & Biden Will Make America Lead Again; We Need a President with Decency and a Sense of Respect & WSJ & 2020 & DomesticPolicy & 1 & 1 & 0\\
\hline
\cellcolor{gray!6}{Allen, Thad} & \cellcolor{gray!6}{Trump is Failing to Provide For the Common Defense} & \cellcolor{gray!6}{WaPO} & \cellcolor{gray!6}{2020} & \cellcolor{gray!6}{DomesticPolicy} & \cellcolor{gray!6}{1} & \cellcolor{gray!6}{0} & \cellcolor{gray!6}{0}\\
\hline
Hayden, Michael et. Al & A Dangerous Purge & WaPO & 2020 & Warfighting & 1 & 0 & 0\\
\hline
\cellcolor{gray!6}{89 Former Defense Officials} & \cellcolor{gray!6}{The Military Must Never Be Used to Violate Constitutional Rights} & \cellcolor{gray!6}{WaPO} & \cellcolor{gray!6}{2020} & \cellcolor{gray!6}{Warfighting} & \cellcolor{gray!6}{1} & \cellcolor{gray!6}{0} & \cellcolor{gray!6}{0}\\
\hline
\end{longtable}
\endgroup{}
\doublespace

First, 17 of the 49 publications are, upon a closer reading, determined to not reflect possible violations of military professionalism. This is worth showcasing in order to draw attention to the fact that there are instances of retired military officers authoring opinion pieces which address the subjects of social policy related to the military, domestic policy, and figures by name while simultaneously upholding standards of military professionalism. However, the data reveals that in practice, such instances are questionable. The practical implication is simple and straightforward: if military officers decide to write about topics other than war fighting and operations, troop support and institutional culture, or strategic advice, they need to carefully craft their argument so as to not violate norms of military professionalism.

Second, and resulting from the first, is that across the entire data set, instances of potential violations of military professionalism in the op-eds published by retired military officers in major US newspapers is a somewhat rare occurrence. Indeed, the entire count of potentially questionable publications with respect to military professionalism across the data set is just 32 of 344 total observations, or 9.3\% of the entire sample. Thus, ample support exists to confirm the first hypothesis, that the vast majority of opinion pieces authored by retired military officers uphold standards of military professionalism.

A third and interesting finding, however, is that the first publishing year in which an opinion piece penned by a retired officer potentially failed to uphold professional standards is 1998, nearly halfway across the sample in terms of publishing year. Why is this the case? We will deal more with the suspected culprits of this trend, rising political polarization and rising institutional credibility of the military, later in this paper. The point to make is that without question, retired military officers before 1998 had plenty of opportunities either to air partisan preferences and/or to address topics that may have similarly pushed the limits of military professionalism. In fact, many of these retired officers would have been veterans of the Vietnam and Korean Wars, and perhaps, like Maxwell Taylor, an author who is in the sample data, veterans of World War Two. Moreover, the data shows that these retirees wrote opinion pieces ( a total of 46 observations were written before the first observation in Table 4 raises a question about military professionalism), but that in doing so, they stayed clearly within the bounds of military professionalism.

Fourth, Table 4 reveals that several retired military officers wrote publications not from the perspective of a retired military officer, but from the perspective of other professional and political roles that these officers served in after retiring from the military. These include op-eds by: Retired Air Force Lieutenant General Claudius Watts, who wrote an op-ed regarding the advantages of all-male education while serving as the president of the Citadel; Retired Army General Barry McCaffrey, who penned an opinion piece while serving in the administration of Bill Clinton as the Director of the Office of National Drug Control Policy; and Retired Marine Corps Colonel John Murtha, who also served in the US House of Representatives. Each of these authors addressed topics that one should reasonably expect them to on the basis of the roles each was serving in. This would include, in the case of John Murtha, making obviously partisan agreements and thus, revealing political preferences.

Table 4 reveals, fifth, that several publications are deemed as possibly having violated military professionalism in that they address topics that may be considered outside of the realm of proper military expertise. It is worth briefly commenting on several of these examples. In his \emph{Washington Post} editorial, ``Home Should Not Be a War Zone,'' Retired Army General Stanley McChrystal addresses the tragic nature of mass shootings in America. He states in the article, ``\ldots{}as a combat veteran and proud American, I believe we need a national response to the gun violence that threatens so many of our communities'' \autocite{mcchrystal_home_2016}. He later, in 2018, pens an op-ed entitled ``Good Riddance'' in which he discusses the evolution of his views on Robert E. Lee and other Confederate soldiers and symbols \autocite{mcchyrstal_stanley_good_2018}. Retired Army Brigadier General Ty Seidule similarly, in 2020, advocates for the renaming of several Army installations \autocite{seidule_ty_what_2020}. Other examples of observations coded as potentially violating military professionalism based on the topic of the publication include articles penned by Retired Navy Admiral Mike Mullen (entitled ``The Refugees We Need'') and co-authors Retired Army Lieutenant General Mark Hertling and Retired Naval Admiral Robert Natter (entitled ``Cutting Refugee Admissions Will Have Severe Consequences for the US Military'').

These articles are certainly not unreasonable in that they, in the opinion of the author of this article, express reasonable thoughts and ideas to many of the nation's pressing problems, including gun violence, the appropriate legacy of Confederate symbols, and national immigration policy. To be clear, many Americans agree with the stance these officers take in the articles they have published. One can even make a strong, perhaps even a compelling argument, why these officers might be uniquely qualified to address aspects of some of these topics in that they possess expertise commanding large and powerful organizations charged to carry out violence on behalf of the country (related to the issue of gun control), that they may have experience serving and leading troops on bases named after Confederate generals (related to the issue of Confederate symbols and bases named after Confederate generals), and that they have have considerable experience serving alongside soldiers of multiple ethnicities as well as considerable experience in Iraq and Afghanistan (related to the issue of improving American immigration policy).

But I contend that it is also not wrong to insist that these and other contentious national issues are most appropriately solved by elected civilian officials and not by retired military officers. Interestingly, in July of 2020, General Mark Milley, the Chairman of the Joint Chiefs of Staff, was asked by US Representative Anthony Brown, D-MD, during a hearing held by the House Armed Services Committee, to ``comment on the naming of Army installations after Confederate soldiers,'' and whether the continued use of the names of these bases ``reflects the values'' that the military is trying to instill in the nation's armed forces \autocite{milley_mark_general_2020}? Milley's response was instructive. He offered the Congressmen the idea that even as the Department of Defense continues to take a hard look at the issue, the decision to remove those names should reflect the decision to adopt those names in the first place in that such a decision was ``a political decision'' \autocite{milley_mark_general_2020}. Milley's answer was the appropriate one, I contend, because it recognized that regardless of how he feels about this issue, the issue is not for him (or the military) to solve. Many other topics addressed by publications listed in Table 4, to include gun violence and immigration, fall into the same realm.

A sixth finding is that there are roughly three clusters of observations in Table Four with respect to the year of publication. In the first cluster (1998-1999), the publications reveal multiple grievances between certain retired military officers and the manner in which President Clinton was planning or using military force abroad. Only two of the five observations in this first cluster are coded as possibly having violated standards of military professionalism.

The second cluster refers to publications penned between 2004-2006. Most of these comprise part of or are related to the so-called ``Revolt of the Generals,'' which occurred near the 2004 US midterm elections and referred to sharp disagreements that entered the public arena between multiple US generals and admirals and President George W. Bush's Secretary of Defense, Donald Rumsfeld. These disagreements were directly related to these military leaders' disagreement about Rumsfeld's leadership regarding the planning and execution of the early years of the war in Iraq. Retired Army General Tommy Franks penned two observations that essentially provided public support for the President, George W. Bush, while Retired Army Major General Paul Eaton, a critic of Rumsfeld, authored two of the pieces published in this cluster. Of the seven observations published in Table 4 between 2004-2006, all but one contained possible violations of the concept of military professionalism.

The third and largest cluster occurs between 2016-2019. The 23 observations in this cluster reveals concern over the candidate and later President Donald Trump. The titles alone of many of these opinion pieces reveal, at a minimum, clear partisan preferences by many of the authors. Retired Air Force General Michael Hayden, Navy Admirals Mike Mullen and William McRaven, and Army General Stanley McChrystal are among the retired officers who authored several publications during this time period. All but four of the observations in this cluster are coded as having potentially violated standards of military professionalism. Of note, one observation in this cluster entitled, ``Fists Raised at West Point'' and published in the Washington Post is actually a compendium of two anonymous letters, each one taking an opposite stance on the appropriateness of an incident in which a photograph at West Point's 2016 graduation ceremony captured a graduating female cadet raising a clenched a fist, an action which some saw as an inappropriate political gesture related to the growing Black Lives Matter movement \autocite{unknown_fists_2016}.

\hypertarget{variation-in-military-professionalism-over-time}{%
\subsection{Variation in Military Professionalism Over Time}\label{variation-in-military-professionalism-over-time}}

Figure 3, below, plots both the total number and proportion of yearly opinion pieces authored by retired military officers that address topics of social policy affecting the military, domestic policy, and political administrations or figures. The points are the actual terms (number and proportion, respectively), while the blue line represents a smoothed line of best fit for the data. Both plots reveal that beginning with the first year of the data, 1979, there has a moderate increase until approximately 2005, followed by a decline until approximately 2011, followed by a sharp increase that has persisted until the present day. What explains these dynamics - both the increase that occurred between 1979 - 2005 as well as the sharper increase between 2012 - today - and the decline that occurred between roughly 2005 - 2011? According to the theory developed at the beginning of this paper, a significant reason stems from changes in political polarization and institutional credibility that have occurred over the past four decades.

\includegraphics{Tahkv1_files/figure-latex/figure 3-1.pdf}

To understand how measurements of political polarization have changed over time, I rely on the DW-NOMINATE (dynamic-weight, nominal three step estimation) scores that were originally developed by the political scientists Keith Poole and Howard Rosenthal in the early 1980s \autocite{jeffrey_b_lewis_voteview_2020}. The DW-NOMINATE data are one measure that uses roll-call voting data for each member of congress to assign a score to the member of congress on a liberal-conservative dimension, ranging from -1 (extreme liberal) to 1 (extreme conservative) \autocite{jeffrey_b_lewis_voteview_2020}. The data is frequently relied upon by academics, journalists, and the media to understand how a variety of political bodies, to include Congress and political parties, have changed over time. Over the time period examined in this paper, polarization in each chamber of Congress, the House of Representatives and the Senate, has increased \autocite{jeffrey_b_lewis_voteview_2020}. The distance in party mean DW-NOMINATE scores in the late 1970s was just less than .6, while the distance today for both chambers is above .8 \autocite{jeffrey_b_lewis_voteview_2020}.

Changes in the institutional credibility of the American military institution can be seen by examining public opinion data. Pollsters survey the American public annually to gauge their trust in a variety of institutions. Using Gallup opinion polling, one sees that since the late 1970s, the credibility of the American military as an institution has gradually increased \autocite{gallup_polls_confidence_2020}. In 1979, 54\% of those surveyed expressed either a great deal or quite a lot of confidence in the military, while in 2020, 72\% expressed the same \autocite{gallup_polls_confidence_2020}. These trends in political polarization and confidence in the military are displayed in Figure Four, below. The top and middle graph show changes in means between Republicans and Democrats in the House of Representatives (top graph) and Senate (middle graph) since 1979, while the bottom graph shows changes in the percentage of surveyed Americans who reported either a great deal of or a lot of confidence in the military institution over the same time period.

\includegraphics{Tahkv1_files/figure-latex/figure 4-1.pdf}

While it is possible that the trend in increasing institutional credibility of the US military over the past four decades may be at least partially responsible for the increasing trend in the number of opinion pieces published by retired military officers, a trend that the data supported earlier as shown in Figure 2, rising institutional credibility alone does not adequately explain the trends shown in Figure 3: that there has been a sharp increase over the past four decades in retired military officers penning pieces that address non-traditional military topics that lay outside the bounds of relevant military expertise, or pieces that express partisan sentiment. The theory laid out earlier in this paper predicted that as political polarization increases, three mechanisms are at work which have strong impacts on the military professional: it becomes increasingly difficult to remain politically neutral as partisan conflicts subsume a greater number of issues; as this occurs, strong feelings of dislike or disdain surface against those who hold on opposing view; and alterations in one's sense of loyalty arise, particularly as issues that individuals hold to most dearly raise the salience of one's identity.

To test if changes in political polarization adequately explain the changes in the qualitative nature (increasing trends of potential violations of military professionalism) of opinion pieces authored by retired military officers, I conduct a logit regression of the variable \emph{questionable\_professionalism} while holding institutional trust in the military constant at its mean (71 percent) and varying the value of the mean ideological distance between both major political parties in the House between 1979-2020. Conducting such a regression allows one to examine, given the data assembled, the predictive likelihood of publishing an opinion piece that violates military professionalism given varying levels of political polarization and constant institutional credibility in the military. These results are displayed in Figure 5, below.

\% Table created by stargazer v.5.2.2 by Marek Hlavac, Harvard University. E-mail: hlavac at fas.harvard.edu
\% Date and time: Thu, Dec 03, 2020 - 19:06:06
\% Requires LaTeX packages: dcolumn

\begin{table}[!htbp] \centering 
  \caption{Table 3: Logistic Regression Results} 
  \label{} 
\begin{tabular}{@{\extracolsep{5pt}}lD{.}{.}{-3} D{.}{.}{-3} D{.}{.}{-3} D{.}{.}{-3} } 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{4}{c}{\textit{Dependent variable:}} \\ 
\cline{2-5} 
\\[-1.8ex] & \multicolumn{1}{c}{leg\_crit\_part} & \multicolumn{1}{c}{crit\_part} & \multicolumn{1}{c}{leg\_crit\_part} & \multicolumn{1}{c}{crit\_part} \\ 
 & \multicolumn{1}{c}{Model 3} & \multicolumn{1}{c}{Model 4} & \multicolumn{1}{c}{Model 5} & \multicolumn{1}{c}{Model 6} \\ 
\\[-1.8ex] & \multicolumn{1}{c}{(1)} & \multicolumn{1}{c}{(2)} & \multicolumn{1}{c}{(3)} & \multicolumn{1}{c}{(4)}\\ 
\hline \\[-1.8ex] 
 Institutional Credibility & -0.057^{*} & -0.047 & -0.024 & -0.019 \\ 
  & (0.031) & (0.036) & (0.024) & (0.029) \\ 
  & & & & \\ 
 Congressional Polarization & 0.084^{***} & 0.088^{***} &  &  \\ 
  & (0.027) & (0.031) &  &  \\ 
  & & & & \\ 
 Affective Polarization &  &  & 0.033^{**} & 0.037^{**} \\ 
  &  &  & (0.013) & (0.015) \\ 
  & & & & \\ 
 Male & -0.708 & -0.856 & -0.848 & -0.976 \\ 
  & (1.256) & (1.254) & (1.262) & (1.263) \\ 
  & & & & \\ 
 Mixed Authorship & 1.340 & 1.265 & 1.324 & 1.253 \\ 
  & (1.755) & (1.739) & (1.759) & (1.746) \\ 
  & & & & \\ 
 Unknown Authorship & -16.178 & -15.996 & -16.366 & -16.206 \\ 
  & (787.186) & (798.975) & (789.107) & (799.686) \\ 
  & & & & \\ 
 Rank & 0.171^{**} & 0.022 & 0.167^{**} & 0.018 \\ 
  & (0.076) & (0.078) & (0.076) & (0.078) \\ 
  & & & & \\ 
 Constant & -4.977^{**} & -4.946^{**} & -2.567 & -2.256 \\ 
  & (2.155) & (2.477) & (1.945) & (2.173) \\ 
  & & & & \\ 
\hline \\[-1.8ex] 
Observations & \multicolumn{1}{c}{389} & \multicolumn{1}{c}{389} & \multicolumn{1}{c}{388} & \multicolumn{1}{c}{388} \\ 
Log Likelihood & \multicolumn{1}{c}{-180.151} & \multicolumn{1}{c}{-149.394} & \multicolumn{1}{c}{-182.477} & \multicolumn{1}{c}{-150.757} \\ 
Akaike Inf. Crit. & \multicolumn{1}{c}{374.301} & \multicolumn{1}{c}{312.787} & \multicolumn{1}{c}{378.954} & \multicolumn{1}{c}{315.514} \\ 
\hline 
\hline \\[-1.8ex] 
\textit{Note:}  & \multicolumn{4}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
\end{tabular} 
\end{table}

The graph clearly shows a non-linear relationship between increasing political polarization and the likelihood that an opinion piece authored by a retired military officer will potentially violate aspects of military professionalism. While the confidence intervals at each point estimate are relatively large, the graph demonstrates that the predicted probability of violating military professionalism is different between relative extremes along the x axis. In other words, the political polarization levels that exist today, as measured by the ideological distance between members of the two prominent political parties in the house, is predicted to have resoundingly different impacts on military professionalism than the polarization levels that existed several decades ago. The polarization levels that existed several decades ago, in the late 1970s and 1980s, may have had negligible, if any impact, on the likelihood of a retired military officer violating military professionalism in the content of a published opinion piece, but the political landscape is simply different today. Contemporary polarization levels, insofar as the predictive power of the model is concerned, changed the environment: at contemporary polarization levels (the far right of the x axis), there is roughly a 17\% chance of a retired military officer's published opinion piece violating the standards and norms of military professionalism, with confidence intervals that range from roughly 11\% to nearly 25\%.

Given these dynamics, as well as the results of both Figures 3 and 5, ample evidence exists to confirm the second hypothesis, which stated that as a result of increasing political polarization, there will be an increase in recent years in the number of op-eds written about topics that are not relevant to military expertise, or that reveal inappropriate partisan sentiments, among authors which are a vital component of the institutional military, retired military officers.

\hypertarget{does-combat-attenuate-polarizations-impact-on-military-professionalism}{%
\subsection{Does Combat Attenuate Polarization's Impact on Military Professionalism?}\label{does-combat-attenuate-polarizations-impact-on-military-professionalism}}

Why did Figure 3 reveal a sharp drop between approximately 2005-2011 in both the real number and proportion of annual opinion pieces authored by retired military officers that address non-traditional military topics, such as social policy that impacts the military, domestic policy, and political administrations/figures? What was it about this time period that made it unlike any other span of years across the past four decades?

One compelling answer is that between these years, roughly 2005-2011, the military was heavily involved in combat operations to a point such that the institution was taxed tremendously, thereby necessitating that it remain focused on its wartime mission. Repeated deployments during the Global War on Terrorism, which included substantial surges of forces to both Iraq (2007-2008) and Afghanistan (2010-2011), as well as sustained US casualties resulting from these conflicts, possibly shaped the military institution so strongly such that even retired military officers acknowledged that the military's institutional energy required a laser-sharp focus on combat requirements. This is not to suggest the military was not heavily engaged in conflicts nor active around the glove at other times throughout the past four decades (which we know is not the case), but to instead suggest that even retired military officers are perhaps more sensitive to combat requirements and missions undertaken by the military as higher levels of continued operational strain and duress are endured. Though more data is required to fully test this argument, other scholars have argued similarly in that periods of transition, especially away from major combat operations and into eras of downsizing and relative peace, often seem to test the very essence of military professions in significant ways \autocite{snider_once_2012}.

\hypertarget{conclusion-will-the-military-remain-professional-in-a-polarized-setting}{%
\section{Conclusion: Will the Military Remain Professional in a Polarized Setting?}\label{conclusion-will-the-military-remain-professional-in-a-polarized-setting}}

The goal of this paper was to explore and examine the ways in which political polarization and institutional credibility have impacted the professionalism of the military over the past fifty years. Examining the published opinion pieces authored by retired military officers in major US newspapers revealed that although the vast number of total opinion publications uphold the military's norms and traditions concerning military professionalism, a couple of worrisome trends are evident in the data: first, retired military officers are increasingly writing about topics outside of the umbrella of traditional national security topics, and second, retired officers are expressing partisan preferences at a greater rate than they previously did. What do these trends mean, and should they cause concern? If so, what should be done about it?

First, these findings illuminate the power and pervasiveness of increasing political polarization, and in particular, its impact not only on political institutions, but also on professions, which the military prides itself on being. Thus, if the military, which has for decades adopted and promulgated norms of apolitical and nonpartisan expression, has been impacted by rising political polarization, it stands to reason that other political institutions and professions also have been impacted. While the specific manifestations of political polarization's increase and spread in non-military institutions and professions might be different, the fact that the military is experiencing a noticeable impact as a result of increasing political polarization speaks to its significant power and influence.

Indeed, evident in the content of opinion pieces authored by retired military officers over the past four decades are the three features of contemporary American political polarization developed in the theory proffered earlier in this paper: the loss of politically neutral space as more issues enter the partisan battleground, the creation of strong feelings of dislike for those who hold to opposing views on these issues, and perhaps worst for the military, alterations in an individual's concept of loyalty, which arises as the salience of one's identity is raised as the issues that an individual holds most dearly to are increasingly at stake. Moreover, the data supported what most scholars of American politics have argued for: a realization that polarizing politics across America's political landscape will likely not disappear anytime soon, regardless of which party holds the nation's highest office.

Second, these findings underscore the challenge that the military as a profession has in ``reigning in'' its retired officers and compelling them to stay out of partisan battles, and out of politics more broadly. Among the features and traits of a profession is that it is self-policing; however, the data indicates that current efforts by active military leaders to police and clean up the behavior of its retired elites is not working. Indeed, the letters and pleas of previous military leaders to retired military officers, including those written by now Retired Army General Martin Dempsey and Retired Marine Corps General Joseph Dunford, both of which were referenced in this paper, are simply not working. This suggests that stronger and more creative efforts are required.

Third, these findings demonstrate that the phrase ``politicization of the military,'' which has long been a proper concern of academics, journalists, the military, and citizens alike, should not be intellectually confined to notions of how politicians or other political leaders political leaders attempt to harness or exploit the military for political purposes or advantages. Though such efforts do occur and are a genuine cause for concern that continue to deserve both scholarly and practical attention, the evidence presented in this paper shows that ``politicization of the military'' is also occurring at the hands and through the words of particular retired military officers. Gaining a better understanding of the impacts of political polarization on the military profession remains a critical area for future research. Other scholars such as Michael Robinson have explored the specific ways that polarization is impacting the interactions between the military and the public \autocite{robinson_michael_danger_2018}. Though this paper focused on the ways that polarization may impact professional military behavior, it recognized but stopped short of discussing polarization's subsequent impact on various ``military outcomes,'' which may arguably include even more decisive events, such as achieving victory or defeat in battle, or the ability to formulate sound strategic advice. These outcomes and more, and the role of polarization in influencing them, deserve to be explored further.

Maintaining credibility and trust with the American public is crucial for the military, but should a polarized political climate endure, the military needs to also understand that credibility and trust can be abused to the point that they also can constitute significant risks. One such risk is that as polarization worsens, military leaders, to include military retirees, abuse the high level of credibility they currently enjoy with the American public to advocate for obviously partisan positions. Such actions, though, in a polarized setting, might actually result in an increase in the credibility of the military institution, at least among citizens who share the same partisan sentiments advocated for by military leaders, which might, somewhat perversely, entice more of the same behavior. A second risk, stemming from the first, is that a polarized setting sets the stage for the military institution to eventually openly jockey with civilian leaders, either for power or for control, a concern that generations of scholars have warned us about \autocites{finer_man_1962}{brooks_shaping_2008-1}. Were either of these risks to materialize, the character and nature of the American military as an institution, and likely, American democracy, would be irrevocably changed.

There are no simple solutions for the military with respect to the risks posed by rising political polarization. Two possible ideas, however, are apropos to this discussion. One involves identity, and the other, credibility. First, our nation's military is composed of soldiers who come from every walk of life throughout America, and accordingly, in the years to come, many future soldiers and leaders will bear the features of a polarized society. Thus, the military's challenge, and one which it must embrace itself for immediately, is integrating the multiple identities each soldier carries with them as people in such a way that each can still proudly wear the uniform and bear the identity of an American service member. Though this challenge is not new for the US military, increasing political polarization has significantly raised the level of difficulty the military faces in tackling such a vital challenge. Scholars who have studied multi-ethnic militaries across the world have argued that it is military leaders, not civilians, who ultimately devise effective solutions to the problems posed by soldiers who carry multiple loyalties, and interestingly, only as a result of combat requirements \autocite{peled_question_1998}. Perhaps the American challenge is even greater in that at least for the near term future, there is no major war that will necessitate the coming together of disparate loyalties and identities among future soldiers.

The second idea, and one that may sound like anathema to modern military leaders, is that perhaps the level of trust and credibility between the American public and the military institution should not be the ultimate indicator of the military's performance as a profession. In suggesting this, I am not suggesting that credibility and trust are not important or that the concept should be disregarded entirely. The American public is, after all, the client of the American military and ultimately, it is their opinion that counts.I am suggesting, however, that for the reasons laid out in this paper, credibility and trust are influenced by the mechanisms through which polarization works, to include the loss of politically-neutral space, increased negative feelings toward those who hold differing opinions, and most sinister of all, alterations in one's concept of loyalty. For these reasons, I am simply offering the notion that \emph{credibility can be deceiving} in polarized times. Now, more than ever before, military leaders will need to rely on individual moral courage over and above any other trait in order to perform their duties as military professionals.\\
\singlespace
\pagebreak

\printbibliography

\end{document}

### 2020-12-04 code/prospectusv2.Rmd
commit

@@
 ---
-title: "Servants or Syndicates?: The Rise of Domestic Political Interventionism by Senior Military Leaders (Prospectus)"
+title: "Servants, Syndicates, and Saviors: The Rise of Domestic Political Interventionism by Senior Military Leaders (Prospectus)"
 author: "Peter M. Erickson"
@@
 To support my argument, I trace changes in levels and types of distinct forms of political activism by military officers, including authoring opinion-editorials in major newspapers, as domestic polarization and military institutional credibility rises.
-I ultimately argue that the increasing political activism of retired military leaders is not merely a response to the hyper-partisanship present in the current political atmosphere, but rather the logical outworking of historical trends that have been building for decades.
+I ultimately argue that the increasing political activism of military leaders is not merely a response to the hyper-partisanship present in the current political atmosphere, but rather the logical outworking of trends involving the prestige of the military institution as well as political polarization that have been building for decades.
 I conclude by describing the implications of changing beliefs among military leaders regarding the role the military should have in domestic politics.
-This dissertation contributes to the fields of international relations and civil-military relations by advanving an empirical and historical analysis of a growing phenomenon: military leaders intervening in domestic American politics.
+This dissertation contributes to the fields of international relations and civil-military relations by advancing an empirical and historical analysis of a growing phenomenon: military leaders intervening in domestic American politics.
 _
@@
 However, as almost every civil-military relations scholar points out, the absence of coups does not mean that the relationship between civilian and military leaders is healthy, nor absent of conflict altogether.
-Indeed, a cursory examination of the relationships between Lincoln and MacClellan, Truman and MacArthur, and Obama and McChrystal reveals that disagreement between those who serve at the highest levels of civilian government and military organizations is neither rare nor understudied.
+Indeed, a cursory examination of the relationships between Lincoln and McClellan, Truman and MacArthur, and Obama and McChrystal reveals that disagreement between those who serve at the highest levels of civilian government and military organizations is neither rare nor understudied.
 
-What has changed, however, is the degree to which a very specific subset of the American military institution - retired senior military officials - have participated in the domestic political process post retirement.
+What has changed, however, is the degree to which a very specific subset of the American military institution - senior military leaders - have participated in the domestic political process, especially post retirement.
 Indeed, whereas General Omar Bradley, one of the military's five star generals and a hero of World War Two, famously remarked in 1959, "I am convinced that the best service a retired general can perform is to turn in his tongue along with his suit and mothball his opinions," [@times_bradley_1959], retired generals and admirals today appear frequently in the media, endorse Presidential candidates at higher frequency [@griffiths_not_2019], and frequently serve in government after their retirement from military service [@brooks_dont_2016].
-To provide one such example, following protests across the United States in the summer of 2020 in which President Donald Trump nearly ordered the use of active duty military forces in American cities, former Chairman of the Joint Chiefs of Staff Admiral (Ret) Michael Mullen was one of at least ten four star retired officers who publicly aired concern about the President's potential use of force.
+To provide only one such example, following protests across the United States in the summer of 2020 in which President Donald Trump nearly ordered the intervention of active duty military forces into American cities, former Chairman of the Joint Chiefs of Staff, Admiral (Ret) Michael Mullen, was one of at least ten four star retired officers who publicly aired concerns through a variety of mediums about the President's potential use of active-duty military force.
 ^[Other officers included General (Ret) James Mattis, General (Ret) Mark Kelly, Admiral (Ret) William McRaven, General (Ret) Martin Dempsey, General (Ret) Colin Powell, General (Ret) Michael Hayden, General (Ret) David Petraeus, General (Ret) John Allen, and General (Ret) Vincent Brooks.].
-During an interview in the days which followed, when asked by NPR reporter why he spoke out, Mullen stated, "we were getting to a point where the country was in jeopardy, which is why I spoke out" [@isaacson_ret_2020; for further analysis, see also @brooks_let_2020].
+In fact, during an interview in the days which followed the incident, when asked by an NPR reporter why he decided to publicly share his thoughts on the situation, Mullen stated, "we were getting to a point where the country was in jeopardy, which is why I spoke out" [@isaacson_ret_2020; for further analysis, see also @brooks_let_2020].
 
-Mullen's words are incredible from a number of important perspectives, and raise important normative questions regarding the military's role in a democratic society [@brooks_let_2020] and the conditions under which the military can and should express dissent to its civilian bosses [@cohn_what_2019].
-Few, if any, scholars have attempted to systematically examine and explain the rise of political activism by retired military officers over time.
-Such an effort would better illumine the understanding of scholars, as well as the American military and society, of why a powerful group of what were once non-political actors have, at least in some cases, suddenly become one.
-More important, a systematic study of this trend would contribute to a much-needed discussion regarding the role of the military institution and its relationship to the government and American society in a contentious political age.
+Mullen's words are incredible from a number of perspectives, and raise important normative questions regarding the military's role in a democratic society [@brooks_let_2020] and the conditions under which the military can and should express dissent to its civilian bosses [@cohn_what_2019].
+And while a small but dedicated number of civil-military relations scholars do examine these and related issues quite frequently in today's charged political environment, few, if any, have attempted to systematically examine and explain over time the conditions under which senior military leaders have broken and will break normative traditions and engage in domestic politics.
+Such an effort would better illumine the understanding of scholars, as well as of the military and society more broadly, concerning why a powerful group of what were once non-political actors have, at least in some cases, suddenly become one.
 
-The central argument made in this dissertation is that increasing political activism at the hands of military officers is the result of two longstanding and interconnected trends: rising political polarization within the domestic political sphere, and rising prestige and credibility of military institutions afforded by American society.
-The interaction of these trends explains the remarkable increase in political activity by retired senior military leaders, and portends a shift in the perceived role that these senior military leaders fill vis-a-vis civilian authority - a shift away from professional and obedient servants and towards "guardians of democracy" who pick and choose when and how to intervene in American politics.
-Such a shift builds on previous work regarding normative "role beliefs" [@fitch_armed_1998] of the military institution, and has significant ramifications on the organizational culture of the military, its relationship with government and society, and American national security.
+The central argument made in this dissertation is that increasing political activism at the hands of military officers indicates a shift in the beliefs of the military regarding its central role and purpose in society, or "role beliefs" [@fitch_armed_1998].
+This shift, moreover, is the result of two longstanding and interconnected trends: rising political polarization within the domestic political sphere, and the rising prestige and credibility of military institutions afforded by American society.
+The interaction of these trends explains the remarkable increase in political activity by retired senior military leaders, and portends a shift in the perceived role that these senior military leaders fill vis-a-vis civilian authority - a shift away from professional and obedient "servants" and towards that of "syndicates" or even "saviors" of the American state.
+This shift, moreover, holds enormous ramifications regarding the organizational culture of the military, its relationship with government and society, and American national security.
 
@@
 
-My proposed dissertation topic features within the subfield of civil-military relations, a broad academic subfield that is generally placed within the International Relations and Comparative Politics subfield (though within the American context, the subfield of American Politics also certainly applies).
+My proposed dissertation topic features within the sub-field of civil-military relations, a broad academic sub-field that is generally placed within the International Relations and Comparative Politics sub-field (though within the American context, the sub-field of American Politics is entirely relevant and appropriate).
 Civil-military relations involve a wide host of significant and important theoretical questions that involve the pull and push of power between institutions, the delineation of authority between military and civilian leaders, and the relationship between a state's military and society.
@@
 The next level of intervention Finer labels "pressures or blackmail," describing efforts by the military to "convince the civil power by threat of some sanction" [@finer_man_1962, 77].
-Finer adds that military intervention via this level is "on the margin," referring to his belief that while some cases may be unproblematic, many other forms, such as the "threat to stage a coup," are normatively wrong [@finer_man_1962, 77].
+Finer adds that military intervention via this level is "on the margin," referring to his belief that while some cases may not be problematic, many other forms, such as the "threat to stage a coup," are normatively wrong [@finer_man_1962, 77].
 The third level of intervention, "displacement," refers to the "removal" of government officials by the military as a result of "violence or the threat of violence," and the fourth level, "supplantment," occurs when the military "sweeps away the civilian regime and establishes the military in its place" [@finer_man_1962, 78].
@@
 The first is that there is increasing difficulty for military leaders to remain politically neutral, non-partisan, or otherwise apolitical - a hallmark of Huntingtonian military professionalism.
-Fewer and fewer issue areas and policy matters are outside of the "spell" of polarization, owing to a phenomenon schoalrs of American politis refer to as "conflict extension" [@layman_party_2002].
+Fewer and fewer issue areas and policy matters are outside of the "spell" of polarization, owing to a phenomenon scholars of American politics refer to as "conflict extension" [@layman_party_2002].
 In previous eras of polarization, whereas a single issue may have served as the "battleground" for partisan conflict, this is no longer true today, as research has consistently shown that voters who identify as liberals are more liberal across multiple issue categories and that the same trend holds for conservatives [@abramowitz_great_2018].
@@
 
-With respect to the military institution itself, one of the most massive and influential changes that has occurred over the past roughly seventy years is the increase in overall prestige, credibility, and influence of the military institution.
-While key legislative acts passed since World War Two, to include the National Security Act of 1947 and the Goldwater-Nichols Act of 1946, may have fixed important gaps in America's ability to meet its national security requirements [@inboden_national_2017], they also cemented the military as one of if not the dominant institution and actor within American politics.
-Over the same time period, while there have been spikes as a result of exogenous events such as wars and scandals that have involved the military, public opinion towards the military has risen sharpy.
+With respect to the military institution itself, one of the most massive and influential changes that has occurred over the past roughly seventy years is the increase in overall prestige, influence, and public credibility of the military institution.
+While key legislative acts passed since World War Two, to include the National Security Act of 1947 and the Goldwater-Nichols Act of 1946, may have fixed important gaps in America's ability to meet its national security requirements [@inboden_national_2017], they also cemented the military as one of the dominant institutions and actors within American politics.
+Over the same time period, while there have been spikes as a result of exogenous events such as wars and scandals that have involved the military, public opinion towards the military has increased.
 Consider that in the late 1970s, as a likely result of public opinion over the Vietnam War, 54% of Americans indicated that they trusted the military "a great deal or quite a lot," whereas in 2019, 73% of Americas expressed the same sentiments [@gallup_polls_confidence_2007].
-Unsurprisingly given these developments, compared to other government institutions, today the American public trusts the military to a remarkably high degree.
+Compared to other government institutions, today the American public trusts the military to a remarkably high degree.
 According to a recent Pew Research Center poll, 80% of the American public expects the military to act in the best interests of the American public, compared to business leaders (45%), the media (40%), and elected officials (25%) [@johnson_trust_2018].
 
-While military leaders are rightfully cognizant of the benefits of holding the public's trust, a number of scholars and practitioners have noted that the trend of "military worship" is not only unhealthy [@fallows_tragedy_2014], but likely harmful for democracy [@friend_military_2018].
-Esteeming the military more than it ought can perpetuate, according to Mara Karlin and Alice Friend, a number of harmful myths, including the idea that military service is somehow superior than other, non-military forms of service [@friend_military_2018].
-Over the long haul, a military that is perceived to be more prestigious than its normative civilian superiors can, according to numerous scholars, be problematic for civilian control of the miltiary [@friend_military_2018].
+Coinciding with these developments, scholars began exploring whether there was an unhealthy "gap" forming between American society and the military, especially after the US ended conscription and formed the all-volunteer force in 1973 [@feaver_soldiers_2001; @schake_warriors_2016].
+While there is no definitive answer as to whether a broad "gap" exists, scholars have noted the existence of several smaller "gaps," including a "partisan gap" [@golby_duty_2011] between an officer corps that strongly (though not exclusively) identifies with the Republican party and potential Democratic presidential administrations, and a "familiarity gap" between an all-volunteer military that has endured repeated deployments since September 11, 2001 and American society, especially younger Americans [@golby_thanks_2016; @colford_young_2016].
+A major concern of scholars is that to the degree that these gaps exist, these potentially facilitate the formation of unhealthy attitudes regarding the superiority of military service over and above other types of public service, and further reinforces the military's isolation from society [@fallows_tragedy_2014] while also negatively impacting democracy [@friend_military_2018].
+According to Mara Karlin and Alice Friend, the act of "military worship" - esteeming the military more than society ought to - can, over the long haul, be problematic for civilian control of the military [@friend_military_2018].
 
-Scholars have also shown that when elites possess a high degree of credibility, their ability to influence and shape public opinion is likewise high.
+We further know, as scholars have argued, that when elites possess a high degree of credibility, they possess the ability to influence and shape public opinion.
 In his influential study, _The Nature and Origins of Mass Opinion_, John Zaller argues that formations in mass public opinion are the result of three interacting components: "variation in the information carried in elite discourse, individual differences in attention to this information, and individual differences in political values and other predispositions" [@zaller_nature_1992].
@@
 
-Therefore, as the public's trust in the military has grown in the years since Vietnam, so has the military's potential to impact public opinion.
-This is especially true of military elites, who hold substantial positions of authority.
-An increase in credibility of the institution may not indicate that the military or particular military elites will actually engage in behavior to deliberately impact public opinion, but rather that the potential to do so has increased, as the American people tend to listen to those whom they trust and revere.
+Therefore, as the public's trust in the military has grown, as gaps between segments of society and the all-volunteer force have emerged, and as military service has come to be viewed as somehow superior to other forms of civilian service, the latent power and influence of the military has grown significantly.
+This is especially true of military elites, who hold substantial positions of authority and responsibility.
+However, an increase in the power and influence of the military institution may not indicate that the military will actually use it in particularly harmful ways, but merely that it has the potential to do so, as the American people tend to listen to those whom they trust and revere.
 
@@
 
-The central question I wish to explore in my dissertation is whether and why the US military, or portions of it, have become increasingly politically active, and whether such a shift portends a shift in the military's own "role beliefs" concerning how it relates to American government and society.
-As such, my dissertation will center on fundamental and longstanding debates about the concepts of authority, loyalty, and the proper role of the military with respect to government.
-What is different from previous works in the field is that I wish to explore changes in these concepts, particularly within the US military, over time and in historical perspective.
+The central question I wish to explore in my dissertation is whether and why the US military, or portions of it, have become increasingly politically active, and whether such a shift portends a shift in the military's own "role beliefs" concerning how it should and does relate to American government and society.
+As such, my dissertation will center on fundamental and longstanding debates about the proper role of the military with respect to government.
+I wish to explore changes in these concepts over time and in historical perspective within the United States.
 
-While there have been several vital studies, most of them incorporating original survey data, exploring the beliefs that military members have regarding forms of political behavior [@dempsey_our_2010; @urben_party_2013; @urben_wearing_2014], these studies have not sought to explain these changes.
-Similarly, there have also been a number of recent and important studies regarding the impacts of political polarization on the credibility society places in the military as an instition [@robinson_michael_danger_2018; @robinson_who_2019], as well as a host of impressionistic pieces noting the potentially detrimental impacts of polarization and politicization on the military institution [@feaver_we_2016; @barno_how_2016; @golby_jim_keep_2020; @reid_retired_2020 ].
-However, these studies also have not empirically explored the impact of increasing polarization and institutioncal prestige on the military as an institution over time.
-I wish to close that gap, as I think doing so will help not only scholars, but also practitioners, to include members of the military, understand the answers to some very important questions.
-These questions include:
+While there have been several rich and important studies that explore the attitudes of military officers towards various forms of political behavior [@dempsey_our_2010; @urben_party_2013; @urben_wearing_2014], these studies have not primarily sought to explain why these attitudes have changed over time.
+Similarly, while there have been a number of important studies exploring the impacts of political polarization on the credibility society places in the military as an institution [@robinson_michael_danger_2018; @robinson_who_2019], the dependent variable in these studies was the credibility that society places in the military.
+There have been a number of impressionistic pieces noting the potentially detrimental impacts of polarization and politicization on the military institution [@feaver_we_2016; @barno_how_2016; @golby_jim_keep_2020; @reid_retired_2020], but these works, too, have not sought to empirically explain the impacts of these changes over time.
+It is my opinion that civil-military relations scholars need grounding in theory, specific to the context of the United States, that seeks to explore how rising political polarization and military influence interact, and that such a theory needs to be supported by rigorous empirical work.
+I wish to close this gap, as doing so will enable scholars and practitioners, to include members of the military, to understand the answers to some very important questions, including:
 
+* Has the answer to the question, "what does it mean to be a military professional?" changed over the past several decades, and if so, why?
 * Among whom within the military institution is a shift in normative "role beliefs" occurring, and why?
-* When did such a shift start, and why? 
-* If there is a shift in the military's normative "role beliefs," how will that impact the military's organizational culture?
-* Are some of the most polarized domestic issues compatible with historically normative conceptions of military professionalism, such as the principles of civilian supremacy and a non-partisan ethos?
-why or why not?
+* When did such a shift start, and have there been similar periods in American history?
+* If there is a shift among portions of the military in "role beliefs," how does such a shift impact the military's organizational culture?
+* Are some of the most contemporary polarizing domestic issues compatible with historically normative conceptions of military professionalism, such as the principles of civilian supremacy and a non-partisan ethos?
+Why or why not?
 * What are the key challenges or obstacles to the military regaining and sustaining an effective non-partisan ethos? 
+* Should we expect these challenges and obstacles to endure?
 
-Specifically, I desire to explore the relationship between polarizing domestic politics and institutional credibility on the professional identity and character of the military.
-In the post World War Two era in particular, the military's institutional power, social prestige, and level of credibility has generally grown, making the military a recognized and respected actor within American politics.
-And, especially since the 1970s, political polarization has also increased dramatically, to the point that America is now as polarized as it was at the end of the American Civil-War [@hare_polarization_2014].
-I believe that these two trends, occurring at least in a broad sense simultaneously, have significantly impacted the military's beliefs about its proper role in politics, and that a systematic investigation of this relationship should be undertaken.
-
----
-### Figure this out
-
-In the context of the US military, to my knowledge, loyalty has rarely been used as a conceptual construct in the political science literature, likely and thankfully because the US military has not yet considered, or been involved, in a coup.
-Nonetheless, I contend that there is a rich opportunity to study strategic loyalty as a concept, especially in the contemporary context of the United States' domestic political climate, as political polarization has increased extensively.
-
-I argue that this third challenge is especially relevant for the contemporary military, and specifically, with respect to how members of the military exercise the concept of loyalty.
-As an increasing number of what individuals believe to be core issues are raised and debated in the partisan arena, individuals face decisions of how to prioritize the persons or objects of their loyalty and further, what actions to take on the basis of loyalty to these persons or objects.
-I further theorize (though this has to be tested, of course) that while all individuals will likely face some version of this challenge of loyalty in a polarizing age, the challenges will be more acute for military leaders.
-Not only is the de-facto setting in which military leaders operate marked by multiple and often conflicting loyalties due to the structure of and statutes set forth in the US Constitution, but also, the potential consequences are higher for military leaders than average citizens should they fail to discharge the duties that correspond to their roles as professional officers.
-These consequences include failure to achieve tangible national security goals, such as loss on the battlefield, or less tangible but still important outcomes such as maintaining trust with the American public.
-
-As important as these consequences are, they do not speak to another set of consequences individual military leaders also may face in a polarizing age.
-Indeed, in addition to consequences to the country that may occur should the military fail to discharge its duties to safeguard America, military leaders also potential consequences if confronted with a situation in which they have to execute orders which are technically legal, but morally borderline at best.
-Military leaders are still "moral agents" who "remain responsible for the decisions and actions they take" [@dubik_just_2016, 160].
-But the considerations of what constitutes moral right and wrong to begin with may be changing as polarization worsens [@graham_liberals_2009].
-
-Therefore, I argue that if polarization continues to heighten the identities of groups that people associate with, as well as come closer to the boundaries of how individuals think about moral right and wrong, there is an increased likelihood of military leaders having to choose between conflicting loyalties.
-What happens if and when the objects of a military leader's multiple loyalties not merely disagree, but become stubbornly divided ideologically?
-Can such an outcome exist?
-Is there empirical proof that such a trend is happening?
-If so, what are the normative implications, both for the military as a profession, and for the nation's security, and are there actions that can be taken to halt such a trend?
-
- In the past several years for example, there has been a concerted effort by the services to identify and punish senior leaders who engage in misconduct, to demonstrate to Congress that senior military leaders do not need assistance to handle sexual assault, and to otherwise prove that it can be trusted to accomplish its missions [@us_army_repairing_2014].
-Former Chief of Staff of the Army, Retired General Ray Odierno, told fellow Army generals that "of great importance within the Army Profession is the duty of each of us to be a steward of our profession and to maintain its essential and hard won trust with society [Ray Odierno in @us_army_repairing_2014].
-
----
-
- I am specifically interested in the question of if and how political polarization impacts the concept of loyalty in the military.
-This is an interesting research question because it brings together at least three relevant strands of literatures which, when combined, generates an interesting set of puzzles and questions.
-First, one of the most fundamental normative hallmarks of the military's professional nature, at least by Huntingtonian standards, is that it ought to be an apolitical and non-partisan institution.
-Second, one of the features of the US military is that it prizes and elevates the trait of loyalty among those who serve and, by statute as well as normative tradition, exercises simultaneous dual loyalties to its civilian heads, the President as the Commander in Chief, and Congress.
-And third, we know from the American politics literature that ideological polarization and partisan loyalty are connected in that those citizens who hold to stronger ideological convictions tend to be the most loyal partisans in terms of voting and other political behaviors [@abramowitz_great_2018, 112-116].
-
-If these things are true, then a question is whether rising political polarization will generate instances in which military officers must choose between which objects of loyalty to ultimately elevate and side with in the event that there is conflict between and among these objects.
-Moreover, at the level of the military institution, a similar problem may also emerge: will the military institution have to choose between competing objects of loyalty, such as the President as the Commander in Chief, the Congress, and even its own professional ethic?^[Loyalty had to be a prime consideration in several prominent, recent cases involving the military and the government.
-Consider the case of now retired Army Lieutenant Colonel Alex Vindman, who testified before Congress against the US President, Donald Trump, regarding what Vindman believed to be behavior by the President that may have been illegal, or the case of Navy Captain Brett Crozier, who leaked a letter to the press after his pleas for help in battling the coronavirus aboard the ship he commanded, the USS Theodore Roosevelt, went unheeded for what he felt was too long.
-Certainly these instances involved individuals making choices about to whom or to what they would be most loyal to - the President, the Constitution, the ideals and obligations of the military, the operational mission, themselves as individual moral agents, and perhaps even more objects of loyalty.]
-
-Other questions I wish to explore include whether there patterns in the way that loyalty obligations held by the military are in fact challenged by governmental actors during periods of increased political polarization, and if so, what are they, and how is the military reacting to these patterns?
-Are there impacts on the military's professional identity as a result of political polarization, and if so, from who or what?
-Are partisan/ideological loyalties displacing other loyalty relationships within the military?
-Through what mechanisms do these changes operate?
-What sorts of changes are they driving, if any, in the military institution?
-Is this problematic for the military profession, and why?
-These and other questions will shed light on how rising political polarization is impacting the professional identity of the military.
-
-The gaps in the current literature, which I hope to advance and fill, involve the impact of political polarization directly on the military institution itself, and why the military experiences such impacts.
-A number of excellent and recent scholarly works have looked at aspects of this question, but from an angle that examines not the military directly, but rather society's perception of the military or various aspects of the military as a result of polarized politics.
-A number of these works have directly examined the impact of partisan affiliation on the credibility of military actors, resulting in some very interesting findings.
-Michael Robinson, for instance, found that members of the public did not sanction military elites for making public messages that accorded with a member's partisan views, and that the public only sanctioned those elites whose messages were in line with the opposite political party [@robinson_danger_2018].
-Previously, Heidi Urben found that among active military officers, there was weakening support for the norm that retired military officers should remain apolitical or non-partisan in retirement [@urben_party_2013].
-These scholars' findings suggest, as others have pointed out [@brooks_paradoxes_2020], that political polarization at least coincides with, if not causes, a weakening in support among both the public and among service members for the military's normative apolitical ethos.
-Other scholars have also unpacked a trove of information regarding public opinion and the military, including how partisan affiliation shapes differences in support for the use of military force [@golby_jim_elite_2017] and military elite endorsement of political candidates [@golby_military_2012].
-
-Focusing attention on the military institution will help scholars in the field of civil-military relations and international relations more broadly better understand if and how the military, as a unique body charged with the defense of the nation, changes in fundamental ways as a result of a polarized political climate.
-On one hand, one would expect that it might, in the sense that the American military is, after all, a microcosm of American society.
-Therefore, one could expect that if society truly is becoming polarized, then at least to some degree, so is the military, or at least those who volunteer for service and at the point in time that they join.
-On the other hand, one might reasonably expect that given the unique purpose of the military, combined with its long-standing professional ethic, might insulate the military institution from some of the potentially negative impacts of polarized politics.
-These differing expectations need to be more clearly articulated and explored.
-
-To be clear, scholars have long studied how domestic politics impact the military.
-And as scholars have pointed out, military and civilian relationships are in tension by their very nature in that the leaders of a democracy allow a most undemocratic institution, the military, to protect the polity [@feaver_civil-military_1996].
-And as American history proves, there have been countless instances in which civilian and military leaders disagreed.
-Many of these rise to the level of common knowledge: Lincoln fired McClellan because he lost confidence in him as a wartime Commander; Truman sacked MacArthur because the latter was insubordinate regarding the future of war in Asia; and Obama and McChrystal clashed about troop levels in Afghanistan, ultimately resulting in the former firing the latter.
-Beyond these episodes in which President and wartime commander locked horns, consider several other instances of "civilians" and "military" clashing.
-One episode is certainly the way in which the Vietnam war escalated in ways that, at least according to H.R.
-McMaster, military generals at the time knew were destined to fail [@mcmaster_dereliction_1998].
-A more recent example, and one much studied and commented upon by civil-military relations scholars, is the so-called "Revolt of the Generals" that occurred following the early years of the invasion of Iraq, which involved several recently-retired generals and admirals sharply criticizing Secretary of Defense Rumsfeld's leadership dynamics [@cook_revolt_2008].
-
-My intended research topic is different than these other, still very important topics in which disagreements between the executive and the military occur during wartime.
-It is my assessment that polarization is causing conflict, or certainly has the potential to do so, between the military and civilian leaders, but in most instances, the subject is about topics other than the use of military force, which is somewhat puzzling, given the military's reason d'etre.
-For example, I believe we are seeing portions of the military institution (particular retired military officers) openly disagree with civilian leaders on topics that I believe are, even by the widest of academic definitions, outside the bounds of what constitutes appropriate military expertise.^[I hope to prove this, empirically, in or as part of a chapter in my dissertation]
-
-This raises, then, an important question: does the military's essential purpose risk evolving or changing significantly during polarized times?
-While I believe that these potential developments are catching the attention of civil-military relations scholars, none to my knowledge have specifically set out to systematically explore if and how polarized partisan politics are impacting the military's professional identity and character, and in particular, the role of the military institution itself in facilitating these developments.
-Most scholars acknowledge the potentially damaging impacts of the politicization of the military [@golby_listening_2013; @brooks_paradoxes_2020] at the hands of civilian leaders, and I share these concerns.
-Fewer scholars, if any, are exploring the potential role of the military in politicizing itself, and the mechanisms through which these developments occur.
-
-## Military Professions and Organizational Loyalty
-
-As a subject or concept, loyalty features prominently in the international relations and civil-military relations literatures, often in a comparative setting and in the context of coups [@marcum_overthrowing_2016; @johnson_squeaky_2018], civil wars [@mclauchlin_loyalty_2018], and the difficulties that multi-ethnic states face in maintaining a professional military [@peled_question_1998].
-This literature deals with a host of important issues that stem from problems associated with ethnic identification and loyalty, either to a regime, cause, or leader.
-Though the US context is very different from the contexts usually explored in this literature, its insights may help me identify related dynamics and/or mechanisms that are unfolding in the US: how are changes in partisan or ideological identification (instead of ethnic) as a result of political polarization impacting conceptions of loyalty within the US military, if at all?
-To even begin answering this question, one has to first define what loyalty is and how it works.
-For this fundamental task, the fields of psychology, ethics, management, and sociology are helpful.
-
-Loyalty has been conceptualized as both a virtue [@coleman_problems_2009] and as a moral emotion [@connor_military_2019], the main difference between the two being that the former is a normative aspirational trait, while the latter is a primed behavioral response.
-For the purposes of my dissertation, these differences are not consequential, because both assert that loyalty is something that is ultimately felt by an individual and subsequently reflected in action, such as, in the military context, "the desire to protect each other from harm and to defend each other from attack" [@coleman_problems_2009, 109].
-Oldenquist, a philosopher, further describes loyalty as a devotion or attachment that an individual has to someone or something specific such that one refers to the object of loyalty as "mine" and is inclined to feel "pride when it prospers, shame when it declines, and anger or indignation when it is harmed" [@oldenquist_loyalties_1982, 175].
-Loyalty also generates a willingness in the individual to assume "obligations" [@oldenquist_loyalties_1982, p.
-175] on behalf of the object of one's loyalty, and these obligations can take multiple forms, such as a willingness of the individual to endure hardships and make sacrifices [@fielder_organizational_1992].
-Scholars agree that individuals can hold multiple loyalties simultaneously, and recognize that this often can and does cause problems [@coleman_problems_2009].
-For organizations, and by extension, professions, the concept of loyalty is also very important.
-Conner et al.
-state, for example, that "having loyal members is vital to perpetuating the group" [@connor_military_2019].
-
-We also know that the military services in the United States place a premium on loyalty as a trait and concept.
-Indeed, of the different military services, the US Army in particular has extensively developed, in a normative sense, a framework entitled "The Army Ethic," in which loyalty plays a significant role within the Army profession [@us_department_of_the_army_adp_2019].
-This ethic, which will be described in detail in later parts of this dissertation, is formally laid out and described in the Army's doctrinal publication entitled, _Army Leadership and the Profession, ADP 6-22_, and holds that as "stewards of the Army profession" and as "committed professionals," Army leaders "embrace and uphold the Army Values and standards of the profession" [@us_department_of_the_army_adp_2019].
-There are a total of seven "Army Values," which together spell the acronym, l-d-r-s-h-i-p, which is short for the word "leadership," and the first of these values, loyalty, is followed by six others: duty, respect, selfless service, honor, integrity, and personal courage [@us_department_of_the_army_adp_2019, p.
-1-9.].
-
-ADP 6-22 further states that the Army Values "embody the practical application of the Army Ethic.
-They encompass the enduring moral principles, beliefs, and laws that guide Army professionals in accomplishing the mission as well as their conduct in all aspects of life" [@us_department_of_the_army_adp_2019, p.
-1-12].
-The manual defines loyalty as, "loyalty: bear true faith and allegiance to the Constitution of the United States, the Army, your unit and other Soldiers" [@us_department_of_the_army_adp_2019].
-The virtue of loyalty is likewise extolled in a host of codes and mantras that have played a central part of the American military since its inception.
-In the middle of the West Point Cadet Prayer is the eloquently written plea, "make us to choose the harder right instead of the easier wrong" [@wheat_west_1918], indicating that to be loyal to what is right is not only preferable but often a difficult endeavor.
-Even Samuel Huntington recognized that in order for his model of civilian control to be effective, the military profession had to organize itself as a "hierarchy of obedience," which required "loyalty and obedience" to stand out as "the highest military virtues" [@huntington_soldier_1957, 73].
-
-Loyalty is further important in a military setting in a broad and strategic sense, especially when one considers that the US military, both by statute and by normative standards, serves as institution with multiple, simultaneous loyalties.
-The Constitution of the United States, to which military officers and soldiers alike ascribe allegiance when they take their respective oaths of commissioning and enlistment, contrives the military as an institution possessing multiple loyalties: to the President as "the Commander in Chief of the Army and the Navy of the United States" [@noauthor_us_1787], but also to Congress in that they likewise hold unique powers and responsibilities pertaining to the Armed Forces [@noauthor_us_1787-1].
-The tensions that inherently exist in American's constitutional framework with respect to national security are explored in a robust scholarly literature [for instance, see @howell_while_2007; @milner_sailing_2015].
+Answers to these questions will be of particular importance to furthering our understanding of how the US military has responded to epochs of polarizing politics before, and how changes in levels of institutional prestige impact these responses.
+I suspect that my dissertation will show that any contemporary shift in military role beliefs is not the first shift that has occurred in American history, that such a shift has long been underway, and that particular aspects of today's polarized climate portend that the military will wrestle with the issue of shifting role beliefs for years to come.
+For these reasons, and more, a systematic investigation of the impact of polarizing politics and institutional prestige on the US military's beliefs concerning its role in government should be undertaken.
 
@@
 
-## Political Polarization Impacts Conceptions of Loyalty
+## The Impact of Polarization and Military Prestige on US Military Role Beliefs
 
-In this section of the prospectus, I will lay out an initial argument, as well as accompanying hypotheses, which I intend to explore and test in my dissertation.
-Figure 2, below, shows an initial conception of the linkages between several of the variables discussed in the previous section of the prospectus.
-From left to right, this conceptual model suggests that worsening political polarization is driving changes in both the institutional credibility of the military (how society views the military) and changes in the way members of the military institution think about and exercise loyalty, and by extension, how they think of their professional identity.
-These two independent variables are, I suggest, driving subsequent changes in the concept of military professionalism, the main dependent variable in my analysis.
+In this section of the prospectus, I will lay out an initial argument which I intend to explore and test in my dissertation.
+Figure 2, below, shows an initial conception of the linkages between the variables discussed in the previous section of the prospectus.
+From left to right, the model suggests that political polarization, prestige and credibility, and war (independent variables) all shape how the military views its fundamental role and purpose within the government and society (the dependent variable).
 
-In line with Huntington's model as well as a host of other scholarship, I suggest that changes in military professionalism can impact a variety of military outcomes, varying from effectiveness on the battlefield [@reiter_democracies_2002; @biddle_democracy_2004] to achieving strategic effectiveness [@brooks_creating_2007; @brooks_shaping_2008].
-While I intend to examine the broad system of variables shown in Figure 2, the focus of my dissertation is the relationship between political polarization and changes in the concept of loyalty and identity, and how such changes manifest themselves in the variation of military professionalism.
-For emphasis and clarity, I do not think that the other link - that between political polarization and institutional credibility - is unimportant.
-Rather, this link seems to have been the focus of several recent scholarship efforts [@robinson_michael_danger_2018; ].
+Borrowing from Fitch, I will use the term military "role beliefs," but I will alter the concept he sets forth ever so slightly.
+Whereas he used the term broadly to describe the set of norms, beliefs, and attitudes that the military holds regarding its "role in politics" [@fitch_armed_1998], I argue that we can be more specific in the US context.
+In particular, the "role beliefs" of the military comprise the answer to the question, "what does it mean to be a military professional?"
+
+After surveying the rich normative literature on the topic of military professionalism, to include critiques of the concept, there are, I argue, two central tenets that stand out above others when asking the question, "what does it mean to be a military professional?" These tenets are adherence to the principle of civilian control of the military, and confinement of one's knowledge and skills to relevant military expertise.
+The first principle of course deals with who is in charge of the military, and the second deals with for what purpose the military exists.
+The degree to which the military deviates from one or both of these central tenets indicates a shift in the role beliefs the military holds.
+Figure Two, below, is a visual illustration of the variables of interest in this study.
+They include the dependent variable, military role beliefs, and the independent variables of war, political polarization, and prestige/credibility of the military.
 
@@
 problem_dag <- dagify(
-    MO ~ MP,
-    MP ~ LI + IC,
-    IC~PP,
-    LI~PP,
+    RB ~ PC + PP + W,
+    RB~PP,
+    PC ~ PP,
     coords = tribble(
       ~ name,      ~ x,    ~ y,
-      "MO", 1, .5,
-      "MP", .7, .5,
-      "IC", .3, .3,
-      "LI", .3, .7,
-      "PP", .1,.5,
+      "RB", 1, .5,
+      "PC", .6, .5,
+      "PP", .3, .7,
+      "W", .8,.7,
     ))
@@
 problem_dag %>% tidy_dagitty() %>% 
-  dag_label(labels=c("MO"="Changes in\n Military\n Outcomes", 
-                     "MP"="Military\n Professionalism", 
-                     "PP"="Increasing\n Political\n Polarization", 
-                     "IC"="Institutional\n Credibility",
-                     "LI"="Concepts of\n Loyalty and Identity")) %>% 
+  dag_label(labels=c("RB"="Military Role Beliefs", 
+                     "PC"="Prestige and Credibility", 
+                     "PP"="Political Polarization",
+                     "W"="War")) %>% 
   ggplot(aes(x=x, y=y, xend=xend, yend=yend))+
@@
   geom_dag_text()+ 
-  geom_dag_label_repel(aes(label=label, fill=label), col="white", show.legend=FALSE)+ theme_dag() + labs(title="Figure 2", subtitle="The Impact of Political Polarization on Military Professionalism")
+  geom_dag_label_repel(aes(label=label, fill=label), col="white", show.legend=FALSE)+ theme_dag() + labs(title="Figure 2", subtitle="Determinants of US Military Role Beliefs")
 
@@
 
-A primary contention from the system visualized above is that as political polarization in the domestic sphere increases, the military faces increased challenges in how both individual military leaders as well as the institution conceptualizes, adopts, and exercises loyalty, and by extension, how it sees its own professional identity.
-This expectation stems from several features that scholars of American politics have raised regarding the current politically polarized climate within the United States.
-First, scholars have noted that political polarization raises the number and scope of issues that are subsumed by partisan debate and disagreement (a phenomenon referred to among scholars of American Politics as "conflict extension" [@layman_party_2002]).
-A second feature of contemporary political polarization is the prominence of  negative feelings held by one partisan group towards another, a phenomenon referred to as "negative partisanship" [@abramowitz_great_2018].
-A third feature is an increase in the salience of debates over issues that individual partisans consider core to their individual identities [@white_values_2002; @mann_its_2012; @sides_identity_2018].
-I contend that these features impact individuals serving in the military, including military leaders, as they do members of the domestic public writ large, and that these impacts are problematic for members of an institution that normatively exists as an apolitical and non-partisan actor.
-Table 1, below, lays out how these three prominent features of domestic political polarization may impact the non-partisan and apolitical nature of the normative military professional and institution.
+Beginning with the impact of war on military role beliefs, I have chosen to draw an arrow directly from war to role beliefs, and not from war to political polarization, or vice versa.
+There is sound theoretical expectation that while the military is actively engaged in fighting a conflict, they will focus on prevailing in that conflict, and thus, that active combat has a dampening effect on the military's propensity to inappropriately intervene in politics.
+Such an argument follows a similar (though not identical) argument made by Michael Desch, who theorized that civilians would be in a better position to control the military when external threat levels are high [@desch_civilian_2001].
+
+War, vice threat, is shown in this model, however, following recent scholarship that argues that threats in themselves hold little if any unifying power, and thus, do not attenuate polarization levels [@myrick_external_2020].
+Other scholars have argued that combat operations and war tend to focus the military inward.
+Don Snider, for instance, argues that military institutions tend to diverge away from their professional nature as war fighters and towards that of a bureaucracy as combat operations wind down and as peacetime ushers in changes such as force reductions and a return to training for combat [@snider_once_2012].
+I follow Snider's expectation and argument, and conclude that war is the appropriate variable to show here.
+
+In terms of the two central tenets of the professional nature of the military, I argue that war can theoretically impact both, but its main impact is on expertise.
+Winning in combat demands that the military develop and implement unique military expertise to prevail in war.
+Conversely, winning in combat does not mean that the military will necessarily adhere to the principle of civilian control, as American history aptly demonstrates (e.g., Truman and MacArthur), but the presence of active combat does predict that military intervention in politics during wartime at the hands of the military will likely center on the conduct of the war itself, and not on domestic political matters that fall outside of military expertise.
+
+The next variable to theoretically unpack is political polarization, and admittedly, this is a difficult endeavor.
+However, previous literature in American politics enables the development of two plausible relationships and corresponding mechanisms, each stemming from a particular conception or definition of polarization.
+If one conceives of and focuses on polarization as the ideological widening mainly between political elites, such as members of Congress [@hare_polarization_2014], then one possible mechanism by which the military intervenes in politics is by adopting a belief that elites from both dominant political parties have failed to advance and implement an effective governing agenda.
+Thus, the military steps in to do what political elites could not do.
+Scholars of comparative politics have noted such a dynamic occurring among militaries in other parts of the world [@fitch_armed_1998].
+
+A second, and I believe the more likely, relationship between political polarization and military role beliefs, particularly in the contemporary United States, results from a conception of polarization as the strength of feelings concerning a number of ideological issues.
+This is a view of polarization that is consistent with the concept of affective polarization, or "negative partisanship" [@abramowitz_great_2018].
+If the variables of polarization and military role beliefs are linked in this way, the military would not intervene as a monolithic block due to a frustration with the political class's inability writ large to effectively govern.
+Rather, certain military elites, who share beliefs with a polarized electorate on any number of ideological issues, such as race, gender, and immigration [@abramowitz_great_2018], will selectively intervene in order to advance these very beliefs.
+In other words, this mechanism posits that certain military elites feel as strongly as average citizens about certain issues, and that they will break from historical apolitical or non-partisan military traditions and norms in order to advance these causes.
+
+This conception adopts a view such that despite the military organization's emphasis on apolitical traditions, beliefs, and norms, the forces of political polarization and the ideologies that underlie it have the potential to prove stronger.
+Additionally, political polarization may impact both of the tenets of military professionalism (adherence to the principle of civilian control, and adherence to relevant military expertise).
+High polarization may impact the principle of civilian control because specific military leaders may hold differing ideological beliefs than their civilian bosses, and it may impact expertise because certain military leaders may adopt a view that they can perform the functions of their civilian leaders better, even if these functions fall well outside the confines of what traditionally constitutes military expertise.
+
+In short, military leaders are experiencing and reacting to the very same phenomena that scholars of American politics have identified as impacting American society more broadly: the phenomena of "conflict extension" [@layman_party_2002] and "negative partisanship" [@abramowitz_great_2018], and the fact that polarizing domestic disputes involve so many issues that individuals consider a "core" interest or value [@white_values_2002; @mann_its_2012; @sides_identity_2018].
+As a result of these dynamics, political polarization has provided the _"motive"_ for certain military leaders to intervene in domestic politics, which is exactly what Finer warned us about 58 years ago [@finer_man_1962].
+When polarization is not high, there simply is less of a motive for the military, to include military elites, to deviate from their historical non-partisan and apolitical traditions.
 
@@
 \begin{footnotesize}
-\caption{How Features of Contemporary Political Polarization Impact Normative Military Professionalism} % title of Table
+\caption{How Features of Contemporary Political Polarization Impact Military Role Beliefs} % title of Table
 \centering % used for centering table
@@
     \hline
-    \textbf{Features of Increasing Political Polarization} & \textbf{Associated Mechanism} & \textbf{Potential Impact on Military Professionalism}  \\ \hline
+    \textbf{Features of Increasing Political Polarization} & \textbf{Associated Mechanism} & \textbf{Potential Impact on Military Role Beliefs}  \\ \hline
     \textbf{Conflict Extension} & Less politically-neutral space exists as partisan disagreements subsume a greater number/scope of issues.
-& Military Professionals and the Institution find it increasingly difficult to "stay out" of partisan politics as the territory of what is considered partisan politics grows.
+& Military Professionals and the Institution find it increasingly difficult to "stay out" of partisan politics as the territory of what is considered partisan politics expands.
 \\ \hline
@@
 There is a belief that those on the other side of the political aisle are not just mistaken over policy, but greatly wrong and flawed as individuals.
-& Even among military leaders, there is a greater willingness to confront or take issue with those who hold opposing views.\\ \hline
+& Even among military leaders, there is a greater willingness to confront or take issue with those who hold opposing views.
+Like average citizens, military elites feel they should speak up and against those whom they feel hold threatening political views.\\ \hline
     \textbf{Individuals believe that "core" issues are increasingly at stake} & The salience of one's identity and core values is raised as partisans believe essential issues are at stake in the political arena.
 Associated with this is a belief that not just policy preference, but moral right and wrong, are at stake.
-& Military professionals, who hold multiple institutional loyalties, face an increasingly difficult task of balancing these loyalties, especially as the objects of these loyalties are increasingly associated with or attached to notions of moral right and wrong.
+& Military professionals face an increasingly difficult task of balancing the multiple loyalties they owe, especially as the objects of these loyalties are increasingly associated with specific ideological or partisan causes, and are attached to notions of moral right and wrong.
 \\ \hline
@@
 
-## How Polarization Interacts with Institutional Credibility
+Figure Two also argues that the level of prestige and credibility afforded to the military also impact military role beliefs.
+Here, I view prestige and credibility in both an _institutional_ or _organizational_ sense, as well as an _individualistic_ sense.
+It is important to distinguish the two because of the variation that can occur within each.
+For example, General (Ret.) David Petraeus enjoyed an outsized level of personal prestige and credibility during the height of the surge in Iraq, whereas Lieutenant General (Ret.) Ricardo Sanchez likely lost, at least temporarily, some individual prestige and credibility after being removed from commanding US forces during the first year of Operation Iraqi Freedom, and as allegations of detainee abuse at the hands of US forces at the abu-Ghraib prison, near Baghdad, surfaced [@freed_last_2011].
+Both officers, however, served in their positions at a time when the US military commanded a relatively high level of institutional prestige and credibility relative to other US institutions [@gallup_polls_confidence_2020].
 
-The previous section described how increasing political polarization presents significant challenges to aspects of the normative concept of military professionalism.
-But whether these challenges enter the public domain, or remain private challenges that specific individual actors contend with on a personal level only, matters a great deal because it "changes the character...from a matter of principle to a political matter" [@dubik_principled_2014] In fact, in much of the field of civil-military relations, this dynamic of private disagreement or discontent potentially entering the public realm is seen in the scholarly literature on senior officer resignation.
-This debate examines various aspects of resignation, including whether there can ever be conditions under which resignation is justified or appropriate, with scholars taking various sides on the issue, both for [@dubik_principled_2014; @snider_strategic_2014] and against [@feaver_should_2014; @golby_beyond_2015].
-I contend that polarization and credibility of the military interact, however, in at least two important ways.
-First, as loyalties are challenged as a result of increasing political polarization, individual military leaders may not necessarily remain quiet despite the existence of a strong professional norm of doing so because the objects of loyalty may demand that their adherents publicly speak out.
-This of course, depends specifically on what the object of loyalty is and what it demands from those who hold closely to it.
-The second way in which polarization and credibility may interact arises more directly from what can be termed a credibility effect, and has more to do with the place of prominence (or lack thereof) that the military or individual leaders hold in the perception of American society.
-As recent scholarship has shown, the credibility of the military institution in a variety of contexts depends heavily on partisan affiliation [@urben_party_2013; @golby_listening_2013; @golby_jim_elite_2017; @robinson_danger_2018].
+Whereas high political polarization provides the "motive" to intervene, I contend that high levels of prestige and credibility provide the other missing link - the _"opportunity"_ - to do so [@finer_man_1962].
+High levels of prestige and credibility, often correlated with high experience levels in a specialized area, serve as latent power to influence a number of important audiences, to include the general public  [@zaller_nature_1992], key government figures [@saunders_no_2017], and other organizations and institutions.
+Zaller argues that most citizens, as non-elites, must rely heavily on the messages that come from elites regarding complex topics and issue areas [@zaller_nature_1992].
+Zaller further argues that the messages that elites send is necessarily "stereotyped" in some way: that is, the messages that elites send must be "short, simple, and highly thematic" [@zaller_nature_1992].
+In developing this argument, Zaller is not claiming that regular citizens are uninformed or careless people who are incapable of processing complex information, but rather acknowledging that most citizens care most about a world that is not "beyond their personal experience" [@zaller_nature_1992].
+Thus, his main point is simply that elites play a significant role in forming and swaying public opinion.
+Therefore, when the military institution or when military elites command a high degree of prestige and credibility, they are in a position to heavily influence those around them.
+Therefore, high prestige and credibility can, like political polarization, impact both the military's adherence to the principle of civilian control and its willingness to stick to relevant military expertise.
 
-I contend that when institutional credibility is high, a demand signal is created by the American public such that future interactions between the military institution and the American public are encouraged and welcomed.
-From an institutional perspective, the military might respond to the public's demand signal for increased military interaction by increasingly participating in public ceremonies, parades, and other showcases of military prowess (at sporting venues, for instance).
-From an individual perspective, military elites, especially retired military officers, would be psychologically primed to increasingly dialogue with the American public when military credibility is high.
-By extension, other outlets such as the press would also be primed to showcase military elites when military credibility runs high, because these elites represent an institution that commands a high degree of public respect.
+What is less straightforward, however, is unpacking how political polarization impacts the levels of prestige and credibility of the military institution, as Figure Two illustrates.
+While it is theoretically possible that as polarization rises, the public views the military more favorably relative to other institutions precisely because of its non-partisan ethic and traditions, recent scholarship has shown that the public views and responds to specific military actions almost entirely along partisan lines [@golby_listening_2013; @golby_jim_elite_2017; @robinson_michael_danger_2018].
+This has further complicated the military's role in politics, as it essentially means that one's partisan identity heavily conditions one's opinion of the military, and in particular, any incident that involves the military's politicization: if the military or a particular military leader says or does something that aligns with one's ideological views, that particular individual is likely to view the military favorably, while the opposite is also true [@obinson_michael_danger_2018].
 
-When military credibility is low, the theory contends the opposite will occur.
-Since demand for interactions between the American public and the military is muted when institutional credibility for the military is low, the institution will not engage with the public as frequently nor as intimately because the strong demand signal that existed when institutional credibility was high no longer exists.
-At the individual level, military elites will also not engage as often with the public, nor will the press ask these elites to.
-In sum, institutional credibility and respect for the military institution impacts the likelihood of the military institution and military elites interacting with the American public.
-Such interaction can take a variety of forms.
-When respect for the military is high, the military is likely to interact, speak to, and dialogue with the American public, and when institutional credibility is low, the military is less likely to do so.
+This dynamic may impact military role beliefs in important ways.
+The military has long cherished its ability to hold the public's trust.
+In fact, former Army Chief of Staff, General (Ret) Ray Odierno used to say, "trust is the bedrock of our profession" @us_army_repairing_2014].
+But if the degree of prestige and credibility that the military commands splinters along partisan or ideological lines, the military may believe that it should or must act to repair that gap in trust.
+Yet doing so, especially along partisan or ideological lines, could cause even more divisions with opposing partisan groups, and the spiral in credibility among segments of the public may continue.
+Therefore, it is reasonable to conclude that while polarization may impact military prestige and credibility, it is less clear to conclude precisely how so.
 
-Ultimately, four possible worlds exist when the impacts of military credibility and political polarization are considered in tandem (see Table 2, below).
-The first is when political polarization and military credibility are both low.
-When these conditions hold, there is no anticipated drop in military professionalism by the military institution nor its individual elites.
-With these conditions, the military does not have an out-sized platform to interact with the American public, nor are there outstanding incentives for the military to raise significant disagreement over any domestic political issues.
+When the levels of military prestige and credibility are high, the American public approves of the military's performance and behavior, and thus, there is somewhat of a demand signal for the military to continue doing whatever it has been doing.
+In the contemporary environment, this might equate to the military continuing to interact or dialogue with the public in a number of ways, such as military participation in public ceremonies, parades, or other high visibility events, or from an individual perspective, though the press, social media, or other outlets.
+When military credibility and prestige are low, there is less of a demand signal from the public to the military institution or to particular military elites, and thus, there will be fewer interactions between the military or military elites and the public, and thus, fewer opportunities for the military to extend their influence.
+
+## When Polarization and Prestige Collide
+
+With the view that rising political polarization in the US may provide the "motive" and that increasing prestige and credibility may offer the "opportunity" for military intervention in politics [@finer_man_1962], it is now possible to determine how the presence of these factors combine to actually manifest in the central role beliefs that militaries adopt.
+Ultimately, four possible worlds exist (see Table 2, below).
+The first is when political polarization and military prestige are both relatively low.
+When these conditions hold, there is little if any expectation that the military will deviate from its non-partisan ethos.
+Because polarization is low, there are few if any ideological issues or concerns that may serve as a motive for military intervention.
+Further, owing to a low degree of prestige and credibility, the military does not have a high degree of latent power, and thus, its ability to have an outsized influence on the public, government figures, and other organizations and institutions, is minimal.
 
@@
 \begin{footnotesize}
-\caption{Four Types of "Military Professionalism Worlds" Where Institutional Credibility and Political Polarization Vary} % title of Table
+\caption{US Military Role Beliefs: A Function of Institutional Prestige and Political Polarization} % title of Table
 \centering % used for centering table
@@
     \hline
-    Influence of Variables & \textbf{Low Institutional Credibility} & \textbf{High Institutional Credibility}  \\ \hline
-    \textbf{Low Political Polarization} & Negligible anticipated impacts.
-Traditional Professionalism is Expected by the Military, Civilian Leaders, and the American Public & Possible Impacts.
-There is a demand for military voice, but little substantive matter stemming from US domestic politics that can cause crucial disagreement.
-Little incentive to deviate from traditional concepts of loyalty and duty.
+    Influence of Variables & \textbf{Low Prestige / Credibility} & \textbf{High Prestige / Credibility}  \\ \hline
+    \textbf{Low Political Polarization} & Unlikely deviation from an apolitical and non-partisan ethos.
+Traditional notions of military professionalism are expected by the military, civilian Leaders, and the public.
+The military does not hold outsized power to influence the public or those within government, and there are few if any issues that motivate the military to deviate from its traditional roles & Little deviation from an apolitical and non-partisan ethos.
+The public respects the military institution and there is a demand for military voice, but there are few polarizing issues in domestic politics.
+The military may weigh in on matters, but it will focus on national security issues that directly impact the military.
+There is little incentive for the military to take visible stances on domestic topics that do not impact the military directly.
 \\ \hline
-    \textbf{High Political Polarization} & Possible Impacts.
-Polarization raises the stakes of domestic politics, but military professionals are constrained from voicing their opinions and/or interfacing with the public due to low institutional credibility with the American public &  Highly Likely Impacts to Professionalism.
-Likely impacts.
-Military professionals have a "platform with the public" due to high institutional credibility, and high polarization raises the likelihood of disagreement between individual military elites and civilian leaders on a number of crucial topics.\\ \hline
+    \textbf{High Political Polarization} & Moderate deviation from an apolitical and non-partisan ethos.
+Polarization provides a motive for the military to intervene in domestic politics, but the military is constrained from holding an outsized influence in debates on these matters due to its low institutional prestige and credibility &  Highly Likely Deviation from an apolitical and non-partisan ethos.
+Polarizing issues provide a motive for the military or certain elites to intervene, and high levels of prestige and credibility make such intervention possible\\ \hline
 \end{tabular}
@@
 
-The second possible set of conditions is when political polarization is low and institutional credibility is high.
-When these set of conditions exist, the military has a platform to interact with the public, as the public highly respects the military institution.
-The military enjoys this high level of trust and does not want to lose it.
-Further, because the degree of political polarization is low, there is little incentive for the military as an institution or military elites to use the trust they enjoy with the American public to adopt or denounce an issue associated with any partisan entity.
-The high institutional credibility increases the chance of the military interacting with the public in an unprofessional way, but the low political polarization removes the incentives of doing so.
+The second possible set of conditions is when political polarization is low and institutional prestige is high.
+When these set of conditions exist, we can expect little deviation by the military from its traditionally non-partisan ethos.
+The military has the potential to wield an outsized influence on the public and within the government, but because the degree of political polarization is low, there is little opportunity nor incentive for the military to adopt or denounce an issue associated with any partisan or ideological cause.
+When these conditions are present, the military may be active, but likely on matters that directly relate to American national security.
 
-The third type occurs when political polarization is high, but institutional credibility of the military is low.
-Under these conditions, the military does not command the respect of the American public like it does when credibility of the institution is high, and thus, does not possess a considerable outlet to interact with the American public.
-However, the incentives of loyalty and duty still loom large in the minds of some military elites as a result of the high degree of political polarization, and thus, a dip in professionalism is still possible.
+The third general set of conditions occurs when political polarization is high, but institutional credibility of the military is low.
+Under these conditions, there is a moderate expectation that the military will deviate from its non-partisan and apolitical traditions.
+The military does not command the respect of the American public to the degree that it does when the prestige of the institution is high, but because polarization levels are high under these set of conditions, certain military leaders may be motivated to intervene in politics to advance certain ideological or partisan causes.
+There may be advancement of these causes by certain military elites, but such activities are not likely to feature prominently in the public eye.
+This may involve more "behind the scenes" activities by particular members of the military or military elites.
 
-The final set of conditions exists when political polarization and institutional credibility are high.
-It is in this world that I contend the American military finds itself in now in 2020.
-Under these conditions, the military faces dual dangers.
-High institutional credibility risks the military institution unilaterally acting or another body co-opting the military institution to speak out an a number of issue areas, and simultaneously, the highly polarized environment risks the military advocating for or denouncing partisan sentiments in the name of upholding the very virtue it is most normatively adhered to - loyalty.
+The final set of conditions exists when political polarization and institutional prestige and credibility are both high.
+Under such conditions, the military is highly likely to deviate from its apolitical and non-partisan norms, and thus, to intervene in politics.
+The high degree of prestige the military institution enjoys means that it possesses the ability to influence a number of important audiences by its words and actions, and the high levels of polarization indicate that there are motives to act.
+It is under this dangerous set of conditions, I argue, that the American military finds itself in now in 2020, as it did at the end of the American Civil War.
 
-# The Analytical Framework - Hypotheses
+# Hypotheses, Observables, and Plan of the Dissertation
 
-In order to test my argument, I will have to examine and test several hypotheses, which are inter-related and linked. They include the following:
+In this section of the prospectus, I will elaborate on what a political scientist would expect to see given the theory advanced in the preceding pages, and describe the overall plan for my dissertation.
+In short, I plan to implement a mixed-methods approach that incorporates quantitative and qualitative analysis, as well as data from an original survey experiment, over the course of what I have now planned to take seven chapters.
+My dissertation will hinge on the examination of several hypotheses which must be examined, tested, and evaluated, as doing so will enable me to strengthen the overall argument I wish to make.
 
-\singlespace
+In broad terms, I need to look for evidence that will confirm or, alternately, disprove, my theory.
+I have made the central claim that changes in the way a professional military involves itself in politics, and thus, changes in its central role beliefs, will challenge one or both of two central guiding principles: the principle of civilian control of the military, and confinement of knowledge and advice to the realm of relevant military expertise.
+The theory advanced thus far further argues that although the presence of active war limits the military's deviation from professional role beliefs, two other domestic factors, political polarization and military prestige/credibility, provide what Finer previously identified as the "motive" and "opportunity" for the military's intervention in politics [@finer_man_1962].
+Therefore, I need to show that political polarization, and not another variable or set of variables, truly provides sufficient incentive for military leaders to deviate from non-partisan or apolitical traditions.
+I also need to show that it is the level of military prestige and credibility which affords military elites the opportunity to engage in politics inappropriately, and not some other variable.
+As a scope condition, I also need to find evidence that the presence of war truly does limit or constrain the military's ability or desire to intervene inappropriately in politics by the way that I have, building on other scholars, argued: that active war keeps the military focused on predominantly military expertise.
 
-_H1: In highly polarizing times, there will be an increase in the frequency of military officers breaking normative civil-military relationship behaviors, such as criticizing the President, members of Congress, or those who hold to different political views on particularly-polarizing issues outside of the conduct of war._
-\doublespace
-\singlespace
-
-_H2: These breaks in normative civil-military behaviors will occur because political polarized periods feature events or situations which test or challenge the military's ability to juggle and balance multiple simultaneous loyalties: to the President as Commander in Chief, to Congress in its oversight role of the military, to the underlying ideologies or ideas inherent in issues that prevail during polarized times, to the normative professional ideals of the military, and to the individual's bonds of loyalty to his or fellow soldiers and the mission._
-\doublespace
-\singlespace
-
-_H3: Because of the normative strength and prominence of place that the ideal and concept of loyalty has within the profession of arms, military leaders are more sensitive than regular civilians to appeals of loyalty._
-\doublespace
-\singlespace
-
-_H4: Following from the above hypothesis, polarized periods will feature events in which governmental actors appeal to the military's sense of loyalty in order to persuade them to adopt a stance on a particular issue area.
-These can be viewed as "exploits of loyalty."_
-\doublespace
-\singlespace
-
-_H5: In highly polarized time periods, it is the senior military officers as the "stewards of the profession" [@us_department_of_the_army_adp_2019, p.1-9] whose actions will most influence to what degree and how the military institution reacts to and changes as a result of "exploits of loyalty." If changes occur to the military institution, they are bound to occur to the professional identity of the military, which includes the degree to which it functions as a profession (vice bureaucracy [@snider_once_2012]) and the degree to which it is obedient to its civilian leaders (vice acting as an activist on behalf of a particular cause)._
-\doublespace
-
-# Plan of the Dissertation and the Data Generating Process
-
-To complete my dissertation, I plan on incorporating a mixed-methods approach.
-In broad conceptual terms, I plan to propose a theory and then test the theory using a combination of quantitative analyses and qualitative case studies.
-The one sentence thesis of my dissertation is that political polarization significantly impacts how the military thinks about and exercises loyalty, a chief military trait, and that these impacts can and often do result in changes to the military's professional identity and character.
-I plan to execute a book style project comprised of seven chapters.
+## Chapter 1-2 : Introduction and Theory
 
@@
 This will involve synthesizing several of the key aspects and debates within the field of civil-military relations, much in the same way I have done in the first part of this dissertation prospectus.
-I will focus my attention on introducing and summarizing the concepts of military professionalism, organizational loyalty, and political polarization, and spend time orienting the reader to the debate within scholarly literature regarding a Huntingtonian conception of military professionalism.
+I will focus my attention on orienting the reader to debates regarding a Huntingtonian conception of military professionalism.
 I will emphasize the long-running nature of this debate going all the way back to when Huntington first published his book, but point to some of the prominent contemporary critiques regarding his theory [@feaver_civil-military_1996-1; @rapp_civil-military_2015; @golby_why_2017; @karlin_case_2020; @brooks_paradoxes_2020].
-The main point of chapter one is to set the stage for my dissertation and signal to the reader that modern critiques of Huntington may be overlooking a key part of his concept of military professionalism: imperfect as it may be, it may also insulate the military from the most pernicious effects of polarizing politics.
+In chapter one, I will also introduce the growth of political polarization and military prestige in the latter half of the twentieth century in order to motivate the fundamental question of my dissertation.
 
-In the second chapter, I will introduce, develop, and explain my theory that links increasing political polarization to conceptions of loyalty and military professionalism (see Figure 2, above).
-I will argue that specific features of the polarized environment strain and challenge both individual and institutional conceptions of loyalty, and by extension, normative aspects of military professionalism (Table 1).
-I will further argue that rising political polarization can interact with changes in the military's institutional credibility to increase the likelihood that the public becomes aware of how the military chooses to respond to the dynamics brought on as the military faces challenges to how it conceptualizes and practices loyalty.
-Using Table 2 as a framework, I will place epochs of historical American civil-military relations into one of four distinct quadrants, and signal that these quadrants will comprise the qualitative case studies that I later examine in my dissertation.
+In the second chapter, I will introduce, develop, and explain a theory that links increasing political polarization and military prestige to the role beliefs that militaries adopt.
+Much in the same way that I have summarized in the theory section of this prospectus, I will emphasize that I am building off of the theories advanced by previous scholars, such as Finer and Fitch, and applying these theories to the 21st century US military.
+Using Table Two as a framework, I will argue that to determine whether the military will stray from its traditional non-partisan and apolitical traditions, the key variables to consider are war, polarization, and prestige, and that the key observables are how the military thinks and behaves about two central tenets: the principle of civilian control of the military, and adherence to relevant military expertise.
+When war is present, the military will likely focus on winning that war, and developing and focusing on the military expertise to do so.
+When high levels of polarization and prestige are present, however, there is a strong likelihood that certain military actors may use the credibility they enjoy to advance certain ideological causes, potentially bucking both the principle of civilian control and straying outside of relevant military expertise.
 
-In the third chapter, I will argue that today, in 2020, the American public is witnessing a decline in professional military conduct among retired military officers.
-I plan to demonstrate this empirically by drawing upon an original data set consisting of more than 340 op-ed pieces authored by retired military officers since the late 1970s and published in the top five major print newspapers in the United States, including the Wall Street Journal, the New York Times, the Washington Post, the USA Today, and the Los Angeles Times.
-I believe that I can show several things from this analysis, including that there has been a sharp uptick in the number of op-eds published by retired military officers during this time period, and more relevant to the theory I propose in the dissertation, that these published op-eds reveal a decline in the normative professional ideals of the military institution.
-Unlike previous recent works which focus on how polarization impacts society's view of the military, this chapter introduces an empirical record of actors within the military institution (retired senior officers) also changing as a result of increasing political polarization.
+## Chapters 3-5: The Heart of the Empirical Analysis
 
-Critical to this chapter is acknowledgement that coding for what is professional versus what is not is a somewhat subjective process.
-Thus, I will need to develop and implement a clear working definition of the concept of military professionalism in Chapter 1.
-However, I believe that I can show in Chapter 3 that whether one adopts a strictly Huntingtonian version of military professionalism (strict separation of political and military spheres) or a less strict version that allows for an increasing shared space, the retired military officer corps is not living up to the non-partisan ethic that the military espouses.
+Chapters 3-5 will comprise the core of the empirical analysis of my dissertation.
+In Chapter 3, I plan to examine variation in opinion-editorial pieces published by retired military generals and admirals in major newspapers since the end of the Vietnam War.
+The hypotheses that I will examine and test in this chapter are as follows:
 
-I will argue that while there are certainly other confounding variables, to include changes in the conservative-liberal tilt of newspapers over time, the chief explanation for the variation (decline) in professionalism stems from the combined effects of increasing political polarization and rising institutional credibility of the military since the late 1970s, and the challenges that these changes introduce to a military leader's conception of loyalty.
-Moreover, my ultimate point is to point out a blatant inconsistency in the way that the military currently thinks about the political behavior by those within the ranks in that the institution has rules that governs the conduct of those on active duty [@department_of_defense_political_2008], but those same rules, and the spirit behind those rules, are disregarded by the institution's vanguards, its retired multistar officers and "stewards of the profession" [@us_department_of_the_army_adp_2019, p.
-1-9] of arms.
-I believe that this chapter will result in finding support for _H1._
+_H1: When political polarization is high, the military (or certain military figures) will increasingly criticize civilian leadership (indicating a drop in the principle of civilian control)._
 
-In Chapter four, I plan to analyze the results of an original survey experiment in which I survey a nationally representative sample of civilians and compare the results to a sample of either a) veterans, or b) active duty military officers (pending a lengthy approval process through the Army).
-I intend to design a survey experiment that measures whether arguments that appeal to one's individual or collective sense of loyalty are more effective than those messages that do not incorporate such language, and more directly, whether such arguments are acutely effective among military audiences.
-Such an experiment will build off of recent surveys that have examined the impacts of "moral reframing" [@feinberg_moral_2019] and "moral messaging" [@luttrell_challenging_2019], which describe efforts senders of messages employ to persuade their audiences by appealing to the audience's sense of morality.
+_H2: When political polarization is low, the military (or certain military figures) will publicly weigh in on traditional national security debates (and not on domestic political concerns or issues outside of relevant military expertise)._
 
-Moral reframing is an extension of the broader _Moral Foundations Theory_, a paradigm developed by psychologists several years ago [@graham_liberals_2009] that seeks to explain the moral domain in terms of five distinct moral foundations: "harm/care, fairness/reciprocity, ingroup/loyalty, authority/respect, and purity/sanctity" [@graham_liberals_2009].
-Several follow on studies have confirmed that political liberals and conservatives rely on these five moral foundations in profoundly different ways, particularly when making "moral judgements" [@graham_liberals_2009]: liberals tend to rely on a morality based primarily on the foundations of harm/care and fairness/reciprocity, while conservatives tend to rely on a "more even distribution" of the five values, to include the "binding" foundations of ingroup/loyalty, authority/respect, and purity/sanctity [@graham_liberals_2009].
-My intended survey experiment will advance the field because it will explore the degree to which one particular moral foundation - ingroup/loyalty - resonates within an organization that emphasizes and values this particular moral foundation, as I have laid out in the preceeding pages.
-To my knowledge, while there have been many studies that examine the extent to which partisans rely on differing moral foundations [need cite], there have not been studies that examine the degree to which one particular moral foundation binds members of an organization.
-The military serves as a potentially very interesting case study, particularly concerning the moral foundation of ingroup/loyalty.
+_H3: During periods of high prestige and credibility, the frequency of military involvement in quasi-political activities (such as writing op-eds) will increase relative to periods of low prestige and credibility._
 
-Therefore, I will test whether an appeal to loyalty works stronger among members of the military than among members of the civilian population, as I have hypothesized in _H3._ The survey, after controlling for gender, partisan identification, and age (experience for military), will present the respondent with 6-8 hypothetical situations and a corresponding argument made by military voice (the biographical particulars of which I will also vary).
-Members of the treatment group will receive an appeal to loyalty as part of the military voice's argument as to why a certain course of action should be followed regarding the hypothetical situation presented.
+_H4: When both political polarization and high prestige are present, the military will be increasingly vocal and at higher frequency about certain domestic ideological or partisan causes than when these factors are not present._
 
-The survey will be structured such that the scenario and corresponding course of action contain references to issues that are, at least in some of the cases, highly correlated with partisan / ideological affiliation.
-I will also vary the treatment such that while each treatment will be an appeal to loyalty, the specific object of loyalty within the treatment condition will be varied.
-For example, some questions will appeal to the military member's sense of loyalty to the Constitution, while another may appeal to the military member's sense of loyalty to the mission and/or the member's fellow soldiers.
-An example of one of these questions is as follows:
+_H5: During periods of active ground combat, military elites will likely comment on the war, but with less frequency on other domestic issues._
+
+To test these hypotheses, I plan to conduct a quantitative analysis of more than 300 opinion-editorials published by retired military officers in the New York Times, Wall Street Journal, and Washington Post since the early 1970s.
+I plan to hand code these op-eds and classify each according to a categorical subject type (such as whether the piece covers a topic involving foreign policy, war fighting, military budgets, institutional military culture, social policy (such as who serves in the military), or domestic policy (not directly related to the military), and to also code for several dichotomous variables, such as whether the op-ed sharply criticizes a partisan figure (President or other named official), whether the op-ed clearly adopts a partisan stance on an issue, and whether the op-ed strongly calls for definitive legislative or political activity (such as Congress funding an initiative or signing a treaty).
+
+The dependent variables in my analysis will include the frequency of certain subject types of op-eds (such as domestic policy) and the frequency of the dichotomous variables outlined above.
+Additionally, I will include controls for the level of polarization present in Congress (using DW-NOMINATE scores) in each year of my analysis as a proxy for the intensity of polarization present in society, and the level of military prestige (measured by Gallup public opinion polling that measures the level of trust in the military institution by American society)^[I may also use alternate measurements, such as measures of affective polarization, which will involve measuring the difference in presidential approval ratings between Republicans and Democrats for each year of my analysis.] Because the data covers the past 45 years, from the 1970s to the present day, the data allow for one to examine variation of a quasi-political behavior by the military (writing op-eds) while holding the sources constant across a wide period of time.
+Moreover, the period of time examined allows the researcher to observe the impacts of exogenous shocks on such behavior (such as the presence of wars, conflict, etc.).
+Moreover, over the same time period, both of my independent variables of interest, polarization and prestige of the military, have varied considerably, enabling for a fairly clear measurement strategy.
+
+What the data and the ensuing analysis from Chapter 3 cannot answer, however, is the degree to which variation in the observed military behavior is a function of changes in the editorial sources themselves over the same time period.
+In other words, it will be difficult to determine the degree to which the New York Times, for instance, has shifted to the political left over the past five decades (if any), and what impact these changes may have in determining the content type of its editorials.
+As far as connecting these initial five hypotheses to my overall theory and argument, I do not think that one hypothesis serves as more critical than any of the others, but rather that together, all five of the hypotheses lend convincing support for my theory.
+Writing an op-ed is a uniquely quasi-political activity in the sense that an op-ed can be very political or, alternately, entirely apolitical, depending on the content of the op-ed.
+But unlike tweets, facebook posts, are other modern forms of military elite speech that scholars have analyzed [@robinson_michael_danger_2018], op-eds must be deliberately written, and contain an overall argument.
+Publishing an op-ed in a major newspaper is also a somewhat competitive process, indicating that the author likely desired or greatly desired to make his or her opinions known.
+By analyzing the content of op-eds authored by retired flag officers over time and by controlling for the variables outlined above, I will be able to detect the degree to which there have been changes in the willingness of military elites to openly criticize civilian leaders and to weigh in on matters outside of their traditional military expertise, which are both critical measures of the military's role beliefs.
+
+In Chapter 4, I plan to conduct a mixed-methods analysis of the beliefs of senior military officers regarding the role of the military in politics in historical perspective.
+In order to do this, I plan to examine the behavior of individual high ranking officers - in fact, the highest ranking officers - Chairmen of the Joint Chiefs of Staff, and military service chiefs (such as the Chief of Staff of the Army, the Chief of Naval Operations, the Commandant of the Marine Corps, the Chief of Staff of the Air Force, and the Commandant of the Coast Guard) since World War Two.
+Examining the behavior of these leaders (roughly 120 in total), particularly during their tenure in the position as head of their respective services (or the military in the case of the Chairman of the Joint Chiefs of Staff) and in each's retirement years, will shed light on what each leader thought about the military's role in politics, and whether each officer lived out such beliefs in their retirement years.
+
+Focusing on the highest ranking military officers is especially important, as it enables a researcher to examine if there have been changes in what these leaders - as the "stewards of the profession" [@us_department_of_the_army_adp_2019, p.
+1-9] - have thought regarding the military's role in politics.
+Additionally, focusing on the individual enables the researcher to control for individual-level impacts, such as shocks to a specific officer's level of prestige or credibility, that may impact the dependent variable, individual acts of political behavior by military elites.
+The hypotheses that I wish to examine and test in this chapter include:
+
+_H6: Officers who endure a decline in either their individual or institutional level of prestige will be less politically active than those who experience either a relative boost or no change._
+
+_H7: When polarization and military prestige are low, senior military elites are less likely to weigh in on turbulent domestic political events than when polarization and military prestige are high._
+
+_H8: As polarization and prestige increase, there will be a gap between the "role beliefs" espoused by senior military leaders while on active duty and the "role beliefs" these same leaders live out during retirement._
+
+The dependent variables in this chapter will include political acts of behavior or seemingly political behavior, to include: writing opinion editorials (building off of data in Chapter 3), providing congressional testimony (on active duty as well as during retirement), and weighing in on turbulent political events in the form of speeches, articles, commentary, etc.^[I would include other actions, such as endorsing a Presidential candidate for office or speaking at a political convention, but these more "obvious" forms of political behavior did not start until at the late 1980s (publicly endorsing a presidential candidate), and even later (for speaking at a national political convention).
+Thus, I need to account for this type of "discontinuity" in my results, i.e.
+somewhere in the late 1980s, military officers began these particular types of behavior in earnest.] I will also record the occupation or primary role that each officer held in retirement (for example, political appointee, private citizen, defense industry consultant, etc.) The officer's post-retirement occupation is another indicator of how politically active the officer decided to be.
+
+For independent variables, I have to control again for the aggregate levels of institutional military prestige and political polarization, both of which data is available for using Gallup public opinion polling data and DW-NOMINATE scores.
+But in this chapter, since I am examining the individual officer, I will also control for whether a particular military officer experienced a relative decline, boost, or "no change" to his/her individual status of military prestige while serving in their senior military role.
+This will require coding for whether the officer experienced reputational damage as a result of an exogenous event, such as: being fired by a President (likely resulting in a loss of prestige), enjoying a tremendous increase in public opinion for directing successful combat operations (likely resulting in an increase in prestige) or, conversely, unsuccessful combat operations (loss of prestige) during war or a major military engagement.
+Alternatively, I suspect that the vast majority of these senior officers experience neither a boost or decline in individual prestige during their tenure as senior military leaders, i.e.
+each does his or her duty and then retires.
+But I suspect that the data will nonetheless include sufficient variation in this variable, over time, that there may be at least a correlation between an officer's individual reputation and their post-retirement political activity.
+
+Obtaining the data for these dependent variables will require classifying what primary role the officer played in retirement (i.e.
+political appointee, regular civilian, corporate board member), searching the congressional record by the officer's last name and categorizing the number and type of congressional hearings the officer participated in, both while on active duty and during retirement, and searching historical newspaper databases for any articles or speeches the author might have given, both on active duty and during retirement, particularly regarding turbulent or polarizing events (such as implementing the end to segregation (both within the military and in society), national protests during 1968, the response to Hurricane Katrina, etc.).
+^[I need to determine which events these are.
+Similar to the ICB, I have to find the 30 or so "most turbulent domestic events" that have occurred in the United States in the past 70 years, i.e.
+protests and riots, etc., and focus my research to answer the question, what did these military leaders I am examining have to say about these events?
+I am scouring the American politics literature for a database or a consensus of what these events are or might be, but as of yet, I have not found one.]
+
+I am not yet clear whether a strict quantitative or qualitative analysis, or a mixture of both, is most appropriate for this chapter.
+I think that one of the issues this will hinge on is the degree of clarity with which some of the coding decisions can and should be made.
+In the international relations literature, leader and state reputation is a topic that has been studied, particularly regarding the length of tenure for heads of state as a result of initiation of conflict [@croco_deciders_2011; @croco_war_2016].
+What made these studies possible, in my opinion, was clear selection criteria, and particularly the use of data sets involving Militarized Interstate Disputes or International Crises proved helpful.
+Unfortunately, to my knowledge, there is not a similar data set in the American politics literature that consists of relevant domestic turbulent events of the 20th century, for instance.
+I am potentially up to the task of building such a list myself, but would require some assistance from an American Politics scholar to do so in a way that passes muster for academic rigor.
+For example, I can imagine that I would want to include all instances in which military force was used domestically, instances in which there was large, sustained use of the military for responding to domestic natural disasters, and instances in which there was rioting, etc.
+But there may be more that I need to include, and I would want to bounce this idea off of other scholars to determine how to appropriately selecting the right events.
+
+In chapter five, I plan to present the results of an original survey experiment in which I survey a nationally representative sample of civilians and compare the results to a sample of either a) veterans, or b) active duty military officers (pending a lengthy approval process through the Army, which, quite frankly, may not prove possible given the time constraints I am working with in this particular dissertation).
+The goal of this survey is to test one of the micro-foundations for the theory that I have laid out, particularly, the role of political polarization in impacting military role beliefs.
+
+The survey will not test the impact of military prestige on military role beliefs, as I assume that everyone taking the survey, in 2020 or 2021, is operating from a standpoint that the current military institution enjoys a relatively high level of prestige and credibility (in other words, the survey assumes that the credibility of the military institution is endogenous).
+This assumption allows me to focus more on the impact of political polarization on military role beliefs, and to design questions that can further isolate its impact on the dependent variable.
+
+In the survey, the dependent variable I will measure is support for the military engaging in a range of overtly political behaviors in response to controversial situations involving the military.
+These behaviors span the spectrum of possible outcomes, include obeying the orders, writing an opinion piece, giving an interview, to resignation in protest.
+I further plan to frame the controversial situations in such a way that each involves the independent variable (IV) of interest, political polarization, through a hypothetical but reasonable (i.e.
+based on historic events or events that could reasonably occur) situation that is likely to awaken or stimulate two conflicting responses from the survey respondent: a desire to engage in the normatively-correct, non-partisan ethic of the military officer, and a desire to engage in the behavior that is likely affiliated with a particular partisan identity or ideological group.
+In the series of questions that I plan to ask, half of the population will comprise the treatment group, which will receive an additional piece of information that is highly charged or potentially polarizing that the control group does not receive.
+I then plan to measure the difference in responses between those whose ideological affiliations are primed (the treatment group) and those whose who were not (the control group), as well as test for differences between the civilian sample and the military/veteran sample.
+Designing and implementing such a survey will enable me to test two important hypotheses:
+
+_H9: When the salience of one's ideological/partisan identity is raised, there is a greater willingness to engage in overtly political behavior than when one's ideological identity is not raised._
+
+_H10: Relative to civilians, military officers/veterans believe that their identity as a military professional compels them to engage in overtly political behavior when the salience of their ideological/partisan identity is raised._
+
+To test these hypotheses, my survey design will incorporate arguments made by military actors that appeal to the military officer's sense of loyalty.
+As recent scholarship has argued, the contemporary domestic political environment introduces unique challenges in part because a military officer holds multiple "loyalty structures," or several simultaneous relationships that all thrive in part because a military officer is loyal to the object of these relationships.
+For example, the military officer is loyal to the President as the Commander in Chief, but also loyal to the Constitution of the United States.
+
+Invoking conceptions of loyalty during the survey is deliberate, and the idea is to appeal to the normative foundation that members of the military are socialized into during their years of service.
+Each military service emphasizes loyalty as a trait and concept.
+Of the different military services, the US Army in particular has extensively developed this concept in a framework entitled "The Army Ethic," in which loyalty plays a significant role within the Army profession [@us_department_of_the_army_adp_2019].
+This ethic is formally laid out and described in the Army's doctrinal publication entitled, _Army Leadership and the Profession, ADP 6-22_, which charges Army leaders to "embrace and uphold the Army Values and standards of the profession" [@us_department_of_the_army_adp_2019].
+There are a total of seven "Army Values," which together spell the acronym, "l-d-r-s-h-i-p," which is short for the word "leadership," and the first of these values, loyalty, is followed by six others: duty, respect, selfless service, honor, integrity, and personal courage [@us_department_of_the_army_adp_2019, p.
+1-9.].
+By appealing to a military officer's sense of loyalty, the argument made in each question will likely arouse a strong response.
+
+Before asking several questions (likely 3-4), I plan to ask for the respondent's age, sex, partisan affiliation, ideological affiliation, veteran status, and other potential covariates.
+I will then engage in a series of 4-6 questions, each of which follows the same pattern: presentation of a situation, followed by an elite voice, and an elite opinion.
+I then will measure the degree to which the respondent agrees with the opinion of the military elite, and the degree to which the respondent agrees with the manner in which the military elite makes his or her opinion known.
+
+Four of these hypothetical situations/questions are as follows:
 
 \singlespace
-* Situation: The President of the United States, a [Republican, Democrat] recently commented that he is seriously contemplating ordering active duty military Troops to assist local police and National Guard units in several US cities in maintaining law and order as a result of initially-peaceful protests which turned violent in some, but not all, cases.
-These protests have been going on for more than two weeks and are largely, but not exclusively, affiliated with [the Black Lives Matter Organization, a pro-2nd Amendment organization].
+* Situation #1 (Domestic Use of Force): The President of the United States, a Republican, recently commented that he is seriously contemplating ordering active duty military Troops to assist local police and National Guard units in several US cities in maintaining law and order.
+Multiple news outlets have shown that these protests, while mostly peaceful, have turned violent in about ten percent of cases over the past several weeks.
+[Treatment group, add: _These protests are largely affiliated with the Black Lives Matter Organization._]
 
@@
 
-* “The military must not be ordered to put down riots and protests, even if a small number of them have, unfortunately, turned violent.
-Even if the President is legally able to issue such an order, the military’s fundamental purpose is to defend the nation from external threats, and not to endanger the Constitutional rights of everyday Americans.
-[Treatment group, add ]: Should the President order the active military to deploy to cities where these protests have turned violent, military officers should resist such an order and if necessary, resign their commissions.
-They owe a higher loyalty to the Constitution, and the President’s order would clearly violate the oath of office these officers have sworn to uphold.”
+* Elite Opinion: “If the President orders the military to deploy to cities where these protests have turned violent, military officers should resign their commissions.
+Military officers owe a higher loyalty - a sacred loyalty - to the Constitution, which protects the rights of protesters, than they do to any particular President.”
 
-* Do you find this argument compelling (7 point scale of (1) not at all - (7) absolutely, very compelling).
+* Do you agree with the retired general's opinion? (7 point scale, strongly disagree - strongly agree)
+
+* To what degree was it appropriate for the retired general to make his opinion public, i.e. to write the op-ed? (7 point scale, strongly disagree - strongly agree)
 
@@
 
-Such a survey will enable me to measure at least three things.
-First, it will allow me to measure whether appeals to loyalty are truly more effective than standard arguments that do not invoke obligations of loyalty.
-This result can be measured by comparing the answers between the treatment and control groups for both populations.
-The second object this survey will enable me to measure is whether members of the military (or veterans) react differently to appeals based in the trait of loyalty than regular civilians, as I have hypothesized in _H3_.
-The third thing this survey will allow me measure is the relative strength, if any, of appeals made to differing objects of loyalty (to the Constitution, to the Commander in Chief, to the sense of mission, to one's fellow soldiers, etc.) in relation to a particular course of action.
-The results of this chapter will inform a micro-foundational understanding of the strength of loyalty as it pertains to the decision-making of individuals.^[Of particular note, I am planning to also pursue a survey among actively-serving military officers, but the lead time for obtaining approval to do this can be very long, perhaps 9-12 months or more.
-If I am able to obtain approval for this, I will then have three different populations to measure: civilians, veterans, and actively-serving military officers.]
+\singlespace
+* Situation #2 (Withdrawal of Troops from Overseas): The President of the United States is considering pulling all US Troops from a middle Eastern country in which American forces have been deployed to for many years as part of the ongoing Global War on Terrorism.
+Congress is split on this decision.
+some feel that the President has the authority to make and implement this decision unilaterally, while others have vocally argued that the President should consult Congress before making such an influential decision.
+Several foreign policy think-tanks have argued that a sudden withdrawal of US forces will result in the harm of innocent civilians by local terrorist groups.
+[Treatment group, add: _The President is a Democrat and both the House and Senate are controlled by Republicans._]
 
-In the fifth chapter, I want to show that the military has also faced institutional challenges to the concept of loyalty as political polarization has worsened.
-To do this, I plan to examine, via textual analysis, the records of Congressional Committee hearings in which military officers, active and retired, have participated.
-My goal is to detect if and how patterns of questioning and providing answers have changed, particularly in how language related to the concept of loyalty has been used.
-I still have some work to do in fleshing this out some more, but my initial idea is to examine how the military is both asked and answers questions by members of Congress, and how these questions and answers invoke language that speaks to various loyalty obligations.
-An example of this might be analyzing the ways in which terms such as "moral obligation" and "sacred duty" have been used over time.
-I might be able to associate causes or subjects that are associated with this type of language, and correlate these causes and subjects to the level of polarization present in the country, as well as to the partisan identification of the specific member of Congress asking the question.
-My goal is to detect patterns that will help support my overall argument, and specifically, _H2_.
+* Elite Voice: Larry Ferguson is a retired US Navy Admiral, who has extensive overseas experience. He recently gave an interview on a prominent cable news channel in which he stated the following:
 
-In the sixth chapter, I will turn to a series of qualitative case studies to explore and test my theory.
-Key in this chapter is finding and examining historical periods in American civil-military relations in which there existed varying levels of political polarization and institutional credibility, and then, to walk the reader through how challenges of loyalty during these periods manifested themselves in different ways among specific members of the military and the institutional military at large, in line with the theory that I propose in Chapter 2 and specifically, _H2_.I then hope to find evidence for how the actions of the military and its leaders during these periods impacted the military's professional identity and character, in line with _H4_ and _H5_.
+* Elite Opinion: “The Chairman of the Joint Chiefs of Staff is the President's chief military adviser.
+But the Chairman also has an obligation to support and defend the Constitution of the United States, which clearly stipulates that Congress has a duty to exercise an oversight role of the military.
+If the Chairman does not advise the President to talk to Congress first, he will be failing to uphold his professional duties and responsibilities.”
 
-I have not yet settled on what the specific case studies be, particularly because I am wresting with whether I need to confine my examination to the time period in which the US military has generally been considered a profession (depending on who you ask, this is generally around World War One, or perhaps a little bit earlier, at the tail end of the 19th Century).
-My initial reaction is that I do not need to confine myself to this period.
-My initial projection of what these case studies will be and how they will align with varying levels of political polarization and institutional credibility are as follows:
+* Do you agree with the retired general's opinion? (7 point scale, strongly disagree - strongly agree)
 
-* High Polarization, High Institutional Credibility: civil-military relations in the Andrew Johnson administration after the American Civil War.
-I will explore how General Ulysses Grant confronted challenges of loyalty to multiple institutions and objects as he dealt with a President who was by many accounts sympathetic to southern interests as the US Army played a prominent role in reconstruction in the south [@hyman_santon_1960].
+* To what degree was it appropriate for the retired general to make his opinion public, i.e. to make these comments during an interview? (7 point scale, strongly disagree - strongly agree)
 
-* High Polarization, Low institutional credibility: civil-military relations after World War One, during the 1920s.
+\doublespace
 
-* Low Polarization, High Institutional Credibility: civil-military relations after World War Two, in the late 1940s and early 1950s.
+\singlespace
+* Situation #3 (Federal Budget): The President's top economic advisers spent the last several months developing a proposal for next year's federal budget, which was revealed yesterday.
+Several government agencies experienced funding cuts, to include the Center for Disease Control (CDC), the National Aeronautics and Space Administration (NASA), and the National Academy of Sciences (NAS).
+[Treatment group, add: _The President has come under fire from some predominantly left-leaning organizations for stating that while he believes in climate change, "the science is not conclusive that it is harming the earth to the degree that some have claimed."_]
 
-* Low Polarization, Low Institutional Credibility: civil-military relations in the post-Vietnam years, roughly 1972-1980.
+* Elite Voice: former Chief of Staff of the Air Force recently gave an interview with NPR radio in which he said, "I truly do not understand why the President would cut the funding from organizations whose job it is to promote science in one form or another.
+This is par for the course with this President, unfortunately, and it insults the military, too, because we now have to figure out how to work with a President who seems to not value scientific fact.
+As military professionals, we must uphold facts and seek the truth."
+
+* Do you agree with the retired general's opinion? (7 point scale, strongly disagree - strongly agree)
+
+* To what degree was it appropriate for the retired general to make his opinion public, i.e. to write the op-ed? (7 point scale, strongly disagree - strongly agree)
+
+\doublespace
+
+\singlespace
+* Situation #4 (Endorsement of a Presidential Candidate): In the 2024 election, the Republican and Democratic Candidate are very different, particularly regarding their stances on the contentious issue of immigration.
+[Treatment group, add: _The Democratic candidate has vocally supported a policy of "amnesty" for all immigrants who are currently in the country illegally, while the Republican candidate has stated that, "America is a nation of laws, and without laws, the foundation of society crumbles._"]
+
+* Elite Voice: A retired Marine Corps general spoke at the Democratic National Convention very briefly, introducing the Democratic candidate.
+He said, "I wrestled with whether I should throw my voice into the political fray, because typically, that's not something we as retired military officers do.
+But I could not stay silent when I heard the Republican candidate talk about his plans for immigration.
+And so tonight, I am proud to support the right candidate to be the next President - a candidate who, like me, believes that all people have a right to live in a society that is free from fear.
+One that is safe, and one that is secure.
+It is not in spite of the fact that I wore the uniform for 40 years that I make this statement.
+It is because of it."
+
+* Do you agree with the retired general's opinion? (7 point scale, strongly disagree - strongly agree)
+
+* To what degree was it appropriate for the retired general to make his opinion public, i.e. to speak at the convention? (7 point scale, strongly disagree - strongly agree)
+
+\doublespace
+
+\singlespace
+* Situation #5 (Domestic Show of Military Presence/Force): As part of the upcoming Independence Day holiday on July 4th, the President and his administration are mulling the idea of putting on a large parade showcasing the capabilities of the US Military.
+Some members in Congress are opposed to such a move, as they think it is a political stunt to refocus the nation’s attention away from recent domestic political baggage the President has been dealing with over the past several weeks.
+[Treatment group, add: The President is a Republican, and the domestic political baggage he has been dealing with refers to new revelations that he and other members of his cabinet may have downplayed the arrival and potential harm caused by a strain of a flu virus that killed more than 10,000 Americans last year.]
+
+* Elite Voice: The Chairman of the Joint Chiefs of Staff, during a news conference at the Pentagon, was recently asked what he felt about the overall situation and the plan to showcase the military.
+He said: “The military’s duty in this situation is to obey the President, salute the flag, and execute.
+Period and end of story.
+There is nothing illegal, immoral, or unethical about what the President wants us to do here, so the military will keep our mouths shut and execute the order.”
+
+* Do you agree with the retired general's opinion? (7 point scale, strongly disagree - strongly agree)
+
+* To what degree was it appropriate for the retired general to make his opinion public, i.e. to speak at the convention? (7 point scale, strongly disagree - strongly agree)
+
+\doublespace
+
+In the sixth chapter, I will turn to a series of qualitative case studies to explore and test the generalizability of my theory.
+Key in this chapter is finding and examining historical periods in American civil-military relations that are not in the contemporary period, which chapter three and four focus on.
+I then hope to find evidence for how political polarization and military prestige and credibility impacted the role beliefs adopted by the military during this particular time periods.
+My goal is to examine all four possible scenarios that Table 2 in the previous pages described, i.e.
+high and low periods of both polarization and credibility.
+While I am not settled on the specific historical cases to examine, I have determined an initial breakdown as described in the following paragraphs.
+
+To examine military role beliefs when both political polarization and institutional credibility and prestige are high, I will examine civil-military relations and the role of the military in the post Civil War era.
+This era is historically fascinating, as it deals with how General Ulysses Grant and other military leaders who enjoyed enormous levels of prestige following the Union's victory in the Civil War, behaved relative to civilian leaders such as President Andrew Johnson during the monumental challenges of reconstruction, and in particular, the Army's role in it, in the southern United States [@hyman_santon_1960].
+
+To examine military role beliefs when polarization is high but credibility and prestige are low, I will focus my attention on the period between roughly 1870 and the start of World War One, or "the gilded age." The military was, at best, in the very beginning of becoming a profession, and the credibility that it had enjoyed following the end of the Civil War had largely if not altogether dissipated.
+Politically, however, the nation was polarized.
+
+To examine when polarization is low, but the credibility and prestige of the military is high, I may focus on the period after World War Two and the Korean War, from about 1953-1958.
+Similarly, the period right before World War Two, from about 1933-1940, offers an interesting period of when polarization and military credibility and prestige were both low.
+A comparison of the military's perceived role in society between these two time periods will help round out the qualitative case study chapter.
 
 In the seventh and final chapter, I will conclude by exploring the implications of my argument.
-Note: one idea I am considering is doing an extra chapter or at least an extended portion of Chapter 6 on how the Black Lives Matter movement is impacting the military profession.
-I think there could be some interesting work here.
 
@@
 
-My main concern at this point is that I harness the relevant outside literature, from the fields of psychology, ethics, sociology, and management, and still focus my puzzle on an empirical, political science question.
-To that end, I am concerned that I have a topic that is not sufficiently narrow, and thus, I may need to hone in on a specific subset of questions that I have presented here.
+In terms of a plan to finish this dissertation by the desired time of May 2022, I have to set out in earnest.
+The following bulletized points capture what I believe are the critical milestones that must be completed.
 
-A second concern is that I do my best to make this dissertation relevant and interesting for international relations, and not only for the military itself (though that is certainly a focus of my intended work).
-To that end, I may need to adopt a more comparative approach that examines if and how other militaries have been impacted by political polarization in their countries, and thus attempt some sort of cross-national study, should one of my intended chapters be deemed insufficient.
-I note this especially because the concept of loyalty, as mentioned in this prospectus, features so prominently in international relations in the context of coups, ethnic exclusion, etc., and I want to make sure I am not missing an opportunity to bring these matters to bear on my research project.
+* Between now and 15 December 2020 - Continue to work on my prospectus.
+Work with Jonathan Renshon and Jon Pevehouse to develop and field my intended survey experiment.
+I have already obtained IRB approval through UW-Madison to field a survey, but I need to apply for funding (through the Army), finalize the experiment's questions, and prepare for data collection.
+Continue collecting data on chapter three.
 
-In terms of timeline, and this is still very fluid, I intend to reach the following milestones:
+* January-February 2021 - No later than date to defend my dissertation prospectus and complete my methods comprehensive exam (second field). Continue collecting data for chapter four.
 
-* Between now and 15 December 2021: Continue to work on my prospectus.
-Work with Jonathan Renshon and Jon Pevehouse to develop and field my intended survey experiment (I have already obtained IRB approval through UW-Madison, but I need to apply for funding (through the Army), finalize the experiment, and prepare for data collection.) Plan to take my methods (second field) comprehensive exam.
+* February 2021 - field my survey experiment (ideally before the inauguration, if possible!).
 
-* No later than 15 January 2021: defend my dissertation prospectus.
+* March 2021 - Submit chapter three of my dissertation and submit for review to the dissertation committee.
 
-* No later than 15 May 2021: fully write Chapters 1-3 of my dissertation and submit for review to the dissertation committee.
-Prepare to field the original survey experiment for Chapter 4.
-Begin data collection for Chapter 5 of the dissertation, and settle on which case studies to write for Chapter 6 of the dissertation.
+* July 2021 -  Submit chapter four of my dissertation for review to the dissertation committee.
 
-* Summer 2021: continue working on Chapters 4 and 5 of the dissertation.
+* September 2021 - Submit chapter five of my dissertation for review to the dissertation committee
 
-* Fall 2021: submit drafts of Chapters 4 and 5 for review to the committee, and begin writing Chapters 6 and 7.
+* January 2022 - Submit chapter six of my dissertation for review to the dissertation committee.
 
-* January 2022: submit initial drafts of Chapters 6 and 7 to the committee for review.
+* March 2022 - Submit chapter seven of my dissertation for review to the dissertation committee.
 
-*February - May 2022: improve all chapters of the dissertation.
-
-* No later than May 2022: defend dissertation.
+* No later than May 2022 - defend dissertation.
 
@@
 
-The purpose of my dissertation is to explore how polarizing domestic politics affects normative concepts of military professionalism, particularly in the context of the United States and among the leaders of the US military.
-Certain features of increasing political polarization - in particular, the way in which individuals associate their partisan beliefs with salient features of identity [@sides_identity_2018] and right and wrong [@graham_liberals_2009], I argue, introduce immense challenges to the way military leaders conceive of and exercise loyalty, which has long been a foundational trait of military professions [@us_department_of_the_army_adp_2019].
-In particular, military officers in polarized time periods face challenges of conflicting loyalties [@coleman_problems_2009] that are more acute than during non-polarized periods, resulting in situations where military officers must prioritize the multiple objects of their loyalty (to the President as Commander in Chief, to Congress, to the operational mission, to subordinates and superiors along the chain-of-command, and to one's self).
+The purpose of my dissertation is to explore how polarizing domestic politics and increasing military prestige impact the normative conceptions of role beliefs among military leaders.
+How military leaders view their role within the government will likely impact the degree to which military leaders intervene in politics.
+My dissertation contends that the two major forces impacting how military leaders view their central roles within government are the degree of political polarization and the level of prestige or credibility of both the military institution and individual military actors.
+While recent scholars have suspected these forces impact the professional aspects of the military, my dissertation will provide a significant amount of empirical understanding to clarifying the mechanisms through which these trends interact.
+Moreover, a careful and rigorous analysis will better inform our understanding of how the US military in particular has weathered changes to its role beliefs in the past.
+Ultimately, this dissertation will yield knowledge that scholars, policymakers, and military officers can each benefit from.
 
-These decisions, though made by individual military leaders, often result in changes to the military as a profession and as an institution, with ramifications that bear on civil-military relations, the military's ability to defend the nation, and other important areas.
-By focusing my dissertation on this topic, I will contribute to scholarship by examining the specific ways in which polarization impacts the military institution, thus helping scholars of civil-military relations and international relations better understand the mechanisms through which polarization changes American institutions, particularly among the one institution charged with maintaining national security.
-
-\singlespace
 \pagebreak

### 2020-12-05 code/Tahkv1.Rmd
commit

--- 2020-12-04 code/Tahkv1.tex
@@
-\PassOptionsToPackage{unicode=true}{hyperref} % options for packages loaded elsewhere
-\PassOptionsToPackage{hyphens}{url}
-\PassOptionsToPackage{dvipsnames,svgnames*,x11names*}{xcolor}
-%
-\documentclass[12pt,]{article}
-\usepackage{lmodern}
-\usepackage{amssymb,amsmath}
-\usepackage{ifxetex,ifluatex}
-\usepackage{fixltx2e} % provides \textsubscript
-\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
-  \usepackage[T1]{fontenc}
-  \usepackage[utf8]{inputenc}
-  \usepackage{textcomp} % provides euro and other symbols
-\else % if luatex or xelatex
-  \usepackage{unicode-math}
-  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
-\fi
-% use upquote if available, for straight quotes in verbatim environments
-\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
-% use microtype if available
-\IfFileExists{microtype.sty}{%
-\usepackage[]{microtype}
-\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
-}{}
-\usepackage{xcolor}
-\usepackage{hyperref}
-\hypersetup{
-            pdftitle={When the Pen Becomes a Sword: Retired US Military Officers and Op-Ed Use in Contemporary America},
-            pdfauthor={Peter M. Erickson},
-            colorlinks=true,
-            linkcolor=magenta,
-            filecolor=Maroon,
-            citecolor=black,
-            urlcolor=blue,
-            breaklinks=true}
-\urlstyle{same}  % don't use monospace font for urls
-\usepackage[margin = 1.15in]{geometry}
-\usepackage{longtable,booktabs}
-% Fix footnotes in tables (requires footnote package)
-\IfFileExists{footnote.sty}{\usepackage{footnote}\makesavenoteenv{longtable}}{}
-\usepackage{graphicx,grffile}
-\makeatletter
-\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
-\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
-\makeatother
-% Scale images if necessary, so that they will not overflow the page
-% margins by default, and it is still possible to overwrite the defaults
-% using explicit options in \includegraphics[width, height, ...]{}
-\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
-\setlength{\emergencystretch}{3em}  % prevent overfull lines
-\providecommand{\tightlist}{%
-  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
-\setcounter{secnumdepth}{0}
-% Redefines (sub)paragraphs to behave more like sections
-\ifx\paragraph\undefined\else
-\let\oldparagraph\paragraph
-\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
-\fi
-\ifx\subparagraph\undefined\else
-\let\oldsubparagraph\subparagraph
-\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
-\fi
+---
+title: "When the Pen Becomes a Sword: Retired US Military Officers and Op-Ed Use in Contemporary America"
+author: "Peter M. Erickson"
+date: |
+  `r format(Sys.time(), '%B %d, %Y')`
+abstract: |
 
-% set default figure placement to htbp
-\makeatletter
-\def\fps@figure{htbp}
-\makeatother
-
-% there are a few ways to include latex options.
-% 1: as a preamble file (like this one)
-% 2: directly into the header-includes YAML variable (global)
-% 3: output templates. For more info on templates,
-%    see <https://bookdown.org/yihui/rmarkdown/rticles-templates.html>
-%    and <https://bookdown.org/yihui/rmarkdown/rticles-usage.html>
-
-% These packages control the fonts used in PDF output.
-% The standard LaTeX engine (PDFLaTeX) has a limited selection of fonts.
-% See <https://www.tug.org/FontCatalogue/>
-
-\usepackage{libertine}                 % serif font
-\usepackage{libertinust1math}          % matching math font
-\usepackage[scaled = 0.95, varqu]{zi4} % teletype (monospace/code) font
-
-% The XeLaTeX engine gives you more control to specify fonts
-%   at the expense of having poorer typesetting principles
-%   (but xelatex typesetting is getting better)
-
-% this package enables decimal-aligned table columns
-\usepackage{dcolumn}
-
-% the default PDF template for R Markdown loads a lot of convenient packages
-% so you don't actually have to mess with them all that much.
-\usepackage{setspace}
-\usepackage{multirow}
-\usepackage{indentfirst}
-\usepackage{caption}
-\usepackage{booktabs}
-\usepackage{longtable}
-\usepackage{array}
-\usepackage{multirow}
-\usepackage{wrapfig}
-\usepackage{float}
-\usepackage{colortbl}
-\usepackage{pdflscape}
-\usepackage{tabu}
-\usepackage{threeparttable}
-\usepackage{threeparttablex}
-\usepackage[normalem]{ulem}
-\usepackage{makecell}
-\usepackage{xcolor}
-\usepackage[style=authoryear,]{biblatex}
-\addbibresource{researchpaper-ps904.bib}
-
-\title{When the Pen Becomes a Sword: Retired US Military Officers and Op-Ed Use in Contemporary America}
-\author{Peter M. Erickson}
-\date{December 03, 2020}
-
-\begin{document}
-\maketitle
-\begin{abstract}
-\emph{This chapter explores variation in the professionalism of opinion pieces authored by retired US military officers and published in major US newspapers since 1979.
+  _This chapter explores variation in the professionalism of opinion pieces authored by retired US military officers and published in major US newspapers since 1979.
 Drawing on an original data set almost 400 opinion-editorial pieces published by retired military officers in The Wall Street Journal, The Washington Post, The New York Times, The Los Angeles Times, and the USA Today since 1979, the paper argues there has been an increase in published opinion pieces which fail to uphold the concept of military professionalism as defined by historical military norms and standards.
@@
 The author argues that increases in political polarization and the prestige of the military account for the increase in the number of op-eds which challenge traditional conceptions and norms of military professionalism.
-The paper concludes by arguing that the concept of ``politicization of the military'' --- something that previous civil-military relations scholars have identified and described --- must be widened to account not only for efforts by politicians to use the military for political advantages, but also for the phenomenon of retired military officers engaging the public to achieve political ends.
-The author argues that such a development, if uncorrected, will cause significant damage to the military as a profession.}\\
-\end{abstract}
+The paper concludes by arguing that the concept of "politicization of the military" --- something that previous civil-military relations scholars have identified and described --- must be widened to account not only for efforts by politicians to use the military for political advantages, but also for the phenomenon of retired military officers engaging the public to achieve political ends.
+The author argues that such a development, if uncorrected, will cause significant damage to the military as a profession._
+geometry: margin = 1.15in              # margin widths
+fontsize: 12pt                         # text size
+bibliography: researchpaper-ps904.bib # bibliography file
+biblio-style: authoryear               # citation style
+urlcolor: blue                         # links to external URLs
+citecolor: black                       # links to bibliography
+linkcolor: magenta                     # internal figures, footnotes, etc.
+indent: true                          # paragraph format
+                                       # true for indented paragraphs
+
+header-includes:
+    - \usepackage{setspace}
+    - \usepackage{multirow}
+    - \usepackage{indentfirst}
+    - \usepackage{caption}
+
+output: 
+  bookdown::pdf_document2:
+    keep_tex: true
+    latex_engine: pdflatex
+    number_sections: false
+    toc: false
+    fig_caption: true
+    citation_package: biblatex
+    includes: 
+      in_header: 
+        - latex-preamble.tex
+  bookdown::htmldocument2: default
+  bookdown::word_document2:
+      toc: false
+---
 
@@
 
+```{r setup, include=FALSE}
+knitr::opts_chunk$set(echo = TRUE)
+```
+
+```{r, include = FALSE, echo = FALSE}
+library("here")       # file paths
+library("tidyverse")  # workhorse package
+library("tidylog")
+library("kableExtra")
+library("knitr")
+library("ggdag")
+library("dagitty")
+library("gridExtra")
+library("broom")
+library("dplyr")
+library("reshape2")
+library("stargazer")
+library("rstatix")
+```
 \newpage
 
-\hypertarget{introduction}{%
-\section{Introduction}\label{introduction}}
+# Introduction
 
@@
 In some cities and communities across the nation, peaceful protests were followed by periods of rioting and looting.
-On June 1, after giving remarks at the Rose Garden, President Donald Trump walked to St.~John's Church after security forces visibly displaced throngs of protesters in neighboring Lafayette Square.
+On June 1, after giving remarks at the Rose Garden, President Donald Trump walked to St.
+John's Church after security forces visibly displaced throngs of protesters in neighboring Lafayette Square.
 Flanked by members of his cabinet, to include the Secretary of Defense, and with the Chairman of the Joint Chiefs of Staff not far behind, the President held up a Bible at the church, stared into the cameras of the media, and, several minutes later, returned to the White House.
-In the days and weeks that followed, a civil-military relations crisis ensued \autocite{feaver_military_2020}.
-The Secretary of Defense, Mark Esper, publicly disagreed with the President's threat to invoke the Insurrection Act of 1807 to use active-duty forces to quell domestic riots \autocite{esper_secretary_2020}, and General Mark Milley, the Chairman of the Joint Chiefs of Staff, later told senior military officers graduating from the National Defense University that he ``should not have been there'' (accompanying the President on the walk to St.~John's Church), as his presence ``created a perception of the military involved in domestic politics'' \autocite{milley_official_2020}.
+In the days and weeks that followed, a civil-military relations crisis ensued  [@feaver_military_2020].
+The Secretary of Defense, Mark Esper, publicly disagreed with the President's threat to invoke the Insurrection Act of 1807 to use active-duty forces to quell domestic riots [@esper_secretary_2020], and General Mark Milley, the Chairman of the Joint Chiefs of Staff, later told senior military officers graduating from the National Defense University that he "should not have been there" (accompanying the President on the walk to St.
+John's Church), as his presence "created a perception of the military involved in domestic politics" [@milley_official_2020].
 
-Furthermore, several retired military officers weighed in not only on the Trump administration's handling of specific events near the White House, but also on the broader issues of race in America.\footnote{These officers included several retired four star generals and admirals}.
-However, there was marked variation in the tone and tenor of the remarks made by these officers: although each expressed deep concern about the prospect of active-duty forces confronting peaceful protesters, and thus, violating protections held by the First Amendment of the US Constitution \autocites{brooks_dismay_2020}{dempsey_former_2020}, some officers went much further in their criticisms, leveling direct insults of the President, including his fitness to lead \autocite{goldberg_james_2020}, his administration's broader policies on race in America \autocite{allen_moment_2020}, and his general leadership abilities \autocite{mullen_i_2020}.
-Moreover, the comments made by these retired senior leaders renewed interest, at least for a while, in the topic of retired military officer speaking out on current events - a topic that civil-military relations scholars have debated \autocites{kagan_let_2006}{brooks_let_2020}, but one that has yet, in the author's opinion, to be studied empirically and systematically.
+Furthermore, several retired military officers weighed in not only on the Trump administration's handling of specific events near the White House, but also on the broader issues of race in America.^[These officers included several retired four star generals and admirals].
+However, there was marked variation in the tone and tenor of the remarks made by these officers: although each expressed deep concern about the prospect of active-duty forces confronting peaceful protesters, and thus, violating protections held by the First Amendment of the US Constitution [@brooks_dismay_2020; @dempsey_former_2020], some officers went much further in their criticisms, leveling direct insults of the President, including his fitness to lead [@goldberg_james_2020], his administration's broader policies on race in America [@allen_moment_2020], and his general leadership abilities [@mullen_i_2020].
+Moreover, the comments made by these retired senior leaders renewed interest, at least for a while, in the topic of retired military officer speaking out on current events - a topic that civil-military relations scholars have debated [@kagan_let_2006; @brooks_let_2020], but one that has yet, in the author's opinion, to be studied empirically and systematically.
 In a domestic political age in which polarization is high, the topic of retired military officer speech is important for a variety of reasons.
@@
 More importantly, the topics that military leaders choose to address in the public square, and the tone of their remarks, is reflective of the conceptualizations these officers have regarding their role in society and ultimately, their purpose and professional identity as military officers.
-Indeed, in 2020, retired as well as active duty military officers grabbed the pen and argued in support of a variety of topics that clearly pushed the traditional boundaries of professional conduct by military officers by openly arguing for a variety of topics related to President Donald Trump, including including race in America \autocite{askew_anti-racist_2020} and even removing the President from the White House \autocite{nagl__2020}.
+Indeed, in 2020, retired as well as active duty military officers grabbed the pen and argued in support of a variety of topics that clearly pushed the traditional boundaries of professional conduct by military officers by openly arguing for a variety of topics related to President Donald Trump, including including race in America [@askew_anti-racist_2020] and even removing the President from the White House [@nagl__2020].
 
@@
 This paper proceeds in three parts.
-First, the paper connects the concept of military ``role beliefs'' to retired military officer speech.
+First, the paper connects the concept of military "role beliefs" to retired military officer speech.
 Second, the paper explores and analyzes an original data set of nearly 390 opinion-editorial pieces published in major US newspapers by retired military officers since the late 1970s.
@@
 \singlespace
+# How the Military Views Itself
+\doublespace
+## Military Role Beliefs
+The concept of military "role beliefs" was introduced by J.
+Samuel Fitch in his book, _The Armed Forces and Democracy in Latin America._ In the book, Fitch describes military "role beliefs" as the perceptions held by military officers regarding their normative role in politics [@fitch_armed_1998, 61].
+Though Fitch's focus is on militaries in Latin America, it is possible to use the concept of "role beliefs" to develop a normative starting point applicable to military officers in the United States.
 
-\hypertarget{how-the-military-views-itself}{%
-\section{How the Military Views Itself}\label{how-the-military-views-itself}}
-
-\doublespace
-
-\hypertarget{military-role-beliefs}{%
-\subsection{Military Role Beliefs}\label{military-role-beliefs}}
-
-The concept of military ``role beliefs'' was introduced by J.
-Samuel Fitch in his book, \emph{The Armed Forces and Democracy in Latin America.} In the book, Fitch describes military ``role beliefs'' as the perceptions held by military officers regarding their normative role in politics \autocite[61]{fitch_armed_1998}.
-Though Fitch's focus is on militaries in Latin America, it is possible to use the concept of ``role beliefs'' to develop a normative starting point applicable to military officers in the United States.
-
-Two foundations for developing a starting point for military ``role beliefs'' in the United States are the tandem norms of civilian control of the military, and the idea of the military acting as a non-partisan institution.
-Though such norms have been episodically challenged throughout American history in a variety of ways and during specific instances, these norms have nonetheless been both foundational and aspirational, particularly in the post World War Two period \autocite{weigley_american_2001}.
-The norm of civilian control of the military stems directly from the US Constitution, which expressly grants control of the military (as well as specific responsibilities associated with maintaining control) to the President, Congress, and the individual states \autocite{noauthor_us_1787-2}.
+Two foundations for developing a starting point for military "role beliefs" in the United States are the tandem norms of civilian control of the military, and the idea of the military acting as a non-partisan institution.
+Though such norms have been episodically challenged throughout American history in a variety of ways and during specific instances, these norms have nonetheless been both foundational and aspirational, particularly in the post World War Two period [@weigley_american_2001].
+The norm of civilian control of the military stems directly from the US Constitution, which expressly grants control of the military (as well as specific responsibilities associated with maintaining control) to the President, Congress, and the individual states [@noauthor_us_1787-2].
 Unlike the norm of civilian control, the norm of the military functioning as a non-partisan institution is not expressly spelled out in the Constitution, but rather the product of the American military's collective experiences and traditions in fighting and prevailing in the nation's wars and conflicts.
@@
 Further, a number of scholars have underscored the importance of these two norms, and the relationship between these norms and the military's ability to prevail during times of war.
-In his influential book, \emph{The Soldier and the State: The Theory and Politics of Civil-Military Relations}, Samuel Huntington argues in favor of maintaining separate operating spheres between the political leaders of a democracy and the military as an institution in order to ensure civilian control of the military by rendering the military ``politically sterile and neutral,'' or non-partisan \autocite[84]{huntington_soldier_1957}.
-Huntington goes on to argue that only by such an arrangement, which he called ``objective control of the military,'' will the military be capable of defeating significant security threats while also remaining apolitical \autocite[83-84]{huntington_soldier_1957}.
+In his influential book, _The Soldier and the State: The Theory and Politics of Civil-Military Relations_, Samuel Huntington argues in favor of maintaining separate operating spheres between the political leaders of a democracy and the military as an institution in order to ensure civilian control of the military by rendering the military "politically sterile and neutral," or non-partisan [@huntington_soldier_1957, 84].
+Huntington goes on to argue that only by such an arrangement, which he called "objective control of the military," will the military be capable of defeating significant security threats while also remaining apolitical [@huntington_soldier_1957, 83-84].
 
@@
 This requires balancing these norms with the fact that military officers are nonetheless servants of the Republic who have accumulated knowledge and experiences over the course of sometimes lengthy careers (in the case of retired military officers, at least 20 and in some cases, upwards of 40 or even more years of experience), and thus, in many cases are qualified to weigh in on circumstances and situations that impact the nation.
-For instance, in the aftermath of the ``Revolt of the Generals'' in 2006, in which several retired military officers publicly criticized and called for the removal of Defense Secretary Donald Rumsfeld, Frederick Kagan argued in support of retired military officers having the freedom to weigh in policy debates, particularly those involving an active armed conflict, stating, ``there is no danger to the republic in a handful of retired generals speaking their minds.
-There is great danger in making vital decisions about an on-going armed struggle without hearing the views of all available experts'' \autocite{kagan_let_2006}.
-Kagan's remarks lend support to the notion that in democracies in particular, transparent and flowing information is a precious resource that helps leaders make decisions and, more importantly, the public to gain awareness of the state's actions {[}need cite{]}.
+For instance, in the aftermath of the "Revolt of the Generals" in 2006, in which several retired military officers publicly criticized and called for the removal of Defense Secretary Donald Rumsfeld, Frederick Kagan argued in support of retired military officers having the freedom to weigh in policy debates, particularly those involving an active armed conflict, stating, "there is no danger to the republic in a handful of retired generals speaking their minds.
+There is great danger in making vital decisions about an on-going armed struggle without hearing the views of all available experts" [@kagan_let_2006].
+Kagan's remarks lend support to the notion that in democracies in particular, transparent and flowing information is a precious resource that helps leaders make decisions and, more importantly, the public to gain awareness of the state's actions [need cite].
 
 Nonetheless, there have been both formal and informal boundaries placed on various forms of political activities by military officers.
-Formally speaking, Department of Defense (DoD) Directive 1344.10 contains a ``standing list of political behavior do's and don'ts for members of the active duty military'' \autocite[1]{urben_wearing_2014}.
-This list includes authorizing members of the military to engage in activities such as voting and joining a partisan or non-partisan club and to attend this club's meetings when not in uniform, and prohibits members of the military from activities such as participating in partisan fundraising, or advocating for or against a partisan cause or figure on radio or television \autocite{department_of_defense_political_2008}.
+Formally speaking, Department of Defense (DoD) Directive 1344.10 contains a "standing list of political behavior do's and don'ts for members of the active duty military" [@urben_wearing_2014, 1].
+This list includes authorizing members of the military to engage in activities such as voting and joining a partisan or non-partisan club and to attend this club's meetings when not in uniform, and prohibits members of the military from activities such as participating in partisan fundraising, or advocating for or against a partisan cause or figure on radio or television [@department_of_defense_political_2008].
 Informally, a number of recent Chairmen of the Joint Chiefs of Staff have likewise urged members of the military to refrain from engaging in overly political activities, particularly in election years, when such activity is more likely to negatively reflect on the military.
-For example, prior to the 2016 Presidential Election, then-Chairman of the Joint Chiefs of Staff General Joseph Dunford reminded the military that the basis upon which it is able to maintain the public's trust is it's ability to remain ``apolitical and neutral'' \autocite{dunford_upholding_2016}.
+For example, prior to the 2016 Presidential Election, then-Chairman of the Joint Chiefs of Staff General Joseph Dunford reminded the military that the basis upon which it is able to maintain the public's trust is it's ability to remain "apolitical and neutral" [@dunford_upholding_2016].
 
@@
 However, there is a point at which a criticism levied at a political leader by a military officer goes beyond criticism.
-Finally, it is possible to conclude that activities which build on the appropriate expertise and knowledge of the military officer are appropriate, provided that these activities do not violate the stipulations previously mentioned, i.e.~a non-partisan military, and civilian control of the military.
+Finally, it is possible to conclude that activities which build on the appropriate expertise and knowledge of the military officer are appropriate, provided that these activities do not violate the stipulations previously mentioned, i.e.
+a non-partisan military, and civilian control of the military.
 
-\hypertarget{opinion-editorial-publication-as-a-political-activity}{%
-\subsection{Opinion-Editorial Publication as a Political Activity?}\label{opinion-editorial-publication-as-a-political-activity}}
-
+## Opinion-Editorial Publication as a Political Activity?
 We can now focus on the actual behavior of the military authoring op-ed pieces.
@@
 Members of the active military do not frequently write op-eds, but there have been notable exceptions over the years, perhaps none so notable as several written by then-Chairman of the Joint Chiefs of Staff General Colin Powell in the early 1990s.
-For instance, Powell presented the public with his view regarding the potential use of American military force in several developing ``hot spots'' across the world, including the Balkans, Somalia, and other places, during an October 8, 1992 piece entitled, ``Why Generals Get Nervous.'' Published in the \emph{New York Times}, Powell wrote in part, ``decisive means and results are always to be preferred, even if they are not always possible.
+For instance, Powell presented the public with his view regarding the potential use of American military force in several developing "hot spots" across the world, including the Balkans, Somalia, and other places, during an October 8, 1992 piece entitled, "Why Generals Get Nervous." Published in the _New York Times_, Powell wrote in part, "decisive means and results are always to be preferred, even if they are not always possible.
 So you bet I get nervous when so-called experts suggest that all we need is a little surgical bombing or a limited attack.
 When the desired result isn't obtained, a new set of experts then comes forward with talk of a little esclalation.
-History has not been kind to this approach'' \autocite{powell_opinion_1992}.
+History has not been kind to this approach" [@powell_opinion_1992].
 
@@
 Bush, against Democratic challenger Bill Clinton.
-The two candidates differed with respect to foreign policy on a number of issues, and Bill Clinton was not highly respected by the military establishment, a fact that would impact the military's relationship with Clinton during his early years in office {[}need cite{]}.
+The two candidates differed with respect to foreign policy on a number of issues, and Bill Clinton was not highly respected by the military establishment, a fact that would impact the military's relationship with Clinton during his early years in office [need cite].
 Whether it was proper for Powell to write such an op-ed truly depends: one could argue that the nation's top general was simply informing the public based on his years of individual experience serving in the military, in which case, from a democratic transparency standpoint, the op-ed well served the Republic; alternately, one could also argue that by writing during an election year, Powell's remarks brazenly staked the claims of the military ahead of a potential Presidential transition (which turned out to be the case), and in the process, limited the President's options for potentially using force abroad during his early days in office.
@@
 
-\hypertarget{retired-military-officers}{%
-\subsection{Retired Military Officers}\label{retired-military-officers}}
-
+## Retired Military Officers
 One might argue that the logic employed thus far is seemingly plausible for members of the active military, but that the cautionary tone and reservations regarding certain types of political activity should not apply for members of the retired military community.
 And indeed, there are obvious distinctions between those who are on active duty and those who have retired from the service.
-These distinctions are made clear in fact in portions of DoD 1344.10 \autocite{department_of_defense_political_2008}.
+These distinctions are made clear in fact in portions of DoD 1344.10 [@department_of_defense_political_2008].
 But the fact of the matter is that, especially for retired military officers who achieved general or flag officer rank, these retired officers sometimes carry outsized influence and impact when they engage in activities that approach on the political.
-During the Presidential Campaign of 2016, Retired Army General Martin Dempsey published an opinion-editorial piece in \emph{The Washington Post} which criticized retired Army Major General Mike Flynn and Marine Corps General John Allen for participating in the respective Republican and Democratic conventions.
-Dempsey wrote, ``As generals, they have an obligation to uphold our apolitical traditions.
-They have just made the task of their successors -- who continue to serve in uniform and are accountable for our security - more complicated.
+During the Presidential Campaign of 2016, Retired Army General Martin Dempsey published an opinion-editorial piece in _The Washington Post_ which criticized retired Army Major General Mike Flynn and Marine Corps General John Allen for participating in the respective Republican and Democratic conventions.
+Dempsey wrote, "As generals, they have an obligation to uphold our apolitical traditions.
+They have just made the task of their successors – who continue to serve in uniform and are accountable for our security - more complicated.
 It was a mistake for them to participate as they did.
-It was a mistake for our presidential candidates to ask them to do so'' \autocite{dempsey_military_2016}.
+It was a mistake for our presidential candidates to ask them to do so" [@dempsey_military_2016].
 
@@
 
-\begin{quote}
-\textbf{Yours is the profession of arms, the will to win, the sure knowledge that in war there is no substitute for victory, that if you lose, the Nation will be destroyed, that the very obsession of your public service must be Duty, Honor, Country.
-Others will debate the controversial issues, national and international, which divide men's minds.
-But serene, calm, aloof, you stand as the Nation's war guardians, as its lifeguards from the raging tides of international conflict, as its gladiators in the arena of battle.
+> **Yours is the profession of arms, the will to win, the sure knowledge that in war there is no substitute for victory, that if you lose, the Nation will be destroyed, that the very obsession of your public service must be Duty, Honor, Country.
+Others will debate the controversial issues, national and international, which divide men’s minds.
+But serene, calm, aloof, you stand as the Nation’s war guardians, as its lifeguards from the raging tides of international conflict, as its gladiators in the arena of battle.
 For a century and a half you have defended, guarded and protected its hallowed traditions of liberty and freedom, of right and justice.
@@
 These great national problems are not for your professional participation or military solution.
-Your guidepost stands out like a tenfold beacon in the night: Duty, Honor, Country} \autocite{macarthur_duty_1962}.
-\end{quote}
+Your guidepost stands out like a tenfold beacon in the night: Duty, Honor, Country** [@macarthur_duty_1962].
 
 \doublespacing
-
-By systematically examining the op-eds written by retired military officers in major US publications over time, it is possible, I contend, to draw inferences about how the concept of a ``role belief'' - conceptions of one's normative role in politics - has changed among retired US military officers (if at all) \autocite{fitch_armed_1998}.
-Op-eds are an interesting and appropriate medium to examine if differentiation in ``role beliefs'' is an object of interest.
+By systematically examining the op-eds written by retired military officers in major US publications over time, it is possible, I contend, to draw inferences about how the concept of a "role belief" - conceptions of one's normative role in politics - has changed among retired US military officers (if at all) [@fitch_armed_1998].
+Op-eds are an interesting and appropriate medium to examine if differentiation in "role beliefs" is an object of interest.
 The first reason why this is the case is because as previously stated, the act of crafting and publishing an op-ed does not necessarily constitute a political activity, though depending on content, they can be.
@@
 A deliberate and carefully-crafted message further underscores the degree to which the author believes the message he or she is sending.
-It is difficult to ``misspeak'' during or to be ``misquoted'' from an opinion-editorial.
+It is difficult to "misspeak" during or to be "misquoted" from an opinion-editorial.
 The author generally agrees with the message he or she is sending.
@@
 
-\hypertarget{hypotheses}{%
-\subsection{Hypotheses}\label{hypotheses}}
+## Hypotheses
 
@@
 
-\emph{H1: When political polarization is high, the military (or certain military figures) will increasingly criticize civilian leadership (indicating a drop in the principle of civilian control).}
+_H1: When political polarization is high, the military (or certain military figures) will increasingly criticize civilian leadership (indicating a drop in the principle of civilian control)._
 
-\emph{H2: When political polarization is low, the military (or certain military figures) will publicly weigh in on traditional national security debates (and not on domestic political concerns or issues outside of relevant military expertise).}
+_H2: When political polarization is low, the military (or certain military figures) will publicly weigh in on traditional national security debates (and not on domestic political concerns or issues outside of relevant military expertise)._
 
-\emph{H3: During periods of high prestige and credibility, the frequency of military involvement in quasi-political activities (such as writing op-eds) will increase relative to periods of low prestige and credibility.}
+_H3: During periods of high prestige and credibility, the frequency of military involvement in quasi-political activities (such as writing op-eds) will increase relative to periods of low prestige and credibility._
 
-\emph{H4: When both political polarization and high prestige are present, the military will be increasingly vocal and at higher frequency about certain domestic ideological or partisan causes than when these factors are not present.}
+_H4: When both political polarization and high prestige are present, the military will be increasingly vocal and at higher frequency about certain domestic ideological or partisan causes than when these factors are not present._
 
-\emph{H5: During periods of active ground combat, military elites will likely comment on the war, but with less frequency on other domestic issues.}
+_H5: During periods of active ground combat, military elites will likely comment on the war, but with less frequency on other domestic issues._
 
@@
 
-\hypertarget{methodological-approach-and-the-data}{%
-\section{Methodological Approach and the Data}\label{methodological-approach-and-the-data}}
+# Methodological Approach and the Data
+
+```{r introduction, include = FALSE}
+
+# Read in the Data
+
+dataset <- read_csv(here("data", "MilEditorials.csv"))
+dataset %>% print()
+# Only read in the first 390 rows - is there a better way to get rid of N/A data?
+
+# Replace NA with 0 in the data for variables letter_ed to online_only, and leg_act #to role
+adjdataset <- dataset[1:390,] %>%
+  mutate_at(
+    .vars = vars(letter_ed:online_only, Leg_Act:Role),
+    .funs = ~ case_when(is.na(.) ~0, TRUE ~ 1)
+) %>% mutate(
+    PubYr = lubridate::year(PubDate)
+) %>% view()
+
+summary(adjdataset)
+names(adjdataset)
+
+# count the number of observations by type of binary variable and store as new dataframe
+countbytype  <- adjdataset %>% 
+  count(Topic) %>% view()
+
+# count the number of observations by source
+countbysource <- adjdataset %>% 
+  count(Source) %>% view()
+
+#count the number of times each source occurs in the data
+WSJcount <- adjdataset %>% count(Source=="WSJ")
+WSJcount
+
+NYTcount <- adjdataset %>% count(Source=="NYT")
+NYTcount
+
+WaPOcount <- adjdataset %>% count(Source=="WaPO")
+WaPOcount
+
+USATodaycount <- adjdataset %>% count(Source=="USA Today")
+USATodaycount
+
+LATimescount <- adjdataset %>% count(Source=="LA Times") %>% print()
+LATimescount
+
+# Create a new variable called Yrcount, which is a count of the yearly number of published op-eds
+
+adjdataset <- adjdataset %>% 
+  group_by(PubYr) %>% 
+  mutate(Yrcount=n()) %>% 
+  view()
+
+# number of authors and author teams
+adjdataset %>% 
+  group_by(Author) %>% 
+  count() %>% 
+  view()
+```
 
@@
 The data consists of opinion-editorial and letters to the editor publications authored by retired military officers and published in major US newspapers from 1979-2020.
-These newspapers consist of \emph{The Wall Street Journal} (WSJ), \emph{The New York Times} (NYT), \emph{The Washington Post} (WaPo), \emph{The Los Angeles Times}, and \emph{The USA Today}.
+These newspapers consist of _The Wall Street Journal_ (WSJ), _The New York Times_ (NYT), _The Washington Post_ (WaPo), _The Los Angeles Times_, and _The USA Today_.
 
@@
 I do not include published interviews with retired military officers, nor reprinted editorials that were first published before 1979, such as those reprinted for the benefit of readers sometime after 1979.
-To assemble the data, I use a series of key-word searches, such as ``retired military officer,'' ``retired from the Army, Navy, Marine Corps, Air Force,'' ``former Chairman of the Joint Chiefs of Staff,'' ``General,'' ``Admiral,'' and so forth.
+To assemble the data, I use a series of key-word searches, such as "retired military officer," "retired from the Army, Navy, Marine Corps, Air Force," "former Chairman of the Joint Chiefs of Staff," "General," "Admiral," and so forth.
 I conduct these key word searches in several different databases, including Factiva, Lexus-Uni, and ProQuest Historical Newspapers.
-\footnote{For exact replication of the database search parameters, please see the Appendix.} I include in my search publications that were made available both in print and online format, though I did not double-count: if the same op-ed was published both in print and online and published by the same author, I count the article only one time.
+^[For exact replication of the database search parameters, please see the Appendix.] I include in my search publications that were made available both in print and online format, though I did not double-count: if the same op-ed was published both in print and online and published by the same author, I count the article only one time.
 
@@
 
-Although social media has taken over as the primary means in which a majority of Americans receive their news today \autocite{martin_how_2018}, the methodological approach adopted in this analysis allows for an examination of the same sources over time.
+Although social media has taken over as the primary means in which a majority of Americans receive their news today [@martin_how_2018], the methodological approach adopted in this analysis allows for an examination of the same sources over time.
 More importantly, even after accounting for the advent of online editions of newspapers in the late 1990s and early 2000s, the overall circulation of newspapers (print and online editions combined) is still down today compared to earlier years.
-As a means of rough comparison, estimated total daily newspaper circulation in the US in 2000 was 62 million for weekday and 54 million for weekend editions, respectively, while in 2018, the same figures were 29 million and 31 million \autocite{barthel_leading_2019}.
+As a means of rough comparison, estimated total daily newspaper circulation in the US in 2000 was 62 million for weekday and 54 million for weekend editions, respectively, while in 2018, the same figures were 29 million and 31 million [@barthel_leading_2019].
 
@@
 
-\hypertarget{summary-statistics}{%
-\subsection{Summary Statistics}\label{summary-statistics}}
-
+## Summary Statistics 
 The data set includes a total of 390 observations, penned by 216 distinct different authors or author teams.
@@
 Included in the data are cases in which at least one retired military officer co-authored a piece with one or more civilian authors.
-\footnote{There are several co-authored pieces.
-For example, where General (Retired) David Petraeus wrote one publication and another one was written by General (Retired) David Petraeus and Michael O'Hanlon, I counted these as two different authors for coding purposes.} There are also several retired military officers who penned several observations.
+^[There are several co-authored pieces.
+For example, where General (Retired) David Petraeus wrote one publication and another one was written by General (Retired) David Petraeus and Michael O'Hanlon, I counted these as two different authors for coding purposes.] There are also several retired military officers who penned several observations.
 The most prolific of these officers includes Retired Army Lieutenant General William Odom (22 publications in total), Retired Air Force General Michael Hayden (20 publications in total), and Retired Marine Lieutenant Colonel and Former National Security Advisor Robert McFarlane (18 publications).
 
-Figure One, below, plots the yearly count of publications authored by retired military officers across all of the major newspapers.
+```{r summary stats, include=FALSE, echo=FALSE}
+# show summary stats for yearly count, rank, and then a breakdown according to source
+# yearly count
+yearstats <- adjdataset %>% 
+  select(PubYr, Yrcount) %>% 
+  distinct(PubYr, Yrcount) %>% 
+  print()
+
+summary(yearstats)
+sd(yearstats$Yrcount)
+var(yearstats$Yrcount)
+
+#rank
+rankstats <- adjdataset %>% 
+  filter(AuthRank <=10) %>% 
+  select(Author, AuthRank) %>% print()
+
+summary(rankstats)
+sd(rankstats$AuthRank)
+var(rankstats$AuthRank)
+
+#sex
+adjdataset %>% group_by(AuthSex) %>% count()
+
+#service
+adjdataset %>% group_by(Service) %>% count()
+
+Descriptive_Stats <- tribble(
+  ~Variable, ~Min, ~Max, ~Mean, ~SD, ~N, 
+  'Author Rank', 4, 10, 8.34, 1.99, 390,
+  'Number of Yearly Publications', 1, 27, 9.51, 6.84, 390,
+  'Male Authors', NA, NA, NA,NA, 374,
+  'Female Authors', NA, NA, NA, NA, 3,
+  'Mixed Authorship', NA, NA, NA, NA, 5,
+  'Unknown Authorship', NA, NA, NA, NA, 8,
+  'Army', NA, NA, NA, NA, 205,
+  'Navy', NA, NA, NA, NA, 56,
+  'Marine Corps', NA, NA, NA, NA, 47,
+  'Air Force', NA, NA, NA, NA, 63,
+  'Coast Guard', NA, NA, NA, NA, 3,
+  'Mixed Service Authorship', NA, NA, NA, NA, 16)
+
+T1 <- kable(Descriptive_Stats, 
+            format = 'simple', 
+            booktabs = TRUE, 
+            caption = 'Descriptive Statistics (Author Rank and Yearly Number of Publications, and breakdown by Author Sex, Service',
+            align = c('l', 'c', 'c', 'c'),
+            digits = 2)
+
+ rankdistrib <- adjdataset %>% group_by(AuthRank) %>% count()
+
+```
+
+In addition to coding each publication for the source and a count of publications in a given year, several bits of biographical information about each author is collected, including the service in which the author served, the rank achieved by the author during his or her time in service, and the sex of the author.
+Summary statistics for the variables of author rank and the annual count of publications are displayed in Table 1, below, along with a breakdown of the observations by sex and service of the author.
+Mixed authorship refers to publications that were authored by a mixture of male and female authors, whereas mixed service authorship denotes publications that were authored by more than one individual and together, as an author team, had served in multiple service branches - for example, a retired Colonel who had served in the Air Force and a retired Admiral who had served in the Navy write an op-ed together.
+
+```{r table1, echo=FALSE, include=TRUE}
+T1
+```
+
+Of note, the variable for author rank is coded equivalently to the officer's military pay grade.
+For example, a rank of "10" means the officer achieved the pay grade of O-10, which is a four star general or admiral.
+A rank value of 7 refers to a one star general or admiral.
+A rank of 4 corresponds to a pay grade of O-4, which in the Army, Air Force, and Marine Corps, refers to the rank of Major.
+In the Navy and Coast Guard, a pay grade of O-4 refers to the rank of Lieutenant Commander.
+A rank of 5 in the Army, Air Force, and Marine Corps refers to a Lieutenant Colonel, whereas in the Navy and Coast Guard, a rank of 5 denotes a Commander.
+Finally, a rank of 6 in the Army, Air Force, and Marine Corps refers to a Colonel, whereas in the Navy and Coast Guard, a rank of 6 corresponds to the rank of Captain.
+
+It is worthy to briefly note the maximum, minimum, and mean values for the number of yearly publications from Table 1, as these numbers can be interpreted in a couple of different ways.
+The overall number of opinion pieces published by retired military officers in a given year may reflect some combination of willingness on the part of the press to publish the viewpoints of officers, or desire by the retired military community to make its viewpoints known, or some combination of these "demand" and "supply" effects.
+
+Figure 1, below, plots the yearly count of publications authored by retired military officers across all of the major newspapers.
 The smoothed trendline indicates that the average yearly count of publications rose steadily from 1979 until the early 2000s, at which time, a decline occurred until approximately 2008.
@@
 In the late 1970s, the military had ended the draft only years before (1973) following the conclusion of the Vietnam War.
-The debates over ``Don't Ask, Don't Tell'' (instituted in 1994 and repealed in 2011), as well as policies regarding women in the combat arms (2015) are also included in the data, as are numerous annual discussions regarding the size of military budgets.
+The debates over "Don't Ask, Don't Tell" (instituted in 1994 and repealed in 2011), as well as policies regarding women in the combat arms (2015) are also included in the data, as are numerous annual discussions regarding the size of military budgets.
 
-\begin{verbatim}
-## `geom_smooth()` using formula 'y ~ x'
-\end{verbatim}
+```{r figure1, fig.height=4, fig.width=5, fig.align="center", warning=FALSE, include=TRUE, echo=FALSE}
 
-\begin{center}\includegraphics{Tahkv1_files/figure-latex/figure one-1} \end{center}
+# construct a scatter plot with line and smoothing
 
-In addition to collecting the yearly count of publications, biographical information of the author is collected, including the service in which the author served, the rank achieved by the author during his or her time in service, and the sex of the author.
-Summary statistics for these variables are included in Table 1, below.
+ggplot(data=adjdataset, aes(x=PubYr, y=Yrcount)) + geom_point() + geom_line() + geom_smooth(method="loess") + labs(
+  x="Year", 
+  y="yearly count", 
+  title="Figure 1: Number of Op-Eds Published by Retired Military\n Officers (WSJ, NYT, WaPo, USA Today, LA Times)",
+  subtitle= "1979-2020",
+  caption="Note: As of 1 December 2020"
+  ) + scale_x_continuous(limits=c(1979, 2020)) + scale_y_continuous(limits=c(-7, 28)) + 
+geom_vline(xintercept = 1981, size=.2, colour="tomato") + 
+  geom_vline(xintercept=1989, size=.2, colour="tomato") + 
+  geom_vline(xintercept=1993, size=.2, colour="dodgerblue") + geom_vline(xintercept=2001, size=.2, colour="tomato") + 
+  geom_vline(xintercept=2009, size=.2, colour="dodgerblue") +
+  geom_vline(xintercept=2017, size=.2, colour="tomato") +
+  geom_text(aes(x=1981, label="Reagan",y=10), colour="tomato", angle=90, vjust=1, text=element_text(size=9)) + 
+  geom_text(aes(x=1989, label="Bush 41",y=15), colour="tomato", angle=90, vjust=1, text=element_text(size=9)) + 
+  geom_text(aes(x=1993, label="Clinton",y=20), colour="dodgerblue", angle=90, vjust=1, text=element_text(size=9)) + 
+  geom_text(aes(x=2001, label="Bush 43",y=-2), colour="tomato", angle=90, vjust=1, text=element_text(size=9)) + 
+  geom_text(aes(x=2009, label="Obama",y=-2), colour="dodgerblue", angle=90, vjust=1, text=element_text(size=9)) + 
+  geom_text(aes(x=2017, label="Trump",y=-2), colour="tomato", angle=90, vjust=1, text=element_text(size=9))
 
-Insert table of yearly count, rank, and sex, as well as a panel facet graph of Figure 1, above, by source.
+```
 
-Describe the other variables and coding scheme.
+Returning to a description of the statistics, Figure 2 below displays the distribution of publications by author rank (left) and source (right).
+As the distribution plot according to author rank shows, the publications tend to be authored by those of a higher rank.
+185 of the 390 publications (47.4%) were authored by retired 4 star general or flag officers (Generals or Admirals).
+Generals or Admirals account for 276 of the total 390 publications (70.8%).
+The fact that high ranking officers tend to publish in the newspapers examine is unsurprising, and confirms the competitive nature of publishing an opinion piece in one of the premier newspapers in the nation.
+At the same time, there are enough authors who are not retired generals or admirals (29.2%) to show that those who do not attain the rank of general or admiral still have a respectable opportunity to publish their views in the examined sources.
 
-A bar graph by decade across all sources collectively of \# of publication by topic count. Or do this later.
+The distribution according to source, shown on the right side of Figure 2, reveals that the overwhelming number (379, or 97.2%) of opinion publications were authored by retired military officers in the _New York Times_, _Washington Post_, or _Wall Street Journal._ Only 11 of the 390 publications (2.8%) were authored in either the _USA Today_ or the _Los Angeles Times_.
+This is not surprising.
+Though they distribute worldwide and are read across the country, the _Washington Post_, _New York Times_, and _Wall Street Journal_ are based in the nation's capital and New York City.
+Think tanks and international organizations abound in both cities, and thus, these three newspapers are likely more read by those who work in the defense or foreign affairs community.
 
-Of note, when assembling the data, some authors are explicitly identified in some of their publications as retired military officers, but not in others.\footnote{This was true of several individuals, including Alexander Haig, a retired Army General who later served as Secretary of State under Ronald Reagan; Robert McFarlane, a retired Marine Corps Lieutenant Colonel who later served as the National Security Adviser to President Reagan, Andrew Bacevich, a retired Army Colonel who then became a prominent historian and author; and Michael Hayden, a retired Air Force General who served as director of the National Security Agency and the Central Intelligence Agency}.
-I conduct a separate search for all opinion pieces penned by these authors, regardless of whether they were identified as a retired military officer or not, concluding that it is best do so for coding purposes.
+```{r figure2, include=TRUE, echo=FALSE}
+colplot1 <- ggplot(data=rankdistrib, aes(x=AuthRank, y=n)) + labs(x="Pay Grade", y="Number of Publications") + geom_col() + geom_text(aes(label=n), position=position_dodge(width=.9), vjust=-.25)
 
-Attempting to reconcile the fact that the military as an institution prides itself on being a non-partisan and apolitical institution, that newspaper circulations (even after accounting for digital offerings) are substantially down over the past 40 years, and that the US political environment is increasingly polarized, I adopt the two following hypotheses:
+colplot2 <- ggplot(data=countbysource, aes(x=Source, y=n)) + labs(x="Source", y="Number of Publications") + geom_col() + geom_text(aes(label=n), position=position_dodge(width=.9), vjust=-.25)
 
-\emph{\(H_1\): The majority of op-eds published by retired military officers over the time period examine will be apolitical and non-partisan in tone.}
+grid.arrange(colplot1, colplot2, ncol=2, top="Figure 2: Publications authored by Retired Military Officers\n by Author Rank (L) and Source (R), 1979-2020")
 
-\emph{\(H_2\): Given that polarization has become increasingly pronounced in the United States over the time period examined, there will be an increase in recent years in the number of op-eds written about topics that are not relevant to military expertise, or that reveal inappropriate partisan sentiments.
-Such a trend may indicate that increasing political polarization is impacting military professionalism among a small but important part of the military institution, retired military officers.}
+```
 
-\hypertarget{first-layer-analysis-topics-addressed}{%
-\section{First Layer Analysis: Topics Addressed}\label{first-layer-analysis-topics-addressed}}
+## First Layer Analysis: What Are Retired Military Officers Addressing?
 
-In the first level of analysis, I code each observation into one of several binary categories that corresponds to the overall topic or subject of the publication.
-The goal is to create a coding scheme that captures a concise number of topical descriptions that are still sufficiently broad enough to reveal variation in the topic addressed by retired military authors.
-I develop a total of six topical variables, and each observation takes on the value of 1 if the publication corresponds to the general definition of the variable, and 0 otherwise.
+Having established that there has been a relative increase in the overall number of opinion pieces authored by retired military officers and published in major US newspapers since 1979, the next critical step is to determine whether there has been change, and to what degree, in what retired military officers have actually been writing about.
+This requires classifying each of the observations into one of several possible types of topic.
+The possible range of topics should be broad enough to classify every observation, while maintaining sufficient distinction between each of the topics so as to make useful comparisons between the topics.
 
-These topical variables are: war fighting and operations, if the piece advocates for or denounces a particular course of action directly related to operational or war fighting decisions of a clear military nature; troop support and institutional culture, if the publication generally highlights the work and sacrifice of members of the armed forces, defends the work of specific members of the Armed Forces, or makes a call for action to improve a service's institutional culture; strategic advice, if the piece is generally written to offer some sort of strategic advice for the nation, to include the wisdom of pursuing particular alliances or agreements, or adopting a specific weapons platform or military budget; military social policy, if the piece advocates for or denounces a particular social policy that \emph{directly influences} the armed forces (Examples include: debates over keeping or removing the ``Don't Ask, Don't Tell'' policy, opening up combat roles to women, the role of Transgender members of the armed forces); domestic policy, if the piece advocates for or denounces a position on a policy that \emph{does not directly} influence the armed forces in its day to day activities (Examples include: support for the second amendment, national immigration policy, and whether confederate statues should be taken down or left up); and political figures, if the piece advocates for or denounces a particular position held by a political administration, or \emph{if the main point of the piece} is to explicitly defend or attack a political figure by name (Examples: endorsing a Presidential candidate for office, or for attacking a President or member of Congress for a decision that was made).
+I generate a categorical variable, "topic", that can take one of eight possible categories.
+The first is **foreign policy**, if the publication generally refers to American foreign policy or overarching National Security policy.
+The next category of publication is **warfighting and operations**, which I code those publications that address an ongoing conflict, or a central aspect of the military's ability to fight and win on the battlefield.
+Of note, I also include those publications that deal with most intelligence procedures, as in most cases, the authors of these publications tied intelligence to the military's ability to win on the battlefield.
+The third category is **social policy**, which include those publications that address questions such as who should serve, and why.
+In the data, this included publications that addressed subjects such as the "Don't Ask, Don't Tell" policy, recruitment, and the integration of women into the combat arms branches of the military.
 
-Up front, separating opinion-editorial publications into one of several distinct categories is an imperfect process.
-It requires a thoroughly objective reading of the publication as possible.
-To aid the coding process, I identify the thesis statement of each opinion piece, which in most cases is identifiable in either the lead or concluding paragraph of the opinion piece.
-In most cases, identifying and re-reading the thesis of each op-ed makes it possible to reasonably classify the publication by topic.
-Two additional points, however, are worth emphasizing here.
+A fourth category is **domestic policy**, for which I include those publications that address issues that do not have a _direct_ impact on the military, or those that as the main points of emphasis, describe the general operating environment in Washington D.C.
+To be clear, especially in a day and age when the military has taken on greater roles, it is easy for a military author to logically connect almost any topic with the nation's defense.
+And though there may be convincing times when these connections need to be made, the determination used when classifying publications according to this variable was whether the military can reasonably claim that it either is or should be an expert on a particular topic addressed.
+In the data, for instance, publications that first and foremost addressed the national debt or the domestic economic impacts of a given policy, are classified according to this category.
+The fifth categorical variable for topic is **budget/weapons/troops**, which I use to classify those publications that primarily address issues such as military or defense budgets, or why the military or a particular service should or should not adopt a particular weapons platform.
 
-First, opinion pieces are intended to argue for something, and thus are, to an extent, designed to be at least mildly provocative.
-With this in mind, I did not necessarily code pieces that called for a political figure or institution to take a specific action into the category of \emph{figures by name}.
-Again, the identification of the publication's thesis statement aided significantly in identifying the main thrust of the argument, and thus, in classifying the argument by topic.
-Second, there were some instances in which I had to choose between placing a publication into one of two categories.
-However, in instances of choosing between two categories, the choice to classify between categories was usually between categories that were reasonably close.
-For example, there were a few instances of choosing to classify a publication between the war fighting and operations category and the strategic advice category, but no instances of choosing between the war fighting and operations category and the domestic policy category.
-Table 3 reveals the breakdown of all examined publications by article type.
+The sixth category of topic used to classify publications is **support**, which I classify those publications that are written in support of the collective military (such as op-eds that ask the nation to remember fallen Troops and their families on Memorial Day), as well as those that actually support or weigh in on specific military leaders (such as those that recall a prominent officer who passed away recently).
+In the data, most patriotic calls to think about the sacrifice that troops make or have made are classified in this category.
+The seventh category of possible topics is **civil-military relations,** which I use to classify those topics that mainly address how civilian and military leaders should or actually do share power, or how generals and admirals should offer advice in an improved manner.
+In the data, this also includes publications which comment on how the military and government organizations may be better organized or rearranged in order to better  implement defense policies.
+The eighth and final category for publication topic is **service culture**, and refers to those publications that primarily address an issue or issues that are ongoing within the military services.
+^[See the Appendix for a more-detailed explanation of each variable and classification.]
 
-\begin{table}
+In order to classify each observation by topic, I read the op-ed or publication, usually multiple times, and identify what I think is the author's main point or thesis statement.
+To be clear, separating these publications into one of the eight possible categories is not a perfect process.
+Some classifications address multiple topics, or straddle the line between two or more categories.
+When this occurred, the central question made when determining an observation's ultimate topic category was, "what is the central argument made in the publication?"
 
-\caption{\label{tab:table3}Publications by Subject}
-\centering
-\fontsize{10}{12}\selectfont
-\begin{tabu} to \linewidth {>{\raggedright}X>{\raggedleft}X}
-\hline
-\textbf{Topic} & \textbf{n}\\
-\hline
-\cellcolor{gray!6}{BdgtWpnsTrps} & \cellcolor{gray!6}{37}\\
-\hline
-CivMilBalance & 19\\
-\hline
-\cellcolor{gray!6}{DomesticPolicy} & \cellcolor{gray!6}{38}\\
-\hline
-ForeignPolicy & 112\\
-\hline
-\cellcolor{gray!6}{ServiceCulture} & \cellcolor{gray!6}{13}\\
-\hline
-SocialPolicy & 9\\
-\hline
-\cellcolor{gray!6}{Support} & \cellcolor{gray!6}{44}\\
-\hline
-Warfighting & 118\\
-\hline
-\end{tabu}
-\end{table}
+Table 2, below, displays the breakdown of observations by topic category.
+The topics of **warfighting and operations** and **foreign policy** account for a total of 230 of the 390 observations (59% of the sample).
+The categories of **support**, **domestic policy**, and **budget/weapons/troops** account for a total of 119 observations, or 30.5% of the sample.
+Finally, the topics of **social policy**, **civil-military relations**, and **service culture** account for a combined 41 observations, or 10.5% of the sample.
 
-An overwhelming majority (293 of 344, or 85.2\%) of the total number of opinion pieces in the data set fall into the categories of war fighting and operations, troop support and institutional culture, and strategic advice.
-Moreover, I conclude that these observations pose no problem for norms of military professionalism, as they are written on topics that I contend are clearly related to one's relevant military experience.
-The remaining 51 publications (14.8\% of the observations) are coded as falling under the other three categories of topics: social policy directly related to the military, domestic policy, and political figures by name.
-These observations require a second and more-thorough examination to determine whether the content of these publications indicate a drop in the norm of military professionalism.
+```{r table2, fig.align="center", include=TRUE, echo=FALSE}
+countbytype %>%
+  kable(format="latex", caption="Publications by Subject") %>%
+  kable_styling(latex_options = "striped", font_size=10) %>% 
+  row_spec(0, bold=T)
+```
+
+Is this breakdown by itself surprising in any way, especially as it relates to the topic of military role beliefs?
+I contend that the answer is not especially.
+It is unsurprising that retired military officers comment heavily on matters related to warfighting, operations, and national security policy, as these matters directly relate to the experiences of retired military leaders.
+If anything, it is somewhat surprising, though not necessarily cause for concern, that the number of observations addressing domestic policy matters is as high as it is, and that the number addressing civil-military relations is as low as it is.
+
+However, what is relevant and applicable to the issue of military role beliefs is whether there has been a change in the way retired military officers have been addressing domestic political matters over time.
+Figure 3, below, plots the proportion of annual publications authored by retired military leaders that address domestic policy since 1979, and reveals several interesting trends.
+The first is that with the exception of an outlier year (1983), no publications addressed domestic policy until 1999.
+The first two decades worth of data, in other words, saw retired military officers nearly shun domestic politics altogether.
+The second interesting trend is that since that time period, retired military officers have become increasingly active in addressing domestic political matters when authoring opinion publications.
+The trend line in blue, moreover, drops between roughly the years 2009 and 2012, during the first Obama Administration.
+The final trend is that since 2012, the proportion of annual opinion publications authored by retired military officers which address domestic political matters has increased significantly.
+While this trend has increased under the Trump administration, it's origin can be found under the Obama and George W.
+Bush Administrations.
+
+```{r, include=FALSE, echo=FALSE}
+## Create several new variables.
+
+# The first is a variable called "case-count" that counts the number of times by publishing year an #article published that calls for legislative action, criticizes or is #insubordinate, or is blatantly #partisan.
+
+#The second variable, prop_yr, is simply a 
+# proportion of annual publications where these conditions hold true.
+
+# The third variable is called dom_pol_count which counts the number of publications
+# by publishing year that are written on topics of domestic policy, and the fourth 
+# variable is a proportion of this type of article by year.
+
+adjdataset <- adjdataset %>% group_by(PubYr) %>% 
+  mutate(
+    case_count=sum(Leg_Act ==1 | Crit_Insub==1 | Endorse_Partisan==1), 
+    prop_yr=(case_count/Yrcount),
+    dom_pol_count=sum(Topic=="DomesticPolicy"),
+    dom_pol_prop=(dom_pol_count/Yrcount)
+  ) %>% view()
+```
+
+```{r figure3, include=TRUE, echo=FALSE, warning=FALSE, error=FALSE}
+ggplot(data=adjdataset, mapping=aes(x=PubYr, y=dom_pol_prop)) + geom_point() + geom_smooth() + labs(title="Figure 3: Proportion of Annual Publications authored by Retired Military Officers\n that Address Domestic Political Matters, 1979-2020", x="Year", y="Proportion", color=NULL)+ geom_vline(xintercept = 1981, size=.2, colour="tomato") +
+  geom_vline(xintercept=1989, size=.2, colour="tomato") + 
+  geom_vline(xintercept=1993, size=.2, colour="dodgerblue") + geom_vline(xintercept=2001, size=.2, colour="tomato") + 
+  geom_vline(xintercept=2009, size=.2, colour="dodgerblue") +
+  geom_vline(xintercept=2017, size=.2, colour="tomato") +
+  geom_text(aes(x=1981, label="Reagan",y=.25), colour="tomato", angle=90, vjust=1, text=element_text(size=9)) + 
+  geom_text(aes(x=1989, label="Bush 41",y=.25), colour="tomato", angle=90, vjust=1, text=element_text(size=9)) + 
+  geom_text(aes(x=1993, label="Clinton",y=.25), colour="dodgerblue", angle=90, vjust=1, text=element_text(size=9)) + 
+  geom_text(aes(x=2001, label="Bush 43",y=.3), colour="tomato", angle=90, vjust=1, text=element_text(size=9)) + 
+  geom_text(aes(x=2009, label="Obama",y=.3), colour="dodgerblue", angle=90, vjust=1, text=element_text(size=9)) + 
+  geom_text(aes(x=2017, label="Trump",y=.4), colour="tomato", angle=90, vjust=1, text=element_text(size=9))
+
+```
+
+##Second Layer Analysis - A More Brazen Retired Officer Corps?
+
+Having classified what retired military officers have been addressing in publications according to topic, the next step empirically is to determine to what degree retired military officers have been upholding the two foundational norms of civilian control and non-partisanship.
+To measure adherence to these norms, I develop three binary indicator variables that shed light on how well the retired military officers who publish in major US newspapers adhere to these norms.
+
+The first indicator measures whether a publication severely criticizes a sitting President or political figure to the point that such criticism goes beyond making recommendations or pointing out shortcomings.
+Inclusion of such an indicator is designed to measure how willing a retired military officer is to blatantly "take a shot" at a civilian leader.
+Publications that go far beyond making a recommendation, and that actually are designed to sway the public's opinion regarding a specific political leader, are coded positively.
+Including such an indicator, moreover, enables measurement of how well retired military officers adhere to the principle of civilian control of the military.
+This is not to suggest that every single instance of a retired military officer criticizing a civilian leader is a crisis of civilian control of the military.
+However, including such an indicator will allow how frequent blatant criticism occurs.
+
+A second indicator is whether a publication clearly endorses or rebukes a partisan politician or party, including candidates for office.
+Including this indicator allows the researcher to measure the degree to which the principle of non-partisanship of the military has been upheld in the publications authored by retired military officers over time.
+
+The third and final indicator measures whether a publication calls on Congress or a President to undertake specific legislative actions, such as ratifying a treaty or passing a bill or budget.
+This indicator does not directly measure either non-partisanship or civilian control of the military, but it does measure how comfortable retired military officers feel being involved in a part of the political process, that of calling for and/or making legislative action, which is an experience that virtually all military officers do not develop during their time in uniform.^[A small number of officers may have experience working on the hill, or serving as a Congressional liasion, but I argue that this is different from actually calling upon a politician to engage in adopting a stance regarding a specific piece of legislation.] This is not to say that military officers on active duty or during retirement either should be or in fact are oblivious to this aspect of the political process.
+Rather, including an indicator for assertiveness (calls to pursue legislative action) measures to what extent the retired military author believes his or her military experience serves as an appropriate basis for calling for legislative action.
+
+Because some retired military officers in fact do go on to serve in political administrations, or run for office and serve as politicians themselves, it is important to denote whether a retired military officer wrote a publication while serving in a role that legitimately would authorize the military officer to play a more prominent role in the legislative process.
+Therefore, a fourth and final indicator control for whether the author serves in such a role at the time of publication is included.
+
+Figure 4, below, shows over time the proportion of observations in a given year which either heavily criticizes a civilian leader, takes a clear partisan stance, or calls for specific legislative action to be taken.
+
+```{r second level analysis, include = FALSE}
+
+# Show all articles and authors who published articles on Domestic Policy and Social Policy 
+adjdataset %>% arrange(PubYr) %>% 
+  filter(Topic=="DomesticPolicy" | Topic=="SocialPolicy")  %>% 
+  select (Author, Title, Source, PubYr, Topic) %>% 
+  print()
+
+# Show all articles where Leg_Act, Crit_Insub, or Endorse_Partisan was equal to 1
+## (in other words, articles in which the author called for legislative activity, criticized, or made
+## an explicitly partisan statement)
+
+adjdataset %>% arrange(PubYr) %>% 
+  filter(Leg_Act== 1 | Crit_Insub== 1 | Endorse_Partisan ==1)  %>% 
+  select (Author, Title, Source, PubYr, Topic, Leg_Act, Crit_Insub, Endorse_Partisan) %>% 
+  print()
+
+# create a df called quest_prof that contains articles that either criticized or were #expressly partisan, as well as whether someone was in a role at the time that may have # made such behavior ok
+
+quest_prof <-adjdataset %>% arrange(PubYr, Topic) %>% 
+  filter(Crit_Insub== 1 | Endorse_Partisan ==1)  %>% 
+  select (Author, Title, Source, PubYr, Topic, Crit_Insub, Endorse_Partisan, Role) %>% 
+  view()
+
+# Now create two variables, leg_crit_part and crit_part that take a 1 if certain 
+# conditions are met.  This will come in handy for the glm regressions that
+# will be conducted later.
+
+adjdataset <- adjdataset %>% 
+  mutate(
+    leg_crit_part=case_when(
+      (Leg_Act==1 | Crit_Insub ==1 | Endorse_Partisan==1 & Role ==0) ~ 1,
+      (Leg_Act != 1 & Crit_Insub != 1 & Endorse_Partisan !=1 | Role==1) ~ 0
+    ),
+    crit_part=case_when(
+      (Crit_Insub ==1 | Endorse_Partisan==1 & Role ==0) ~ 1,
+      (Crit_Insub != 1 & Endorse_Partisan != 1 | Role ==1) ~ 0
+    )
+  ) %>% 
+  view()
+
+```
+
+```{r build plots, include = TRUE, warning = FALSE, error=FALSE, echo=FALSE}
+ggplot(data=adjdataset, mapping=aes(x=PubYr, y=prop_yr)) + geom_point() + geom_smooth() + labs(title="Figure 4: Proportion of Op-Eds Published Which Call for Legislative Action, Criticize Civilian Leaders, or Take Clear Partisan Stances", x="Year", y="Annual Proportion", color=NULL) + geom_vline(xintercept = 1981, size=.2, colour="tomato") +
+  geom_vline(xintercept=1989, size=.2, colour="tomato") + 
+  geom_vline(xintercept=1993, size=.2, colour="dodgerblue") + geom_vline(xintercept=2001, size=.2, colour="tomato") + 
+  geom_vline(xintercept=2009, size=.2, colour="dodgerblue") +
+  geom_vline(xintercept=2017, size=.2, colour="tomato") +
+  geom_text(aes(x=1981, label="Reagan",y=.2), colour="tomato", angle=90, vjust=1, text=element_text(size=9)) + 
+  geom_text(aes(x=1989, label="Bush 41",y=.2), colour="tomato", angle=90, vjust=1, text=element_text(size=9)) + 
+  geom_text(aes(x=1993, label="Clinton",y=.4), colour="dodgerblue", angle=90, vjust=1, text=element_text(size=9)) + 
+  geom_text(aes(x=2001, label="Bush 43",y=.3), colour="tomato", angle=90, vjust=1, text=element_text(size=9)) + 
+  geom_text(aes(x=2009, label="Obama",y=.4), colour="dodgerblue", angle=90, vjust=1, text=element_text(size=9)) + 
+  geom_text(aes(x=2017, label="Trump",y=.05), colour="tomato", angle=90, vjust=1, text=element_text(size=9))
+
+```
+The smoothed trend line in blue shows a story that is similar to the frequency at which retired military officers address domestic political topics.
+Not until the 1990s did retired military officers frequently criticize civilian leaders, call for express legislative action, or take clear partisan stances in the publications they authored.
+Additionally, there is a localized minimum in the trend line around the year 2012 before a sharp increase that has persisted until the present day.
 
 \singlespace
+# Determinants of Changing Military Role Beliefs
+\doublespace
 
-\hypertarget{the-second-layer---failing-to-uphold-professionalism}{%
-\section{The Second Layer - Failing to Uphold Professionalism?}\label{the-second-layer---failing-to-uphold-professionalism}}
+The story told in Figures 3 and 4, in short, is that retired military officers are addressing domestic political topics today more frequently and at a greater rate relative to other topics than they did in previous decades, and that they are criticizing civilian leaders, taking partisan stances, and calling for legislative action at greater rates than they previously did.
+Thus, the body of retired military officers is more assertive, at least in how it engages with the five major newspapers examined in the data, than it was in previous time periods.
+Having established these trends, the next task is to explain why they have formed.
+
+Theoretically, there are at least two major drivers of military role beliefs, including the level of prestige held by the military in American society, and the degree of political polarization that exists.^[In the theory chapter of my dissertation, I derive why these are the main drivers of military role beliefs.]
+
+To understand how measurements of political polarization have changed over time, I rely on the DW-NOMINATE (dynamic-weight, nominal three step estimation) scores that were originally developed by the political scientists Keith Poole and Howard Rosenthal in the early 1980s [@jeffrey_b_lewis_voteview_2020].
+The DW-NOMINATE data are one measure that uses roll-call voting data for each member of congress to assign a score to the member of congress on a liberal-conservative dimension, ranging from -1 (extreme liberal) to 1 (extreme conservative) [@jeffrey_b_lewis_voteview_2020].
+The data is frequently relied upon by academics, journalists, and the media to understand how a variety of political bodies, to include Congress and political parties, have changed over time.
+Over the time period examined in this paper, polarization in each chamber of Congress, the House of Representatives and the Senate, has increased [@jeffrey_b_lewis_voteview_2020].
+The distance in party mean DW-NOMINATE scores in the late 1970s was just less than .6, while the distance today for both chambers is above .8 [@jeffrey_b_lewis_voteview_2020].
+
+Changes in the institutional credibility of the American military institution can be seen by examining public opinion data.
+Pollsters survey the American public annually to gauge their trust in a variety of institutions.
+Using Gallup opinion polling, one sees that since the late 1970s, the credibility of the American military as an institution has gradually increased [@gallup_polls_confidence_2020].
+In 1979, 54% of those surveyed expressed either a great deal or quite a lot of confidence in the military, while in 2020, 72% expressed the same [@gallup_polls_confidence_2020].
+These trends in political polarization and confidence in the military are displayed in Figure Four, below.
+The top and middle graph show changes in means between Republicans and Democrats in the House of Representatives (top graph) and Senate (middle graph) since 1979, while the bottom graph shows changes in the percentage of surveyed Americans who reported either a great deal of or a lot of confidence in the military institution over the same time period.
+
+```{r figure 4, include=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
+#adjust the data to account for polarization by creating a new dataset called polaradjdataset.  
+#In this dataset, create variables that capture ideological distance in each the House and Senate along each of the two dimensions according to Poole and Rosenthal's DW-NOMINATE data
+
+# scale the data by multiplying by 100
+
+polaradjdataset <-adjdataset %>% mutate (
+  House_distance_dim1=abs(House_Dem_mean_nom1-House_Rep_mean_nom1)*100,
+  Sen_distance_dim1=abs(Sen_Dem_mean_nom1-Sen_Rep_mean_nom1)*100,
+  House_distance_dim2=abs(House_Dem_mean_nom2-House_Rep_mean_nom2)*100,
+  Sen_distance_dim2=abs(Sen_Dem_mean_nom2-Sen_Rep_mean_nom2)*100
+) %>% view ()
+
+# Create a variable called polar_and_cred that uses the melt function (from the library reshape2")
+#note that the id.var will serve as the common axis for these 
+polar_and_cred <- melt(subset(polaradjdataset, select=c(PubYr, House_distance_dim1, Sen_distance_dim1, Inst_Cred_Gallup)), id.var="PubYr")
+
+# 
+ggplot(polar_and_cred, aes(x=PubYr, y=value)) + geom_line(aes(color=variable)) + facet_grid(variable~., scales="free_y") + theme(legend.position = "none") +
+  labs(
+    title = "Figure Four: Distance Between Party Means in the House,\n the US Senate,\n and American Confidence in the Military, 1979-2020"
+  )
+```
+
+While it is possible that the trend in increasing institutional credibility of the US military over the past four decades may be at least partially responsible for the increasing trend in the number of opinion pieces published by retired military officers, a trend that the data supported earlier as shown in Figure 2, rising institutional credibility alone does not adequately explain the trends shown in Figure 3: that there has been a sharp increase over the past four decades in retired military officers penning pieces that address non-traditional military topics that lay outside the bounds of relevant military expertise, or pieces that express partisan sentiment.
+The theory laid out earlier in this paper predicted that as political polarization increases, three mechanisms are at work which have strong impacts on the military professional: it becomes increasingly difficult to remain politically neutral as partisan conflicts subsume a greater number of issues; as this occurs, strong feelings of dislike or disdain surface against those who hold on opposing view; and alterations in one's sense of loyalty arise, particularly as issues that individuals hold to most dearly raise the salience of one's identity.
+
+To test if changes in political polarization adequately explain the changes in the qualitative nature (increasing trends of potential violations of military professionalism) of opinion pieces authored by retired military officers, I conduct a logit regression of the variable _questionable_professionalism_ while holding institutional trust in the military constant at its mean (71 percent) and varying the value of the mean ideological distance between both major political parties in the House between 1979-2020.
+Conducting such a regression allows one to examine, given the data assembled, the predictive likelihood of publishing an opinion piece that violates military professionalism given varying levels of political polarization and constant institutional credibility in the military.
+These results are displayed in Figure 5, below.
+
+```{r logit analysis, include=FALSE, echo=FALSE}
+
+#next, conduct logit regressions where the DV is first, values of leg_crit_part and
+# then, where the DV is crit_part
+
+#model 1
+
+mod1 <- glm(leg_crit_part ~ Inst_Cred_Gallup + House_distance_dim1 + AuthSex + AuthRank + Topic, 
+  family=binomial(link="logit"), 
+  data=polaradjdataset
+  )
+
+summary(mod1)
+tidy(mod1)
+glance(mod1)
+
+# model 2
+mod2 <- glm(crit_part ~ Inst_Cred_Gallup + House_distance_dim1 + AuthSex + AuthRank + Topic, 
+    family=binomial(link="logit"),
+    data=polaradjdataset
+    )
+
+summary(mod2)
+tidy(mod2)
+glance(mod2)
+
+# model 3 - remove topic as an IV because there is significant correlation between
+# one type of topic (domestic policy) and the DV
+
+mod3 <- glm(leg_crit_part ~ Inst_Cred_Gallup + House_distance_dim1 + AuthSex + AuthRank,
+    family=binomial(link="logit"),
+    data=polaradjdataset
+    )
+
+summary(mod3)
+tidy(mod3)
+glance(mod3)
+
+# model 4 - same as mod3, but change the DV to crit_part
+
+mod4 <- glm(crit_part ~ Inst_Cred_Gallup + House_distance_dim1 + AuthSex + AuthRank,
+    family=binomial(link="logit"),
+    data=polaradjdataset
+    )
+
+summary(mod4)
+tidy(mod4)
+glance(mod4)
+
+#model 5 - now check for robustness using aff_polar as the polarization variable
+mod5 <- glm(leg_crit_part ~ Inst_Cred_Gallup + aff_polar + AuthSex + AuthRank,
+    family=binomial(link="logit"),
+    data=polaradjdataset
+    )
+
+summary(mod5)
+
+mod6 <- glm(crit_part ~ Inst_Cred_Gallup + aff_polar + AuthSex + AuthRank,
+    family=binomial(link="logit"),
+    data=polaradjdataset
+    )
+
+summary(mod6)
+```
+\singlespace
+```{r regression table, include=TRUE, echo=FALSE, results="asis"}
+stargazer(mod3, mod4, mod5, mod6,
+          title="Table 3: Logistic Regression Results", 
+          column.labels = c('Model 3','Model 4','Model 5', 'Model 6'),
+          covariate.labels = c('Institutional Credibility', 'Congressional Polarization', 'Affective Polarization', 'Male', 'Mixed Authorship', 'Unknown Authorship', 'Rank'),
+          align=TRUE)
+
+```
 
@@
 
-Conclusively determining whether an observation failed to uphold military professionalism is once again an imperfect and somewhat subjective evaluation.
-However, the work of academic theorists and decades of military norms make clear that at a minimum, military professionalism emphasizes behavior by military officers that is relevant to one's military expertise, as well as absent of partisan expression or preferences.
-Recall that Retired Army General Douglas MacArthur cautioned the cadets at West Point in 1962 about involving themselves in matters that do not require a military solution (in MacArthur's words, of ``politics grown to corrupt,'' ``morals grown too low,'' and ``whether our personal liberties are as firm and complete as they should be'') \autocite{macarthur_duty_1962}, and that more recently, in 2016, Retired Army General Martin Dempsey lamented the participation of retired generals in the political process by endorsing political candidates and appearing at the conventions of political parties:
+The graph clearly shows a non-linear relationship between increasing political polarization and the likelihood that an opinion piece authored by a retired military officer will potentially violate aspects of military professionalism.
+While the confidence intervals at each point estimate are relatively large, the graph demonstrates that the predicted probability of violating military professionalism is different between relative extremes along the x axis.
+In other words, the political polarization levels that exist today, as measured by the ideological distance between members of the two prominent political parties in the house, is predicted to have resoundingly different impacts on military professionalism than the polarization levels that existed several decades ago.
+The polarization levels that existed several decades ago, in the late 1970s and 1980s, may have had negligible, if any impact, on the likelihood of a retired military officer violating military professionalism in the content of a published opinion piece, but the political landscape is simply different today.
+Contemporary polarization levels, insofar as the predictive power of the model is concerned, changed the environment: at contemporary polarization levels (the far right of the x axis), there is roughly a 17% chance of a retired military officer's published opinion piece violating the standards and norms of military professionalism, with confidence intervals that range from roughly 11% to nearly 25%.
+
+Given these dynamics, as well as the results of both Figures 3 and 5, ample evidence exists to confirm the second hypothesis, which stated that as a result of increasing political polarization, there will be an increase in recent years in the number of op-eds written about topics that are not relevant to military expertise, or that reveal inappropriate partisan sentiments, among authors which are a vital component of the institutional military, retired military officers.
+
+## Does Combat Attenuate Polarization's Impact on Military Professionalism?
+
+Why did Figure 3 reveal a sharp drop between approximately 2005-2011 in both the real number and proportion of annual opinion pieces authored by retired military officers that address non-traditional military topics, such as social policy that impacts the military, domestic policy, and political administrations/figures?
+What was it about this time period that made it unlike any other span of years across the past four decades?
+
+One compelling answer is that between these years, roughly 2005-2011, the military was heavily involved in combat operations to a point such that the institution was taxed tremendously, thereby necessitating that it remain focused on its wartime mission.
+Repeated deployments during the Global War on Terrorism, which included substantial surges of forces to both Iraq (2007-2008) and Afghanistan (2010-2011), as well as sustained US casualties resulting from these conflicts, possibly shaped the military institution so strongly such that even retired military officers acknowledged that the military's institutional energy required a laser-sharp focus on combat requirements.
+This is not to suggest the military was not heavily engaged in conflicts nor active around the glove at other times throughout the past four decades (which we know is not the case), but to instead suggest that even retired military officers are perhaps more sensitive to combat requirements and missions undertaken by the military as higher levels of continued operational strain and duress are endured.
+Though more data is required to fully test this argument, other scholars have argued similarly in that periods of transition, especially away from major combat operations and into eras of downsizing and relative peace, often seem to test the very essence of military professions in significant ways [@snider_once_2012].
+
+# Conclusion: Will the Military Remain Professional in a Polarized Setting?
+
+The goal of this paper was to explore and examine the ways in which political polarization and institutional credibility have impacted the professionalism of the military over the past fifty years.
+Examining the published opinion pieces authored by retired military officers in major US newspapers revealed that although the vast number of total opinion publications uphold the military's norms and traditions concerning military professionalism, a couple of worrisome trends are evident in the data: first, retired military officers are increasingly writing about topics outside of the umbrella of traditional national security topics, and second, retired officers are expressing partisan preferences at a greater rate than they previously did.
+What do these trends mean, and should they cause concern?
+If so, what should be done about it?
+
+First, these findings illuminate the power and pervasiveness of increasing political polarization, and in particular, its impact not only on political institutions, but also on professions, which the military prides itself on being.
+Thus, if the military, which has for decades adopted and promulgated norms of apolitical and nonpartisan expression, has been impacted by rising political polarization, it stands to reason that other political institutions and professions also have been impacted.
+While the specific manifestations of political polarization's increase and spread in non-military institutions and professions might be different, the fact that the military is experiencing a noticeable impact as a result of increasing political polarization speaks to its significant power and influence.
+
+Indeed, evident in the content of opinion pieces authored by retired military officers over the past four decades are the three features of contemporary American political polarization developed in the theory proffered earlier in this paper: the loss of politically neutral space as more issues enter the partisan battleground, the creation of strong feelings of dislike for those who hold to opposing views on these issues, and perhaps worst for the military, alterations in an individual's concept of loyalty, which arises as the salience of one's identity is raised as the issues that an individual holds most dearly to are increasingly at stake.
+Moreover, the data supported what most scholars of American politics have argued for: a realization that polarizing politics across America's political landscape will likely not disappear anytime soon, regardless of which party holds the nation's highest office.
+
+Second, these findings underscore the challenge that the military as a profession has in "reigning in" its retired officers and compelling them to stay out of partisan battles, and out of politics more broadly.
+Among the features and traits of a profession is that it is self-policing; however, the data indicates that current efforts by active military leaders to police and clean up the behavior of its retired elites is not working.
+Indeed, the letters and pleas of previous military leaders to retired military officers, including those written by now Retired Army General Martin Dempsey and Retired Marine Corps General Joseph Dunford, both of which were referenced in this paper, are simply not working.
+This suggests that stronger and more creative efforts are required.
+
+Third, these findings demonstrate that the phrase "politicization of the military," which has long been a proper concern of academics, journalists, the military, and citizens alike, should not be intellectually confined to notions of how politicians or other political leaders political leaders attempt to harness or exploit the military for political purposes or advantages.
+Though such efforts do occur and are a genuine cause for concern that continue to deserve both scholarly and practical attention, the evidence presented in this paper shows that "politicization of the military" is also occurring at the hands and through the words of particular retired military officers.
+Gaining a better understanding of the impacts of political polarization on the military profession remains a critical area for future research.
+Other scholars such as Michael Robinson have explored the specific ways that polarization is impacting the interactions between the military and the public [@robinson_michael_danger_2018].
+Though this paper focused on the ways that polarization may impact professional military behavior, it recognized but stopped short of discussing polarization's subsequent impact on various "military outcomes," which may arguably include even more decisive events, such as achieving victory or defeat in battle, or the ability to formulate sound strategic advice.
+These outcomes and more, and the role of polarization in influencing them, deserve to be explored further.
+
+Maintaining credibility and trust with the American public is crucial for the military, but should a polarized political climate endure, the military needs to also understand that credibility and trust can be abused to the point that they also can constitute significant risks.
+One such risk is that as polarization worsens, military leaders, to include military retirees, abuse the high level of credibility they currently enjoy with the American public to advocate for obviously partisan positions.
+Such actions, though, in a polarized setting, might actually result in an increase in the credibility of the military institution, at least among citizens who share the same partisan sentiments advocated for by military leaders, which might, somewhat perversely, entice more of the same behavior.
+A second risk, stemming from the first, is that a polarized setting sets the stage for the military institution to eventually openly jockey with civilian leaders, either for power or for control, a concern that generations of scholars have warned us about [@finer_man_1962; @brooks_shaping_2008-1].
+Were either of these risks to materialize, the character and nature of the American military as an institution, and likely, American democracy, would be irrevocably changed.
+
+There are no simple solutions for the military with respect to the risks posed by rising political polarization.
+Two possible ideas, however, are apropos to this discussion.
+One involves identity, and the other, credibility.
+First, our nation's military is composed of soldiers who come from every walk of life throughout America, and accordingly, in the years to come, many future soldiers and leaders will bear the features of a polarized society.
+Thus, the military's challenge, and one which it must embrace itself for immediately, is integrating the multiple identities each soldier carries with them as people in such a way that each can still proudly wear the uniform and bear the identity of an American service member.
+Though this challenge is not new for the US military, increasing political polarization has significantly raised the level of difficulty the military faces in tackling such a vital challenge.
+Scholars who have studied multi-ethnic militaries across the world have argued that it is military leaders, not civilians, who ultimately devise effective solutions to the problems posed by soldiers who carry multiple loyalties, and interestingly, only as a result of combat requirements [@peled_question_1998].
+Perhaps the American challenge is even greater in that at least for the near term future, there is no major war that will necessitate the coming together of disparate loyalties and identities among future soldiers.
+
+The second idea, and one that may sound like anathema to modern military leaders, is that perhaps the level of trust and credibility between the American public and the military institution should not be the ultimate indicator of the military's performance as a profession.
+In suggesting this, I am not suggesting that credibility and trust are not important or that the concept should be disregarded entirely.
+The American public is, after all, the client of the American military and ultimately, it is their opinion that counts.I am suggesting, however, that for the reasons laid out in this paper, credibility and trust are influenced by the mechanisms through which polarization works, to include the loss of politically-neutral space, increased negative feelings toward those who hold differing opinions, and most sinister of all, alterations in one's concept of loyalty.
+For these reasons, I am simply offering the notion that _credibility can be deceiving_ in polarized times.
+Now, more than ever before, military leaders will need to rely on individual moral courage over and above any other trait in order to perform their duties as military professionals.
+\singlespace
+\pagebreak
+
+#Appendix 1
+-sort out the table with Kable and make it look good-
+***Move the Actual Table of these observations to the Appendix, and then drive on with the Regression.  Good work!***
 
 \singlespace
+```{r table3, fig.align = "center", include=TRUE, echo=FALSE}
+#Need to adjust the width and spacing of this table
+quest_prof %>%
+  kable(format="latex", caption="Op-Eds authored by Retired Military Officers that Strongly Criticized Civilian Leaders, or were Expressly Partisan, 1979-2020", longtable=T) %>%
+  kable_styling(font_size=10, latex_options="striped", repeat_header_continued=T) %>% 
+  column_spec(1, width="9.5em") %>% 
+  column_spec(2, width="15em") %>% 
+  column_spec(3:4, width="2.5em") %>%
+  column_spec(5, width="5em") %>%
+  column_spec(6, width="6em") %>%
+  row_spec(0, bold=T)
+```
 
-\begin{quote}
-\textbf{The American people should not wonder where their military leaders draw the line between military advice and political preference.
-And our nation's soldiers, sailors, airmen and Marines should not wonder about the political leanings and motivations of their leaders.
-As generals, they have an obligation to uphold our apolitical traditions.
-They have just made the task of their successors -- who continue to serve in uniform and are accountable for our security - more complicated} \autocite{dempsey_military_2016}.
-\end{quote}
-
-\doublespace
-
-Relying on these two central ideals of military professionalism, relevant military expertise and non-partisan/apolitical expression, allows the researcher to ask two important questions: first, is the author addressing a topic related and applicable to the officer's military expertise, and second, is the author making an argument without clearly conveying or expressing political or partisan preferences?
-If the answer to either of these two questions is ``no,'' then the publication might violate the concept of military professionalism.
-I therefore develop two new variables, \emph{questionable professionalism} and \emph{justification}.
-The first variable is binary and codes as `1' the observations which may violate the concept of military professionalism, and the second variable briefly captures the basis for a possible violation, i.e.~whether it is a violation of topic or of expression of political sentiment.
-The results of this analysis on the remaining 51 observations are displayed in Table 4, below.
-Several observations are worth noting.
-
-\singlespace
-\begingroup\fontsize{10}{12}\selectfont
-
-\begin{longtable}[t]{>{\raggedright\arraybackslash}p{9.5em}|>{\raggedright\arraybackslash}p{15em}|>{\raggedright\arraybackslash}p{2.5em}|>{\raggedleft\arraybackslash}p{2.5em}|>{\raggedright\arraybackslash}p{5em}|>{\raggedleft\arraybackslash}p{6em}|r|r}
-\caption{\label{tab:table4}Op-Eds authored by Retired Military Officers that Strongly Criticized Civilian Leaders, or were Expressly Partisan, 1979-2020}\\
-\hline
-\textbf{Author} & \textbf{Title} & \textbf{Source} & \textbf{PubYr} & \textbf{Topic} & \textbf{Crit\_Insub} & \textbf{Endorse\_Partisan} & \textbf{Role}\\
-\hline
-\cellcolor{gray!6}{Yancey, William} & \cellcolor{gray!6}{General Bashing, Then and Now} & \cellcolor{gray!6}{WSJ} & \cellcolor{gray!6}{1989} & \cellcolor{gray!6}{CivMilBalance} & \cellcolor{gray!6}{1} & \cellcolor{gray!6}{0} & \cellcolor{gray!6}{0}\\
-\hline
-Beckwith, Charlie & Somalia's Needless Deaths & WSJ & 1993 & Warfighting & 1 & 0 & 0\\
-\hline
-\cellcolor{gray!6}{Odom, William} & \cellcolor{gray!6}{Invade, Don't Bomb} & \cellcolor{gray!6}{WSJ} & \cellcolor{gray!6}{1994} & \cellcolor{gray!6}{Warfighting} & \cellcolor{gray!6}{1} & \cellcolor{gray!6}{0} & \cellcolor{gray!6}{0}\\
-\hline
-McCain, John & In Bosnia, Another Mistake & WSJ & 1994 & Warfighting & 1 & 0 & 1\\
-\hline
-\cellcolor{gray!6}{Odom, William} & \cellcolor{gray!6}{One Year?
-In Bosnia?} & \cellcolor{gray!6}{NYT} & \cellcolor{gray!6}{1995} & \cellcolor{gray!6}{Warfighting} & \cellcolor{gray!6}{1} & \cellcolor{gray!6}{0} & \cellcolor{gray!6}{0}\\
-\hline
-Odom, William & Chechnya, Freedom, and the Voice of Yeltsin Past & WaPO & 1996 & ForeignPolicy & 1 & 0 & 0\\
-\hline
-\cellcolor{gray!6}{Mcdonough, James} & \cellcolor{gray!6}{Clinton's Contempt for U.S.
-Soldiers} & \cellcolor{gray!6}{WSJ} & \cellcolor{gray!6}{1998} & \cellcolor{gray!6}{Warfighting} & \cellcolor{gray!6}{1} & \cellcolor{gray!6}{0} & \cellcolor{gray!6}{0}\\
-\hline
-Peters, Ralph & It's Wonks vs. Warlords, and Guess Who Wins Again? & WSJ & 1998 & Warfighting & 1 & 0 & 0\\
-\hline
-\cellcolor{gray!6}{Peters, Ralph} & \cellcolor{gray!6}{A Question of Leadership} & \cellcolor{gray!6}{WSJ} & \cellcolor{gray!6}{1998} & \cellcolor{gray!6}{Warfighting} & \cellcolor{gray!6}{1} & \cellcolor{gray!6}{0} & \cellcolor{gray!6}{0}\\
-\hline
-Krulak, Charles & Don't Politicize the Joint Chiefs & WSJ & 2000 & CivMilBalance & 0 & 1 & 0\\
-\hline
-\cellcolor{gray!6}{Odom, William} & \cellcolor{gray!6}{Buchanan Has it Backwards on Globalization} & \cellcolor{gray!6}{WSJ} & \cellcolor{gray!6}{2000} & \cellcolor{gray!6}{ForeignPolicy} & \cellcolor{gray!6}{0} & \cellcolor{gray!6}{1} & \cellcolor{gray!6}{0}\\
-\hline
-Shalikashvili, John & Kerry Proceeds With Caution & USA Today & 2004 & BdgtWpnsTrps & 0 & 1 & 0\\
-\hline
-\cellcolor{gray!6}{Clark, Wesley} & \cellcolor{gray!6}{Medals of Honor} & \cellcolor{gray!6}{NYT} & \cellcolor{gray!6}{2004} & \cellcolor{gray!6}{DomesticPolicy} & \cellcolor{gray!6}{0} & \cellcolor{gray!6}{1} & \cellcolor{gray!6}{0}\\
-\hline
-Shalikashvili, John & Old Soldiers Don't Have to Fade Away & WSJ & 2004 & DomesticPolicy & 0 & 1 & 0\\
-\hline
-\cellcolor{gray!6}{Franks, Tommy} & \cellcolor{gray!6}{Right Leader, Right Time} & \cellcolor{gray!6}{WSJ} & \cellcolor{gray!6}{2004} & \cellcolor{gray!6}{DomesticPolicy} & \cellcolor{gray!6}{0} & \cellcolor{gray!6}{1} & \cellcolor{gray!6}{0}\\
-\hline
-Whitlow, William & The Price of Giving Bad Advice & WaPO & 2004 & Warfighting & 1 & 0 & 0\\
-\hline
-\cellcolor{gray!6}{Franks, Tommy} & \cellcolor{gray!6}{War of Words} & \cellcolor{gray!6}{NYT} & \cellcolor{gray!6}{2004} & \cellcolor{gray!6}{Warfighting} & \cellcolor{gray!6}{0} & \cellcolor{gray!6}{1} & \cellcolor{gray!6}{0}\\
-\hline
-Clark, Wesley & Before It's Too Late in Iraq & WaPO & 2005 & Warfighting & 1 & 0 & 0\\
-\hline
-\cellcolor{gray!6}{Clark, Wesley} & \cellcolor{gray!6}{The Next Iraq Offensive} & \cellcolor{gray!6}{NYT} & \cellcolor{gray!6}{2005} & \cellcolor{gray!6}{Warfighting} & \cellcolor{gray!6}{1} & \cellcolor{gray!6}{0} & \cellcolor{gray!6}{0}\\
-\hline
-Eaton, Paul & A Top-Down Review for the Pentagon & NYT & 2006 & CivMilBalance & 1 & 0 & 0\\
-\hline
-\cellcolor{gray!6}{Crosby, John and McInerney, Thomas and Moore, Burton and Vallely, Paul} & \cellcolor{gray!6}{In Defense of Donald Rumsfeld} & \cellcolor{gray!6}{WSJ} & \cellcolor{gray!6}{2006} & \cellcolor{gray!6}{Support} & \cellcolor{gray!6}{1} & \cellcolor{gray!6}{0} & \cellcolor{gray!6}{0}\\
-\hline
-Murtha, John & Confessions of a 'Defeatocrat' & WaPO & 2006 & Warfighting & 1 & 0 & 1\\
-\hline
-\cellcolor{gray!6}{Eaton, Paul} & \cellcolor{gray!6}{An Army of One Less} & \cellcolor{gray!6}{NYT} & \cellcolor{gray!6}{2006} & \cellcolor{gray!6}{Warfighting} & \cellcolor{gray!6}{1} & \cellcolor{gray!6}{0} & \cellcolor{gray!6}{0}\\
-\hline
-Scales, Robert & A War the Pentagon Doesn't Want & WaPO & 2013 & Warfighting & 1 & 0 & 0\\
-\hline
-\cellcolor{gray!6}{Hayden, Michael and Mukasey, Michael} & \cellcolor{gray!6}{NSA Reform that Only ISIS Could Love} & \cellcolor{gray!6}{WSJ} & \cellcolor{gray!6}{2014} & \cellcolor{gray!6}{Warfighting} & \cellcolor{gray!6}{1} & \cellcolor{gray!6}{0} & \cellcolor{gray!6}{0}\\
-\hline
-McSally, Martha & Saving a Plane that Saves Lives & NYT & 2015 & BdgtWpnsTrps & 1 & 0 & 1\\
-\hline
-\cellcolor{gray!6}{Scales, Robert} & \cellcolor{gray!6}{Our Army is Breaking} & \cellcolor{gray!6}{WaPO} & \cellcolor{gray!6}{2015} & \cellcolor{gray!6}{BdgtWpnsTrps} & \cellcolor{gray!6}{1} & \cellcolor{gray!6}{0} & \cellcolor{gray!6}{0}\\
-\hline
-Petraeus, David and O'Hanlon, Michael & Afghanistan After Obama & WaPO & 2015 & Warfighting & 1 & 0 & 0\\
-\hline
-\cellcolor{gray!6}{Hayden, Michael} & \cellcolor{gray!6}{Classified Briefings and Candidates} & \cellcolor{gray!6}{NYT} & \cellcolor{gray!6}{2016} & \cellcolor{gray!6}{CivMilBalance} & \cellcolor{gray!6}{1} & \cellcolor{gray!6}{0} & \cellcolor{gray!6}{0}\\
-\hline
-Davis, Morris & Stop Meddling in the Bergdahl Case & NYT & 2016 & CivMilBalance & 1 & 0 & 0\\
-\hline
-\cellcolor{gray!6}{Hayden, Michael} & \cellcolor{gray!6}{Russia's Useful Fool} & \cellcolor{gray!6}{WaPO} & \cellcolor{gray!6}{2016} & \cellcolor{gray!6}{DomesticPolicy} & \cellcolor{gray!6}{1} & \cellcolor{gray!6}{0} & \cellcolor{gray!6}{0}\\
-\hline
-McCain, John & We Have a Stake in Syria, Yet We Have Done Nothing & WaPO & 2016 & ForeignPolicy & 1 & 0 & 0\\
-\hline
-\cellcolor{gray!6}{Hayden, Michael} & \cellcolor{gray!6}{Trump's Most Important New Partner} & \cellcolor{gray!6}{WaPO} & \cellcolor{gray!6}{2016} & \cellcolor{gray!6}{Warfighting} & \cellcolor{gray!6}{1} & \cellcolor{gray!6}{0} & \cellcolor{gray!6}{0}\\
-\hline
-Hayden, Michael & A Damaging Disregard for Intel & WaPO & 2016 & Warfighting & 1 & 0 & 0\\
-\hline
-\cellcolor{gray!6}{Hayden, Michael} & \cellcolor{gray!6}{The Travel Ban Hurts American Spies - and America} & \cellcolor{gray!6}{WaPO} & \cellcolor{gray!6}{2017} & \cellcolor{gray!6}{DomesticPolicy} & \cellcolor{gray!6}{1} & \cellcolor{gray!6}{0} & \cellcolor{gray!6}{0}\\
-\hline
-Hayden, Michael & Trump Proves He's Russia's Useful Fool & WaPO & 2017 & DomesticPolicy & 1 & 0 & 0\\
-\hline
-\cellcolor{gray!6}{Mullen, Mike} & \cellcolor{gray!6}{The Refugees We Need} & \cellcolor{gray!6}{NYT} & \cellcolor{gray!6}{2017} & \cellcolor{gray!6}{DomesticPolicy} & \cellcolor{gray!6}{1} & \cellcolor{gray!6}{0} & \cellcolor{gray!6}{0}\\
-\hline
-Mullen, Mike & Bannon Has No Place on the NSC & NYT & 2017 & ForeignPolicy & 1 & 0 & 0\\
-\hline
-\cellcolor{gray!6}{Hayden, Michael} & \cellcolor{gray!6}{Donald Trump is Undermining Intelligence Gathering} & \cellcolor{gray!6}{NYT} & \cellcolor{gray!6}{2017} & \cellcolor{gray!6}{Warfighting} & \cellcolor{gray!6}{1} & \cellcolor{gray!6}{0} & \cellcolor{gray!6}{0}\\
-\hline
-Hayden, Michael & The End of Intelligence & NYT & 2018 & DomesticPolicy & 1 & 0 & 0\\
-\hline
-\cellcolor{gray!6}{McRaven, William} & \cellcolor{gray!6}{Take My Security Clearance, too, Mr.
-President} & \cellcolor{gray!6}{WaPO} & \cellcolor{gray!6}{2018} & \cellcolor{gray!6}{DomesticPolicy} & \cellcolor{gray!6}{1} & \cellcolor{gray!6}{0} & \cellcolor{gray!6}{0}\\
-\hline
-Wilkerson, Lawrence and Wilson III, Isaiah and Adams, Gordon & Trump's Border Stunt is a Profound Betrayal of Our Military & NYT & 2018 & Warfighting & 1 & 0 & 0\\
-\hline
-\cellcolor{gray!6}{Nagl, John} & \cellcolor{gray!6}{Retired Generals Warned Us About Rumsfeld.
-Now They're Warning Us About Trump} & \cellcolor{gray!6}{WaPO} & \cellcolor{gray!6}{2019} & \cellcolor{gray!6}{DomesticPolicy} & \cellcolor{gray!6}{1} & \cellcolor{gray!6}{0} & \cellcolor{gray!6}{0}\\
-\hline
-Allen, John and Victor, David & Despite What Trump Says, Climate Change Threatens Our National Security & NYT & 2019 & DomesticPolicy & 1 & 0 & 0\\
-\hline
-\cellcolor{gray!6}{McRaven, William} & \cellcolor{gray!6}{Our Republic is Under Attack From the President} & \cellcolor{gray!6}{NYT} & \cellcolor{gray!6}{2019} & \cellcolor{gray!6}{DomesticPolicy} & \cellcolor{gray!6}{1} & \cellcolor{gray!6}{1} & \cellcolor{gray!6}{0}\\
-\hline
-McRaven, William & If Good Men Can’t Speak the Truth, We Should Be Deeply Afraid & WaPO & 2020 & DomesticPolicy & 1 & 0 & 0\\
-\hline
-\cellcolor{gray!6}{VanLandingham, Rachel and Corn, Geoffrey} & \cellcolor{gray!6}{Military Brass Finally Speaks Up on Trump} & \cellcolor{gray!6}{USA Today} & \cellcolor{gray!6}{2020} & \cellcolor{gray!6}{DomesticPolicy} & \cellcolor{gray!6}{1} & \cellcolor{gray!6}{0} & \cellcolor{gray!6}{0}\\
-\hline
-Seidule, Ty & West Point and its Cadets are not Campaign Props & WaPO & 2020 & DomesticPolicy & 1 & 0 & 0\\
-\hline
-\cellcolor{gray!6}{Vindman, Alexander} & \cellcolor{gray!6}{Coming Forward Ended My Career.
-I Still Believe Doing What's Right Matters} & \cellcolor{gray!6}{WaPO} & \cellcolor{gray!6}{2020} & \cellcolor{gray!6}{DomesticPolicy} & \cellcolor{gray!6}{1} & \cellcolor{gray!6}{0} & \cellcolor{gray!6}{0}\\
-\hline
-McRaven, William & Trump is Working to Actively Undermine the Postal Service - and Every Major U.S. Institution & WaPO & 2020 & DomesticPolicy & 1 & 1 & 0\\
-\hline
-\cellcolor{gray!6}{Vindman, Alexander and Gans, John} & \cellcolor{gray!6}{Trump Has Sold off America's Credibility for His Personal Gain} & \cellcolor{gray!6}{NYT} & \cellcolor{gray!6}{2020} & \cellcolor{gray!6}{DomesticPolicy} & \cellcolor{gray!6}{1} & \cellcolor{gray!6}{0} & \cellcolor{gray!6}{0}\\
-\hline
-McRaven, William & Biden Will Make America Lead Again; We Need a President with Decency and a Sense of Respect & WSJ & 2020 & DomesticPolicy & 1 & 1 & 0\\
-\hline
-\cellcolor{gray!6}{Allen, Thad} & \cellcolor{gray!6}{Trump is Failing to Provide For the Common Defense} & \cellcolor{gray!6}{WaPO} & \cellcolor{gray!6}{2020} & \cellcolor{gray!6}{DomesticPolicy} & \cellcolor{gray!6}{1} & \cellcolor{gray!6}{0} & \cellcolor{gray!6}{0}\\
-\hline
-Hayden, Michael et. Al & A Dangerous Purge & WaPO & 2020 & Warfighting & 1 & 0 & 0\\
-\hline
-\cellcolor{gray!6}{89 Former Defense Officials} & \cellcolor{gray!6}{The Military Must Never Be Used to Violate Constitutional Rights} & \cellcolor{gray!6}{WaPO} & \cellcolor{gray!6}{2020} & \cellcolor{gray!6}{Warfighting} & \cellcolor{gray!6}{1} & \cellcolor{gray!6}{0} & \cellcolor{gray!6}{0}\\
-\hline
-\end{longtable}
-\endgroup{}
 \doublespace
@@
 Second, and resulting from the first, is that across the entire data set, instances of potential violations of military professionalism in the op-eds published by retired military officers in major US newspapers is a somewhat rare occurrence.
-Indeed, the entire count of potentially questionable publications with respect to military professionalism across the data set is just 32 of 344 total observations, or 9.3\% of the entire sample.
+Indeed, the entire count of potentially questionable publications with respect to military professionalism across the data set is just 32 of 344 total observations, or 9.3% of the entire sample.
 Thus, ample support exists to confirm the first hypothesis, that the vast majority of opinion pieces authored by retired military officers uphold standards of military professionalism.
@@
 It is worth briefly commenting on several of these examples.
-In his \emph{Washington Post} editorial, ``Home Should Not Be a War Zone,'' Retired Army General Stanley McChrystal addresses the tragic nature of mass shootings in America.
-He states in the article, ``\ldots{}as a combat veteran and proud American, I believe we need a national response to the gun violence that threatens so many of our communities'' \autocite{mcchrystal_home_2016}.
-He later, in 2018, pens an op-ed entitled ``Good Riddance'' in which he discusses the evolution of his views on Robert E.
-Lee and other Confederate soldiers and symbols \autocite{mcchyrstal_stanley_good_2018}.
-Retired Army Brigadier General Ty Seidule similarly, in 2020, advocates for the renaming of several Army installations \autocite{seidule_ty_what_2020}.
-Other examples of observations coded as potentially violating military professionalism based on the topic of the publication include articles penned by Retired Navy Admiral Mike Mullen (entitled ``The Refugees We Need'') and co-authors Retired Army Lieutenant General Mark Hertling and Retired Naval Admiral Robert Natter (entitled ``Cutting Refugee Admissions Will Have Severe Consequences for the US Military'').
+In his _Washington Post_ editorial, "Home Should Not Be a War Zone," Retired Army General Stanley McChrystal addresses the tragic nature of mass shootings in America.
+He states in the article, "...as a combat veteran and proud American, I believe we need a national response to the gun violence that threatens so many of our communities" [@mcchrystal_home_2016].
+He later, in 2018, pens an op-ed entitled "Good Riddance" in which he discusses the evolution of his views on Robert E.
+Lee and other Confederate soldiers and symbols [@mcchyrstal_stanley_good_2018].
+Retired Army Brigadier General Ty Seidule similarly, in 2020, advocates for the renaming of several Army installations [@seidule_ty_what_2020].
+Other examples of observations coded as potentially violating military professionalism based on the topic of the publication include articles penned by Retired Navy Admiral Mike Mullen (entitled "The Refugees We Need") and co-authors Retired Army Lieutenant General Mark Hertling and Retired Naval Admiral Robert Natter (entitled "Cutting Refugee Admissions Will Have Severe Consequences for the US Military").
 
@@
 But I contend that it is also not wrong to insist that these and other contentious national issues are most appropriately solved by elected civilian officials and not by retired military officers.
-Interestingly, in July of 2020, General Mark Milley, the Chairman of the Joint Chiefs of Staff, was asked by US Representative Anthony Brown, D-MD, during a hearing held by the House Armed Services Committee, to ``comment on the naming of Army installations after Confederate soldiers,'' and whether the continued use of the names of these bases ``reflects the values'' that the military is trying to instill in the nation's armed forces \autocite{milley_mark_general_2020}?
+Interestingly, in July of 2020, General Mark Milley, the Chairman of the Joint Chiefs of Staff, was asked by US Representative Anthony Brown, D-MD, during a hearing held by the House Armed Services Committee, to "comment on the naming of Army installations after Confederate soldiers," and whether the continued use of the names of these bases "reflects the values" that the military is trying to instill in the nation's armed forces [@milley_mark_general_2020]?
 Milley's response was instructive.
-He offered the Congressmen the idea that even as the Department of Defense continues to take a hard look at the issue, the decision to remove those names should reflect the decision to adopt those names in the first place in that such a decision was ``a political decision'' \autocite{milley_mark_general_2020}.
+He offered the Congressmen the idea that even as the Department of Defense continues to take a hard look at the issue, the decision to remove those names should reflect the decision to adopt those names in the first place in that such a decision was "a political decision" [@milley_mark_general_2020].
 Milley's answer was the appropriate one, I contend, because it recognized that regardless of how he feels about this issue, the issue is not for him (or the military) to solve.
@@
 The second cluster refers to publications penned between 2004-2006.
-Most of these comprise part of or are related to the so-called ``Revolt of the Generals,'' which occurred near the 2004 US midterm elections and referred to sharp disagreements that entered the public arena between multiple US generals and admirals and President George W.
+Most of these comprise part of or are related to the so-called "Revolt of the Generals," which occurred near the 2004 US midterm elections and referred to sharp disagreements that entered the public arena between multiple US generals and admirals and President George W.
 Bush's Secretary of Defense, Donald Rumsfeld.
@@
 All but four of the observations in this cluster are coded as having potentially violated standards of military professionalism.
-Of note, one observation in this cluster entitled, ``Fists Raised at West Point'' and published in the Washington Post is actually a compendium of two anonymous letters, each one taking an opposite stance on the appropriateness of an incident in which a photograph at West Point's 2016 graduation ceremony captured a graduating female cadet raising a clenched a fist, an action which some saw as an inappropriate political gesture related to the growing Black Lives Matter movement \autocite{unknown_fists_2016}.
-
-\hypertarget{variation-in-military-professionalism-over-time}{%
-\subsection{Variation in Military Professionalism Over Time}\label{variation-in-military-professionalism-over-time}}
-
-Figure 3, below, plots both the total number and proportion of yearly opinion pieces authored by retired military officers that address topics of social policy affecting the military, domestic policy, and political administrations or figures.
-The points are the actual terms (number and proportion, respectively), while the blue line represents a smoothed line of best fit for the data.
-Both plots reveal that beginning with the first year of the data, 1979, there has a moderate increase until approximately 2005, followed by a decline until approximately 2011, followed by a sharp increase that has persisted until the present day.
-What explains these dynamics - both the increase that occurred between 1979 - 2005 as well as the sharper increase between 2012 - today - and the decline that occurred between roughly 2005 - 2011?
-According to the theory developed at the beginning of this paper, a significant reason stems from changes in political polarization and institutional credibility that have occurred over the past four decades.
-
-\includegraphics{Tahkv1_files/figure-latex/figure 3-1.pdf}
-
-To understand how measurements of political polarization have changed over time, I rely on the DW-NOMINATE (dynamic-weight, nominal three step estimation) scores that were originally developed by the political scientists Keith Poole and Howard Rosenthal in the early 1980s \autocite{jeffrey_b_lewis_voteview_2020}.
-The DW-NOMINATE data are one measure that uses roll-call voting data for each member of congress to assign a score to the member of congress on a liberal-conservative dimension, ranging from -1 (extreme liberal) to 1 (extreme conservative) \autocite{jeffrey_b_lewis_voteview_2020}.
-The data is frequently relied upon by academics, journalists, and the media to understand how a variety of political bodies, to include Congress and political parties, have changed over time.
-Over the time period examined in this paper, polarization in each chamber of Congress, the House of Representatives and the Senate, has increased \autocite{jeffrey_b_lewis_voteview_2020}.
-The distance in party mean DW-NOMINATE scores in the late 1970s was just less than .6, while the distance today for both chambers is above .8 \autocite{jeffrey_b_lewis_voteview_2020}.
-
-Changes in the institutional credibility of the American military institution can be seen by examining public opinion data.
-Pollsters survey the American public annually to gauge their trust in a variety of institutions.
-Using Gallup opinion polling, one sees that since the late 1970s, the credibility of the American military as an institution has gradually increased \autocite{gallup_polls_confidence_2020}.
-In 1979, 54\% of those surveyed expressed either a great deal or quite a lot of confidence in the military, while in 2020, 72\% expressed the same \autocite{gallup_polls_confidence_2020}.
-These trends in political polarization and confidence in the military are displayed in Figure Four, below.
-The top and middle graph show changes in means between Republicans and Democrats in the House of Representatives (top graph) and Senate (middle graph) since 1979, while the bottom graph shows changes in the percentage of surveyed Americans who reported either a great deal of or a lot of confidence in the military institution over the same time period.
-
-\includegraphics{Tahkv1_files/figure-latex/figure 4-1.pdf}
-
-While it is possible that the trend in increasing institutional credibility of the US military over the past four decades may be at least partially responsible for the increasing trend in the number of opinion pieces published by retired military officers, a trend that the data supported earlier as shown in Figure 2, rising institutional credibility alone does not adequately explain the trends shown in Figure 3: that there has been a sharp increase over the past four decades in retired military officers penning pieces that address non-traditional military topics that lay outside the bounds of relevant military expertise, or pieces that express partisan sentiment.
-The theory laid out earlier in this paper predicted that as political polarization increases, three mechanisms are at work which have strong impacts on the military professional: it becomes increasingly difficult to remain politically neutral as partisan conflicts subsume a greater number of issues; as this occurs, strong feelings of dislike or disdain surface against those who hold on opposing view; and alterations in one's sense of loyalty arise, particularly as issues that individuals hold to most dearly raise the salience of one's identity.
-
-To test if changes in political polarization adequately explain the changes in the qualitative nature (increasing trends of potential violations of military professionalism) of opinion pieces authored by retired military officers, I conduct a logit regression of the variable \emph{questionable\_professionalism} while holding institutional trust in the military constant at its mean (71 percent) and varying the value of the mean ideological distance between both major political parties in the House between 1979-2020.
-Conducting such a regression allows one to examine, given the data assembled, the predictive likelihood of publishing an opinion piece that violates military professionalism given varying levels of political polarization and constant institutional credibility in the military.
-These results are displayed in Figure 5, below.
-
-\% Table created by stargazer v.5.2.2 by Marek Hlavac, Harvard University. E-mail: hlavac at fas.harvard.edu
-\% Date and time: Thu, Dec 03, 2020 - 19:06:06
-\% Requires LaTeX packages: dcolumn
-
-\begin{table}[!htbp] \centering 
-  \caption{Table 3: Logistic Regression Results} 
-  \label{} 
-\begin{tabular}{@{\extracolsep{5pt}}lD{.}{.}{-3} D{.}{.}{-3} D{.}{.}{-3} D{.}{.}{-3} } 
-\\[-1.8ex]\hline 
-\hline \\[-1.8ex] 
- & \multicolumn{4}{c}{\textit{Dependent variable:}} \\ 
-\cline{2-5} 
-\\[-1.8ex] & \multicolumn{1}{c}{leg\_crit\_part} & \multicolumn{1}{c}{crit\_part} & \multicolumn{1}{c}{leg\_crit\_part} & \multicolumn{1}{c}{crit\_part} \\ 
- & \multicolumn{1}{c}{Model 3} & \multicolumn{1}{c}{Model 4} & \multicolumn{1}{c}{Model 5} & \multicolumn{1}{c}{Model 6} \\ 
-\\[-1.8ex] & \multicolumn{1}{c}{(1)} & \multicolumn{1}{c}{(2)} & \multicolumn{1}{c}{(3)} & \multicolumn{1}{c}{(4)}\\ 
-\hline \\[-1.8ex] 
- Institutional Credibility & -0.057^{*} & -0.047 & -0.024 & -0.019 \\ 
-  & (0.031) & (0.036) & (0.024) & (0.029) \\ 
-  & & & & \\ 
- Congressional Polarization & 0.084^{***} & 0.088^{***} &  &  \\ 
-  & (0.027) & (0.031) &  &  \\ 
-  & & & & \\ 
- Affective Polarization &  &  & 0.033^{**} & 0.037^{**} \\ 
-  &  &  & (0.013) & (0.015) \\ 
-  & & & & \\ 
- Male & -0.708 & -0.856 & -0.848 & -0.976 \\ 
-  & (1.256) & (1.254) & (1.262) & (1.263) \\ 
-  & & & & \\ 
- Mixed Authorship & 1.340 & 1.265 & 1.324 & 1.253 \\ 
-  & (1.755) & (1.739) & (1.759) & (1.746) \\ 
-  & & & & \\ 
- Unknown Authorship & -16.178 & -15.996 & -16.366 & -16.206 \\ 
-  & (787.186) & (798.975) & (789.107) & (799.686) \\ 
-  & & & & \\ 
- Rank & 0.171^{**} & 0.022 & 0.167^{**} & 0.018 \\ 
-  & (0.076) & (0.078) & (0.076) & (0.078) \\ 
-  & & & & \\ 
- Constant & -4.977^{**} & -4.946^{**} & -2.567 & -2.256 \\ 
-  & (2.155) & (2.477) & (1.945) & (2.173) \\ 
-  & & & & \\ 
-\hline \\[-1.8ex] 
-Observations & \multicolumn{1}{c}{389} & \multicolumn{1}{c}{389} & \multicolumn{1}{c}{388} & \multicolumn{1}{c}{388} \\ 
-Log Likelihood & \multicolumn{1}{c}{-180.151} & \multicolumn{1}{c}{-149.394} & \multicolumn{1}{c}{-182.477} & \multicolumn{1}{c}{-150.757} \\ 
-Akaike Inf. Crit. & \multicolumn{1}{c}{374.301} & \multicolumn{1}{c}{312.787} & \multicolumn{1}{c}{378.954} & \multicolumn{1}{c}{315.514} \\ 
-\hline 
-\hline \\[-1.8ex] 
-\textit{Note:}  & \multicolumn{4}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
-\end{tabular} 
-\end{table}
-
-The graph clearly shows a non-linear relationship between increasing political polarization and the likelihood that an opinion piece authored by a retired military officer will potentially violate aspects of military professionalism.
-While the confidence intervals at each point estimate are relatively large, the graph demonstrates that the predicted probability of violating military professionalism is different between relative extremes along the x axis.
-In other words, the political polarization levels that exist today, as measured by the ideological distance between members of the two prominent political parties in the house, is predicted to have resoundingly different impacts on military professionalism than the polarization levels that existed several decades ago.
-The polarization levels that existed several decades ago, in the late 1970s and 1980s, may have had negligible, if any impact, on the likelihood of a retired military officer violating military professionalism in the content of a published opinion piece, but the political landscape is simply different today.
-Contemporary polarization levels, insofar as the predictive power of the model is concerned, changed the environment: at contemporary polarization levels (the far right of the x axis), there is roughly a 17\% chance of a retired military officer's published opinion piece violating the standards and norms of military professionalism, with confidence intervals that range from roughly 11\% to nearly 25\%.
-
-Given these dynamics, as well as the results of both Figures 3 and 5, ample evidence exists to confirm the second hypothesis, which stated that as a result of increasing political polarization, there will be an increase in recent years in the number of op-eds written about topics that are not relevant to military expertise, or that reveal inappropriate partisan sentiments, among authors which are a vital component of the institutional military, retired military officers.
-
-\hypertarget{does-combat-attenuate-polarizations-impact-on-military-professionalism}{%
-\subsection{Does Combat Attenuate Polarization's Impact on Military Professionalism?}\label{does-combat-attenuate-polarizations-impact-on-military-professionalism}}
-
-Why did Figure 3 reveal a sharp drop between approximately 2005-2011 in both the real number and proportion of annual opinion pieces authored by retired military officers that address non-traditional military topics, such as social policy that impacts the military, domestic policy, and political administrations/figures?
-What was it about this time period that made it unlike any other span of years across the past four decades?
-
-One compelling answer is that between these years, roughly 2005-2011, the military was heavily involved in combat operations to a point such that the institution was taxed tremendously, thereby necessitating that it remain focused on its wartime mission.
-Repeated deployments during the Global War on Terrorism, which included substantial surges of forces to both Iraq (2007-2008) and Afghanistan (2010-2011), as well as sustained US casualties resulting from these conflicts, possibly shaped the military institution so strongly such that even retired military officers acknowledged that the military's institutional energy required a laser-sharp focus on combat requirements.
-This is not to suggest the military was not heavily engaged in conflicts nor active around the glove at other times throughout the past four decades (which we know is not the case), but to instead suggest that even retired military officers are perhaps more sensitive to combat requirements and missions undertaken by the military as higher levels of continued operational strain and duress are endured.
-Though more data is required to fully test this argument, other scholars have argued similarly in that periods of transition, especially away from major combat operations and into eras of downsizing and relative peace, often seem to test the very essence of military professions in significant ways \autocite{snider_once_2012}.
-
-\hypertarget{conclusion-will-the-military-remain-professional-in-a-polarized-setting}{%
-\section{Conclusion: Will the Military Remain Professional in a Polarized Setting?}\label{conclusion-will-the-military-remain-professional-in-a-polarized-setting}}
-
-The goal of this paper was to explore and examine the ways in which political polarization and institutional credibility have impacted the professionalism of the military over the past fifty years.
-Examining the published opinion pieces authored by retired military officers in major US newspapers revealed that although the vast number of total opinion publications uphold the military's norms and traditions concerning military professionalism, a couple of worrisome trends are evident in the data: first, retired military officers are increasingly writing about topics outside of the umbrella of traditional national security topics, and second, retired officers are expressing partisan preferences at a greater rate than they previously did.
-What do these trends mean, and should they cause concern?
-If so, what should be done about it?
-
-First, these findings illuminate the power and pervasiveness of increasing political polarization, and in particular, its impact not only on political institutions, but also on professions, which the military prides itself on being.
-Thus, if the military, which has for decades adopted and promulgated norms of apolitical and nonpartisan expression, has been impacted by rising political polarization, it stands to reason that other political institutions and professions also have been impacted.
-While the specific manifestations of political polarization's increase and spread in non-military institutions and professions might be different, the fact that the military is experiencing a noticeable impact as a result of increasing political polarization speaks to its significant power and influence.
-
-Indeed, evident in the content of opinion pieces authored by retired military officers over the past four decades are the three features of contemporary American political polarization developed in the theory proffered earlier in this paper: the loss of politically neutral space as more issues enter the partisan battleground, the creation of strong feelings of dislike for those who hold to opposing views on these issues, and perhaps worst for the military, alterations in an individual's concept of loyalty, which arises as the salience of one's identity is raised as the issues that an individual holds most dearly to are increasingly at stake.
-Moreover, the data supported what most scholars of American politics have argued for: a realization that polarizing politics across America's political landscape will likely not disappear anytime soon, regardless of which party holds the nation's highest office.
-
-Second, these findings underscore the challenge that the military as a profession has in ``reigning in'' its retired officers and compelling them to stay out of partisan battles, and out of politics more broadly.
-Among the features and traits of a profession is that it is self-policing; however, the data indicates that current efforts by active military leaders to police and clean up the behavior of its retired elites is not working.
-Indeed, the letters and pleas of previous military leaders to retired military officers, including those written by now Retired Army General Martin Dempsey and Retired Marine Corps General Joseph Dunford, both of which were referenced in this paper, are simply not working.
-This suggests that stronger and more creative efforts are required.
-
-Third, these findings demonstrate that the phrase ``politicization of the military,'' which has long been a proper concern of academics, journalists, the military, and citizens alike, should not be intellectually confined to notions of how politicians or other political leaders political leaders attempt to harness or exploit the military for political purposes or advantages.
-Though such efforts do occur and are a genuine cause for concern that continue to deserve both scholarly and practical attention, the evidence presented in this paper shows that ``politicization of the military'' is also occurring at the hands and through the words of particular retired military officers.
-Gaining a better understanding of the impacts of political polarization on the military profession remains a critical area for future research.
-Other scholars such as Michael Robinson have explored the specific ways that polarization is impacting the interactions between the military and the public \autocite{robinson_michael_danger_2018}.
-Though this paper focused on the ways that polarization may impact professional military behavior, it recognized but stopped short of discussing polarization's subsequent impact on various ``military outcomes,'' which may arguably include even more decisive events, such as achieving victory or defeat in battle, or the ability to formulate sound strategic advice.
-These outcomes and more, and the role of polarization in influencing them, deserve to be explored further.
-
-Maintaining credibility and trust with the American public is crucial for the military, but should a polarized political climate endure, the military needs to also understand that credibility and trust can be abused to the point that they also can constitute significant risks.
-One such risk is that as polarization worsens, military leaders, to include military retirees, abuse the high level of credibility they currently enjoy with the American public to advocate for obviously partisan positions.
-Such actions, though, in a polarized setting, might actually result in an increase in the credibility of the military institution, at least among citizens who share the same partisan sentiments advocated for by military leaders, which might, somewhat perversely, entice more of the same behavior.
-A second risk, stemming from the first, is that a polarized setting sets the stage for the military institution to eventually openly jockey with civilian leaders, either for power or for control, a concern that generations of scholars have warned us about \autocites{finer_man_1962}{brooks_shaping_2008-1}.
-Were either of these risks to materialize, the character and nature of the American military as an institution, and likely, American democracy, would be irrevocably changed.
-
-There are no simple solutions for the military with respect to the risks posed by rising political polarization.
-Two possible ideas, however, are apropos to this discussion.
-One involves identity, and the other, credibility.
-First, our nation's military is composed of soldiers who come from every walk of life throughout America, and accordingly, in the years to come, many future soldiers and leaders will bear the features of a polarized society.
-Thus, the military's challenge, and one which it must embrace itself for immediately, is integrating the multiple identities each soldier carries with them as people in such a way that each can still proudly wear the uniform and bear the identity of an American service member.
-Though this challenge is not new for the US military, increasing political polarization has significantly raised the level of difficulty the military faces in tackling such a vital challenge.
-Scholars who have studied multi-ethnic militaries across the world have argued that it is military leaders, not civilians, who ultimately devise effective solutions to the problems posed by soldiers who carry multiple loyalties, and interestingly, only as a result of combat requirements \autocite{peled_question_1998}.
-Perhaps the American challenge is even greater in that at least for the near term future, there is no major war that will necessitate the coming together of disparate loyalties and identities among future soldiers.
-
-The second idea, and one that may sound like anathema to modern military leaders, is that perhaps the level of trust and credibility between the American public and the military institution should not be the ultimate indicator of the military's performance as a profession.
-In suggesting this, I am not suggesting that credibility and trust are not important or that the concept should be disregarded entirely.
-The American public is, after all, the client of the American military and ultimately, it is their opinion that counts.I am suggesting, however, that for the reasons laid out in this paper, credibility and trust are influenced by the mechanisms through which polarization works, to include the loss of politically-neutral space, increased negative feelings toward those who hold differing opinions, and most sinister of all, alterations in one's concept of loyalty.
-For these reasons, I am simply offering the notion that \emph{credibility can be deceiving} in polarized times.
-Now, more than ever before, military leaders will need to rely on individual moral courage over and above any other trait in order to perform their duties as military professionals.\\
-\singlespace
-\pagebreak
-
-\printbibliography
-
-\end{document}
+Of note, one observation in this cluster entitled, "Fists Raised at West Point" and published in the Washington Post is actually a compendium of two anonymous letters, each one taking an opposite stance on the appropriateness of an incident in which a photograph at West Point's 2016 graduation ceremony captured a graduating female cadet raising a clenched a fist, an action which some saw as an inappropriate political gesture related to the growing Black Lives Matter movement [@unknown_fists_2016].

### 2020-12-05 code/Tahkv1.Rmd
commit

@@
 A second and releated reason is because unlike other forms of engaging with the press, such as giving interviews on television, crafting an op-ed requires the author to deliberately prepare and submit a viewpoint for publication.
-In major newspapers, this is often a competitive process.
-It is not uncommon to see opinion pieces in prominent newspapers published by heads of state, members of congress, governmental figures, and leaders of industry and academia.
 A deliberate and carefully-crafted message further underscores the degree to which the author believes the message he or she is sending.
@@
 
+#Determinants of Military Role Beliefs
+
+Theoretically, there are at least two major determinants of military role beliefs, or the conceptions held by members of the military regarding their normative role in politics [@fitch_armed_1998].
+In the most general terms, the military's involvement in politics requires both the _motive_ and the _opportunity_ to do so [@finer_man_1962].
+I argue that rising political polarization serves as the _motive_ for the military's intervention in politics, while high levels of prestige, or credibility of the military institution, provide the _opportunity_ for doing so.
+It is important to elaborate why each of these offers the military in particular the ability to intervene in politics.^[For a richer elaboration on the mechanisms afforded by political polarization and institutional levels of prestige for intervention in politics, please see the author's theory chapter in his dissertation.
+The proceeding paragraphs here are intended to summarize these mechanisms.]
+
+With respect to the first determinant, political polarization, the literature in American politics enables the development of two plausible relationships and corresponding mechanisms that connect polarization to  military role beliefs.
+These relationships and mechanisms further stem from a particular conception or definition of what polarization is and how it workd.
+One possible conception of polarization is that it marks the ideological widening between political elites, to include members of Congress or the leaders of political parties [@hare_polarization_2014].
+According to this view of polarization, one possible mechanism by which the military intervenes in politics is by adopting a belief that elites from both dominant political parties have failed to advance and implement an effective governing agenda, and thus, the military steps in to do what political elites are not doing.
+This particular dynamic has been noted, in fact, by scholars in the comparative politics subfield describing the actions of militaries in other parts of the world [@fitch_armed_1998].
+
+Yet a second, and I believe the more likely, relationship between political polarization and military role beliefs, particularly in the contemporary United States, stems from a conception of polarization as the strength of feelings that one has concerning a number of ideological issues.
+This view of polarization is more consistent with a concept scholars have labeled "affective polarization", or "negative partisanship" [@abramowitz_great_2018].
+If the variables of polarization and military role beliefs are linked in this way, the military does not intervene as a monolithic block due to a frustration with the political class's inability writ large to effectively govern.
+Rather, certain military elites, who share beliefs with a polarized electorate on any number of ideological issues, such as those regarding race, gender, and immigration [@abramowitz_great_2018], will selectively intervene in order to advance their personal beliefs regarding these topics.
+In other words, this mechanism posits that certain military elites feel as strongly as average citizens about certain polarizing issues, to the point that they will break from historical apolitical or non-partisan military traditions and norms in order to advance their beliefs.
+Thus, when polarization is high, there is a _motive_ among certain members of the the military, to include retired military elites, to deviate from their historical non-partisan and apolitical traditions.
+In short, increasing political polarization may make it increasingly difficult for members of the military to remain politically neutral as partisan conflicts subsume a greater number of issues, and as strong feelings of dislike or disdain surface against those who hold on opposing view.
+
+The second determinant of military role beliefs is the level of prestige, or credibility, that the public places in the military institution.
+When these levels are high, I contend that members of the military possess the _opportunity_ to engage with the public.
+This includes engaging in behaviors such as making public arguments in the press, publishing opinion pieces, or more generally, sharing one's views.
+When society overall esteems the military at low levels, the expectation is that there would be fewer instances of the military publicly sharing its views, because there is less desire from the public (and thus, less desire from press outlets) to hear from members of the military community.
+There may also be, during periods when the military commands low levels of prestige and credibility, less inclination among members of the military to make public pronouncements.
+When the opportunities to engage with the public are more prevalent as a result of high levels of prestige, certain members of the military, I contend, are more likely to feel empowered to take on increasingly active roles, particularly in helping to weigh in or solve problems.
+Thus, we would expect to see a greater level of overall engagement with the press, and therefore, a greater number of overall opinion pieces authored by retired military officers, when military credibility is high.
+
 ## Hypotheses
 
-The hypotheses I wish to examine, then, are as follows:
+Given the expectations derived from theories regarding military role beliefs, the overall purpose of opinion-editorial publications, and retired military officers as members of the military community, five distinct hypotheses are derived.
+These include:
 
-_H1: When political polarization is high, the military (or certain military figures) will increasingly criticize civilian leadership (indicating a drop in the principle of civilian control)._
+\singlespace
+**H1: When political polarization is high, the military (or certain military figures) will increasingly criticize civilian leadership (indicating a drop in the principle of civilian control).**
 
-_H2: When political polarization is low, the military (or certain military figures) will publicly weigh in on traditional national security debates (and not on domestic political concerns or issues outside of relevant military expertise)._
+\doublespace
 
-_H3: During periods of high prestige and credibility, the frequency of military involvement in quasi-political activities (such as writing op-eds) will increase relative to periods of low prestige and credibility._
+\singlespace
+**H2: When political polarization is low, the military (or certain military figures) will publicly weigh in on traditional national security debates (and not on domestic political concerns or issues outside of relevant military expertise).**
 
-_H4: When both political polarization and high prestige are present, the military will be increasingly vocal and at higher frequency about certain domestic ideological or partisan causes than when these factors are not present._
+\doublespace
 
-_H5: During periods of active ground combat, military elites will likely comment on the war, but with less frequency on other domestic issues._
+\singlespace
+**H3: During periods of high prestige and credibility, the frequency of military involvement in quasi-political activities (such as writing op-eds) will increase relative to periods of low prestige and credibility.**
+
+\doublespace
+
+\singlespace
+**H4: When both political polarization and high prestige are present, the military will be increasingly vocal about certain domestic ideological or partisan causes than when these factors are not present.**
+
+\doublespace
+
+\singlespace
+**H5: During periods of active ground combat, military elites will likely comment on the war, but with less frequency on other domestic issues.**
+
+\doublespace
 
@@
 
-The methodological approach used in this analysis consists of two steps.
-The first is to analyze the topics which retired military officers have been writing about, and the second is to determine, regardless of the topic addressed in a publication, whether the op-ed has violated norms of non-partisanship or civilian control of the military.
-The data consists of opinion-editorial and letters to the editor publications authored by retired military officers and published in major US newspapers from 1979-2020.
-These newspapers consist of _The Wall Street Journal_ (WSJ), _The New York Times_ (NYT), _The Washington Post_ (WaPo), _The Los Angeles Times_, and _The USA Today_.
+The methodological approach used in this analysis consists of two broad steps.
+The first is to analyze the topics which retired military officers have been writing about, and the second is to determine, regardless of the topic addressed in a publication, whether an op-ed has violated the norms of non-partisanship or civilian control of the military.
+The critical question is whether there are patterns in the data that can be detected, and whether these patterns can be explained.
 
-I include both opinion-editorials and letters to the editor because both types of publications display a viewpoint that required a deliberate effort by the officer to craft and submit to the particular newspaper.
-I do not include published interviews with retired military officers, nor reprinted editorials that were first published before 1979, such as those reprinted for the benefit of readers sometime after 1979.
-To assemble the data, I use a series of key-word searches, such as "retired military officer," "retired from the Army, Navy, Marine Corps, Air Force," "former Chairman of the Joint Chiefs of Staff," "General," "Admiral," and so forth.
-I conduct these key word searches in several different databases, including Factiva, Lexus-Uni, and ProQuest Historical Newspapers.
-^[For exact replication of the database search parameters, please see the Appendix.] I include in my search publications that were made available both in print and online format, though I did not double-count: if the same op-ed was published both in print and online and published by the same author, I count the article only one time.
+The data consists of 390 opinion-editorial publications and letters to the editor authored by retired military officers and published in major US newspapers from 1979-2020.
+I include both opinion-editorials and letters to the editor because both display a viewpoint that required the author to craft a deliberate message, and to submit the piece for consideration to be published by an editorial board at a particular newspaper.
+Because I am interested in messages that required the officer to craft a deliberate message, I do not include published interviews with retired military officers, nor any reprinted speeches or remarks that a retired military officer gave at another venue (while searching for the data, there were a few instances in which newspapers published remarks of officers made at events).
+
+The five newspapers that I examine in depth include the _The Wall Street Journal_ (WSJ), _The New York Times_ (NYT), _The Washington Post_ (WaPo), _The Los Angeles Times_, and _The USA Today_.
+With the exception of _The USA Today_, all of the newspapers have long and established histories (the _USA Today_ began in 1982 [@noauthor_about_nodate]).
+Moreover, publishing an opinion piece or letter in these newspapers is often a competitive process.
+It is not uncommon to see opinion pieces in these prominent newspapers published by heads of state, members of congress, governmental figures, and leaders of industry and academia.
+Therefore, examining these particular five sources - as opposed to smaller, less known newspapers - helps select the most prominent viewpoints that retired military officers greatly wanted to convey to the public.
+Finally, these newspapers, particularly _The Wall Street Journal_ (WSJ), _The New York Times_ (NYT), and _The Washington Post_ (WaPo) are likely the most circulated among the defense and national security communities, as they are located in New York City and Washington D.C., where think tanks, international organizations, and the Pentagon are based.
 
@@
 The primary advantage is that these sources offer longevity across a time period of nearly forty-two years; in doing so, they provide the researcher with an opportunity to detect patterns across several decades.
-The primary disadvantage is that by limiting myself to only the five newspapers I examine, I am not including a large number of modern social media and blogging sites, contemporary webpages, and other digital offerings that showcase the opinions of retired military officers.
+The primary disadvantage is that by limiting myself to only the five newspapers I examine, I am not including a large number of modern social media and blogging sites, contemporary webpages, and other digital offerings that likewise showcase a large number of opinions of retired military officers.
 
-Although social media has taken over as the primary means in which a majority of Americans receive their news today [@martin_how_2018], the methodological approach adopted in this analysis allows for an examination of the same sources over time.
-More importantly, even after accounting for the advent of online editions of newspapers in the late 1990s and early 2000s, the overall circulation of newspapers (print and online editions combined) is still down today compared to earlier years.
+Although this is a limitation in terms of selection, the methodological approach adopted in this analysis allows for an examination of the same newspaper sources over time.
+This is important, as it affords the researcher the opportunity to detect potential within-source variation.
+However, it should be noted that the overall influence and impact of newspapers is likely less than it was several decades ago.
+Social media has taken over as the primary means in which a majority of Americans receive their news today [@martin_how_2018], and even after accounting for the advent of online editions of newspapers in the late 1990s and early 2000s, the overall circulation of newspapers (print and online editions combined) is still down today compared to earlier years.
 As a means of rough comparison, estimated total daily newspaper circulation in the US in 2000 was 62 million for weekday and 54 million for weekend editions, respectively, while in 2018, the same figures were 29 million and 31 million [@barthel_leading_2019].
+Still, the five newspapers examined are reasonably well known, and while it can't be proven in this paper, most individuals would consider it an accomplishment to author an opinion piece published in any one of these five examined sources.
+
+Actually obtaining the data is a time consuming process.
+I use a series of key-word searches, such as "retired military officer," "retired from the Army, Navy, Marine Corps, Air Force," "former Chairman of the Joint Chiefs of Staff," "General," "Admiral," and so forth in several different databases, including Factiva, Lexus-Uni, and ProQuest Historical Newspapers.
+^[For exact replication of the database search parameters, please see the Appendix.] Each of these databases, I learn, classifies historical newspaper articles differently: some databases include editorials as a separate source type to search for (making the acquisition of data easier), while others do not, requiring a search of all articles that contain the key words (which therefore results in a significantly larger pool of articles to search through and find the op-eds and letters I am looking for).
+I include in my search publications that were made available both in print and online format, though I did not double-count: if the same op-ed was published both in print and online and published by the same author, I count the article only one time.
+^[To my knowledge, there is no data set that yet includes the data assembled for this research project.
+Therefore, I cannot conclusively say that I actually found all of the opinion pieces and letters authored by retired military officers and published in these sources within the time periods examined in this paper.
+It is possible, for instance, that I left out an important search parameter that did not trigger a certain number of op-eds that were actually authored by retired military officers.
+While it is difficult to put a precise degree of confidence in the method employed to obtain the data, my personal reaction is that the method employed caught greater than 95% of the actual observations.
+I offer this as a "ballpark" assessment based off several searches of individual newspapers in a given year in which I compared the results I obtained using the search parameters described against a method in which I literally looked through every op-ed published in a given calendar year.]
 
@@
 
-```{r table2, fig.align="center", include=TRUE, echo=FALSE}
+```{r table2, fig.align="center", include=TRUE, echo=FALSE, warning=FALSE}
 countbytype %>%
@@
 ```{r figure3, include=TRUE, echo=FALSE, warning=FALSE, error=FALSE}
-ggplot(data=adjdataset, mapping=aes(x=PubYr, y=dom_pol_prop)) + geom_point() + geom_smooth() + labs(title="Figure 3: Proportion of Annual Publications authored by Retired Military Officers\n that Address Domestic Political Matters, 1979-2020", x="Year", y="Proportion", color=NULL)+ geom_vline(xintercept = 1981, size=.2, colour="tomato") +
+ggplot(data=adjdataset, mapping=aes(x=PubYr, y=dom_pol_prop)) + geom_point() + geom_smooth() + labs(title="Figure 3: Proportion of Annual Publications authored by Retired\n Military Officers that Address Domestic Political Matters, 1979-2020", x="Year", y="Proportion", color=NULL)+ geom_vline(xintercept = 1981, size=.2, colour="tomato") +
   geom_vline(xintercept=1989, size=.2, colour="tomato") + 
@@
 \singlespace
-# Determinants of Changing Military Role Beliefs
+## Explaining A More Assertive Retired Military Officer Corps 
 \doublespace
 
-The story told in Figures 3 and 4, in short, is that retired military officers are addressing domestic political topics today more frequently and at a greater rate relative to other topics than they did in previous decades, and that they are criticizing civilian leaders, taking partisan stances, and calling for legislative action at greater rates than they previously did.
-Thus, the body of retired military officers is more assertive, at least in how it engages with the five major newspapers examined in the data, than it was in previous time periods.
-Having established these trends, the next task is to explain why they have formed.
+Collectively, Figures 3 and 4 tell an interesting story: retired military officers are addressing domestic political topics more frequently and at a greater rate relative to other topics than they did in previous decades, and they are criticizing civilian leaders, taking partisan stances, and calling for legislative action at greater rates than they previously did.
+Said differently, retired military officers are exuding boldness through the opinion pieces they publish in ways that differ from previous decades.
+Having established that these trends exist, the next task is to attempt to explain their development.Fortunately, measurements for political polarization and credibility of the military exist, offering the researcher the ability conduct a quantitative analysis of the data.
 
-Theoretically, there are at least two major drivers of military role beliefs, including the level of prestige held by the military in American society, and the degree of political polarization that exists.^[In the theory chapter of my dissertation, I derive why these are the main drivers of military role beliefs.]
+One measurement of political polarization are DW-NOMINATE (dynamic-weight, nominal three step estimation) scores, which were originally developed by the political scientists Keith Poole and Howard Rosenthal in the early 1980s [@jeffrey_b_lewis_voteview_2020].
+These scores use roll-call voting data to assign member of Congress a score on the liberal-conservative dimension, ranging from -1 (extremely liberal) to 1 (extremely conservative) [@jeffrey_b_lewis_voteview_2020].^[The data is frequently relied upon by academics, journalists, and the media to understand how a variety of political bodies, to include Congress and political parties, have changed over time.
+Over the time period examined in this paper, polarization in each chamber of Congress, the House of Representatives and the Senate, has increased [@jeffrey_b_lewis_voteview_2020].] For instance, the distance in party mean DW-NOMINATE scores in the late 1970s was just less than .6, while the distance today for both chambers is above .8 [@jeffrey_b_lewis_voteview_2020].
 
-To understand how measurements of political polarization have changed over time, I rely on the DW-NOMINATE (dynamic-weight, nominal three step estimation) scores that were originally developed by the political scientists Keith Poole and Howard Rosenthal in the early 1980s [@jeffrey_b_lewis_voteview_2020].
-The DW-NOMINATE data are one measure that uses roll-call voting data for each member of congress to assign a score to the member of congress on a liberal-conservative dimension, ranging from -1 (extreme liberal) to 1 (extreme conservative) [@jeffrey_b_lewis_voteview_2020].
-The data is frequently relied upon by academics, journalists, and the media to understand how a variety of political bodies, to include Congress and political parties, have changed over time.
-Over the time period examined in this paper, polarization in each chamber of Congress, the House of Representatives and the Senate, has increased [@jeffrey_b_lewis_voteview_2020].
-The distance in party mean DW-NOMINATE scores in the late 1970s was just less than .6, while the distance today for both chambers is above .8 [@jeffrey_b_lewis_voteview_2020].
+A second measurement of political polarization is a measurement of affective polarization, which was first conducted by Iyengar et al using time series data from the American National Election Study [@iyengar_affect_2012; @iyengar_origins_2019].
+Using a series of "feeling thermometer" questions that guage a person's feelings about those on the opposite side of one's political spectrum, this measurement roughly captures how much "dislike" exists among the electorate for each other [@iyengar_strengthening_2018].
+This measurement does not reflect an encouraging story: while the level at which citizens like their own party has remained remarkably stable for the past 40 years, the level of affection for those of the opposite political party precipitously dropped around the year 2000, a trend that has continued [@iyengar_strengthening_2018].
+I include both measures of polarization - among members of Congress and among the electorate - in the analysis to follow.
 
-Changes in the institutional credibility of the American military institution can be seen by examining public opinion data.
-Pollsters survey the American public annually to gauge their trust in a variety of institutions.
-Using Gallup opinion polling, one sees that since the late 1970s, the credibility of the American military as an institution has gradually increased [@gallup_polls_confidence_2020].
+Using Gallup public opinion data, I include a measurement of the credibility of the American military as an institution.
+These surveys have asked Americans to rate the level of trust that they place in a variety of American institutions, such as the Presidency, Congress, Churches, and the Police.
+Since the late 1970s, the credibility of the American military as an institution has gradually increased [@gallup_polls_confidence_2020].
 In 1979, 54% of those surveyed expressed either a great deal or quite a lot of confidence in the military, while in 2020, 72% expressed the same [@gallup_polls_confidence_2020].
-These trends in political polarization and confidence in the military are displayed in Figure Four, below.
-The top and middle graph show changes in means between Republicans and Democrats in the House of Representatives (top graph) and Senate (middle graph) since 1979, while the bottom graph shows changes in the percentage of surveyed Americans who reported either a great deal of or a lot of confidence in the military institution over the same time period.
+Trends in political polarization (using DW-Nominate Scores and a measure of Affective Polarization) and confidence in the military are displayed in Figure Four, below.
 
-```{r figure 4, include=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
+```{r figure 5, include=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
 #adjust the data to account for polarization by creating a new dataset called polaradjdataset.  
 #In this dataset, create variables that capture ideological distance in each the House and Senate along each of the two dimensions according to Poole and Rosenthal's DW-NOMINATE data
+# and a variable for affective polarization
 
-# scale the data by multiplying by 100
+# scale the data for congressional polarization by multiplying by 100
 
@@
   House_distance_dim2=abs(House_Dem_mean_nom2-House_Rep_mean_nom2)*100,
-  Sen_distance_dim2=abs(Sen_Dem_mean_nom2-Sen_Rep_mean_nom2)*100
+  Sen_distance_dim2=abs(Sen_Dem_mean_nom2-Sen_Rep_mean_nom2)*100,
+  aff_polar=dem_diff_aff+`rep-diff_aff`
 ) %>% view ()
@@
 #note that the id.var will serve as the common axis for these 
-polar_and_cred <- melt(subset(polaradjdataset, select=c(PubYr, House_distance_dim1, Sen_distance_dim1, Inst_Cred_Gallup)), id.var="PubYr")
+polar_and_cred <- melt(subset(polaradjdataset, select=c(PubYr, House_distance_dim1, aff_polar_alt, Inst_Cred_Gallup)), id.var="PubYr")
 
@@
   labs(
-    title = "Figure Four: Distance Between Party Means in the House,\n the US Senate,\n and American Confidence in the Military, 1979-2020"
+    title = "Figure 5: Polarization Levels in the US House of Representatives,\n Affective Polarization, and American Confidence in the Military, 1979-2020"
   )
@@
 
-While it is possible that the trend in increasing institutional credibility of the US military over the past four decades may be at least partially responsible for the increasing trend in the number of opinion pieces published by retired military officers, a trend that the data supported earlier as shown in Figure 2, rising institutional credibility alone does not adequately explain the trends shown in Figure 3: that there has been a sharp increase over the past four decades in retired military officers penning pieces that address non-traditional military topics that lay outside the bounds of relevant military expertise, or pieces that express partisan sentiment.
-The theory laid out earlier in this paper predicted that as political polarization increases, three mechanisms are at work which have strong impacts on the military professional: it becomes increasingly difficult to remain politically neutral as partisan conflicts subsume a greater number of issues; as this occurs, strong feelings of dislike or disdain surface against those who hold on opposing view; and alterations in one's sense of loyalty arise, particularly as issues that individuals hold to most dearly raise the salience of one's identity.
+Because there are only three observations written exclusively by retired female officers, and because there is no compelling logic for why the service of the author would increase the likelihood of violating norms when publishing an opinion piece, I do not control for the sex or service affiliation of the author.
+Controls are included, however, for the author's rank and the source in which a publication appears.
 
-To test if changes in political polarization adequately explain the changes in the qualitative nature (increasing trends of potential violations of military professionalism) of opinion pieces authored by retired military officers, I conduct a logit regression of the variable _questionable_professionalism_ while holding institutional trust in the military constant at its mean (71 percent) and varying the value of the mean ideological distance between both major political parties in the House between 1979-2020.
-Conducting such a regression allows one to examine, given the data assembled, the predictive likelihood of publishing an opinion piece that violates military professionalism given varying levels of political polarization and constant institutional credibility in the military.
-These results are displayed in Figure 5, below.
+The dependent variable is an indicator of whether the opinion piece either called for legislative action, blatantly criticized a political leader, or clearly took a partisan stance on an issue.
+As some may argue that calling for legislative action in itself does not violate norms of civilian control of the military or the non-partisanship of the military institution, the dependent variable in an alternate model is an indicator for whether publication criticized a political leader or took a clearly partisan stance.
+The estimated model therefore takes the form:
 
-```{r logit analysis, include=FALSE, echo=FALSE}
+$$y = \beta_1 credibility + \beta_2polarization + \beta3source +\epsilon $$
+The results of a logistic regression of this model are displayed in Table 3, below.
+
+```{r logit analysis, include=FALSE, echo=FALSE, warning=FALSE}
 
@@
 
-mod1 <- glm(leg_crit_part ~ Inst_Cred_Gallup + House_distance_dim1 + AuthSex + AuthRank + Topic, 
+mod1 <- glm(leg_crit_part ~ Inst_Cred_Gallup + House_distance_dim1 + Source + AuthRank + Topic, 
   family=binomial(link="logit"), 
@@
 # model 2
-mod2 <- glm(crit_part ~ Inst_Cred_Gallup + House_distance_dim1 + AuthSex + AuthRank + Topic, 
+mod2 <- glm(crit_part ~ Inst_Cred_Gallup + House_distance_dim1 + Source + AuthRank + Topic, 
     family=binomial(link="logit"),
@@
 
-mod3 <- glm(leg_crit_part ~ Inst_Cred_Gallup + House_distance_dim1 + AuthSex + AuthRank,
+mod3 <- glm(leg_crit_part ~ Inst_Cred_Gallup + House_distance_dim1 + AuthRank,
     family=binomial(link="logit"),
@@
 
-mod4 <- glm(crit_part ~ Inst_Cred_Gallup + House_distance_dim1 + AuthSex + AuthRank,
+mod4 <- glm(crit_part ~ Inst_Cred_Gallup + House_distance_dim1 + AuthRank,
     family=binomial(link="logit"),
@@
 #model 5 - now check for robustness using aff_polar as the polarization variable
-mod5 <- glm(leg_crit_part ~ Inst_Cred_Gallup + aff_polar + AuthSex + AuthRank,
+mod5 <- glm(leg_crit_part ~ Inst_Cred_Gallup + aff_polar_alt + AuthRank,
     family=binomial(link="logit"),
@@
 
-mod6 <- glm(crit_part ~ Inst_Cred_Gallup + aff_polar + AuthSex + AuthRank,
+mod6 <- glm(crit_part ~ Inst_Cred_Gallup + aff_polar_alt + AuthRank,
     family=binomial(link="logit"),
@@
 \singlespace
-```{r regression table, include=TRUE, echo=FALSE, results="asis"}
+```{r regression table, include=TRUE, echo=FALSE, results="asis", warning=FALSE}
 stargazer(mod3, mod4, mod5, mod6,
@@
           column.labels = c('Model 3','Model 4','Model 5', 'Model 6'),
-          covariate.labels = c('Institutional Credibility', 'Congressional Polarization', 'Affective Polarization', 'Male', 'Mixed Authorship', 'Unknown Authorship', 'Rank'),
+          covariate.labels = c('Military Credibility', 'Congressional Polarization', 'Affective Polarization', 'Rank'),
           align=TRUE)
@@
 
-The graph clearly shows a non-linear relationship between increasing political polarization and the likelihood that an opinion piece authored by a retired military officer will potentially violate aspects of military professionalism.
+The results shown in Table 3 indicate that both measurements of polarization - Congressional and affective - are statistically significant at the .01 level.
+The substantive interpretation of the coefficients for polarization, however, are a bit tricky.
+A coefficient of .083 for Congressional Polarization in Model indicates that a 1-unit increase in the average distance between political parties is associated with a 8.7 percent increase in the likelihood of a publication authored by a retired military officer calling for legislative action, blatantly criticizing a civilian official, or expressing a clearly partisan stance.
+As Figure 5 shows, though, the differences in polarization levels vary with each Congress.
+For instance, within the House of Representatives in the 112th Congress, which began in January 2011, polarization levels were nearly six points higher than the level that existed in the 111th Congress.
+The estimated effects of such a drastic change in the polarization level in Congress would be associated with more than a 50 percent increase in the predicted likelihood of a publication retired by a military officer calling for legislative action, criticizing a civilian leader, or taking a clear partisan stance.
+
+Similary, the coefficient in Model 6 for affective polarization of .033 indicates that a 1-unit increase in the level of affective polarization is associated with a 3.4 percent increase in the likelihood of a publication authored by a retired military officer blatantly criticizing a civilian leader or taking a clear partisan stance on a given issue.
+And as Figure 5 also shows, changes in the level of affective polarization over the past four decades has varied.
+However, in between elections held every four years, it is not uncommon to see changes of 5-7 points, which would indicate a corresponding 17 - 23.8 percent increase in the likelihood of a publication authored by a retired military officer criticizing a civilian official or taking a clear partisan stance.
+
+Interestingly, the coefficient on military credibility is statistically significant for Model 1 only.
+Across other model specifications, the coefficient remains negative, indicating than an increase in the credibility of the military would be associated with a reduction in the likelihood of a publication authored by a retired military officer violating norms of civilian control and non-partisanship.
+But because the coefficient is not statistically signficant in Models 4-6, we cannot draw conclusions that such a relationship exists in the data.
+
+Finally, the coefficient for rank of the author is statistically significant in Models 3 and 5, when the dependent variable includes calling for legislative action.
+In Models 2 and 4, when the dependent variable is whether a publication criticizes a civilian official or takes an expressly partisan stance, the coefficient is not statistically significant.
+This suggests that the act of calling upon Congress or the President to take specific action is clearly more associated with the rank of an author.
+This makes sense, as those authors who possess a greater rank likely have had experiences that have worked alongside or dealt with legislative policymakers, and because these authors feel more comfortable in calling for legislative action than those of a lower rank who probably lack such professional experience.
+
+## Describe the interpretation of the author, the press being willing to publish, etc. Pull from the paras below...
+
+## Describe whether you found support for your H's or not
+
+## Mixed Models and Source Random / Fixed Effects
+
+a  polarization change However, is statistically significant across a graph clearly shows a non-linear relationship between increasing political polarization and the likelihood that an opinion piece authored by a retired military officer will potentially violate aspects of military professionalism.
 While the confidence intervals at each point estimate are relatively large, the graph demonstrates that the predicted probability of violating military professionalism is different between relative extremes along the x axis.
@@
 Of note, one observation in this cluster entitled, "Fists Raised at West Point" and published in the Washington Post is actually a compendium of two anonymous letters, each one taking an opposite stance on the appropriateness of an incident in which a photograph at West Point's 2016 graduation ceremony captured a graduating female cadet raising a clenched a fist, an action which some saw as an inappropriate political gesture related to the growing Black Lives Matter movement [@unknown_fists_2016].
+
+\newpage

### 2020-12-18 code/Erickson-dissprospect.v3.Rmd
commit

--- 2020-12-04 code/prospectusv2.Rmd
@@
 ---
-title: "Servants, Syndicates, and Saviors: The Rise of Domestic Political Interventionism by Senior Military Leaders (Prospectus)"
+title: "Servants, Syndicates, and Saviors: How Polarization and Prestige Impact the Political Behavior of Military Elites  (Prospectus)"
 author: "Peter M. Erickson"
@@
 
-  _This dissertation explores changes in how the contemporary military institution intervenes in domestic politics.
-I advance a theory which argues that normative beliefs of military officers regarding political activism is a result of two interacting factors: the level of domestic political polarization within a society, and the degree of prestige and credibility the military holds.
-To support my argument, I trace changes in levels and types of distinct forms of political activism by military officers, including authoring opinion-editorials in major newspapers, as domestic polarization and military institutional credibility rises.
-I ultimately argue that the increasing political activism of military leaders is not merely a response to the hyper-partisanship present in the current political atmosphere, but rather the logical outworking of trends involving the prestige of the military institution as well as political polarization that have been building for decades.
-I conclude by describing the implications of changing beliefs among military leaders regarding the role the military should have in domestic politics.
-This dissertation contributes to the fields of international relations and civil-military relations by advancing an empirical and historical analysis of a growing phenomenon: military leaders intervening in domestic American politics.
-_
+  _This dissertation explores variation in the political behavior of the US military.
+The main theory that I advance is that across time, the political behavior of the US military is mainly impacted by two distinct factors: the level of domestic political polarization within a society, and the degree of prestige and credibility the military holds.
+These factors greatly influence the normative beliefs that militaries hold regarding their role in politics, and place at risk both the military's non-partisan ethic and the norm of civilian control of the military.
+As a result, the presence of these factors in domestic society impact the type, frequency, and other characteristics of political behaviors the military engages in.
+To support my theoretical argument, I conduct a mixed-methods empirical strategy that includes quantitative analysis and case study investigation.
+This dissertation advances scholarly understanding of how domestic politics impacts civil-military relations and international security._
 
@@
 
-Across the world and in almost nearly every country, discussions regarding the participation of the military in politics are nothing new.
-Indeed, few topics evoke such strong opinions and emotions than whether and to what degree the military, both as a corporate body and as individual military officers, should be political vice remain separate from certain aspects of and activities associated with politics.
-These normative questions rightly form a significant portion of the study of civil-military relations, and intellectuals, practitioners, and pundits have frequently weighed in on these topics over the years.
+Across the world and in almost nearly every country, scholars, civilians, diplomats, and military leaders regularly discuss why and how militaries intervene in politics.
+In fact, it might be difficult to find a subject area other than civil-military relations where the practical behavior of soldiers and civilians, occurring before our very eyes, is so closely scrutinized and compared to normative conceptions of how we think these interactions between soldiers and statesmen should occur.
+For this reason, especially in the United States context, the political activity of the Armed Forces is a subject that has undergone intense scrutiny in recent years as polarization has, upon several occasions, thrust the military into the political spotlight.
 
-Thankfully, within the United States, unlike many other countries across the world and throughout history, there has never been a coup, nor a serious attempt made by the military to dislodge political control from their duly-constituted elected leaders and to hold it themselves.
-However, as almost every civil-military relations scholar points out, the absence of coups does not mean that the relationship between civilian and military leaders is healthy, nor absent of conflict altogether.
-Indeed, a cursory examination of the relationships between Lincoln and McClellan, Truman and MacArthur, and Obama and McChrystal reveals that disagreement between those who serve at the highest levels of civilian government and military organizations is neither rare nor understudied.
+From a now-retired Army Lieutenant Colonel providing Congressional testimony during President Donald Trump's impeachment proceedings, to a Navy Captain's removal for leaking a letter to the press that described his frustration with the military chain of command's response to the onset of the COVID-19 pandemic, to a remarkable panoply of retired military elites blasting the President's threat to use active duty forces to dispel protesters in cities across America during the summer of 2020, the events of the past few years have ignited an impassioned debate regarding the military's role in politics.
+Moreover, the near-ubiquitous use of militaries to respond to the COVID-19 pandemic has similarly awakened a desire to better understand the role of militaries in society, what they are used for, and how they should and actually do engage with civilian leaders.
+These healthy scholarly pursuits and debates, as well as the instances that justify their existence, will likely not end soon.
 
-What has changed, however, is the degree to which a very specific subset of the American military institution - senior military leaders - have participated in the domestic political process, especially post retirement.
-Indeed, whereas General Omar Bradley, one of the military's five star generals and a hero of World War Two, famously remarked in 1959, "I am convinced that the best service a retired general can perform is to turn in his tongue along with his suit and mothball his opinions," [@times_bradley_1959], retired generals and admirals today appear frequently in the media, endorse Presidential candidates at higher frequency [@griffiths_not_2019], and frequently serve in government after their retirement from military service [@brooks_dont_2016].
-To provide only one such example, following protests across the United States in the summer of 2020 in which President Donald Trump nearly ordered the intervention of active duty military forces into American cities, former Chairman of the Joint Chiefs of Staff, Admiral (Ret) Michael Mullen, was one of at least ten four star retired officers who publicly aired concerns through a variety of mediums about the President's potential use of active-duty military force.
-^[Other officers included General (Ret) James Mattis, General (Ret) Mark Kelly, Admiral (Ret) William McRaven, General (Ret) Martin Dempsey, General (Ret) Colin Powell, General (Ret) Michael Hayden, General (Ret) David Petraeus, General (Ret) John Allen, and General (Ret) Vincent Brooks.].
-In fact, during an interview in the days which followed the incident, when asked by an NPR reporter why he decided to publicly share his thoughts on the situation, Mullen stated, "we were getting to a point where the country was in jeopardy, which is why I spoke out" [@isaacson_ret_2020; for further analysis, see also @brooks_let_2020].
+And while numerous scholarly works describe important facets of the political behavior of the military [@huntington_soldier_1957; @finer_man_1962; @fitch_armed_1998; @brooks_militaries_2009; @griffiths_not_2019; @brooks_paradoxes_2020], few if any fully describe how the presence of high degrees of political polarization and military prestige impact the manner in which the US military engages in distinct forms of political behavior, and importantly, how these forms of political behavior vary over time.
+This dissertation seeks to fill this important gap.
+In so doing, the intended audience groups of this dissertation are scholars, civilian officials, military officers, and (I hope) citizens who care about the connection between American society and its military.
 
-Mullen's words are incredible from a number of perspectives, and raise important normative questions regarding the military's role in a democratic society [@brooks_let_2020] and the conditions under which the military can and should express dissent to its civilian bosses [@cohn_what_2019].
-And while a small but dedicated number of civil-military relations scholars do examine these and related issues quite frequently in today's charged political environment, few, if any, have attempted to systematically examine and explain over time the conditions under which senior military leaders have broken and will break normative traditions and engage in domestic politics.
-Such an effort would better illumine the understanding of scholars, as well as of the military and society more broadly, concerning why a powerful group of what were once non-political actors have, at least in some cases, suddenly become one.
+The central argument made in this dissertation is that the conditions of high political polarization and a high degree of military prestige are present within a society, the result is an increasingly politically active military _if_ it chooses to be.
+This choice will largely depend on the motives and decisions of a small group of powerful military elites who, for better or for worse, are in charge of guiding the military through turbulent domestic waters.
+This dissertation shows that although the conditions of high polarization and military prestige have been prevalent within American society for the better part of the post-9/11 era, and especially in the past several years, this is not the first time these conditions have existed in the United States.
+Indeed, military elites have acted both nobly and dishonorably in the presence of these conditions at various time in the nation's past.
+It is important to learn from their mistakes so that they can be avoided again in the future, if possible.
 
-The central argument made in this dissertation is that increasing political activism at the hands of military officers indicates a shift in the beliefs of the military regarding its central role and purpose in society, or "role beliefs" [@fitch_armed_1998].
-This shift, moreover, is the result of two longstanding and interconnected trends: rising political polarization within the domestic political sphere, and the rising prestige and credibility of military institutions afforded by American society.
-The interaction of these trends explains the remarkable increase in political activity by retired senior military leaders, and portends a shift in the perceived role that these senior military leaders fill vis-a-vis civilian authority - a shift away from professional and obedient "servants" and towards that of "syndicates" or even "saviors" of the American state.
-This shift, moreover, holds enormous ramifications regarding the organizational culture of the military, its relationship with government and society, and American national security.
+This prospectus guides the reader in four parts.
+First, the prospectus introduces the reader to the general field of civil-military relations.
+In doing so, it emphasizes the importance of the dependent variable of study: political behavior of the military.
+Second, the prospectus briefly points out where the intended dissertation fits within the scholarly literature.
+Third, the prospectus lays out a theory of polarization and prestige and their impact on political behavior of the military.
+Fourth, the prospectus walks the reader through an empirical strategy for collecting the data required to test the theory laid out.
+In the conclusion, the prospectus addresses some immediate concerns and questions.
 
@@
 
-My proposed dissertation topic features within the sub-field of civil-military relations, a broad academic sub-field that is generally placed within the International Relations and Comparative Politics sub-field (though within the American context, the sub-field of American Politics is entirely relevant and appropriate).
-Civil-military relations involve a wide host of significant and important theoretical questions that involve the pull and push of power between institutions, the delineation of authority between military and civilian leaders, and the relationship between a state's military and society.
+The proposed dissertation topic regarding the political behavior by the military features within civil-military relations, a broad academic sub-field that is studied by scholars of International Relations, Comparative Politics, and American Politics (concerning the US military in particular).
+A broad range of scholarly interests comprises the study of civil-military relations, including coups, the role of domestic institutions, strategic effectiveness, the relationship between the military and society, and a host of other important questions.
 
-At its root, the study of civil-military relations examines what Peter Feaver calls the "problematique," or the question of how to ensure that a military is strong enough to defeat threats to the polity without threatening the polity itself [@feaver_civil-military_1996; @feaver_civil-military_1999].
-Feaver describes this "problematique" as "vexing" and a great "paradox" in that the military must be "strong enough to prevail in war," yet submissive to authority such that it cannot and will not "destroy or prey on the society it is intended to protect" [@feaver_civil-military_1999, p.
+At its root, the study of civil-military relations examines the question of how to ensure that a military is strong enough to defeat threats to the state without threatening the state itself.
+In the terminology of Peter Feaver, one prominent scholar, this question is indeed the "problematique" or central question within the field of all civil-military relations scholarship [@feaver_civil-military_1996; @feaver_civil-military_1999].
+Feaver notes that this "problematique" is "vexing" and, actually a great "paradox" in that the military must be "strong enough to prevail in war," yet submissive to authority such that it cannot and will not "destroy or prey on the society it is intended to protect" [@feaver_civil-military_1999, p.
 214].
-Moreover, the study of civil-military relations often focuses on the relationship between military and government elites, or at what Risa Brooks calls "leadership at the state's apex" [@brooks_risa_integrating_2019, p.
-380].
-Indeed, it is at this level of interaction between government and military leaders that vital decisions are made, including decisions whether to go to war, how to achieve and implement effective strategy, how to prepare for the next conflict, and how large the military and its budget should be.
 
-A number of scholars have offered solutions to the"problematique" raised by Feaver, but none has been more lasting and consequential than the late Samuel Huntington, whose 1957 seminal work, _The Soldier and the State: The Theory and Politics of Civil-Military Relations_, addresses this fundamental problem.
-Indeed, even more than 60 years after he wrote his book, Huntington's work often serves as the point of departure for many scholarly works on topics concerning civil-military relations.
-Although there have been many critiques of Huntington over the years [for example, see @janowitz_professional_1960; @finer_man_1962; @cohen_supreme_2003; @feaver_armed_2003; @brooks_paradoxes_2020], it is absolutely essential to understand Huntington's normative conception of effective civil-military relations, and what lies at the heart of it - a theory of civilian control of the military that maximizes a concept that Huntington called "military professionalism" [@huntington_soldier_1957].
+The focus of study within civil-military relations can occur in several places, but as one might expect, there tend to be patterns depending on the specific area of scholarship.
+Sociologists studying the relationship between the military and society tend to focus on state-military relations, while political scientists often focus on the relationship between the state and militaries.
+Among the focus areas of political scientists is the relationships between prominent state officials and military leaders or elites.
+The civil-military relations scholar Risa Brooks, for instance, has rightly coined this relationship "leadership at the state's apex" [@brooks_risa_integrating_2019, p.
+380], for it is at this level or set of interactions that government leaders and military officials make and implement vital decisions.
+These decisions include whether to go to war, how to implement effective strategy, how to prepare for the next conflict, and how large the military and its budget should be.
 
-In _The Soldier and the State_, Huntington asserts that two distinct forces would always be in tension with one another in a liberal democracy such as the United States.
-The first was a "functional imperative," which Huntington said drove the requirement that a military be strong enough and capable of defeating the military threats it faced, and the second was a "societal imperative," which Huntington claims actually constrains the military in that the military can never become so alienated from society such that the military no longer represents the liberal democracy it defends [@huntington_soldier_1957, 1-3].
-The societal imperative, Huntington claimed, was a further function of two distinct features of the United States.
-The first was an "ideological constant," which referred to America's political ideology of liberalism [@huntington_soldier_1957, 143-162].
-Huntington argued that liberalism as a political ideology stressed "individualism" [@huntington_soldier_1957, 90], and that American society was, since its founding, suspicious of standing military organizations and professional military forces.
-The second feature of the societal imperative, however, was the US Constitution, which Huntington claimed to be "conservative," and written by "men who feared concentrated political power and who provided for the widespread dispersion of that power among numerous governmental units" [@huntington_soldier_1957, 143].
-Huntington claimed that the tension generated by these two imperatives - one functional stemming from external threats, the other societal stemming from the character of American political ideology and structure - needed to be solved.
-Writing in the context of the seemingly long and dangerous Cold War, Huntington's point is that America needs strong military that can win on the battlefield, but cautions that such a military cannot fundamentally alienate the character of American liberal society.
-Both goals - military security and civilian control - had to be obtained.
+A number of scholars have offered solutions, warnings, and caveats to the "problematique" Feaver describes - how the military can protect the state without threatening it - but none has been more lasting and consequential than the late Samuel Huntington, whose 1957 seminal work, _The Soldier and the State: The Theory and Politics of Civil-Military Relations_, often serves as the point of departure for a number of seminal debates within civil-military relations.
+Although there have been many critiques of Huntington over the years [for example, see @janowitz_professional_1960; @finer_man_1962; @cohen_supreme_2003; @feaver_armed_2003; @brooks_paradoxes_2020], it is absolutely essential to understand Huntington's normative conception of effective civil-military relations, and what lies at the heart of it - a theory of civilian control of the military that maximizes "military professionalism" [@huntington_soldier_1957].
 
-Figure 1, below, shows a visual diagram of Huntington's logic in a political science framework [@feaver_civil-military_1996].
-As Peter Feaver points out in a 1996 article, Huntington did not use such language when developing and explaining his concept.
-According to Huntington, the ultimate goal, or dependent variable, of civil-military relations in the American context is achieving both military security and maintaining civilian supremacy over the military.
-Huntington viewed the key independent variable that enables the achievement of these goals while also resolving the tension generated by the opposing functional and societal imperatives as the adoption of a type of civil-military relations framework which he calls, "objective civilian control" of the military [@huntington_soldier_1957, 83].
-In such a model of "objective civilian control," Huntington envisioned civilians granting the military a high degree of professional autonomy, which he believed would simultaneously be "most conducive to the emergence of professional attitudes and behavior among members of the officer corps" as well as the best way to minimize the military's political power by rendering the military "politically sterile and neutral" [@huntington_soldier_1957, 83-84], and thus preserving civilian supremacy.
+## Civilian Control of the Military and the Non-Partisan Military Ethic
 
+In _The Soldier and the State_, Huntington asserts that liberal democracies such as the United States have to contend with, and ultimately balance, the simultaneous demands of military security and political liberalism.
+Huntington argued that in the aftermath of World War Two, American security requirements generated a "functional imperative" to maintain a military strong enough to defeat the  external threats of the Cold War, while the liberal character of American society generated a "societal imperative" such that the military as an institution could never become so alienated from society that it ceased to represent America [@huntington_soldier_1957, 1-3].
+In making his argument, Huntington discusses a number of crucial differences between the military officer corps and the rest of American society.
+He claims that those who serve in the military hold to a distinct "mind" and "ethic" that views conflict as part of a "universal pattern throughout nature," and thus, that "the military view of man is...decidedly pessimistic" [@huntington_soldier_1957, 62-63].
+On the other hand, American political liberalism, Huntington claims, is distinctly different.
+Huntington conceives of American liberal ideology as stressing "individualism," and that liberalism as an ideology "either denies the existence of power, minimizes its importance, or castigates it as inherently evil" [@huntington_soldier_1957, 90].
+It is this juxtaposition between the unique character and ethic of militaries and the attributes of political liberalism in America, claims Huntington, that explains why America had been suspicious of standing military forces since its founding [@huntington_soldier_1957, 143-151].
+
+In essence, Huntington is claiming that changes in the international security environment after World War Two are confronting the United States to adapt, whether it wants to or not, its relationship with its military.
+The United States had to learn, claims Huntington, to live with a sizable professional force capable of inflicting a significant degree of coercive power.
+At the same time, warns Huntington, America will not agree to such a relationship if its unique, liberal political ideology is threatened.
+Any solution to solving this tension must simultaneously render the military strong enough to defeat external threats, but completely subservient to civilian control.
+
+Figure 1, below, displays the solution that Huntington puts forth to solve the tensions created in the strategic situation facing the US after World War Two.^[As Feaver points out in a 1996 article, Huntington did not use modern political science language when developing this concept.
+I constructed this puzzle using Feaver's 1996 recapitulation of Huntington.] The ultimate goal, or dependent variable, of civil-military relations in the American context is achieving both military security and maintaining civilian supremacy over the military.
+Huntington's proposed solution is to adopt a form of civilian control of the military known as "objective civilian control" in which civilian leaders grant the military a high degree of professional autonomy in exchange for the military staying out of partisan politics and instead, focusing on the professional expertise required to achieve American security [@huntington_soldier_1957, 83].
 ```{r figure one, include=FALSE, echo=FALSE}
@@
 ```
-
-Huntington's argument in support of "objective control" of the military stands in contra-distinction, as he explains, to an alternate model of "subjective control" in which civilian leaders limit the power of the military by treating it as any other civilian interest group [@huntington_soldier_1957, 80-83].
-In fact, Huntington admits that the US Constitution actually envisioned and established a form of subjective vice objective scheme of civilian control of the military [@huntington_soldier_1957, 163-168].
-What the framers did not see, nor could they, argues Huntington, was the rise of the military as a profession in response to ever-increasing and dangerous external threats.
-It was because of these pernicious threats, argues Huntington, which the US military had witnessed, confronted, and overcome during the 19th and 20th centuries in World War One, World War Two, and as Huntington writes, the Cold War, as well as the destructive capability of modern weapons and warfare in general, that the military profession had emerged [@huntington_soldier_1957].
-Huntington claims that Cold-War military officers, unlike the corporate body of officers that existed at the nation's founding, constitute a profession and thus, are part of a unique and distinct organization that differs from other occupations: professions practice and develop a special body of knowledge and expertise; professions exercise a  "social responsibility" by providing a vital service to society; and professions have a degree of "corporateness" in that they develop and formalize standards by which members of the profession are certified to practice [@huntington_soldier_1957, 11-18].
+The reason that such an arrangement would work, claimed Huntington, was because of the rise of the military as a profession that had unique corporate attributes.
+Unlike the military that had existed at the founding of America, which had been largely comprised of citizen-soldiers armed with small arms, innovation in wartime weapons and technologies had given rise to a professional body of officers skilled in what Harold Lasswell called the "management of violence" [Lasswell in @huntington_soldier_1957, 11].^[Of note, the topic of when the US military truly became a "profession" is a lively discussion.
+Historians tend to peg the dates at which the individual services embodied themselves as professions differently, ranging from...For more on that, see @need cite...] This was a development, Huntington claims, the American founders simply could not see.
+Because professions had unique corporate attributes, including that of developing a special body of knowledge and expertise, exercising a "social responsibility" by providing a vital service to society (defense), and fostering a degree of "corporateness" in that they develop and formalize standards by which members of the profession are certified to practice [@huntington_soldier_1957, 11-18], allowing the military a high degree of autonomy to exist qua profession would help ensure that it stay out of politics.
+It would be more interested in thriving as a profession than as a political body.
 
@@
 Most of these critiques, in one way or another, charge that a strict theoretical and practical separation between military and political operating spheres, which Huntington to a certain degree certainly argued for, is fundamentally flawed.
-Many, if not all, of these critiques cite Clausewitz's famous dictum that war is a "mere continuation of policy by other means" [@clausewitz_war_1976, 69, 605], and thus point out that by its very nature, war requires the highest integration - not segregation - of political and military activities.
+Many, if not all, of these critiques cite Clausewitz's famous dictum that war is a "mere continuation of policy by other means" [@clausewitz_war_1976, 69, 605], and thus point out that by its very nature, war requires the highest integration, as opposed to segregation, of political and military activities.
 Moreover, these critiques often point out, correctly in this author's opinion, that the conduct of war is more complicated than civilians simply telling the military the goals it wishes to achieve, and then allowing the military to determine how best to accomplish these goals.
 Rather, the development and implementation of strategy, especially during a time of war, requires an "iterative" dialogue between military leaders and civilians to ensure that effective policies are adopted [@rapp_civil-military_2015, 20].
+
 Others argue that Huntington's conception of military professionalism is too narrow to be of practical use in the contemporary period in that such a model actually enables, rather than constrains, inappropriate political activity by military officers [@brooks_paradoxes_2020].
-Risa Brooks charges that military officers who strongly identify with an ethic that stresses remaining apolitical at all costs, which she alleges follows from a Huntingtonian conception of military professionalism, may develop "blind spots such that they fail to recognize the political content or impact of their actions" [@brooks_paradoxes_2020, 17].
-Whether military leaders recognize it or not, Brooks argues, military leaders comprise a profession that is inherently political, and such a reality ought to be acknowledged and embraced [@brooks_paradoxes_2020].
+Risa Brooks, for instance, charges that military officers who strongly identify with an ethic that stresses remaining apolitical at all costs may in practice develop "blind spots such that they fail to recognize the political content or impact of their actions" [@brooks_paradoxes_2020, 17].
+Her argument is simply that whether military leaders recognize it or not, military leaders comprise a profession that is irreducibly and inherently political to a certain extent.
+Rather than pretend this truth is not so, such a reality ought to be acknowledged and embraced [@brooks_paradoxes_2020].
 
-Regardless of whether and the degree to which one agrees or disagrees with Huntington's concept of military professionalism, a significant number of scholars note that an enduring issue involves attempting to define where the line between appropriate political expression and inappropriate partisan activity should be drawn [@owens_military_2015; @golby_beyond_2015; @brooks_paradoxes_2020].
+# Framing the DV: Variation in the Political Behavior of the Military
+
+Regardless of whether and the degree to which one agrees or disagrees with Huntington's concept of military professionalism, there is nonetheless unity among scholars that in normative terms, civilian control of the military and a corresponding non-partisan ethic among military members are essential norms that should be observed and adhered to.
+Nonetheless, attempting to define where the line between appropriate political expression and inappropriate partisan activity should be drawn constitutes a  [@owens_military_2015; @golby_beyond_2015; @brooks_paradoxes_2020].
 It is a difficult task for many reasons because, as Mackubin Thomas Owens points out, civil-military relations in the United States can be viewed as a "bargain" [@owens_us_2011] between the American people, the military, and the government that has to account for constantly-changing "political, social, technological, or geopolitical changes" [@owens_is_2016].
@@
 
-# Framing the DV: Military Intervention in Politics and Military "Role Beliefs"
-
-Given Huntington's normative model and the critiques his model has been subjected to, it is important to understand what frameworks exist for understanding the debates surrounding why the military would or should intervene in politics, and to what degree.
-Here again, the literature is vast.
-Scholars from the field of comparative politics have generally tackled the question of why militaries intervene in political matters, likely as a result of the variation that is available in country-specific or cross-national studies.
+Given Huntington's normative model and the critiques his model has been subjected to, it is important to understand what frameworks exist for understanding debates involving why and under what circumstances the military intervenes in politics.
 It is vital to briefly understand a couple of these important frameworks.
@@
 The British political scientist, S.E.
-Finer, argues that military intervention in politics is ultimately a function of both the opportunity and desire to do so [@finer_man_1962].
+Finer, argues that military intervention in politics is ultimately a function of the military having both the "opportunity" and "motive" to do so [@finer_man_1962].
 Finer first identifies several factors that should inhibit the military from intervening in politics, including the concept of military professionalism (as espoused by Huntington), strict adherence to the principle of civilian supremacy, and the military's fear that intervention in politics would lead to it fracturing internally, and thus failing to remain a cohesive fighting force [@finer_man_1962].
@@
 problem_dag <- dagify(
+    PB ~ RB,
     RB ~ PC + PP + W,
-    RB~PP,
-    PC ~ PP,
     coords = tribble(
       ~ name,      ~ x,    ~ y,
-      "RB", 1, .5,
-      "PC", .6, .5,
-      "PP", .3, .7,
-      "W", .8,.7,
+      "PB", 1,.5,
+      "RB", .7, .5,
+      "PC", .3, .7,
+      "PP", .3, .5,
+      "W", .3,.3,
     ))
@@
 problem_dag %>% tidy_dagitty() %>% 
-  dag_label(labels=c("RB"="Military Role Beliefs", 
-                     "PC"="Prestige and Credibility", 
-                     "PP"="Political Polarization",
+  dag_label(labels=c("PB"="Political\n Behavior",
+                     "RB"="Role Beliefs\n and Attitudes", 
+                     "PC"="Prestige", 
+                     "PP"="Polarization",
                      "W"="War")) %>% 
@@
   geom_dag_text()+ 
-  geom_dag_label_repel(aes(label=label, fill=label), col="white", show.legend=FALSE)+ theme_dag() + labs(title="Figure 2", subtitle="Determinants of US Military Role Beliefs")
+  geom_dag_label_repel(aes(label=label, fill=label), col="white", show.legend=FALSE)+ theme_dag() + labs(title="Figure 2", subtitle="Determinants of Political Behavior by the Military")
 
@@
 \pagebreak
+
+testimony of a now-retired Army Lieutenant Colonel  United States military  there has never been a coup, nor a serious attempt made by the military to dislodge political control from their duly-constituted elected leaders and to hold it themselves.
+However, as almost every civil-military relations scholar points out, the absence of coups does not mean that the relationship between civilian and military leaders is healthy, nor absent of conflict altogether.
+Indeed, a cursory examination of the relationships between Lincoln and McClellan, Truman and MacArthur, and Obama and McChrystal reveals that disagreement between those who serve at the highest levels of civilian government and military organizations is neither rare nor understudied.
+
+What has changed, however, is the degree to which a very specific subset of the American military institution - senior military leaders - have participated in the domestic political process, especially post retirement.
+Indeed, whereas General Omar Bradley, one of the military's five star generals and a hero of World War Two, famously remarked in 1959, "I am convinced that the best service a retired general can perform is to turn in his tongue along with his suit and mothball his opinions," [@times_bradley_1959], retired generals and admirals today appear frequently in the media, endorse Presidential candidates at higher frequency [@griffiths_not_2019], and frequently serve in government after their retirement from military service [@brooks_dont_2016].
+To provide only one such example, following protests across the United States in the summer of 2020 in which President Donald Trump nearly ordered the intervention of active duty military forces into American cities, former Chairman of the Joint Chiefs of Staff, Admiral (Ret) Michael Mullen, was one of at least ten four star retired officers who publicly aired concerns through a variety of mediums about the President's potential use of active-duty military force.
+^[Other officers included General (Ret) James Mattis, General (Ret) Mark Kelly, Admiral (Ret) William McRaven, General (Ret) Martin Dempsey, General (Ret) Colin Powell, General (Ret) Michael Hayden, General (Ret) David Petraeus, General (Ret) John Allen, and General (Ret) Vincent Brooks.].
+In fact, during an interview in the days which followed the incident, when asked by an NPR reporter why he decided to publicly share his thoughts on the situation, Mullen stated, "we were getting to a point where the country was in jeopardy, which is why I spoke out" [@isaacson_ret_2020; for further analysis, see also @brooks_let_2020].
+
+Mullen's words are incredible from a number of perspectives, and raise important normative questions regarding the military's role in a democratic society [@brooks_let_2020] and the conditions under which the military can and should express dissent to its civilian bosses [@cohn_what_2019].
+And while a small but dedicated number of civil-military relations scholars do examine these and related issues quite frequently in today's charged political environment, few, if any, have attempted to systematically examine and explain over time the conditions under which senior military leaders have broken and will break normative traditions and engage in domestic politics.
+Such an effort would better illumine the understanding of scholars, as well as of the military and society more broadly, concerning why a powerful group of what were once non-political actors have, at least in some cases, suddenly become one.
+
+which they have done so.
+
+understanding of how polarization and prestige impact the political behaviors of the military, and why.
+as been the case for much of the post-9/11 era, but especially in the past several years, This increase in political activism is most concerning, moreover, among a small but powerful group of military elites who hold tremendously positive reputations among a large segment of the increasing political activism at the hands of military officers indicates a shift in the beliefs of the military regarding its central role and purpose in society, or "role beliefs" [@fitch_armed_1998].
+This shift, moreover, is the result of two longstanding and interconnected trends: rising political polarization within the domestic political sphere, and the rising prestige and credibility of military institutions afforded by American society.
+The interaction of these trends explains the remarkable increase in political activity by retired senior military leaders, and portends a shift in the perceived role that these senior military leaders fill vis-a-vis civilian authority - a shift away from professional and obedient "servants" and towards that of "syndicates" or even "saviors" of the American state.
+This shift, moreover, holds enormous ramifications regarding the organizational culture of the military, its relationship with government and society, and American national security.

### 2020-12-18 code/Tahkv1.Rmd
commit

@@
 
-\captionsetup[table]{labelformat=empty}
-
 ```{r setup, include=FALSE}
@@
 library("rstatix")
-```
+library("lme4")
+library("brglm")
+library("bucky")
+```
+
 \newpage
@@
 There are two main arguments of this chapter.
-The first argument is that the variation in the number, topic addressed, and tone of opinion-editorial publications published by retired military officers indicates a shift over the past 40 years in the way a number of retired military officers think about their normative roles as members of the military institution and how they interact with the public in America.
-The second argument is that the driver of the variation in the number, topic, and tone of op-eds authored by retired military officers is mainly the increase in political polarization that has occurred in America since the 1970s.
-While there are other factors that influence also responsible, such as the increase in institutional prestige of the military over the same time period, the empirical story reveals that political polarization holds more explanatory power than other factors in explaining the variation of retired military officer speech in America.
+The first is that the data clearly demonstrates a significant change in the number, topic, and tone of opinion-editorial publications published by retired military officers.
+The second argument holds that the primary driver of this change is the increase in political polarization that has occurred in America over the same time period.
+While other factors certainly play a role in influencing retired military officers to pick up the pen and write pieces that challenge various norms of officership, such as non-partisanship and civilian control of the military, the data clearly reveals that political polarization holds more explanatory power than other factors in explaining the variation of retired military officer speech in America over the past four decades.
+Together, this analysis highlights the extent to which retired military officer speech has strayed from professional normative standards.
 
 This paper proceeds in three parts.
-First, the paper connects the concept of military "role beliefs" to retired military officer speech.
-Second, the paper explores and analyzes an original data set of nearly 390 opinion-editorial pieces published in major US newspapers by retired military officers since the late 1970s.
-This analysis ultimately demonstrates that although the vast majority of op-eds published by retired military officers maintain professional standards, there is clear evidence that retired military officers are using the pen both to argue about topics that fall outside of the realm of traditional national security topics, and to violate the principle of civilian control of the military when doing so.
-The analysis also shows that these trends have worsened as the American political climate has become increasingly polarized over the past several decades.
-Third, thechapter concludes by exploring the implications of these results on the military as an institution, and specifically, on its ability to remain professional in a highly polarized setting that will likely endure.
+First, the paper connects the behavior of retired military officer speech to the larger concept of military "role beliefs," a concept that is, I argue, worthy of reflection today, when both political polarization and the degree of societal prestige afforded to the military are at relatively high levels.
+Second, the paper explores and analyzes an original data set of nearly 390 opinion-editorial pieces published in major US newspapers by retired military officers since the late 1970s, and concludes that although the vast majority of op-eds published by retired military officers maintain professional standards, there is clear evidence that retired military officers are using the pen both to argue about topics that fall outside of the realm of traditional national security topics, and to violate norms of civilian control of the military and non-partisanship when doing so.
+Third, this paper discusses the implications of these findings on the challenges that face the military in what is perhaps the most difficult challenge it has faced in a long time: remaining professional in a highly polarized and divided society.
 
@@
 ## Military Role Beliefs 
-The concept of military "role beliefs" was introduced by J.
+The concept of military "role beliefs" was first introduced by J.
 Samuel Fitch in his book, _The Armed Forces and Democracy in Latin America._ In the book, Fitch describes military "role beliefs" as the perceptions held by military officers regarding their normative role in politics [@fitch_armed_1998, 61].
-Though Fitch's focus is on militaries in Latin America, it is possible to use the concept of "role beliefs" to develop a normative starting point applicable to military officers in the United States.
-
-Two foundations for developing a starting point for military "role beliefs" in the United States are the tandem norms of civilian control of the military, and the idea of the military acting as a non-partisan institution.
-Though such norms have been episodically challenged throughout American history in a variety of ways and during specific instances, these norms have nonetheless been both foundational and aspirational, particularly in the post World War Two period [@weigley_american_2001].
-The norm of civilian control of the military stems directly from the US Constitution, which expressly grants control of the military (as well as specific responsibilities associated with maintaining control) to the President, Congress, and the individual states [@noauthor_us_1787-2].
-Unlike the norm of civilian control, the norm of the military functioning as a non-partisan institution is not expressly spelled out in the Constitution, but rather the product of the American military's collective experiences and traditions in fighting and prevailing in the nation's wars and conflicts.
-
-Further, a number of scholars have underscored the importance of these two norms, and the relationship between these norms and the military's ability to prevail during times of war.
-In his influential book, _The Soldier and the State: The Theory and Politics of Civil-Military Relations_, Samuel Huntington argues in favor of maintaining separate operating spheres between the political leaders of a democracy and the military as an institution in order to ensure civilian control of the military by rendering the military "politically sterile and neutral," or non-partisan [@huntington_soldier_1957, 84].
-Huntington goes on to argue that only by such an arrangement, which he called "objective control of the military," will the military be capable of defeating significant security threats while also remaining apolitical [@huntington_soldier_1957, 83-84].
-
-Having established the foundational norms of civilian control of the military and the military operating as an apolitical institution, the next step is to extrapolate how these norms might impact the public speech of military officers.
-This requires balancing these norms with the fact that military officers are nonetheless servants of the Republic who have accumulated knowledge and experiences over the course of sometimes lengthy careers (in the case of retired military officers, at least 20 and in some cases, upwards of 40 or even more years of experience), and thus, in many cases are qualified to weigh in on circumstances and situations that impact the nation.
-For instance, in the aftermath of the "Revolt of the Generals" in 2006, in which several retired military officers publicly criticized and called for the removal of Defense Secretary Donald Rumsfeld, Frederick Kagan argued in support of retired military officers having the freedom to weigh in policy debates, particularly those involving an active armed conflict, stating, "there is no danger to the republic in a handful of retired generals speaking their minds.
-There is great danger in making vital decisions about an on-going armed struggle without hearing the views of all available experts" [@kagan_let_2006].
-Kagan's remarks lend support to the notion that in democracies in particular, transparent and flowing information is a precious resource that helps leaders make decisions and, more importantly, the public to gain awareness of the state's actions [need cite].
-
-Nonetheless, there have been both formal and informal boundaries placed on various forms of political activities by military officers.
-Formally speaking, Department of Defense (DoD) Directive 1344.10 contains a "standing list of political behavior do's and don'ts for members of the active duty military" [@urben_wearing_2014, 1].
-This list includes authorizing members of the military to engage in activities such as voting and joining a partisan or non-partisan club and to attend this club's meetings when not in uniform, and prohibits members of the military from activities such as participating in partisan fundraising, or advocating for or against a partisan cause or figure on radio or television [@department_of_defense_political_2008].
-Informally, a number of recent Chairmen of the Joint Chiefs of Staff have likewise urged members of the military to refrain from engaging in overly political activities, particularly in election years, when such activity is more likely to negatively reflect on the military.
-For example, prior to the 2016 Presidential Election, then-Chairman of the Joint Chiefs of Staff General Joseph Dunford reminded the military that the basis upon which it is able to maintain the public's trust is it's ability to remain "apolitical and neutral" [@dunford_upholding_2016].
-
-Based on these formal and informal prohibitions on the political activities conducted by military officers, it is possible to conclude that visible behaviors that are overtly partisan are to be avoided at all costs, and are inappropriate for military officers to engage in.
-It is also possible to extrapolate that the norm of civilian control of the military results in a corresponding prohibition on members of the military criticizing political leaders to the point that civilian control of the military is jeopardized or fails to be upheld.
-Where this line should be drawn in practice is not clear, however, because as Kagan has pointed out, criticisms can and often do point out blind spots in a plan or otherwise reveal shortcomings that may not otherwise have been identified.
-However, there is a point at which a criticism levied at a political leader by a military officer goes beyond criticism.
-Finally, it is possible to conclude that activities which build on the appropriate expertise and knowledge of the military officer are appropriate, provided that these activities do not violate the stipulations previously mentioned, i.e.
-a non-partisan military, and civilian control of the military.
+At the center of the concept of role beliefs are the ideas that military officers have about the extent to which they should participate in various aspects of the political process and engage in various forms of political behavior.
+While Fitch's focus is on militaries in Latin America, it is possible to use the concept of "role beliefs" to develop a normative starting point applicable to military officers in the United States.
+
+Theoretically, there are at least two major determinants of military role beliefs, or the conceptions held by members of the military regarding their normative role in politics [@fitch_armed_1998].
+In the most general terms, the military's involvement in politics requires both the _motive_ and the _opportunity_ to do so [@finer_man_1962].
+I argue that in today's domestic political environment,  rising political polarization serves as the _motive_ for the military's intervention in politics, while high levels of prestige, or credibility of the military institution, provides the _opportunity_ for doing so.
+It is important to briefly elaborate how each of these offers the military in particular the ability to intervene in politics.
+
+With respect to the first determinant, political polarization, the literature in American politics enables the development of two plausible relationships and corresponding mechanisms that connect polarization to  military role beliefs.
+These relationships and mechanisms further stem from a particular conception or definition of what polarization is and how it works.
+One possible conception of polarization is that it marks the ideological widening between political elites, to include members of Congress or the leaders of political parties [@hare_polarization_2014].
+According to this view of polarization, one possible mechanism by which the military intervenes in politics is by adopting a belief that elites from both dominant political parties have failed to advance and implement an effective governing agenda, and thus, the military steps in to do what political elites are not doing.
+This particular dynamic has been noted, in fact, by scholars in the comparative politics sub-field describing the actions of militaries in other parts of the world [@fitch_armed_1998].
+
+Yet a second, and I believe the more likely, relationship between political polarization and military role beliefs, particularly in the contemporary United States, stems from a conception of polarization as the strength of feelings that one has concerning a number of ideological issues.
+This view of polarization is more consistent with a concept scholars have labeled "affective polarization", or "negative partisanship" [@abramowitz_great_2018].
+If the variables of polarization and military role beliefs are linked in this way, the military does not intervene as a monolithic block due to a frustration with the political class's inability writ large to effectively govern.
+Rather, certain military elites, who share beliefs with a polarized electorate on any number of ideological issues, such as those regarding race, gender, and immigration [@abramowitz_great_2018], will selectively intervene in order to advance their personal beliefs regarding these topics.
+In other words, this mechanism posits that certain military elites feel as strongly as average citizens about certain polarizing issues, to the point that they will break from historical apolitical or non-partisan military traditions and norms in order to advance their beliefs.
+Thus, when polarization is high, there is a _motive_ among certain members of the the military, to include retired military elites, to deviate from their historical non-partisan and apolitical traditions.
+In short, increasing political polarization may make it increasingly difficult for members of the military to remain politically neutral as partisan conflicts subsume a greater number of issues, and as strong feelings of dislike or disdain surface against those who hold on opposing view.
+
+The second determinant of military role beliefs is the level of prestige, or credibility, that the public places in the military institution.
+When these levels are high, I contend that members of the military possess the _opportunity_ to engage with the public.
+This includes engaging in behaviors such as making public arguments in the press, publishing opinion pieces, or more generally, sharing one's views.
+When society overall esteems the military at low levels, the expectation is that there would be fewer instances of the military publicly sharing its views, because there is less desire from the public (and thus, less desire from press outlets) to hear from members of the military community.
+There may also be, during periods when the military commands low levels of prestige and credibility, less inclination among members of the military to make public pronouncements.
+When the opportunities to engage with the public are more prevalent as a result of high levels of prestige, certain members of the military, I contend, are more likely to feel empowered to take on increasingly active roles, particularly in helping to weigh in or solve problems.
+Thus, we would expect to see a greater level of overall engagement with the press, and therefore, a greater number of overall opinion pieces authored by retired military officers, when military credibility is high.
+
+## Civilian Control, and Non-Partisanship
+
+There are at two seminal principles that have helped guide and influence the military's "role beliefs" for most of US history: the tandem norms of civilian control of the military, and the non-partisanship of the military as an institution.
+Though these norms have been episodically challenged in a variety of specific instances and events, both norms have prominently featured within the military's organizational culture, and increasingly so as the military has modeled the behaviors of a profession, particularly in the post World War Two period [@weigley_american_2001].
+The first norm of civilian control of the military stems directly from several provisions outlined in the US Constitution, which expressly grants control of the military (as well as a division of specific roles and responsibilities associated with maintaining control) to the President, Congress, and the individual states [@noauthor_us_1787-2].
+The second norm of non-partisanship, while not expressly spelled out in the Constitution, has been cultivated throughout the military's collective experience of fighting and prevailing in the nation's wars and conflicts.
+
+A number of scholars have likewise underscored the importance of and relationship between the norms of civilian control and non-partisanship of the military.
+Perhaps the most influential of these scholars has been Samuel Huntington, whose book, _The Soldier and the State: The Theory and Politics of Civil-Military Relations_, argued that maintaining the norms of civilian control of the military and non-partisanship would also ensure that the military prevailed against the dangerous threat environment of the Cold War.
+Huntington specifically argues in favor of a model of civilian control of the military known as "objective control" in which the military and civilian occupy largely separate spheres of operation such that civilian leaders do not interfere with the details of military operations, and likewise, the military does not intervene in politics [@huntington_soldier_1957].
+Only by such an arrangement, Huntington argues, would the military exist as a "politically sterile and neutral," and thus, non-partisan, institution, while simultaneously ensuring civilian control of the military [@huntington_soldier_1957, 83-84].
 
 ## Opinion-Editorial Publication as a Political Activity?
-We can now focus on the actual behavior of the military authoring op-ed pieces.
-Should these considered a form of political activity?
-I argue that it depends on what message the op-ed actually sends or is intended to send.
-An op-ed that strictly points out the pitfalls of the nation's combat strategy written by an experienced military officer may not be inappropriate, whereas an op-ed written by a retired military officer that argues in favor of a specific Presidential candidate is, at least by the standards put forth in this paper, entirely inappropriate.
-Therefore, at least two questions should be asked when examining an opinion piece authored by a retired military officer: first, does the piece violate either the norm of civilian control of the military or the military as an apolitical institution, and second, does the op-ed address an appropriate topic that a military officer is qualified to address?
-
-Beyond these two questions, however, opinion-editorials are important because they serve as an informative barometer of how the military institution sees itself, and particularly, its role with respect to the American public.
-Members of the active military do not frequently write op-eds, but there have been notable exceptions over the years, perhaps none so notable as several written by then-Chairman of the Joint Chiefs of Staff General Colin Powell in the early 1990s.
-For instance, Powell presented the public with his view regarding the potential use of American military force in several developing "hot spots" across the world, including the Balkans, Somalia, and other places, during an October 8, 1992 piece entitled, "Why Generals Get Nervous." Published in the _New York Times_, Powell wrote in part, "decisive means and results are always to be preferred, even if they are not always possible.
-So you bet I get nervous when so-called experts suggest that all we need is a little surgical bombing or a limited attack.
-When the desired result isn't obtained, a new set of experts then comes forward with talk of a little esclalation.
-History has not been kind to this approach" [@powell_opinion_1992].
-
-Powell clearly felt compelled to inform the public of how he, as the military's top officer, viewed the potential use of American military force abroad during the prevailing international context at the time.
-What is less known, however, is that Powell was also writing against a domestic backdrop in which there was an election coming up in less than a month that pitted the incumbent, George W.
-Bush, against Democratic challenger Bill Clinton.
-The two candidates differed with respect to foreign policy on a number of issues, and Bill Clinton was not highly respected by the military establishment, a fact that would impact the military's relationship with Clinton during his early years in office [need cite].
-Whether it was proper for Powell to write such an op-ed truly depends: one could argue that the nation's top general was simply informing the public based on his years of individual experience serving in the military, in which case, from a democratic transparency standpoint, the op-ed well served the Republic; alternately, one could also argue that by writing during an election year, Powell's remarks brazenly staked the claims of the military ahead of a potential Presidential transition (which turned out to be the case), and in the process, limited the President's options for potentially using force abroad during his early days in office.
-The first explanation argues that from the standpoint of civil-military relations, there was no problem with Powell's opinion-editorial, while the second argues that the op-ed was nothing but problematic.
-
-## Retired Military Officers
-One might argue that the logic employed thus far is seemingly plausible for members of the active military, but that the cautionary tone and reservations regarding certain types of political activity should not apply for members of the retired military community.
-And indeed, there are obvious distinctions between those who are on active duty and those who have retired from the service.
-These distinctions are made clear in fact in portions of DoD 1344.10 [@department_of_defense_political_2008].
-But the fact of the matter is that, especially for retired military officers who achieved general or flag officer rank, these retired officers sometimes carry outsized influence and impact when they engage in activities that approach on the political.
-During the Presidential Campaign of 2016, Retired Army General Martin Dempsey published an opinion-editorial piece in _The Washington Post_ which criticized retired Army Major General Mike Flynn and Marine Corps General John Allen for participating in the respective Republican and Democratic conventions.
-Dempsey wrote, "As generals, they have an obligation to uphold our apolitical traditions.
-They have just made the task of their successors – who continue to serve in uniform and are accountable for our security - more complicated.
-It was a mistake for them to participate as they did.
-It was a mistake for our presidential candidates to ask them to do so" [@dempsey_military_2016].
-
-Of course, speaking at a political convention on behalf of a Presidential candidate is, if not formally, at least informally the textbook definition of a political activity.
-Not every opinion-editorial piece supports a political candidate's agenda, or disparages the political views of another.
-Thus, it is important to remember that although an op-ed can be written in such a way that it is a politically inappropriate form of political behavior, it need not be.
-It is precisely on this point that the norms of civilian control of the military, the military as an apolitical institution, and retired members of the military expressing views that are in accord with the knowledge and skills gained during their time in uniform are appropriate.
-
-Moreover, because multistar generals and admirals in the contemporary military possess in many cases upwards of 30 and sometimes 40 years of service, it is entirely reasonable to conclude that these high-ranking officers in particular have been privy to decisions and experiences occurring at or near the highest levels of government.
-Having interacted with Presidents, members of the cabinet, members of Congress, and foreign dignitaries, the experiences of retired general and flag officers are not limited to only tactical matters.
-Though they possess an abundance of expertise in many tactical matters that pertain to warfighting, they are also familiar with issues ranging from how wars should be fought, to future security threats looming on the horizon, to fixing the institutional culture of large organizations such as the military services, and much more.
-
-Nonetheless, they are not experts in everything, nor should they try to be.
-Even General of the Army Douglas MacArthur, one of several men who ever wore the five star rank, and a hero of World War Two who later clashed with President Truman regarding American strategic goals in Asia, recognized that the military had focus areas, even if he did not personally exhibit these limitations while on active duty.
+
+Scholars have long commented on where the line should be drawn that separates appropriate vs.
+inappropriate political behavior by military officers, to include military retirees [@brooks_militaries_2009; @owens_military_2015; @golby_beyond_2015].
+Most scholars note that this is in many ways a balancing act: while the norms of civilian control of the military and non-partisanship are as close to what can reasonably be considered "sacred" within and among the military, military officers are (especially in the case of those who achieve high rank) have accumulated unique knowledge and experiences over the course of lengthy careers (in the case of retired military officers, at least 20 and in some cases, upwards of 40 or even more years of service).
+In many cases, these military elites have interacted with Presidents, members of the cabinet, members of Congress, and foreign dignitaries.
+Though they possess an abundance of expertise in many tactical matters that pertain to war-fighting, they are also familiar with issues ranging from how wars should be fought, to future security threats looming on the horizon, to fixing the institutional culture of large organizations such as the military services, and much more.
+Moreover, especially in a democracy such as the United States, there is a reasonable expectation that these leaders use their military experience to contribute to the free flow of information that helps inform the citizens of a Republic and enables elected leaders to weigh the decisions before them.
+Nonetheless, military elites are not experts in everything, nor should they try to be.
+Even General of the Army Douglas MacArthur, one of several men who ever wore the five star rank, and a hero of World War Two who later clashed with President Truman regarding American strategic goals in Asia, recognized that the military should have focus areas.
 Addressing the Corps of Cadets at West Point in 1962, the then-82 year old retired General told the Corps:
@@
 \doublespacing
-By systematically examining the op-eds written by retired military officers in major US publications over time, it is possible, I contend, to draw inferences about how the concept of a "role belief" - conceptions of one's normative role in politics - has changed among retired US military officers (if at all) [@fitch_armed_1998].
-Op-eds are an interesting and appropriate medium to examine if differentiation in "role beliefs" is an object of interest.
-The first reason why this is the case is because as previously stated, the act of crafting and publishing an op-ed does not necessarily constitute a political activity, though depending on content, they can be.
-A second and releated reason is because unlike other forms of engaging with the press, such as giving interviews on television, crafting an op-ed requires the author to deliberately prepare and submit a viewpoint for publication.
+
+There have likewise been both formal and informal boundaries placed on various forms of political activities by military officers over the years.
+Formally, Department of Defense (DoD) Directive 1344.10 contains a "standing list of political behavior do's and don'ts for members of the active duty military" [@urben_wearing_2014, 1].
+This list includes authorizing members of the military to engage in activities such as voting and joining a partisan or non-partisan club and to attend this club's meetings when not in uniform, and prohibits members of the military from activities such as participating in partisan fundraising, or advocating for or against a partisan cause or figure on radio or television [@department_of_defense_political_2008].
+Informally, a number of recent Chairmen of the Joint Chiefs of Staff have likewise urged members of the military to refrain from engaging in overly political activities, particularly in election years, when such activity is more likely to negatively reflect on the military.
+For example, prior to the 2016 Presidential Election, then-Chairman of the Joint Chiefs of Staff General Joseph Dunford reminded the military that the basis upon which it is able to maintain the public's trust is it's ability to remain "apolitical and neutral" [@dunford_upholding_2016].
+Clearly present in both the formal regulations and informal admonition regarding the political behavior of military officers is concern for the military's non-partisan ethos.
+
+The specific act of authoring opinion commentary has occasionally garnered increased attention in recent years, perhaps starting in 1992 when then-Chairman of the Joint Chiefs of Staff General Colin Powell authored a piece that presented his views regarding the potential use of American military force in several developing "hot spots" across the world (including the Balkans, Somalia, and other places) entitled, "Why Generals Get Nervous." Published in the _New York Times_ less than a month away from the 1992 Presidential Election, Powell wrote in part, "decisive means and results are always to be preferred, even if they are not always possible.
+So you bet I get nervous when so-called experts suggest that all we need is a little surgical bombing or a limited attack.
+When the desired result isn't obtained, a new set of experts then comes forward with talk of a little escalation.
+History has not been kind to this approach" [@powell_opinion_1992].
+More than a decade later, in 2006, the so-called "Revolt of the Generals" occurred in which several retired military officers publicly criticized and called for the removal of Defense Secretary Donald Rumsfeld as a result of his leadership and its impacts on the war in Iraq.
+Scholars noted that the event was unprecedented in the way that multiple officers "devoid of partisan political motive, felt the need or obligation to speak out publicly regarding policy and leadership during an ongoing conflict" [@cook_revolt_2008, 7].
+While the "Revolt of the Generals" generated concern that the military's professional ethic had been strained, others applauded the generals' honesty and forthrightness and claimed that concerns about a dip in the professionalism among military officers was overblown.
+Frederick Kagan, for instance, claimed, "there is no danger to the republic in a handful of retired generals speaking their minds.
+There is great danger in making vital decisions about an on-going armed struggle without hearing the views of all available experts" [@kagan_let_2006].
+
+Whether one agrees with Kagan's general assessment regarding the impact of opinion pieces authored by military officers, it is important to understand the ways in which opinion commentary is different from other types of speech made by military officers, such as that provided during formal testimony to Congress.
+Risa Brooks, for instance, argues that opinion or commentary is distinguished by both its "public nature" as well as the "scope and nature of the commentary" [@brooks_militaries_2009, 218].
+Building off of Brooks, I would add that the action of authoring opinion commentary is completely voluntary, and thus, a deliberate action.
+Crafting a written op-ed is different from even giving interviews on television, because writing an op-ed requires the author to deliberately prepare and submit a viewpoint for publication.
 A deliberate and carefully-crafted message further underscores the degree to which the author believes the message he or she is sending.
 It is difficult to "misspeak" during or to be "misquoted" from an opinion-editorial.
-The author generally agrees with the message he or she is sending.
-
-When examining op-eds authored by retired military officers over time to detect potential shifts in normative role beliefs, the key area of focus must be the degree to which retired military officers have upheld vice strayed from the norms of civilian control and non-partisanship of the military, as well as the degree to which retired officers have addressed topics that are appropriately rooted in relevant military expertise.
+
+In short, we should conclude that military officers who engage in authoring commentary or opinion pieces genuinely believes the message he or she is sending.
+As Brooks points out, however, the content of the specific message matters, and thus, the act of writing an op-ed in and of itself is not problematic.
+An op-ed that strictly points out the pitfalls of the nation's combat strategy while avoiding blatant criticism of a civilian official may not be inappropriate at all, whereas an op-ed written by a retired military officer that argues in favor of a specific Presidential candidate or policy is entirely inappropriate, at least by the normative standards of non-partisanship and civilian control of the military.
+Therefore, at least two questions should be asked when examining an opinion piece authored by a retired military officer: first, does the piece violate either the norm of civilian control of the military or the military as an apolitical institution, and second, does the op-ed address an appropriate topic that the military officer, on the basis of his or her professional expertise, is qualified to address?
+
+## The Role of Retired Military Officers
+
+Beyond these two questions that an individual officer (active or retired) should ask when thinking about writing an op-ed, there is, I argue, an even larger and broader question: what message or messages is the Republic receiving from the collective body of military leaders who provide opinion-laden commentary?
+Because military elites tend to provide this type of commentary more often than those on active duty, this question is particularly relevant for retired military elites.
+The collective tone, nature, and message sent by retired military elites serves as a barometer of sorts regarding the role that the retired officer corps sees for itself in interacting with the American public.
+Are retired military elites upholding norms of non-partisanship and civilian control of the military?
+When they are not, should we view these instances as a rare aberration, or are there clear trends that can be detected that should legitimately cause concern?
+
+One reason to believe that a shift in the "role beliefs" among retired military officers may be occurring is the increasing political behavior of retired military elites during recent national political events.
+During the Presidential Campaign of 2016, for instance, Retired Army General Martin Dempsey, who previously served as Chairman of the Joint Chiefs of Staff, published an opinion-editorial piece in _The Washington Post_ which criticized retired Army Major General Mike Flynn and Marine Corps General John Allen for participating in the respective Republican and Democratic conventions.
+Dempsey wrote, "As generals, they have an obligation to uphold our apolitical traditions.
+They have just made the task of their successors – who continue to serve in uniform and are accountable for our security - more complicated.
+It was a mistake for them to participate as they did.
+It was a mistake for our presidential candidates to ask them to do so" [@dempsey_military_2016].
+Thus, there was both a clear expectation and a concern among some prominent military elites that their fellow retired generals and admirals were not adhering to the norms of civilian control and non-partisanship of the military institution.
+Additionally, after four years of the Trump administration, during which several retired military officers served in prominent positions within the administration, and now that President-Elect Biden has, at least as of the time of writing this paper, nominated another retired general for the position of Secretary of Defense, there is a renewed interest in the normative roles that retired military elites should and actually do have in America.
+
+By systematically examining the op-eds written by retired military officers in major US publications over time, I contend that it is possible to identify potential changes (or none) in the role beliefs held by retired US military officers.
+In doing so, the key area of focus must be the degree to which retired military officers have upheld vice strayed from the norms of civilian control and non-partisanship of the military, as well as the degree to which retired officers have addressed topics that are appropriately rooted in relevant military expertise.
 If there are patterns of violating these norms that can be detected, it may indicate that the role beliefs of certain members of the retired military community have likewise changed.
@@
 
-#Determinants of Military Role Beliefs
-
-Theoretically, there are at least two major determinants of military role beliefs, or the conceptions held by members of the military regarding their normative role in politics [@fitch_armed_1998].
-In the most general terms, the military's involvement in politics requires both the _motive_ and the _opportunity_ to do so [@finer_man_1962].
-I argue that rising political polarization serves as the _motive_ for the military's intervention in politics, while high levels of prestige, or credibility of the military institution, provide the _opportunity_ for doing so.
-It is important to elaborate why each of these offers the military in particular the ability to intervene in politics.^[For a richer elaboration on the mechanisms afforded by political polarization and institutional levels of prestige for intervention in politics, please see the author's theory chapter in his dissertation.
-The proceeding paragraphs here are intended to summarize these mechanisms.]
-
-With respect to the first determinant, political polarization, the literature in American politics enables the development of two plausible relationships and corresponding mechanisms that connect polarization to  military role beliefs.
-These relationships and mechanisms further stem from a particular conception or definition of what polarization is and how it workd.
-One possible conception of polarization is that it marks the ideological widening between political elites, to include members of Congress or the leaders of political parties [@hare_polarization_2014].
-According to this view of polarization, one possible mechanism by which the military intervenes in politics is by adopting a belief that elites from both dominant political parties have failed to advance and implement an effective governing agenda, and thus, the military steps in to do what political elites are not doing.
-This particular dynamic has been noted, in fact, by scholars in the comparative politics subfield describing the actions of militaries in other parts of the world [@fitch_armed_1998].
-
-Yet a second, and I believe the more likely, relationship between political polarization and military role beliefs, particularly in the contemporary United States, stems from a conception of polarization as the strength of feelings that one has concerning a number of ideological issues.
-This view of polarization is more consistent with a concept scholars have labeled "affective polarization", or "negative partisanship" [@abramowitz_great_2018].
-If the variables of polarization and military role beliefs are linked in this way, the military does not intervene as a monolithic block due to a frustration with the political class's inability writ large to effectively govern.
-Rather, certain military elites, who share beliefs with a polarized electorate on any number of ideological issues, such as those regarding race, gender, and immigration [@abramowitz_great_2018], will selectively intervene in order to advance their personal beliefs regarding these topics.
-In other words, this mechanism posits that certain military elites feel as strongly as average citizens about certain polarizing issues, to the point that they will break from historical apolitical or non-partisan military traditions and norms in order to advance their beliefs.
-Thus, when polarization is high, there is a _motive_ among certain members of the the military, to include retired military elites, to deviate from their historical non-partisan and apolitical traditions.
-In short, increasing political polarization may make it increasingly difficult for members of the military to remain politically neutral as partisan conflicts subsume a greater number of issues, and as strong feelings of dislike or disdain surface against those who hold on opposing view.
-
-The second determinant of military role beliefs is the level of prestige, or credibility, that the public places in the military institution.
-When these levels are high, I contend that members of the military possess the _opportunity_ to engage with the public.
-This includes engaging in behaviors such as making public arguments in the press, publishing opinion pieces, or more generally, sharing one's views.
-When society overall esteems the military at low levels, the expectation is that there would be fewer instances of the military publicly sharing its views, because there is less desire from the public (and thus, less desire from press outlets) to hear from members of the military community.
-There may also be, during periods when the military commands low levels of prestige and credibility, less inclination among members of the military to make public pronouncements.
-When the opportunities to engage with the public are more prevalent as a result of high levels of prestige, certain members of the military, I contend, are more likely to feel empowered to take on increasingly active roles, particularly in helping to weigh in or solve problems.
-Thus, we would expect to see a greater level of overall engagement with the press, and therefore, a greater number of overall opinion pieces authored by retired military officers, when military credibility is high.
-
 ## Hypotheses
 
-Given the expectations derived from theories regarding military role beliefs, the overall purpose of opinion-editorial publications, and retired military officers as members of the military community, five distinct hypotheses are derived.
-These include:
+Given the theoretical importance of the norms of civilian control and non-partisanship of the military institution, the rise of domestic political polarization and prestige of the US military over the past several decades, and the unique contributions that retired military elites make in a democratic society, four distinct hypotheses are derived, which will be examined and tested in the analysis to follow.
+These hypotheses include:
 
 \singlespace
-**H1: When political polarization is high, the military (or certain military figures) will increasingly criticize civilian leadership (indicating a drop in the principle of civilian control).**
+**H1: Increasing political polarization is associated with an increase in commentary authored by retired military officers that criticizes civilian leadership or makes clear partisan statements (potentially indicating a drop in adherence among retired officers to the principle of civilian control).**
 
@@
 \singlespace
-**H2: When political polarization is low, the military (or certain military figures) will publicly weigh in on traditional national security debates (and not on domestic political concerns or issues outside of relevant military expertise).**
+**H2: Increasing political polarization is associated with an increase in commentary authored by retired military officers that addresses topics that are outside of traditional national security debates.**
 
@@
 \singlespace
-**H3: During periods of high prestige and credibility, the frequency of military involvement in quasi-political activities (such as writing op-eds) will increase relative to periods of low prestige and credibility.**
+**H3: During periods of high prestige and credibility, the aggregate level of commentary offered by retired military officers is higher relative to lower periods of prestige and credibility.**
 
@@
 \singlespace
-**H4: When both political polarization and high prestige are present, the military will be increasingly vocal about certain domestic ideological or partisan causes than when these factors are not present.**
+**H4: During periods of active ground combat, military elites will increasingly comment on topics related to the war, and comment less on other domestic issues.**
 
@@
 \singlespace
-**H5: During periods of active ground combat, military elites will likely comment on the war, but with less frequency on other domestic issues.**
 
 \doublespace
-
-Some caution must be taken up front so as to not overpromise and under-deliver concerning what the data is actually capable of presenting.
-The first caution to consider is that opinion-editorial publications by nature reflect the demand of the press to publish the viewpoints of retired military officers.
-Moreover, because op-eds are generally provocative and at least to some degree designed to advocate for or against a certain course of action, it is possible that op-eds may generally present somewhat contentious material.
-Additionally, this means that under periods of harmonious civil-military relations, there is either less of a demand from the press to publish the viewpoints of retired military officers, or less of a supply side effect for retired military officers to write about topics in the first place.
-Certain exogenous events, such as wars, will impact whether retired military officers speak out.
-A second cautionary note is that over the course of several decades, underlying societal changes - such as political polarization - have occurred that likely impact either or both the willingess of the press to publish the viewpoints of military officers, or the desire of military officers to write about certain topics in the first place.
-It is important that these societal-level effects be identified and, if possible, controlled for.
 
@@
 The first is to analyze the topics which retired military officers have been writing about, and the second is to determine, regardless of the topic addressed in a publication, whether an op-ed has violated the norms of non-partisanship or civilian control of the military.
-The critical question is whether there are patterns in the data that can be detected, and whether these patterns can be explained.
-
 The data consists of 390 opinion-editorial publications and letters to the editor authored by retired military officers and published in major US newspapers from 1979-2020.
@@
 Because I am interested in messages that required the officer to craft a deliberate message, I do not include published interviews with retired military officers, nor any reprinted speeches or remarks that a retired military officer gave at another venue (while searching for the data, there were a few instances in which newspapers published remarks of officers made at events).
+I also do not include short forms of analysis, such as that provided by the "Monkey Cage" from _The Washington Post_.
 
@@
 
-Actually obtaining the data is a time consuming process.
-I use a series of key-word searches, such as "retired military officer," "retired from the Army, Navy, Marine Corps, Air Force," "former Chairman of the Joint Chiefs of Staff," "General," "Admiral," and so forth in several different databases, including Factiva, Lexus-Uni, and ProQuest Historical Newspapers.
-^[For exact replication of the database search parameters, please see the Appendix.] Each of these databases, I learn, classifies historical newspaper articles differently: some databases include editorials as a separate source type to search for (making the acquisition of data easier), while others do not, requiring a search of all articles that contain the key words (which therefore results in a significantly larger pool of articles to search through and find the op-eds and letters I am looking for).
+To obtain the data, I use a series of key-word searches, such as "retired military officer," "retired from the Army, Navy, Marine Corps, Air Force," "former Chairman of the Joint Chiefs of Staff," "General," "Admiral," and so forth in several different databases, including Factiva, Lexus-Uni, and ProQuest Historical Newspapers.
+^[For exact replication of the database search parameters, please see the Appendix.] Each of these databases classifies historical newspaper articles differently: some databases include editorials as a separate source type to search for (making the acquisition of data easier), while others do not, requiring a search of all articles that contain the key words (which therefore results in a significantly larger pool of articles to search through and find the op-eds and letters I am looking for).
 I include in my search publications that were made available both in print and online format, though I did not double-count: if the same op-ed was published both in print and online and published by the same author, I count the article only one time.
@@
 This occurred in several instances in which the officer was, ostensibly, writing from the perspective of someone other than a retired military officer, such as that of a political appointee to a high level governmental position (after retirement from the military, Army General Barry McCaffrey, for instance, served as President Clinton's Director of the Office of National Drug Control Policy).
-Nonethless, because these officers had previously identified themselves as retired military officers in earlier publications, and thus, because members of the public might reasonably recognize the author as a prominent retired military figure, I include the subsequent publications in the data, and simply control for whether the author identified him or herself as a retired military officer in a specific publication.
+Nonetheless, because these officers had previously identified themselves as retired military officers in earlier publications, and thus, because members of the public might reasonably recognize the author as a prominent retired military figure, I include the subsequent publications in the data, and simply control for whether the author identified him or herself as a retired military officer in a specific publication.
+
+Some caution must be taken up front so as to not over-promise and under-deliver concerning what the data is actually capable of presenting.
+The first caution to consider is that opinion-editorial publications to some degree reflect not only the views of retired military officers, but also the desire of the press, or in economic terms, the "demand" of the press to publish the views of these officers.
+Moreover, because op-eds are to some degree designed to advocate for or against a certain course of action, op-eds will likely present somewhat provocative material.
+As a result, under periods of harmonious civil-military relations, there may be either less of a demand from the press to publish the viewpoints of retired military officers, or less of a supply side effect for retired military officers to write about topics in the first place.
+Moreover, both the supply and demand for commentary written by retired military elites will likely be somewhat driven by exogenous events, such as wars.
+A second cautionary note is that over the course of several decades, underlying societal changes, including political polarization, have occurred.
+These changes have likewise likely impacted the willingness of the press to publish certain viewpoints of military officers, and/or the desire of military officers to write about certain topics in the first place.
+It is important that these societal-level effects be identified and, if possible, controlled for.
 
@@
 For example, where General (Retired) David Petraeus wrote one publication and another one was written by General (Retired) David Petraeus and Michael O'Hanlon, I counted these as two different authors for coding purposes.] There are also several retired military officers who penned several observations.
-The most prolific of these officers includes Retired Army Lieutenant General William Odom (22 publications in total), Retired Air Force General Michael Hayden (20 publications in total), and Retired Marine Lieutenant Colonel and Former National Security Advisor Robert McFarlane (18 publications).
-
-```{r summary stats, include=FALSE, echo=FALSE}
+The most prolific of these officers includes Retired Army Lieutenant General William Odom (22 publications in total), Retired Air Force General Michael Hayden (20 publications in total), and Retired Marine Lieutenant Colonel and Former National Security Adviser Robert McFarlane (18 publications).
+
+```{r, include=FALSE, echo=FALSE}
 # show summary stats for yearly count, rank, and then a breakdown according to source
@@
   'Author Rank', 4, 10, 8.34, 1.99, 390,
-  'Number of Yearly Publications', 1, 27, 9.51, 6.84, 390,
-  'Male Authors', NA, NA, NA,NA, 374,
-  'Female Authors', NA, NA, NA, NA, 3,
-  'Mixed Authorship', NA, NA, NA, NA, 5,
-  'Unknown Authorship', NA, NA, NA, NA, 8,
-  'Army', NA, NA, NA, NA, 205,
-  'Navy', NA, NA, NA, NA, 56,
-  'Marine Corps', NA, NA, NA, NA, 47,
-  'Air Force', NA, NA, NA, NA, 63,
-  'Coast Guard', NA, NA, NA, NA, 3,
-  'Mixed Service Authorship', NA, NA, NA, NA, 16)
-
-T1 <- kable(Descriptive_Stats, 
-            format = 'simple', 
-            booktabs = TRUE, 
-            caption = 'Descriptive Statistics (Author Rank and Yearly Number of Publications, and breakdown by Author Sex, Service',
-            align = c('l', 'c', 'c', 'c'),
-            digits = 2)
+  'Number of Yearly Publications', 1, 27, 9.51, 6.84, 390)
+
+Sex_Service_Stats <- tribble(~Breakdown, ~Observations, ~Percentage,
+   'Male Authors', 374, 95.9,
+  'Female Authors', 3, .8,
+  'Mixed  Authorship', 5, 1.3,
+  'Sex Unknown', 8, 2,
+  'Army', 205, 52.5,
+  'Navy',  56, 14.3,
+  'Marine Corps', 47, 12.1,
+  'Air Force', 63, 16.2,
+  'Coast Guard',3,.8,
+  'Mixed Service Authorship', 16, 4.1)
 
@@
 In addition to coding each publication for the source and a count of publications in a given year, several bits of biographical information about each author is collected, including the service in which the author served, the rank achieved by the author during his or her time in service, and the sex of the author.
-Summary statistics for the variables of author rank and the annual count of publications are displayed in Table 1, below, along with a breakdown of the observations by sex and service of the author.
-Mixed authorship refers to publications that were authored by a mixture of male and female authors, whereas mixed service authorship denotes publications that were authored by more than one individual and together, as an author team, had served in multiple service branches - for example, a retired Colonel who had served in the Air Force and a retired Admiral who had served in the Navy write an op-ed together.
-
-```{r table1, echo=FALSE, include=TRUE}
-T1
-```
-
+Summary statistics for the variables of author rank and the annual count of publications are displayed in Table 1, below.
 Of note, the variable for author rank is coded equivalently to the officer's military pay grade.
@@
 
-It is worthy to briefly note the maximum, minimum, and mean values for the number of yearly publications from Table 1, as these numbers can be interpreted in a couple of different ways.
-The overall number of opinion pieces published by retired military officers in a given year may reflect some combination of willingness on the part of the press to publish the viewpoints of officers, or desire by the retired military community to make its viewpoints known, or some combination of these "demand" and "supply" effects.
+```{r table1, echo=FALSE, include=TRUE, warning=FALSE, error=FALSE}
+T1 <- kable(Descriptive_Stats, 
+            format = 'simple', 
+            booktabs = TRUE, 
+            caption = "Descriptive Statistics for Author Rank and Yearly Number of Publications Breakdown of Observations by Author Sex and Service Affiliation",
+            align = c('l', 'c', 'c', 'c'),
+            digits = 2)
+
+T1
+```
+
+Table 2, below, shows the breakdown of observations by author sex and service affiliation.
+Mixed authorship refers to publications that were authored by a mixture of male and female authors, whereas mixed service authorship denotes publications that were authored by more than one individual and together, as an author team, had served in multiple service branches - for example, a retired Colonel who had served in the Air Force and a retired Admiral who had served in the Navy write an op-ed together.
+
+```{r table2, echo=FALSE, include=TRUE, warning=FALSE, error=FALSE}
+T1 <- kable(Sex_Service_Stats, 
+            format = 'simple', 
+            booktabs = TRUE, 
+            caption = "Breakdown of Observations by Author Sex and Service Affiliation",
+            align = c('l', 'c', 'c', 'c'),
+            digits = 2)
+
+T1
+```
+
+Unsurprisingly, as the military's largest service, retired military elites who served in the Army comprise the largest segment (52.5%) of authors in the sample.
+Authors who served in the Navy and Air Force, whose active duty populations in 2020 were of comparable size, featured similarly in the sample (14.3% and 16.2%, respectively).
+Representation from authors who served in the Marine Corps (12.1%) and Coast Guard (.8%) round out the sample.
+In terms of the gender of the author, again unexpectedly, male authors dominate the sample (95.9%).
+Because the sample includes publications dating back to 1979-2020, a probable reason for this is the difference in gender at the highest ranks of service - a feature that will likely change in the coming years as the restrictions on women serving in specific positions within the military have been removed.
 
 Figure 1, below, plots the yearly count of publications authored by retired military officers across all of the major newspapers.
-The smoothed trendline indicates that the average yearly count of publications rose steadily from 1979 until the early 2000s, at which time, a decline occurred until approximately 2008.
+The smoothed trend line indicates that the average yearly count of publications rose steadily from 1979 until the early 2000s, at which time, a decline occurred until approximately 2008.
 From 2008 until the present day, the average yearly count of publications authored by retired military officers in the major US newspapers examined in this study rose heavily during the Obama administration years before briefly falling, but then resuming the high count during the Trump administration.
-
-It is noteworthy, from a macro perspective, to consider all of the impacts on the US military over the past four decades and more.
-Numerous substantive events, debates, and changes have occurred in this time period for retired military officers to weigh in on.
-In terms of external threats and wars, this period includes the final years of the Cold War against the Soviet Union (through 1991), the Gulf War in Iraq (1990-1991), operations in the Balkans (primarily 1993-1995), and the post-9/11 wars in Iraq (2003-present day) and Afghanistan (2001-present day), as well as a host of smaller but important military operations in Grenada (1983), Panama (1989), Somalia (1992-1993), and Haiti (1994-1995).
-The military has also undergone significant social and institutional changes in the time period examined.
-In the late 1970s, the military had ended the draft only years before (1973) following the conclusion of the Vietnam War.
-The debates over "Don't Ask, Don't Tell" (instituted in 1994 and repealed in 2011), as well as policies regarding women in the combat arms (2015) are also included in the data, as are numerous annual discussions regarding the size of military budgets.
-
-```{r figure1, fig.height=4, fig.width=5, fig.align="center", warning=FALSE, include=TRUE, echo=FALSE}
+```{r plot1, fig.align="center", warning=FALSE, error=FALSE, message=FALSE, include=TRUE, echo=FALSE, fig.cap="Number of Op-Eds and/or Letters to the Editor Authored by Retired Military Officers and Published in Major US Newspapers (WSJ, NYT, WaPo, USA Today, LA Times), 1979-2020"}
 
@@
   y="yearly count", 
-  title="Figure 1: Number of Op-Eds Published by Retired Military\n Officers (WSJ, NYT, WaPo, USA Today, LA Times)",
-  subtitle= "1979-2020",
-  caption="Note: As of 1 December 2020"
+  caption="Note: Data as of 1 December 2020"
   ) + scale_x_continuous(limits=c(1979, 2020)) + scale_y_continuous(limits=c(-7, 28)) + 
@@
 ```
+It is noteworthy, from a macro perspective, to consider all of the impacts on the US military over the past four decades and more.
+Numerous substantive events, debates, and changes have occurred in this time period for retired military officers to weigh in on.
+In terms of external threats and wars, this period includes the final years of the Cold War against the Soviet Union (through 1991), the Gulf War in Iraq (1990-1991), operations in the Balkans (primarily 1993-1995), and the post-9/11 wars in Iraq (2003-present day) and Afghanistan (2001-present day), as well as a host of smaller but important military operations in Grenada (1983), Panama (1989), Somalia (1992-1993), and Haiti (1994-1995).
+The military has also undergone significant social and institutional changes in the time period examined.
+In the late 1970s, the military had ended the draft only years before (1973) following the conclusion of the Vietnam War.
+The debates over "Don't Ask, Don't Tell" (instituted in 1994 and repealed in 2011), as well as policies regarding women in the combat arms (2015) are also included in the data, as are numerous annual discussions regarding the size of military budgets.
 
@@
 
-```{r figure2, include=TRUE, echo=FALSE}
+```{r plot2, include=TRUE, echo=FALSE, warning=FALSE, fig.cap="Opinion Pieces authored by Retired Military Officers Broken Down by Author Rank (L) and Newspaper Source (R), 1979-2020"}
 colplot1 <- ggplot(data=rankdistrib, aes(x=AuthRank, y=n)) + labs(x="Pay Grade", y="Number of Publications") + geom_col() + geom_text(aes(label=n), position=position_dodge(width=.9), vjust=-.25)
@@
 
-grid.arrange(colplot1, colplot2, ncol=2, top="Figure 2: Publications authored by Retired Military Officers\n by Author Rank (L) and Source (R), 1979-2020")
+grid.arrange(colplot1, colplot2, ncol=2)
 
@@
 The first is **foreign policy**, if the publication generally refers to American foreign policy or overarching National Security policy.
-The next category of publication is **warfighting and operations**, which I code those publications that address an ongoing conflict, or a central aspect of the military's ability to fight and win on the battlefield.
+The next category of publication is **war-fighting and operations**, which I code those publications that address an ongoing conflict, or a central aspect of the military's ability to fight and win on the battlefield.
 Of note, I also include those publications that deal with most intelligence procedures, as in most cases, the authors of these publications tied intelligence to the military's ability to win on the battlefield.
@@
 Table 2, below, displays the breakdown of observations by topic category.
-The topics of **warfighting and operations** and **foreign policy** account for a total of 230 of the 390 observations (59% of the sample).
+The topics of **war-fighting and operations** and **foreign policy** account for a total of 230 of the 390 observations (59% of the sample).
 The categories of **support**, **domestic policy**, and **budget/weapons/troops** account for a total of 119 observations, or 30.5% of the sample.
@@
 
-```{r table2, fig.align="center", include=TRUE, echo=FALSE, warning=FALSE}
+```{r table3, fig.align="center", include=TRUE, echo=FALSE, warning=FALSE}
 countbytype %>%
-  kable(format="latex", caption="Publications by Subject") %>%
+  kable(format="latex", caption="Publication Breakdown by Subject") %>%
   kable_styling(latex_options = "striped", font_size=10) %>% 
@@
 I contend that the answer is not especially.
-It is unsurprising that retired military officers comment heavily on matters related to warfighting, operations, and national security policy, as these matters directly relate to the experiences of retired military leaders.
+It is unsurprising that retired military officers comment heavily on matters related to war-fighting, operations, and national security policy, as these matters directly relate to the experiences of retired military leaders.
 If anything, it is somewhat surprising, though not necessarily cause for concern, that the number of observations addressing domestic policy matters is as high as it is, and that the number addressing civil-military relations is as low as it is.
@@
 
-```{r figure3, include=TRUE, echo=FALSE, warning=FALSE, error=FALSE}
-ggplot(data=adjdataset, mapping=aes(x=PubYr, y=dom_pol_prop)) + geom_point() + geom_smooth() + labs(title="Figure 3: Proportion of Annual Publications authored by Retired\n Military Officers that Address Domestic Political Matters, 1979-2020", x="Year", y="Proportion", color=NULL)+ geom_vline(xintercept = 1981, size=.2, colour="tomato") +
+```{r plot3, include=TRUE, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE, fig.cap="Proportion of Opinion Publications authored by Retired Military Officers Which Primarily Address Domestic Politics, 1979-2020"}
+ggplot(data=adjdataset, mapping=aes(x=PubYr, y=dom_pol_prop)) + geom_point() + geom_smooth() + labs(x="Year", y="Proportion", color=NULL)+ geom_vline(xintercept = 1981, size=.2, colour="tomato")+
   geom_vline(xintercept=1989, size=.2, colour="tomato") + 
@@
 
-##Second Layer Analysis - A More Brazen Retired Officer Corps?
+## Second Layer Analysis - A More Brazen Retired Officer Corps?
 
@@
 The third and final indicator measures whether a publication calls on Congress or a President to undertake specific legislative actions, such as ratifying a treaty or passing a bill or budget.
-This indicator does not directly measure either non-partisanship or civilian control of the military, but it does measure how comfortable retired military officers feel being involved in a part of the political process, that of calling for and/or making legislative action, which is an experience that virtually all military officers do not develop during their time in uniform.^[A small number of officers may have experience working on the hill, or serving as a Congressional liasion, but I argue that this is different from actually calling upon a politician to engage in adopting a stance regarding a specific piece of legislation.] This is not to say that military officers on active duty or during retirement either should be or in fact are oblivious to this aspect of the political process.
+This indicator does not directly measure either non-partisanship or civilian control of the military, but it does measure how comfortable retired military officers feel being involved in a part of the political process, that of calling for and/or making legislative action, which is an experience that virtually all military officers do not develop during their time in uniform.^[A small number of officers may have experience working on the hill, or serving as a Congressional liaison, but I argue that this is different from actually calling upon a politician to engage in adopting a stance regarding a specific piece of legislation.] This is not to say that military officers on active duty or during retirement either should be or in fact are oblivious to this aspect of the political process.
 Rather, including an indicator for assertiveness (calls to pursue legislative action) measures to what extent the retired military author believes his or her military experience serves as an appropriate basis for calling for legislative action.
@@
 
-Figure 4, below, shows over time the proportion of observations in a given year which either heavily criticizes a civilian leader, takes a clear partisan stance, or calls for specific legislative action to be taken.
-
-```{r second level analysis, include = FALSE}
+```{r, include = FALSE, warning=FALSE, echo=FALSE}
 
@@
 
-# create a df called quest_prof that contains articles that either criticized or were #expressly partisan, as well as whether someone was in a role at the time that may have # made such behavior ok
-
-quest_prof <-adjdataset %>% arrange(PubYr, Topic) %>% 
-  filter(Crit_Insub== 1 | Endorse_Partisan ==1)  %>% 
-  select (Author, Title, Source, PubYr, Topic, Crit_Insub, Endorse_Partisan, Role) %>% 
-  view()
+# create a df called quest_prof that contains articles that either called for legislation 
+# or criticized or were expressly partisan, as well as whether someone was 
+# in a role at the time that may have made such behavior ok
+
+quest_prof <-adjdataset %>% arrange(PubYr) %>% 
+  filter(Crit_Insub== 1 | Endorse_Partisan ==1 | Leg_Act ==1)  %>% 
+  rename(Crit=Crit_Insub, Endorse=Endorse_Partisan, Leg=Leg_Act
+  ) %>% 
+  select (Author, Title, Source, PubYr, Topic, Leg, Crit, Endorse, Role) %>% 
+  view ()
 
@@
 
-```
-
-```{r build plots, include = TRUE, warning = FALSE, error=FALSE, echo=FALSE}
-ggplot(data=adjdataset, mapping=aes(x=PubYr, y=prop_yr)) + geom_point() + geom_smooth() + labs(title="Figure 4: Proportion of Op-Eds Published Which Call for Legislative Action, Criticize Civilian Leaders, or Take Clear Partisan Stances", x="Year", y="Annual Proportion", color=NULL) + geom_vline(xintercept = 1981, size=.2, colour="tomato") +
+# create a variable decade
+authorcalcs <- adjdataset %>% 
+  mutate(
+  decade =case_when(
+    (PubYr < 1980) ~ "1970s",
+    (PubYr >= 1981 & PubYr < 1991) ~ "1980s",
+    (PubYr >= 1991 & PubYr < 2001) ~ "1990s",
+    (PubYr >= 2001 & PubYr < 2011) ~ "2000s",
+    (PubYr >= 2011 & PubYr < 2021) ~ "2010s"
+  )
+) %>% select(Author:Service, Title:Topic, leg_crit_part, crit_part, decade) %>% print()
+
+#count the number of publications by GO/FO author in each decade
+authorcalcs %>% group_by(decade) %>% filter(AuthRank >=7) %>% select(Author) %>% count(Author) %>% view()
+
+#create a table for the Appendix
+TA1 <- quest_prof %>%
+  kable(format="latex", caption="Op-Eds authored by Retired Military Officers that Called for Legislative Action, Strongly Criticized Civilian Leaders, or were Expressly Partisan, 1979-2020", longtable=T) %>%
+  kable_styling(font_size=9, latex_options="striped", repeat_header_continued=T) %>% 
+  column_spec(1, width="9em") %>% 
+  column_spec(2, width="14em") %>% 
+  column_spec(3:4, width="2.5em") %>%
+  column_spec(5, width="5.6em") %>%
+  column_spec(6:7, width="1.4em") %>%
+  column_spec(8:9, width="3.1em") %>% 
+  row_spec(0, bold=T)
+```
+
+```{r plot4, include = TRUE, echo=FALSE, warning = FALSE, error=FALSE, message=FALSE, fig.cap="Proportion of Op-Eds Which Either Argues for Legislative Action, Expressly Criticizes Civilian Leadership, or Takes Clearly Partisan Stances, 1979-2020"}
+
+ggplot(data=adjdataset, mapping=aes(x=PubYr, y=prop_yr)) + geom_point() + geom_smooth() + labs(x="Year", y="Annual Proportion", color=NULL) + geom_vline(xintercept = 1981, size=.2, colour="tomato") + 
   geom_vline(xintercept=1989, size=.2, colour="tomato") + 
@@
   geom_text(aes(x=2017, label="Trump",y=.05), colour="tomato", angle=90, vjust=1, text=element_text(size=9))  
-
-```
+```
+
+Figure 4, above, shows over time the proportion of observations in a given year which either heavily criticizes a civilian leader, takes a clear partisan stance, or calls for specific legislative action to be taken.
 The smoothed trend line in blue shows a story that is similar to the frequency at which retired military officers address domestic political topics.
@@
 A second measurement of political polarization is a measurement of affective polarization, which was first conducted by Iyengar et al using time series data from the American National Election Study [@iyengar_affect_2012; @iyengar_origins_2019].
-Using a series of "feeling thermometer" questions that guage a person's feelings about those on the opposite side of one's political spectrum, this measurement roughly captures how much "dislike" exists among the electorate for each other [@iyengar_strengthening_2018].
+Using a series of "feeling thermometer" questions that gauge a person's feelings about those on the opposite side of one's political spectrum, this measurement roughly captures how much "dislike" exists among the electorate for each other [@iyengar_strengthening_2018].
 This measurement does not reflect an encouraging story: while the level at which citizens like their own party has remained remarkably stable for the past 40 years, the level of affection for those of the opposite political party precipitously dropped around the year 2000, a trend that has continued [@iyengar_strengthening_2018].
@@
 
-```{r figure 5, include=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
+```{r, include=FALSE, echo=FALSE, warning=FALSE, message=FALSE}
 #adjust the data to account for polarization by creating a new dataset called polaradjdataset.  
@@
 polar_and_cred <- melt(subset(polaradjdataset, select=c(PubYr, House_distance_dim1, aff_polar_alt, Inst_Cred_Gallup)), id.var="PubYr")
-
+```
+
+```{r plot5, inculde=TRUE, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE, fig.cap="Polarization Levels in the US House of Representatives, Affective Polarization Levels Among the US Electorate, and American Confidence in the Military, 1979-2020"}
 # 
-ggplot(polar_and_cred, aes(x=PubYr, y=value)) + geom_line(aes(color=variable)) + facet_grid(variable~., scales="free_y") + theme(legend.position = "none") +
-  labs(
-    title = "Figure 5: Polarization Levels in the US House of Representatives,\n Affective Polarization, and American Confidence in the Military, 1979-2020"
-  )
+ggplot(polar_and_cred, aes(x=PubYr, y=value)) + geom_line(aes(color=variable)) + facet_grid(variable~., scales="free_y") + theme(legend.position = "none")
 ```
@@
 Because there are only three observations written exclusively by retired female officers, and because there is no compelling logic for why the service of the author would increase the likelihood of violating norms when publishing an opinion piece, I do not control for the sex or service affiliation of the author.
-Controls are included, however, for the author's rank and the source in which a publication appears.
-
-The dependent variable is an indicator of whether the opinion piece either called for legislative action, blatantly criticized a political leader, or clearly took a partisan stance on an issue.
-As some may argue that calling for legislative action in itself does not violate norms of civilian control of the military or the non-partisanship of the military institution, the dependent variable in an alternate model is an indicator for whether publication criticized a political leader or took a clearly partisan stance.
+I do include source fixed effects in case there is concern that increasing polarization may reflect the fact that individual newspaper sources or other media outlets may have taken on a more-pronounced ideological bent along the political left-right spectrum as polarization has worsened, and thus, the data merely reflect the fact that that certain newspaper sources are more inclined to publish pieces that espouse a point of view reflecting the source's political leanings.^[For example, the _New York Times_ in 2020 is deemed to lean further to the political left than the current  _Wall Street Journal_, which some see as aligned with the political center or slightly right of center [for an informative view, see the Media Bias Chart created by allsides.com [@noauthor_allsides_nodate].]
+
+The dependent variable is a binary indicator of whether the opinion piece either called for legislative action, blatantly criticized a political leader, or clearly took a partisan stance on an issue.
+Because some may argue that calling for legislative action in itself does not violate norms of civilian control of the military or the non-partisanship of the military institution, the dependent variable in alternate models is an indicator for whether a publication either criticized a political leader or took a clearly partisan stance (and does not include instances in which a publication called for legislative action).
 The estimated model therefore takes the form:
 
-$$y = \beta_1 credibility + \beta_2polarization + \beta3source +\epsilon $$
-The results of a logistic regression of this model are displayed in Table 3, below.
+$$y = \beta_0 + \beta_1 MilitaryCredibility + \beta_2Polarization + \beta3AuthorRank + \beta_4Source +\epsilon $$
+
+Because the dependent variable is dichotomous, a logistic regression is most appropriate.
+I run several model specifications including a regular logit model, a logit model that clusters standard errors at the author level, and a mixed model that introduces random effects at the author level.
+The intent of including models which cluster standard errors or introduce random effects at the author level and introduce is to determine the degree to which the data can distinguish if the individual authors themselves are an important determinant of whether an opinion piece criticizes civilians, calls for legislation, or takes a partisan stance.
+In other words, the logit model without clustered errors at the author level assumes that authors have no distinguishing characteristics that would incline them towards behavior that would result in the dependent variable taking on a value of 1, whereas both the mixed model with author random effects and the logit model that clusters errors at the author level allows for the possibility that individual author characteristics may explain the dependent variable in important ways that the model does not otherwise account for.
+The regressions proceed on 379 of the 390 observations (97.2%) of the sample.
+Due to complete separation between the dependent variable involving sources with a few number of observations (_Los Angeles Times_ and _USA Today_), observations from these sources are dropped.
+
+The results are displayed in Table 4, below.
 
 ```{r logit analysis, include=FALSE, echo=FALSE, warning=FALSE}
+# check for separation (complete or quasi-complete)
+polaradjdataset %>% 
+  filter(Leg_Act == 1 | Crit_Insub ==1 | Endorse_Partisan == 1) %>% group_by(Source) %>% 
+  view()
 
@@
 # then, where the DV is crit_part
-
-#model 1
-
-mod1 <- glm(leg_crit_part ~ Inst_Cred_Gallup + House_distance_dim1 + Source + AuthRank + Topic, 
+# model 1
+
+mod1 <- glm(leg_crit_part ~ Inst_Cred_Gallup + House_distance_dim1 + AuthRank,
     family=binomial(link="logit"),
@@
     )
-
 summary(mod1)
@@
 
-# model 2
-mod2 <- glm(crit_part ~ Inst_Cred_Gallup + House_distance_dim1 + Source + AuthRank + Topic, 
+# model 2 - logit with different DV
+
+mod2 <- glm(crit_part ~ Inst_Cred_Gallup + House_distance_dim1 + AuthRank,
     family=binomial(link="logit"),
@@
 
-# model 3 - remove topic as an IV because there is significant correlation between
-# one type of topic (domestic policy) and the DV
-
-mod3 <- glm(leg_crit_part ~ Inst_Cred_Gallup + House_distance_dim1 + AuthRank,
+#model 3 - logit using aff polar 
+mod3 <- glm(leg_crit_part ~ Inst_Cred_Gallup + aff_polar_alt + AuthRank,
     family=binomial(link="logit"),
@@
 summary(mod3)
-tidy(mod3)
-glance(mod3)
-
-# model 4 - same as mod3, but change the DV to crit_part
-
-mod4 <- glm(crit_part ~ Inst_Cred_Gallup + House_distance_dim1 + AuthRank,
+
+#model 4 - logit using aff polar with different DV
+
+mod4 <- glm(crit_part ~ Inst_Cred_Gallup + aff_polar_alt + AuthRank,
     family=binomial(link="logit"),
@@
 summary(mod4)
-tidy(mod4)
-glance(mod4)
-
-#model 5 - now check for robustness using aff_polar as the polarization variable
-mod5 <- glm(leg_crit_part ~ Inst_Cred_Gallup + aff_polar_alt + AuthRank,
+
+# models 5-8 include Source as an IV, but the standard error estimates are too large
+# likely because there is not enough variation within Sources (LA Times is the problem)
+
+mod5 <- glm(leg_crit_part ~ Inst_Cred_Gallup + House_distance_dim1 + AuthRank + Source,
     family=binomial(link="logit"),
@@
     )
-
-summary(mod5)
-
-mod6 <- glm(crit_part ~ Inst_Cred_Gallup + aff_polar_alt + AuthRank,
+summary (mod5)
+
+mod6 <- glm(crit_part ~ Inst_Cred_Gallup + House_distance_dim1 + AuthRank + Source,
     family=binomial(link="logit"),
@@
     )
-
-summary(mod6)
+summary (mod6)
+
+mod7 <- glm(leg_crit_part ~ Inst_Cred_Gallup + aff_polar_alt + AuthRank + Source,
+    family=binomial(link="logit"),
+    data=polaradjdataset
+    )
+summary(mod7)
+
+mod8 <- glm(crit_part ~ Inst_Cred_Gallup + aff_polar_alt + AuthRank + Source,
+    family=binomial(link="logit"),
+    data=polaradjdataset
+    )
+summary(mod8)
+
+# I am now going to drop observations from the LA Times and USA Today  which will enable me to obtain some #better, more #precise estimates of the data.
+
+xtabs(leg_crit_part ~ Source, polaradjdataset)
+xtabs(crit_part ~ Source, polaradjdataset)
+slimdata <- polaradjdataset %>% filter(Source != "LA Times" & Source !="USA Today") %>% print()
+xtabs(leg_crit_part ~ Source, slimdata)
+xtabs(crit_part ~ Source, slimdata)
+
+slimdata %>% view()
+
+slimdata %>% filter(leg_crit_part ==1) %>% group_by(Author) %>% count(Author) %>% arrange(n) %>% view()
+
+slimdata %>% filter(crit_part ==1) %>% group_by(Author) %>% count(Author) %>%  arrange(n) %>% view ()
+
+# Now run a straght logit GLM including which includes Source as an IV
+# Them conduct model with robust standard errors at the author level
+# Then conduct a mixed model where we include author random effects
+
+# DV is leg_Crit_part, house_distance_dim1 is IV
+mod9 <- glm(leg_crit_part ~ Inst_Cred_Gallup + House_distance_dim1 + AuthRank + Source, family=binomial(link="logit"), data=slimdata)
+summary(mod9)
+
+mod9robust <- robustify(mod9, cluster=Author)
+summary(mod9robust)
+
+mod9mixed <- glmer(leg_crit_part ~ Inst_Cred_Gallup + House_distance_dim1 + AuthRank + Source + (1 | Author), family=binomial, data=slimdata)
+summary(mod9mixed)
+
+# DV is leg_Crit_part, aff_polar_alt is IV
+
+mod11 <- glm(leg_crit_part ~ Inst_Cred_Gallup + aff_polar_alt + AuthRank + Source, family=binomial(link="logit"), data=slimdata)
+summary(mod11)
+
+mod11robust <- robustify(mod11, cluster=Author)
+summary(mod11robust)
+
+mod11mixed <- glmer(leg_crit_part ~ Inst_Cred_Gallup + aff_polar_alt + AuthRank + Source + (1 | Author), family=binomial, data=slimdata)
+summary(mod11mixed)
+
+# DV is crit_part, IV is house_distance_dim1
+
+mod10 <- glm(crit_part ~ Inst_Cred_Gallup + House_distance_dim1 + AuthRank + Source, family=binomial(link="logit"), data=slimdata)
+summary(mod10)
+
+mod10robust <- robustify(mod10, cluster=Author)
+summary(mod10robust)
+
+mod10mixed <- glmer(crit_part ~ Inst_Cred_Gallup + House_distance_dim1 + AuthRank + Source + (1 | Author), family=binomial, data=slimdata)
+summary(mod10mixed)
+
+# DV is crit_part, IV is aff_polar_alt
+
+mod12 <- glm(crit_part ~ Inst_Cred_Gallup + aff_polar_alt + AuthRank + Source, family=binomial(link="logit"), data=slimdata)
+summary(mod12)
+
+mod12robust <- robustify(mod12, cluster=Author)
+summary(mod12robust)
+
+mod12mixed <- glmer(crit_part ~ Inst_Cred_Gallup + aff_polar_alt + AuthRank + Source + (1 | Author), family=binomial, data=slimdata)
+summary(mod12mixed)
+
 ```
 \singlespace
-```{r regression table, include=TRUE, echo=FALSE, results="asis", warning=FALSE}
-stargazer(mod3, mod4, mod5, mod6,
-          title="Table 3: Logistic Regression Results", 
-          column.labels = c('Model 3','Model 4','Model 5', 'Model 6'),
-          covariate.labels = c('Military Credibility', 'Congressional Polarization', 'Affective Polarization', 'Rank'),
+```{r table4, include=TRUE, echo=FALSE, results="asis", warning=FALSE, message=FALSE}
+stargazer(mod9, mod11robust, mod9mixed, mod12, mod10robust, mod12mixed,
+          type="latex",
+          title="Regression Results", 
+          covariate.labels = c('Mil Cred', 'Cong. Polar', 'Aff. Polar', 'Rank', 'SourceWaPo'),
+          model.names=TRUE,
+          multicolumn=FALSE,
+          column.sep.width = "-15pt",
+          dep.var.caption = "Odds (log) of Opinion Publications Authored by Retired Military Officers",
+          dep.var.labels.include=FALSE,
+          column.labels = c('Legislation, Criticize Civilians, or Partisan', 'Criticize Civilians or Partisan'),
+          column.separate=c(3, 3),
           align=TRUE)
@@
 
-The results shown in Table 3 indicate that both measurements of polarization - Congressional and affective - are statistically significant at the .01 level.
-The substantive interpretation of the coefficients for polarization, however, are a bit tricky.
-A coefficient of .083 for Congressional Polarization in Model indicates that a 1-unit increase in the average distance between political parties is associated with a 8.7 percent increase in the likelihood of a publication authored by a retired military officer calling for legislative action, blatantly criticizing a civilian official, or expressing a clearly partisan stance.
-As Figure 5 shows, though, the differences in polarization levels vary with each Congress.
+The results shown in Table 3 indicate that across five of six model specifications, the coefficient for polarization, regardless of how it is measured, is statistically significant.
+In models 1-3, where the dependent variable is an indicate for whether an opinion publications calls for legislation, criticizes civilians, or takes an expressly partisan stance, all three model specifications show that the measure of polarization plays a significant role in determining the odds of the dependent variable occurring.
+Substantively, a coefficient of .087 in Model 1 (logistic regression) indicates that a one unit increase in the mean ideological distance between political parties as measured by roll call voting is associated with a 9.1 percent increase in the odds that a publication authored by a retired military officer calls for legislative action, blatantly criticizes a civilian official, or expresses a clearly partisan stance over an issue.
+In model 2 (logistic regression with clustered standard errors at the author level), a coefficient of .029 indicates that a one unit increase in the level of affective polarization among the electorate is associated with a 2.9 percent increase in the odds of a publication calling for legislative action, criticizing a civilian official, or expressing a clearly partisan stance over an issue.
+
+While these coefficients appear to be of different size, scaling the measures of polarization are important.
 For instance, within the House of Representatives in the 112th Congress, which began in January 2011, polarization levels were nearly six points higher than the level that existed in the 111th Congress.
-The estimated effects of such a drastic change in the polarization level in Congress would be associated with more than a 50 percent increase in the predicted likelihood of a publication retired by a military officer calling for legislative action, criticizing a civilian leader, or taking a clear partisan stance.
-
-Similary, the coefficient in Model 6 for affective polarization of .033 indicates that a 1-unit increase in the level of affective polarization is associated with a 3.4 percent increase in the likelihood of a publication authored by a retired military officer blatantly criticizing a civilian leader or taking a clear partisan stance on a given issue.
-And as Figure 5 also shows, changes in the level of affective polarization over the past four decades has varied.
-However, in between elections held every four years, it is not uncommon to see changes of 5-7 points, which would indicate a corresponding 17 - 23.8 percent increase in the likelihood of a publication authored by a retired military officer criticizing a civilian official or taking a clear partisan stance.
-
-Interestingly, the coefficient on military credibility is statistically significant for Model 1 only.
-Across other model specifications, the coefficient remains negative, indicating than an increase in the credibility of the military would be associated with a reduction in the likelihood of a publication authored by a retired military officer violating norms of civilian control and non-partisanship.
-But because the coefficient is not statistically signficant in Models 4-6, we cannot draw conclusions that such a relationship exists in the data.
-
-Finally, the coefficient for rank of the author is statistically significant in Models 3 and 5, when the dependent variable includes calling for legislative action.
-In Models 2 and 4, when the dependent variable is whether a publication criticizes a civilian official or takes an expressly partisan stance, the coefficient is not statistically significant.
-This suggests that the act of calling upon Congress or the President to take specific action is clearly more associated with the rank of an author.
-This makes sense, as those authors who possess a greater rank likely have had experiences that have worked alongside or dealt with legislative policymakers, and because these authors feel more comfortable in calling for legislative action than those of a lower rank who probably lack such professional experience.
-
-## Describe the interpretation of the author, the press being willing to publish, etc. Pull from the paras below...
-
-## Describe whether you found support for your H's or not
-
-## Mixed Models and Source Random / Fixed Effects
-
-a  polarization change However, is statistically significant across a graph clearly shows a non-linear relationship between increasing political polarization and the likelihood that an opinion piece authored by a retired military officer will potentially violate aspects of military professionalism.
-While the confidence intervals at each point estimate are relatively large, the graph demonstrates that the predicted probability of violating military professionalism is different between relative extremes along the x axis.
-In other words, the political polarization levels that exist today, as measured by the ideological distance between members of the two prominent political parties in the house, is predicted to have resoundingly different impacts on military professionalism than the polarization levels that existed several decades ago.
-The polarization levels that existed several decades ago, in the late 1970s and 1980s, may have had negligible, if any impact, on the likelihood of a retired military officer violating military professionalism in the content of a published opinion piece, but the political landscape is simply different today.
-Contemporary polarization levels, insofar as the predictive power of the model is concerned, changed the environment: at contemporary polarization levels (the far right of the x axis), there is roughly a 17% chance of a retired military officer's published opinion piece violating the standards and norms of military professionalism, with confidence intervals that range from roughly 11% to nearly 25%.
-
-Given these dynamics, as well as the results of both Figures 3 and 5, ample evidence exists to confirm the second hypothesis, which stated that as a result of increasing political polarization, there will be an increase in recent years in the number of op-eds written about topics that are not relevant to military expertise, or that reveal inappropriate partisan sentiments, among authors which are a vital component of the institutional military, retired military officers.
-
-## Does Combat Attenuate Polarization's Impact on Military Professionalism?
-
-Why did Figure 3 reveal a sharp drop between approximately 2005-2011 in both the real number and proportion of annual opinion pieces authored by retired military officers that address non-traditional military topics, such as social policy that impacts the military, domestic policy, and political administrations/figures?
-What was it about this time period that made it unlike any other span of years across the past four decades?
-
-One compelling answer is that between these years, roughly 2005-2011, the military was heavily involved in combat operations to a point such that the institution was taxed tremendously, thereby necessitating that it remain focused on its wartime mission.
-Repeated deployments during the Global War on Terrorism, which included substantial surges of forces to both Iraq (2007-2008) and Afghanistan (2010-2011), as well as sustained US casualties resulting from these conflicts, possibly shaped the military institution so strongly such that even retired military officers acknowledged that the military's institutional energy required a laser-sharp focus on combat requirements.
-This is not to suggest the military was not heavily engaged in conflicts nor active around the glove at other times throughout the past four decades (which we know is not the case), but to instead suggest that even retired military officers are perhaps more sensitive to combat requirements and missions undertaken by the military as higher levels of continued operational strain and duress are endured.
-Though more data is required to fully test this argument, other scholars have argued similarly in that periods of transition, especially away from major combat operations and into eras of downsizing and relative peace, often seem to test the very essence of military professions in significant ways [@snider_once_2012].
-
-# Conclusion: Will the Military Remain Professional in a Polarized Setting?
-
-The goal of this paper was to explore and examine the ways in which political polarization and institutional credibility have impacted the professionalism of the military over the past fifty years.
-Examining the published opinion pieces authored by retired military officers in major US newspapers revealed that although the vast number of total opinion publications uphold the military's norms and traditions concerning military professionalism, a couple of worrisome trends are evident in the data: first, retired military officers are increasingly writing about topics outside of the umbrella of traditional national security topics, and second, retired officers are expressing partisan preferences at a greater rate than they previously did.
-What do these trends mean, and should they cause concern?
-If so, what should be done about it?
-
-First, these findings illuminate the power and pervasiveness of increasing political polarization, and in particular, its impact not only on political institutions, but also on professions, which the military prides itself on being.
-Thus, if the military, which has for decades adopted and promulgated norms of apolitical and nonpartisan expression, has been impacted by rising political polarization, it stands to reason that other political institutions and professions also have been impacted.
-While the specific manifestations of political polarization's increase and spread in non-military institutions and professions might be different, the fact that the military is experiencing a noticeable impact as a result of increasing political polarization speaks to its significant power and influence.
-
-Indeed, evident in the content of opinion pieces authored by retired military officers over the past four decades are the three features of contemporary American political polarization developed in the theory proffered earlier in this paper: the loss of politically neutral space as more issues enter the partisan battleground, the creation of strong feelings of dislike for those who hold to opposing views on these issues, and perhaps worst for the military, alterations in an individual's concept of loyalty, which arises as the salience of one's identity is raised as the issues that an individual holds most dearly to are increasingly at stake.
-Moreover, the data supported what most scholars of American politics have argued for: a realization that polarizing politics across America's political landscape will likely not disappear anytime soon, regardless of which party holds the nation's highest office.
-
-Second, these findings underscore the challenge that the military as a profession has in "reigning in" its retired officers and compelling them to stay out of partisan battles, and out of politics more broadly.
-Among the features and traits of a profession is that it is self-policing; however, the data indicates that current efforts by active military leaders to police and clean up the behavior of its retired elites is not working.
-Indeed, the letters and pleas of previous military leaders to retired military officers, including those written by now Retired Army General Martin Dempsey and Retired Marine Corps General Joseph Dunford, both of which were referenced in this paper, are simply not working.
-This suggests that stronger and more creative efforts are required.
-
-Third, these findings demonstrate that the phrase "politicization of the military," which has long been a proper concern of academics, journalists, the military, and citizens alike, should not be intellectually confined to notions of how politicians or other political leaders political leaders attempt to harness or exploit the military for political purposes or advantages.
-Though such efforts do occur and are a genuine cause for concern that continue to deserve both scholarly and practical attention, the evidence presented in this paper shows that "politicization of the military" is also occurring at the hands and through the words of particular retired military officers.
-Gaining a better understanding of the impacts of political polarization on the military profession remains a critical area for future research.
-Other scholars such as Michael Robinson have explored the specific ways that polarization is impacting the interactions between the military and the public [@robinson_michael_danger_2018].
-Though this paper focused on the ways that polarization may impact professional military behavior, it recognized but stopped short of discussing polarization's subsequent impact on various "military outcomes," which may arguably include even more decisive events, such as achieving victory or defeat in battle, or the ability to formulate sound strategic advice.
-These outcomes and more, and the role of polarization in influencing them, deserve to be explored further.
-
-Maintaining credibility and trust with the American public is crucial for the military, but should a polarized political climate endure, the military needs to also understand that credibility and trust can be abused to the point that they also can constitute significant risks.
-One such risk is that as polarization worsens, military leaders, to include military retirees, abuse the high level of credibility they currently enjoy with the American public to advocate for obviously partisan positions.
-Such actions, though, in a polarized setting, might actually result in an increase in the credibility of the military institution, at least among citizens who share the same partisan sentiments advocated for by military leaders, which might, somewhat perversely, entice more of the same behavior.
-A second risk, stemming from the first, is that a polarized setting sets the stage for the military institution to eventually openly jockey with civilian leaders, either for power or for control, a concern that generations of scholars have warned us about [@finer_man_1962; @brooks_shaping_2008-1].
-Were either of these risks to materialize, the character and nature of the American military as an institution, and likely, American democracy, would be irrevocably changed.
-
-There are no simple solutions for the military with respect to the risks posed by rising political polarization.
-Two possible ideas, however, are apropos to this discussion.
-One involves identity, and the other, credibility.
-First, our nation's military is composed of soldiers who come from every walk of life throughout America, and accordingly, in the years to come, many future soldiers and leaders will bear the features of a polarized society.
-Thus, the military's challenge, and one which it must embrace itself for immediately, is integrating the multiple identities each soldier carries with them as people in such a way that each can still proudly wear the uniform and bear the identity of an American service member.
+Such a drastic change in the polarization level within Congress would roughly predict a 54.6 percent increase in the odds of a publication retired by a military officer calling for legislative action, criticizing a civilian leader, or taking a clear partisan stance.
+Regarding changes in the levels of affective polarization, data from the 2012 ANES indicate that the aggregate level of affective polarization increased by just over 16 points from the previous Presidential Election year (2008), which would correspond to a 46.4 percent increase in the odds of a publication authored by a retired military officer calling for legislative action, criticizing a civilian official, or taking a clear partisan stance.
+Thus, the coefficients for the respective levels of polarization are substantively similar.
+Even in Model 3 (Mixed Model with author random effects), the coefficient for polarization increases from its estimate in Model 1 to a value of .091, which corresponds to a 9.5 percent increase in the odds of a publication calling for legislative action, criticizing a civilian official, or taking a clear partisan stance for every one unit increase in the mean ideological distance between political parties as measured by roll call voting in Congress.
+Because the coefficient in Model 3 is still statistically significant even after introducing author random effects, we would interpret this as signifying that the individual author characteristics do not substantively weaken the impact of polarization.
+
+When the dependent variable no longer includes whether a publication calls for distinct legislative action to be taken, and thus, indicates either criticism of a civilian official or taking a clearly partisan stance, the models (models 4-6) tell a very similar story with the exception of the mixed model (model 6).
+Compared to the estimates of the coefficients in Models 1 and 2, the coefficients in Models 4 and 5 are relatively larger.
+A coefficient for polarization in Model 5 of .100 which corresponds to a 10.5 percent increase in the odds of a publication criticizing a civilian official or taking a clear partisan position with every one unit increase as measured by Congressional roll call voting.
+A coefficient of .037 in Model 4 corresponds to a 3.7 percent increase in the odds of a publication criticizing a civilian official or taking a clear partisan position with every one unit increase in the level of affective polarization among the electorate.
+
+In Model 6, which includes author random effects, the coefficient for affective polarization is no longer statistically significant, and thus, we cannot rule out the possibility that individual author characteristics are driving changes in the dependent variable, as opposed to the level of affective polarization itself.
+Such an interpretation also suggests that not every individual author reacts in the same way to increasing levels of polarization - which actually confirms much of the theoretical literature about polarization.
+Still, the fact that coefficients measuring polarization in two ways are significant in five of six models confirms the role of polarization in presenting publications authored by retired military officers that deviate from the norms of civilian control of the military and non-partisanship.
+If we conceptualize the commentary proffered by retired military officers in major newspapers as a significant portion of the overall messaging from the military community, even if it is offered by a relatively small number of total retired military officers, the impact of polarization on adherence to military norms is striking.
+
+In terms of the other coefficients besides measurements of polarization, three points are striking.
+The first is that coefficients for the estimated impacts of military credibility are statistically significant in only two of six models, but that in all six models, the sign of the coefficient is negative.
+Relative to the consistent statistical significance of the coefficients for measurements of polarization, we conclude that changes in military credibility have less explanatory power, if any, in whether a publication authored by retired military officers fails to adhere to norms of civilian control and non-partisanship.
+The second interesting point is that the author's rank is significant in also just two of six models, and so we cannot conclude that author rank explains significant changes in our dependent variable of interest.
+The final point is that across all model specifications, not a single coefficient for source is statistically significant.
+This suggests that even if all of the three examined sources for the regression (Washington Post, Wall Street Journal, and New York Times) have changed over time as a result of polarization in terms of the types of content each source is willing to publish, changes in an opinion or commentary publication authored by a retired military officer breaking norms of civilian control or non-partisanship have not depended on which particular source the author publishes in.
+Said differently, regardless of the specific newspaper source, the changes in commentary or opinion publications authored by military retirees have been driven more by changes in polarization than any other variable controlled for across the different models.
+
+We can also use the regression results from Table 3 to transform the results of our logistic regressions into predicted probabilities.
+Using the results of the logistic regression in Model 4, which included a measure for affective polarization, we set the source as the _New York Times_ and hold values of institutional credibility of the military and author rank at their respective mean values of 71.1 and 8.39.
+These predicted probabilities are graphed in Figure 6, below.
+
+```{r predicted probabilities, include=FALSE, warning=FALSE, message=FALSE, error=FALSE}
+# Use Model 4 (Logit) from Above, where DV is crit_part and there are no Source Fixed Effects
+# Hold values of Author Rank and Institutional Credibility at their mean
+# Vary the Level of Affective Polarization from its minimum value to an arbitrary 
+# level of 120, which is higher than it currently is but not theoretically impossible.
+
+predictiondata <- tibble(aff_polar_alt=seq(min(slimdata$aff_polar_alt), 100, by=1), Inst_Cred_Gallup = mean(slimdata$Inst_Cred_Gallup), Source="NYT", AuthRank=mean(slimdata$AuthRank)) %>% crossing(crit_part = c(0,1))
+
+predictiondata
+augment(mod12, newdata = predictiondata)
+
+# prediction data
+predictions <- augment(mod12, newdata=predictiondata) %>% 
+  mutate(
+    .prob=plogis(.fitted),
+    conf.low = plogis(.fitted -(qnorm(.975) * .se.fit)),
+    conf.high = plogis(.fitted + (qnorm(.975) * .se.fit)),
+    pred_final_rec = case_when(
+      .prob >= .5 ~ "Likely",
+      .prob < .5 ~ " Not likely"
+    )
+  )
+
+predictions %>% arrange(.prob, .fitted, crit_part, desc(.fitted)) %>% 
+  view()
+
+# now let's try some graphs
+# notice that the predictions are not in (0, 1)
+ggplot(predictions) +
+  aes(x = .fitted) +
+  geom_histogram()
+
+# if we wanted predictions on the probability scale,
+#   we predict on the log scale,
+#   then apply inverse-logit (logistic CDF)
+# use plogis(): cumulative probability of the logistic distribution
+
+# demonstration: scatter log odds (x) and predicted probability (y)
+# with 95% CIs
+# (CIs don't look smooth in this example because .se.fit depends on X data)
+predictions %>%
+  ggplot() +
+  aes(x = .fitted, y = .prob) +
+  geom_pointrange(aes(ymin = conf.low, ymax = conf.high)) +
+  geom_smooth()
+
+```
+
+```{r plot6, include=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.cap="Predicted Probability of Polarization\n on Retired Military Elite Opinion Publications"}
+predictions%>%
+  ggplot() +
+  aes(x = aff_polar_alt, y = .prob) +
+  geom_pointrange(aes(ymin = conf.low, ymax = conf.high)) +
+  geom_smooth() + labs(
+    x="Level of Affective Polarization\n Among the US Electorate",
+    y="Predicted Probability of an Op-Ed Published in NYT\n by Retired Military Elites which Criticizes Civilians\n or Takes a Clear Partisan Stance"
+  ) + scale_y_continuous(breaks=seq(-.1, .8, .1)) + 
+  geom_vline(xintercept = 48.81, size=.2, colour="tomato") + 
+  geom_vline(xintercept=59.62, size=.2, colour="tomato") + 
+  geom_vline(xintercept=68.001, size=.2, colour="dodgerblue") + 
+  geom_vline(xintercept=84.4, size=.2, colour="dodgerblue") +
+  geom_vline(xintercept=79.55, size=.2, colour="tomato") +
+  geom_text(aes(x=48.81, label="Reagan, 1980 ANES",y=.3), colour="tomato", angle=90, vjust=1, text=element_text(size=9)) + 
+  geom_text(aes(x=59.62, label="Bush 43, 2000 ANES",y=.4), colour="tomato", angle=90, vjust=1, text=element_text(size=9)) + 
+  geom_text(aes(x=68.001, label="Obama, 2008 ANES",y=.4), colour="dodgerblue", angle=90, vjust=1, text=element_text(size=9)) + 
+  geom_text(aes(x=84.4, label="Obama, 2012 ANES",y=.52), colour="dodgerblue", angle=90, vjust=1, text=element_text(size=9)) + 
+  geom_text(aes(x=79.55, label="Trump, 2016 ANES",y=.50), colour="tomato", angle=90, vjust=1, text=element_text(size=9))  
+```
+
+As the x axis of Figure 6 reveals, the highest levels of affective polarization within the electorate were recorded during the ANES administered in 2012, in the months leading up to the start of President Obama's second term.
+In fact, the level of affective polarization actually dropped as measured by the ANES fielded in 2016, as President Trump won the Presidency.
+Figure 6 shows a clear increase in the predicted point estimate of the probability of a publication authored by a retired military officer failing to adhere to the norms of civilian control of the military and/or non-partisanship as the level of affective polarization among the electorate varies.
+Using this model shows, in fact, that at ANES levels that existed at the time of the election of President Reagan in 1980, the predicted probability of an opinion publication in _The New York Times_ authored by a retired military officer failing to adhere norms of professionalism or non-partisanship was roughly 8.0 percent.
+At ANES levels that existed when President Trump was elected, that probability increased to 21.7 percent.
+It remains to be seen how much worse polarization will become, but the predicted impacts of any increase in polarization do not bode well for the future of military professionalism.
+If we project that the nation is now back at the level of affective polarization as that which existed as measured at the time of President Obama being elected to a second term, and should the same level of change in affective polarization occur that took place between President Obama's first and second terms (roughly 16 points in the level of affective polarization), the model would predict a nearly 37% probability that an opinion publication published in the the _New York Times_ and authored by a retired military officer would violate norms of civilian control or non-partisanship.
+
+In terms of the four hypotheses this paper set out to explore and examine, the data seems to tell a convincing story about the relative impact of increasing political polarization on the propensity of retired military officers adhering to the norms of civilian control of the military and non-partisanship, and particularly whether these norms are followed in the opinion publications retired officers seek to publish in major US newspapers.
+The regression analysis indicated that in fact, the level of political polarization is a statistically significant indicator of the likelihood of retired military elites criticizing civilian officials or making clearly partisan statements, thus lending confirmation of _H1_.
+Additionally, the descriptive statistics, and particularly the trend in Figure 3, indicated that increasing levels of political polarization are associated with military retirees addressing domestic political topics, thus providing support for _H2_.
+Figure 3 also revealed in a general sense that the frequency with which retired military elites write commentary or opinion pieces in major US newspapers is associated with an increase in the level of prestige and credibility the military institution enjoys among American society.
+Finally, the data indicated that periods of active ground combat tend to focus retired military elites to write on topics related to combat.
+This finding is supported by Figures 1, 3 and 4, each of which revealed a clear local minimum in the number of publications authored by retired military officers (Figure 1), the proportion of publications that addressed domestic topics (Figure 3), and the proportion of publications calling for legislative action, criticizing civilian leaders, or making boldly partisan claims.
+These local minima occurred during the periods of surge operations in Iraq and/or Afghanistan during the late 2000s through early 2010s, when the US military had been engaged overseas for several years with frequent deployments, enduring a steady frequency of American casualties.
+This confirms the broad insight previous scholars have advanced regarding how the demands of combat and war-fighting tend to focus certain professional behaviors of the military [@snider_once_2012].
+
+# Can the Military Remain Professional?
+
+This study demonstrated that although the vast number of opinion publications authored by retired military officer uphold the military's norms of civilian control and non-partisanship, worrisome trends are evident that suggest retired military elites are increasingly failing to uphold the central norms of civilian control of the military and the non-partisanship of the military institution in a significant portion of the messaging that they conduct with the American public: deliberately written opinion and/or commentary pieces which retired military officers author in major US newspapers.
+The data clearly reveal that the manner in which retired military officers address the American public has changed with respect to upholding military norms and standards over the past four decades, even while holding the source of a newspaper constant.
+These trends are alarming for at least three reasons, and speak not only to the degree of seriousness of the problem, but the role that political polarization has played in driving certain members of the retired military officer community to engage in this behavior.
+
+A primary reason that these trends are alarming stems from the influence that even a small number of retired military elites can potentially have.
+While one can argue in theory that Americans should not be concerned if a handful or even several dozen retired military officers break certain norms of professionalism on the basis that there are more than 7,000 living retired general and flag officers, such a figure should not serve as a baseline comparison when analyzing the particular remarks of retired military officers that are published in prominent US newspapers.
+Only a small number of pieces are chosen to be published by the editorial boards of major newspapers whose circulations range in the tens of millions, and it is important that this small number of retired military officers showcase to the American public if not a unifying message, at least a unifying adherence to the fundamental norms of civilian control of the military and non-partisanship that all members of the military institution are expected to follow.
+
+A second reason why these trends are concerning, I argue, is because members of the active duty force still look, in many cases, to these retired officers who are violating norms of professionalism as role models and figures whose behavior is worthy of emulation.
+This is not to suggest that every single instance of a retired military officer breaching a professional norm has a major impact on the military profession, but it is to suggest that at some point, many instances of behavior constitute a pattern.
+The evidence presented in this paper, I argue, presents clear patterns of failing to uphold foundational norms of professionalism among the retired military community.
+At this point, considering that several active duty military generals have previously attempted to "reign in" or at least plead with members of the retired military community to limit behaviors which would tarnish the image of a professional military, it is doubtful that anything less than numerous and urgent pleas from fellow retired military elites and current four star generals and admirals would work to reign some of the most blatant offenders of professional norms.
+Even then, however, the outcome is likely uncertain whether such actions would attenuate these sorts of behaviors.
+
+A third and final reason that the data presented in this paper is concerning is that amid all of the cries from pundits and scholars alike of civilian leaders acting to "politicize the military," a fair bit of politicization of the military is occurring via the written words of retired military officers.
+For an institution such as the military that espouses to be a profession that looks inward and polices the behavior of those who stray from professional norms, it is time that retired military officers - and especially, the community of retired flag and general officers - look inward, and ask both themselves and their peers, "are we upholding professional norms in everything that we do and say?" If the answer to this question is "no," these retired military officers should put partisan preferences aside and commit themselves to solving the problem.
+In an age marked by increasing polarization and divisiveness, however, this may just require retired military officers to perhaps do the hardest things of all: put the pen down and hold one's tongue.
+
+One of the long-term challenges facing the active military in an age of political polarization is integrating the multiple identities each soldier carries with them in such a way that each can still proudly wear the uniform and bear the identity of an American service member.
 Though this challenge is not new for the US military, increasing political polarization has significantly raised the level of difficulty the military faces in tackling such a vital challenge.
-Scholars who have studied multi-ethnic militaries across the world have argued that it is military leaders, not civilians, who ultimately devise effective solutions to the problems posed by soldiers who carry multiple loyalties, and interestingly, only as a result of combat requirements [@peled_question_1998].
-Perhaps the American challenge is even greater in that at least for the near term future, there is no major war that will necessitate the coming together of disparate loyalties and identities among future soldiers.
-
-The second idea, and one that may sound like anathema to modern military leaders, is that perhaps the level of trust and credibility between the American public and the military institution should not be the ultimate indicator of the military's performance as a profession.
-In suggesting this, I am not suggesting that credibility and trust are not important or that the concept should be disregarded entirely.
-The American public is, after all, the client of the American military and ultimately, it is their opinion that counts.I am suggesting, however, that for the reasons laid out in this paper, credibility and trust are influenced by the mechanisms through which polarization works, to include the loss of politically-neutral space, increased negative feelings toward those who hold differing opinions, and most sinister of all, alterations in one's concept of loyalty.
-For these reasons, I am simply offering the notion that _credibility can be deceiving_ in polarized times.
-Now, more than ever before, military leaders will need to rely on individual moral courage over and above any other trait in order to perform their duties as military professionals.
+Scholars who have studied multi-ethnic militaries across the world have argued that it is military leaders, not civilians, who ultimately devise the most effective solutions to the problems posed by soldiers who carry multiple loyalties, and interestingly, only as a result of combat requirements [@peled_question_1998].
+Perhaps the American challenge is even greater in that at least for the near term future, there is no major war on the immediate horizon that will bring about the coming together of disparate loyalties and identities among future soldiers.
+While the knowledge, skills, and experiences of retired military officers are important, and can and should be used to help solve myriad problems facing the country, the impacts of political polarization are threatening to undermine the commitment of certain members of the military, to include military retirees, to professional norms.
+Without addressing and correcting the trend of retired military officers violating professional norms swiftly, there could be significant repercussions on the military's ability to gain and maintain public trust, as well as impacts on how future generations of military officers conceive of their normative role in politics.
 \singlespace
@@
 
-#Appendix 1
--sort out the table with Kable and make it look good-
-***Move the Actual Table of these observations to the Appendix, and then drive on with the Regression.  Good work!***
+# Appendix 1
+
+\doublespace
+## Word and Data-Base Searches to Obtain Data
+
+I did two different sets of searches.
+This first set of search terms was applied to the Factiva and Nexus-Uni data bases.
+In each of these databases, I had to select the "type of article" and included options such as commentary/opinion and letters to the editor.
+The specific terms of search included:
+
+"retired Army General" or "former Army General" or "retired Navy Admiral" or "former Navy Admiral" or "retired Marine Corps General" or "former Marine Corps General" or "former Air Force General" or "retired Air Force General" or "retired Coast Guard Admiral" or "former Coast Guard Admiral" or "retired from the Marine Corps" or "retired from the Army" or "retired from the Navy" or "retired from the Air Force" or "retired from the Coast Guard" or "supreme commander of" or "commander of" or "Chairman of the Joint Chiefs" or "chairman of the joint chiefs" or "Supreme Commander of" or "Chief of Staff" or "Chief of Naval Operations" or "retired Army" or "retired Navy" or "retired Air Force" or "retired Marine Corps" or "retired Coast Guard" or "retired military" or "commandant of" or "director of" or "retired from the military"
+
+The second set of search terms was applied to Pro Quest Historical Newspapers, which the University of Wisconsin had access to for both the Wall Street Journal and the New York Times.
+Factiva included article from the Wall Street Journal only back to the middle of 1994, so I relied heavily on Pro Quest to search through the Wall Street Journal from 1979-1994.
+Unlike the search terms used for Factiva and Nexus-Uni, however, I came to quickly realize that the opinion pieces I sought were not categorized as opinion-editorials or letters to the editor, but rather as "articles."  This required significantly more time searching for the desired data, as more articles had to be searched through and sifted.
+The specific terms of search included:
+
+(retired OR former OR “chief of staff”) AND (“Army General” OR Gen.
+OR general OR "Navy Admiral" OR Adm.
+OR admiral OR "Marine Corps General" OR "Coast Guard Admiral" OR "Air Force General" OR "Chairman of the Joint Chiefs" OR "colonel" OR "captain" OR "commander" OR "chief of naval operations" OR "commandant" OR "director of")
+
+## Observations used for Regression Analysis
+
+The following table contains those observations whose content either called for legislative action, criticized civilian leaders, or took clearly partisan stances.
 
 \singlespace
-```{r table3, fig.align = "center", include=TRUE, echo=FALSE}
-#Need to adjust the width and spacing of this table
-quest_prof %>%
-  kable(format="latex", caption="Op-Eds authored by Retired Military Officers that Strongly Criticized Civilian Leaders, or were Expressly Partisan, 1979-2020", longtable=T) %>%
-  kable_styling(font_size=10, latex_options="striped", repeat_header_continued=T) %>% 
-  column_spec(1, width="9.5em") %>% 
-  column_spec(2, width="15em") %>% 
-  column_spec(3:4, width="2.5em") %>%
-  column_spec(5, width="5em") %>%
-  column_spec(6, width="6em") %>%
-  row_spec(0, bold=T)
-```
-
-\doublespace
+
+```{r table5, include=TRUE, echo=FALSE}
+TA1
+```
 
@@
 All but four of the observations in this cluster are coded as having potentially violated standards of military professionalism.
-Of note, one observation in this cluster entitled, "Fists Raised at West Point" and published in the Washington Post is actually a compendium of two anonymous letters, each one taking an opposite stance on the appropriateness of an incident in which a photograph at West Point's 2016 graduation ceremony captured a graduating female cadet raising a clenched a fist, an action which some saw as an inappropriate political gesture related to the growing Black Lives Matter movement [@unknown_fists_2016].
-
-\newpage
+Of note, one observation in this cluster entitled, "Fists Raised at West Point" and published in the Washington Post is actually a compendium of two anonymous letters, each one taking an opposite stance on the appropriateness of an incident in which a photograph at West Point's 2016 graduation ceremony captured a graduating female cadet raising a clenched a fist, an action which some saw as an inappropriate political gesture related to the growing Black Lives Matter movement [@anonymous_author_fists_2016].
+
+\singlespace
+\pagebreak

### 2021-02-05 code/01-intro.Rmd
commit

---
title: "Servants, Syndicates, and Saviors: How Polarization and Prestige Impact the Political Behavior of the Military (Prospectus)"
author: "Peter M. Erickson"
date: |
  `r format(Sys.time(), '%B %d, %Y')`
abstract: |

  _This dissertation explores variation in the political behavior of the US military. I advance a theory in which the political behavior of the US military is impacted by two distinct factors: the level of domestic political polarization within a society, and the degree of prestige the military institution commands. I argue that these two factors greatly shape the constraining influence of three central norms of civil-military relations: civilian control of the military, the non-partisanship of the military institution, and the degree to which the military fills an active role in domestic politics. As the constraining influence of these norms varies, distinct political behaviors conducted by or involving the military also change. To support my theoretical argument, I conduct a mixed-methods empirical strategy focusing on military behavior within the United States across time. This dissertation advances the scholarly understanding of US civil-military relations, and provides a unique perspective about how domestic structure influences the ability of the military to act professionally._

geometry: margin = 1.15in              # margin widths
fontsize: 12pt                         # text size
bibliography: researchpaper-ps904.bib # bibliography file
biblio-style: authoryear               # citation style
urlcolor: blue                         # links to external URLs
citecolor: black                       # links to bibliography
linkcolor: magenta                     # internal figures, footnotes, etc.
indent: true                          # paragraph format
                                       # true for indented paragraphs

header-includes:
    - \usepackage{setspace}
    - \usepackage{multirow}
    - \usepackage{indentfirst}

output: 
  bookdown::pdf_document2:
    keep_tex: true
    latex_engine: pdflatex
    number_sections: false
    toc: false
    fig_caption: true
    citation_package: biblatex
    includes: 
      in_header: 
        - latex-preamble.tex
  bookdown::htmldocument2: default
  bookdown::word_document2:
      toc: false
---

\doublespacing

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo=FALSE,
  include=FALSE,
  message=FALSE,
  error=FALSE,
  fig.height=3,
  fig.width=4,
  out.width="90%")
```

```{r, include = FALSE, echo = FALSE}
library("here")       # file paths
library("tidyverse")  # workhorse package
library("tidylog")
library("kableExtra")
library("knitr")
library("ggdag")
library("dagitty")
library("gridExtra")
library("broom")
library("dplyr")
library("reshape2")
library("pacman")
library("ggrepel")
```

\newpage

# Introduction and Research Question

Scholars, civilians, diplomats, and military leaders regularly discuss why and how militaries intervene in politics. These discussions often invoke the theoretical question of whether military actors _should_ intervene in politics, and for what purpose. Especially in the United States context, but certainly not exclusively, the political activity of the military is a subject that has undergone intense scrutiny in recent years.

In the United States, increasing political polarization has thrust the contemporary military institution into the political spotlight on numerous occasions. From a now-retired Army Lieutenant Colonel testifying against President Donald Trump during impeachment proceedings before Congress, to a Navy Captain's removal for leaking a letter to the press in which he castigated the chain of command's response to the onset of the COVID-19 pandemic, to a remarkable panoply of retired military officers blasting President Donald Trump for threatening to use active duty forces to dispel protesters in cities across America during the summer of 2020, instances in recent years abound which illustrate the domestic challenges militaries face during polarized political periods. While numerous scholarly works describe important facets of the political behavior of the military, few if any have systematically set out to explore the impact of structural domestic factors on distinct forms of political behaviors either conducted by or involving the military. This dissertation seeks to fill this important gap.

The central argument made in this dissertation is that the domestic political behavior of the US military is a function of two main factors: the level of polarization in a society, and the overall level of prestige of the military institution. Building upon a general "motives" and "opportunity" framework to describe military intervention in politics first advanced by Finer (1962) and later by Taylor (2003), I argue that in the American context, the relative levels of polarization and prestige, across time, are the variables that have provided the same functions for political involvement by the US military [@finer_man_1962; @taylor_politics_2003]. Moreover, the presence or absence of these two factors strongly influence _how_ militaries and military actors intervene in the political space. Further, this dissertation sheds light on what is, I argue, an under-appreciated task: the changing difficulty that military leaders face in harnessing, managing, and orchestrating the domestic behavior of the military as an institution. Far from being the hierarchical and monolithic actor most people think that it is, even the US military, at times, can be divided regarding its normative role in politics.

Examining the patterns of political behavior by the military is an important undertaking. In a normative sense, it is important because despite the normative tradition of militaries staying above or outside of partisan politics, the reality is that professional military actors can and often do create political costs that civilian leaders have to contend with. These costs include constraining the choices of civilian leaders, leveraging public opinion against elected officials, and otherwise stultifying the ability of elected officials to implement their intended foreign _and_ domestic policy choices. On the other hand, there are also potential benefits to some instances of political behavior by military actors. As Risa Brooks has argued, military actors possess tremendous informational and expertise advantages compared to the public, and thus, hearing from military actors can increase the public's awareness of an issue, improve public debate over the nation's defense policies, and spur civilian leaders to think through alternate courses of action [@brooks_militaries_2009, 233-236]. 
Exploring military behavior in politics is also an important empirical undertaking. If there are patterns of military behavior that can be detected over time, and if it is possible to better understand the variables that drive these patterns, scholars, civilian leaders, and military officers will have a better appreciation for the difficulties that particular domestic structures have on the ability to maintain norms of civil-military relations. More importantly, having an increased awareness for these challenges might also cause actors - whether they be from the military, within government, or from elsewhere within American public - to reconsider following through on a particular course of action in the short term that might harm civil-military and democratic norms in the long run. While this particular hope may be wishful thinking, deriving and testing a theory that explains military intervention in politics in the United States nonetheless helps us explore a pivotal issue, and one that is likely to remain relevant well into the future.

This introductory chapter of the dissertation proceeds in three main parts. First, this introduction introduces the general field of civil-military relations. In doing so, emphasis is given to underscore three central norms of American civil-military relations: civilian control of the military instrument, non-partisanship of the military institution, and non-interference of the military in the space and roles of partisan political actors. Second, the introduction describes where the intended dissertation topic fits within the scholarly literature as well as the gap it intends to fill. The third and final part of this introductory chapter describes the methodology and the plan of the dissertation.

# The Field of Civil-Military Relations

Truly one of the interdisciplinary topics of scholarly research, civil-military relations spans the fields of political science, history, and sociology. Questions involving the set of relationships that exist and should exist between militaries and governments, the relationships between militaries and societies, strategic and military effectiveness, the occurrence of coups, and a host of other pivotal topics comprise the study of civil-military relations. Within the field of political science in particular, numerous scholars of civil-military relations frequently study the relationship between prominent state officials and military leaders. Indeed, this "leadership at the state's apex" is responsible for making and implementing numerous weighty decisions that impact national security, including whether to go to war, how to form and implement effective strategy, how to prepare for the next conflict, and how large the military and its budget should be [@brooks_risa_integrating_2019, p. 380].

In the most basic sense, the study of civil-military relations examines the question of how to ensure that a military is strong enough to defeat threats to the state without threatening the state itself. Peter Feaver refers to this central question of civil-military relations as the "problematique," and notes that this dilemma is both "vexing" and a great "paradox" in that the military must be "strong enough to prevail in war," yet submissive to authority such that it cannot and will not "destroy or prey on the society it is intended to protect" [@feaver_civil-military_1996; @feaver_civil-military_1999, p. 214]. With this understanding in mind, it is clear that civil-military relations are fraught with a natural tension that, while perhaps capable of attenuation, is never entirely absent.

A number of scholars have offered solutions for solving the "problematique" Feaver describes, but none has been more lasting or consequential than the late Samuel Huntington, whose 1957 work, _The Soldier and the State: The Theory and Politics of Civil-Military Relations_, often serves as the point of departure for a number of debates within the field of civil-military relations. Huntington's theory has endured important and substantive critiques over the years [for example, see @janowitz_professional_1960; @finer_man_1962; @cohen_supreme_2003; @feaver_armed_2003; @brooks_paradoxes_2020]. Nonetheless, it is essential to understand Huntington's conception of effective civil-military relations, and what lies at the heart of it - a theory of civilian control of the military that maximizes what Huntington called "military professionalism" [@huntington_soldier_1957].

In the section to follow, I will briefly summarize Huntington's argument and logic, as well as summarize several prominent critiques made by other scholars of Huntington. The purpose of summarizing both Huntington and his critics is to argue that in spite of the important differences between them, Huntington and his critics share common ground regarding the prominence of three central norms that exist to normatively govern both civilian and military behavior.  These norms, which I articulate and describe in the section to follow, include the principles of civilian control of the military, the non-partisanship of the military institution, and what I will term the "non-interference" of the military into realms of politics that are the _perogative_ of duly-elected and appointed civilian leaders. While these principles are related to each other, each is distinct and holds different normative implications for both civilian and military behavior.

## Huntington, His Critics, and Fundamental Norms of Civil-Military Relations

In _The Soldier and the State_, the problem that Huntington is trying to solve is how can liberal democracies such as the United States balance the simultaneous demands of military security and the ideology of political liberalism. Writing after the Korean War, Huntington argues that American security requirements constitute a "functional imperative" that required the US to maintain a strong military capable of defeating the external threats of the Cold War [@huntington_soldier_1957, 1-3]. At the same time, Huntington argues that the liberal character of American society generated a "societal imperative" such that the pursuit of American security requirements could not cause America to deviate from its liberal political ideology in fundamental ways, including abandoning the principle of civilian control of the military [@huntington_soldier_1957, 1-3]. In essence, Huntington aims to solve how to render the military strong enough to prevail during the Cold War, but sufficiently subservient and obedient to civilian power.

Huntington argues that this challenge is made more acute by the fact that the military officer corps is different from the rest of America in several important ways. He claims that those who serve in the military hold to a distinct "mind" and "ethic" that views conflict as part of a "universal pattern throughout nature," and thus, that "the military view of man is...decidedly pessimistic" [@huntington_soldier_1957, 62-63]. On the other hand, Huntington describes American liberal ideology as stressing "individualism," and claims that liberalism "either denies the existence of power, minimizes its importance, or castigates it as inherently evil" [@huntington_soldier_1957, 90]. Huntington adds that this juxtaposition between the unique character and ethic of militaries and the attributes of political liberalism in America explains why America has been suspicious of standing military forces since its founding [@huntington_soldier_1957, 143-151]. However, the Cold War poses a unique problem in terms of the security requirements facing America. Standing, sizable military forces are a must.

The solution Huntington offers is for America to adopt a system of civil-military relations which Huntington terms "objective control" of the military [@huntington_soldier_1957, 83]. Under such an arrangement, civilian leaders grant the military a high degree of professional autonomy, which allows the military to develop and hone its professional expertise, which further causes military leaders to stay out of partisan politics. Figure 1, below, graphically represents the overall problem and solution Huntington attempts to solve.^[As Feaver points out in a 1996 article, Huntington did not use modern political science language such as dependent variable and independent variable when developing this concept. I constructed this puzzle using Feaver's 1996 recapitulation of Huntington.]

```{r figure one, include=FALSE, echo=FALSE}
huntington_dag <- dagify(
    MSCS ~ CC,
    CC ~ FI + SI,
    SI~I + S,
    coords = tribble(
      ~ name,      ~ x,    ~ y,
      "MSCS", 1, .5,
      "CC", .7, .5,
      "FI", .5, .6,
      "SI", .5, .4,
      "I", .3, .3,
      "S",.7, .3,
    ))

huntington_dag %>% 
  ggplot(aes(x=x, y=y, xend=xend, yend=yend))+
  geom_dag_edges()+
  geom_dag_point()+
  geom_dag_text_repel(aes(label=name), show.legend=FALSE)+
  theme_dag()
```

```{r include=TRUE, echo=FALSE, fig.height=5, fig.width=7}

huntington_dag %>% tidy_dagitty() %>% 
  dag_label(labels=c("MSCS"="Military Security and\n Civilian Supremacy", 
                     "CC"="Patterns of\n Civilian Control\n 'Objective Control'", 
                     "FI"="Functional\n Imperative", 
                     "SI"="Societal\n Imperative",
                     "I"="Liberal Ideology",
                     "S"="Constitutional Structure")) %>% 
  ggplot(aes(x=x, y=y, xend=xend, yend=yend))+
  geom_dag_edges()+
  geom_dag_point()+
  geom_dag_text()+ 
  geom_dag_label_repel(aes(label=label, fill=label), col="white", show.legend=FALSE)+ theme_dag(plot.title=element_text(hjust=.5), plot.subtitle=element_text(hjust=.5)) + labs(title="Figure 1: Samuel Huntington's Theory Of Civilian Control of the Military", subtitle="As Explained in The Soldier and The State, 1957")

```

The reason that such an arrangement would work, claimed Huntington, was because of the rise of the military as a profession - a development that occurred during the 19th and 20th centuries. Unlike the military that had existed at America's founding, which had been comprised of citizen-soldiers armed with very basic weapons, the military by the end of World War Two employed sophisticated weapons systems such as battleships, fighter jets, and nuclear weapons such that the officer corps was a body truly skilled in, in Harold Lasswell's terms, the "management of violence" [Lasswell in @huntington_soldier_1957, 11]. These developments in war, and the rise of a professional military, were developments America's founding fathers simply could not predict. ^[It should be noted, however, that determining a precise date for when the US military became a "profession" is a matter of some debate. Huntington makes the point in his book that he sees the rise of the military profession as dating back to as early as 1808, when the Prussian government issued a decree pointing to the content of material that officers in its Army were expected to study. But he does not provide a definitive date for the beginning of the US military as a profession. My reading of Huntington is that he infers the "professionalization" of the US military to have begun in the post-US Civil War period.]

Huntington argued that as a profession, the military's foremost desire is to develop and foster its own professional attributes and characteristics. These attributes included the development of a special body of knowledge and expertise, as well as the establishment of a degree of corporateness, which refers to the way that other professions, such as law and medicine, formalize and govern the standards by which members of the profession are certified [@huntington_soldier_1957, 11-18]. Huntington argued that allowing the military to develop a high degree of professional autonomy would cause the military to stay out of politics, thus preserving civilian control of the military. Simultaneously, the further advancement of the military profession would ensure that the military developed the skills required to ensure American supremacy during the Cold War.

Huntington's concept of "objective civilian control" of the military has generated a number of important and meaningful critiques over the years. The major theme of these critiques is that a stark separation between military and political spheres, which Huntington advocated for, is unwise, impractical, and damaging to US security. Citing Clausewitz's famous dictum that war is a "mere continuation of policy by other means", these critics of Huntington point out that the military cannot truly conduct military operations in a vacuum, so to speak, absent from political influence [@clausewitz_war_1976, 69, 605]. Further, critics of Huntington argue that by virtue of their superior position within democratic governments, rightfully elected civilian officials must ultimately shoulder the responsibility for what militaries do and fail to do, and thus, should not abdicate these crucial responsibilities to the military [@cohen_supreme_2003]. Peter Feaver has parsimoniously captured the essence of this critique of Huntington by noting that the principle of civilian supremacy over the military means that civilian preferences should prevail even if they are wrong [@feaver_armed_2003].

Other recent critiques of Huntington have further developed the possible domestic and international consequences that can arise if military officers strictly adopt his concept of professionalism. For instance, critics warn that adoption of a strict Huntingtonian concept of professionalism can lead to military officers failing to recognize the inherent political nature of their roles, which can subsequently lead to military officers unknowingly engaging in political behaviors under the guise of acting in an apolitical manner [@brooks_paradoxes_2020; see also @janowitz_professional_1960; @feaver_civil-military_1996]. Other critics warn that military officers who have been socialized to expect separation between military and civilian spheres in wartime - which reflects a Huntingtonian understanding of professionalism - will fail to help civilians develop, implement, and achieve sound strategic goals, which has disastrous and negative effects on overall US national security policy [@rapp_civil-military_2015].

In spite of these insightful concerns and critiques of Huntington, which are important, I argue that Huntington and his critics nonetheless share important common ground concerning the appropriate behavior of civilians and the military in politics. Specifically, both Huntington and his critics agree in the prominence of at least three central norms that normatively govern how militaries and civilian leaderes should relate. While the precise definitions and boundaries of these norms may in fact be disputed, these three central norms should be viewed as the baseline set of principles that inform an understanding of whether civilian or military actors are behaving appropriately with respect to the other actor.

The first is a norm of civilian control of the military. While Huntington certainly did not invent the norm of civilian control, the overarching purpose of Huntington's theory of civil-military relations is to ensure both America's security _and_ civilian control of the military. Civilian control of the military is a key foundation of democratic theory, and Huntington's critics explicitly support this foundational principle.

The second norm of civil-military relations that both Huntington and his critics reinforce is a norm of non-partisanship by the military institution. To be clear, Huntington and his critics differ on _how_ this principle can and should be achieved. Huntington believed that non-partisanship of the military institution would follow as a consequence of enabling the military to maximize its professionalism, which required a stark separation of political and military spheres. While critics have pointed out the problems and consequences of Huntington's concept, they nonetheless share the aspirational goal and concern for the military operatins as a non-partisan actor.

The third and final norm that Huntington and his critics share is what I am terming the _non-interference_ of the military into the realm of politics. This norm of _non-interference_ simply means that from the perspective of military actors, there are areas of politics which should be "off limits." These areas involve policy arenas that are unrelated or at best tangentially-related from the world of military and defense policy. In other words, while military officers may in fact be inherently political actors [@brooks_paradoxes_2020] who must be aware of the political implications of their behavior, this does not mean that the military should have a say in every aspect of state policy. In the United States and in other democracies, the military needs to understand the political ramifications of their pending defense budget request, for example, but at the same time, democratic citizens would likely not be pleased if the military sought out or was placed in roles that set them in charge of establishing elementary education guidelines. If such a far-fetched example were to actually occur, democratic citizens would likely at that point agree that the military is being used for areas of state policy for which they were not designed. While the boundaries of where the appropriate versus inappropriate role of the military begin and ends is a matter of intense debate, I assert that Huntington and his critics would nonetheless agree in principle of _non-interference_ by the military into civilian partisan politics and roles [@brooks_shaping_2008; @owens_military_2015; @brooks_paradoxes_2020].

In summary, Huntington's _The Soldier and the State_ has served for decades as a starting point for numerous debates involving civil-military relations. Despite the important critiques of Huntington's work, scholars of civil-military relations have generally agreed upon the importance of three foundational norms which influence the political behavior of both the military and civilian leaders. These norms include principles of civilian control of the military, non-partisanship of the military institution, and _non-interference_ of the military into state policy arenas. Furthermore, adherence to these three critical norms serve as the _baseline_ for determining instances in which the military or civilian actors have deviated from appropriate and important behaviors required to sustain healthy and vibrant civil-military relations.

# Role Beliefs and Political Behaviors of the Military

Although most scholars accept the foundational norms of civilian control, non-partisanship, and military expertise, there is widespread debate about where the line should be drawn in practice between appropriate political expression and inappropriate partisan activity by the military [@owens_military_2015; @golby_beyond_2015; @brooks_paradoxes_2020]. Indeed, establishing conceptual boundaries around the political behavior of the military is a difficult task because, as Mackubin Thomas Owens points out, civil-military relations in the United States can be viewed as a "bargain" [@owens_us_2011] between the American people, the military, and the government that has to account for constantly-changing "political, social, technological, or geopolitical changes" [@owens_is_2016]. Therefore, rendering a normative evaluation of political behavior by the military has to account for the specific context in which these behaviors take place.

In this section, I will summarize three frameworks for understanding why, under what circumstances, and how the military might intervene in politics. The first addresses the concept of military "role beliefs" [@fitch_armed_1998], while the second and third address political behaviors of the military. The concepts of "role beliefs" and political behaviors are similar, but not the same. "Role beliefs" refer to the normative conceptions and attitudes that military actors adopt regarding their role in politics [@fitch_armed_1998], while political behaviors comprise the distinct, observable acts committed by the military in the realm of politics. These concepts are linked in that the conception a military actor adopts regarding his or her role in politics affects _how_ a particular military actor behaves politically.

## Role Beliefs

J. Samuel Fitch introduced the concept of military "role beliefs" in his book, _The Armed Forces and Democracy in Latin America._ Through rich interview data conducted in the 1990s, Fitch traces how military leaders in Argentina and Ecuador adopted one of five distinct "role beliefs," referring to the perceptions and attitudes held by these officers regarding their normative role in politics, and importantly, their roles as military officers with respect to democracy as a political system [@fitch_armed_1998, 61]. These "role beliefs" vary in several dimensions, most notable of which are the degree of subordination and assertiveness militaries should espouse relative to civilian authority, especially in times of crises, and the duties and responsibilities militaries should fulfill within the government.

The first type of role belief Fitch identifies is "democratic professionalism," and involves both "a subjective commitment to democracy as the preferred political system," and a "belief in a professional ethic that differentiates and separates military and political affairs" [@fitch_armed_1998, 66]. This type of "role belief", I would point out, is closely aligned with the normative standards of _military professionalism_ raised by Huntington. A second type of "role belief" Fitch identifies is "classical professionalism," which differs from "democratic professionalism" in that while it "argues for the subordination of the military...to the constitutionally designated civilian authorities", it does not espouse a "normative commitment to a democratic regime" [@fitch_armed_1998, 66-67]. Where a clear departure from American Huntingtonian norms emerges is beginning with a third type of "role belief" that Fitch identifies. When militaries conceive of themselves as a "power factor," they adopt a belief that their defense role within society entitles them to actively participate in "the national project" [@fitch_armed_1998, 67]. Though the boundaries of participation are ambiguous, this third belief suggests that military officers should and do seek to fulfill their "rightful place" within the regime [@fitch_armed_1998, 68].

Continuing a trend of increasing assertiveness of the military institution, Fitch identifies that a fourth type of "role belief" is one of "conditional separation" in which the military stays out of politics "in normal circumstances" but not necessarily if and when crises exist [@fitch_armed_1998, 68]. This type of role belief warrants the military to "safeguard the highest interests of the nation" during times of crisis, "even if that entails overthrowing the incumbent government and assuming power temporarily" [@fitch_armed_1998, 68]. A final possible role belief militaries can adopt is a "tutelary" model in which the civilian government grants the military both a "high degree of autonomy" and the ability to become a regular and "institutional voice in policy matters" [@fitch_armed_1998, 68-69]. Military actors who espouse this role belief deem the military the final "guardian of national security," and thus, that the military has "the right" to participate in a wide variety of policy debates that impact the state [@fitch_armed_1998, 69]. Fitch argues that in the Latin American context, the particular role beliefs militaries adopt are a function of three primary factors: the lessons militaries draw from previous domestic intervention in politics, a wide-ranging assessment of current domestic indicators (the economy, the effectiveness of institutions, etc.), and the prevailing international context [@fitch_armed_1998].

## Political Behaviors of the Military

Two other frameworks are helpful in rounding out a discussion regarding how the attitudes and beliefs that militaries adopt regarding their role in politics impact the specific political behaviors that militaries engage in. The first of these frameworks, that espoused by the British political scientist S.E. Finer in 1962, argues that military intervention in politics is ultimately a function of militaries possessing both the "opportunity" and "motive" to do so [@finer_man_1962]. Regarding possible motives that could spur political behavior by the military, Finer argues that soldiers might adopt a specific form of "manifest destiny," or a worldview in which the military acts as "saviors" of the country [@finer_man_1962, 29]. Finer adds that such a motive can arise as a result of the confluence of the military's unique mission, sense of power, and claim of being "outside party politics" [@finer_man_1962, 29].

A second motive that could cause the military to intervene in politics is what Finer calls the military's "unique identification with the national interest" [@finer_man_1962, 30]. When this motive is present, Finer argues that militaries see themselves as the "custodians" of the "national interest" [@finer_man_1962, 30]. Finer adds that militaries assuming the role of custodian can take different forms, including a "duty to arbitrate or veto...to exercise it when some convulsion or decision of the civil authorities seems to them [the military] to threaten what they think are the permanent interests of the nation" [@finer_man_1962, 31]. Finally, Finer argues that militaries intervene in politics because of their own bureaucratic interests. These can include, Finer says, any of a wide-ranging set of "sectional interests," including those based on social class, regional affiliation, personal career motivations, or organizational interests [@finer_man_1962, 34-53].

Like Fitch, Finer introduces a typology of political interventionism by the military in which he describes four possible levels ranging from the least harmful (and therefore, appropriate) to the most damaging (and therefore, inappropriate). Militaries intervene politically at the lowest level when they "influence," which involves  making appeals "to reason or emotion" in an attempt to persuade the civilian government to adopt a certain course of action [@finer_man_1962, 77]. Finer adds that political intervention by the military at this level is not problematic and entirely "consistent with the supremacy of the civil power" [@finer_man_1962, 77]. Militaries can also intervene politically by "pressuring" or "blackmailing" civilian leaders in an effort to "convince the civil power by threat of some sanction" [@finer_man_1962, 77]. Finer adds that militaries intervening in this way are not normatively wrong in every circumstance, but certain forms of behavior, such as threatening to stage a coup, are clearly wrong from a normative perspective [@finer_man_1962, 77]. A third level of intervention is that of "displacement," and refers to the "removal" of government officials by the military as a result of "violence or the threat of violence" [@finer_man_1962, 77]. The fourth level, that of "supplantment," occurs when the military "sweeps away the civilian regime and establishes the military in its place" [@finer_man_1962, 78].

In describing the ways in which militaries can intervene politically, Finer is providing one of the early critiques of Huntington. Finer specifically takes issue with Huntington's concept of _military professionalism_, or the idea that the military's professional nature ensures that military actors will act in a way that does not threaten norms of civil-military relations. Finer actually points out that while the concept of military professionalism should focus the military on itself by developing the skills and expertise it need to fight and win wars, and thus, away from politics, the danger of the concept of military professionalism is that it can also be abused, particularly by military leaders who adopt the view that they serve the larger concept of "the state" rather than a particular civilian government [@finer_man_1962, 22-23]. Pointing to numerous military leaders who intervened inappropriately in politics, including von Seeckt and Gröner under the German Weimar Republic and MacArthur under President Truman in the United States, Finer argues that the reason these leaders intervened was precisely because of how they conceptualized their duties and responsibilities as professional officers [@finer_man_1962]. Finer's point is that the principle of civilian control, and not military professionalism, ought to serve as the only bulwark to keep the military out of politics.

A third typology of political behavior by the military focuses on the US context in particular, which helps because some of the broad political behaviors that Finer discusses, such as coups, have not yet (thankfully) occurred in an advanced democracy such as the United States. In an edited volume, Risa Brooks outlines and discusses five specific political behaviors that the American military engages in. These five behaviors are shown in Table 1, above.

\singlespace

```{r table1, echo=FALSE, include=TRUE, error=FALSE, message=FALSE}

text_tbl1 <- tribble(~"Tactic", ~"Nature of Activity", ~"Audience", ~"Examples",
  "Public Appeal", "Public, value-laden commentary", "Mass public",  "Colin Powell's 1992 statements on intervention in Bosnia", "Grandstanding", "Threat or actual resignation in protest of policy", "Mass public, Congress, executive branch", "Hypothetical; if Eric Shinseki had resigned in 2003",
  "Politicking",  "Retired officer endorsements; organizing vote drives", "Mass public", "Crowe's endorsement of Clinton in 1992; Frank's endorsement of Bush in 2004", 
  "Alliance Building",  "Military leaders form ties with civilian interest groups", "Congress", "Air Force and Lockheed Martin over procurement of F-22 raptor",
  "Shoulder Tapping", "Military leaders setting an agenda by bringing issues to attention of politicians and engaging in lobbying-like activities on behalf of those issues", "Congress", "Military Mobilization of key members of Congress over gays in the military in early 1990s")

text_tbl1 %>% kable(booktabs=T, caption="Political Activities of the Military") %>%
  kable_styling(full_width = F, font_size=8, latex_options="striped" ) %>% 
  column_spec(1, width="10em") %>% 
  column_spec(2, width="15em") %>% 
  column_spec(3, width="10em") %>% 
  column_spec(4, width="15em") %>% 
  add_footnote(label = "The above table is reproduced from Brooks, Risa (2009), 'Political Activities of the Military' in\n American Civil-Military Relations: The Soldier and the State in a New Era, Eds. Suzanne Nielsen and Don Snider.")

```

\doublespace

Brooks' typology is useful for several reasons. First, her typology illustrates that there are myriad ways in which the norms of civil-military relations can be threatened by different types of political behavior. It is not merely the cataclysmic collisions that occur between the President and wartime Commander that threaten civil-military relations. To be sure, the instances of Lincoln clashing with McClellan, Truman with MacArthur, and Obama with McChrystal are important, but we should consider the other means that are available to the military for creating or raising the costs political leaders face when thinking about civil-military relations. Second, unlike the typologies set forth by Fitch and Finer, Brooks does not rank order certain political behaviors as more harmful than others. She instead suggests that each type of behavior has the potential to threaten norms of healthy civil-military relations depending on the prevailing domestic political context [@brooks_militaries_2009, 215]. This is helpful because it should cause military actors to think through the differing ramifications their political behaviors may have depending on the context in which these behaviors occur. Finally, Brooks' typology draws attention to the fact that it is the public nature of various forms of political behavior that causes certain types of costs for political actors. Along these lines, Brooks points out that during times in which the military is highly respected, the military possesses a considerable "latent political influence" in that it can _choose_ when to engage in political behaviors [@brooks_militaries_2009, 231].

Brooks also recognizes that despite the pitfalls of the military engaging in certain forms of political behavior, there are potentially positive repercussions that can occur as a result of the military acting politically. For instance, Brooks notes that when militaries engage in public forms of commentary, the revelation of information and opinions by military leaders can help citizens better understand and process information that is often complex and disconnected from everyday life [@brooks_militaries_2009, 225-231; see also @schultz_democratic_1999]. Further, when militaries alert Congress to potentially destructive decisions an executive may be considering, the military may be helping the nation avoid negative consequences [@brooks_militaries_2009, 225-231]. Brooks ultimately concludes that military actors must weigh the potential benefits of behaving politically against the costs and risks of doing so. Her personal conclusion, and one I share, is that from the perspective of maintaining healthy civil-military relations in the long run, military officers should error on the side of accepting short-term risks (for example, a citizenry remaining less informed) by not engaging in forms of political behavior that diminish civil-military norms [@brooks_militaries_2009, 236-238].

# My Puzzle, Where it Fits, and Why It Matters

The central question I wish to explore in my dissertation is how political behaviors of the military vary as a result of a) the level of domestic political polarization, and b) the level of prestige the military commands within society. These attributes of domestic politics now exist in contemporary America and will likely remain for some time, and to my knowledge, no study exists that systematically examines the impact of these particular domestic structural factors on the political behavior of the military.

In the scholarship, such a study would make an important contribution to civil-military relations, security studies, and international relations. Fitch's study explored military "role beliefs" among senior members of the military, but from a Latin American context, where the history and traditions of military professionalism are not quite as rich as those that prevail in the US [@fitch_armed_1998]. Finer noted the role that "opportunity" and "motive" play in causing political behavior by the military, but his work, partially due to the fact that it is nearly six decades old, does not provide the granular detail that would be more helpful for application in a contemporary American context. Brooks' typology of political behavior is incredibly detailed and focused on the US military, but does not seek to explain or account for the sources that cause variation in this behavior.

This is not to say that polarization and the military have not been studied. On the contrary, and especially in the past several years during the Trump Administration, a number of scholars have repeatedly commented on what scholars think the influence of polarization is on a number of facets of civil-military relations [@feaver_we_2016; @barno_how_2016; @golby_jim_americas_2020; @reid_retired_2020]. Further, a number of civil-military relations scholars have focused directly on the presence of polarization and how it impacts the military, but the dependent variable in most of these studies have been aspects of the public's trust and confidence in the military as a result of increasing polarization, and not actual instances of the military committing political behavior [@robinson_michael_danger_2018; @burbach_partisan_2019]. We do know, as a result of several rich studies, that current attitudes of military members are changing regarding certain forms of political behavior, such as voting, social media use, senior military leader endorsements of presidential candidates, and the willingness to criticize civilian officials [@dempsey_our_2010; @urben_party_2013; @urben_wearing_2014; @griffiths_not_2019]. These are important studies, but they have not sought to test how changes in the domestic political environment, and particularly, the factors of polarization and prestige, cause changes in the way military actors commit political behaviors.

For these reasons, this dissertation focuses squarely on the political behavior of the military. In doing so, it seeks to advance the understanding of scholars, military officers, civilian leaders, and citizens of how the presence or absence of structural domestic factors, namely political polarization and the prestige of the military, affect how military institutions behave in the political sphere, and why. In desiring to close this gap, this dissertation seeks to answer several important questions, including:

* How does the presence of political polarization and prestige impact the answer to the question, "what does it mean to be a military professional?"
* What political behaviors would we expect to see when polarization and military prestige are low? Why? Does the historical record bear this out?
* What political behaviors would we expect to see when either polarization and military prestige are high, but the other factor is low? Why? Does the historical record bear this out?
* Of the foundational norms required for healthy civil-military relations - civilian control of the military, non-partisanship, and military expertise - which are most at risk in the presence of high (or low) levels of polarization and prestige? Why?
* Within the military institution, which actors are most likely to engage in certain forms of political behavior when polarization and/or prestige are high? 
* Have there been similar periods in American history when polarization and prestige of the military were high? If so, when were they, and did the military commit political behaviors then? In what ways are these periods similar to what we are now experiencing?
* How does the presence or absence of high levels of polarization and/or prestige impact the role of senior military leaders in orchestrating, leveraging, and harnessing the military institution as a political actor?

# Initial Theory

In this third part of the prospectus, I develop an initial theoretical framework for how the domestic levels of political polarization and military prestige impact the political behavior of the military. A graphical representation of this theory is shown in Figure 2, below. From left to right, the model suggests that the independent variables of political polarization and military prestige shape an intervening variable, the beliefs and attitudes of the military regarding their normative role in politics. These beliefs and attitudes subsequently influence the distinct political behaviors of the the military in important ways. I will walk the reader through the theory in two steps.

```{r figure two, include=FALSE, echo=FALSE}
problem_dag <- dagify(
    PB ~ RB,
    RB ~ MP + PP,
    coords = tribble(
      ~ name,      ~ x,    ~ y,
      "PB", 1,.5,
      "RB", .7, .5,
      "PP", .3, .7,
      "MP", .3, .3,
    ))

problem_dag %>% 
  ggplot(aes(x=x, y=y, xend=xend, yend=yend))+
  geom_dag_edges()+
  geom_dag_point()+
  geom_dag_text_repel(aes(label=name), show.legend=FALSE)+
  theme_dag()
```

```{r include=TRUE, echo=FALSE, fig.height=5, fig.width=7}

problem_dag %>% tidy_dagitty() %>% 
  dag_label(labels=c("PB"="DV\n Political\n Behaviors",
                     "RB"="Int Var\n Role Beliefs\n and Attitudes\n re: Civ-Mil\n Norms", 
                     "MP"="IV\n Military\ Prestige", 
                     "PP"="IV\n Political\n Polarization"
                      )) %>% 
  ggplot(aes(x=x, y=y, xend=xend, yend=yend))+
  geom_dag_edges()+
  geom_dag_point()+
  geom_dag_text()+ 
  geom_dag_label_repel(aes(label=label, fill=label), col="white", show.legend=FALSE)+ theme_dag(plot.title=element_text(hjust=.5)) + labs(title="Figure 2: Determinants of Political Behavior by the Military")

```

First, I will explain how the independent variables of polarization and prestige impact and influence the beliefs and attitudes of military actors regarding the inviolability of three foundational norms described in the previous section: civilian control of the military, non-partisanship of the military institution, and military expertise. In this first section, building off of the American politics literature and military history, I trace the rise of political polarization and military prestige in the contemporary US context, as well as describe the presence of war as a scope condition. This step concludes with the development of a basic "2 x 2" table in which I assign an overall level of risk that military actors pose to the foundational norms of civil-military relations as political polarization and military prestige vary. In determining these overall levels of risk, I focus on four factors: whether military actors are likely to be unified or divided in their ideological beliefs; whether the military can leverage either public opinion or their positional influence in government in challenging norms of civil-military relations; what is the main source of corporate motivation among military actors under these conditions, and which of the three foundational civil-military norms are most likely to be challenged as a result of the varying levels of polarization and prestige.

Second, I then project how these overall levels of risk to the foundational norms of civil-military relations under varying conditions of polarization and prestige translate into observable characteristics or attributes of military actors engaging in forms of political behavior. These observable characteristics build off of the typologies Finer and Brooks previously develop, but the goal is to determine how the presence (or absence) of the independent variables of political polarization and military prestige ultimately influence which military actors (who), commit which forms of political behavior (what), for what purpose (why), with what frequency (when), and with what other actors from across government or society ("where" are these behaviors targeted, i.e. Congress, the Media, interest groups, etc.). At the conclusion of this step, I present a total of eleven hypotheses to be tested, analyzed, and explored throughout the course of the dissertation.

## Political Polarization

Within the American Politics literature, there is broad recognition that political polarization levels in contemporary America are higher now than at any point since the end of the American Civil War [@hare_polarization_2014; @jeffrey_b_lewis_voteview_2020]. Figure 3, below, shows American polarization levels as measured using roll-call voting data from the US House and Senate since the later part of the 19th Century. As Figure 3 demonstrates, American polarization levels were high at the end of the American Civil War, and remained high before gradually decreasing early in the 20th century. Polarization reached it's nadir in the period before World War Two, and remained at relatively low levels until approximately the 1970s. Since that time, however, polarization in both chambers of the US Congress has increased dramatically.

```{r figure three, include=FALSE, echo=FALSE}
nom_dat <- read_csv("https://voteview.com/static/data/out/members/HSall_members.csv")

south <- c(40:49,51,53)
polar_dat <- nom_dat %>% 
    filter(congress>45 & 
           chamber != "President") %>%
    mutate( 
      year = 2*(congress-1) + 1789,
    ) %>%
    group_by(chamber,congress,year) %>% 
    summarize(
      party.mean.diff.d1 = mean(nominate_dim1[party_code==200],na.rm=T) - 
                           mean(nominate_dim1[party_code==100],na.rm=T),
      prop.moderate.d1 = mean(abs(nominate_dim1)<0.25,na.rm=T),
      prop.moderate.dem.d1 = mean(abs(nominate_dim1[party_code==100])<0.25,na.rm=T),
      prop.moderate.rep.d1 = mean(abs(nominate_dim1[party_code==200])<0.25,na.rm=T),
      overlap = (sum(nominate_dim1[party_code==200] <
                       max(nominate_dim1[party_code==100],na.rm=T),na.rm=T)  +
                 sum(nominate_dim1[party_code==100] >
                       min(nominate_dim1[party_code==200],na.rm=T),na.rm=T))/
                 (sum(!is.na(nominate_dim1[party_code==100]))+
                  sum(!is.na(nominate_dim1[party_code==200]))),
      chamber.mean.d1 = mean(nominate_dim1,na.rm=T),
      chamber.mean.d2 = mean(nominate_dim2,na.rm=T),
      dem.mean.d1 = mean(nominate_dim1[party_code==100],na.rm=T),
      dem.mean.d2 = mean(nominate_dim2[party_code==100],na.rm=T),
      rep.mean.d1 = mean(nominate_dim1[party_code==200],na.rm=T),
      rep.mean.d2 = mean(nominate_dim2[party_code==200],na.rm=T),
      north.rep.mean.d1 = mean(nominate_dim1[party_code==200 & 
                                             !(state_icpsr %in% south)],na.rm=T),    
      north.rep.mean.d2 = mean(nominate_dim2[party_code==200 & 
                                             !(state_icpsr %in% south)],na.rm=T),    
      south.rep.mean.d1 = mean(nominate_dim1[party_code==200 & 
                                              (state_icpsr %in% south)],na.rm=T),    
      south.rep.mean.d2 = mean(nominate_dim2[party_code==200 & 
                                             (state_icpsr %in% south)],na.rm=T),    
      north.dem.mean.d1 = mean(nominate_dim1[party_code==100 & 
                                              !(state_icpsr %in% south)],na.rm=T),    
      north.dem.mean.d2 = mean(nominate_dim2[party_code==100 & 
                                              !(state_icpsr %in% south)],na.rm=T),    
      south.dem.mean.d1 = mean(nominate_dim1[party_code==100 & 
                                              (state_icpsr %in% south)],na.rm=T),    
      south.dem.mean.d2 = mean(nominate_dim2[party_code==100 & 
                                              (state_icpsr %in% south)],na.rm=T),    
    )

head(polar_dat)
write_csv(polar_dat,path="voteview_polarization_data.csv")

```

```{r include=FALSE, echo=FALSE}
# Here are up to date versions of the famous party means over time plots for each chamber:
polar_dat_long <- polar_dat %>% gather(score,value,-chamber,-year,-congress)
labels <- c("dem.mean.d1"="DEM",
            "rep.mean.d1"="REP",
            "north.dem.mean.d1"="N. DEM",
            "south.dem.mean.d1"="S. DEM")

polarized_plot <- function(chamb) {
  pdatl <- polar_dat_long %>% 
                filter(chamber==chamb,
                       score %in% c("dem.mean.d1","rep.mean.d1",
                                    "north.dem.mean.d1","south.dem.mean.d1")) %>%
                mutate(party=labels[score]) %>%
                ungroup()

  gg <- ggplot(data=pdatl,
               aes(x=year,y=value,group=party,col=party)) +
               scale_x_continuous(expand = c(0.15, 0), 
                                 breaks=seq(1880, max(pdatl$year), by=8)) +
               geom_line() + geom_point(size=0.7) + 
               xlab("Year") + ylab("Liberal-Conservative") + 
               geom_text_repel(data=pdatl %>%
                    filter(year == min(year)),
                  aes(label = party, color = party),
                  size = 3,
                  nudge_x = -8,
                  point.padding = 0.1,
                  segment.color = NA,
                  show.legend = FALSE) +
              theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
              scale_color_manual(values=c("REP"="#c70828","DEM"="#1372ad",
                                         "N. DEM"="#6194F4",
                                         "S. DEM"="#81c4e4"),guide="none") +
              theme_bw()
  gg
}
```

```{r include= FALSE, echo=FALSE}
# House Polarization Plot
gg1 <- polarized_plot("House")
ggsave("voteview_house_party_means.png", plot=gg1, device="png", 
       width=7, height=4, dpi=300)
ggsave("voteview_house_party_means.pdf", plot=gg1, device="pdf", 
       width=7, height=4)
gg1
```

```{r include = FALSE, echo=FALSE}
# Senate Plot
gg2 <- polarized_plot("Senate")
ggsave("voteview_senate_party_means.png", plot=gg2, device="png", 
       width=7, height=4, dpi=300)
ggsave("voteview_senate_party_means.pdf", plot=gg2, device="pdf", 
       width=7, height=4)
gg2

```

```{r include=TRUE, echo=FALSE, fig.align='center'}
# liberal-conserative polarization by chamber
gg3 <- ggplot(polar_dat_long %>% 
              filter(score == "party.mean.diff.d1"),
              aes(x=year,y=value,group=chamber,col=chamber)) +
              scale_x_continuous(expand = c(0.10, 0), breaks=seq(1880,max(polar_dat_long$year),by=8)) +
              geom_line() + geom_point(size=0.5) + 
              xlab("Year") + ylab("Distance between party means") + 
              geom_text_repel(data=polar_dat_long %>%
                    ungroup() %>%
                    filter(year == min(year) & score=="party.mean.diff.d1"),
                  aes(label = chamber, color = chamber),
                  size = 3,
                  nudge_x = -10,
                  point.padding = 0.1,
                  segment.color = NA,
                  show.legend = FALSE) +
              theme(axis.text.x = element_text(angle = 60, hjust = 1), plot.title=element_text(hjust=.5), plot.subtitle=element_text(hjust=.5)) + scale_color_manual(values=c("House"="#c70828","Senate"="#1372ad"), guide="none") +
               labs(title="Figure 3: Liberal-Conservative Partisan\nPolarization by Chamber", subtitle="1880-present", caption="Data Available from Lewis et. al. (2020) at Voteview.com")

gg3

```

Polarization is a difficult concept to precisely define. Generally speaking, polarization refers to the movement of political thoughts and attitudes away from the political center and towards more ideological extremes. While scholars are not unified in explaining the causes of polarization, which range from a growing divide over matters pertaining to race and religion [@abramowitz_great_2018], the size and role of government [@sides_two_2015], disagreements over specific policy issues [@mann_its_2012], and competing visions of personal and national identity [@sides_identity_2018], they are united in their assessment that contemporary levels of political polarization pose significant problems for the health of American democracy. More than this, scholars have noted that the impacts of political polarization reach far beyond the realm of domestic politics. A polarized domestic environment makes it more difficult to determine what threats America faces, what lessons America should learn from its foreign policy failures, and what role in the world America should fill [@schultz_perils_2017; @myrick_external_2020].

For the purposes of the theory advanced here, it is important to understand at least three of the features that are associated with the contemporary polarized domestic environment. These are provided in no particular order, as each of these features are related and provide insight into the overall phenomenon of polarization. The first feature is that polarization in the US goes far beyond mere ideological differences between political elites. Instead, the polarized domestic political environment is marked by increasing levels of disdain among regular citizens for those who identify with the opposite side of the political spectrum. This phenomenon is known in the American politics literature as "affective polarization" [@iyengar_affect_2012; @iyengar_origins_2019] or "negative partisanship" [@abramowitz_great_2018].

Moreover, these feelings of dislike that citizens have towards others that identify with a different political party have not risen in a vacuum, but likely reflect trends in which an increasing number of issues have become part of partisan battlegrounds with little if any prospect for resolution. Moreover, the scope of the issues that comprise these battlegrounds continues to grow. Scholars of American politics describe this dynamic of increasing political warfare as "conflict extension," a second feature of contemporary political polarization in the US [@layman_party_2002]. Scholars note that in an environment marked by conflict extension, there is little to any prospect of meaningful cooperation across partisan aisles [@layman_activists_2010].

A third feature of contemporary political polarization is the belief among partisans that many of the issues at stake today are fundamental and essential to both their individual and national identity [@sides_identity_2018]. As the rise of American "identity politics" illustrates, the normative roles of race, gender, religion, and immigration invoke very strong commitments among those on either side of the political spectrum [@white_values_2002; @mann_its_2012; @sides_identity_2018]. There is an overall belief that the issues at stake in polarizing political environments are not peripheral interests, but rather fundamental and core concerns. It is little wonder then, that both candidates for US President in 2020 claimed, to one degree or another, that the very "soul of the nation" was at stake [@dias_biden_2020].

When these conditions are present, I argue that all three of the foundational civil-military norms face a heightened level of risk. Not only do these conditions increase the potential for civilian leaders "politicizing the military," a dynamic that scholars rightly warn about [@barno_increasingly_2019; @karlin_case_2020], but the dynamics that exist during polarizing political times also increase the likelihood that the military politicizes itself.

Why would this be the case? To begin with, the dynamic of "conflict extension" previously described may decrease the available politically-neutral space upon which the military as an institution can stand [@layman_party_2002]. In other words, a normatively non-partisan actor such as the military will likely find it increasingly difficult to stay out of politics during polarizing times even though it genuinely wants to. However, the dynamic of "conflict extension" can also reflect a more insidious development in that it could describe a condition in which certain military actors strongly believe in a certain issue or set of issues that are raised to the political forefront. These military actors, in turn, may deliberately challenge civilian authority or take a clear partisan side on these issues or set of issues, which could threaten norms of civilian control and/or non-partisanship of the military institution.

Reflecting the dynamics of "affective polarization," these military actors may believe it is more important to advance the causes they care deeply about than it is to fulfill their responsibility qua military professionals who remain subservient to civilian control and who uphold the non-partisan ethic. Indeed, recent experimental research has shown that a significant majority of Americans believe it is more important to advance candidates with whom they share a partisan or ideological affiliation, even if these same candidates violate democratic norms [@graham_democracy_2020]. Other research has shown that citizens find those military elites credible conditional on these elites sharing similar ideological and partisan convictions [@robinson_michael_danger_2018].

Finally, polarization could threaten the norm of military expertise and in at least two ways. First, polarization could lead to political gridlock and a prevailing perception among some military leaders that elected civilian leaders are unable to implement an effective governing agenda. These certain military actors, out of a sense of frustration with civilian leaders' inability to overcome political gridlock, may seek a larger role for the military to fulfill. Such a dynamic is similar to the "role belief" of "conditional separation" identified in Fitch's typology, in which military actors view the emergence of a national crisis as the vehicle for entry into roles that normatively belong to civilian leaders [@fitch_armed_1998,68]. A second and more subtle way that polarization can threaten the norm of expertise is if certain military actors, in accord with the strong beliefs they hold regarding a number of polarizing political issues, strongly believe that the military should play a greater role in helping to solve or alleviate a particular issue that prevails. The differences between these two possible ways of polarization threatening the norm of military expertise are subtle but important. In the first, certain military leaders seek to "take over" the roles of civilian leaders, either by weighing in on political topics that have little if anything to deal with the military, or by offering unsolicited advice to civilian leaders. The primary motivation of these military actors for doing so is frustration with civilian leadership. In the second way, certain military actors do not necessarily want to "take over" the role of civilian leaders, but they want the military to play a greater role in supporting the ideological causes with which they identify. Taking on these greater roles involves the military fulfilling duties that are outside of the scope of traditional military expertise.

In summary, the dynamics of polarization - "conflict extension" [@layman_party_2002], "affective polarization" [@iyengar_affect_2012; @iyengar_origins_2019], and the belief that core values or issues are at stake [@white_values_2002; @mann_its_2012; @sides_identity_2018] -  create multiple mechanisms and ways through which military actors can threaten norms of civil-military relations. Not only does the loss of politically neutral space make it increasingly difficult for military actors to stay out of politics, but polarization may actually be associated with certain military actors desiring to support causes which they strongly identify with and believe in. Depending on the specific political issue or set of issues at stake during polarizing times, certain military actors may believe that the morally and ethically right thing to do is to support these issues, even if doing so threatens norms of civilian control and non-partisanship [@graham_democracy_2020]. Further, if military actors believe that the military should be used to solve crises that are politically controversial, or if these military leaders believe they should offer solutions to solving political problems, these military actors can also threaten norms military expertise.

## The Military's Prestige

The second independent variable I explore in the theory is the military's overall level of prestige. In defining the military's prestige, I am not referring to any single indicator but rather a conglomerate of factors including the military's relative level of influence within the government, it's standing and relationship with society, and the overall level of public credibility and trust that the military possesses relative to other institutions, such as Congress and the Presidency. It is important to emphasize here that public credibility and trust in the military, while an important if not the seminal part of the variable of military prestige, does not fully capture the concept of prestige, mainly because I do not think it goes far enough conceptually. By including bureaucratic influence in a concept of prestige, I am accounting for the power that the military holds as a bureaucratic actor within the government relative to other actors.  I acknowledge up front that many of these sub-component parts of military prestige are difficult to measure. As Richard Betts has stated, "influence...is one of the most important and most elusive concepts in the study of politics. In decisions on use of force it is impossible to detect objectively whose influence is greatest; it can only be inferred because only the President - and maybe not even he - really knows what convinces him or changes his mind" [@betts_soldiers_1977, 5].

A similar concept to that of prestige in the international relations literature is the concept of reputation. In this literature, the concept of reputation almost always refers to the reputation of either a state or head of state [@lupton_reexamining_2018; @renshon_leader_2018]. Additionally, the substance of one's reputation in the international relations literature generally refers to their level of resolve, a concept that is highly connected to theories of deterrence [@powell_nuclear_2003].Two points are worth briefly highlighting in carrying over portions of the concept of reputation into a working definition of military prestige. The first is that similar to the international relations literature, the prestige of the military has both an institutional or corporate component as well as an individual component. Not only is it possible for the military writ large to possess a level of prestige; it is also possible for individual military elites to also have a level of prestige attached to them. The most glaring instance of this is when individual military elites, such as high ranking generals and admirals, become familiar household names: Grant, Pershing, Eisenhower, MacArthur, and Petraeus are all examples who, in each's respective times, commanded incredible prestige as a result of the public's perception of each's contributions to ongoing wartime efforts. The second point to make concerns the substance of the military's prestige. I argue that unlike the literature on reputation which generally focuses on the single factor of an individual's or state's level of resolve, there is no single substantive factor that captures the prestige of the military. How the military is perceived to perform during war, to what extent the military is a powerful actor within the governmental bureaucracy, to what extent the public perceives the military behaving and acting ethically, and to what extent the military upholds foundational norms of civil-military relations are all important aspects of the military's prestige. In very general terms, we would expect the military's prestige to be high during or after achieving victory in times of war (post World War Two, post Gulf War), or when the military is perceived by the public to have filled a significant role in responding to a devastating natural disaster (after Hurricane Katrina), and to be low when the military is perceived to have either lost a war (post Vietnam) or to have been plagued by troubling scandals (the Navy and the Marine Corps, specifically, after the 1991 "Tailhook" Scandal).

With a general understanding of what military prestige is and what it refers to, the next task is to briefly summarize the growth in the military's prestige in the post World War Two era. Two main drivers of the military's prestige after World War Two have been, I argue, the growing institutional influence of the military within the national security apparatus, and the military's growing exclusiveness between itself and American society. Each of these trends have slowly developed. After summarizing these trends, I then discuss how high levels of military prestige can affect the foundational civil-military norms of civilian control, non-partisanship, and military expertise.

After World War Two, the government instituted several institutional and legislative changes in order to integrate American national security efforts. The first of these legislative reforms was the National Security Act of 1947, which was signed by President Harry Truman and which sought to unify the defense and intelligence establishments. In doing so, this legislation created new institutions, including the National Security Council (NSC), the Central Intelligence Agency (CIA), and the US Air Force [@noauthor_national_1947]. Formally, the National Security Act also merged the Departments of War and Navy into what would be first called the "National Military Establishment," which was essentially a precursor to what became in 1949 the Department of Defense [@locher_has_2001]. For the next several decades, the reforms wrought by the National Security Act helped the defense and intelligence communities prevail in a number of conflicts, including the Cold War, and to ultimately ensure the security of the United States. So influential was the National Security Act that fifty years after its signing, President Bill Clinton stated that as a result of the institutions created by the National Security Act and the people who had served in them, a generation had "helped to secure the peace and prosperity that America enjoys...the success of their efforts and of the historic legislation enacted half a century ago is reflected in an outstanding record of achievement: nuclear war averted, the Cold War won, and the nations of the world turning to democracy and free markets" [@clinton_weekly_1997].

Like most institutional reforms, however, the National Security Act of 1947 could not nor foresee all of the difficulties that the future would bring. A second set of major institutional defense reforms was undertaken in the mid-1980s. Known as the Goldwater-Nichols Act, and named after Senator Barry Goldwater and Congressman William "Bill" Nichols, several incidents made clear that the military lacked the requisite capabilities to plan and operate in tactical environments which required military personnel from across the various services [@kitfield_prodigal_1997].^[This included the the failed Iranian hostage rescue mission in 1980. For an excellent account of this event and the demand for "joint" services afterwards, see @kitfield_prodigal_1997.] Especially as military technologies and equipment grew increasingly complex and sophisticated, the Goldwater-Nichols legislation sought to ensure that national defense interests were identified and prioritized over and above the parochial interests of the individual military services [@locher_has_2001]. Among the most significant of the changes wrought by the Goldwater-Nichols Act was codification of the Chairman of the Joint Chiefs of Staff as the President's principal military adviser, a change which in essence meant that all advice from the respective individual service chiefs of the Army, Navy, Air Force, and Marine Corps would flow through one single point of contact: the Chairman of the Joint Chiefs [@locher_has_2001; @wills_effect_2016]. Additionally, the Goldwater-Nichols Act created several regional combatant commanders who would report directly to the Secretary of Defense, thus bypassing the military service chiefs [@locher_has_2001; @wills_effect_2016].

The National Security Act of 1947 and the Goldwater-Nichols Act of 1986 had the collective impact of significantly strengthening the military relative to other civilian institutions. Indeed, one scholar claims that the institutional changes wrought by these pieces of legislation had the practical effect of causing the military and the Pentagon to replace the State Department as the government's "organizational center of gravity for U.S. national security and foreign policy" [@schmidt_silent_2019]. Without question, in the decades since World War Two, the institutional influence of the US military has grown considerably.

Yet a second driver of the military's growing prestige in the post World War Two years is the growing perception of the military as an exclusive institution that is too separate from American society. While the military is rightly recognized as unique among institutions, there is growing concern that too much of American society can no longer relate to those who wear the uniform. Within the scholarly literature, this topic is often referred to as the _gap_ between society and the military, and numerous debates exist concerning the _gaps's_ causes, size, and potential solutions. At first glance, one might think that a _gap_ between society and the military might diminish, rather than enhance, the military's prestige. Depending on the circumstances, such as the late years of the Vietnam War, this could be true. However, the contemporary concern among most scholars of civil-military relations who study the _gap_ have determined the opposite to be true: the separation of the military from American society has resulted in the military being elevated to positions of prominence that have numerous deleterious impacts.

Since the nation's founding, Americans have generally been suspicious of large numbers of standing military forces, tracing back to suspicions held by American colonists over the establishment of British military garrisons [@weigley_american_2001]. The early founders even held significant differences of opinion over how large America's military forces should be, and whether these forces should be comprised of volunteers (favored by Thomas Jefferson) or a standing military (favored by Alexander Hamilton). In the end, the founders chose a concept that established a moderately-sized standing Army that could be augmented, as required, by state militias comprised of "citizen soldiers" during times of war [@weigley_american_2001]. Further, America has relied on this "citizen soldier" concept for most of its history. It was predominantly not until the the 20th century when the government faced wartime demands that required instituting mandatory military service, especially during World War Two and later, Vietnam [@rostker_i_2006].

However, the Vietnam War changed the way Americans and military leaders viewed the Armed Forces, as a result of developments abroad and at home. The war proved to be tremendously unpopular on the domestic front as the length of the war increased, and Americans came to view the existing draft system as inequitable, both from racial and economic perspectives [@rostker_i_2006]. Additionally, the growing population of the United States, combined with severe disciplinary issues that plagued the military at the time, led many Americans as well as military leaders to believe that an all-volunteer force was at least a viable alternative to the draft [@rostker_i_2006]. Finally, in 1973, the United States abandoned the draft and instituted the All-Volunteer Force (AVF), the result of recommendations made by the Gates Commission, which had been established by President Richard Nixon [@rostker_i_2006].

It is difficult to find a scholar or pundit who does not laud the tactical proficiency of the AVF. Indeed, there is near-universal acceptance of the tremendous war-fighting skill levels reached by the AVF since its inception. And yet, multiple scholars assert that in the years since Vietnam, when the military draft ended, there have been noteworthy gaps that have developed between the military and American society.

Nonetheless, there are at least three specific gaps that have been identified by scholars of civil-military relations: a geographical gap, a partisan gap, and a generational gap. Regarding the geographical gap, Golby and Liebert note that a disproportionate number of those serving in the military hail from rural and southern states, and that a large proportion of military recruitment stations, Reserve Officer Training Corps (ROTC) programs, and military installations are located in the southern United States [@liebert_midlife_2017]. The partisan gap is a little more complicated. Although the military officer corps does in fact disproportionately identify with the Republican Party versus the Democratic Party [@golby_duty_2011; @liebert_midlife_2017], studies have also shown that differences between the military and civilians become smaller if not altogether diminished after conditioning on partisan affiliation [@golby_duty_2011]. Finally, studies have shown that a third gap exists between the military and younger Americans, a gap that has grown as fewer Americans have a firsthand connection to someone either serving or who has served in the military [@colford_young_2016].

To the extent that these gaps should and actually do matter is a matter of some debate. Critics of the AVF allege that when gaps between society and the military become too large, it is too easy to send the military to war, as there are little if any corresponding burdens society must bear [@bacevich_breach_2014; @fallows_tragedy_2014]. These critics also allege that the logical consequence of gross unfamiliarity between American society and its military is fighting of wars of choice, which ultimately result in indecisive outcomes and an altogether feckless foreign policy [@bacevich_breach_2014]. Finally, these critics argue that dynamics have unfortunately borne themselves out in the post-9/11 era as the military has endured repeated deployments to Iraq and Afghanistan [@bacevich_breach_2014]. Pointing out that President George W. Bush lowered taxes and encouraged Americans to travel and take vacations even as the American military was invading Afghanistan and Iraq, Bacevich for instance goes so far as to say that "the state now owns war, with the country consigned to observer status" [@bacevich_breach_2014, 44]. The journalist James Fallows, in an influential article, further describes the consequences of a military-societal gap grown too large. He alleges that Americans in the post-9/11 era have developed a "reverent but disengaged attitude toward the military - we love the Troops, but we'd rather not think about them." [@fallows_tragedy_2014].

For better or for worse, Americans indicate that they greatly respect their military. For instance, according to Gallup public opinion polls, 72 percent of Americans indicated that they trust the military a "great deal" or "quite a lot" - the most of any institution, to include Congress (13%), the Presidency (39%), the police (48%), the Supreme Court (40%), and the church/organized religion (42%) [@noauthor_confidence_2020]. Moreover, trends in trust of various American institutions over time reveal that whereas trust in the military has generally increased over the past 45 years, trust in the other institutions have generally declined or remained the same.

The norms of civilian control of the military, and to a lesser degree, the norm of military expertise, are the most likely to be impacted by the growth in the military's institutional and societal prestige. There are at least two ways in which high military prestige can lead to the threatening of these foundational civil-military norms.

A first possible way in which high military prestige can threaten civilian control of the military is if military leaders willingly assert themselves too brazenly in the civil-military dialogue process. When interacting with civilian leaders on any number of issues, military actors may adopt an attitude such that they believe their military service entitles them to be heard in a certain deferential manner [@beehner_matter_2020]. A similar dynamic occurs if civilian leaders fail to sufficiently assert themselves into the policy process when interacting with military actors who are incredibly popular. Karlin and Friend, for instance, note a dynamic of "military worship" in which military service is somehow viewed as better, more relevant, or more appropriate than other types of civilian service which ultimately harms civilian control of the military and other democratic norms [@friend_military_2018]. As a result, military actors may insist on solving problems in the manner that they want the job done, and not in a way that aligns with the desire of civilian leaders. Or, in the case of "military worship," civilians will not hold the military accountable to the degree that they should be [@friend_military_2018].

Even if "military worship," does not occur, the fact that the military commands more resources may make it more difficult for civilians, in some cases, to assert themselves. Over the past several decades, for instance, the budget of the State Department has remained relatively constant when measured in real dollars, whereas the Defense Department's Budget has exploded [@matisek_crisis_2018]. Over the same time period, the military has taken on an expansive set of missions and responsibilities that are far outside the traditional scope of conventional combat operations [@brooks_how_2016]. Thus, even if military leaders do not willingly seek to assert themselves over civilians because they believe that their opinions are somehow "better" than that of civilians, their position of influence makes them, by default, very powerful actors capable of subverting civilian control with relative ease. Additionally, trends in the military accumulating responsibility for tasks that are outside the scope of traditional combat operations might lead to the military advocating for assuming responsibility for additional roles and tasks, which likely bring with it increased budgetary power and institutional influence, developments which could further threaten the norm of military expertise.

These concerns are amplified, moreover, when civilian leaders are literally not present in the room and at the table when civil-military dialogue occurs. Indeed, during the Trump administration, numerous civilian positions (more than 1/3rd) in the Pentagon went unfilled, prompting concern from Congress and other lawmakers that the first half of the civilian-military dialogue process was notably absent [@seligman_civilian_2020]. While the vacancy of these positions was not a result of high levels of military prestige, a major concern was the insufficient civilian oversight of a very powerful military establishment.

A second, and equally powerful way in which high levels of military prestige can threaten the norm of civilian control of the military is if military actors hijack the public credibility they possess in order to subvert civilian leaders. There are a number ways in which this can occur, and to varying degrees. Leaking information or making public statements on purpose, in one way or another, brings the American public squarely into the fold of any actual or perceived disagreement between civilian officials and military leaders. Engaging in this type of behavior is likely a very deliberate decision by military leaders, because one of the reasons that the military likely enjoys a high degree of public prestige is precisely because they do not routinely engage in this sort of behavior. However, depending on the circumstances, the military may conclude that the particular situation warrants making a public appeal that relies on the high credibility the military enjoys.

In summary, when militaries possess a high degree of prestige, military actors are increasingly willing to challenge norms of civilian control. Military leaders may believe that their experience entitles them to fill a certain role in the policy process, or to provide advice that deserves to be listened to over other types of advice. Similarly, civilians may fail to push back or challenge military actors when the military enjoys a powerful degree of prestige and influence. Moreover, as the government continues to rely on the military to solve problems at the exclusion of other governmental agencies and institutions, the scope of the military's duties and responsibilities grow. Thus, when military prestige is high, the military may increasingly believe that it has the solution to how to solve a particular problem, and may seek a further expansion of its roles and responsibilities. This could further threaten the norm of military expertise.

# When Polarization and Prestige Collide

Given the impacts of political polarization and the prestige of the military on the beliefs and attitudes of military actors, it is possible to now assess the overall level of risk that military actors pose to the foundational norms of civil-military relations in environments where political polarization and military prestige vary. Borrowing the language Finer uses, I argue that high polarization levels serve as the "motive" for military officers to intervene in politics, and thus, to engage in political behavior, while high levels of military prestige provide the "opportunity" to do so [@finer_man_1962]. These levels of risk, and a brief description of these environments, are displayed in Table 2, above.

\singlespace

```{r table2, echo=FALSE, include=TRUE, error=FALSE, message=FALSE}

text_tbl2 <- tribble(~"Environment", ~"Low Polarization", ~"High Polarization",
  "Low Prestige", "Low Overall Risk. Military actors are united ideologically. Military actors are unable to harness/mobilize public opinion. The military acts as a unified institution and adopts a belief that its members are servants of the Republic. The norm at greatest risk is civilian control.", 
  "Moderate Overall Risk. Military actors may be divided ideologically, but are unable to harness public opinion or bureaucratic power in pursuit of ideological causes. The military acts mostly as a unified institution and adopts a belief that its members are servants of the Republic. The norms at greatest risk are civilian control and non-partisanship.",  
  "High Prestige", "Moderate Overall Risk. Military actors are united ideologically. The military can harness their influence in government and public opinion, but are likely to do so only when their prestige is threatened. The military acts as a unified institution and adopts a belief that while they are servants of the Republic, they also can and should advocate for specific positions related to military prestige. The norm at greatest risk is civilian control.", 
  "High Overall Risk. Military actors are divided ideologically and can harness both public opinion and bureaucratic power. Some military actors believe that the pursuit of ideological causes outweighs adherence to foundational civil-military norms. The military is less likely to act as a unified institution, but certain individual military actors emerge as saviors or guardians of the national interest. All foundational civil-military norms are at risk." 
  )

  kable(text_tbl2, booktabs=T, caption="Overall Risk Level to Norms of Civil-Military Relations and Role Beliefs of Military Actors As Political Polarization and Military Prestige Vary") %>%
  kable_styling(full_width = F, font_size=8, latex_options="striped") %>% 
  column_spec(1, width="10em") %>% 
  column_spec(2:3, width="20em") 
```
\doublespace
The overall risk level posed by military actors to foundational norms of civil-military relations is low when polarization and prestige are both at relatively low levels. Under these conditions, military actors possess neither a motive nor opportunity to intervene in politics [@finer_man_1962]. When polarization is low, there is little if any ideological substance for military actors to engage in political behaviors that threaten norms of civil-military relations. Moreover, because the level of military prestige is also low under these conditions, the military has little institutional influence to leverage, and it lacks the ability to harness widespread public opinion. Therefore, the military is able to act as a unified institution. This is not to suggest that there is in fact no risk at all to civil-military norms under these conditions. Rather, conflict between civilians and military actors is likely to occur in a manner that is less visible to the public, and over substance areas that are directly related to the military's traditional duties and responsibilities, such as maintaining military readiness, and fighting and winning during combat. The norm of civil-military relations at greatest risk under these conditions is civilian control of the military.

Moving to the top right of the table, when polarization is high but military is prestige is low, the overall risk level stemming from military actors to foundational civil-military norms is moderate. Stemming from conditions of high polarization, military actors may in fact possess a "motive" to intervene in politics, but lack the "opportunity" to do so as a result of low institutional prestige [@finer_man_1962]. Under these conditions, military actors may be divided ideologically. Some military actors may desire to advance ideological causes while others believe it is more important that the military remain apolitical and non-partisan at all costs. As a result of low levels in military's prestige, however, military actors who wish to advance the causes with which they strongly identify are unable to harness public opinion or institutional influence in support of these causes. For these reasons, the military will be mostly capable of acting as a unified institution. The norms at greatest risk under these conditions are civilian control and non-partisanship of the military.

The bottom left of the table depicts a world in which domestic polarization levels are low, but military prestige is high. Under these conditions, the military possesses an opportunity to intervene in politics, but lacks the motive to do so outside of protecting its high degree of prestige. Thus, the overall risk level to civil-military norms is moderate [@finer_man_1962]. When these conditions exist, there are few partisan or ideological issues that serve to divide military actors, and thus, military actors are relatively unified. Because military actors enjoy a high level of prestige, they are willing to challenge norms of civilian control and are more assertive in their overall behavior. However, the substance of these challenges to civilian control are less likely to involve ideological concerns than they are debates about traditional military topics or matters, such as military readiness and the use of force abroad. The civil-military norm at greatest risk in this environment is civilian control of the military.

The bottom right of the table reflects the conditions of high polarization and high military prestige. Under these conditions, the overall risk to norms of civil-military relations posed by military actors is high. Military actors have both a motive and an opportunity to intervene in politics [@finer_man_1962]. Under these conditions, the military institution faces a collective action problem and therefore, cannot act as a unified institution. Some military actors may deem it more important to support ideological causes than it is to uphold the foundational norms of civil-military relations. Due to the high level of polarization that exists in society, other military actors who wish to uphold norms of civilian control and non-partisanship at all costs will find it increasingly difficult to do so. Furthermore, because the military's prestige is also high under this set of conditions, some military actors are likely to behave in a way that harnesses the military's high level of influence in order to support or advance the causes which they support. In some cases, the actors who desire to advance the ideological causes with which they identify may seek an expanded role for the military, or themselves fill a role that is traditionally the purview of civilian leaders. Some actors believe that the military's role, above all else, is to remain subservient to civilian control and outside the realm of partisan politics. Other military actors, however, believe that their normative role as members of a prestigious institution compels them to support specific causes, even if the causes with which these actors identify are heavily conditioned by their partisan affiliation.

## How Prestige and Polarization Impact Political Behaviors by Military Actors

Having now identified an overall level of risk to the norms of civilian control of the military, non-partisanship of the military institution, and military expertise under varying conditions of domestic political polarization and military prestige, it is now possible to develop a set of predictions and observable indicators about specific political behaviors that military actors may commit when these conditions prevail. Based on the theory laid out in this prospectus thus far, the goal is to make predictions that are as granular as possible without extending past what the theoretical groundwork allows. I build off of the specific political activities highlighted by Brooks but now project how the combined influence of polarization and prestige impact several of these activities [@brooks_militaries_2009].

The first type of political behavior, or "tactic," that military actors can commit is a "public appeal" [@brooks_militaries_2009]. In this category, I include writing op-eds, giving interviews, and even leaking information. In environments where polarization and prestige are both low, we would expect military actors to make few public appeals. If military actors do make public appeals under these conditions, the topical substance of these appeals will likely center less on ideological issues, but rather directly on topics that are related to traditional matters of military expertise, such as military deployments, readiness, weapons systems, etc. Under conditions of low prestige and high polarization, military actors would still likely conduct a limited number of public appeals, but the few they do make might occasionally address a topic outside of the traditional realm of military expertise, or overtly criticize a civilian leader for ideological differences. The overall number and frequency of public appeals under conditions of low military prestige, however, is likely to be low relative to conditions of high military prestige.

Under conditions of high prestige and low polarization, the military will appeal to the public more often than they would under conditions of low military prestige. However, in terms of content, we expect that unified military actors will limit the topical content of their public appeals to issues or concerns that are directly related to traditional matters of military expertise. Under conditions of high polarization and high military prestige, however, military actors (who are no longer unified ideologically) might frequently make public appeals. The content of these public appeals, will include a higher frequency and overall level of appeals that stray outside of topics that are directly related to traditional matters of military expertise. Some of the public appeals made by military actors will adopt an increasingly partisan tone, and some will criticize civilian leaders in topical areas that stray further from the topical expertise of military actors.

The second type of political behavior is "grandstanding" [@brooks_militaries_2009], which refers to the resignation (or threat of resignation) of a military actor. In political environments of both low polarization and military prestige, there would exist little if any ideological basis for military actors to resign, and so, any instances of resignation would likely be for disagreeing with civilians on distinctly military issues or topics. But owing to low levels of military prestige, there would be relatively little public impact if a military actor were to resign. Thus, we would expect few instances of resignation, and the few instances that may occur to be based not on partisan ideological differences, but on concerns regarding traditional military topics. When polarization is high but military prestige is still low, there is still little if any public impact to be gained by a military actor for resigning. However, the likelihood increases that the basis for a military actor's resignation includes ideological differences with civilian leaders.

Under conditions of high prestige and low polarization, military actors now possess the ability to leverage public opinion to impose political costs on civilian leaders through resignation. Thus, there is a reasonable expectation to see more resignation or threats of resignation under these conditions. However, there is little if any reason to expect an ideological basis for the resignation of a military actor. Rather, instances of resignation are likely to involve disagreement over traditional military topics. These expectations change in environments marked by both high polarization and high military prestige. Under these conditions, there is an expectation that an increased number of officers may threaten to or actually resign relative to conditions of low military prestige, and also an expectation that the basis for the resignation of these military actors will include ideological issues that are beyond the traditional realm of military expertise.

The third type of political behavior Brooks identifies is "politicking," which refers to the act of retired military officers making official endorsements of partisan political candidates [@brooks_militaries_2009]. I include in this category behaviors such as military officers speaking at political conventions, or campaigning on behalf of a political candidate. Under conditions of low military prestige and low polarization, the expectation is that few officers participate in this type of behavior. There exists little if any demand from civilian political candidates for the endorsement of active or retired military officers because military actors do not possess a high level of prestige. There is also under these conditions little if any ideological motivation to propel an officer to support a partisan candidate for office over another candidate. Under conditions of high polarization and low military prestige, some officers might actually want to put their support behind a candidate with whom they share similar ideological or partisan convictions, but a candidate for office still has little incentive to gain from the public in obtaining a military actor's endorsement. Under both of these conditions, the expectation is that there are few instances of military actor endorsement of political candidates.

Under conditions of low polarization and high military prestige, expectations are different. Because of the high prestige of the military institution, political candidates will seek the endorsement of military actors. Although some (perhaps even most) military actors will avoid making an endorsement of a political candidate so as to not jeopardize the norm of non-partisanship of the military institution, other actors will agree to endorse. However, because the level of political polarization is low, visible endorsements may be lopsided. There is a greater likelihood under conditions of low polarization relative to high polarization that military actors (who are relatively ideologically united) will coalesce around a candidate whom they view as supportive of the military, i.e. a candidate who will increase military spending, take care of the troops, etc. Endorsements made by military actors will focus on what a particular candidate will do for the military. In environments marked by high military prestige and high political polarization, several differences are likely to emerge. I expect the endorsement of political candidates by military actors not to be lopsided, but more evenly spread. Moreover, under these conditions, military actors are more likely to base their endorsements not on narrow conceptions of what a political candidate will do for the military, but rather on shared ideological convictions between the military actor and the political candidate. Under these conditions, the question of which candidate is better for the military thus takes on an entirely different meaning than it does under conditions of high military prestige and low polarization. Endorsements by military actors under these conditions are likely to involve content that focuses less on why the candidate whom a military actor is endorsing is better for the military institution, and more on why the candidate is better for America. In making endorsements under these conditions, military actors will thus likely discuss topics that fall outside of the actor's traditional military expertise, and are likely to attack or criticize the other political candidate on partisan ideological grounds.

In her typology, Brooks identifies two additional types of political behaviors, each of which has to deal with how the military engages Congress.  The first is that of "alliance building," in which "military leaders form ties with interest groups" in an effort to lobby Congress, and the second is "shoulder tapping," which refers to "military leaders setting an agenda by bringing issues to the attention of politicians and engaging in lobbying-like activities on behalf of those issues" [@brooks_militaries_2009, 219]. I combine these behaviors under a more general description of the military engaging Congress as a third party. When political polarization and military prestige are low, the military is likely to engage in this sort of behavior to obtain funding or weapons systems that will enable the military to win on the battlefield in order to improve its overall level of prestige. Because military prestige is low, there is less reason to expect that members of Congress will listen to the military than if the level of military prestige is high. When political polarization is high and prestige is low, some military actors may engage in lobbying efforts to advance causes other than the military's ability to fight and win on the battlefield, but these efforts will likely be limited and occur out of the public eye.

When military prestige is high but political polarization is low, military actors are likely to lobby Congress for matters that the military insists are directly related to their combat mission, such as readiness. Moreover, such lobbying efforts are likely to be successful because members of Congress do not want to be viewed by their constituents as failing to support the military. Additionally, military actors will likely focus their lobbying efforts on pro-military, influential members of Congress who can subsequently put pressure on other members of Congress to support the items for which the military is lobbying. When polarization and military prestige are high, some military actors may engage in efforts to advance ideological causes not directly related to the military's combat mission. These efforts may result in visible connections between members of the military and members of Congress that that fall outside of the political center.

These political behaviors, key indicators, and observable descriptions are displayed in Table 3, below, along with corresponding historical periods of the United States in which general conditions of polarization and military prestige prevailed.

```{r table3, echo=FALSE, include=TRUE, error=FALSE, message=FALSE}

text_tbl3 <- tribble(~" Political Behavior", ~"Key Indicators", ~"Low Polarization, Low Prestige", ~"High Polarization, Low Prestige", ~"Low Polarization, High Prestige", ~"High Polarization, High Prestige",
  "Public Appeals and Leaks", "What topics do the public appeals address? Do they address matters of military expertise? Do they criticize civilian leaders or take partisan sides? What is the frequency/number of public appeals made?", "Topics directly relate to military expertise. Criticism of civilians or taking partisan sides is rare. The level and frequency of public appeals is low. Leaks are uncommon.", "Topics mostly address matters of military expertise. Public criticism of civilians or taking partisan sides is uncommon. The level and frequency of public appeals is low. Leaks are possible but infrequent.", "Topics directly relate to matters of military expertise. Criticism of civilians is possible, but related to military affairs. Taking ideological partisan sides unlikely. Level and frequency of public appeals is high. Leaks possible.", "Topics include matters of military expertise and ideological concerns that are tangential to military affairs. Criticism of civilians and taking partisan sides more frequent. Level and frequency of public appeals is high. Leaks likely.",
  "Resignation", "What is the reason or basis of resignation (threat of resignation)? How often do they occur, and at what level of military leadership?", "Reasons for resignation are directly related to traditional military expertise. Resignations are rare and occur at the pinnacle of military leadership.", "Some resignations or threats to resign may occur for ideological differences tangentially related to military expertise. Resignations are rare, but may occur at subordinate levels of military leadership.", "Reasons for resignation are directly related to matters of military expertise. Resignations or threats to resign are possible and impose real costs to civilian leaders, and occur at the pinnacle of military leadership.","Reasons for resignation may include ideological differences between military and civilian leaders or matters directly related to military expertise. Resignation or threats of resignation may occur at subordinate levels of military command.",
  "Political Candidate Endorsement", "Do endorsements occur at all? Are endorsements one-sided, or evenly distributed? Is the basis of the military actor's endorsement that the candidate is 'pro-military,' or is the basis ideological?", "Few endorsements made, as little demand among civilian actors exists for military actor endorsement.", "Few endorsements made, as little demand among civilians exists for military actor endorsement.", "Endorsements exist, but are lopsided. Military actors endorse the candidate whom they think is better for the military.", "Endorsements exist, and are evenly spread. Military actors endorse the candidate whom they align with ideologically and whom they believe will be better for America.",
  "Militaries Use Congress as a 3rd Party", "For what purpose does the military lobby Congress (related directly to military expertise)? Do members of Congress face pressure from constituents to support the military? Is the lobbying effort of the military effective?", "Lobbying efforts by the military are directly related to traditional military expertise. Members of Congress are not significantly pressured to support. Efforts are somewhat effective.", "Lobby efforts are mostly related to traditional military expertise, with some deviation. Members of Congress are not significantly pressured to support. Efforts are somewhat effective.", "Lobby efforts are directly related to traditional military expertise. Members of Congress feel pressure to support the military. Efforts are very effective.", "Lobby efforts are mostly related to traditional military expertise, with some deviation. Efforts are very effective, and members of Congress feel pressure to support the military.",
  "Historical Eras", "","Post Vietnam; Pre World War Two", "Gilded Age (1875) Through Pre-World War One", "Post World War Two; Post Gulf War", "Reconstruction after the Civil War; 2010-Present Day"
  )

  kable(text_tbl3, booktabs=T, caption="Expectation of Political Behaviors by Military Actors as Political Polarization and Military Prestige Vary") %>%
  kable_styling(full_width = F, font_size=8, latex_options="striped") %>% 
  column_spec(1, width="6em") %>% 
  column_spec(2:4, width="13em") %>% 
  column_spec(5, width="15em") %>%
  column_spec(6, width="17em") %>%
    landscape()

```

At a minimum, eleven hypotheses can be uncovered from the predictions posited in Table 3. These are:

_H1: When political polarization is high, the military (or certain military figures) will increasingly criticize civilian leadership (indicating a drop in the principle of civilian control)._

_H2: When political polarization is low, the military (or certain military figures) will publicly weigh in on traditional national security debates (and not on domestic political concerns or issues outside of relevant military expertise)._

_H3: During periods of high prestige and credibility, the frequency of military involvement in quasi-political activities (such as writing op-eds) will increase relative to periods of low prestige and credibility._

_H4: When both political polarization and high prestige are present, the military will be increasingly vocal and at higher frequency about certain domestic ideological or partisan causes than when these factors are not present._

_H5: During periods of active ground combat, military elites will likely comment on the war, but with less frequency on other domestic issues._

_H6: Resignations (or threats of) by military actors increase as the prestige of the military rises_.

_H7: The basis of resignation (or threat of) becomes increasingly ideological as polarization increases_.  
_H8: Endorsements by military actors become more evenly distributed as polarization increases_.

_H9: The number and level of endorsements made by military actors increases as military prestige increases_.

_H10: The effectiveness of military lobbying efforts with Congress increases as military prestige rises._

_H11: The ideological purpose of military lobbying efforts with Congress will widen as polarization increases. When polarization is low, the purpose of military lobbying efforts will center on matters of traditional military expertise._

Before laying out the plan of the dissertation, a few scope conditions are required to help focus the dissertation. The first involves which military actors this dissertation plans to focus on. I will focus my attention in the dissertation on _military elites_, whom I take to mean those active _and_ retired career military officers who serve in influential and visible command and staff positions. The term _military elite_ has a strong scholarly basis, though previous scholars have defined it differently and in rather loose terms. Huntington and Janowitz saw military elites as the entire officer active duty officer corps, a group that for the purposes of this study, is both too broad and yet also not broad enough [@huntington_soldier_1957; @janowitz_professional_1960].

The entire active duty officer corps includes junior officers who are, I argue, still in the process of adopting and learning professional norms, and importantly, what adherence to these norms looks like in practice. The political behavior of junior lieutenants and captains, while potentially interesting and certainly worth scholarly exploration, is not likely to impose costs on political leaders that are high and which result in political leaders adjusting course. However, active duty officers in the rank of roughly Lieutenant Colonel and higher, who have served close to a full-length career of twenty years or more, have learned and know the normative standards of civilian control, non-partisanship, and military expertise such that deviation from these standards can and should be considered as qualitatively different from junior officers who do the same.

In the same vein, I argue that _military elites_ must include retired military officers, and particularly those who have achieved the rank of general or admiral. Commonly referred to as flag officers (referring to the literal flag accompanying these officers while on active duty, which bears the number of stars of their rank), these officers play a substantial role in representing the military institution to America. Today, they increasingly dialogue with the media, and as other scholars have noted, many if not most Americans cannot distinguish the behavior of retired military officers from those on active duty, which has generated concern that the American public perceives retired military officers as speaking on behalf of the military [@golby_military_2012; @golby_jim_americas_2020]. Therefore, in line with dozens of works of previous scholarship, the term _military elite_ refers to both active _and_ retired career military officers, and especially to those who have achieved the rank of general or admiral.

The second and final scope condition involves the time horizon of the study. Here, I think there are two potential options, both of which have relative strengths and weaknesses. The primary option is to focus entirely on the post-World War Two period. Limiting the study to this time still allows the researcher to observe variation in both of the key independent variables. In addition, the military institution has more or less existed during this time period as a profession. While the threat environment has changed drastically since World War Two, the United States has more or less remained the world's prominent power in this time period.

A second option is to consider the period from the post-US Civil War until the present day. The advantage of going back this far in time is that it enables testing of the theory under multiple different conditions. For example, although the specific content of America's polarizing politics is likely different today than that which existed at the end of the US Civil War, the level of polarization, at least by some measurements, is remarkably similar [@hare_polarization_2014; @jeffrey_b_lewis_voteview_2020]. However, by most accounts, the formative developmental years of the military in terms of becoming a profession occurred after the Civil War. Thus, one concern of bounding the scope of research to this larger time horizon is that doing so includes the formative years in which the military presumably ceased engaging in political activities in order to professionalize itself, which could introduce selection bias into the design. On the other hand, this point raises a potentially more important question, and one that is understudied: did the military's ability to become a profession hinge on domestic political factors (i.e., levels of political polarization and military prestige). I think that either time horizon will potentially work in this dissertation, but the plan laid out in the following section of the dissertation incorporates both.

# Plan of the Dissertation

In this section of the prospectus, I will describe the overall plan for my dissertation. In short, I plan to implement a mixed-methods approach that incorporates quantitative and qualitative analysis over the course of what I have now planned to take seven chapters.

In broad terms, the goal of this dissertation is to show that polarization and prestige are the main factors responsible for variation in the political behavior of the military. The dissertation should strive to show, moreover, that the level and substance of polarization serves as the "motive" for the military to behave politically, and that the overall level of prestige of the military serves as the "opportunity" for the military's intervention in politics [@finer_man_1962].

## Chapters 1-2 : Introduction and Theory

In the first chapter, I will frame my research question, lay out my research agenda, and review the relevant literature for my project. There is one critical step to this first chapter, which is deriving (what I argue are) the three foundational norms of civil-military relations of civilian control of the military, the non-partisanship of the military institution, and military expertise. I plan to accomplish this by laying out the concepts espoused by Samuel Huntington, and by summarizing several of the prominent critiques of Huntington [@feaver_civil-military_1996; @rapp_civil-military_2015; @golby_why_2017; @karlin_case_2020; @brooks_paradoxes_2020]. Together, this will show that while there is much disagreement among scholars regarding Huntington's conception of practicing civil-military relations, there is much unity regarding the sanctity of the foundational norms laid out in this paper. In this chapter, I will also make it clear that these norms, while aspirational, have not always been historically upheld by the military, either in contemporary times or closer to America's founding.

In the second chapter, I will introduce, develop, and explain the theory of political behavior by the military that I have laid out in this prospectus. I will begin by summarizing the evolution of contemporary political polarization and the military's prestige, taking care to define these terms. Using Fitch's typology, I will also explain the intervening variable of military "role beliefs," and will argue that these "role beliefs" are largely the product of how military elites think about the inviolability of the foundational norms of civil-military relations, and describe the normative role that military leaders believe they should adopt with respect to politics [@fitch_armed_1998]. Introducing typologies by Finer and Brooks, I will describe the potential ways that militaries can behave politically. I will then argue that while these typologies are useful, they fall short of accounting for the variation in political behavior by American military actors.

I will address this shortcoming in chapter two by advancing a theory that first shows how environments marked by varying levels of polarization and prestige place foundational norms of civil-military relations at differing levels of risk. Next, I will show how these environments produce distinct changes in forms of political behavior by military actors. Chapter two will conclude with the construction of a graphic similar to Table 3 of this prospectus, which clearly identifies the observable indicators and evidence that the remainder of the research project attempts to uncover.

## Chapters 3-5: The Core of the Empirical Analysis

Chapter 3 will consist of a quantitative analysis of one form of political behavior - that of "public appeals" [@brooks_militaries_2009]. Of the four types of political behavior summarized in Table 3, public appeals are visible, constant (in the sense that they always occur, at some level), and available. These qualities indicate that a quantitative analysis is not only possible, but appropriate. 
I have been assembling a data set for approximately eight months consisting of approximately 390 opinion-editorial pieces published by retired military officers in _The New York Times, The Washington Post, and the Wall Street Journal_ since roughly 1979 through the end of 2020. This roughly 40 year period thus begins in a period marked by relatively low military prestige and relatively low political polarization, and ends in a period marked by high military prestige and high political polarization. In the middle (roughly 1990-2004), the period is generally marked by relatively low polarization and high military prestige.

Focusing on appeals made by retired military officers (instead of active duty military officers) produces more variance in the patterns of public appeals because active military officers do not often engage in this form of behavior. Retired military officers, on the other hand, face less constraints with respect to engaging in this form of political behavior than active officers and therefore frequently engage in making public appeals. And yet, the theory posited in this prospectus should still fully apply to retired military officers.

I will examine the first five hypotheses in this chapter, which include the following:

_H1: When political polarization is high, the military (or certain military figures) will increasingly criticize civilian leadership (indicating a drop in the principle of civilian control)._

_H2: When political polarization is low, the military (or certain military figures) will publicly weigh in on traditional national security debates (and not on domestic political concerns or issues outside of relevant military expertise)._

_H3: During periods of high prestige and credibility, the frequency of military involvement in quasi-political activities (such as writing op-eds) will increase relative to periods of low prestige and credibility._

_H4: When both political polarization and high prestige are present, the military will be increasingly vocal and at higher frequency about certain domestic ideological or partisan causes than when these factors are not present._

_H5: During periods of active ground combat, military elites will likely comment on the war, but with less frequency on other domestic issues._

To test these hypotheses, I will first classify each op-ed according to subject type (such as whether the piece primarily covers foreign policy, war fighting, military budgets, institutional military culture, or domestic policy), and code each op-ed for several dichotomous variables, such as whether the op-ed sharply criticizes a partisan figure (the President or other named official), whether the op-ed clearly adopts a partisan stance on an issue, or whether the op-ed strongly calls for definitive legislative or political activity (such as Congress funding a particular initiative or signing a treaty).

Following the coding of each op-ed, I will conduct several regression analyses (likely using logit models), where the dependent variable is a constructed measure of the likelihood of an op-ed piece criticizing a civilian official and/or taking a clearly partisan stance. Data for the levels of polarization exist (measured by either DW-Nominate Scores or a measure for affective polarization). A proxy measure for military prestige also exists using public opinion data on trust in the military (with the caveat that, as laid out in this prospectus, a measure of prestige includes more than trust but is ultimately very difficult to obtain). In addition to these controls, I will control for a couple of biographical pieces of information of the author, such as rank and service affiliation. The results of this regression analyses will reveal what roles, if any, the levels of prestige and polarization are associated with an increased likelihood of an op-ed written by a retired military officer criticizing a civilian official.

In Chapters 4 and 5, instead of focusing squarely on one type of political behavior, I will instead focus on two distinct historical eras, which together cover the years since the end of the Vietnam War. The first era, spanning from the end of the Vietnam War (1973) until the start of President Bill Clinton's administration (1993), is a period that begins with relatively low or moderate political polarization and low military prestige, and ends with relatively low or moderate polarization and high military prestige. The second era, spanning from the beginning of President Clinton's administration (1993) through the end of President Donald Trump's administration (2021), is a period that begins with relatively low political polarization and high military prestige, and ends with high polarization and high military prestige. The goal of each chapter is to show descriptive evidence for the roles that polarization and military prestige played in impacting the distinct forms of political behavior committed by the military within each time period.

Chapter 4 will focus on the period between 1973 and 1993, and thus will focus on an environment which changed from low political polarization and low military prestige to low polarization and high military prestige (following the end of the Gulf War). Tables 2 and 3 together offer several predictions. First, there should be increasing endorsements made by military actors over this time period, but these endorsements should be lopsided in their distribution among candidates. Second, any resignations or threats of resignation should be solely on the basis of matters that are directly related to traditional military expertise, such as military deployment or the use of force abroad. And third, there should be clear evidence of the military engaging Congress as a third party that becomes increasingly successful as military prestige rises, as well as evidence of the military engaging Congress on the basis of maintaining the military's war fighting edge (as opposed to, for instance, ideological reasons or purposes).

Thus, chapter 4 should confirm the following hypotheses:

_H6: Resignations (or threats of) by military actors increase as the prestige of the military rises_.

_H9: The number and level of endorsements made by military actors increases as military prestige increases_.

_H10: The effectiveness of military lobbying efforts with Congress increases as military prestige rises._

Chapter 5 will focus on the period between 2003 and 2021, and thus will focus on an environment that changed from one marked by relatively low political polarization and high military prestige, to one marked by high political polarization and high military prestige. Relying once again on the predictions that stem from Tables 2 and 3, there should be clear evidence of several changes in the political behaviors the military institution commits. First, there should be a steady number of endorsements made by members of the military over this period, and the distribution of these endorsements should even out as polarization increases. Second, instances of resignation or threats of resignation during this period should hold steady in terms of the overall number or level. But as polarization increases, the basis for these resignations should incorporate reasons that are based on ideological differences (and not differences of opinion over matters that are directly related to traditional military expertise). Third, there should also be clear evidence that when the military engages Congress, the military's effectiveness at achieving what it wants holds steady. There should also be evidence of the military (or military actors) engaging Congress to achieve items that are not only directly related to matters of traditional military expertise, but also to achieve ideological goals that align with the preferences of specific military individual actors.

Thus, chapter 5 should confirm the remaining hypotheses:

_H7: The basis of resignation (or threat of) becomes increasingly ideological as polarization increases_.

_H8: Endorsements by military actors become more evenly distributed as polarization increases_.

_H11: The ideological purpose of military lobbying efforts with Congress will widen as polarization increases. When polarization is low, the purpose of military lobbying efforts will center on matters of traditional military expertise._

The data for chapters 4 and 5 will come from a variety of primary and secondary sources, including the memoirs of senior military and government officials, Congressional testimonies, and other government sources. Data on military actor endorsement for Chapter 4 exists, building upon a paper written by Griffiths and Simon (2019). In their paper, they collect data on the level of retired flag officer endorsements of presidential candidates. However, although they hypothesized the reasons why retired flag officers make endorsements, they did not qualitatively scrutinize the reasons that accompanied the endorsements made by these military actors, which I intend to do. Additionally, I will draw attention to the resignations or threats of resignations that occur within each time period for chapters 4 and 5. Since this type of political behavior does not occur often and in numerically high levels, the key is rather to describe the few instances of resignation (or threats of resignation) in great detail, and to show how the presence or absence of political polarization and military prestige impacted these behaviors.

## Chapters 6-7: Outside the Post-WW2 Period and Conclusion

For chapter 6, I plan to examine a couple of periods of US history that are outside of the post World War Two period in order to determine how valid my theory is, and specifically, whether the predictions of the theory apply to eras in which the US military profession was not fully developed. Including this chapter is also important because holistically, chapters 3 - 5 only "touch" three of the four possible environments in which political polarization and military prestige vary. Of the four environments possible, the one that is not examined before chapter 6 is when polarization is high, but military prestige is low.

Therefore, chapter 6 will focus on a period in which these conditions of high polarization and low military prestige prevail. As the data on political polarization shows, the period of time spanning from the end of the American Civil War until roughly 1900 was marked by incredibly high polarization. Additionally, the prestige of the US military dropped in the years after the American Civil War, and as the military not only became smaller, but took up a more constabulary role on the American Frontier in the late 1800s. Therefore, this period makes for an excellent case study of the political behaviors of the military.

Data for this chapter will come from a variety of historical sources, including biographies and memoirs of key military leaders, Congressional testimonies, and historical newspapers. While I do not plan to assemble an original data set for quantitative analysis in this chapter, I do believe that sufficient data exists to argue in support of my theory [@hyman_santon_1960].

In the seventh and final chapter, I will conclude by summarizing my findings and exploring the implications of my argument.

# Conclusion

The purpose of my dissertation is to explore how polarizing domestic politics and increasing military prestige impact the political behaviors of military actors. My initial theory contends that these two independent variables impact the political behavior of military actors by shaping how military actors view their normative roles as military actors in politics. While recent scholars have suspected that polarization and prestige each change the calculus of military actors and how they behave politically, little empirical work has examined how the presence (or absence) of these variables alters the behavior of military actors, and why - a gap this dissertation hopes to fill. Moreover, a careful and rigorous analysis will better inform our understanding of how the US military in particular has endured the forces of polarization and institutional prestige in the past. While we often think of the military as a structured, hierarchical actor, there are theoretical reasons to believe that polarization and prestige may divide members of the military institution, raising questions regarding the difficulties military leaders face in managing and harnessing the military institution as a political actor. Ideally, this dissertation will provide an original contribution that scholars, policymakers, and military leaders can each benefit from.

\pagebreak
\singlespace

### 2021-02-05 code/Tahkv1.Rmd
commit

@@
   _This chapter explores variation in the professionalism of opinion pieces authored by retired US military officers and published in major US newspapers since 1979.
-Drawing on an original data set almost 400 opinion-editorial pieces published by retired military officers in The Wall Street Journal, The Washington Post, The New York Times, The Los Angeles Times, and the USA Today since 1979, the paper argues there has been an increase in published opinion pieces which fail to uphold the concept of military professionalism as defined by historical military norms and standards.
-What explains this variation?
-The author argues that increases in political polarization and the prestige of the military account for the increase in the number of op-eds which challenge traditional conceptions and norms of military professionalism.
+Drawing on an original data set of 390 opinion-editorial pieces published by retired military officers in The Wall Street Journal, The Washington Post, The New York Times, The Los Angeles Times, and the USA Today since 1979, the paper argues there has been an increase in published opinion pieces which fail to uphold the baseline norms of civil-military relations.
+This paper argues that this variation is the result of changes in two domestic political factors: political polarization and the overall prestige of the military institution.
 The paper concludes by arguing that the concept of "politicization of the military" --- something that previous civil-military relations scholars have identified and described --- must be widened to account not only for efforts by politicians to use the military for political advantages, but also for the phenomenon of retired military officers engaging the public to achieve political ends.
@@
 These surveys have asked Americans to rate the level of trust that they place in a variety of American institutions, such as the Presidency, Congress, Churches, and the Police.
-Since the late 1970s, the credibility of the American military as an institution has gradually increased [@gallup_polls_confidence_2020].
-In 1979, 54% of those surveyed expressed either a great deal or quite a lot of confidence in the military, while in 2020, 72% expressed the same [@gallup_polls_confidence_2020].
+Since the late 1970s, the credibility of the American military as an institution has gradually increased [@noauthor_confidence_2020].
+In 1979, 54% of those surveyed expressed either a great deal or quite a lot of confidence in the military, while in 2020, 72% expressed the same [@noauthor_confidence_2020].
 Trends in political polarization (using DW-Nominate Scores and a measure of Affective Polarization) and confidence in the military are displayed in Figure Four, below.
